### 2017-12-21 svrpg.tex
initial commit

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2018}

% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{icml2018}

%%%USEFUL PACKAGES%%%
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{amsmath,amssymb}
\usepackage{mathtools}
%%Theorems
\usepackage{amsthm}
\theoremstyle{remark}
\newtheorem*{rem}{Remark}
\newtheorem*{note}{Note}
\newtheorem{case}{Case}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{conj}{Conjecture}
\newtheorem{example}{Example}
%Restatable
\usepackage{thmtools, thm-restate}
\declaretheorem[numberwithin=section]{thm}
\declaretheorem[sibling=thm]{lemma}
\declaretheorem[sibling=thm]{corollary}
\declaretheorem[sibling=thm]{assumption}
\declaretheorem[sibling=thm]{theorem}
%e.g. ...
\usepackage{xspace}
\DeclareRobustCommand{\eg}{e.g.,\@\xspace}
\DeclareRobustCommand{\ie}{i.e.,\@\xspace}
\DeclareRobustCommand{\wrt}{w.r.t.\@\xspace}
%%%%%%

%%%CUSTOM COMMANDS%%%
%Math
\newcommand{\realspace}{\mathbb R}      % realspace
\newcommand{\transpose}[1]{{#1}^\texttt{T}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\EV}{\mathbb{E}}
\newcommand{\EVV}[2][\ppvect \in \ppspace]{\EV_{#1}\left[{#2}\right]}
\newcommand{\norm}[2][\infty]{\left\|#2\right\|_{#1}}
\newcommand{\Dij}[2]{\frac{\partial^{2}{#1}}{\partial{#2}_i\partial{#2}_j}}
\newcommand{\de}{\,\mathrm{d}}
%RL
\newcommand{\vtheta}{\boldsymbol{\theta}}
\newcommand{\Aspace}{\mathcal{A}}
\newcommand{\Sspace}{\mathcal{S}}
\newcommand{\stationary}{d_{\mu}^{\pi_{\vtheta}}(s)}
\newcommand{\policy}{\pi_{\vtheta}(a \mid s)}
\newcommand{\pol}{\pi_{\vtheta}}
\newcommand{\score}[1]{\nabla_{#1}\log\policy}
\newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
\newcommand{\Vfun}{V^{\pi_{\vtheta}}(s)}
\newcommand{\vTheta}{\boldsymbol{\Theta}}
\newcommand{\vphi}{\boldsymbol{\phi}}
\newcommand{\gradJ}[1]{\nabla_{#1}J_\mu(\vtheta)}
\newcommand{\gradApp}[1]{\hat{\nabla}_{#1}J_{\mu}(\vtheta)}
\newcommand{\eqdef}{\mathrel{\mathop:}=}
%%%%%%

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{SVRPG}

\begin{document}

\twocolumn[
\icmltitle{Stochastic Variance Reduced Policy Gradient}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2018
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Damiano Binaghi}{equal,polimi}
\icmlauthor{Giuseppe Canonaco}{equal,polimi}
\icmlauthor{Matteo Papini}{polimi}
\icmlauthor{Matteo Pirotta}{inria}
\icmlauthor{Marcello Restelli}{polimi}
\end{icmlauthorlist}

\icmlaffiliation{polimi}{Politecnico di Milano, Milano, Italy}
\icmlaffiliation{inria}{Inria, Lille, France}

\icmlcorrespondingauthor{Matteo Papini}{matteo.papini@polimi.it}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Machine Learning, ICML}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}

\end{abstract}

\section{Introduction}

\section{Preliminaries}

\section{Algorithm}

\section{Convergence Guarantees}

\section{Practical Variants}

\section{Related Work}

\section{Experiments}

\section{Discussion}

\bibliography{paper}
\bibliographystyle{icml2018}

\appendix

\end{document}

### 2017-12-30 svrpg.tex
Preliminaries: written policy gradient subsection

@@
 \newcommand{\Sspace}{\mathcal{S}}
-\newcommand{\stationary}{d_{\mu}^{\pi_{\vtheta}}(s)}
+\newcommand{\Tspace}{\mathcal{T}}
+\newcommand{\Transition}{\mathcal{P}}
+\newcommand{\Reward}{\mathcal{R}}
+\newcommand{\stationary}{d_{\rho}^{\pi_{\vtheta}}(s)}
 \newcommand{\policy}{\pi_{\vtheta}(a \mid s)}
 \newcommand{\pol}{\pi_{\vtheta}}
-\newcommand{\score}[1]{\nabla_{#1}\log\policy}
+\newcommand{\score}[2]{\nabla_{#1}\log\pi_{#1}(#2)}
 \newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
@@
 \newcommand{\vphi}{\boldsymbol{\phi}}
-\newcommand{\gradJ}[1]{\nabla_{#1}J_\mu(\vtheta)}
-\newcommand{\gradApp}[1]{\hat{\nabla}_{#1}J_{\mu}(\vtheta)}
+\newcommand{\gradJ}[1]{\nabla_{#1}J(\vtheta)}
+\newcommand{\gradApp}[1]{\hat{\nabla}_{#1}J(\vtheta)}
 \newcommand{\eqdef}{\mathrel{\mathop:}=}
+\newcommand{\Dataset}{\mathcal{D}}
 %%%%%%
@@
 \section{Preliminaries}
+In this section we provide essential background on policy gradient methods for Reinforcement Learning and variance-reduced stochastic gradient methods for finite sum optimization.
+In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
+
+\subsection{Policy Gradient for Reinforcement Learning}
+A Reinforcement Learning task can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space, $\Aspace$ is a continuous action space, $\Transition$ is a Markovian transition model, where $\Transition(s'\mid s,a)$ defines the transition density from state $s$ to $s'$ under action $a$, $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward, $\gamma\in[0,1)$ is the discount factor and $\rho$ is the initial state distribution.
+The agent's behavior is modeled as a policy $\pi$, where $\pi(a\mid s)$ is the density distribution over $\Aspace$ of the action to take in state $s$.
+We consider episodic tasks, \ie, tasks composed of episodes of length $H$, also called time horizon.
+In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
+With abuse of notation, we denote with $\pi(\tau)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories, and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
+\begin{align*}
+\Reward(\tau) = \sum_{t=0}^{H-1}\gamma^t\Reward(s_t,a_t).
+\end{align*}
+Performance $J$ allows to rank policies in terms of expected reward:
+\begin{align*}
+	J(\pi) = \int_{\Tspace}\pi(\tau)\Reward(\tau)\de \tau.
+\end{align*}
+Solving an MDP means finding a policy $\pi^*$ maximizing $J$.
+\par
+Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol : \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable in the $m$-dimensional parameter vector $\vtheta$.
+We denote the performance of a parametric policy with $J(\vtheta)$ for brevity.
+The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is:
+\begin{align*}
+\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
+\end{align*}
+We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment.
+In this setting, stochastic gradient descent is typically employed: after sampling a trajectory $\tau$ or a batch of $N$ trajectories $\Dataset_{N}$,
+policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}$, where $\alpha$ is a step size and $\gradApp{\vtheta}$ is an estimate of the true gradient.
+The REINFORCE gradient estimator provides a simple, unbiased way of estimating the gradient:
+\begin{align*}
+\gradApp{\vtheta} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
+\end{align*}
+where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
+The G(PO)MDP gradient estimator is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
+\begin{align*}
+\gradApp{\vtheta} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\gamma^h r_h^n - b\right).
+\end{align*}
 

### 2018-01-05 svrpg.tex
Started writing SVRG subsection

@@
 \section{Preliminaries}
-In this section we provide essential background on policy gradient methods for Reinforcement Learning and variance-reduced stochastic gradient methods for finite sum optimization.
+In this section we provide essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite sum optimization.
 In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
 
-\subsection{Policy Gradient for Reinforcement Learning}
+\subsection{Policy Gradient}
 A Reinforcement Learning task can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space, $\Aspace$ is a continuous action space, $\Transition$ is a Markovian transition model, where $\Transition(s'\mid s,a)$ defines the transition density from state $s$ to $s'$ under action $a$, $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward, $\gamma\in[0,1)$ is the discount factor and $\rho$ is the initial state distribution.
@@
 
+\subsection{Stochastic Variance-Reduced Gradient}
+Finite sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $g_i(\vtheta)$:
+\begin{align*}
+\max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g_i(\vtheta).
+\end{align*}
+This kind of optimization is very common in machine learning, where each $g_i$ may correspond to a data sample $x_i$. In this case, we adopt the following, more meaningful notation:
+\begin{align*}
+\max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g_{\vtheta}(x_i).
+\end{align*}
+A common requirement is that $g$ must be smooth and convex in $\vtheta$.
+Under this hypothesis, full gradient ascent with a constant step size achieves a convergence rate of $O(\nicefrac{1}{k})$, \ie $f(\vtheta^*) - f(\vtheta_k) = O(\nicefrac{1}{k})$, where $\vtheta^*$ is the optimal parameter and $\vtheta_k$ is the parameter after $k$ gradient updates.
+The problem with full gradient is that each update can be very expensive, especially when $N$ is large.
+Stochastic gradient ascent overcomes this problem, at the cost introducing some variance.
+As a result, the (expected) convergence rate drops to $O(\nicefrac{1}{\sqrt{k}})$.
+
 \section{Algorithm}

### 2018-01-05 svrpg.tex
added references, switched to IFO complexity formulation

@@
 \subsection{Policy Gradient}
-A Reinforcement Learning task can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space, $\Aspace$ is a continuous action space, $\Transition$ is a Markovian transition model, where $\Transition(s'\mid s,a)$ defines the transition density from state $s$ to $s'$ under action $a$, $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward, $\gamma\in[0,1)$ is the discount factor and $\rho$ is the initial state distribution.
+A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'\mid s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
 The agent's behavior is modeled as a policy $\pi$, where $\pi(a\mid s)$ is the density distribution over $\Aspace$ of the action to take in state $s$.
@@
 We denote the performance of a parametric policy with $J(\vtheta)$ for brevity.
-The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is:
+The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
 \begin{align*}
@@
 policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}$, where $\alpha$ is a step size and $\gradApp{\vtheta}$ is an estimate of the true gradient.
-The REINFORCE gradient estimator provides a simple, unbiased way of estimating the gradient:
+The REINFORCE gradient estimator \cite{williams1992simple} provides a simple, unbiased way of estimating the gradient:
 \begin{align*}
@@
 where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
-The G(PO)MDP gradient estimator is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
+The G(PO)MDP gradient estimator \cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
 \begin{align*}
@@
 A common requirement is that $g$ must be smooth and convex in $\vtheta$.
-Under this hypothesis, full gradient ascent with a constant step size achieves a convergence rate of $O(\nicefrac{1}{k})$, \ie $f(\vtheta^*) - f(\vtheta_k) = O(\nicefrac{1}{k})$, where $\vtheta^*$ is the optimal parameter and $\vtheta_k$ is the parameter after $k$ gradient updates.
-The problem with full gradient is that each update can be very expensive, especially when $N$ is large.
-Stochastic gradient ascent overcomes this problem, at the cost introducing some variance.
-As a result, the (expected) convergence rate drops to $O(\nicefrac{1}{\sqrt{k}})$.
+Under this hypothesis, full gradient ascent \cite{cauchy1847methode} with a constant step size achieves an IFO complexity \cite{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.
+This can be too expensive for large values of $N$.
+Stochastic gradient ascent \cite{robbins1951stochastic} \cite{bottou2004large} overcomes this problem, at the cost introducing some variance.
+The IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation: it does not depend on $N$, but it has a worse dependency on $\epsilon$.
 
@@
 
-\bibliography{paper}
+\bibliography{svrpg}
 \bibliographystyle{icml2018}

### 2018-01-08 svrpg.tex
written convex SVRG part; some stuff will probably be moved to the introduction; may add pseudocode for SVRG

@@
 \end{align*}
-This kind of optimization is very common in machine learning, where each $g_i$ may correspond to a data sample $x_i$. In this case, we adopt the following, more meaningful notation:
+This kind of optimization is very common in machine learning, where each $g_i$ may correspond to a data sample $x_i$ from a dataset $\mathcal{D}_N$ of size $N$.
+In this case, we adopt the following, more meaningful notation:
 \begin{align*}
@@
 This can be too expensive for large values of $N$.
-Stochastic gradient ascent \cite{robbins1951stochastic} \cite{bottou2004large} overcomes this problem, at the cost introducing some variance.
-The IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation: it does not depend on $N$, but it has a worse dependency on $\epsilon$.
+Stochastic gradient ascent \cite{robbins1951stochastic} \cite{bottou2004large} overcomes this problem by sampling a single $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling.
+The resulting IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$.
+Starting from SAG \cite{roux2012stochastic}, a series of variations to stochastic gradient ascent have been proposed to achieve a better trade-off between convergence speed and cost per iteration.
+This family of algorithms includes SAG, SVRG \cite{johnson2013accelerating}, SAGA \cite{defazio2014saga} and Finito \cite{defazio2014finito}, the common intuition being the reuse of previous gradient computations to reduce the variance of the current one.
+In particular, Stochastic Variance Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which can become a problem when neural networks with several parameters are employed.
 
+The idea of SVRG is to alternate full and stochastic gradient updates.
+Each $m$ iterations, a snapshot $\tilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\tilde{\vtheta})$.
+Otherwise, the parameter is updated with a corrected stochastic gradient estimate of the form:
+\begin{align*}
+\widehat{\nabla}_{SVRG}f(\vtheta) = 
+	\nabla f(\tilde{\vtheta}) + 
+	\nabla g_{\vtheta}(x_i) -
+	\nabla g_{\tilde{\vtheta}}(x_i),
+\end{align*} 
+where $x_i$ is sampled uniformly at random from $\mathcal{D}_N$.
+The corrected gradient is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient.
+The resulting IFO complexity is $O(N+\nicefrac{\sqrt{N}}{\epsilon})$ \cite{reddi2016stochastic}.
 \section{Algorithm}

### 2018-01-08 svrpg.tex
Section 2 completed; some stuff will be moved to introduction

@@
 The idea of SVRG is to alternate full and stochastic gradient updates.
-Each $m$ iterations, a snapshot $\tilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\tilde{\vtheta})$.
+Each $m = O(N)$ iterations, a snapshot $\tilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\tilde{\vtheta})$.
 Otherwise, the parameter is updated with a corrected stochastic gradient estimate of the form:
@@
 where $x_i$ is sampled uniformly at random from $\mathcal{D}_N$.
-The corrected gradient is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient.
+The corrected gradient $\widehat{\nabla}_{SVRG}f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient.
 The resulting IFO complexity is $O(N+\nicefrac{\sqrt{N}}{\epsilon})$ \cite{reddi2016stochastic}.
+
+More recently, some extensions of variance reduction algorithms to the non-convex objectives have been proposed \cite{reddi2016stochastic} \cite{allen2016variance} \cite{reddi2016stochastic} \cite{reddi2016fast}.
+In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
+Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the convex case \cite{nesterov2013introductory} \cite{ghadimi2013stochastic}.
+Also in this case, SVRG yields an advantageous trade-off, with an IFO complexity of $O(N + \nicefrac{N^{\nicefrac{2}{3}}}{N})$ \cite{reddi2016stochastic}.
+The only additional requirement is to select $\vtheta^*$ uniformly at random among all the $\vtheta_k$ instead of simply setting it to the final value.
+Pseudocode of SVRG for both the convex and the non-convex case is provided in Algorithm \ref{alg:svrg}.
+
+\begin{algorithm}[tb]
+	\caption{SVRG}
+	\label{alg:svrg}
+	\begin{algorithmic}
+		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_0$
+		\FOR{$s=0$ {\bfseries to} $S-1$}
+		\STATE $\vtheta_{s+1}^0 = \tilde{\vtheta} = \vtheta_s$
+		\FOR{$t=0$ {\bfseries to} $m-1$}
+		\STATE Pick $i$ uniformly at random from $[1,N]$
+		\STATE $\nabla_{SVRG} = 
+					\nabla f(\tilde{\vtheta}) + 
+					\nabla g_{\vtheta_{s+1}^t}(x_i) -
+					\nabla g_{\tilde{\vtheta}}(x_i)$
+		\STATE $\vtheta_{s+1}^{t+1} = \vtheta_{s+1}^{t} + \alpha\nabla_{SVRG}$
+		\ENDFOR
+		\STATE $\vtheta_{s+1} = \vtheta_{s+1}^m$
+		\ENDFOR
+		\STATE {\bfseries Convex case: return} $\vtheta_S$
+		\STATE {\bfseries Non-Convex case: return} $\vtheta_s^t$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
+	\end{algorithmic}
+\end{algorithm}
+
 \section{Algorithm}

### 2018-01-10 svrpg.tex
started writing section 3 -Algorithm; some changes in the notation

@@
 \newcommand{\pol}{\pi_{\vtheta}}
+\newcommand{\trajdistr}{\pi_{\vtheta}(\tau)}
 \newcommand{\score}[2]{\nabla_{#1}\log\pi_{#1}(#2)}
@@
 \newcommand{\gradJ}[1]{\nabla_{#1}J(\vtheta)}
-\newcommand{\gradApp}[1]{\hat{\nabla}_{#1}J(\vtheta)}
+\newcommand{\gradApp}[2]{\hat{\nabla}_{#2}J(#1)}
 \newcommand{\eqdef}{\mathrel{\mathop:}=}
@@
 \par
-Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol : \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable in the $m$-dimensional parameter vector $\vtheta$.
-We denote the performance of a parametric policy with $J(\vtheta)$ for brevity.
+Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable in the $m$-dimensional parameter vector $\vtheta$.
+We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we denote $\pol(\tau)$ with $\pi(\tau\mid\vtheta)$ for the sake of readability.
 The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
 \begin{align*}
-\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
+\gradJ{} = \int_{\Tspace}\pol(\tau)\score{}{\tau}\Reward(\tau)\de \tau.
 \end{align*}
@@
 In this setting, stochastic gradient descent is typically employed: after sampling a trajectory $\tau$ or a batch of $N$ trajectories $\Dataset_{N}$,
-policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}$, where $\alpha$ is a step size and $\gradApp{\vtheta}$ is an estimate of the true gradient.
+policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient.
 The REINFORCE gradient estimator \cite{williams1992simple} provides a simple, unbiased way of estimating the gradient:
 \begin{align*}
-\gradApp{\vtheta} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
 \end{align*}
@@
 \begin{align*}
-\gradApp{\vtheta} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\gamma^h r_h^n - b\right).
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\gamma^h r_h^n - b\right).
 \end{align*}
@@
 \begin{align*}
-\max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g_{\vtheta}(x_i).
+\max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g(x_i \mid \vtheta).
 \end{align*}
@@
 \begin{align*}
-\widehat{\nabla}_{SVRG}f(\vtheta) = 
+\blacktriangledown f(\vtheta) = 
 	\nabla f(\tilde{\vtheta}) + 
-	\nabla g_{\vtheta}(x_i) -
-	\nabla g_{\tilde{\vtheta}}(x_i),
+	\nabla g(x_i\mid\vtheta) -
+	\nabla g(x_i \mid \tilde{\vtheta}),
 \end{align*} 
 where $x_i$ is sampled uniformly at random from $\mathcal{D}_N$.
-The corrected gradient $\widehat{\nabla}_{SVRG}f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient.
+The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient.
 The resulting IFO complexity is $O(N+\nicefrac{\sqrt{N}}{\epsilon})$ \cite{reddi2016stochastic}.
@@
 		\STATE Pick $i$ uniformly at random from $[1,N]$
-		\STATE $\nabla_{SVRG} = 
+		\STATE $\blacktriangledown f(\vtheta_{s+1}^t) = 
 					\nabla f(\tilde{\vtheta}) + 
-					\nabla g_{\vtheta_{s+1}^t}(x_i) -
-					\nabla g_{\tilde{\vtheta}}(x_i)$
-		\STATE $\vtheta_{s+1}^{t+1} = \vtheta_{s+1}^{t} + \alpha\nabla_{SVRG}$
+					\nabla g(x_i\mid\vtheta_{s+1}^t) -
+					\nabla g(x_i \mid \tilde{\vtheta})$
+		\STATE $\vtheta_{s+1}^{t+1} = \vtheta_{s+1}^{t} + \alpha\blacktriangledown f(\vtheta_{s+1}^t)$
 		\ENDFOR
@@
 \section{Algorithm}
+In on-line policy gradient problems, stochastic gradient ascent (SGA) is not really a choice, but is dictated by the necessity of interacting with an unknown environment.
+We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
+In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$.
+Compared to the typical finite sum optimization scenario, we have the following additional challenges:
+\begin{itemize}
+	\item The objective $J(\vtheta)$ is typically non-convex;
+	\item The dataset $\mathcal{T}$ can be infinite;
+	\item The value of the parameter $\vtheta$ influences the sampling of trajectories;
+\end{itemize}
+To deal with non-convexity, we require $J(\vtheta)$ to be L-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and Softmax [appendix?].
+Because of the infinite dataset, we can only rely on an estimate of the full gradient.
+Finally, we employ importance weighting \cite{rubinstein1981simulation} \cite{precup2000eligibility} to guarantee proper sampling of trajectories.
+We focus on the REINFORCE gradient estimator for simplicity, but the extension to G(PO)MDP is straightforward.
+The proposed SVRG-like gradient estimate, which we call Stochastic Variance Reduced Policy Gradient (SVRPG), is the following:
+\begin{align*}
+	\blacktriangledown J(\vtheta) = &\frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta}) \\
+		&+ \omega(\tau_i)\nabla\log\pi(\tau_i \mid \vtheta) 
+		- \nabla\log\pi(\tau_i \mid \tilde{\vtheta}),
+\end{align*}
+where trajectories $\tau_j$ for $j=1,\dots,N$ and $\tau_i$ are all sampled independently from the snapshot policy $\pi_{\tilde{\vtheta}}$, and $\omega(\tau) = \frac{\pi(\tau\mid\vtheta)}{\pi(\tau\mid\tilde{\vtheta})}$ is an importance weight from $\pi_{\tilde{\vtheta}}$ to the current policy $\pol$.
+We immediately have the following (proof in the Appendix):
+\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
+The SVRPG gradient estimator is unbiased:
+\[
+\mathop{E}_{\tau\sim\pi(\cdot\mid\tilde{\vtheta})}\left[\blacktriangledown J(\vtheta)\right] = \gradJ{}.
+\]
+\end{restatable}
+The resulting SVRPG algorithm is detailed in [todo:pseudocode].
+[Next: explain which details are left open and briefly outline next sections; deeper explanation of algorithm?]
 

### 2018-01-11 svrpg.tex
added svrpg pseudocode

@@
 	\begin{algorithmic}
-		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_0$
+		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m-1}^0$
 		\FOR{$s=0$ {\bfseries to} $S-1$}
-		\STATE $\vtheta_{s+1}^0 = \tilde{\vtheta} = \vtheta_s$
+		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m-1}^s$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
 		\STATE Pick $i$ uniformly at random from $[1,N]$
-		\STATE $\blacktriangledown f(\vtheta_{s+1}^t) = 
-					\nabla f(\tilde{\vtheta}) + 
-					\nabla g(x_i\mid\vtheta_{s+1}^t) -
-					\nabla g(x_i \mid \tilde{\vtheta})$
-		\STATE $\vtheta_{s+1}^{t+1} = \vtheta_{s+1}^{t} + \alpha\blacktriangledown f(\vtheta_{s+1}^t)$
+		\STATE \begin{align*}
+			\blacktriangledown f(\vtheta_t^{s+1}) = 
+			&\nabla f(\tilde{\vtheta}^s) \\ + 
+			&\nabla g(x_i\mid\vtheta_t^{s+1}) -
+			\nabla g(x_i \mid \tilde{\vtheta}^{s+1})
+		\end{align*}
+		\STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha\blacktriangledown f(\vtheta_t^{s+1})$
 		\ENDFOR
-		\STATE $\vtheta_{s+1} = \vtheta_{s+1}^m$
 		\ENDFOR
-		\STATE {\bfseries Convex case: return} $\vtheta_S$
-		\STATE {\bfseries Non-Convex case: return} $\vtheta_s^t$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
+		\STATE {\bfseries Convex case: return} $\vtheta_{m-1}^S$
+		\STATE {\bfseries Non-Convex case: return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
 	\end{algorithmic}
@@
 \end{restatable}
-The resulting SVRPG algorithm is detailed in [todo:pseudocode].
+The resulting SVRPG algorithm is detailed in Algorithm \ref{alg:svrpg}.
 [Next: explain which details are left open and briefly outline next sections; deeper explanation of algorithm?]
+
+\begin{algorithm}[tb]
+	\caption{SVRPG}
+	\label{alg:svrpg}
+	\begin{algorithmic}
+		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m-1}^0$, batch size $N$
+		\FOR{$s=0$ {\bfseries to} $S-1$}
+		\STATE Sample N trajectories $\tau_j$ from $\pi(\cdot\mid\tilde{\vtheta}^{s})$
+		\STATE $\gradApp{\tilde{\vtheta}^{s}}{N} = \frac{1}{N}\sum_{j=0}^{N-1}\score{}{\tau_j\mid\tilde{\vtheta}^{s}}$
+		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m-1}^s$
+		\FOR{$t=0$ {\bfseries to} $m-1$}
+		\STATE Sample trajectory $\tau_i$ from 				$\pi(\cdot\mid\tilde{\vtheta}^{s})$
+		\STATE 
+		\begin{align*}
+		\blacktriangledown J(\vtheta_t^{s+1}) = 
+		&\gradApp{\tilde{\vtheta}^s}{N}
+		+ \omega(\tau_i)\score{}{t_i\mid\vtheta_t^{s+1}} \\
+		& - \score{}{\tau_i \mid \tilde{\vtheta}^{s+1}}
+		\end{align*}
+		\STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha\blacktriangledown J(\vtheta_t^{s+1})$
+		\ENDFOR
+		\ENDFOR
+		\STATE {\bfseries return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
+	\end{algorithmic}
+\end{algorithm}
 

### 2018-01-11 svrpg.tex
first complete draft of Section 3; the algorithm could be explained in more detail

@@
 \end{restatable}
-The resulting SVRPG algorithm is detailed in Algorithm \ref{alg:svrpg}.
-[Next: explain which details are left open and briefly outline next sections; deeper explanation of algorithm?]
+The resulting SVRPG policy optimization method is detailed in Algorithm \ref{alg:svrpg}.
+To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient and the step size $\alpha$.
+Adaptive selection of these meta-parameters is also possible.
+However, the next section will focus on providing converge guarantees for the plain SVRPG algorithm.
 

### 2018-01-15 svrpg.tex
added R(\tau) in SVRPG gradient definition and pseudocode
switched to method A; old version stored in separate branch

@@
 \begin{align*}
-	\blacktriangledown J(\vtheta) = &\frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta}) \\
-		&+ \omega(\tau_i)\nabla\log\pi(\tau_i \mid \vtheta) 
-		- \nabla\log\pi(\tau_i \mid \tilde{\vtheta}),
+	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta})\Reward(\tau_j) \\
+		&+ \nabla\log\pi(\tau_i \mid \vtheta)\Reward(\tau_i) 
+		- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta})\Reward(\tau_i),
 \end{align*}
-where trajectories $\tau_j$ for $j=1,\dots,N$ and $\tau_i$ are all sampled independently from the snapshot policy $\pi_{\tilde{\vtheta}}$, and $\omega(\tau) = \frac{\pi(\tau\mid\vtheta)}{\pi(\tau\mid\tilde{\vtheta})}$ is an importance weight from $\pi_{\tilde{\vtheta}}$ to the current policy $\pol$.
+where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from the snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\mid\tilde{\vtheta})}{\pi(\tau\mid\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$.
 We immediately have the following (proof in the Appendix):
@@
 \end{restatable}
+[May also add Lemma about convergence of the variance to zero, since it provides an intuitive justification of some design choices]
 The resulting SVRPG policy optimization method is detailed in Algorithm \ref{alg:svrpg}.
@@
 		\STATE Sample N trajectories $\tau_j$ from $\pi(\cdot\mid\tilde{\vtheta}^{s})$
-		\STATE $\gradApp{\tilde{\vtheta}^{s}}{N} = \frac{1}{N}\sum_{j=0}^{N-1}\score{}{\tau_j\mid\tilde{\vtheta}^{s}}$
+		\STATE $\gradApp{\tilde{\vtheta}^{s}}{N} = \frac{1}{N}\sum_{j=0}^{N-1}\score{}{\tau_j\mid\tilde{\vtheta}^{s}}\Reward(\tau_j)$
 		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m-1}^s$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
-		\STATE Sample trajectory $\tau_i$ from 				$\pi(\cdot\mid\tilde{\vtheta}^{s})$
+		\STATE Sample trajectory $\tau_i$ from 				$\pi(\cdot\mid\vtheta_t^{s+1})$
 		\STATE 
@@
 		&\gradApp{\tilde{\vtheta}^s}{N}
-		+ \omega(\tau_i)\score{}{t_i\mid\vtheta_t^{s+1}} \\
-		& - \score{}{\tau_i \mid \tilde{\vtheta}^{s+1}}
+		+ \score{}{\tau_i\mid\vtheta_t^{s+1}}\Reward(\tau_i) \\
+		& - \omega(\tau_i)\score{}{\tau_i \mid \tilde{\vtheta}^{s}}\Reward(\tau_i)
 		\end{align*}

### 2018-01-15 svrpg.tex
added appendix for proofs

@@
 where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from the snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\mid\tilde{\vtheta})}{\pi(\tau\mid\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$.
-We immediately have the following (proof in the Appendix):
+We immediately have the following (proof in Appendix \ref{app:proofs}):
+
 \begin{restatable}{lemma}{unbias}\label{lemma:unbias}
@@
 \end{restatable}
+
 [May also add Lemma about convergence of the variance to zero, since it provides an intuitive justification of some design choices]
@@
 
+\section{Proofs}\label{app:proofs}
+
+\unbias*
+\begin{proof}
+	...
+\end{proof}
+
 \end{document}

### 2018-01-15 svrpg.tex
written main theorem

@@
 \DeclareMathOperator*{\EV}{\mathbb{E}}
+\DeclareMathOperator{\Tr}{Tr}
+\DeclareMathOperator*{\Cov}{\mathbb{C}ov}
 \newcommand{\EVV}[2][\ppvect \in \ppspace]{\EV_{#1}\left[{#2}\right]}
@@
 \end{align*}
-where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from the snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\mid\tilde{\vtheta})}{\pi(\tau\mid\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$.
+where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from a snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\mid\tilde{\vtheta})}{\pi(\tau\mid\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$.
 We immediately have the following (proof in Appendix \ref{app:proofs}):
@@
 \[
-\mathop{E}_{\tau\sim\pi(\cdot\mid\tilde{\vtheta})}\left[\blacktriangledown J(\vtheta)\right] = \gradJ{}.
+\mathop{E}_{\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}),\tau_i\sim\pi(\cdot\mid\vtheta)}
+\left[\blacktriangledown J(\vtheta)\right] = \gradJ{}.
 \]
@@
 
+\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
+The parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some constants $\psi>0, b>0$ and $c_{t}>0$ for $t=1,\dots,m$, the following property:
+\[
+	\EVV[]
+	{\norm[]{\nabla J(\vtheta_A)}^2} \leq
+		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
+		\frac{Z}{N},
+\]
+where
+\begin{align*}
+	Z &= b\max_{s}\EVV[]{\norm[]{\gradApp{\tilde{\vtheta^s}}{1}}^2} \\
+	&+ \max_{s,t}\left|\Tr\Cov\left[
+		\gradApp{\tilde{\vtheta^s}}{1},
+		c_{t+1}\vtheta_t^{s+1} + \alpha_t\nabla J(\vtheta_t^{s+1})\right]\right|
+\end{align*}
+\end{restatable}
+
 \section{Practical Variants}
@@
 
+\convergence*
+\begin{proof}
+	...
+\end{proof}
 \end{document}

### 2018-01-17 svrpg.tex
added definitions of constants in proof of main theorem!
started writing proof of Lemma 1

@@
 \usepackage{mathtools}
+\usepackage{pifont}
 %%Theorems
@@
 \newcommand{\de}{\,\mathrm{d}}
+\newcommand{\dotprod}[2]{\left\langle#1,#2\right\rangle}
 %RL
@@
 \newcommand{\vphi}{\boldsymbol{\phi}}
-\newcommand{\gradJ}[1]{\nabla_{#1}J(\vtheta)}
+\newcommand{\gradJ}[1]{\nabla J(#1)}
 \newcommand{\gradApp}[2]{\hat{\nabla}_{#2}J(#1)}
@@
 \newcommand{\Dataset}{\mathcal{D}}
+%Specific
+\newcommand{\Ets}[2][t]{\mathbb{E}_{#1\mid s}\left[#2\right]}
+\newcommand{\Covts}[3][t]{{\mathbb{C}\text{ov}}_{#1\mid s}\left(#2,#3\right)}
+\newcommand{\Varts}[2][t]{{\mathbb{V}\text{ar}}_{#1\mid s}\left[#2\right]}
+\newcommand{\gradBlack}[1]{\blacktriangledown J(#1)}
+\newcommand{\gradIdeal}[1]{\overline{\blacktriangledown} J(#1)}
 %%%%%%
@@
 \begin{align*}
-\gradJ{} = \int_{\Tspace}\pol(\tau)\score{}{\tau}\Reward(\tau)\de \tau.
+\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{}{\tau}\Reward(\tau)\de \tau.
 \end{align*}
@@
 \[
-\mathop{E}_{\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}),\tau_i\sim\pi(\cdot\mid\vtheta)}
+\mathop{\mathbb{E}}
 \left[\blacktriangledown J(\vtheta)\right] = \gradJ{}.
@@
 \begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
-The parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some constants $\psi>0, b>0$ and $c_{t}>0$ for $t=1,\dots,m$, the following property:
+Provided $J(\vtheta)$ is L-smooth with Lipschitz constant $L$ and step sizes are sufficiently small, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some constants $\psi>0, D>0$ and $c_{t}>0$ for $t=1,\dots,m$, the following property:
 \[
@@
 \begin{align*}
-	Z &= b\max_{s}\EVV[]{\norm[]{\gradApp{\tilde{\vtheta^s}}{1}}^2} \\
+	Z &= D\max_{s}\EVV[]{\norm[]{\gradApp{\tilde{\vtheta^s}}{1}}^2} \\
 	&+ \max_{s,t}\left|\Tr\Cov\left[
@@
 
+\newpage
+\mbox{}
+\newpage
+\onecolumn
 \appendix
+
+\section{Definitions}
+We give some additional definitions which will be useful in the proofs of Appendix \ref{app:proofs}.
+\begin{definition}
+For a random variable $X$:
+\begin{align*}
+	&\mathbb{E}_{t\mid s}\left[X\right] \coloneqq 
+		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}^s) \\ \tau_i\sim\pi(\cdot\mid\vtheta^{s+1}_i)\text{ for $i=0,\dots,t-1$}}}{\left[X \mid \tilde{\vtheta^s}\right]} \\
+	&\coloneqq \EVV[\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}^s)]{
+			\EVV[\tau_0\sim\pi(\cdot\mid\vtheta_0^{s+1})]
+				{\dots
+					\EVV[\tau_t\sim\pi(\cdot\mid\vtheta_t^{s+1})]
+						{X\mid\vtheta_t^{s+1}}
+				 \dots
+			\mid\vtheta_0^{s+1}}
+		\mid\tilde{\vtheta}^s},
+\end{align*}
+where the sequence $\tilde{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}. 
+\end{definition}
+Intuitively, the $\Ets{\cdot}$ operator computes the expected value with respect to the sampling of trajectories from the snapshot $\tilde{\vtheta}^s$ up to the $t$-th iteration included.
+Note that the order in which expected values are taken is important since each $\vtheta_{t}^{s+1}$ is function of previously sampled trajectories and is used to sample new ones.
+
+\begin{definition}
+The full gradient estimation error is:
+\[
+	e_s \coloneqq \gradApp{\tilde{\vtheta}^s}{N} - \gradJ{\tilde{\vtheta}^s} 
+\]
+\end{definition}
+
+\begin{definition}\label{def:ideal}
+The ideal SVRPG gradient estimate is:
+\begin{align*}
+	\overline{\blacktriangledown}J(\vtheta_t^{s+1}) &\coloneqq 
+	\gradJ{\tilde{\vtheta}^s}
+	+ \nabla\log\pi(\tau_i \mid \vtheta_t^{s+1})\Reward(\tau_i) 
+	- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta}^s)\Reward(\tau_i) \\
+	&= \gradBlack{\vtheta_t^{s+1}} - \gradApp{\tilde{\vtheta}^s}{N} + \gradJ{\tilde{\vtheta}^s} \\
+	&= \gradBlack{\vtheta_t^{s+1}} - e_s
+\end{align*}
+\end{definition}
 
@@
 \begin{proof}
+We prove the theorem for the following values of the constants:
+\begin{align*}
+& \psi \coloneqq \min_t\Psi_t, \\
+& D \coloneqq \max_t d_t \\
+& c_m \coloneqq 0 \\
+& c_t \coloneqq c_{t+1}(1+\alpha_t\beta+3\alpha_t^2L^2), \\
+\end{align*}
+where:
+\begin{align*}
+	& \beta > 0 \\
+	& \Psi_t \coloneqq (\alpha_t - \frac{c_{t+1}\alpha_t}{\beta} - \frac{3}{2}\alpha_t^2L - 3c_{t+1}\alpha_t^2) \\
+	& d_t \coloneqq 3(\alpha_t\frac{L}{2} + c_{t+1}\alpha_t^2), \\
+\end{align*}
+and $\beta,\alpha_t$ are chosen so that $\psi_t > 0$ for $t=1,\dots,m$.
+We use two auxiliary lemmas, whose proofs are provided later on:
+%Lemma 1
+\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
+\[
+	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
+	\frac{R_{t+1}^{s+1} - R_t^{s+1} + Z_t}{\Psi_t},
+\]
+where
+\[
+	R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) + c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+\]
+\end{restatable}
+%Lemma 2
+\begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
+	...
+\end{restatable}
+
+\end{proof}
+
+%Proof of Lemma 1
+\auxone*
+\begin{proof}
+We have:
+\begin{align}
+	\Ets{J(\vtheta_{t+1}^{s+1})} 
+	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
+	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
+	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradIdeal{\vtheta_t^{s+1}}+e_s} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
+	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}  + \alpha_t\Ets[t-1]{\dotprod{e_s}{\gradJ{\vtheta_t^{s+1}}}} \\
+	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} +
+	\alpha_t\Covts[t-1]{e_s}{\gradJ{\vtheta_t^{s+1}}} \\
+	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} -
+	\alpha_t\left|\Covts[t-1]{e_s}{\gradJ{\vtheta_t^{s+1}}}\right| \\
+	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
+	&+
+	\alpha_t\frac{\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}\Varts{\gradJ{\vtheta_t^{s+1}}}}}{\sqrt{N}}
+\end{align}
+where (1) is from the L-smoothness of $J(\vtheta)$, (2) is from the SVRPG gradient ascent update, (3) is from Definition \ref{def:ideal}, (4) is from the fact that $\EVV[\tau_t\sim\pi(\cdot\mid \vtheta_t^{s+1})]{\gradBlack{\vtheta_t^{s+1}}\mid\vtheta_t^{s+1}} = \EVV[\tau_t\sim\pi(\cdot\mid \vtheta_t^{s+1})]{\gradIdeal{\vtheta_t^{s+1}}\mid\vtheta_t^{s+1}}$, (5) is from the fact that $\Ets[t-1]{e_s} = 0$, and (7)+(8) is from the definition of $\gradApp{\vtheta}{N}$.
+
 ...
 \end{proof}
+
 \end{document}

### 2018-01-18 svrpg.tex
done writing first proposition of Lemma 1

@@
 
-\section{Definitions}
-We give some additional definitions which will be useful in the proofs of Appendix \ref{app:proofs}.
+\section{Proofs}\label{app:proofs}
+
+\subsection*{Definitions}
+We give some additional definitions which will be useful in the proofs.
 \begin{definition}
@@
 
-\section{Proofs}\label{app:proofs}
+\subsection*{Lemmas}
+Before addressing the main convergence theorem, we prove some useful lemmas.
 
@@
 \end{proof}
+
+%Lemma 1
+\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
+	\[
+	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
+	\frac{R_{t+1}^{s+1} - R_t^{s+1} + Z_t}{\Psi_t},
+	\]
+	where
+	\[
+	R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) + c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	\]
+\end{restatable}
+\begin{proof}
+	We have:
+	\begin{align}
+	\Ets{J(\vtheta_{t+1}^{s+1})} 
+	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
+	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
+	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradIdeal{\vtheta_t^{s+1}}+e_s} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
+	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}  + \alpha_t\Ets[t-1]{\dotprod{e_s}{\gradJ{\vtheta_t^{s+1}}}} \nonumber\\
+	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
+	&\qquad+
+	\alpha_t\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\gradJ{\vtheta_t^{s+1}}} \\
+	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-
+	\alpha_t\Tr\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\gradJ{\vtheta_t^{s+1}}}\right| \nonumber \\
+	%
+	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-\alpha_t \Tr\left(\sqrt{\Varts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{\gradJ{\vtheta_t^{s+1}}}}\right) \nonumber \\
+	%
+	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-\frac{\alpha_t}{N} \Tr\left(\sqrt{\Varts[t-1]{\gradApp{\tilde{\vtheta}^s}{1}}}\circ\sqrt{\Varts[t-1]{\gradJ{\vtheta_t^{s+1}}}}\right) \\
+	%
+	&=
+	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(\Tr\Varts[t-1]{\gradApp{\tilde{\vtheta}^s}{1}} + \Tr\Varts[t-1]{\gradJ{\vtheta_t^{s+1}}}\right) \\
+	%
+	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(\Tr\Varts[t-1]{\gradApp{\tilde{\vtheta}^s}{1}} + \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1} - \gradJ{\tilde{\vtheta}^s}}}^2}\right) \nonumber\\
+	%
+	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-\frac{\alpha_tL^2}{2N}\left(\Tr\Varts[t-1]{\gradApp{\tilde{\vtheta}^s}{1}} + \Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}\right)
+\end{align}
+where (1) is from the L-smoothness of $J(\vtheta)$, (2) is from the SVRPG gradient ascent update, (3) is from Definition \ref{def:ideal}, (4) is from the fact that $\EVV[\tau_t\sim\pi(\cdot\mid \vtheta_t^{s+1})]{\gradBlack{\vtheta_t^{s+1}}\mid\vtheta_t^{s+1}} = \EVV[\tau_t\sim\pi(\cdot\mid \vtheta_t^{s+1})]{\gradIdeal{\vtheta_t^{s+1}}\mid\vtheta_t^{s+1}}$, (5) is from the fact that $\Ets[t-1]{e_s} = 0$, (6) is from the definition of $\gradApp{\vtheta}{N}$, (7) is from Young's inequality and (8) is again from L-smoothness.
+
+	...
+\end{proof}
+
+%Lemma 2
+\begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
+	...
+\end{restatable}
+
+\subsection*{Main theorem}
 
@@
 and $\beta,\alpha_t$ are chosen so that $\psi_t > 0$ for $t=1,\dots,m$.
-We use two auxiliary lemmas, whose proofs are provided later on:
-%Lemma 1
-\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
-\[
-	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
-	\frac{R_{t+1}^{s+1} - R_t^{s+1} + Z_t}{\Psi_t},
-\]
-where
-\[
-	R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) + c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-\]
-\end{restatable}
-%Lemma 2
-\begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
-	...
-\end{restatable}
-
-\end{proof}
-
-%Proof of Lemma 1
-\auxone*
-\begin{proof}
-We have:
-\begin{align}
-	\Ets{J(\vtheta_{t+1}^{s+1})} 
-	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
-	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
-	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradIdeal{\vtheta_t^{s+1}}+e_s} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
-	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}  + \alpha_t\Ets[t-1]{\dotprod{e_s}{\gradJ{\vtheta_t^{s+1}}}} \\
-	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} +
-	\alpha_t\Covts[t-1]{e_s}{\gradJ{\vtheta_t^{s+1}}} \\
-	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} -
-	\alpha_t\left|\Covts[t-1]{e_s}{\gradJ{\vtheta_t^{s+1}}}\right| \\
-	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
-	&+
-	\alpha_t\frac{\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}\Varts{\gradJ{\vtheta_t^{s+1}}}}}{\sqrt{N}}
-\end{align}
-where (1) is from the L-smoothness of $J(\vtheta)$, (2) is from the SVRPG gradient ascent update, (3) is from Definition \ref{def:ideal}, (4) is from the fact that $\EVV[\tau_t\sim\pi(\cdot\mid \vtheta_t^{s+1})]{\gradBlack{\vtheta_t^{s+1}}\mid\vtheta_t^{s+1}} = \EVV[\tau_t\sim\pi(\cdot\mid \vtheta_t^{s+1})]{\gradIdeal{\vtheta_t^{s+1}}\mid\vtheta_t^{s+1}}$, (5) is from the fact that $\Ets[t-1]{e_s} = 0$, and (7)+(8) is from the definition of $\gradApp{\vtheta}{N}$.
-
-...
 \end{proof}

### 2018-01-20 svrpg.tex
almost finished lemma 1

--- 2018-01-17 svrpg.tex
@@
 \mathop{\mathbb{E}}
-\left[\blacktriangledown J(\vtheta)\right] = \gradJ{}.
+\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
 \]
@@
 \end{proof}
+
+%Lemma 2
+\begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
+	...
+\end{restatable}
+
+%Lemma 0
+\begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
+The expected scalar product between $\gradBlack{\vtheta_t^{s+1}}$ and any function $f(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$ can be bounded as follows:
+\begin{align*}
+\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} &\geq
+\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \\
+&\qquad-
+\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
+\end{align*}
+\end{restatable}
+\begin{proof}
+\begin{align}
+	\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
+	&=
+	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
+	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
+	&=
+	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
+	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
+	&=
+	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
+	&\qquad+
+	\dotprod{\Ets{e_s}}{\Ets{f(\vtheta_t^{s+1})}}
+	+\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}  \nonumber\\
+	&= 
+	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
+	&\qquad+
+	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})} \\
+	&\geq  
+	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
+	&\qquad-
+	\Tr\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right| \nonumber \\
+	%
+	&\geq
+	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
+	&\qquad-\Tr\left(\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}}\right) \nonumber\\
+	%
+	&\geq	
+	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
+	&\qquad-
+	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} -\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\\
+	%
+	&=
+	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
+	&\qquad-
+	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \\
+	%
+	&\geq
+	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
+	&\qquad-
+	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},\nonumber
+\end{align}
+where (1) is from Definition \ref{def:ideal}; (2) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (3) is from $\Ets{e_s}=0$; (4) is from Young's inequality; and (5) is from the definition of $\gradApp{\vtheta}{N}$.
+\end{proof}
+
+%Lemma 1
+\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
+The expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$ can be bounded as follows:
+\[
+	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
+	\frac{R_{t+1}^{s+1} - R_t^{s+1} + Z_t}{\Psi_t},
+\]
+	where
+\[
+	R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) + c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+\]
+\end{restatable}
+\begin{proof}
+	We have:
+	\begin{align}
+	\Ets{J(\vtheta_{t+1}^{s+1})} 
+	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
+	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
+	&\geq
+	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-
+	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2},
+\end{align}
+where (6) is from the L-smoothness of $J(\vtheta)$, (7) is from the SVRPG update, and (8) is from Lemma \ref{lemma:aux0}.
+
+Next we have:
+\begin{align}
+\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} 
+&= \Ets{\norm[]{\vtheta_{t+1}^{s+1}- \vtheta_t^{s+1} + \vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+&=\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_{t}^{s+1}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\dotprod{\vtheta_{t+1}^{s+1}-\vtheta_{t}^{s+1}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber \\
+&= \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \\
+&= \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber\\ 
+&\qquad-
+\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \\
+%
+&= \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+-2\alpha_t\Ets{\left|\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}\right|} \nonumber\\ 
+&\qquad-
+\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+%
+&\geq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+-2\alpha_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber\\ 
+&\qquad-
+\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+%
+&\geq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+-2\alpha_t\Ets{\frac{1}{2\beta}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta}{2}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\ 
+&\qquad-
+\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2},
+\end{align}
+where (9) is from the SVRPG update, (10) is from Lemma \ref{lemma:aux0}, and (11) is from Young's inequality in the 'Peter-Paul' variant.
+
+...[next: put (11),(8) and Lemma A.1 together to finish proof]
+\end{proof}
+
+\subsection*{Main theorem}
 
@@
 and $\beta,\alpha_t$ are chosen so that $\psi_t > 0$ for $t=1,\dots,m$.
-We use two auxiliary lemmas, whose proofs are provided later on:
-%Lemma 1
-\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
-\[
-	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
-	\frac{R_{t+1}^{s+1} - R_t^{s+1} + Z_t}{\Psi_t},
-\]
-where
-\[
-	R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) + c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-\]
-\end{restatable}
-%Lemma 2
-\begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
-	...
-\end{restatable}
-
-\end{proof}
-
-%Proof of Lemma 1
-\auxone*
-\begin{proof}
-We have:
-\begin{align}
-	\Ets{J(\vtheta_{t+1}^{s+1})} 
-	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
-	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
-	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradIdeal{\vtheta_t^{s+1}}+e_s} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
-	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}  + \alpha_t\Ets[t-1]{\dotprod{e_s}{\gradJ{\vtheta_t^{s+1}}}} \\
-	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} +
-	\alpha_t\Covts[t-1]{e_s}{\gradJ{\vtheta_t^{s+1}}} \\
-	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} -
-	\alpha_t\left|\Covts[t-1]{e_s}{\gradJ{\vtheta_t^{s+1}}}\right| \\
-	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
-	&+
-	\alpha_t\frac{\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}\Varts{\gradJ{\vtheta_t^{s+1}}}}}{\sqrt{N}}
-\end{align}
-where (1) is from the L-smoothness of $J(\vtheta)$, (2) is from the SVRPG gradient ascent update, (3) is from Definition \ref{def:ideal}, (4) is from the fact that $\EVV[\tau_t\sim\pi(\cdot\mid \vtheta_t^{s+1})]{\gradBlack{\vtheta_t^{s+1}}\mid\vtheta_t^{s+1}} = \EVV[\tau_t\sim\pi(\cdot\mid \vtheta_t^{s+1})]{\gradIdeal{\vtheta_t^{s+1}}\mid\vtheta_t^{s+1}}$, (5) is from the fact that $\Ets[t-1]{e_s} = 0$, and (7)+(8) is from the definition of $\gradApp{\vtheta}{N}$.
 

### 2018-01-20 svrpg.tex
Working on preliminaries

--- 2018-01-18 svrpg.tex
@@
 \DeclareRobustCommand{\wrt}{w.r.t.\@\xspace}
+% graphics
+\usepackage[usenames,dvipsnames]{xcolor}
+\usepackage{tikz,pgfplots}
+\pgfplotsset{compat=newest}
+%% TODOs
+% \setlength{\marginparwidth}{2cm}
+\usepackage[colorinlistoftodos, textwidth=20mm]{todonotes}
+\definecolor{citrine}{rgb}{0.89, 0.82, 0.04}
+\definecolor{blued}{RGB}{70,197,221}
+%%todo by Marcello
+\newcommand{\todomarc}[1]{\todo[color=green, inline]{\small #1}}
+\newcommand{\todomarcout}[1]{\todo[color=green]{\scriptsize #1}}
+%%todo by Matteo Papini
+\newcommand{\todomat}[1]{\todo[color=citrine, inline]{\small #1}}
+\newcommand{\todomatout}[1]{\todo[color=citrine]{\scriptsize #1}}
+%%todo by Matteo Pirotta
+\newcommand{\todopir}[1]{\todo[color=blued, inline]{\small #1}}
+\newcommand{\todopirout}[1]{\todo[color=blued]{\scriptsize #1}}
 %%%%%%
@@
 \begin{abstract}
-
+Why we need variance-reduced gradient in RL?
 \end{abstract}
@@
 \section{Introduction}
+\begin{itemize}
+        \item Similarities between RL and SL
+        \item SL needs SVRG so RL
+        \item Potential advantages of SVRG
+        \item Difficulties in RL \wrt SL
+\end{itemize}
 
 \section{Preliminaries}
-In this section we provide essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite sum optimization.
+In this section we provide essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
 In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
@@
 \subsection{Policy Gradient}
-A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'\mid s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
-The agent's behavior is modeled as a policy $\pi$, where $\pi(a\mid s)$ is the density distribution over $\Aspace$ of the action to take in state $s$.
-We consider episodic tasks, \ie, tasks composed of episodes of length $H$, also called time horizon.
+A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
+The agent's behavior is modeled as a policy $\pi$, where $\pi(\cdot|s)$ is the density distribution over $\Aspace$ in state $s$.
+We consider episodic tasks, \ie tasks composed of episodes of length $H$, also called time horizon.
 In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
+\todopir{This definition of episodic task is not precise!
+        An episodic task is a task in which there is an absorbing state (at least one) that is optimal.
+This means that the optimal policy reaches the target state in a finite number of steps.
+A part from that, there is no other property on the problem.
+The previous sentences may be misinterpreted and can lead to think that the problem is finite horizon but in that case we know that the optimal policy is non-stationary (bad)!
+        I would rewrite it in this way:
+        We consider episodic MDPs with effective horizon $H$, \ie optimal policy is able to reach the target state (\ie absorbing state) in at most $H$ steps (the true length depend on the initial state).
+This has not to be confused with a finite horizon problem where the optimal policy is non-stationary.
+In this setting, we can limit our attention to trajectories of length $H$.
+A trajectory $\tau$ is a sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following a stationary policy, where $s_0 \sim \rho$.
+}
 With abuse of notation, we denote with $\pi(\tau)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories, and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
@@
 \end{align*}
-Solving an MDP means finding a policy $\pi^*$ maximizing $J$.
-\par
+Solving an MDP means finding a policy $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
+
 Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable in the $m$-dimensional parameter vector $\vtheta$.

### 2018-01-20 svrpg.tex
written Lemma 2

--- 2018-01-20 svrpg.tex
@@
 \newcommand{\trajdistr}{\pi_{\vtheta}(\tau)}
-\newcommand{\score}[2]{\nabla_{#1}\log\pi_{#1}(#2)}
+\newcommand{\score}[2]{\nabla\log\pi_{#1}(#2)}
 \newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
@@
 \begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
-	...
+The expected squared norm of the SVRPG gradient can be bounded as follows:
+\[
+\Ets{\gradBlack{\vtheta_t^{s+1}}} \leq
+\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
++L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
++\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+\nonumber 
++dG^2R^2\left(M_2-1\right)
+\]
 \end{restatable}
+\begin{proof}
+
+	\begin{align}
+	\Ets{\gradBlack{\vtheta_t^{s+1}}} 
+	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
+			+ \score{\vtheta_t^{s+1}}{\tau_i} -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber\\
+	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
+			+ \score{\vtheta_t^{s+1}}{\tau_i} -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}-\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
+			+\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}}^2} \nonumber\\
+	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
+	+\Ets[]{\norm[]{\gradApp{\tilde{\vtheta}^s}{N} - \Ets[]{\gradApp{\tilde{\vtheta}^s}{N}}}^2} \nonumber\\
+	&\qquad+ 
+	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i} -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}- \Ets{\score{\vtheta_t^{s+1}}{\tau_i} -
+				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}}^2} \nonumber\\
+	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	\\
+	&\qquad+ 
+	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i} -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}- \Ets{\score{\vtheta_t^{s+1}}{\tau_i} -
+				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}}^2} \nonumber\\
+	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i} -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \\
+	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	\nonumber\\
+	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}
+			-\score{\tilde{\vtheta}^s}{\tau_i}
+			+\score{\tilde{\vtheta}^s}{\tau_i} 
+			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber\\
+	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	\nonumber\\
+	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}
+			-\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber
+	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i} 
+			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber\\
+	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	\nonumber\\
+	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i} 
+			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \\
+	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	\nonumber\\
+	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+dG^2R^2\Ets{(\omega(\tau_i)-1)^2} \\
+	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	\nonumber\\
+	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+dG^2R^2\left(\Ets{\omega^2(\tau_i)}-1\right) \nonumber\\
+	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	\nonumber\\
+	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+dG^2R^2\left(M_2-1\right),
+\end{align}
+where (1) is from the definition of $\gradApp{\vtheta}{N}$, (2) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (3) is from L-smoothness, 
+(4) is from [todo], and (5) is from [todo].
+\end{proof}
 
@@
 \end{align}
-where (1) is from Definition \ref{def:ideal}; (2) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (3) is from $\Ets{e_s}=0$; (4) is from Young's inequality; and (5) is from the definition of $\gradApp{\vtheta}{N}$.
+where (6) is from Definition \ref{def:ideal}; (7) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (8) is from $\Ets{e_s}=0$; (9) is from Young's inequality; and (10) is from the definition of $\gradApp{\vtheta}{N}$.
 \end{proof}
@@
 \end{align}
-where (6) is from the L-smoothness of $J(\vtheta)$, (7) is from the SVRPG update, and (8) is from Lemma \ref{lemma:aux0}.
+where (11) is from the L-smoothness of $J(\vtheta)$, (12) is from the SVRPG update, and (13) is from Lemma \ref{lemma:aux0}.
 
@@
 \end{align}
-where (9) is from the SVRPG update, (10) is from Lemma \ref{lemma:aux0}, and (11) is from Young's inequality in the 'Peter-Paul' variant.
+where (14) is from the SVRPG update, (15) is from Lemma \ref{lemma:aux0}, and (16) is from Young's inequality in the 'Peter-Paul' variant.
 
-...[next: put (11),(8) and Lemma A.1 together to finish proof]
+...[next: put (16),(13) and Lemma A.1 together to finish proof]
 \end{proof}

### 2018-01-21 svrpg.tex
Inserted pratical experiment

@@
 \section{Experiments}
+We design our experiments to investigate if:
+\begin{enumerate}
+\item Algorithm reaches high value of the reward function faster then GPOMDP.
+\item Averagely SVRPG maintains the reward function more stable then GPOMDP.
+\end{enumerate}
+
+In order to analyze these topics, we evaluate the behavior of rewards function mediated by 10 runs for each experiment.
+
+We decide to use the following models:
+\begin{enumerate}
+\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, the cart velocity $\dot{x}$ , and the pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body.
+Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$.
+Episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time step T is major than 100.
+\item \emph{Mujoco Swimmer}: 8-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
+2-dimensional action space: the two momentums applied on actuated joints.
+Reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2] a^2$.
+Episodes terminates when the number of time step T is major than 500.
+\item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles.
+6-dimensional action space: the 6 momentums applied on actuated joints.
+Reward function is defined as $r(s, a) := v_x - 0.05\norm[2] a^2$.
+Episodes terminates when the number of time step T is major than 500.
+\end{enumerate}
+
+In all the experiments we face with GPOMDP standard algorithm. GPOMDP uses a number of trajectories per batch that is equal to the number of trajectories used in SVRPO mini-batch size.
+In SVRPO we implemented two different Adam parameters: the first one for Snapshot and the second one for sub-iterations.
+The stopping condition depends on the maximum number of sub-iterations and on the step size of sub-iteration: in particular we stop if the ratio of step size to number of trajectories used results smaller in sub-iteration than in snapshot.
+We fixed the maximum number of trajectories that both the algorithms can use.
+
+All the experiments' parameters are reported in the following table:
+
+\begin{tabular}{ l | c  c  c }	
+  & Cart-Pole & Swimmer & Half Cheetah \\
+  \hline
+  State space dim  & 4 & 8 & 20 \\
+  Control space dim & 1 & 2 & 6 \\
+  NN Hidden layers & 8 & 32x32 & 32x32 \\
+  Learning rate & $10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
+  Snapshot traj & 100 & 100 & 100 \\
+  Sub-iter traj & 10 & 10 & 10 \\
+  Max sub-iter & 50 & 20 & 20 \\
+  Traj lengths& 100 & 500 & 500 \\
+  Discount $(\gamma)$& 0.99 & 0.995 & 0.99 \\
+  Total traj& 10000 & 10000 & 10000 \\
+  \hline  
+\end{tabular}
 

### 2018-01-22 svrpg.tex
Modified experiment part and added the experimental details section into appendices

@@
 
-We decide to use the following models:
-\begin{enumerate}
-\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, the cart velocity $\dot{x}$ , and the pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body.
-Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$.
-Episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time step T is major than 100.
-\item \emph{Mujoco Swimmer}: 8-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
-2-dimensional action space: the two momentums applied on actuated joints.
-Reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2] a^2$.
-Episodes terminates when the number of time step T is major than 500.
-\item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles.
-6-dimensional action space: the 6 momentums applied on actuated joints.
-Reward function is defined as $r(s, a) := v_x - 0.05\norm[2] a^2$.
-Episodes terminates when the number of time step T is major than 500.
-\end{enumerate}
+We decide to use the following models of Rllab: \emph{Cart-Pole Balancing} ,\emph{Mujoco Swimmer},\emph{Mujoco Half Cheetah}.
 
-In all the experiments we face with GPOMDP standard algorithm. GPOMDP uses a number of trajectories per batch that is equal to the number of trajectories used in SVRPO mini-batch size.
-In SVRPO we implemented two different Adam parameters: the first one for Snapshot and the second one for sub-iterations.
+In all the experiments we face with GPOMDP standard algorithm. GPOMDP uses a number of trajectories per batch that is equal to the number of trajectories used in SVRPG mini-batch size.
+The neural network and the learning rate are the same for GPOMDP and SVRPG.
+In SVRPG we implemented two different Adam parameters: the first one for Snapshot and the second one for sub-iterations.
 The stopping condition depends on the maximum number of sub-iterations and on the step size of sub-iteration: in particular we stop if the ratio of step size to number of trajectories used results smaller in sub-iteration than in snapshot.
@@
 
-All the experiments' parameters are reported in the following table:
-
-\begin{tabular}{ l | c  c  c }	
-  & Cart-Pole & Swimmer & Half Cheetah \\
-  \hline
-  State space dim  & 4 & 8 & 20 \\
-  Control space dim & 1 & 2 & 6 \\
-  NN Hidden layers & 8 & 32x32 & 32x32 \\
-  Learning rate & $10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
-  Snapshot traj & 100 & 100 & 100 \\
-  Sub-iter traj & 10 & 10 & 10 \\
-  Max sub-iter & 50 & 20 & 20 \\
-  Traj lengths& 100 & 500 & 500 \\
-  Discount $(\gamma)$& 0.99 & 0.995 & 0.99 \\
-  Total traj& 10000 & 10000 & 10000 \\
-  \hline  
-\end{tabular}
+The neural networks is composed by one layer with 8 neurons for Cart-Pole environment and by two layers with 32x32 neurons for Swimmer and for Half Cheetah environments.
+In each experiment, the length of episodes is 500 steps.
 
@@
 
+\section{Experimental Details}\label{app:proofs}
+Environments of our experiments are:
+\begin{enumerate}
+	\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, the cart velocity $\dot{x}$ , and the pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body.
+Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$.
+Episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time step T is major than 100.
+	\item \emph{Mujoco Swimmer}: 12-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
+2-dimensional action space: the two momentums applied on actuated joints.
+Reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2] a^2$.
+Episodes terminates when the number of time step T is major than 500.
+	\item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles.
+6-dimensional action space: the 6 momentums applied on actuated joints.
+Reward function is defined as $r(s, a) := v_x - 0.05\norm[2] a^2$.
+Episodes terminates when the number of time step T is major than 500.
+\end{enumerate}
+
+All the experiments' parameters are reported in the following table:
+
+\centering
+\begin{tabular}{| l | c  c  c |}
+	\hline	
+	& Cart-Pole & Swimmer & Half Cheetah \\
+	\hline
+	State space dimension  & 4 & 8 & 20 \\
+	Control space dimension & 1 & 2 & 6 \\
+	NN Hidden layers & 8 & 32x32 & 32x32 \\
+	Learning rate & $10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
+	Snapshot trajectories & 100 & 100 & 100 \\
+	Sub-iterations trajectories/GPOMDP batch & 10 & 10 & 10 \\
+	Max sub-iterations & 50 & 20 & 20 \\
+	Trajectories lengths& 500 & 500 & 500 \\
+	Discount $(\gamma)$& 0.99 & 0.995 & 0.99 \\
+	Total trajectories& 10000 & 10000 & 10000 \\
+	\hline  
+\end{tabular}
+
 \end{document}

### 2018-01-22 svrpg.tex
Added Stopping condition section

@@
 
+\section{Stopping Condition}\label{sec:stopping}
+In practice ,the performance of the SVRPG algorithm is sensitive to the number of sub-iterations.
+As the number of sub-iterations increases, the variance of the importance weights and the gradient increases too.
+Therefore, the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
+A fixed number of sub-iterations is not the optimal choice.
+Few sub-iterations lead to few updates, many update lead to increases of variance.
+
+We introduce two stopping condition: maximum number of sub-iterations and learning rate of Adam sub-iterations.
+Adam update rule help to stabilize the gradient estimate. In fact, this method computes adaptive learning rates for each parameter that depends by history of gradients' variance.
+In particular we use different Adam parameters: the first one for snapshot update and the second one for the sub-iterations update.
+We stop if:
+
+\[\frac{\alpha_{snap}}{N}>\frac{\alpha_{sub-iter}}{M}\]
+
+where M is the sub-iterations batch size.
+
 \section{Practical Variants}
@@
 
-We decide to use the following models of Rllab: \emph{Cart-Pole Balancing} ,\emph{Mujoco Swimmer},\emph{Mujoco Half Cheetah}.
+We decide to use the following models of Rllab: \emph{Cart-Pole Balancing}, \emph{Mujoco Swimmer}, \emph{Mujoco Half Cheetah}.
 
@@
 In SVRPG we implemented two different Adam parameters: the first one for Snapshot and the second one for sub-iterations.
-The stopping condition depends on the maximum number of sub-iterations and on the step size of sub-iteration: in particular we stop if the ratio of step size to number of trajectories used results smaller in sub-iteration than in snapshot.
+We use the stopping condition defined in Section \ref{sec:stopping}.
 We fixed the maximum number of trajectories that both the algorithms can use.

### 2018-01-22 svrpg.tex
Updated

--- 2018-01-21 svrpg.tex
@@
 \newcommand{\trajdistr}{\pi_{\vtheta}(\tau)}
-\newcommand{\score}[2]{\nabla\log\pi_{#1}(#2)}
+\newcommand{\score}[2]{\nabla_{#1}\log\pi_{#1}(#2)}
 \newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
@@
 \subsection{Policy Gradient}
-A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'\mid s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
-The agent's behavior is modeled as a policy $\pi$, where $\pi(a\mid s)$ is the density distribution over $\Aspace$ of the action to take in state $s$.
-We consider episodic tasks, \ie, tasks composed of episodes of length $H$, also called time horizon.
-In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
+A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
+The agent's behavior is modeled as a policy $\pi$, where $\pi(\cdot|s)$ is the density distribution over $\Aspace$ in state $s$.
+% We consider episodic tasks, \ie tasks composed of episodes of length $H$, also called time horizon.
+% In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
+We consider episodic MDPs with effective horizon $H$, \ie optimal policy is able to reach the target state (\ie absorbing state) in at most $H$ steps (the true length depend on the initial state).
+This has not to be confused with a finite horizon problem where the optimal policy is non-stationary.
+In this setting, we can limit our attention to trajectories of length $H$.
+A trajectory $\tau$ is a sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following a stationary policy, where $s_0 \sim \rho$.
 With abuse of notation, we denote with $\pi(\tau)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories, and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
@@
 \end{align*}
-Performance $J$ allows to rank policies in terms of expected reward:
+Policies can be ranked based on their expected total reward:
 \begin{align*}
@@
 \end{align*}
-Solving an MDP means finding a policy $\pi^*$ maximizing $J$.
-\par
-Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable in the $m$-dimensional parameter vector $\vtheta$.
-We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we denote $\pol(\tau)$ with $\pi(\tau\mid\vtheta)$ for the sake of readability.
+Solving an MDP means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
+
+Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt the $m$-dimensional parameter vector $\vtheta$.
+We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
 The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
+\todopir{stochastic or mini-batch gradient}
 \begin{align*}
-\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{}{\tau}\Reward(\tau)\de \tau.
+\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
 \end{align*}
 We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment.
-In this setting, stochastic gradient descent is typically employed: after sampling a trajectory $\tau$ or a batch of $N$ trajectories $\Dataset_{N}$,
+In this setting, stochastic gradient ascent is typically employed: after sampling a trajectory $\tau$ or a batch of $N$ trajectories $\Dataset_{N}$,
 policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient.
@@
 \section{Experiments}
-We design our experiments to investigate if:
-\begin{enumerate}
-\item Algorithm reaches high value of the reward function faster then GPOMDP.
-\item Averagely SVRPG maintains the reward function more stable then GPOMDP.
-\end{enumerate}
-
-In order to analyze these topics, we evaluate the behavior of rewards function mediated by 10 runs for each experiment.
-
-We decide to use the following models:
-\begin{enumerate}
-\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, the cart velocity $\dot{x}$ , and the pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body.
-Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$.
-Episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time step T is major than 100.
-\item \emph{Mujoco Swimmer}: 8-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
-2-dimensional action space: the two momentums applied on actuated joints.
-Reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2] a^2$.
-Episodes terminates when the number of time step T is major than 500.
-\item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles.
-6-dimensional action space: the 6 momentums applied on actuated joints.
-Reward function is defined as $r(s, a) := v_x - 0.05\norm[2] a^2$.
-Episodes terminates when the number of time step T is major than 500.
-\end{enumerate}
-
-In all the experiments we face with GPOMDP standard algorithm. GPOMDP uses a number of trajectories per batch that is equal to the number of trajectories used in SVRPO mini-batch size.
-In SVRPO we implemented two different Adam parameters: the first one for Snapshot and the second one for sub-iterations.
-The stopping condition depends on the maximum number of sub-iterations and on the step size of sub-iteration: in particular we stop if the ratio of step size to number of trajectories used results smaller in sub-iteration than in snapshot.
-We fixed the maximum number of trajectories that both the algorithms can use.
-
-All the experiments' parameters are reported in the following table:
-
-\begin{tabular}{ l | c  c  c }	
-  & Cart-Pole & Swimmer & Half Cheetah \\
-  \hline
-  State space dim  & 4 & 8 & 20 \\
-  Control space dim & 1 & 2 & 6 \\
-  NN Hidden layers & 8 & 32x32 & 32x32 \\
-  Learning rate & $10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
-  Snapshot traj & 100 & 100 & 100 \\
-  Sub-iter traj & 10 & 10 & 10 \\
-  Max sub-iter & 50 & 20 & 20 \\
-  Traj lengths& 100 & 500 & 500 \\
-  Discount $(\gamma)$& 0.99 & 0.995 & 0.99 \\
-  Total traj& 10000 & 10000 & 10000 \\
-  \hline  
-\end{tabular}
 
@@
 \begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
-The expected squared norm of the SVRPG gradient can be bounded as follows:
-\[
-\Ets{\gradBlack{\vtheta_t^{s+1}}} \leq
-\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-\nonumber 
-+dG^2R^2\left(M_2-1\right)
-\]
+	...
 \end{restatable}
-\begin{proof}
-
-	\begin{align}
-	\Ets{\gradBlack{\vtheta_t^{s+1}}} 
-	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
-			+ \score{\vtheta_t^{s+1}}{\tau_i} -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber\\
-	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
-			+ \score{\vtheta_t^{s+1}}{\tau_i} -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}-\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
-			+\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}}^2} \nonumber\\
-	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
-	+\Ets[]{\norm[]{\gradApp{\tilde{\vtheta}^s}{N} - \Ets[]{\gradApp{\tilde{\vtheta}^s}{N}}}^2} \nonumber\\
-	&\qquad+ 
-	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i} -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}- \Ets{\score{\vtheta_t^{s+1}}{\tau_i} -
-				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}}^2} \nonumber\\
-	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	\\
-	&\qquad+ 
-	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i} -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}- \Ets{\score{\vtheta_t^{s+1}}{\tau_i} -
-				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}}^2} \nonumber\\
-	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
-	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i} -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \\
-	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	\nonumber\\
-	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}
-			-\score{\tilde{\vtheta}^s}{\tau_i}
-			+\score{\tilde{\vtheta}^s}{\tau_i} 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber\\
-	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	\nonumber\\
-	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}
-			-\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber
-	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i} 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber\\
-	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	\nonumber\\
-	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i} 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \\
-	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	\nonumber\\
-	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+dG^2R^2\Ets{(\omega(\tau_i)-1)^2} \\
-	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	\nonumber\\
-	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+dG^2R^2\left(\Ets{\omega^2(\tau_i)}-1\right) \nonumber\\
-	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	\nonumber\\
-	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+dG^2R^2\left(M_2-1\right),
-\end{align}
-where (1) is from the definition of $\gradApp{\vtheta}{N}$, (2) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (3) is from L-smoothness, 
-(4) is from [todo], and (5) is from [todo].
-\end{proof}
 
@@
 \end{align}
-where (6) is from Definition \ref{def:ideal}; (7) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (8) is from $\Ets{e_s}=0$; (9) is from Young's inequality; and (10) is from the definition of $\gradApp{\vtheta}{N}$.
+where (1) is from Definition \ref{def:ideal}; (2) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (3) is from $\Ets{e_s}=0$; (4) is from Young's inequality; and (5) is from the definition of $\gradApp{\vtheta}{N}$.
 \end{proof}
@@
 \end{align}
-where (11) is from the L-smoothness of $J(\vtheta)$, (12) is from the SVRPG update, and (13) is from Lemma \ref{lemma:aux0}.
+where (6) is from the L-smoothness of $J(\vtheta)$, (7) is from the SVRPG update, and (8) is from Lemma \ref{lemma:aux0}.
 
@@
 \end{align}
-where (14) is from the SVRPG update, (15) is from Lemma \ref{lemma:aux0}, and (16) is from Young's inequality in the 'Peter-Paul' variant.
+where (9) is from the SVRPG update, (10) is from Lemma \ref{lemma:aux0}, and (11) is from Young's inequality in the 'Peter-Paul' variant.
 
-...[next: put (16),(13) and Lemma A.1 together to finish proof]
+...[next: put (11),(8) and Lemma A.1 together to finish proof]
 \end{proof}

### 2018-01-22 svrpg.tex
Added comments and changed minor things

@@
 \usepackage{pifont}
+
 %%Theorems
@@
 We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment. 
-In this setting, stochastic gradient ascent is typically employed: after sampling a trajectory $\tau$ or a batch of $N$ trajectories $\Dataset_{N}$,
+In this setting, $(N>0)$-batch gradient ascent is typically employed: after sampling a batch of $N$ trajectories (\ie $\Dataset_{N}$),
 policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient. 
-The REINFORCE gradient estimator \cite{williams1992simple} provides a simple, unbiased way of estimating the gradient:
+\todopir{step size is scalar or vectorial (just for notation)? Check consistency of $N$ with finite-sum function, we can use even $b$ or $B$ for the batch size.}
+The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
 \begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
 \end{align*}
 where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
-The G(PO)MDP gradient estimator \cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
+\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,anonymous2018variance}.}
+The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
 \begin{align*}
@@
 \end{align*}
+\todopir{Cite Jan's paper somewhere}
+\todopir{At this point we need to connect with the supervised learning framework (we can just define everything inside the first sum as $\nabla g(\tau)$).}
 
@@
 A common requirement is that $g$ must be smooth and convex in $\vtheta$.
-Under this hypothesis, full gradient ascent \cite{cauchy1847methode} with a constant step size achieves an IFO complexity \cite{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.
+Under this hypothesis, full gradient ascent \cite{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.
+\todopir{Explain IFO meaning (footnote) and this holds for deterministic algorithms}
 This can be too expensive for large values of $N$.
-Stochastic gradient ascent \cite{robbins1951stochastic} \cite{bottou2004large} overcomes this problem by sampling a single $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling.
+Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling.
 The resulting IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$.
-Starting from SAG \cite{roux2012stochastic}, a series of variations to stochastic gradient ascent have been proposed to achieve a better trade-off between convergence speed and cost per iteration.
-This family of algorithms includes SAG, SVRG \cite{johnson2013accelerating}, SAGA \cite{defazio2014saga} and Finito \cite{defazio2014finito}, the common intuition being the reuse of previous gradient computations to reduce the variance of the current one.
-In particular, Stochastic Variance Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which can become a problem when neural networks with several parameters are employed.
+\todopir{What is the reference?}
+Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga} and Finito~\cite{defazio2014finito},
+The common idea is to reuse past gradient computations to reduce the variance of the current estimate.
+In particular, Stochastic Variance Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which can become a problem when deep and/or wide neural networks are employed.
 
 The idea of SVRG is to alternate full and stochastic gradient updates. 
+\todopir{here would be nice to have a notation for the single batch}
 Each $m = O(N)$ iterations, a snapshot $\tilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\tilde{\vtheta})$.
@@
 	\nabla f(\tilde{\vtheta}) + 
-	\nabla g(x_i\mid\vtheta) -
-	\nabla g(x_i \mid \tilde{\vtheta}),
+	\nabla g(x_i | \vtheta) -
+	\nabla g(x_i | \tilde{\vtheta}),
 \end{align*} 
@@
 
-More recently, some extensions of variance reduction algorithms to the non-convex objectives have been proposed \cite{reddi2016stochastic} \cite{allen2016variance} \cite{reddi2016stochastic} \cite{reddi2016fast}.
+More recently, some extensions of variance reduction algorithms to the non-convex objectives have been proposed~\citep[\eg][]{reddi2016stochastic,allen2016variance,reddi2016stochastic,reddi2016fast}.
 In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
-Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the convex case \cite{nesterov2013introductory} \cite{ghadimi2013stochastic}.
-Also in this case, SVRG yields an advantageous trade-off, with an IFO complexity of $O(N + \nicefrac{N^{\nicefrac{2}{3}}}{N})$ \cite{reddi2016stochastic}.
+Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the convex case~\citep{nesterov2013introductory,ghadimi2013stochastic}.
+Also in this case, SVRG yields an advantageous trade-off, with an IFO complexity of $O(N + \nicefrac{N^{\nicefrac{2}{3}}}{N})$~\citep{reddi2016stochastic}.
 The only additional requirement is to select $\vtheta^*$ uniformly at random among all the $\vtheta_k$ instead of simply setting it to the final value.
-Pseudocode of SVRG for both the convex and the non-convex case is provided in Algorithm \ref{alg:svrg}.
+Pseudocode of SVRG for both the convex and the non-convex case is provided in Alg.~\ref{alg:svrg}.
 

### 2018-01-23 svrpg.tex
finally done writing the lemmas

--- 2018-01-22 svrpg.tex
@@
 \DeclareRobustCommand{\wrt}{w.r.t.\@\xspace}
-% graphics
-\usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{tikz,pgfplots}
-\pgfplotsset{compat=newest}
-%% TODOs
-% \setlength{\marginparwidth}{2cm}
-\usepackage[colorinlistoftodos, textwidth=20mm]{todonotes}
-\definecolor{citrine}{rgb}{0.89, 0.82, 0.04}
-\definecolor{blued}{RGB}{70,197,221}
-%%todo by Marcello
-\newcommand{\todomarc}[1]{\todo[color=green, inline]{\small #1}}
-\newcommand{\todomarcout}[1]{\todo[color=green]{\scriptsize #1}}
-%%todo by Matteo Papini
-\newcommand{\todomat}[1]{\todo[color=citrine, inline]{\small #1}}
-\newcommand{\todomatout}[1]{\todo[color=citrine]{\scriptsize #1}}
-%%todo by Matteo Pirotta
-\newcommand{\todopir}[1]{\todo[color=blued, inline]{\small #1}}
-\newcommand{\todopirout}[1]{\todo[color=blued]{\scriptsize #1}}
 %%%%%%
@@
 \begin{abstract}
-Why we need variance-reduced gradient in RL?
+
 \end{abstract}
@@
 \section{Introduction}
-\begin{itemize}
-        \item Similarities between RL and SL
-        \item SL needs SVRG so RL
-        \item Potential advantages of SVRG
-        \item Difficulties in RL \wrt SL
-\end{itemize}
 
 \section{Preliminaries}
-In this section we provide essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
+In this section we provide essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite sum optimization.
 In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
@@
 \subsection{Policy Gradient}
-A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
-The agent's behavior is modeled as a policy $\pi$, where $\pi(\cdot|s)$ is the density distribution over $\Aspace$ in state $s$.
-% We consider episodic tasks, \ie tasks composed of episodes of length $H$, also called time horizon.
-% In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
-We consider episodic MDPs with effective horizon $H$, \ie optimal policy is able to reach the target state (\ie absorbing state) in at most $H$ steps (the true length depend on the initial state).
-This has not to be confused with a finite horizon problem where the optimal policy is non-stationary.
-In this setting, we can limit our attention to trajectories of length $H$.
-A trajectory $\tau$ is a sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following a stationary policy, where $s_0 \sim \rho$.
+A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'\mid s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
+The agent's behavior is modeled as a policy $\pi$, where $\pi(a\mid s)$ is the density distribution over $\Aspace$ of the action to take in state $s$.
+We consider episodic tasks, \ie, tasks composed of episodes of length $H$, also called time horizon.
+In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
 With abuse of notation, we denote with $\pi(\tau)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories, and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
@@
 \end{align*}
-Policies can be ranked based on their expected total reward:
+Performance $J$ allows to rank policies in terms of expected reward:
 \begin{align*}
@@
 \end{align*}
-Solving an MDP means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
-
-Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt the $m$-dimensional parameter vector $\vtheta$.
-We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
+Solving an MDP means finding a policy $\pi^*$ maximizing $J$.
+\par
+Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable in the $m$-dimensional parameter vector $\vtheta$.
+We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we denote $\pol(\tau)$ with $\pi(\tau\mid\vtheta)$ for the sake of readability.
 The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
-\todopir{stochastic or mini-batch gradient}
 \begin{align*}
-\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
+\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{}{\tau}\Reward(\tau)\de \tau.
 \end{align*}
 We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment.
-In this setting, stochastic gradient ascent is typically employed: after sampling a trajectory $\tau$ or a batch of $N$ trajectories $\Dataset_{N}$,
+In this setting, stochastic gradient descent is typically employed: after sampling a trajectory $\tau$ or a batch of $N$ trajectories $\Dataset_{N}$,
 policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient.
@@
 \section{Convergence Guarantees}
+\begin{assumption}\label{ass:bounded_score}
+For each trajectory $\tau$, value of $\vtheta$ and gradient component:
+\[
+	|\score{\vtheta}{\tau}| \leq G
+\]
+\end{assumption}
+
+\begin{assumption}\label{ass:M2}
+	There is a constant $M_2<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
+\[
+	\mathbb{V}ar\left(\omega(\tau)\right) \leq (M_2-1)
+\]
+\end{assumption}
+
+\begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
+Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth with Lipschitz constant:
+\[
+	L = \frac{2RG^2}{1-\gamma}
+\]
+\end{restatable}
 
 \begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
-Provided $J(\vtheta)$ is L-smooth with Lipschitz constant $L$ and step sizes are sufficiently small, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some constants $\psi>0, D>0$ and $c_{t}>0$ for $t=1,\dots,m$, the following property:
-\[
-	\EVV[]
-	{\norm[]{\nabla J(\vtheta_A)}^2} \leq
-		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
-		\frac{Z}{N},
-\]
-where
+Under Assumptions \ref{ass:bounded_score} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some constants $\psi>0, D>0$ and $F>0$ and for proper choice of step sizes $\alpha_t$, the following property:
 \begin{align*}
-	Z &= D\max_{s}\EVV[]{\norm[]{\gradApp{\tilde{\vtheta^s}}{1}}^2} \\
-	&+ \max_{s,t}\left|\Tr\Cov\left[
-		\gradApp{\tilde{\vtheta^s}}{1},
-		c_{t+1}\vtheta_t^{s+1} + \alpha_t\nabla J(\vtheta_t^{s+1})\right]\right|
+	&\EVV[]
+	{\norm[]{\nabla J(\vtheta_A)}^2} 
+		\leq
+		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} \\
+		&\qquad+
+		\frac{D\max_s\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}}{\psi N}
+		+\frac{F(M_2-1)}{\psi}
 \end{align*}
@@
 \begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
-The expected squared norm of the SVRPG gradient can be bounded as follows:
+Under Assumption \ref{ass:bounded_score}
+, the expected squared norm of the SVRPG gradient can be bounded as follows:
 \[
@@
 	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
-			+ \score{\vtheta_t^{s+1}}{\tau_i} -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber\\
+			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
 	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
-			+ \score{\vtheta_t^{s+1}}{\tau_i} -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}-\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
+			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)-\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
 			+\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}}^2} \nonumber\\
@@
 	&\qquad+ 
-	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i} -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}- \Ets{\score{\vtheta_t^{s+1}}{\tau_i} -
-				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}}^2} \nonumber\\
+	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)- \Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}}^2} \nonumber\\
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 	&\qquad+ 
-	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i} -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}- \Ets{\score{\vtheta_t^{s+1}}{\tau_i} -
-				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}}^2} \nonumber\\
+	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)- \Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}}^2} \nonumber\\
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
 	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
-	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i} -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \\
+	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
 	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 	\nonumber\\
-	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}
-			-\score{\tilde{\vtheta}^s}{\tau_i}
-			+\score{\tilde{\vtheta}^s}{\tau_i} 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber\\
+	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
+			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
+			+\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
+			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 	\nonumber\\
-	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}
-			-\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber
-	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i} 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber\\
+	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
+			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber
+	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
+			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i} 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \\
+	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
+			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+dG^2R^2\Ets{(\omega(\tau_i)-1)^2} \\
+	+G^2\frac{R^2}{(1-\gamma)^2}\Ets{(\omega(\tau_i)-1)^2} \\
 	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+dG^2R^2\left(\Ets{\omega^2(\tau_i)}-1\right) \nonumber\\
+	+G^2\frac{R^2}{(1-\gamma)^2}\left(\Ets{\omega^2(\tau_i)}-1\right) \nonumber\\
 	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+dG^2R^2\left(M_2-1\right),
+	+G^2\frac{R^2}{(1-\gamma)^2}\left(M_2-1\right),
 \end{align}
 where (1) is from the definition of $\gradApp{\vtheta}{N}$, (2) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (3) is from L-smoothness, 
-(4) is from [todo], and (5) is from [todo].
+(4) is from Assumption \ref{ass:bounded_score}, and (5) is from Assumption \ref{ass:M2}.
 \end{proof}
@@
 \begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
-The expected scalar product between $\gradBlack{\vtheta_t^{s+1}}$ and any function $f(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$ can be bounded as follows:
+Under Assumption \ref{ass:bounded_score},
+the expected scalar product between $\gradBlack{\vtheta_t^{s+1}}$ and any function $f(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$ can be bounded as follows:
 \begin{align*}
@@
 \begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
-The expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$ can be bounded as follows:
+Under Assumptions \ref{ass:bounded_score} ans \ref{ass:M2}, the expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$, for appropriate choices of $\alpha_t\geq0$ and $\beta_t>0$, can be bounded as follows:
 \[
 	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
-	\frac{R_{t+1}^{s+1} - R_t^{s+1} + Z_t}{\Psi_t},
+	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_t\Tr\Varts{\gradApp{\tilde{\vtheta}^s}{1}}}{N\Psi_t}
+	+\frac{f_t(M_2-1)}{\Psi_t},
 \]
 	where
-\[
-	R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) + c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-\]
+\begin{align*}
+	&R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \\
+	&c_{m-1} = 0, \\
+	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\alpha_t^2\right) + \alpha_tL^2+\frac{\alpha_t^2L^3}{2}, \\
+	&\Psi_t = \alpha_t\left(1-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
+	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right), \\
+	&f_t = \alpha_t^2\frac{G^2R^2(L+2c_{t+1})}{2(1-\gamma)^2}.
+\end{align*}
+More precisely, the following constraints on $\alpha_t$ and $\beta_t$ are sufficient:
+\begin{align*}
+&0\leq\alpha_t < \frac{2(1-\nicefrac{c_{t+1}}{\beta_t})}{L+2c_{t+1}} \\
+&\beta_t > c_{t+1}.
+\end{align*}
 \end{restatable}
@@
 	&\geq
-	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-
@@
 &= \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \\
-&= \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber\\ 
-&\qquad-
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \\
+&\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber\\ 
+&\qquad+
+\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \\
 %
-&= \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
--2\alpha_t\Ets{\left|\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}\right|} \nonumber\\ 
-&\qquad-
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+&\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
++2\alpha_t\Ets{\left|\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}\right|} \nonumber\\ 
+&\qquad+
+\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
 %
-&\geq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
--2\alpha_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber\\ 
-&\qquad-
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+&\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
++2\alpha_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber\\ 
+&\qquad+
+\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
 %
-&\geq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
--2\alpha_t\Ets{\frac{1}{2\beta}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta}{2}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\ 
-&\qquad-
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2},
+&\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
++2\alpha_t\Ets{\frac{1}{2\beta}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta}{2}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\ 
+&\qquad
+\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2},
 \end{align}
 where (14) is from the SVRPG update, (15) is from Lemma \ref{lemma:aux0}, and (16) is from Young's inequality in the 'Peter-Paul' variant.
+Finally:
+\begin{align}
+	R_{t+1}^{s+1} &= \Ets{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+	%
+	&\geq	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-
+	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
+	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \\
+	%
+	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
+	-\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	&\qquad -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
+	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+	%
+	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
+	-\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	&\qquad -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2} -c_{t+1}\Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	\nonumber\\
+	&\qquad-2c_{t+1}\alpha_t\Ets{\frac{1}{2\beta}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta}{2}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\ 
+	&\qquad
+	-c_{t+1}\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \\
+	%
+	&= \Ets{J(\vtheta_t^{s+1})} - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta\right)+\alpha_tL^2\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	&\qquad-\alpha_t^2\left(\frac{L}{2}+c_{t+1}\right)\Ets{\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}
+	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	%
+	&\geq  \Ets{J(\vtheta_t^{s+1})} - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta\right)+\alpha_tL^2\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	&\qquad
+	-\alpha_t^2\left(\frac{L}{2}+c_{t+1}\right)\left(\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	\right.\nonumber\\
+	&\left.\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+G^2\frac{R^2}{(1-\gamma)^2}\left(M_2-1\right)\right)
+	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \\
+	%
+	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta+\alpha_t^2\right)+\alpha_tL^2+\frac{\alpha_t^2L^3}{2}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	&\qquad
+	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	&\qquad-\alpha_t^2\frac{G^2R^2(L+2c_{t+1})}{2(1-\gamma)^2}(M_2-1) \\
+	%
+	&= R_t^{s+1}
+	+\Psi_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
+	-\frac{d_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	-f_t(M_2-1),
+\end{align}
 
-...[next: put (16),(13) and Lemma A.1 together to finish proof]
+To complete the proof, besides rearranging terms, we have to ensure that $\Psi_t>0$ for each $t$. This gives the constraints on $\alpha_t$ and $\beta_t$.
 \end{proof}
@@
 & D \coloneqq \max_t d_t \\
-& c_m \coloneqq 0 \\
-& c_t \coloneqq c_{t+1}(1+\alpha_t\beta+3\alpha_t^2L^2), \\
+& F \coloneqq \max_t f_t.
 \end{align*}
-where:
-\begin{align*}
-	& \beta > 0 \\
-	& \Psi_t \coloneqq (\alpha_t - \frac{c_{t+1}\alpha_t}{\beta} - \frac{3}{2}\alpha_t^2L - 3c_{t+1}\alpha_t^2) \\
-	& d_t \coloneqq 3(\alpha_t\frac{L}{2} + c_{t+1}\alpha_t^2), \\
-\end{align*}
-and $\beta,\alpha_t$ are chosen so that $\psi_t > 0$ for $t=1,\dots,m$.
+The values of all other constants and the constraints on the step size are provided in Lemma \ref{lemma:aux1}.
 

### 2018-01-23 svrpg.tex

@@
 \end{align*}
-Solving an MDP means finding a policy $\pi^*$ maximizing $J$.
-\par
-Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable in the $m$-dimensional parameter vector $\vtheta$.
-We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we denote $\pol(\tau)$ with $\pi(\tau\mid\vtheta)$ for the sake of readability.
+Solving an MDP means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
+
+Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt the $m$-dimensional parameter vector $\vtheta$.
+We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
 The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
+\todopir{stochastic or mini-batch gradient}
 \begin{align*}
-\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{}{\tau}\Reward(\tau)\de \tau.
+\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
 \end{align*}
 We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment. 
-In this setting, stochastic gradient descent is typically employed: after sampling a trajectory $\tau$ or a batch of $N$ trajectories $\Dataset_{N}$,
+In this setting, $(N>0)$-batch gradient ascent is typically employed: after sampling a batch of $N$ trajectories (\ie $\Dataset_{N}$),
 policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient. 
-The REINFORCE gradient estimator \cite{williams1992simple} provides a simple, unbiased way of estimating the gradient:
+\todopir{step size is scalar or vectorial (just for notation)? Check consistency of $N$ with finite-sum function, we can use even $b$ or $B$ for the batch size.}
+The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
 \begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
 \end{align*}
 where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
-The G(PO)MDP gradient estimator \cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
+\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,anonymous2018variance}.}
+The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
 \begin{align*}

### 2018-01-24 svrpg.tex
added some explanations to proof
changed minor details in proofs; next:minibatch

@@
 \begin{assumption}\label{ass:M2}
-	There is a constant $M_2<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
+	There is a constant $\Omega<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
 \[
-	\mathbb{V}ar\left(\omega(\tau)\right) \leq (M_2-1)
+	\mathbb{V}ar\left(\omega(\tau)\right) \leq \Omega
 \]
@@
 		\frac{D\max_s\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}}{\psi N}
-		+\frac{F(M_2-1)}{\psi}
+		+\frac{F\Omega}{\psi}
 \end{align*}
@@
 \begin{proof}
-
 	\begin{align}
@@
 			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
-	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
+	%
+	&= \mathbb{E}_{t\mid s}\left[\left\|\gradApp{\tilde{\vtheta}^s}{N}
 			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)-\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
-			+\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}}^2} \nonumber\\
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
+			\right.\right.\nonumber\\&\qquad\left.\left.
+			-\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
+			+\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}\right\|^2\right] \nonumber\\
+	%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
@@
 	&\qquad+ 
-	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)- \Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}}^2} \nonumber\\
+	\mathbb{E}_{t\mid s}\left[\left\|\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
+		\right.\right.\nonumber\\&\qquad\left.\left.
+		- \Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}\right\|^2\right] 
+	\nonumber\\
+	%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 	&\qquad+ 
-	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)- \Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}}^2} \nonumber\\
+		\mathbb{E}_{t\mid s}\left[\left\|\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
+			\right.\right.\nonumber\\&\qquad\left.\left.
+			-\Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}\right\|^2\right]  \nonumber\\
+	%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
+	%
 	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
+	%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
-			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber
+			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
+	&\qquad
 	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
 			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
+	%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2}{(1-\gamma)^2}\left(M_2-1\right),
+	+G^2\frac{R^2\Omega}{(1-\gamma)^2},
 \end{align}
@@
 \begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
-Under Assumption \ref{ass:bounded_score},
-the expected scalar product between $\gradBlack{\vtheta_t^{s+1}}$ and any function $f(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$ can be bounded as follows:
+Under Assumption \ref{ass:bounded_score}, for any function $f(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$:
 \begin{align*}
-\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} &\geq
-\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \\
-&\qquad-
-\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
+\left|\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
+-\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
+\right|
+\leq
+\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
 \end{align*}
@@
 	&\qquad+
-	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})} \\
-	&\geq  
-	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
-	&\qquad-
+	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}
+\end{align}
+Hence:
+\begin{align}
+	\left|\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
+	\right.&-\left.\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}\right| 
+	=
+	\left|\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right|  
+	\nonumber\\
+	&\leq
 	\Tr\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right| \nonumber \\
 	%
-	&\geq
-	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
-	&\qquad-\Tr\left(\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}}\right) \nonumber\\
+	&\leq
+	\Tr\left(\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}}\right) \nonumber\\
 	%
-	&\geq	
-	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
-	&\qquad-
-	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} -\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\\
+	&\leq	
+	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\\
 	%
 	&=
-	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
-	&\qquad-
-	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \\
+	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \\
 	%
-	&\geq
-	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
-	&\qquad-
-	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},\nonumber
+	&\leq
+	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},
+	\nonumber
 \end{align}
@@
 	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_t\Tr\Varts{\gradApp{\tilde{\vtheta}^s}{1}}}{N\Psi_t}
-	+\frac{f_t(M_2-1)}{\Psi_t},
+	+\frac{f_t\Omega}{\Psi_t},
 \]
@@
 \end{align*}
-More precisely, the following constraints on $\alpha_t$ and $\beta_t$ are sufficient:
+In particular, the following constraints on $\alpha_t$ and $\beta_t$ are sufficient:
 \begin{align*}
@@
 	&\left.\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2}{(1-\gamma)^2}\left(M_2-1\right)\right)
+	+G^2\frac{R^2}{(1-\gamma)^2}\Omega\right)
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
@@
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
-	&\qquad-\alpha_t^2\frac{G^2R^2(L+2c_{t+1})}{2(1-\gamma)^2}(M_2-1) \\
+	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2\Omega}{2(1-\gamma)^2} \nonumber\\
 	%
@@
 	-\frac{d_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	-f_t(M_2-1),
+	-f_t\Omega,\nonumber
 \end{align}
-
+where (17) is from (13), (18) is from (16), and (19) is from Lemma \ref{lemma:aux0}.
 To complete the proof, besides rearranging terms, we have to ensure that $\Psi_t>0$ for each $t$. This gives the constraints on $\alpha_t$ and $\beta_t$.

### 2018-01-24 svrpg.tex
Auto stash before merge of "master" and "origin/master"

@@
 We introduce two stopping condition: maximum number of sub-iterations and learning rate of Adam sub-iterations.
-Adam update rule help to stabilize the gradient estimate. In fact, this method computes adaptive learning rates for each parameter that depends by history of gradients' variance.
-In particular we use different Adam parameters: the first one for snapshot update and the second one for the sub-iterations update.
+Adam update rule help to stabilize the gradient estimate. In fact, this method computes adaptive learning rates for each parameter that depends on the history of the gradients' variance.
+In particular we use different Adam parameters: the first one for snapshot updates and the second one for the sub-iterations updates.
 We stop if:
@@
 \section{Experiments}
-We design our experiments to investigate if:
+We have designed our experiments to investigate whether:
 \begin{enumerate}
-\item Algorithm reaches high value of the reward function faster then GPOMDP.
-\item Averagely SVRPG maintains the reward function more stable then GPOMDP.
+\item The algorithm reaches high value of the reward function faster than GPOMDP.
+\item Averagely SVRPG keep the reward function more stable than GPOMDP.
 \end{enumerate}
 
-In order to analyze these topics, we evaluate the behavior of rewards function mediated by 10 runs for each experiment.
+In order to pursue the above mentioned objectives, we evaluate the behavior of the reward function averaged on 10 runs for each experiment.
 
@@
 
-In all the experiments we face with GPOMDP standard algorithm. GPOMDP uses a number of trajectories per batch that is equal to the number of trajectories used in SVRPG mini-batch size.
+In all the experiments we compete with GPOMDP standard algorithm. GPOMDP uses a number of trajectories per batch that is equal to the number of trajectories used in SVRPG mini-batch size.
 The neural network and the learning rate are the same for GPOMDP and SVRPG.
-In SVRPG we implemented two different Adam parameters: the first one for Snapshot and the second one for sub-iterations.
-We use the stopping condition defined in Section \ref{sec:stopping}.
+In SVRPG we used two different Adam parameters: the first one for Snapshot and the second one for sub-iterations.
+We used the stopping condition defined in Section \ref{sec:stopping}.
 We fixed the maximum number of trajectories that both the algorithms can use.
 
-The neural networks is composed by one layer with 8 neurons for Cart-Pole environment and by two layers with 32x32 neurons for Swimmer and for Half Cheetah environments.
+The neural networks is composed of one layer with 8 neurons for Cart-Pole environment and by two layers with 32x32 neurons for Swimmer and for Half Cheetah environments.
 In each experiment, the length of episodes is 500 steps.
@@
 \section{Experimental Details}\label{app:proofs}
-Environments of our experiments are:
+Setting specifications for all the used environments:
 \begin{enumerate}
-	\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, the cart velocity $\dot{x}$ , and the pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body.
+	\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, cart velocity $\dot{x}$ and pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body.
 Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$.
-Episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time step T is major than 100.
+The episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
 	\item \emph{Mujoco Swimmer}: 12-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
 2-dimensional action space: the two momentums applied on actuated joints.
-Reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2] a^2$.
-Episodes terminates when the number of time step T is major than 500.
+The reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2] a^2$.
+The episodes terminates when the number of time steps T is greater than 500.
 	\item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles.
 6-dimensional action space: the 6 momentums applied on actuated joints.
-Reward function is defined as $r(s, a) := v_x - 0.05\norm[2] a^2$.
-Episodes terminates when the number of time step T is major than 500.
+The reward function is defined as $r(s, a) := v_x - 0.05\norm[2] a^2$.
+The episodes terminates when the number of time steps T is greater than 500.
 \end{enumerate}

### 2018-01-25 svrpg.tex
added minibatch to proofs and theorem

--- 2018-01-24 svrpg.tex
@@
 \begin{assumption}\label{ass:M2}
-	There is a constant $\Omega<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
+	There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
 \[
-	\mathbb{V}ar\left(\omega(\tau)\right) \leq \Omega
+	\mathbb{V}ar\left(\omega(\tau)\right) \leq W
+\]
+\end{assumption}
+
+\begin{assumption}\label{ass:REINFORCE}
+There is a constant $V<\infty$ such that, for each policy $\pol$:
+\[
+	\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq V
 \]
@@
 \begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
-Under Assumptions \ref{ass:bounded_score} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some constants $\psi>0, D>0$ and $F>0$ and for proper choice of step sizes $\alpha_t$, the following property:
+Under Assumptions \ref{ass:bounded_score} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,D, F$ and for proper choice of step sizes $\alpha_t$, the following property:
 \begin{align*}
@@
 		\leq
-		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} \\
-		&\qquad+
-		\frac{D\max_s\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}}{\psi N}
-		+\frac{F\Omega}{\psi}
+		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
+		\frac{Z}{N}
+		+\frac{\Omega}{b}
 \end{align*}
@@
 \nonumber 
-+dG^2R^2\left(M_2-1\right)
++dG^2R^2W
 \]
@@
 	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
-			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			+\frac{1}{b}\sum_{i=0}^{b-1} \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) 
+			-\frac{1}{b}\sum_{i=0}^{b-1}
 			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
@@
 	&= \mathbb{E}_{t\mid s}\left[\left\|\gradApp{\tilde{\vtheta}^s}{N}
-			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
+			+\frac{1}{b}\sum_{i=0}^{b-1}\left( 
+			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)
 			\right.\right.\nonumber\\&\qquad\left.\left.
@@
 	&\qquad+ 
-	\mathbb{E}_{t\mid s}\left[\left\|\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
+	\mathbb{E}_{t\mid s}\left[\left\|
+		\frac{1}{b}\sum_{i=0}^{b-1}\left(
+		\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)
 		\right.\right.\nonumber\\&\qquad\left.\left.
-		- \Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}\right\|^2\right] 
+		- \Ets{
+			\frac{1}{b}\sum_{i=0}^{b-1}\left(
+			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}\right\|^2\right] 
 	\nonumber\\
@@
 	&\qquad+ 
-		\mathbb{E}_{t\mid s}\left[\left\|\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
+		\mathbb{E}_{t\mid s}\left[\left\|
+		\frac{1}{b}\sum_{i=0}^{b-1}\left(
+		\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+		\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)
 		\right.\right.\nonumber\\&\qquad\left.\left.
-			-\Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}\right\|^2\right]  \nonumber\\
+		- \Ets{
+			\frac{1}{b}\sum_{i=0}^{b-1}\left(
+			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}\right\|^2\right] 
+		\nonumber\\%
+	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	&\qquad+\Ets{\norm[]{
+			\frac{1}{b}\sum_{i=0}^{b-1}\left(
+			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}^2} \\
 	%
@@
 	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
-	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
+	&\qquad+
+			\frac{1}{b^2}\sum_{i=0}^{b-1}
+			\Ets{\norm[]{
+			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
 	%
@@
 	\nonumber\\
-	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
+	&\qquad+
+			\frac{1}{b^2}\sum_{i=0}^{b-1}
+			\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
 			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
@@
 	\nonumber\\
-	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
+	&\qquad+
+			\frac{1}{b^2}\sum_{i=0}^{b-1}
+			\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
 			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
 	&\qquad
-	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
+			+\frac{1}{b^2}\sum_{i=0}^{b-1}
+			\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
 			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
@@
 	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	\nonumber\\
-	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
+	+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}\nonumber\\
+	&\qquad+
+		\frac{1}{b^2}\sum_{i=0}^{b-1}
+		\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
 			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
+% ...
+\end{align}
+\begin{align}
+% ...
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 	\nonumber\\
-	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2}{(1-\gamma)^2}\Ets{(\omega(\tau_i)-1)^2} \\
+	&\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+G^2\frac{R^2}{(1-\gamma)^2}\frac{1}{b^2}\sum_{i=0}^{b-1}\Ets{(\omega(\tau_i)-1)^2} \\
+	%
 	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 	\nonumber\\
-	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2}{(1-\gamma)^2}\left(\Ets{\omega^2(\tau_i)}-1\right) \nonumber\\
-	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	&\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+G^2\frac{R^2}{(1-\gamma)^2b^2}\sum_{i=0}^{b-1}\Varts{\omega(\tau_i)} \nonumber\\
+	%
+	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
 	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
 	\nonumber\\
-	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2\Omega}{(1-\gamma)^2},
+	&\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+G^2\frac{R^2W}{(1-\gamma)^2b},
 \end{align}
@@
 	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_t\Tr\Varts{\gradApp{\tilde{\vtheta}^s}{1}}}{N\Psi_t}
-	+\frac{f_t\Omega}{\Psi_t},
+	+\frac{f_tW}{b\Psi_t},
 \]
@@
 	&c_{m-1} = 0, \\
-	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\alpha_t^2\right) + \alpha_tL^2+\frac{\alpha_t^2L^3}{2}, \\
+	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\alpha_t^2\right) + \alpha_t\frac{L^2}{b}+\frac{\alpha_t^2L^3}{2}, \\
 	&\Psi_t = \alpha_t\left(1-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
@@
 	\right.\nonumber\\
-	&\left.\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2}{(1-\gamma)^2}\Omega\right)
+	&\left.\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+G^2\frac{R^2}{(1-\gamma)^2b}W\right)
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
@@
 	%
-	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta+\alpha_t^2\right)+\alpha_tL^2+\frac{\alpha_t^2L^3}{2}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta+\alpha_t^2\right)+\alpha_t\frac{L^2}{b}+\frac{\alpha_t^2L^3}{2}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad
@@
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
-	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2\Omega}{2(1-\gamma)^2} \nonumber\\
+	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2W}{2(1-\gamma)^2b} \nonumber\\
 	%
@@
 	-\frac{d_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	-f_t\Omega,\nonumber
+	-\frac{f_t}{b}W,\nonumber
 \end{align}
@@
 \begin{align*}
-& \psi \coloneqq \min_t\Psi_t, \\
-& D \coloneqq \max_t d_t \\
-& F \coloneqq \max_t f_t.
+& \psi \coloneqq \min_t\{\Psi_t\}, \\
+& Z \coloneqq \frac{\max_t\{d_t\}V}{\psi} \\
+& \Omega \coloneqq \frac{\max_t\{f_t\}W}{\psi}.
 \end{align*}

### 2018-01-25 svrpg.tex
finished proof (minus comments); next: add minibatch to algorithm exposition

@@
 \begin{assumption}\label{ass:bounded_score}
-For each trajectory $\tau$, value of $\vtheta$ and gradient component:
-\[
-	|\score{\vtheta}{\tau}| \leq G
-\]
+For each trajectory $\tau$, value of $\vtheta$ and dimensions $i,j$ there are positive constants $G,F<\infty$ such that:
+\begin{align*}
+	&\left|\nabla_{\theta_i}\log\pi_{\vtheta}(\tau)\right| \leq G \\
+	&\left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(\tau)\right| \leq F
+\end{align*}
 \end{assumption}
@@
 \[
-	L = \frac{2RG^2}{1-\gamma}
+	L = \frac{R(G^2+F)}{1-\gamma}
 \]
@@
 \begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
-Under Assumptions \ref{ass:bounded_score} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,D, F$ and for proper choice of step sizes $\alpha_t$, the following property:
+Under Assumptions \ref{ass:bounded_score} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size, the following property:
 \begin{align*}
@@
 		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
-		\frac{Z}{N}
-		+\frac{\Omega}{b}
+		\frac{\zeta}{N}
+		+\frac{\xi}{B}
 \end{align*}
@@
 \nonumber 
-+dG^2R^2W
++\frac{G^2R^2W}{(1-\gamma)^2B}
 \]
@@
 	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
-			+\frac{1}{b}\sum_{i=0}^{b-1} \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) 
-			-\frac{1}{b}\sum_{i=0}^{b-1}
+			+\frac{1}{B}\sum_{i=0}^{B-1} \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) 
+			-\frac{1}{B}\sum_{i=0}^{B-1}
 			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
@@
 	&= \mathbb{E}_{t\mid s}\left[\left\|\gradApp{\tilde{\vtheta}^s}{N}
-			+\frac{1}{b}\sum_{i=0}^{b-1}\left( 
+			+\frac{1}{B}\sum_{i=0}^{B-1}\left( 
 			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
@@
 	\mathbb{E}_{t\mid s}\left[\left\|
-		\frac{1}{b}\sum_{i=0}^{b-1}\left(
+		\frac{1}{B}\sum_{i=0}^{B-1}\left(
 		\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
@@
 		- \Ets{
-			\frac{1}{b}\sum_{i=0}^{b-1}\left(
+			\frac{1}{B}\sum_{i=0}^{B-1}\left(
 			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
@@
 		\mathbb{E}_{t\mid s}\left[\left\|
-		\frac{1}{b}\sum_{i=0}^{b-1}\left(
+		\frac{1}{B}\sum_{i=0}^{B-1}\left(
 		\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
@@
 	&\qquad+\Ets{\norm[]{
-			\frac{1}{b}\sum_{i=0}^{b-1}\left(
+			\frac{1}{B}\sum_{i=0}^{B-1}\left(
 			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
@@
 	&\qquad+
-			\frac{1}{b^2}\sum_{i=0}^{b-1}
+			\frac{1}{B^2}\sum_{i=0}^{B-1}
 			\Ets{\norm[]{
@@
 	&\qquad+
-			\frac{1}{b^2}\sum_{i=0}^{b-1}
+			\frac{1}{B^2}\sum_{i=0}^{B-1}
 			\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
@@
 	&\qquad
-			+\frac{1}{b^2}\sum_{i=0}^{b-1}
+			+\frac{1}{B^2}\sum_{i=0}^{B-1}
 			\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
@@
 	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}\nonumber\\
+	+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}\nonumber\\
 	&\qquad+
-		\frac{1}{b^2}\sum_{i=0}^{b-1}
+		\frac{1}{B^2}\sum_{i=0}^{B-1}
 		\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
@@
 	\nonumber\\
-	&\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2}{(1-\gamma)^2}\frac{1}{b^2}\sum_{i=0}^{b-1}\Ets{(\omega(\tau_i)-1)^2} \\
+	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+G^2\frac{R^2}{(1-\gamma)^2}\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \\
 	%
@@
 	\nonumber\\
-	&\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2}{(1-\gamma)^2b^2}\sum_{i=0}^{b-1}\Varts{\omega(\tau_i)} \nonumber\\
+	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+G^2\frac{R^2}{(1-\gamma)^2b^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
 	%
@@
 	\nonumber\\
-	&\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2W}{(1-\gamma)^2b},
+	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+G^2\frac{R^2W}{(1-\gamma)^2B},
 \end{align}
@@
 	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
-	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_t\Tr\Varts{\gradApp{\tilde{\vtheta}^s}{1}}}{N\Psi_t}
+	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_tV}{N\Psi_t}
 	+\frac{f_tW}{b\Psi_t},
@@
 	-\frac{d_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	-\frac{f_t}{b}W,\nonumber
+	-\frac{f_t}{b}W,\nonumber\\
+	&\geq R_t^{s+1}
+	+\Psi_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
+	-\frac{d_t}{N}V
+	-\frac{f_t}{b}W,
 \end{align}
-where (17) is from (13), (18) is from (16), and (19) is from Lemma \ref{lemma:aux0}.
+where (17) is from (13), (18) is from (16), and (19) is from Lemma \ref{lemma:aux0} and (20) is from Assumption \ref{ass:REINFORCE}.
 To complete the proof, besides rearranging terms, we have to ensure that $\Psi_t>0$ for each $t$. This gives the constraints on $\alpha_t$ and $\beta_t$.
@@
 & \psi \coloneqq \min_t\{\Psi_t\}, \\
-& Z \coloneqq \frac{\max_t\{d_t\}V}{\psi} \\
-& \Omega \coloneqq \frac{\max_t\{f_t\}W}{\psi}.
+& \zeta \coloneqq \frac{\max_t\{d_t\}V}{\psi} \\
+& \xi \coloneqq \frac{\max_t\{f_t\}W}{\psi}.
 \end{align*}
@@
 
-...
+Starting from Lemma \ref{lemma:aux1}, summing over iterations of an epoch $s$ we get:
+\begin{align}
+\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
+ \frac{\sum_{t=0}^{m-1}\left(R_{t+1}^{s+1} - R_t^{s+1}\right)}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B} \\
+ %
+ &= \frac{\sum_{t=0}^{m-1}\left(\Ets[t+1]{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} - \Ets[t]{J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2}\right)}{\psi} \nonumber\\
+ &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \\
+ %
+ &= \frac{\Ets[m]{\sum_{t=0}^{m-1}\left(J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2\right)}}{\psi} \nonumber\\
+ &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \\
+ %
+ &= \frac{\Ets[m]{J(\vtheta_{m}^{s+1}) - c_{t+1}\norm[]{\vtheta_{m}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{0}^{s+1}) - c_{t}\norm[]{\vtheta_{0}^{s+1}-\tilde{\vtheta}^s}^2}}{\psi} \nonumber\\
+ &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \\
+ %
+ &= \frac{\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B}
+ \end{align}
+Next, summing over epochs:
+\begin{align}
+\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
+\frac{\sum_{s=0}^{S-1}\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \\
+%
+&\leq
+\frac{\EVV[]{J(\tilde{\vtheta}^{S}) - J(\tilde{\vtheta}^{0})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}
+ \\
+ %
+ &\leq
+ \frac{\EVV[]{J(\vtheta_{out})} - J(\tilde{\vtheta}^{0})}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \\
+ %
+ &\leq
+ \frac{J(\vtheta^*) - J(\vtheta^0)}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}
+\end{align}
+Finally, we consider the expectation \wrt all sources of randomness, including the uniform sampling of the output parameter:
+\begin{align}
+\EVV[]{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
+&=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} \\
+&\leq
+\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B}
+\end{align}
 \end{proof}

### 2018-01-29 svrpg.tex
Inserted pratical variants

@@
 
-\subsection{Policy Gradient}
+\subsection{Policy Gradient}\label{subsec:PolicyGradient}
 A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
@@
 \begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\gamma^h r_h^n - b\right).
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\mid s_k^n}\right)\left(\gamma^h r_h^n - b\right).
 \end{align*}
@@
 \section{Stopping Condition}\label{sec:stopping}
-In practice ,the performance of the SVRPG algorithm is sensitive to the number of sub-iterations.
+As shown, the performance of the SVRPG algorithm is sensitive to the number of sub-iterations.
 As the number of sub-iterations increases, the variance of the importance weights and the gradient increases too.
@@
 
-We introduce two stopping condition: maximum number of sub-iterations and learning rate of Adam sub-iterations.
+We introduce two stopping condition: the maximum number of sub-iterations and the condition on learning rate of Adam sub-iterations.
 Adam update rule help to stabilize the gradient estimate. In fact, this method computes adaptive learning rates for each parameter that depends on the history of the gradients' variance.
-In particular we use different Adam parameters: the first one for snapshot updates and the second one for the sub-iterations updates.
-We stop if:
+In particular we use two different Adam parameters: the first one for snapshot updates and the second one for the sub-iterations updates.
+We stop with sub-iterations if:
 
-\[\frac{\alpha_{snap}}{N}>\frac{\alpha_{sub-iter}}{M}\]
+\[\frac{\alpha_{snap}}{N}>\frac{\alpha_{sub-iter}}{B}\]
 
-where M is the sub-iterations batch size.
+After that we sample N trajectories for the next Snapshot point.
 
 \section{Practical Variants}
+
+As reported in Section \ref{subsec:PolicyGradient} in practical variants we use G(PO)MDP gradient estimator~\cite{baxter2001infinite}, that brings the same guarantees of Reinforce Estimator.
+One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}:  this allows a reduction of the gradient estimator variance.
+The off-policy gradient estimator becomes:
+\begin{align*}
+\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\mid s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
+\end{align*} 
+where:
+\begin{align*}
+\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\mid s_k^n})
+ }{\pi_{\vtheta^{'}}({a_k^n\mid s_k^n})}
+\end{align*}
+\begin{align*}
+\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\mid s_k^n})
+}{\pi_{\vtheta^{'}}({a_k^n\mid s_k^n})}
+\end{align*}
 
@@
 We used the stopping condition defined in Section \ref{sec:stopping}.
-We fixed the maximum number of trajectories that both the algorithms can use.
+We fixed the maximum number of trajectories that both the algorithms can use as shown in appendices.
 

### 2018-01-29 svrpg.tex
added comments to main theorem proof

@@
 \newcommand{\todopirout}[1]{\todo[color=blued]{\scriptsize #1}}
+%%todo by Damiano
+\newcommand{\tododam}[1]{\todo[color=orange, inline]{\small #1}}
+\newcommand{\tododamout}[1]{\todo[color=orange]{\scriptsize #1}}
+%%todo by Giuseppe
+\newcommand{\todobep}[1]{\todo[color=purple, inline]{\small #1}}
+\newcommand{\todobepout}[1]{\todo[color=purple]{\scriptsize #1}}
 %%%%%%
@@
 \newcommand{\gradIdeal}[1]{\overline{\blacktriangledown} J(#1)}
+\newcommand{\VARRF}{V}
+\newcommand{\GRADLOG}{G}
+\newcommand{\VARIS}{W}
+\newcommand{\HESSLOG}{F}
 %%%%%%
@@
 \section{Algorithm}
+\todomat{Add minibatch}
 In on-line policy gradient problems, stochastic gradient ascent (SGA) is not really a choice, but is dictated by the necessity of interacting with an unknown environment.
@@
 \end{restatable}
-
-[May also add Lemma about convergence of the variance to zero, since it provides an intuitive justification of some design choices]
+\todomat{Add lemma on zero variance at convergence}
 The resulting SVRPG policy optimization method is detailed in Algorithm \ref{alg:svrpg}.
@@
 \begin{align*}
-	&\left|\nabla_{\theta_i}\log\pi_{\vtheta}(\tau)\right| \leq G \\
-	&\left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(\tau)\right| \leq F
+	&\left|\nabla_{\theta_i}\log\pi_{\vtheta}(\tau)\right| \leq \GRADLOG \\
+	&\left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(\tau)\right| \leq \HESSLOG
 \end{align*}
@@
 \[
-	\mathbb{V}ar\left(\omega(\tau)\right) \leq W
+	\mathbb{V}ar\left(\omega(\tau)\right) \leq \VARIS
 \]
@@
 \[
-	\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq V
+	\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq \VARRF
 \]
@@
 \begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
-Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth with Lipschitz constant:
-\[
-	L = \frac{R(G^2+F)}{1-\gamma}
-\]
+Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth.
 \end{restatable}
@@
 	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	\\
+	\nonumber\\
 	&\qquad+ 
@@
 			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}\right\|^2\right] 
-		\nonumber\\%
+		\label{eq:1}\\%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}^2} \\
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}^2} \label{eq:2}\\
 	%
@@
 		\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
+			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \label{eq:3}
 % ...
@@
 	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2}{(1-\gamma)^2}\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \\
+	+\GRADLOG^2\frac{R^2}{(1-\gamma)^2}\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \label{eq:4}\\
 	%
@@
 	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2}{(1-\gamma)^2b^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
+	+\GRADLOG^2\frac{R^2}{(1-\gamma)^2b^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
 	%
@@
 	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2W}{(1-\gamma)^2B},
+	+\GRADLOG^2\frac{R^2\VARRF}{(1-\gamma)^2B}, \label{eq:5}
 \end{align}
-where (1) is from the definition of $\gradApp{\vtheta}{N}$, (2) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (3) is from L-smoothness, 
-(4) is from Assumption \ref{ass:bounded_score}, and (5) is from Assumption \ref{ass:M2}.
+where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (\ref{eq:3}) is from L-smoothness,
+(\ref{eq:4}) is from Assumption \ref{ass:bounded_score}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
 \end{proof}
@@
 	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
-	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
+	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \label{eq:6}\\
 	&=
 	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
-	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
+	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \label{eq:7}\\
 	&=
@@
 	&\qquad+
-	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}
+	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})} \label{eq:8}
 \end{align}
@@
 	&\leq	
-	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\\
+	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\label{eq:9}\\
 	%
 	&=
-	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \\
+	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \label{eq:10}\\
 	%
@@
 \end{align}
-where (6) is from Definition \ref{def:ideal}; (7) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (8) is from $\Ets{e_s}=0$; (9) is from Young's inequality; and (10) is from the definition of $\gradApp{\vtheta}{N}$.
+where (\ref{eq:6}) is from Definition \ref{def:ideal}; (\ref{eq:7}) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (\ref{eq:8}) is from $\Ets{e_s}=0$; (\ref{eq:9}) is from Young's inequality; and (\ref{eq:10}) is from the definition of $\gradApp{\vtheta}{N}$.
 \end{proof}
@@
 	\Ets{J(\vtheta_{t+1}^{s+1})} 
-	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
-	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
+	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \label{eq:11}\\
+	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \label{eq:12}\\
 	&\geq
@@
 	&\qquad-
-	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2},
+	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}, \label{eq:13}
 \end{align}
-where (11) is from the L-smoothness of $J(\vtheta)$, (12) is from the SVRPG update, and (13) is from Lemma \ref{lemma:aux0}.
+where (\ref{eq:11}) is from the L-smoothness of $J(\vtheta)$, (\ref{eq:12}) is from the SVRPG update, and (\ref{eq:13}) is from Lemma \ref{lemma:aux0}.
 
@@
 &=\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_{t}^{s+1}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\dotprod{\vtheta_{t+1}^{s+1}-\vtheta_{t}^{s+1}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber \\
-&= \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \\
+&= \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \label{eq:14}\\
 &\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber\\ 
 &\qquad+
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \\
+\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
 %
@@
 &\qquad
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2},
+\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
 \end{align}
-where (14) is from the SVRPG update, (15) is from Lemma \ref{lemma:aux0}, and (16) is from Young's inequality in the 'Peter-Paul' variant.
+where (\ref{eq:14}) is from the SVRPG update, (\ref{eq:15}) is from Lemma \ref{lemma:aux0}, and (\ref{eq:16}) is from Young's inequality in the 'Peter-Paul' variant.
 Finally:
@@
 	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
-	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \\
+	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:17}\\
 	%
@@
 	&\qquad
-	-c_{t+1}\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \\
+	-c_{t+1}\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:18}\\
 	%
@@
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \label{eq:19}\\
 	%
@@
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
-	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2W}{2(1-\gamma)^2b} \nonumber\\
+	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2\VARIS}{2(1-\gamma)^2b} \nonumber\\
 	%
@@
 	-\frac{d_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	-\frac{f_t}{b}W,\nonumber\\
+	-\frac{f_t}{b}\VARIS,\nonumber\\
 	&\geq R_t^{s+1}
 	+\Psi_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
-	-\frac{d_t}{N}V
-	-\frac{f_t}{b}W,
+	-\frac{d_t}{N}\VARRF
+	-\frac{f_t}{b}\VARIS, \label{eq:20}
 \end{align}
-where (17) is from (13), (18) is from (16), and (19) is from Lemma \ref{lemma:aux0} and (20) is from Assumption \ref{ass:REINFORCE}.
+where (\ref{eq:17}) is from (\ref{eq:13}), (\ref{eq:18}) is from (\ref{eq:16}), (\ref{eq:19}) is from Lemma \ref{lemma:aux0}, and (\ref{eq:20}) is from Assumption \ref{ass:REINFORCE}.
 To complete the proof, besides rearranging terms, we have to ensure that $\Psi_t>0$ for each $t$. This gives the constraints on $\alpha_t$ and $\beta_t$.
@@
 \sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
- \frac{\sum_{t=0}^{m-1}\left(R_{t+1}^{s+1} - R_t^{s+1}\right)}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B} \\
+ \frac{\sum_{t=0}^{m-1}\left(R_{t+1}^{s+1} - R_t^{s+1}\right)}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
  %
  &= \frac{\sum_{t=0}^{m-1}\left(\Ets[t+1]{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} - \Ets[t]{J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2}\right)}{\psi} \nonumber\\
- &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \\
+ &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
  %
  &= \frac{\Ets[m]{\sum_{t=0}^{m-1}\left(J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2\right)}}{\psi} \nonumber\\
- &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \\
+ &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
  %
- &= \frac{\Ets[m]{J(\vtheta_{m}^{s+1}) - c_{t+1}\norm[]{\vtheta_{m}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{0}^{s+1}) - c_{t}\norm[]{\vtheta_{0}^{s+1}-\tilde{\vtheta}^s}^2}}{\psi} \nonumber\\
- &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \\
+ &= \frac{\Ets[m]{J(\vtheta_{m}^{s+1}) - c_{m}\norm[]{\vtheta_{m}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{0}^{s+1}) - c_{0}\norm[]{\vtheta_{0}^{s+1}-\tilde{\vtheta}^s}^2}}{\psi} \nonumber\\
+ &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \label{eq:21}\\
  %
- &= \frac{\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B}
+ &= \frac{\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B}, \label{eq:22}
  \end{align}
+where (\ref{eq:21}) is obtained by telescoping the sum and (\ref{eq:22}) from the fact that $c_m=0$ and $\vtheta_0^{s+1}=\tilde{\vtheta}^s$.
 Next, summing over epochs:
@@
 \sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
-\frac{\sum_{s=0}^{S-1}\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \\
+\frac{\sum_{s=0}^{S-1}\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \nonumber\\
 %
@@
 \frac{\EVV[]{J(\tilde{\vtheta}^{S}) - J(\tilde{\vtheta}^{0})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}
- \\
+ \label{eq:23}\\
 %
 &\leq
- \frac{\EVV[]{J(\vtheta_{out})} - J(\tilde{\vtheta}^{0})}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \\
- %
- &\leq
- \frac{J(\vtheta^*) - J(\vtheta^0)}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}
+\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}, \label{eq:24}
 \end{align}
+where the expectation in (\ref{eq:23}) is \wrt all the trajectories sampled in a run of Algorithm \ref{alg:svrpg} and (\ref{eq:24}) is from the definition of $\vtheta^*$.
+
 Finally, we consider the expectation \wrt all sources of randomness, including the uniform sampling of the output parameter:
@@
 \EVV[]{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
-&=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} \\
+&=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} \nonumber\\
 &\leq
-\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B}
+\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B} \nonumber.
 \end{align}

### 2018-01-30 svrpg.tex
added minibatch to algorithm description

@@
 where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from a snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\mid\tilde{\vtheta})}{\pi(\tau\mid\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$.
+Our update is still fundamentally on-policy, since the weighting concerns only the correction term.
+However, this partial "off-policyness" is enough to introduce variance.
+To mitigate it, we use mini-batches of trajectories of size B to compute the second and the third term, as follows:
+\begin{align*}
+\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta})\Reward(\tau_j) \\
+&+ \frac{1}{B}\left[\sum_{i=0}^{B-1}
+\nabla\log\pi(\tau_i \mid \vtheta)\Reward(\tau_i) 
+- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta})\Reward(\tau_i)\right].
+\end{align*}
+The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case \cite{konevcny2016mini} \cite{harikandeh2015stopwasting}.
 We immediately have the following (proof in Appendix \ref{app:proofs}):
-
 \begin{restatable}{lemma}{unbias}\label{lemma:unbias}
@@
 The resulting SVRPG policy optimization method is detailed in Algorithm \ref{alg:svrpg}.
-To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient and the step size $\alpha$.
+To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient, the mini-batch size $B$, and the step size $\alpha$.
 Adaptive selection of these meta-parameters is also possible.
-However, the next section will focus on providing converge guarantees for the plain SVRPG algorithm.
+However, the next section will focus on providing general converge guarantees for Algorithm \ref{alg:svrpg}.
 
@@
 	\begin{algorithmic}
-		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m-1}^0$, batch size $N$
+		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m-1}^0$, batch size $N$, mini-batch size B
 		\FOR{$s=0$ {\bfseries to} $S-1$}
@@
 		\FOR{$t=0$ {\bfseries to} $m-1$}
-		\STATE Sample trajectory $\tau_i$ from 				$\pi(\cdot\mid\vtheta_t^{s+1})$
+		\STATE Sample $B$ trajectories $\tau_i$ from 				$\pi(\cdot\mid\vtheta_t^{s+1})$
 		\STATE 
@@
 		\blacktriangledown J(\vtheta_t^{s+1}) = 
-		&\gradApp{\tilde{\vtheta}^s}{N}
-		+ \score{}{\tau_i\mid\vtheta_t^{s+1}}\Reward(\tau_i) \\
-		& - \omega(\tau_i)\score{}{\tau_i \mid \tilde{\vtheta}^{s}}\Reward(\tau_i)
+		&\gradApp{\tilde{\vtheta}^s}{N} \\
+		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
+		\score{}{\tau_i\mid\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
+		&\left. - \omega(\tau_i)\score{}{\tau_i \mid \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
 		\end{align*}

### 2018-01-30 svrpg.tex
added lemma on zero variance at convergence

@@
 \section{Algorithm}
-\todomat{Add minibatch}
 In on-line policy gradient problems, stochastic gradient ascent (SGA) is not really a choice, but is dictated by the necessity of interacting with an unknown environment.
@@
 The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case \cite{konevcny2016mini} \cite{harikandeh2015stopwasting}.
-We immediately have the following (proof in Appendix \ref{app:proofs}):
+It is easy to show that the SVRPG gradient estimator has the following, desirable properties (detailed proofs in Appendix \ref{app:proofs}):
 \begin{restatable}{lemma}{unbias}\label{lemma:unbias}
-The SVRPG gradient estimator is unbiased:
+The SVRPG estimator is unbiased:
 \[
@@
 \end{restatable}
-\todomat{Add lemma on zero variance at convergence}
+\begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
+At convergence, the variance of the SVRPG estimator is equal to the variance of the REINFORCE estimator, regardless of the mini-batch size $B$:
+\[
+	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
+	\mathbb{V}ar\left[\gradApp{\vtheta^*}{N}\right].
+\]
+\end{restatable}
+In particular, the latter suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can actually achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance, at least asymptotically.
 The resulting SVRPG policy optimization method is detailed in Algorithm \ref{alg:svrpg}.
@@
 
-\subsection*{Lemmas}
-Before addressing the main convergence theorem, we prove some useful lemmas.
+\subsection*{Basic Lemmas}
+We prove two basic properties of the SVRPG update.
 
@@
 \end{proof}
+
+\zerovar*
+\begin{proof}
+...
+\end{proof}
+
+\subsection*{Ancillary Lemmas}
+Before addressing the main convergence theorem, we prove some useful lemmas.
 
@@
 \subsection*{Main theorem}
+We finally provide the proof of the convergence theorem:
 

### 2018-01-30 svrpg.tex
added proofs of basic lemmas

@@
 \usepackage{pifont}
+\usepackage[makeroom]{cancel}
 
@@
 \DeclareMathOperator*{\Cov}{\mathbb{C}ov}
+\DeclareMathOperator*{\Var}{\mathbb{V}ar}
 \newcommand{\EVV}[2][\ppvect \in \ppspace]{\EV_{#1}\left[{#2}\right]}
@@
 To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient, the mini-batch size $B$, and the step size $\alpha$.
-Adaptive selection of these meta-parameters is also possible.
+Adaptive selection of these meta-parameters is also possible, and is explored in Section \ref{sec:stopping}.
 However, the next section will focus on providing general converge guarantees for Algorithm \ref{alg:svrpg}.
@@
 \begin{proof}
-...
+\begin{align*}
+\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\tilde{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} + \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i)\score{\tilde{\vtheta}}{\tau_i}R(\tau_i)} \\
+&= \gradJ{\tilde{\vtheta}} + \gradJ{\vtheta} - \gradJ{\tilde{\vtheta}} = \gradJ{\vtheta}.
+\end{align*}
+Note that the importance weight is necessary to guarantee unbiasedness, since the $\tau_i$ are sampled from $\pi_{\vtheta}$.
 \end{proof}
@@
 \begin{proof}
-...
+As $\vtheta\to\vtheta^*$, also $\tilde{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
+\begin{align*}
+\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[\score{\vtheta^*}{\tau}R(\tau) - \cancel{\omega(\tau)}\score{\vtheta^*}{\tau}R(\tau)\right] \\
+&= \Var\left[\gradApp{\vtheta^*}{N}\right].
+\end{align*}
+Note that it is important that the trajectories used in the second and the third term are the same for the variance to vanish.
 \end{proof}
@@
 Before addressing the main convergence theorem, we prove some useful lemmas.
-
 %Lemma 2

### 2018-01-30 svrpg.tex
Updated

--- 2018-01-30 svrpg.tex
@@
 \newcommand{\todopirout}[1]{\todo[color=blued]{\scriptsize #1}}
-%%todo by Damiano
-\newcommand{\tododam}[1]{\todo[color=orange, inline]{\small #1}}
-\newcommand{\tododamout}[1]{\todo[color=orange]{\scriptsize #1}}
-%%todo by Giuseppe
-\newcommand{\todobep}[1]{\todo[color=purple, inline]{\small #1}}
-\newcommand{\todobepout}[1]{\todo[color=purple]{\scriptsize #1}}
 %%%%%%
@@
 \newcommand{\gradIdeal}[1]{\overline{\blacktriangledown} J(#1)}
-\newcommand{\VARRF}{V}
-\newcommand{\GRADLOG}{G}
-\newcommand{\VARIS}{W}
-\newcommand{\HESSLOG}{F}
 %%%%%%
@@
 \section{Preliminaries}
-In this section we provide essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
-In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
+In this section we provide the essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
+%In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
 
-\subsection{Policy Gradient}\label{subsec:PolicyGradient}
-A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
+\subsection{Policy Gradient in Reinforcement Learning}
+A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $M = \{ \Sspace,\Aspace,\Transition,\Reward,\gamma,\rho \}$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$;
+% and $R$ is the maximum absolute-value reward;
+$\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
 The agent's behavior is modeled as a policy $\pi$, where $\pi(\cdot|s)$ is the density distribution over $\Aspace$ in state $s$.
@@
 % In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
-We consider episodic MDPs with effective horizon $H$, \ie optimal policy is able to reach the target state (\ie absorbing state) in at most $H$ steps (the true length depend on the initial state).
+We consider episodic MDPs with effective horizon $H$, \ie the optimal policy is able to reach the target state (\ie absorbing state) in at most $H$ steps (the true length depend on the initial state).
 This has not to be confused with a finite horizon problem where the optimal policy is non-stationary.
@@
 With abuse of notation, we denote with $\pi(\tau)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories, and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
-\begin{align*}
-\Reward(\tau) = \sum_{t=0}^{H-1}\gamma^t\Reward(s_t,a_t).
-\end{align*}
+%
+$\Reward(\tau) = \sum_{t=0}^{H-1}\gamma^t\Reward(s_t,a_t).$
+%
 Policies can be ranked based on their expected total reward:
@@
 \end{align*}
-Solving an MDP means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
+Solving an MDP $M$ means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
 
-Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt the $m$-dimensional parameter vector $\vtheta$.
+Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$.
 We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
-The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
-\todopir{stochastic or mini-batch gradient}
-\begin{align*}
-\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
-\end{align*}
-We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment. 
-In this setting, $(N>0)$-batch gradient ascent is typically employed: after sampling a batch of $N$ trajectories (\ie $\Dataset_{N}$),
-policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient. 
-\todopir{step size is scalar or vectorial (just for notation)? Check consistency of $N$ with finite-sum function, we can use even $b$ or $B$ for the batch size.}
-The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
-\begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
-\end{align*}
-where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
-\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,anonymous2018variance}.}
-The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
-\begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\mid s_k^n}\right)\left(\gamma^h r_h^n - b\right).
-\end{align*}
-\todopir{Cite Jan's paper somewhere}
+The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy, Peters2008reinf}:
+\begin{align} \label{E:policygradient}
+\gradJ{\vtheta} =
+\EVV[\tau \sim \pol(\tau)]{\score{\vtheta}{\tau}\Reward(\tau)}
+%\int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
+\end{align}
+Notice that the distribution defining the gradient is induced by the current policy.
+This peculiarity induces a nonstationarity in the sampling process.
+Since the underlying distribution is changing over time, it is necessary to resample at each update or use weighting techniques like importance sampling.
+
+We consider the \emph{on-line learning problem}, where trajectories are sampled by interacting with the environment at each policy change. 
+In this setting, stochastic gradient ascent is typically employed.
+At each iteration $k >0$, after sampling a batch of $N$ ($0<N<+\infty$) trajectories with policy %$\pi_{{\vtheta_k}}$,
+the policy is updated as $\vtheta  = \vtheta + \alpha\gradApp{{\vtheta_k}}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $N$ samples.
+Several estimators have been provided in literature as empirical average over trajectories:
+\begin{align} \label{E:policygradient.estimate}
+        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} \nabla f(\tau^i).
+\end{align}
+The most widespread estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
+We refer the reader to App.~\ref{A:gradient_estimators} for a formal definition.
+Both are unbiased estimate of~\eqref{E:policygradient} but G(PO)MDP is subject to less variance than REINFORCE.
 \todopir{At this point we need to connect with the supervised learning framework (we can just define everything inside the first sum as $\nabla g(\tau)$).}
@@
 where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from a snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\mid\tilde{\vtheta})}{\pi(\tau\mid\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$.
-Our update is still fundamentally on-policy, since the weighting concerns only the correction term.
-However, this partial "off-policyness" is enough to introduce variance.
-To mitigate it, we use mini-batches of trajectories of size B to compute the second and the third term, as follows:
-\begin{align*}
-\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta})\Reward(\tau_j) \\
-&+ \frac{1}{B}\left[\sum_{i=0}^{B-1}
-\nabla\log\pi(\tau_i \mid \vtheta)\Reward(\tau_i) 
-- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta})\Reward(\tau_i)\right].
-\end{align*}
-The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case \cite{konevcny2016mini} \cite{harikandeh2015stopwasting}.
 We immediately have the following (proof in Appendix \ref{app:proofs}):
+
 \begin{restatable}{lemma}{unbias}\label{lemma:unbias}
@@
 \end{restatable}
-\todomat{Add lemma on zero variance at convergence}
+
+[May also add Lemma about convergence of the variance to zero, since it provides an intuitive justification of some design choices]
 The resulting SVRPG policy optimization method is detailed in Algorithm \ref{alg:svrpg}.
-To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient, the mini-batch size $B$, and the step size $\alpha$.
+To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient and the step size $\alpha$.
 Adaptive selection of these meta-parameters is also possible.
-However, the next section will focus on providing general converge guarantees for Algorithm \ref{alg:svrpg}.
+However, the next section will focus on providing converge guarantees for the plain SVRPG algorithm.
 
@@
 	\begin{algorithmic}
-		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m-1}^0$, batch size $N$, mini-batch size B
+		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m-1}^0$, batch size $N$
 		\FOR{$s=0$ {\bfseries to} $S-1$}
@@
 		\FOR{$t=0$ {\bfseries to} $m-1$}
-		\STATE Sample $B$ trajectories $\tau_i$ from 				$\pi(\cdot\mid\vtheta_t^{s+1})$
+		\STATE Sample trajectory $\tau_i$ from 				$\pi(\cdot\mid\vtheta_t^{s+1})$
 		\STATE 
@@
 		\blacktriangledown J(\vtheta_t^{s+1}) = 
-		&\gradApp{\tilde{\vtheta}^s}{N} \\
-		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
-		\score{}{\tau_i\mid\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
-		&\left. - \omega(\tau_i)\score{}{\tau_i \mid \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
+		&\gradApp{\tilde{\vtheta}^s}{N}
+		+ \score{}{\tau_i\mid\vtheta_t^{s+1}}\Reward(\tau_i) \\
+		& - \omega(\tau_i)\score{}{\tau_i \mid \tilde{\vtheta}^{s}}\Reward(\tau_i)
 		\end{align*}
@@
 \begin{align*}
-	&\left|\nabla_{\theta_i}\log\pi_{\vtheta}(\tau)\right| \leq \GRADLOG \\
-	&\left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(\tau)\right| \leq \HESSLOG
+	&\left|\nabla_{\theta_i}\log\pi_{\vtheta}(\tau)\right| \leq G \\
+	&\left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(\tau)\right| \leq F
 \end{align*}
@@
 \[
-	\mathbb{V}ar\left(\omega(\tau)\right) \leq \VARIS
+	\mathbb{V}ar\left(\omega(\tau)\right) \leq W
 \]
@@
 \[
-	\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq \VARRF
+	\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq V
 \]
@@
 \begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
-Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth.
+Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth with Lipschitz constant:
+\[
+	L = \frac{R(G^2+F)}{1-\gamma}
+\]
 \end{restatable}
@@
 \section{Stopping Condition}\label{sec:stopping}
-As shown, the performance of the SVRPG algorithm is sensitive to the number of sub-iterations.
+In practice ,the performance of the SVRPG algorithm is sensitive to the number of sub-iterations.
 As the number of sub-iterations increases, the variance of the importance weights and the gradient increases too.
@@
 
-We introduce two stopping condition: the maximum number of sub-iterations and the condition on learning rate of Adam sub-iterations.
+We introduce two stopping condition: maximum number of sub-iterations and learning rate of Adam sub-iterations.
 Adam update rule help to stabilize the gradient estimate. In fact, this method computes adaptive learning rates for each parameter that depends on the history of the gradients' variance.
-In particular we use two different Adam parameters: the first one for snapshot updates and the second one for the sub-iterations updates.
-We stop with sub-iterations if:
+In particular we use different Adam parameters: the first one for snapshot updates and the second one for the sub-iterations updates.
+We stop if:
 
-\[\frac{\alpha_{snap}}{N}>\frac{\alpha_{sub-iter}}{B}\]
+\[\frac{\alpha_{snap}}{N}>\frac{\alpha_{sub-iter}}{M}\]
 
-After that we sample N trajectories for the next Snapshot point.
+where M is the sub-iterations batch size.
 
 \section{Practical Variants}
-
-As reported in Section \ref{subsec:PolicyGradient} in practical variants we use G(PO)MDP gradient estimator~\cite{baxter2001infinite}, that brings the same guarantees of Reinforce Estimator.
-One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}:  this allows a reduction of the gradient estimator variance.
-The off-policy gradient estimator becomes:
-\begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\mid s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
-\end{align*} 
-where:
-\begin{align*}
-\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\mid s_k^n})
- }{\pi_{\vtheta^{'}}({a_k^n\mid s_k^n})}
-\end{align*}
-\begin{align*}
-\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\mid s_k^n})
-}{\pi_{\vtheta^{'}}({a_k^n\mid s_k^n})}
-\end{align*}
 
@@
 We used the stopping condition defined in Section \ref{sec:stopping}.
-We fixed the maximum number of trajectories that both the algorithms can use as shown in appendices.
+We fixed the maximum number of trajectories that both the algorithms can use.
 
@@
 \appendix
+
+\section{Policy Gradient Estimators} \label{A:gradient_estimators}
+
+The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
+\begin{align*}
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
+\end{align*}
+where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
+\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,anonymous2018variance}.}
+The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
+\begin{align*}
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\gamma^h r_h^n - b\right).
+\end{align*}
 
@@
 	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	\nonumber\\
+	\\
 	&\qquad+ 
@@
 			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}\right\|^2\right] 
-		\label{eq:1}\\%
+		\nonumber\\%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
@@
 			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}^2} \label{eq:2}\\
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}^2} \\
 	%
@@
 		\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \label{eq:3}
+			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
 % ...
@@
 	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\GRADLOG^2\frac{R^2}{(1-\gamma)^2}\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \label{eq:4}\\
+	+G^2\frac{R^2}{(1-\gamma)^2}\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \\
 	%
@@
 	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\GRADLOG^2\frac{R^2}{(1-\gamma)^2b^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
+	+G^2\frac{R^2}{(1-\gamma)^2b^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
 	%
@@
 	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\GRADLOG^2\frac{R^2\VARRF}{(1-\gamma)^2B}, \label{eq:5}
+	+G^2\frac{R^2W}{(1-\gamma)^2B},
 \end{align}
-where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (\ref{eq:3}) is from L-smoothness,
-(\ref{eq:4}) is from Assumption \ref{ass:bounded_score}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
+where (1) is from the definition of $\gradApp{\vtheta}{N}$, (2) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (3) is from L-smoothness, 
+(4) is from Assumption \ref{ass:bounded_score}, and (5) is from Assumption \ref{ass:M2}.
 \end{proof}
@@
 	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
-	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \label{eq:6}\\
+	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
 	&=
 	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
-	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \label{eq:7}\\
+	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
 	&=
@@
 	&\qquad+
-	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})} \label{eq:8}
+	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}
 \end{align}
@@
 	&\leq	
-	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\label{eq:9}\\
+	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\\
 	%
 	&=
-	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \label{eq:10}\\
+	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \\
 	%
@@
 \end{align}
-where (\ref{eq:6}) is from Definition \ref{def:ideal}; (\ref{eq:7}) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (\ref{eq:8}) is from $\Ets{e_s}=0$; (\ref{eq:9}) is from Young's inequality; and (\ref{eq:10}) is from the definition of $\gradApp{\vtheta}{N}$.
+where (6) is from Definition \ref{def:ideal}; (7) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (8) is from $\Ets{e_s}=0$; (9) is from Young's inequality; and (10) is from the definition of $\gradApp{\vtheta}{N}$.
 \end{proof}
@@
 	\Ets{J(\vtheta_{t+1}^{s+1})} 
-	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \label{eq:11}\\
-	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \label{eq:12}\\
+	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
+	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
 	&\geq
@@
 	&\qquad-
-	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}, \label{eq:13}
+	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2},
 \end{align}
-where (\ref{eq:11}) is from the L-smoothness of $J(\vtheta)$, (\ref{eq:12}) is from the SVRPG update, and (\ref{eq:13}) is from Lemma \ref{lemma:aux0}.
+where (11) is from the L-smoothness of $J(\vtheta)$, (12) is from the SVRPG update, and (13) is from Lemma \ref{lemma:aux0}.
 
@@
 &=\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_{t}^{s+1}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\dotprod{\vtheta_{t+1}^{s+1}-\vtheta_{t}^{s+1}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber \\
-&= \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \label{eq:14}\\
+&= \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \\
 &\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber\\ 
 &\qquad+
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
+\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \\
 %
@@
 &\qquad
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
+\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2},
 \end{align}
-where (\ref{eq:14}) is from the SVRPG update, (\ref{eq:15}) is from Lemma \ref{lemma:aux0}, and (\ref{eq:16}) is from Young's inequality in the 'Peter-Paul' variant.
+where (14) is from the SVRPG update, (15) is from Lemma \ref{lemma:aux0}, and (16) is from Young's inequality in the 'Peter-Paul' variant.
 Finally:
@@
 	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
-	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:17}\\
+	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \\
 	%
@@
 	&\qquad
-	-c_{t+1}\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:18}\\
+	-c_{t+1}\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \\
 	%
@@
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \label{eq:19}\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \\
 	%
@@
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
-	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2\VARIS}{2(1-\gamma)^2b} \nonumber\\
+	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2W}{2(1-\gamma)^2b} \nonumber\\
 	%
@@
 	-\frac{d_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
-	-\frac{f_t}{b}\VARIS,\nonumber\\
+	-\frac{f_t}{b}W,\nonumber\\
 	&\geq R_t^{s+1}
 	+\Psi_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
-	-\frac{d_t}{N}\VARRF
-	-\frac{f_t}{b}\VARIS, \label{eq:20}
+	-\frac{d_t}{N}V
+	-\frac{f_t}{b}W,
 \end{align}
-where (\ref{eq:17}) is from (\ref{eq:13}), (\ref{eq:18}) is from (\ref{eq:16}), (\ref{eq:19}) is from Lemma \ref{lemma:aux0}, and (\ref{eq:20}) is from Assumption \ref{ass:REINFORCE}.
+where (17) is from (13), (18) is from (16), and (19) is from Lemma \ref{lemma:aux0} and (20) is from Assumption \ref{ass:REINFORCE}.
 To complete the proof, besides rearranging terms, we have to ensure that $\Psi_t>0$ for each $t$. This gives the constraints on $\alpha_t$ and $\beta_t$.
@@
 \sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
- \frac{\sum_{t=0}^{m-1}\left(R_{t+1}^{s+1} - R_t^{s+1}\right)}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
+ \frac{\sum_{t=0}^{m-1}\left(R_{t+1}^{s+1} - R_t^{s+1}\right)}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B} \\
  %
  &= \frac{\sum_{t=0}^{m-1}\left(\Ets[t+1]{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} - \Ets[t]{J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2}\right)}{\psi} \nonumber\\
- &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
+ &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \\
  %
  &= \frac{\Ets[m]{\sum_{t=0}^{m-1}\left(J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2\right)}}{\psi} \nonumber\\
- &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
+ &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \\
  %
- &= \frac{\Ets[m]{J(\vtheta_{m}^{s+1}) - c_{m}\norm[]{\vtheta_{m}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{0}^{s+1}) - c_{0}\norm[]{\vtheta_{0}^{s+1}-\tilde{\vtheta}^s}^2}}{\psi} \nonumber\\
- &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \label{eq:21}\\
+ &= \frac{\Ets[m]{J(\vtheta_{m}^{s+1}) - c_{t+1}\norm[]{\vtheta_{m}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{0}^{s+1}) - c_{t}\norm[]{\vtheta_{0}^{s+1}-\tilde{\vtheta}^s}^2}}{\psi} \nonumber\\
+ &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \\
  %
- &= \frac{\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B}, \label{eq:22}
+ &= \frac{\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B}
  \end{align}
-where (\ref{eq:21}) is obtained by telescoping the sum and (\ref{eq:22}) from the fact that $c_m=0$ and $\vtheta_0^{s+1}=\tilde{\vtheta}^s$.
 Next, summing over epochs:
@@
 \sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
-\frac{\sum_{s=0}^{S-1}\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \nonumber\\
+\frac{\sum_{s=0}^{S-1}\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \\
 %
@@
 \frac{\EVV[]{J(\tilde{\vtheta}^{S}) - J(\tilde{\vtheta}^{0})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}
- \label{eq:23}\\
+ \\
  %
  &\leq
-\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}, \label{eq:24}
+ \frac{\EVV[]{J(\vtheta_{out})} - J(\tilde{\vtheta}^{0})}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \\
+ %
+ &\leq
+ \frac{J(\vtheta^*) - J(\vtheta^0)}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}
 \end{align}
-where the expectation in (\ref{eq:23}) is \wrt all the trajectories sampled in a run of Algorithm \ref{alg:svrpg} and (\ref{eq:24}) is from the definition of $\vtheta^*$.
-
 Finally, we consider the expectation \wrt all sources of randomness, including the uniform sampling of the output parameter:
@@
 \EVV[]{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
-&=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} \nonumber\\
+&=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} \\
 &\leq
-\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B} \nonumber.
+\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B}
 \end{align}

### 2018-01-30 svrpg.tex
started writing Convergence Guarantees

--- 2018-01-30 svrpg.tex
@@
 
-\section{Algorithm}
+\section{Algorithm}\label{sec:algo}
 In on-line policy gradient problems, stochastic gradient ascent (SGA) is not really a choice, but is dictated by the necessity of interacting with an unknown environment.
@@
 Compared to the typical finite sum optimization scenario, we have the following additional challenges:
-\begin{itemize}
-	\item The objective $J(\vtheta)$ is typically non-convex;
-	\item The dataset $\mathcal{T}$ can be infinite;
-	\item The value of the parameter $\vtheta$ influences the sampling of trajectories;
-\end{itemize}
+\begin{enumerate}
+	\item \textit{Non-convexity}: the objective $J(\vtheta)$ is typically non-convex;
+	\item \textit{Approximation}: the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
+	\item \textit{Non-stationarity}: the value of the parameter $\vtheta$ influences the sampling of trajectories;
+\end{enumerate}
 To deal with non-convexity, we require $J(\vtheta)$ to be L-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and Softmax [appendix?].
@@
 \section{Convergence Guarantees}
+In this section we adapt existing convergence guarantees for nonconvex SVRG \cite{reddi2016stochastic} \cite{allen2016variance} to the Policy Gradient framework.
+Each of the three challenges presented at the beginning of Section \ref{sec:algo} can potentially prevent convergence, so we need additional assumptions:
+\begin{enumerate}
+\item \textit{Non-convexity}:
+	as already mentioned, a common assumption for nonconvex optimization is the L-smoothness of the objective. We actually require the following, which is stronger:
 	\begin{assumption}\label{ass:bounded_score}
@@
 	\end{assumption}
-
+	This property is satisfied by the commonly used Gaussian policy, and is a sufficient condition for the L-smoothness of the objective (see Lemma \ref{lemma:lsmooth} in the appendix).
+\item \textit{Approximation}:
+since we can only compute an estimate of the full gradient, we require its variance to be bounded:
+	\begin{assumption}\label{ass:REINFORCE}
+		There is a constant $V<\infty$ such that, for each policy $\pol$:
+		\[
+		\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq \VARRF.
+		\]
+	\end{assumption}
+This assumption is satisfied by the REINFORCE and G(PO)MDP estimators and Gaussian policies \cite{zhao2011analysis} \cite{pirotta2013adaptive}.
+The problem of full gradient approximation for SVRG has been addressed in the supervised learning framework \cite{harikandeh2015stopwasting} with results similar to ours.
+\item \textit{Non-stationarity}:
+the variance introduced by importance sampling must be bounded:
 	\begin{assumption}\label{ass:M2}
@@
 		\[
-	\mathbb{V}ar\left(\omega(\tau)\right) \leq \VARIS
+		\mathbb{V}ar\left(\omega(\tau)\right) \leq \VARIS.
 		\]
 	\end{assumption}
+Differently from the previous two, this assumption must be enforced by a proper handling of the epoch size $m$.
+\end{enumerate}
 
-\begin{assumption}\label{ass:REINFORCE}
-There is a constant $V<\infty$ such that, for each policy $\pol$:
-\[
-	\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq \VARRF
-\]
-\end{assumption}
-
-\begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
-Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth.
-\end{restatable}
-
+We can now state the following key result (proof in Appendix \ref{app:proofs}):
 \begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
-Under Assumptions \ref{ass:bounded_score} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size, the following property:
+Under Assumptions \ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
 \begin{align*}
@@
 \end{restatable}
+\todomat{Comment theorem}
 
@@
 Before addressing the main convergence theorem, we prove some useful lemmas.
+
+\begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
+	 Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth.
+\end{restatable}
+\begin{proof}
+\todomat{...}
+\end{proof}
+
 %Lemma 2
@@
 
-\section{Experimental Details}\label{app:proofs}
+\section{Applicability to Gaussian Policies}\label{app:pol}
+\todomat{...}
+
+\section{Experimental Details}\label{app:exp}
 Setting specifications for all the used environments:

### 2018-01-31 svrpg.tex
added comment to main theorem

@@
 
-We can now state the following key result (proof in Appendix \ref{app:proofs}):
+We can now state the following key result:
 \begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
@@
 \end{restatable}
-\todomat{Comment theorem}
+Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
+In Section \ref{sec:stopping}, instead, we propose a joint selection of step size $\alpha$ and epoch size $m$ which is more suited for applications.
+The $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on non-convex SVRG, \eg \cite{reddi2016stochastic}; the $\mathcal{O}(\nicefrac{1}{N})$ term is due to full gradient approximation and is analogous to the one in \cite{harikandeh2015stopwasting}; the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
+To achieve true convergence, both the batch size $N$ and the mini-batch size $B$ should increase over time.
+In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by full gradient approximation and importance sampling respectively.
+Since the optimal configuration may be task-dependent, we leave this choice to the empirical analysis of Section \ref{sec:exp}.
 
@@
 
-\section{Experiments}
+\section{Experiments}\label{sec:exp}
 We have designed our experiments to investigate whether:

### 2018-01-31 svrpg.tex
added proof of L-smoothness lemma

@@
 	\end{assumption}
-	This property is satisfied by the commonly used Gaussian policy, and is a sufficient condition for the L-smoothness of the objective (see Lemma \ref{lemma:lsmooth} in the appendix).
+	This property is satisfied by the commonly used truncated Gaussian policy, and is a sufficient condition for the L-smoothness of the objective (see Lemma \ref{lemma:lsmooth} in the appendix).
 \item \textit{Approximation}:
@@
 \begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
-	 Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth.
+	Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth for some positive Lipschitz constant $L$.
 \end{restatable}
 \begin{proof}
-\todomat{...}
+By definition of $J(\vtheta)$:
+\begin{align}
+\Dij{J(\vtheta)}{\theta} 
+&= \int_{\Tspace}\Dij{}{\theta}\pol(\tau)\Reward(\tau)\de \tau
+\nonumber\\ 
+&= \int_{\Tspace}\pol(\tau)\score{\theta}{\tau}\score{\theta}{\tau}^T\Reward(\tau)\de \tau + \int_{\Tspace}\pol(\tau)\Dij{}{\theta}\log\pol(\tau)\Reward(\tau)\de \tau \nonumber\\
+&\leq \sup_{\tau} \left\{\left|\Reward(\tau)\right|\right\} \left(\GRADLOG^2+\HESSLOG\right) \label{eq:0}\\
+&= \frac{R}{1-\gamma}\left(\GRADLOG^2+\HESSLOG\right),\nonumber
+\end{align}
+where \ref{eq:0} is from Assumption \ref{ass:bounded_score}.
+Since the Hessian is bounded, $J(\vtheta)$ is Lipschitz-smooth.
 \end{proof}

### 2018-02-01 svrpg.tex
correzioni sulla sezione  practical variants

@@
 
-As reported in Section \ref{subsec:PolicyGradient} in practical variants we use G(PO)MDP gradient estimator~\cite{baxter2001infinite}, that brings the same guarantees of Reinforce Estimator.
-One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}:  this allows a reduction of the gradient estimator variance.
+As reported in Section \ref{subsec:PolicyGradient} we use G(PO)MDP gradient estimator~\cite{baxter2001infinite}, which gives us the same guarantees we have for the Reinforce Estimator.
+One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}, allowing a reduction of the gradient estimator variance.
 The off-policy gradient estimator becomes:

### 2018-02-01 svrpg.tex

--- 2018-01-31 svrpg.tex
@@
 \section{Preliminaries}
-In this section we provide essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
-In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
+In this section we provide the essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
+%In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
 
 \subsection{Policy Gradient}\label{subsec:PolicyGradient}
-A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
+A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$;
+% and $R$ is the maximum absolute-value reward;
+$\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
 The agent's behavior is modeled as a policy $\pi$, where $\pi(\cdot|s)$ is the density distribution over $\Aspace$ in state $s$.
@@
 \end{align*}
-Solving an MDP means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
+Solving an MDP $M$ means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
 
-Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt the $m$-dimensional parameter vector $\vtheta$.
+Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$.
 We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
-The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
-\todopir{stochastic or mini-batch gradient}
-\begin{align*}
+The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy, Peters2008reinf}:
+\begin{align} \label{E:policygradient}
 \gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
-\end{align*}
-We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment. 
-In this setting, $(N>0)$-batch gradient ascent is typically employed: after sampling a batch of $N$ trajectories (\ie $\Dataset_{N}$),
-policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient. 
-\todopir{step size is scalar or vectorial (just for notation)? Check consistency of $N$ with finite-sum function, we can use even $b$ or $B$ for the batch size.}
-The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
-\begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
-\end{align*}
-where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
-\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,anonymous2018variance}.}
-The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
-\begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\mid s_k^n}\right)\left(\gamma^h r_h^n - b\right).
-\end{align*}
-\todopir{Cite Jan's paper somewhere}
+\end{align}
+Notice that the distribution defining the gradient is induced by the current policy.
+This aspect introduces a nonstationarity in the sampling process.
+Since the underlying distribution is changing over time, it is necessary to resample at each update or use weighting techniques such as importance sampling.
+
+We consider the \emph{on-line learning problem}, where trajectories are sampled by interacting with the environment at each policy change. 
+In this setting, stochastic gradient ascent is typically employed.
+At each iteration $k >0$, after sampling a batch of $N$ ($0<N<+\infty$) trajectories with policy $\pi_{\vtheta_k}$,
+the policy is updated as $\vtheta  = \vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $N$ samples.
+Several estimators have been provided in literature as empirical average over trajectories:
+\begin{align} \label{E:policygradient.estimate}
+        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} \nabla f(\tau^i).
+\end{align}
+The most widespread estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
+We refer the reader to App.~\ref{A:gradient_estimators} for a formal definition.
+Both are unbiased estimate of~\eqref{E:policygradient} but G(PO)MDP is subject to less variance than REINFORCE.
 \todopir{At this point we need to connect with the supervised learning framework (we can just define everything inside the first sum as $\nabla g(\tau)$).}
@@
 
-\section{Algorithm}\label{sec:algo}
+\section{Algorithm}
 In on-line policy gradient problems, stochastic gradient ascent (SGA) is not really a choice, but is dictated by the necessity of interacting with an unknown environment.
@@
 Compared to the typical finite sum optimization scenario, we have the following additional challenges:
-\begin{enumerate}
-	\item \textit{Non-convexity}: the objective $J(\vtheta)$ is typically non-convex;
-	\item \textit{Approximation}: the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
-	\item \textit{Non-stationarity}: the value of the parameter $\vtheta$ influences the sampling of trajectories;
-\end{enumerate}
+\begin{itemize}
+	\item The objective $J(\vtheta)$ is typically non-convex;
+	\item The dataset $\mathcal{T}$ can be infinite;
+	\item The value of the parameter $\vtheta$ influences the sampling of trajectories;
+\end{itemize}
 To deal with non-convexity, we require $J(\vtheta)$ to be L-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and Softmax [appendix?].
@@
 \section{Convergence Guarantees}
-In this section we adapt existing convergence guarantees for nonconvex SVRG \cite{reddi2016stochastic} \cite{allen2016variance} to the Policy Gradient framework.
-Each of the three challenges presented at the beginning of Section \ref{sec:algo} can potentially prevent convergence, so we need additional assumptions:
-\begin{enumerate}
-\item \textit{Non-convexity}:
-	as already mentioned, a common assumption for nonconvex optimization is the L-smoothness of the objective. We actually require the following, which is stronger:
 \begin{assumption}\label{ass:bounded_score}
@@
 \end{assumption}
-	This property is satisfied by the commonly used Gaussian policy, and is a sufficient condition for the L-smoothness of the objective (see Lemma \ref{lemma:lsmooth} in the appendix).
-\item \textit{Approximation}:
-since we can only compute an estimate of the full gradient, we require its variance to be bounded:
+
+\begin{assumption}\label{ass:M2}
+	There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
+\[
+	\mathbb{V}ar\left(\omega(\tau)\right) \leq \VARIS
+\]
+\end{assumption}
+
 \begin{assumption}\label{ass:REINFORCE}
@@
 \[
-		\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq \VARRF.
+	\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq \VARRF
 \]
 \end{assumption}
-This assumption is satisfied by the REINFORCE and G(PO)MDP estimators and Gaussian policies \cite{zhao2011analysis} \cite{pirotta2013adaptive}.
-The problem of full gradient approximation for SVRG has been addressed in the supervised learning framework \cite{harikandeh2015stopwasting} with results similar to ours.
-\item \textit{Non-stationarity}:
-the variance introduced by importance sampling must be bounded:
-	\begin{assumption}\label{ass:M2}
-		There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
-		\[
-		\mathbb{V}ar\left(\omega(\tau)\right) \leq \VARIS.
-		\]
-	\end{assumption}
-Differently from the previous two, this assumption must be enforced by a proper handling of the epoch size $m$.
-\end{enumerate}
 
-We can now state the following key result:
+\begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
+Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth.
+\end{restatable}
+
 \begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
-Under Assumptions \ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
+Under Assumptions \ref{ass:bounded_score} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size, the following property:
 \begin{align*}
@@
 \end{restatable}
-Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
-In Section \ref{sec:stopping}, instead, we propose a joint selection of step size $\alpha$ and epoch size $m$ which is more suited for applications.
-The $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on non-convex SVRG, \eg \cite{reddi2016stochastic}; the $\mathcal{O}(\nicefrac{1}{N})$ term is due to full gradient approximation and is analogous to the one in \cite{harikandeh2015stopwasting}; the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
-To achieve true convergence, both the batch size $N$ and the mini-batch size $B$ should increase over time.
-In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by full gradient approximation and importance sampling respectively.
-Since the optimal configuration may be task-dependent, we leave this choice to the empirical analysis of Section \ref{sec:exp}.
 
@@
 
-\section{Experiments}\label{sec:exp}
+\section{Experiments}
 We have designed our experiments to investigate whether:
@@
 Before addressing the main convergence theorem, we prove some useful lemmas.
-
-\begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
-	 Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth.
-\end{restatable}
-\begin{proof}
-\todomat{...}
-\end{proof}
-
 %Lemma 2
@@
 
-\section{Applicability to Gaussian Policies}\label{app:pol}
-\todomat{...}
-
-\section{Experimental Details}\label{app:exp}
+\section{Experimental Details}\label{app:proofs}
 Setting specifications for all the used environments:

### 2018-02-01 svrpg.tex
Modified Hypermarapeters section

--- 2018-02-01 svrpg.tex
@@
 
-\section{Stopping Condition}\label{sec:stopping}
+\section{Hyperparameters}\label{sec:stopping}
+In Reinforcement learning we have to take into account the gradient's variance.
+A used way for reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2015adam}, a method that computes adaptive learning rates for each parameter.
+Adam help to stabilize the gradient estimate.
+In fact, this method computes adaptive learning rates for each parameter that depends on the history of the gradients' variance.
+We use two different adam hyperparameters: one for updates tied to Snapshots points and the other tied to the updates of sub-iterations points. 
+This is because we're dealing with two different variance dimensions.
+Denote by $\alpha_{snap}$ the last learning rate computed into the snapshot point and denote by  $\alpha_{sub-iter}$ the last learning rate computed into the sub-iteration point.
+
 As shown, the performance of the SVRPG algorithm is sensitive to the number of sub-iterations.
@@
 We introduce two stopping condition: the maximum number of sub-iterations and the condition on learning rate of Adam sub-iterations.
-Adam update rule help to stabilize the gradient estimate. In fact, this method computes adaptive learning rates for each parameter that depends on the history of the gradients' variance.
-In particular we use two different Adam parameters: the first one for snapshot updates and the second one for the sub-iterations updates.
 We stop with sub-iterations if:

### 2018-02-01 svrpg.tex
Corrected the citation

@@
 In Reinforcement learning we have to take into account the gradient's variance.
-A used way for reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2015adam}, a method that computes adaptive learning rates for each parameter.
+A used way for reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter.
 Adam help to stabilize the gradient estimate.

### 2018-02-01 svrpg.tex
started writing intro

@@
 \section{Introduction}
-\begin{itemize}
-        \item Similarities between RL and SL
-        \item SL needs SVRG so RL
-        \item Potential advantages of SVRG
-        \item Difficulties in RL \wrt SL
-\end{itemize}
+%Similarities between RL and SL
+On a very general level, machine learning addresses the problem of artificial agents that must select the right actions to solve a given task.
+The approach of Reinforcement Learning \cite{sutton2011reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal.
+This makes RL fundamentally different from Supervised Learning (SL), where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels).
+However, the two approaches share many challenges and, as a consequence, solution concepts.
+The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or a policy of behavior directly.
+In particular, when the tasks are characterized by large or continuous state-action spaces, RL needs to borrow the powerful function approximators that are the main subject of study of SL, such as neural networks.
+Among RL solutions for learning continuous tasks, policy gradient methods bear the closest similarity to SL, since they allow to formulate the RL problem as the unconstrained optimization of an objective function (\ie a negative loss) of policy parameters and data samples.
+
+% SL needs SVRG so RL
+In supervised learning, stochastic optimization is often employed for reasons of efficiency.
+The idea is to evaluate the objective on a random subset of the available data to obtain an approximate gradient direction.
+This can be much less expensive that taking into account all the data for all parameter updates.
+However, the sub-sampling of data samples is a source of variance that can potentially compromise convergence, so that speed and variance must be traded off with proper handling of meta-parameters.
+Variance-reduced gradient algorithms such as SAG, SVRG and SAGA offer better ways of achieving the optimal trade-off between variance and speed, both in theory and practice. 
+In Reinforcement Learning, stochastic optimization is not even a choice, since samples must be sequentially obtained by interacting with an initially unknown environment.
+In this scenario, containing the variance of the estimates is a problem that cannot be avoided, and methods like SVRG can offer a more efficient alternative
+
+\todomat{work in progress}
+% Potential advantages of SVRG
+% Difficulties in RL \wrt SL
 
@@
 where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
-\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,anonymous2018variance}.}
+\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,wu2018variance}.}
 The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
@@
 
-\section{Hyperparameters}\label{sec:stopping}
-In Reinforcement learning we have to take into account the gradient's variance.
-A used way for reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter.
-Adam help to stabilize the gradient estimate.
-In fact, this method computes adaptive learning rates for each parameter that depends on the history of the gradients' variance.
-We use two different adam hyperparameters: one for updates tied to Snapshots points and the other tied to the updates of sub-iterations points. 
-This is because we're dealing with two different variance dimensions.
-Denote by $\alpha_{snap}$ the last learning rate computed into the snapshot point and denote by  $\alpha_{sub-iter}$ the last learning rate computed into the sub-iteration point.
-
+\section{Stopping Condition}\label{sec:stopping}
 As shown, the performance of the SVRPG algorithm is sensitive to the number of sub-iterations.
@@
 We introduce two stopping condition: the maximum number of sub-iterations and the condition on learning rate of Adam sub-iterations.
+Adam update rule help to stabilize the gradient estimate. In fact, this method computes adaptive learning rates for each parameter that depends on the history of the gradients' variance.
+In particular we use two different Adam parameters: the first one for snapshot updates and the second one for the sub-iterations updates.
 We stop with sub-iterations if:

### 2018-02-01 svrpg.tex

@@
 
-\section{Stopping Condition}\label{sec:stopping}
+\section{Hyperparameters}\label{sec:stopping}
+In Reinforcement learning we have to take into account the gradient's variance.
+A used way for reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter.
+Adam help to stabilize the gradient estimate.
+In fact, this method computes adaptive learning rates for each parameter that depends on the history of the gradients' variance.
+We use two different adam hyperparameters: one for updates tied to Snapshots points and the other tied to the updates of sub-iterations points. 
+This is because we're dealing with two different variance dimensions.
+Denote by $\alpha_{snap}$ the last learning rate computed into the snapshot point and denote by  $\alpha_{sub-iter}$ the last learning rate computed into the sub-iteration point.
+
 As shown, the performance of the SVRPG algorithm is sensitive to the number of sub-iterations.

### 2018-02-01 svrpg.tex
changed notation of ideal gradient
made second paragraph of intro more explicit (still missing citations)

@@
 \newcommand{\dotprod}[2]{\left\langle#1,#2\right\rangle}
+\newcommand{\dnabla}{\nabla\!\!\!\!\nabla}
 %RL
@@
 \newcommand{\gradBlack}[1]{\blacktriangledown J(#1)}
-\newcommand{\gradIdeal}[1]{\overline{\blacktriangledown} J(#1)}
+\newcommand{\gradIdeal}[1]{\dnabla J(#1)}
 \newcommand{\VARRF}{V}
@@
 %Similarities between RL and SL
-On a very general level, machine learning addresses the problem of artificial agents that must select the right actions to solve a given task.
+On a very general level, machine learning addresses the problem of an agent that must select the right actions to solve a given task.
 The approach of Reinforcement Learning \cite{sutton2011reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal.
-This makes RL fundamentally different from Supervised Learning (SL), where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels).
+This makes RL fundamentally different from Supervised Learning (SL), even in its online version, where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels).
 However, the two approaches share many challenges and, as a consequence, solution concepts.
 The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or a policy of behavior directly.
-In particular, when the tasks are characterized by large or continuous state-action spaces, RL needs to borrow the powerful function approximators that are the main subject of study of SL, such as neural networks.
-Among RL solutions for learning continuous tasks, policy gradient methods bear the closest similarity to SL, since they allow to formulate the RL problem as the unconstrained optimization of an objective function (\ie a negative loss) of policy parameters and data samples.
+Moreover, when the tasks are characterized by large or continuous state-action spaces, RL needs to borrow the powerful function approximators that are the main subject of study of SL, such as neural networks.
 
 % SL needs SVRG so RL
-In supervised learning, stochastic optimization is often employed for reasons of efficiency.
+In a typical discriminative supervised learning setting, an objective function $J(x\mid\vtheta)$ has to be optimized \wrt to a parameter vector $\vtheta$.
+The training set $\mathcal{D}_{train}$ is the set of data samples $x$ for which the objective (and its gradient) can be evaluated.
+In most practical cases, $\mathcal{D}_{train}$ is only a subset of the set $\mathcal{D}$ of all the $x$ of interest, hence we resort to stochastic optimization of the expected value of $J$ over an unknown data distribution.
+The special case in which $\mathcal{D}_{train} = \mathcal{D}$ is known as Empirical Risk Minimization (ERM), and allows for deterministic optimization.
+Even in this case, stochastic optimization is often used for reasons of efficiency.
 The idea is to evaluate the objective on a random subset of the available data to obtain an approximate gradient direction.
-This can be much less expensive that taking into account all the data for all parameter updates.
+This can be much less expensive than taking into account all the data for all parameter updates.
 However, the sub-sampling of data samples is a source of variance that can potentially compromise convergence, so that speed and variance must be traded off with proper handling of meta-parameters.
 Variance-reduced gradient algorithms such as SAG, SVRG and SAGA offer better ways of achieving the optimal trade-off between variance and speed, both in theory and practice.
-In Reinforcement Learning, stochastic optimization is not even a choice, since samples must be sequentially obtained by interacting with an initially unknown environment.
-In this scenario, containing the variance of the estimates is a problem that cannot be avoided, and methods like SVRG can offer a more efficient alternative
+Although designed explicitly for ERM, this algorithms address a problem that is equally important, if not more critical, in more general machine learning problems.
+In Reinforcement Learning, as in online learning n general, stochastic optimization is rarely a matter of choice, since data must be actively sampled by interacting with an initially unknown environment.
+In this scenario, containing the variance of the estimates is a problem that cannot be avoided, and methods like SVRG can offer a more efficient alternative, or a complement, to traditional methods such as batch learning.
 
@@
 % Difficulties in RL \wrt SL
+
+% Among RL solutions for learning continuous tasks, policy gradient methods bear the closest similarity to SL, since they allow to formulate the RL problem as the unconstrained optimization of an objective function (\ie a negative loss) of policy parameters and data samples.
 
@@
 \begin{align*}
-	\overline{\blacktriangledown}J(\vtheta_t^{s+1}) &\coloneqq 
+	\gradIdeal{\vtheta_t^{s+1}} &\coloneqq 
 	\gradJ{\tilde{\vtheta}^s}

### 2018-02-01 svrpg.tex
added comments to Hyperparameters section

@@
 
-\section{Convergence Guarantees}
+\section{Convergence Guarantees}\label{sec:conv}
 In this section we adapt existing convergence guarantees for nonconvex SVRG \cite{reddi2016stochastic} \cite{allen2016variance} to the Policy Gradient framework.
@@
 In Reinforcement learning we have to take into account the gradient's variance.
-A used way for reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter.
-Adam help to stabilize the gradient estimate.
-In fact, this method computes adaptive learning rates for each parameter that depends on the history of the gradients' variance.
-We use two different adam hyperparameters: one for updates tied to Snapshots points and the other tied to the updates of sub-iterations points. 
-This is because we're dealing with two different variance dimensions.
+\todomat{$\uparrow$ the opening of the section could connect better with previous sections. Declare which meta-parameters will be addressed in the section}
+A common way to reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter.
+Adam helps to stabilize the gradient estimate by computing adaptive learning rates for each parameter based on the history of the gradients' variance.
+We use two different adam hyperparameters: one for updates tied to snapshot points and the other tied to the updates of sub-iterations points.
+\todomat{$\uparrow$ explain better: it's not only two separate meta-parameters, but two separate learning rate schedules.
+Use notation from previous sections, \eg snaphot is usually associated with $\tilde{\vtheta}$}
+This is because we are dealing with two different variance magnitudes.
+\todomat{$\uparrow$ the reason is this, but try to spend more words on it}
 Denote by $\alpha_{snap}$ the last learning rate computed into the snapshot point and denote by  $\alpha_{sub-iter}$ the last learning rate computed into the sub-iteration point.
 
-As shown, the performance of the SVRPG algorithm is sensitive to the number of sub-iterations.
+As shown in Section \ref{sec:conv}, the performance of the SVRPG algorithm is sensitive to the epoch size $m$.
 As the number of sub-iterations increases, the variance of the importance weights and the gradient increases too.
 Therefore, the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
-A fixed number of sub-iterations is not the optimal choice.
+A fixed epoch size is not the optimal choice.
 Few sub-iterations lead to few updates, many update lead to increases of variance.
 
-We introduce two stopping condition: the maximum number of sub-iterations and the condition on learning rate of Adam sub-iterations.
-We stop with sub-iterations if:
+We introduce two stopping conditions: the maximum number of sub-iterations and the condition on learning rate of Adam sub-iterations.
+\todomat{I find "two stopping conditions" confusing: it's just one with an additional constraint}
+We stop with inner iterations if:
 
@@
 
+\todomat{why? should mention how this formula combines trajectory efficiency and variance control}
 After that we sample N trajectories for the next Snapshot point.

### 2018-02-01 svrpg.tex
Updated

--- 2018-02-01 svrpg.tex
@@
 \newcommand{\gradJ}[1]{\nabla J(#1)}
-\newcommand{\gradApp}[2]{\hat{\nabla}_{#2}J(#1)}
+\newcommand{\gradApp}[2]{\widehat{\nabla}_{#2}J(#1)}
 \newcommand{\eqdef}{\mathrel{\mathop:}=}
@@
 \section{Introduction}
-%Similarities between RL and SL
-On a very general level, machine learning addresses the problem of artificial agents that must select the right actions to solve a given task.
-The approach of Reinforcement Learning \cite{sutton2011reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal.
-This makes RL fundamentally different from Supervised Learning (SL), where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels).
-However, the two approaches share many challenges and, as a consequence, solution concepts.
-The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or a policy of behavior directly.
-In particular, when the tasks are characterized by large or continuous state-action spaces, RL needs to borrow the powerful function approximators that are the main subject of study of SL, such as neural networks.
-Among RL solutions for learning continuous tasks, policy gradient methods bear the closest similarity to SL, since they allow to formulate the RL problem as the unconstrained optimization of an objective function (\ie a negative loss) of policy parameters and data samples.
-
-% SL needs SVRG so RL
-In supervised learning, stochastic optimization is often employed for reasons of efficiency.
-The idea is to evaluate the objective on a random subset of the available data to obtain an approximate gradient direction.
-This can be much less expensive that taking into account all the data for all parameter updates.
-However, the sub-sampling of data samples is a source of variance that can potentially compromise convergence, so that speed and variance must be traded off with proper handling of meta-parameters.
-Variance-reduced gradient algorithms such as SAG, SVRG and SAGA offer better ways of achieving the optimal trade-off between variance and speed, both in theory and practice. 
-In Reinforcement Learning, stochastic optimization is not even a choice, since samples must be sequentially obtained by interacting with an initially unknown environment.
-In this scenario, containing the variance of the estimates is a problem that cannot be avoided, and methods like SVRG can offer a more efficient alternative
-
-\todomat{work in progress}
-% Potential advantages of SVRG
-% Difficulties in RL \wrt SL
+\begin{itemize}
+        \item Similarities between RL and SL
+        \item SL needs SVRG so RL
+        \item Potential advantages of SVRG
+        \item Difficulties in RL \wrt SL
+\end{itemize}
 
 \section{Preliminaries}
-In this section we provide essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
-In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
+In this section we provide the essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
+%In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
 
 \subsection{Policy Gradient}\label{subsec:PolicyGradient}
-A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
-The agent's behavior is modeled as a policy $\pi$, where $\pi(\cdot|s)$ is the density distribution over $\Aspace$ in state $s$.
+A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $M = \{\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\}$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$;
+% and $R$ is the maximum absolute-value reward;
+$\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
+The agent's behaviour is modelled as a policy $\pi$, where $\pi(\cdot|s)$ is the density distribution over $\Aspace$ in state $s$.
 % We consider episodic tasks, \ie tasks composed of episodes of length $H$, also called time horizon.
 % In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
-We consider episodic MDPs with effective horizon $H$, \ie optimal policy is able to reach the target state (\ie absorbing state) in at most $H$ steps (the true length depend on the initial state).
-This has not to be confused with a finite horizon problem where the optimal policy is non-stationary.
-In this setting, we can limit our attention to trajectories of length $H$.
+We consider episodic MDPs with effective horizon $H$.\footnote{The episode duration is a random variable but the optimal policy is able to reach the target state (\ie absorbing state) in at most $H$ steps.
+This has not to be confused with a finite horizon problem where the optimal policy is non-stationary.} In this setting, we can limit our attention to trajectories of length $H$.
 A trajectory $\tau$ is a sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following a stationary policy, where $s_0 \sim \rho$.
 With abuse of notation, we denote with $\pi(\tau)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories, and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
-\begin{align*}
-\Reward(\tau) = \sum_{t=0}^{H-1}\gamma^t\Reward(s_t,a_t).
-\end{align*}
-Policies can be ranked based on their expected total reward:
-\begin{align*}
-	J(\pi) = \int_{\Tspace}\pi(\tau)\Reward(\tau)\de \tau.
-\end{align*}
-Solving an MDP means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
+%
+$\Reward(\tau) = \sum_{t=0}^{H-1}\gamma^t\Reward(s_t,a_t).$
+%
+Policies can be ranked based on their expected total reward: $J(\pi) = \EVV[\tau \sim \pi]{\Reward(\tau)|M}$.
+Solving an MDP $M$ means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
 
-Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt the $m$-dimensional parameter vector $\vtheta$.
-We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
-The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
-\todopir{stochastic or mini-batch gradient}
-\begin{align*}
-\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
-\end{align*}
-We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment. 
-In this setting, $(N>0)$-batch gradient ascent is typically employed: after sampling a batch of $N$ trajectories (\ie $\Dataset_{N}$),
-policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient. 
-\todopir{step size is scalar or vectorial (just for notation)? Check consistency of $N$ with finite-sum function, we can use even $b$ or $B$ for the batch size.}
-The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
-\begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
-\end{align*}
-where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
-\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,wu2018variance}.}
-The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
-\begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\mid s_k^n}\right)\left(\gamma^h r_h^n - b\right).
-\end{align*}
-\todopir{Cite Jan's paper somewhere}
+Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$.
+We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.\todopirout{See if it used}
+The search for a locally optimal policy is performed by means of gradient ascent, where the policy gradient
+%of $J(\vtheta)$ \wrt the policy parameters 
+is \cite{sutton2000policy, Peters2008reinf}:
+\begin{align} \label{E:policygradient}
+        \gradJ{\vtheta} = \EVV[\tau \sim \pol]{\score{\vtheta}{\tau}\Reward(\tau)}
+        %\int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
+\end{align}
+Notice that the distribution defining the gradient is induced by the current policy.
+This aspect introduces a nonstationarity in the sampling process.
+Since the underlying distribution changes over time, it is necessary to resample at each update or use weighting techniques such as importance sampling.
+Here, we consider the \emph{on-line learning scenario}, where trajectories are sampled by interacting with the environment at each policy change. 
+In this setting, stochastic gradient ascent is typically employed.
+At each iteration $k >0$, a batch $\mathcal{D}_k = \{\tau^i\}_{i=0}^N$ of $N>0$ trajectories is collected using policy $\pi_{\vtheta_k}$.
+The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $\mathcal{D}_k$, \ie
+\begin{align} \label{E:policygradient.estimate}
+        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} \nabla f(\tau^i).
+\end{align}
+The most widespread gradient estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
+We refer the reader to App.~\ref{A:gradient_estimators} for a formal definition.
+Both are unbiased estimate of~\eqref{E:policygradient} but G(PO)MDP is subject to less variance than REINFORCE.
 \todopir{At this point we need to connect with the supervised learning framework (we can just define everything inside the first sum as $\nabla g(\tau)$).}

### 2018-02-01 svrpg.tex
aggiornamento sezione A.1

--- 2018-02-01 svrpg.tex
@@
 	&\mathbb{E}_{t\mid s}\left[X\right] \coloneqq 
-		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}^s) \\ \tau_i\sim\pi(\cdot\mid\vtheta^{s+1}_i)\text{ for $i=0,\dots,t-1$}}}{\left[X \mid \tilde{\vtheta^s}\right]} \\
-	&\coloneqq \EVV[\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}^s)]{
-			\EVV[\tau_0\sim\pi(\cdot\mid\vtheta_0^{s+1})]
+		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}^s)\forall j \in {\it {\bf N}} \\ \tau_{i,h}\sim\pi(\cdot\mid\vtheta^{s+1}_i) \forall h \in {\it {\bf B}}, \text{ for $i=0,\dots,t-1$}}}{\left[X \mid \tilde{\vtheta^s}\right]} \\
+	&\coloneqq \EVV[\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}^s)\forall j \in {\it {\bf N}}]{
+			\EVV[\tau_{0,h}\sim\pi(\cdot\mid\vtheta_0^{s+1}) \forall h \in {\it {\bf B}}]
 				{\dots
-					\EVV[\tau_t\sim\pi(\cdot\mid\vtheta_t^{s+1})]
+					\EVV[\tau_{t,h}\sim\pi(\cdot\mid\vtheta_t^{s+1})\forall h \in {\it {\bf B}}]
 						{X\mid\vtheta_t^{s+1}}
@@
 \end{align*}
-where the sequence $\tilde{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}. 
+where the sequence $\tilde{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}, ${\it {\bf N}}$ is the amount of trajectories sampled for the snapshot and ${\it {\bf B}}$ is the amount of trajectories sampled for a sub iteration.
 \end{definition}

### 2018-02-01 svrpg.tex
correzione lemma 3.1

--- 2018-02-01 svrpg.tex
@@
 On a very general level, machine learning addresses the problem of an agent that must select the right actions to solve a given task.
-The approach of Reinforcement Learning \cite{sutton2011reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal.
+The approach of Reinforcement Learning~\citep{sutton1998reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal.
 This makes RL fundamentally different from Supervised Learning (SL), even in its online version, where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels).
 However, the two approaches share many challenges and, as a consequence, solution concepts.
-The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or a policy of behavior directly.
+The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or a policy of behaviour directly.
 Moreover, when the tasks are characterized by large or continuous state-action spaces, RL needs to borrow the powerful function approximators that are the main subject of study of SL, such as neural networks.
@@
 \section{Preliminaries}
-In this section we provide essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
-In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
+In this section we provide the essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
+%In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
 
 \subsection{Policy Gradient}\label{subsec:PolicyGradient}
-A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
-The agent's behavior is modeled as a policy $\pi$, where $\pi(\cdot|s)$ is the density distribution over $\Aspace$ in state $s$.
+A Reinforcement Learning task~\citep{sutton1998reinforcement} can be modelled with a discrete-time continuous Markov Decision Process (MDP) $M = \{\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\}$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$;
+% and $R$ is the maximum absolute-value reward;
+$\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
+The agent's behaviour is modelled as a policy $\pi$, where $\pi(\cdot|s)$ is the density distribution over $\Aspace$ in state $s$.
 % We consider episodic tasks, \ie tasks composed of episodes of length $H$, also called time horizon.
 % In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
-We consider episodic MDPs with effective horizon $H$, \ie optimal policy is able to reach the target state (\ie absorbing state) in at most $H$ steps (the true length depend on the initial state).
-This has not to be confused with a finite horizon problem where the optimal policy is non-stationary.
-In this setting, we can limit our attention to trajectories of length $H$.
+We consider episodic MDPs with effective horizon $H$.\footnote{The episode duration is a random variable but the optimal policy is able to reach the target state (\ie absorbing state) in at most $H$ steps.
+This has not to be confused with a finite horizon problem where the optimal policy is non-stationary.} In this setting, we can limit our attention to trajectories of length $H$.
 A trajectory $\tau$ is a sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following a stationary policy, where $s_0 \sim \rho$.
 With abuse of notation, we denote with $\pi(\tau)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories, and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
-\begin{align*}
-\Reward(\tau) = \sum_{t=0}^{H-1}\gamma^t\Reward(s_t,a_t).
-\end{align*}
-Policies can be ranked based on their expected total reward:
-\begin{align*}
-	J(\pi) = \int_{\Tspace}\pi(\tau)\Reward(\tau)\de \tau.
-\end{align*}
-Solving an MDP means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
-
-Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt the $m$-dimensional parameter vector $\vtheta$.
-We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
-The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
-\todopir{stochastic or mini-batch gradient}
-\begin{align*}
-\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
-\end{align*}
-We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment. 
-In this setting, $(N>0)$-batch gradient ascent is typically employed: after sampling a batch of $N$ trajectories (\ie $\Dataset_{N}$),
-policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient. 
-\todopir{step size is scalar or vectorial (just for notation)? Check consistency of $N$ with finite-sum function, we can use even $b$ or $B$ for the batch size.}
-The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
-\begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
-\end{align*}
-where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
-\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,wu2018variance}.}
-The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
-\begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\mid s_k^n}\right)\left(\gamma^h r_h^n - b\right).
-\end{align*}
-\todopir{Cite Jan's paper somewhere}
-\todopir{At this point we need to connect with the supervised learning framework (we can just define everything inside the first sum as $\nabla g(\tau)$).}
+%
+$\Reward(\tau) = \sum_{t=0}^{H-1}\gamma^t\Reward(s_t,a_t).$
+%
+Policies can be ranked based on their expected total reward: $J(\pi) = \EVV[\tau \sim \pi]{\Reward(\tau)|M}$.
+Solving an MDP $M$ means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
+
+Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$.
+We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.\todopirout{See if it used}
+The search for a locally optimal policy is performed by means of gradient ascent, where the policy gradient
+%of $J(\vtheta)$ \wrt the policy parameters 
+is \cite{sutton2000policy, Peters2008reinf}:
+\begin{align} \label{E:policygradient}
+        \gradJ{\vtheta} = \EVV[\tau \sim \pol]{\score{\vtheta}{\tau}\Reward(\tau)}
+        %\int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
+\end{align}
+Notice that the distribution defining the gradient is induced by the current policy.
+This aspect introduces a nonstationarity in the sampling process.
+Since the underlying distribution changes over time, it is necessary to resample at each update or use weighting techniques such as importance sampling.
+Here, we consider the \emph{online learning scenario}, where trajectories are sampled by interacting with the environment at each policy change. 
+In this setting, stochastic gradient ascent is typically employed.
+At each iteration $k >0$, a batch $\mathcal{D}_N^k = \{\tau_i\}_{i=0}^N$ of $N>0$ trajectories is collected using policy $\pi_{\vtheta_k}$.
+The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $\mathcal{D}_N^k$, \ie
+\begin{align} \label{E:policygradient.estimate}
+        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} \nabla f(\tau_i), \quad \tau_i \in \mathcal{D}_N^k.
+\end{align}
+The most widespread gradient estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
+We refer the reader to App.~\ref{A:gradient_estimators} for a formal definition.
+
+The main limitation of plain policy gradient has been the high variance of these estimators.
+The naive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples by interacting with the environment.
+For this reason, literature has focused on the introduction of baselines aiming to reduce the variance~\citep[\eg][]{Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.\footnote{A baseline is a function $f(s,a)$ such that $\EVV[a \sim \pol]{f(s,a)} = 0$, for any $s\in\Sspace$.}
+There has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
+Altough, these techniques has been mainly derived for finite-sum problems, we will show in Sec.~\ref{sec:algo} how can be used in RL.
+The next section has the aim to describe variance reduction techniques for finite-sum problems. In particular, we will present SVRG algorithm that is at the core of this work.
 
 \subsection{Stochastic Variance-Reduced Gradient}
-Finite sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $g_i(\vtheta)$:
-\begin{align*}
-\max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g_i(\vtheta).
-\end{align*}
-This kind of optimization is very common in machine learning, where each $g_i$ may correspond to a data sample $x_i$ from a dataset $\mathcal{D}_N$ of size $N$.
-In this case, we adopt the following, more meaningful notation:
-\begin{align*}
-\max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g(x_i \mid \vtheta).
-\end{align*}
-A common requirement is that $g$ must be smooth and convex in $\vtheta$.
-Under this hypothesis, full gradient ascent \cite{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.
-\todopir{Explain IFO meaning (footnote) and this holds for deterministic algorithms}
+Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $g_i(\cdot|\vtheta)$:
+\begin{align*}
+        \max_{\vtheta} \left\{ f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g_i(\vtheta)\right\}.
+\end{align*}
+This kind of optimization is very common in machine learning, where each $g_i$ may correspond to a data sample $x_i$ from a dataset $\mathcal{D}_N$ of size $N$ (\ie $g_i(\vtheta) = g(x_i|\vtheta)$). 
+%% In this case, we adopt the following, more meaningful notation:
+%% \begin{align*}
+%% \max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g(x_i \mid \vtheta).
+%% \end{align*}
+A common requirement is that $g$ must be smooth and {\color{red} concave} in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} Under this hypothesis, full gradient ascent \cite{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.\todopirout{Is it $\ell_2$-norm?}
 This can be too expensive for large values of $N$.
@@
 The common idea is to reuse past gradient computations to reduce the variance of the current estimate.
-In particular, Stochastic Variance Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which can become a problem when deep and/or wide neural networks are employed.
+In particular, Stochastic Variance-Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which can become a problem when deep and/or wide neural networks are employed.
 
@@
 
+\section{Policy Gradient Estimators} \label{A:gradient_estimators}
+
+The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
+\begin{align*}
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
+\end{align*}
+where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
+\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,wu2018variance}.}
+The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
+\begin{align*}
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\gamma^h r_h^n - b\right).
+\end{align*}
+
 \section{Proofs}\label{app:proofs}

### 2018-02-01 svrpg.tex

@@
 \begin{align*}
-\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\tilde{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} + \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i)\score{\tilde{\vtheta}}{\tau_i}R(\tau_i)} \\
+\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\tilde{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} - \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i)\score{\tilde{\vtheta}}{\tau_i}R(\tau_i)} \\
 &= \gradJ{\tilde{\vtheta}} + \gradJ{\vtheta} - \gradJ{\tilde{\vtheta}} = \gradJ{\vtheta}.

### 2018-02-02 svrpg.tex
updated
correzioni su una nota in sezione 5

@@
 
-We introduce two stopping conditions: the maximum number of sub-iterations and the condition on learning rate of Adam sub-iterations.
+In order to cope with this behaviour we introduce an adaptive stopping condition: we permorm a new snapshot anytime the learning rate proposed by the adam optimizer in a sub-iteration is 10 times smaller than that one proposed at the immediately previous snapshot.
+This stopping condition is subject to a constraint on the maximum number of sub-iterations performable, which is problem dependent.
 \todomat{I find "two stopping conditions" confusing: it's just one with an additional constraint}
+\todobep{I tried to solve the issue here:hope it's better now}
 We stop with inner iterations if:

### 2018-02-02 svrpg.tex
corrections

@@
 
-In order to cope with this behaviour we introduce an adaptive stopping condition: we permorm a new snapshot anytime the learning rate proposed by the adam optimizer in a sub-iteration is 10 times smaller than that one proposed at the immediately previous snapshot.
+In order to cope with this behaviour we introduce an adaptive stopping condition: we permorm a new snapshot anytime the weighted learning rate proposed by the adam optimizer in a sub-iteration is smaller than the weighted one proposed at the immediately previous snapshot.
 This stopping condition is subject to a constraint on the maximum number of sub-iterations performable, which is problem dependent.

### 2018-02-02 svrpg.tex
corrections

@@
 \todobep{I tried to solve the issue here:hope it's better now}
-We stop with inner iterations if:
+More precisely we stop performing sub-iterations when:
 
@@
 
+These two weights ${\bf{\it N}}$ and ${\bf{\it B}}$ represent, respectively, the number of trajectories used for a snapshot and the number of trajectories used for an update of the parameters in a sub-iteration.
+Therefore, in this stopping condition, we use them to take into account the trajectories efficiency.
+The less trajectories we use to perform an update the better it is from a data usage point of view.
+The variance control offered by this adaptive stopping condition lies within the value of the two different learning rate proposed by adam.
+Whenever the learning rate of a sub-iteration becomes too small with respect to that one proposed at the previous snapshot we are experiencing high variance.
+Hence it could be better to perform a new snapshot so to mitigate this issue.
+
 \todomat{why? should mention how this formula combines trajectory efficiency and variance control}
-After that we sample N trajectories for the next Snapshot point.
+\todobep{also here I tried to solve the issues highlighted}
 

### 2018-02-02 svrpg.tex
correzioni

@@
 \todomat{$\uparrow$ the reason is this, but try to spend more words on it}
-Denote by $\alpha_{snap}$ the last learning rate computed into the snapshot point and denote by  $\alpha_{sub-iter}$ the last learning rate computed into the sub-iteration point.
+Denote by $\alpha_{snap}$ the learning rate associated to the last snapshot and denote by  $\alpha_{sub-iter}$ the learning rate associated to the current sub-iteration.
 

### 2018-02-02 svrpg.tex
correzioni

@@
 Adam helps to stabilize the gradient estimate by computing adaptive learning rates for each parameter based on the history of the gradients' variance.
-We use two different adam hyperparameters: one for updates tied to snapshot points and the other tied to the updates of sub-iterations points.
+Due to this feature we can use two different adam in order to properly leverage the structure of the proposed SVRPG update.
+The first adam will be associated to the snapshot updates and the second one will be associated to the sub-iteration updates.
+By doing so we are decoupling the contribution to the variance introduced by the snapshots from that one introduced by the sub-iterations, which lie, respectively, on two different order of magnitude.Therefore, we are able to properly adapt the learning rate in each situation.
 \todomat{$\uparrow$ explain better: it's not only two separate meta-parameters, but two separate learning rate schedules.

### 2018-02-02 svrpg.tex
correzioni

@@
 The first adam will be associated to the snapshot updates and the second one will be associated to the sub-iteration updates.
-By doing so we are decoupling the contribution to the variance introduced by the snapshots from that one introduced by the sub-iterations, which lie, respectively, on two different order of magnitude.Therefore, we are able to properly adapt the learning rate in each situation.
-\todomat{$\uparrow$ explain better: it's not only two separate meta-parameters, but two separate learning rate schedules.
-Use notation from previous sections, \eg snaphot is usually associated with $\tilde{\vtheta}$}
-This is because we are dealing with two different variance magnitudes.
-\todomat{$\uparrow$ the reason is this, but try to spend more words on it}
+By doing so we are decoupling the contribution to the variance introduced by the snapshots from that one introduced by the sub-iterations.
+These two variance contributions lie, respectively, on two different order of magnitude because the number of trajectories used to perform the estimates for the snapshot is greater than that used for a sub-iteration.Therefore, using two adam optimizers allows us to properly adapt the learning rate in each situation.
 Denote by $\alpha_{snap}$ the learning rate associated to the last snapshot and denote by  $\alpha_{sub-iter}$ the learning rate associated to the current sub-iteration.
@@
 This stopping condition is subject to a constraint on the maximum number of sub-iterations performable, which is problem dependent.
-\todomat{I find "two stopping conditions" confusing: it's just one with an additional constraint}
-\todobep{I tried to solve the issue here:hope it's better now}
 More precisely we stop performing sub-iterations when:
@@
 Hence it could be better to perform a new snapshot so to mitigate this issue.
-
-\todomat{why? should mention how this formula combines trajectory efficiency and variance control}
-\todobep{also here I tried to solve the issues highlighted}
 

### 2018-02-02 svrpg.tex
Correction Pratical Variant parts
correzioni

@@
 \section{Hyperparameters}\label{sec:stopping}
-In Reinforcement learning we have to take into account the gradient's variance.
-\todomat{$\uparrow$ the opening of the section could connect better with previous sections. Declare which meta-parameters will be addressed in the section}
+In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
+The first one is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
+The second one is essetial to control the variance introduced by the importance weights, mainly because the farther you go from the snapshot policy the higher the variance of the importance weights will be.
+
 A common way to reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter.
@@
 These two variance contributions lie, respectively, on two different order of magnitude because the number of trajectories used to perform the estimates for the snapshot is greater than that used for a sub-iteration.Therefore, using two adam optimizers allows us to properly adapt the learning rate in each situation.
-Denote by $\alpha_{snap}$ the learning rate associated to the last snapshot and denote by  $\alpha_{sub-iter}$ the learning rate associated to the current sub-iteration.
-
-As shown in Section \ref{sec:conv}, the performance of the SVRPG algorithm is sensitive to the epoch size $m$.
-As the number of sub-iterations increases, the variance of the importance weights and the gradient increases too.
+For the above mentioned reasons let us denote by $\alpha_{snap}$ the learning rate associated to the last snapshot and by  $\alpha_{sub-iter}$ the learning rate associated to the current sub-iteration.
+
+We have already said that as the number of sub-iterations $m$ increases the variance of the importance weights increases too.
 Therefore, the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
-A fixed epoch size is not the optimal choice.
-Few sub-iterations lead to few updates, many update lead to increases of variance.
-
-In order to cope with this behaviour we introduce an adaptive stopping condition: we permorm a new snapshot anytime the weighted learning rate proposed by the adam optimizer in a sub-iteration is smaller than the weighted one proposed at the immediately previous snapshot.
+
+We have seen, sperimentally, that a fixed epoch size is not the optimal choice, but an adaptive one would better exploit the total amount of data at hand.
+Hence, we permorm a new snapshot anytime the weighted learning rate proposed by the adam optimizer in a sub-iteration is smaller than the weighted one proposed at the immediately previous snapshot.
 This stopping condition is subject to a constraint on the maximum number of sub-iterations performable, which is problem dependent.

### 2018-02-02 svrpg.tex
correzioni

@@
 We have already said that as the number of sub-iterations $m$ increases the variance of the importance weights increases too.
-Therefore, the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
+Therefore, if $m$ is not properly set the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
 
@@
 \section{Practical Variants}
+
+We introduce a term with importance weights to guarantee proper sampling of trajectories.
+This term's variance is subjected to the variance of importance weights. 
+There are some methods to reduce this variance.
 

### 2018-02-02 svrpg.tex
correzioni

@@
 Therefore, in this stopping condition, we use them to take into account the trajectories efficiency.
-The less trajectories we use to perform an update the better it is from a data usage point of view.
-The variance control offered by this adaptive stopping condition lies within the value of the two different learning rate proposed by adam.
+The less trajectories we use to perform an update the better it is from a data usage point of view, hence we want to facilitate sub-iterations with respect to snapshots.
+
+Another insight about this adaptive stopping condition lies within the value of the two different learning rate proposed by adam.
 Whenever the learning rate of a sub-iteration becomes too small with respect to that one proposed at the previous snapshot we are experiencing high variance.

### 2018-02-02 svrpg.tex
correzioni

@@
 Therefore, in this stopping condition, we use them to take into account the trajectories efficiency.
-The less trajectories we use to perform an update the better it is from a data usage point of view, hence we want to facilitate sub-iterations with respect to snapshots.
+The less trajectories we use to perform an update, the better it is from a data usage point of view.
+Hence we want to facilitate sub-iterations with respect to snapshots.
 

### 2018-02-02 svrpg.tex
completed introduction

--- 2018-02-02 svrpg.tex
@@
 \newcommand{\stationary}{d_{\rho}^{\pi_{\vtheta}}(s)}
-\newcommand{\policy}{\pi_{\vtheta}(a \mid s)}
+\newcommand{\policy}{\pi_{\vtheta}(a \vert s)}
 \newcommand{\pol}{\pi_{\vtheta}}
@@
 %Specific
-\newcommand{\Ets}[2][t]{\mathbb{E}_{#1\mid s}\left[#2\right]}
-\newcommand{\Covts}[3][t]{{\mathbb{C}\text{ov}}_{#1\mid s}\left(#2,#3\right)}
-\newcommand{\Varts}[2][t]{{\mathbb{V}\text{ar}}_{#1\mid s}\left[#2\right]}
+\newcommand{\Ets}[2][t]{\mathbb{E}_{#1\vert s}\left[#2\right]}
+\newcommand{\Covts}[3][t]{{\mathbb{C}\text{ov}}_{#1\vert s}\left(#2,#3\right)}
+\newcommand{\Varts}[2][t]{{\mathbb{V}\text{ar}}_{#1\vert s}\left[#2\right]}
 \newcommand{\gradBlack}[1]{\blacktriangledown J(#1)}
@@
 %Similarities between RL and SL
-On a very general level, machine learning addresses the problem of an agent that must select the right actions to solve a given task.
+On a very general level, machine learning addresses the problem of an agent that must select the right actions to solve a task.
 The approach of Reinforcement Learning~\citep{sutton1998reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal.
-This makes RL fundamentally different from Supervised Learning (SL), even in its online version, where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels).
-However, the two approaches share many challenges and, as a consequence, solution concepts.
+This makes RL fundamentally different from Supervised Learning (SL), where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels).
+However, the two approaches share many challenges and tools.
 The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or a policy of behaviour directly.
 Moreover, when the tasks are characterized by large or continuous state-action spaces, RL needs to borrow the powerful function approximators that are the main subject of study of SL, such as neural networks.
-
 % SL needs SVRG so RL
-In a typical discriminative supervised learning setting, an objective function $J(x\mid\vtheta)$ has to be optimized \wrt to a parameter vector $\vtheta$.
-The training set $\mathcal{D}_{train}$ is the set of data samples $x$ for which the objective (and its gradient) can be evaluated.
-In most practical cases, $\mathcal{D}_{train}$ is only a subset of the set $\mathcal{D}$ of all the $x$ of interest, hence we resort to stochastic optimization of the expected value of $J$ over an unknown data distribution.
-The special case in which $\mathcal{D}_{train} = \mathcal{D}$ is known as Empirical Risk Minimization (ERM), and allows for deterministic optimization.
+In a typical SL setting, a performance function $J(\vtheta)$ has to be optimized \wrt to model parameters $\vtheta$.
+The set of data that are available for training is often a subset of all the data of interest, which may even be infinite, leading to optimization of finite sums to approximate expected performance over an unknown data distribution.
+When generalization to the complete dataset is not taken into consideration we talk about Empirical Risk Minization (ERM).
 Even in this case, stochastic optimization is often used for reasons of efficiency.
-The idea is to evaluate the objective on a random subset of the available data to obtain an approximate gradient direction.
-This can be much less expensive than taking into account all the data for all parameter updates.
-However, the sub-sampling of data samples is a source of variance that can potentially compromise convergence, so that speed and variance must be traded off with proper handling of meta-parameters.
-Variance-reduced gradient algorithms such as SAG, SVRG and SAGA offer better ways of achieving the optimal trade-off between variance and speed, both in theory and practice.
-Although designed explicitly for ERM, this algorithms address a problem that is equally important, if not more critical, in more general machine learning problems.
-In Reinforcement Learning, as in online learning n general, stochastic optimization is rarely a matter of choice, since data must be actively sampled by interacting with an initially unknown environment.
-In this scenario, containing the variance of the estimates is a problem that cannot be avoided, and methods like SVRG can offer a more efficient alternative, or a complement, to traditional methods such as batch learning.
-
-\todomat{work in progress}
+The idea of stochastic gradient descent \cite{nesterov2013introductory} is to iteratively focus on a random subset of the available data to obtain an approximate improvement direction.
+At the level of the single iteration, this can be much less expensive than taking into account all the data.
+However, the sub-sampling of data samples is a source of variance that can potentially compromise convergence, so that per-iteration efficiency and convergence rate must be traded off with proper handling of meta-parameters.
+Variance-reduced gradient algorithms such as SAG \cite{roux2012stochastic}, SVRG \cite{johnson2013accelerating} and SAGA \cite{defazio2014saga} offer better ways of addressing this trade-off, with significant results both in theory and practice.
+Although designed explicitly for ERM, this algorithms address a problem that affects more general machine learning problems.
+In RL, stochastic optimization is rarely a matter of choice, since data must be actively sampled by interacting with an initially unknown environment.
+In this scenario, containing the variance of the estimates is a problem that cannot be avoided, which makes variance-reduced algorithms very interesting.
 % Potential advantages of SVRG
+Among RL approaches, policy gradient is the one that bears the closest similarity to SL solutions.
+The fundamental principle of this methods is to optimize a parametric policy by means of stochastic gradient descent.
+Compared to other applications of SGD, the cost of collecting samples can be very high since it requires to interact with the environment.
+This makes SVRG-like methods potentially much more efficient than, \eg batch learning.
 % Difficulties in RL \wrt SL
-
-% Among RL solutions for learning continuous tasks, policy gradient methods bear the closest similarity to SL, since they allow to formulate the RL problem as the unconstrained optimization of an objective function (\ie a negative loss) of policy parameters and data samples.
-
-\section{Preliminaries}
+Unfortunately, RL has a series of difficulties that are not present in ERM.
+First, in SL the objective can often be designed to be strongly convex.
+This is not possible in RL, so we have to deal with non-convexities.
+Then, as mentioned, the dataset is not initially available and may even be infinite, which  makes approximations unavoidable.
+This also rules out SAG and SAGA because of their storage requirements, which leaves SVRG as the most promising choice.
+Finally, the distribution used to sample data is not under direct control of the algorithm designer, but is a function of policy parameters that changes over time as the policy is optimized, which is a form of nonstationarity.
+
+% Related work
+
+%Paper structure
+After providing a background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} we address the aforementioned difficulties to design SVRPG, a variant of SVRG for the policy gradient framework.
+In Section \ref{sec:conv} we provide convergence guarantees for our algorithm which also offers insights on the challenges of variance reduction in RL.
+In Section \ref{sec:stopping} we suggests how to set the hyperparameters of SVRPG, while in Section \ref{sec:prac} we discuss some practical variants of the algorithm.
+Finally, in Section \ref{sec:exp} we empirically evaluate the performance of our method on popular continuous RL benchmark tasks.
+
+\section{Preliminaries}\label{sec:pre}
 In this section we provide the essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
@@
 %% \begin{align*}
-%% \max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g(x_i \mid \vtheta).
+%% \max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g(x_i \vert \vtheta).
 %% \end{align*}
@@
 			&\nabla f(\tilde{\vtheta}^s) \\ + 
-			&\nabla g(x_i\mid\vtheta_t^{s+1}) -
-			\nabla g(x_i \mid \tilde{\vtheta}^{s+1})
+			&\nabla g(x_i\vert\vtheta_t^{s+1}) -
+			\nabla g(x_i \vert \tilde{\vtheta}^{s+1})
 		\end{align*}
@@
 
-\section{Algorithm}\label{sec:algo}
+\section{Algorithm}\label{sec:alg}
 In on-line policy gradient problems, stochastic gradient ascent (SGA) is not really a choice, but is dictated by the necessity of interacting with an unknown environment.
@@
 \begin{align*}
-	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta})\Reward(\tau_j) \\
-		&+ \nabla\log\pi(\tau_i \mid \vtheta)\Reward(\tau_i) 
-		- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta})\Reward(\tau_i),
-\end{align*}
-where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from a snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\mid\tilde{\vtheta})}{\pi(\tau\mid\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$.
+	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
+		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
+		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i),
+\end{align*}
+where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from a snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\vert\tilde{\vtheta})}{\pi(\tau\vert\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$.
 Our update is still fundamentally on-policy, since the weighting concerns only the correction term.
@@
 \begin{align*}
-\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta})\Reward(\tau_j) \\
+\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
 &+ \frac{1}{B}\left[\sum_{i=0}^{B-1}
-\nabla\log\pi(\tau_i \mid \vtheta)\Reward(\tau_i) 
-- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta})\Reward(\tau_i)\right].
+\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
+- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i)\right].
 \end{align*}
@@
 		\FOR{$s=0$ {\bfseries to} $S-1$}
-		\STATE Sample N trajectories $\tau_j$ from $\pi(\cdot\mid\tilde{\vtheta}^{s})$
-		\STATE $\gradApp{\tilde{\vtheta}^{s}}{N} = \frac{1}{N}\sum_{j=0}^{N-1}\score{}{\tau_j\mid\tilde{\vtheta}^{s}}\Reward(\tau_j)$
+		\STATE Sample N trajectories $\tau_j$ from $\pi(\cdot\vert\tilde{\vtheta}^{s})$
+		\STATE $\gradApp{\tilde{\vtheta}^{s}}{N} = \frac{1}{N}\sum_{j=0}^{N-1}\score{}{\tau_j\vert\tilde{\vtheta}^{s}}\Reward(\tau_j)$
 		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m-1}^s$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
-		\STATE Sample $B$ trajectories $\tau_i$ from 				$\pi(\cdot\mid\vtheta_t^{s+1})$
+		\STATE Sample $B$ trajectories $\tau_i$ from 				$\pi(\cdot\vert\vtheta_t^{s+1})$
 		\STATE 
@@
 		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
-		\score{}{\tau_i\mid\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
-		&\left. - \omega(\tau_i)\score{}{\tau_i \mid \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
+		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
+		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
 		\end{align*}
@@
 \section{Hyperparameters}\label{sec:stopping}
-In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
-The first one is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
-The second one is essetial to control the variance introduced by the importance weights, mainly because the farther you go from the snapshot policy the higher the variance of the importance weights will be.
-
+In Reinforcement learning we have to take into account the gradient's variance.
+\todomat{$\uparrow$ the opening of the section could connect better with previous sections. Declare which meta-parameters will be addressed in the section}
 A common way to reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter.
 Adam helps to stabilize the gradient estimate by computing adaptive learning rates for each parameter based on the history of the gradients' variance.
-Due to this feature we can use two different adam in order to properly leverage the structure of the proposed SVRPG update.
-The first adam will be associated to the snapshot updates and the second one will be associated to the sub-iteration updates.
-By doing so we are decoupling the contribution to the variance introduced by the snapshots from that one introduced by the sub-iterations.
-These two variance contributions lie, respectively, on two different order of magnitude because the number of trajectories used to perform the estimates for the snapshot is greater than that used for a sub-iteration.Therefore, using two adam optimizers allows us to properly adapt the learning rate in each situation.
-For the above mentioned reasons let us denote by $\alpha_{snap}$ the learning rate associated to the last snapshot and by  $\alpha_{sub-iter}$ the learning rate associated to the current sub-iteration.
-
-We have already said that as the number of sub-iterations $m$ increases the variance of the importance weights increases too.
-Therefore, if $m$ is not properly set the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
-
-We have seen, sperimentally, that a fixed epoch size is not the optimal choice, but an adaptive one would better exploit the total amount of data at hand.
-Hence, we permorm a new snapshot anytime the weighted learning rate proposed by the adam optimizer in a sub-iteration is smaller than the weighted one proposed at the immediately previous snapshot.
+We use two different adam hyperparameters: one for updates tied to snapshot points and the other tied to the updates of sub-iterations points.
+\todomat{$\uparrow$ explain better: it's not only two separate meta-parameters, but two separate learning rate schedules.
+Use notation from previous sections, \eg snaphot is usually associated with $\tilde{\vtheta}$}
+This is because we are dealing with two different variance magnitudes.
+\todomat{$\uparrow$ the reason is this, but try to spend more words on it}
+Denote by $\alpha_{snap}$ the last learning rate computed into the snapshot point and denote by  $\alpha_{sub-iter}$ the last learning rate computed into the sub-iteration point.
+
+As shown in Section \ref{sec:conv}, the performance of the SVRPG algorithm is sensitive to the epoch size $m$.
+As the number of sub-iterations increases, the variance of the importance weights and the gradient increases too.
+Therefore, the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
+A fixed epoch size is not the optimal choice.
+Few sub-iterations lead to few updates, many update lead to increases of variance.
+
+In order to cope with this behaviour we introduce an adaptive stopping condition: we perform a new snapshot anytime the weighted learning rate proposed by the adam optimizer in a sub-iteration is smaller than the weighted one proposed at the immediately previous snapshot.
 This stopping condition is subject to a constraint on the maximum number of sub-iterations performable, which is problem dependent.
-More precisely we stop performing sub-iterations when:
+\todomat{I find "two stopping conditions" confusing: it's just one with an additional constraint}
+\todobep{I tried to solve the issue here:hope it's better now}
+We stop with inner iterations if:
 
@@
 
-These two weights ${\bf{\it N}}$ and ${\bf{\it B}}$ represent, respectively, the number of trajectories used for a snapshot and the number of trajectories used for an update of the parameters in a sub-iteration.
-Therefore, in this stopping condition, we use them to take into account the trajectories efficiency.
-The less trajectories we use to perform an update the better it is from a data usage point of view.
-The variance control offered by this adaptive stopping condition lies within the value of the two different learning rate proposed by adam.
-Whenever the learning rate of a sub-iteration becomes too small with respect to that one proposed at the previous snapshot we are experiencing high variance.
-Hence it could be better to perform a new snapshot so to mitigate this issue.
-
-\section{Practical Variants}
-
-We introduce a term with importance weights to guarantee proper sampling of trajectories.
-This term's variance is subjected to the variance of importance weights. 
-There are some methods to reduce this variance.
-
-As reported in Section \ref{subsec:PolicyGradient} we use G(PO)MDP gradient estimator~\cite{baxter2001infinite}, which gives us the same guarantees we have for the Reinforce Estimator.
+\todomat{why? should mention how this formula combines trajectory efficiency and variance control}
+After that we sample N trajectories for the next Snapshot point.
+
+\section{Practical Variants}\label{sec:prac}
+\todomat{Missing opening to section}
+As reported in Section \ref{subsec:PolicyGradient} we use the G(PO)MDP gradient estimator~\cite{baxter2001infinite}, which gives us the same guarantees we have for the Reinforce Estimator.
 One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}, allowing a reduction of the gradient estimator variance.
@@
 \begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\mid s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
+\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\vert s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
 \end{align*} 
@@
 \begin{align*}
-\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\mid s_k^n})
- }{\pi_{\vtheta^{'}}({a_k^n\mid s_k^n})}
-\end{align*}
-\begin{align*}
-\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\mid s_k^n})
-}{\pi_{\vtheta^{'}}({a_k^n\mid s_k^n})}
-\end{align*}
-
+\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
+ }{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
+\end{align*}
+\begin{align*}
+\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
+}{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
+\end{align*}
+\todomat{why should we use it? What advantages do we expect? At what cost? Pointer to experiments}
 \section{Related Work}
@@
 \begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
 \end{align*}
@@
 \begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\gamma^h r_h^n - b\right).
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\gamma^h r_h^n - b\right).
 \end{align*}
@@
 \begin{align*}
-	&\mathbb{E}_{t\mid s}\left[X\right] \coloneqq 
-		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}^s)\forall j \in {\it {\bf N}} \\ \tau_{i,h}\sim\pi(\cdot\mid\vtheta^{s+1}_i) \forall h \in {\it {\bf B}}, \text{ for $i=0,\dots,t-1$}}}{\left[X \mid \tilde{\vtheta^s}\right]} \\
-	&\coloneqq \EVV[\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}^s)\forall j \in {\it {\bf N}}]{
-			\EVV[\tau_{0,h}\sim\pi(\cdot\mid\vtheta_0^{s+1}) \forall h \in {\it {\bf B}}]
+	&\mathbb{E}_{t\vert s}\left[X\right] \coloneqq 
+		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\vert\tilde{\vtheta}^s)\forall j \in {\it {\bf N}} \\ \tau_{i,h}\sim\pi(\cdot\vert\vtheta^{s+1}_i) \forall h \in {\it {\bf B}}, \text{ for $i=0,\dots,t-1$}}}{\left[X \vert \tilde{\vtheta^s}\right]} \\
+	&\coloneqq \EVV[\tau_j\sim\pi(\cdot\vert\tilde{\vtheta}^s)\forall j \in {\it {\bf N}}]{
+			\EVV[\tau_{0,h}\sim\pi(\cdot\vert\vtheta_0^{s+1}) \forall h \in {\it {\bf B}}]
 				{\dots
-					\EVV[\tau_{t,h}\sim\pi(\cdot\mid\vtheta_t^{s+1})\forall h \in {\it {\bf B}}]
-						{X\mid\vtheta_t^{s+1}}
+					\EVV[\tau_{t,h}\sim\pi(\cdot\vert\vtheta_t^{s+1})\forall h \in {\it {\bf B}}]
+						{X\vert\vtheta_t^{s+1}}
 				 \dots
-			\mid\vtheta_0^{s+1}}
-		\mid\tilde{\vtheta}^s},
+			\vert\vtheta_0^{s+1}}
+		\vert\tilde{\vtheta}^s},
 \end{align*}
@@
 	\gradJ{\tilde{\vtheta}^s}
-	+ \nabla\log\pi(\tau_i \mid \vtheta_t^{s+1})\Reward(\tau_i) 
-	- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta}^s)\Reward(\tau_i) \\
+	+ \nabla\log\pi(\tau_i \vert \vtheta_t^{s+1})\Reward(\tau_i) 
+	- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta}^s)\Reward(\tau_i) \\
 	&= \gradBlack{\vtheta_t^{s+1}} - \gradApp{\tilde{\vtheta}^s}{N} + \gradJ{\tilde{\vtheta}^s} \\
@@
 	%
-	&= \mathbb{E}_{t\mid s}\left[\left\|\gradApp{\tilde{\vtheta}^s}{N}
+	&= \mathbb{E}_{t\vert s}\left[\left\|\gradApp{\tilde{\vtheta}^s}{N}
 			+\frac{1}{B}\sum_{i=0}^{B-1}\left( 
@@
 	&\qquad+ 
-	\mathbb{E}_{t\mid s}\left[\left\|
+	\mathbb{E}_{t\vert s}\left[\left\|
 		\frac{1}{B}\sum_{i=0}^{B-1}\left(
@@
 	&\qquad+ 
-		\mathbb{E}_{t\mid s}\left[\left\|
+		\mathbb{E}_{t\vert s}\left[\left\|
 		\frac{1}{B}\sum_{i=0}^{B-1}\left(

### 2018-02-02 svrpg.tex

@@
 \section{Hyperparameters}\label{sec:stopping}
-In Reinforcement learning we have to take into account the gradient's variance.
-\todomat{$\uparrow$ the opening of the section could connect better with previous sections. Declare which meta-parameters will be addressed in the section}
+In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
+The first one is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
+The second one is essetial to control the variance introduced by the importance weights, mainly because the farther you go from the snapshot policy the higher the variance of the importance weights will be.
+
 A common way to reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter.
 Adam helps to stabilize the gradient estimate by computing adaptive learning rates for each parameter based on the history of the gradients' variance.
-We use two different adam hyperparameters: one for updates tied to snapshot points and the other tied to the updates of sub-iterations points.
-\todomat{$\uparrow$ explain better: it's not only two separate meta-parameters, but two separate learning rate schedules.
-Use notation from previous sections, \eg snaphot is usually associated with $\tilde{\vtheta}$}
-This is because we are dealing with two different variance magnitudes.
-\todomat{$\uparrow$ the reason is this, but try to spend more words on it}
-Denote by $\alpha_{snap}$ the last learning rate computed into the snapshot point and denote by  $\alpha_{sub-iter}$ the last learning rate computed into the sub-iteration point.
-
-As shown in Section \ref{sec:conv}, the performance of the SVRPG algorithm is sensitive to the epoch size $m$.
-As the number of sub-iterations increases, the variance of the importance weights and the gradient increases too.
-Therefore, the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
-A fixed epoch size is not the optimal choice.
-Few sub-iterations lead to few updates, many update lead to increases of variance.
-
-In order to cope with this behaviour we introduce an adaptive stopping condition: we perform a new snapshot anytime the weighted learning rate proposed by the adam optimizer in a sub-iteration is smaller than the weighted one proposed at the immediately previous snapshot.
+
+Due to this feature we can use two different adam in order to properly leverage the structure of the proposed SVRPG update.
+The first adam will be associated to the snapshot updates and the second one will be associated to the sub-iteration updates.
+By doing so we are decoupling the contribution to the variance introduced by the snapshots from that one introduced by the sub-iterations.
+These two variance contributions lie, respectively, on two different order of magnitude because the number of trajectories used to perform the estimates for the snapshot is greater than that used for a sub-iteration.Therefore, using two adam optimizers allows us to properly adapt the learning rate in each situation.
+For the above mentioned reasons let us denote by $\alpha_{snap}$ the learning rate associated to the last snapshot and by  $\alpha_{sub-iter}$ the learning rate associated to the current sub-iteration.
+
+We have already said that as the number of sub-iterations $m$ increases the variance of the importance weights increases too.
+Therefore, if $m$ is not properly set the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
+
+We have seen, sperimentally, that a fixed epoch size is not the optimal choice, but an adaptive one would better exploit the total amount of data at hand.
+Hence, we permorm a new snapshot anytime the weighted learning rate proposed by the adam optimizer in a sub-iteration is smaller than the weighted one proposed at the immediately previous snapshot.
 This stopping condition is subject to a constraint on the maximum number of sub-iterations performable, which is problem dependent.
-\todomat{I find "two stopping conditions" confusing: it's just one with an additional constraint}
-\todobep{I tried to solve the issue here:hope it's better now}
-We stop with inner iterations if:
+More precisely we stop performing sub-iterations when:
 
@@
 
-\todomat{why? should mention how this formula combines trajectory efficiency and variance control}
-After that we sample N trajectories for the next Snapshot point.
-
-\section{Practical Variants}\label{sec:prac}
+These two weights ${\bf{\it N}}$ and ${\bf{\it B}}$ represent, respectively, the number of trajectories used for a snapshot and the number of trajectories used for an update of the parameters in a sub-iteration.
+Therefore, in this stopping condition, we use them to take into account the trajectories efficiency.
+The less trajectories we use to perform an update, the better it is from a data usage point of view.
+Hence we want to facilitate sub-iterations with respect to snapshots.
+
+Another insight about this adaptive stopping condition lies within the value of the two different learning rate proposed by adam.
+Whenever the learning rate of a sub-iteration becomes too small with respect to that one proposed at the previous snapshot we are experiencing high variance.
+Hence it could be better to perform a new snapshot so to mitigate this issue.
+
+\section{Practical Variants}
 \todomat{Missing opening to section}
-As reported in Section \ref{subsec:PolicyGradient} we use the G(PO)MDP gradient estimator~\cite{baxter2001infinite}, which gives us the same guarantees we have for the Reinforce Estimator.
+We introduce a term with importance weights to guarantee proper sampling of trajectories.
+This term's variance is subjected to the variance of importance weights. 
+There are some methods to reduce this variance.
+
+As reported in Section \ref{subsec:PolicyGradient} we use G(PO)MDP gradient estimator~\cite{baxter2001infinite}, which gives us the same guarantees we have for the Reinforce Estimator.
 One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}, allowing a reduction of the gradient estimator variance.

### 2018-02-04 svrpg.tex
fixed section labels and svrpg gradient update formula
small fixes + added some references

@@
 In a typical SL setting, a performance function $J(\vtheta)$ has to be optimized \wrt to model parameters $\vtheta$.
-The set of data that are available for training is often a subset of all the data of interest, which may even be infinite, leading to optimization of finite sums to approximate expected performance over an unknown data distribution.
+The set of data that are available for training is often a subset of all the cases of interest, which may even be infinite, leading to optimization of finite sums to approximate expected performance over an unknown data distribution.
 When generalization to the complete dataset is not taken into consideration we talk about Empirical Risk Minization (ERM).
 Even in this case, stochastic optimization is often used for reasons of efficiency.
-The idea of stochastic gradient descent \cite{nesterov2013introductory} is to iteratively focus on a random subset of the available data to obtain an approximate improvement direction.
+The idea of stochastic gradient ascent \cite{nesterov2013introductory} is to iteratively focus on a random subset of the available data to obtain an approximate improvement direction.
 At the level of the single iteration, this can be much less expensive than taking into account all the data.
@@
 % Potential advantages of SVRG
-Among RL approaches, policy gradient is the one that bears the closest similarity to SL solutions.
-The fundamental principle of this methods is to optimize a parametric policy by means of stochastic gradient descent.
+Among RL approaches, policy gradient \cite{sutton2000policy} is the one that bears the closest similarity to SL solutions.
+The fundamental principle of this methods is to optimize a parametric policy by means of stochastic gradient ascent.
 Compared to other applications of SGD, the cost of collecting samples can be very high since it requires to interact with the environment.
@@
 Unfortunately, RL has a series of difficulties that are not present in ERM.
-First, in SL the objective can often be designed to be strongly convex.
-This is not possible in RL, so we have to deal with non-convexities.
+First, in SL the objective can often be designed to be strongly concave.
+This is not possible in RL, so we have to deal with non-concavities.
 Then, as mentioned, the dataset is not initially available and may even be infinite, which  makes approximations unavoidable.
@@
 
-% Related work
+% Related work?
 
@@
 There has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
-Altough, these techniques has been mainly derived for finite-sum problems, we will show in Sec.~\ref{sec:algo} how can be used in RL.
-The next section has the aim to describe variance reduction techniques for finite-sum problems. In particular, we will present SVRG algorithm that is at the core of this work.
+Altough these techniques has been mainly derived for finite-sum problems, we will show in Sec.~\ref{sec:alg} how they can be used in RL.
+The next section has the aim to describe variance reduction techniques for finite-sum problems. In particular, we will present the SVRG algorithm that is at the core of this work.
 
@@
 Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling.
-The resulting IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$.
-\todopir{What is the reference?}
+The resulting IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$ \cite{nemirovskii1983problem}\cite{reddi2016stochastic}.
 Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga} and Finito~\cite{defazio2014finito},
@@
 \begin{align*}
-\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
-&+ \frac{1}{B}\left[\sum_{i=0}^{B-1}
+&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) +\\
+&\quad\frac{1}{B}\left[\sum_{i=0}^{B-1}
 \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
@@
 In this section we adapt existing convergence guarantees for nonconvex SVRG \cite{reddi2016stochastic} \cite{allen2016variance} to the Policy Gradient framework.
-Each of the three challenges presented at the beginning of Section \ref{sec:algo} can potentially prevent convergence, so we need additional assumptions:
+Each of the three challenges presented at the beginning of Section \ref{sec:alg} can potentially prevent convergence, so we need additional assumptions:
 \begin{enumerate}
@@
 
-\section{Practical Variants}
+\section{Practical Variants}\label{sec:prac}
 \todomat{Missing opening to section}
@@
 \end{align*}
-where the sequence $\tilde{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}, ${\it {\bf N}}$ is the amount of trajectories sampled for the snapshot and ${\it {\bf B}}$ is the amount of trajectories sampled for a sub iteration.
+where the sequence $\tilde{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}, ${\it {\bf N}} = [0,\dots,N)$, and ${\it {\bf B}} = [0,\dots,B)$. 
 \end{definition}

### 2018-02-04 svrpg.tex
added appendix on applicability to Gaussian policies

@@
 	as already mentioned, a common assumption for nonconvex optimization is the L-smoothness of the objective. We actually require the following, which is stronger:
-	\begin{assumption}\label{ass:bounded_score}
+	\begin{restatable}[On policy derivatives]{assumption}{boundedscore}\label{ass:bounded_score}
 		For each trajectory $\tau$, value of $\vtheta$ and dimensions $i,j$ there are positive constants $G,F<\infty$ such that:
@@
 		\end{align*}
-	\end{assumption}
-	This property is satisfied by the commonly used truncated Gaussian policy, and is a sufficient condition for the L-smoothness of the objective (see Lemma \ref{lemma:lsmooth} in the appendix).
+	\end{restatable}
+	This property is a sufficient condition for the L-smoothness of the objective (see Lemma \ref{lemma:lsmooth} in the appendix).
 \item \textit{Approximation}:
 since we can only compute an estimate of the full gradient, we require its variance to be bounded:
-	\begin{assumption}\label{ass:REINFORCE}
+	\begin{restatable}[On the variance of the gradient estimator]{assumption}{varreinforce}\label{ass:REINFORCE}
 		There is a constant $V<\infty$ such that, for each policy $\pol$:
@@
 		\]
-	\end{assumption}
-This assumption is satisfied by the REINFORCE and G(PO)MDP estimators and Gaussian policies \cite{zhao2011analysis} \cite{pirotta2013adaptive}.
+	\end{restatable}
 The problem of full gradient approximation for SVRG has been addressed in the supervised learning framework \cite{harikandeh2015stopwasting} with results similar to ours.
@@
 the variance introduced by importance sampling must be bounded:
-	\begin{assumption}\label{ass:M2}
+	\begin{restatable}[On the variance of importance weights]{assumption}{varweights}\label{ass:M2}
 		There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
@@
 		\]
-	\end{assumption}
-Differently from the previous two, this assumption must be enforced by a proper handling of the epoch size $m$.
+	\end{restatable}
+Differently from the previous two, this assumption can be enforced by a proper handling of the epoch size $m$.
 \end{enumerate}
 
+In Appendix \ref{app:gauss} we show how the commonly used Gaussian policy satisfies these assumptions.
 We can now state the following key result:
@@
 \end{align*}
-where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
-\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,wu2018variance}.}
-The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
+where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline provided $\EVV[a \sim \pol]{b(s,a)} = 0$, for any $s\in\Sspace$.
+The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance \cite{zhao2011analysis} while preserving the unbiasedness:
 \begin{align*}
@@
 
-\section{Applicability to Gaussian Policies}\label{app:pol}
-\todomat{...}
+\section{Applicability to Gaussian Policies}\label{app:gauss}
+We provide more details on the applicability of Theorem \ref{theo:convergence} on the case of Gaussian policies.
+We start from the case of one-dimensional bounded action space $\mathcal{A}\subset\mathbb{R}$, linear mean $\mu(s) = \vtheta^T\vphi(s)$ and fixed standard deviation $\sigma$:
+\[
+	\pi_{\vtheta}(a\vert s) = \frac{1}{\sqrt{2\pi}\sigma}\exp\left\{
+		-\frac{(\vtheta^T\vphi(s) - a)^2}{2\sigma^2}\right\},
+\]
+where $\vphi(s)\leq M_{\phi}$ is a bounded feature vector, and we see under which conditions the three assumptions of Section \ref{sec:conv} hold.
+
+\boundedscore*
+For the Gaussian policy defined above, it's easy to show that:
+\begin{align*}
+	&\nabla_{\theta_i}\log\pi_{\vtheta}(\tau) =  \phi_i(s)\frac{a-\vtheta^T\phi(s)}{\sigma^2},\\
+	&\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(\tau) = \frac{\phi_i(s)\phi_j(s)}{\sigma^2}.
+\end{align*}
+Hence, Assumption \ref{ass:bounded_score} is automatically satisfied \footnote{This relies on the fact that $\vtheta^T\phi(s)$ lies in bounded $\Aspace$.
+In practice, this is usually enforced by clipping the action selected by $\pi_{\vtheta}$.
+A more rigorous way would be to employ the truncated Gaussian distribution.} by taking $\GRADLOG = \frac{M_{\phi}|\Aspace|}{\sigma^2}$ and $\HESSLOG = \frac{M_{\phi}^2}{\sigma^2}$.
+\par
+\varreinforce*
+As mentioned, \cite{pirotta2013adaptive} provides a bound on the variance of the REINFORCE estimator, adapted from \cite{zhao2011analysis}, which does not depend on $\vtheta$:
+\begin{align*}
+\Var\left[\gradApp{\theta_i}{N}\right] \leq \frac{R^2M_{\phi}^2H(1-\gamma^H)^2}{N\sigma^2(1-\gamma)^2}.
+\end{align*}
+The same authors provide a similar bound for G(PO)MDP.
+
+\varweights*
+It is noted in \cite{cortes2010learning} that, for any two Gaussian distributions $\mathcal{N}(\mu_1,\sigma_1)$ and $\mathcal{N}(\mu_2,\sigma_2)$, the variance of the importance weights from the latter to the former is bounded whenever $\sigma_2 > \frac{\sqrt{2}}{2}\sigma_1$.
+This is automatically satisfied by our fixed-variance Gaussian policies, since $\sigma_2=\sigma_1=\sigma$.
+
+We now briefly examine some generalizations of the simple Gaussian policy defined above that can be found in applications:
+
+\paragraph{Multi-dimensional actions}
+When actions are multi-dimensional, factored Gaussian policies are typically employed, so the results extends trivially from the one-dimensional case.
+Actual multi-variate Gaussian distributions would require more calculations, but we do not expect substantially different results.
+
+\paragraph{Non-linear mean}
+In complex continuous tasks, $\mu(s)$ often represents a deep neural network, or multi-layer perceptron, where $\vtheta$ are the weights of the network.
+The analysis of first and second order log-derivatives in such a scenario is beyond the scope of this paper.
+
+\paragraph{Adaptive variance}
+It is a common practice to learn also the variance of the policy in order to adapt the degree of exploration.
+The variance (or diagonal covariance matrix in the multi-dimensional case) can be learned as a separate parameter or be state-dependent like the mean.
+In any case, adaptive variance must be carefully employed since it can clearly undermine all the three assumptions of Theorem \ref{theo:convergence}.
 

### 2018-02-04 svrpg.tex
added opening to proofs appendix

@@
 \section{Proofs}\label{app:proofs}
+In this Appendix we prove all the claims made in the paper, with the primary objective of proving Theorem \ref{theo:convergence}.
+Our proof is adapted from the one of Theorem 2 from \cite{reddi2016stochastic} and has a very similar structure, but with all the additional challenges and assumptions described in Section \ref{sec:conv}.
 

### 2018-02-04 svrpg.tex
Updated

--- 2018-02-04 svrpg.tex
@@
 In a typical SL setting, a performance function $J(\vtheta)$ has to be optimized \wrt to model parameters $\vtheta$.
-The set of data that are available for training is often a subset of all the cases of interest, which may even be infinite, leading to optimization of finite sums to approximate expected performance over an unknown data distribution.
+The set of data that are available for training is often a subset of all the data of interest, which may even be infinite, leading to optimization of finite sums to approximate expected performance over an unknown data distribution.
 When generalization to the complete dataset is not taken into consideration we talk about Empirical Risk Minization (ERM).
 Even in this case, stochastic optimization is often used for reasons of efficiency.
-The idea of stochastic gradient ascent \cite{nesterov2013introductory} is to iteratively focus on a random subset of the available data to obtain an approximate improvement direction.
+The idea of stochastic gradient descent \cite{nesterov2013introductory} is to iteratively focus on a random subset of the available data to obtain an approximate improvement direction.
 At the level of the single iteration, this can be much less expensive than taking into account all the data.
@@
 % Potential advantages of SVRG
-Among RL approaches, policy gradient \cite{sutton2000policy} is the one that bears the closest similarity to SL solutions.
-The fundamental principle of this methods is to optimize a parametric policy by means of stochastic gradient ascent.
+Among RL approaches, policy gradient is the one that bears the closest similarity to SL solutions.
+The fundamental principle of this methods is to optimize a parametric policy by means of stochastic gradient descent.
 Compared to other applications of SGD, the cost of collecting samples can be very high since it requires to interact with the environment.
@@
 Unfortunately, RL has a series of difficulties that are not present in ERM.
-First, in SL the objective can often be designed to be strongly concave.
-This is not possible in RL, so we have to deal with non-concavities.
+First, in SL the objective can often be designed to be strongly convex.
+This is not possible in RL, so we have to deal with non-convexities.
 Then, as mentioned, the dataset is not initially available and may even be infinite, which  makes approximations unavoidable.
@@
 \section{Preliminaries}\label{sec:pre}
-In this section we provide the essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
+We start providing the essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
 %In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
@@
 The main limitation of plain policy gradient has been the high variance of these estimators.
-The naive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples by interacting with the environment.
+The naive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples, \ie by interacting with the environment.
 For this reason, literature has focused on the introduction of baselines aiming to reduce the variance~\citep[\eg][]{Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.\footnote{A baseline is a function $f(s,a)$ such that $\EVV[a \sim \pol]{f(s,a)} = 0$, for any $s\in\Sspace$.}
 There has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
-Altough these techniques has been mainly derived for finite-sum problems, we will show in Sec.~\ref{sec:alg} how they can be used in RL.
-The next section has the aim to describe variance reduction techniques for finite-sum problems. In particular, we will present the SVRG algorithm that is at the core of this work.
+Altough, these techniques has been mainly derived for finite-sum problems, we will show in Sec.~\ref{sec:alg} how can be used in RL.
+The next section has the aim to describe variance-reduction techniques for finite-sum problems. In particular, we will present the SVRG algorithm that is at the core of this work.
 
@@
 %% \end{align*}
-A common requirement is that $g$ must be smooth and {\color{red} concave} in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} Under this hypothesis, full gradient ascent \cite{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.\todopirout{Is it $\ell_2$-norm?}
+A common requirement is that $g$ must be smooth and {\color{red} concave} in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} Under this hypothesis, full gradient ascent~\citep{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.\footnote{In the following we will omit the subscript for the $\ell_2$-norm.}
 This can be too expensive for large values of $N$.
 Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling.
-The resulting IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$ \cite{nemirovskii1983problem}\cite{reddi2016stochastic}.
+The resulting IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$.
+\todopir{What is the reference?}
 Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga} and Finito~\cite{defazio2014finito},
@@
 The idea of SVRG is to alternate full and stochastic gradient updates. 
-\todopir{here would be nice to have a notation for the single batch}
-Each $m = O(N)$ iterations, a snapshot $\tilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\tilde{\vtheta})$.
-Otherwise, the parameter is updated with a corrected stochastic gradient estimate of the form:
-\begin{align*}
-\blacktriangledown f(\vtheta) = 
+Each $m = O(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i g(x_i|\widetilde{\vtheta})$.
+Between snapshots, the parameter is updated with a gradient estimate corrected using stochastic gradient, for any $t \in \{0,\ldots,m-1\}$
+\begin{align}\label{E:svrg.gradient.correction}
+        \blacktriangledown f(\vtheta_{t+1}) = 
 	\nabla f(\tilde{\vtheta}) + 
-	\nabla g(x_i | \vtheta) -
+	\nabla g(x_i | \vtheta_t) -
 	\nabla g(x_i | \tilde{\vtheta}),
-\end{align*} 
-where $x_i$ is sampled uniformly at random from $\mathcal{D}_N$.
+\end{align} 
+where $x_i$ is sampled uniformly at random from $\mathcal{D}_N$ (\ie $x_i \sim \mathcal{U}(\mathcal{D}_N)$).
 The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient.
@@
 In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
-Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the convex case~\citep{nesterov2013introductory,ghadimi2013stochastic}.
+Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the \textcolor{red}{concave} case~\citep{nesterov2013introductory,ghadimi2013stochastic}.
 Also in this case, SVRG yields an advantageous trade-off, with an IFO complexity of $O(N + \nicefrac{N^{\nicefrac{2}{3}}}{N})$~\citep{reddi2016stochastic}.
@@
 Pseudocode of SVRG for both the convex and the non-convex case is provided in Alg.~\ref{alg:svrg}.
+\todopir{SVRG for non-convex optimization requires to mix the parameter $\widetilde{\vtheta}^{s+1}$?}
 
@@
 	\begin{algorithmic}
-		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m-1}^0$
+            \STATE {\bfseries Input:} a dataset $\mathcal{D}_N$, number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m}^0$
 		\FOR{$s=0$ {\bfseries to} $S-1$}
-		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m-1}^s$
+		\STATE $\vtheta_0^{s+1} := \widetilde{\vtheta}^s := \vtheta_{m}^s$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
-		\STATE Pick $i$ uniformly at random from $[1,N]$
-		\STATE \begin{align*}
-			\blacktriangledown f(\vtheta_t^{s+1}) = 
-			&\nabla f(\tilde{\vtheta}^s) \\ + 
-			&\nabla g(x_i\vert\vtheta_t^{s+1}) -
-			\nabla g(x_i \vert \tilde{\vtheta}^{s+1})
-		\end{align*}
-		\STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha\blacktriangledown f(\vtheta_t^{s+1})$
+%		\STATE Pick $i$ uniformly at random from $[1,N]$
+        \STATE $x_i \sim \mathcal{U}\left(\mathcal{D}_N\right)$
+		\STATE $v^{s+1} = 
+			\nabla f(\widetilde{\vtheta}^s) + 
+			\nabla g(x_i|\vtheta_t^{s+1}) -
+			\nabla g(x_i|\widetilde{\vtheta}^{s})
+		$
+        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}$
 		\ENDFOR
 		\ENDFOR
-		\STATE {\bfseries Convex case: return} $\vtheta_{m-1}^S$
+		\STATE {\bfseries Convex case: return} $\vtheta_{m}^S$
 		\STATE {\bfseries Non-Convex case: return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
@@
 \section{Algorithm}\label{sec:alg}
-In on-line policy gradient problems, stochastic gradient ascent (SGA) is not really a choice, but is dictated by the necessity of interacting with an unknown environment.
-We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
+In online RL problems the usual approach is to tune the batch size of SG in order to find the optimal trade-off between variance and speed.
+Recall that compared to SL, the samples are not fixed in advance but need to be collected at each policy change.
+Since this operation may be costly, we would like to minimize the number of interactions with the environment.
+Due to this issues, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
+However, a direct application of SVRG to RL is not possible due to the following properties:
+% In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence.
+I would say that SG is not a choice by referring to papers that uses batches}
+% We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
 In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$.
 Compared to the typical finite sum optimization scenario, we have the following additional challenges:
-\begin{enumerate}
-	\item \textit{Non-convexity}: the objective $J(\vtheta)$ is typically non-convex;
-	\item \textit{Approximation}: the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
-	\item \textit{Non-stationarity}: the value of the parameter $\vtheta$ influences the sampling of trajectories;
-\end{enumerate}
-To deal with non-convexity, we require $J(\vtheta)$ to be L-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and Softmax [appendix?].
+\begin{description}
+        \item[Non-convexity:] the objective function $J(\vtheta)$ is typically non-convex;
+        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The actual dataset is infinite and coincide with $\pi_{\vtheta}(\tau)$.
+%    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
+        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the parameter $\vtheta$ influences the sampling process.
+\end{description}
+To deal with non-convexity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax (see App.~\ref{app:lsmoothness}.\todopirout{search for ref}
 Because of the infinite dataset, we can only rely on an estimate of the full gradient. 
-Finally, we employ importance weighting \cite{rubinstein1981simulation} \cite{precup2000eligibility} to guarantee proper sampling of trajectories.
+\todopir{This should be argumented more via stop wasting my gradients}
+Finally, the unbiasedness of Eq.~\ref{E:svrg.gradients.correction} is broken by the non-stationarity of the optimization problem.
+To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
+%Finally, we employ importance weighting~\citep[\eg]{rubinstein1981simulation,precup2000eligibility} to guarantee proper sampling of trajectories. 
 We focus on the REINFORCE gradient estimator for simplicity, but the extension to G(PO)MDP is straightforward.
@@
 In this section we adapt existing convergence guarantees for nonconvex SVRG \cite{reddi2016stochastic} \cite{allen2016variance} to the Policy Gradient framework.
-Each of the three challenges presented at the beginning of Section \ref{sec:alg} can potentially prevent convergence, so we need additional assumptions:
+Each of the three challenges presented at the beginning of Section \ref{sec:algo} can potentially prevent convergence, so we need additional assumptions:
 \begin{enumerate}
@@
 \end{align*}
-where the sequence $\tilde{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}, ${\it {\bf N}} = [0,\dots,N)$, and ${\it {\bf B}} = [0,\dots,B)$. 
+where the sequence $\tilde{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}, ${\it {\bf N}}$ is the amount of trajectories sampled for the snapshot and ${\it {\bf B}}$ is the amount of trajectories sampled for a sub iteration.
 \end{definition}

### 2018-02-04 svrpg.tex

--- 2018-02-04 svrpg.tex
@@
 The main limitation of plain policy gradient has been the high variance of these estimators.
-The naive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples by interacting with the environment.
+The naive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples, \ie by interacting with the environment.
 For this reason, literature has focused on the introduction of baselines aiming to reduce the variance~\citep[\eg][]{Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.\footnote{A baseline is a function $f(s,a)$ such that $\EVV[a \sim \pol]{f(s,a)} = 0$, for any $s\in\Sspace$.}
@@
 %% \end{align*}
-A common requirement is that $g$ must be smooth and {\color{red} concave} in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} Under this hypothesis, full gradient ascent \cite{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.\todopirout{Is it $\ell_2$-norm?}
+A common requirement is that $g$ must be smooth and {\color{red} concave} in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} Under this hypothesis, full gradient ascent~\citep{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.\footnote{In the following we will omit the subscript for the $\ell_2$-norm.}
 This can be too expensive for large values of $N$.
 Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling.
-The resulting IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$ \cite{nemirovskii1983problem}\cite{reddi2016stochastic}.
+The resulting IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$~\cite{nemirovskii1983problem,reddi2016stochastic}.
 Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga} and Finito~\cite{defazio2014finito},

### 2018-02-04 svrpg.tex
Updated

@@
 \end{description}
-To deal with non-convexity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax (see App.~\ref{app:lsmoothness}.\todopirout{search for ref}
+To deal with non-convexity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax (see App.~\ref{app:lsmoothness}).\todopirout{search for ref}
 Because of the infinite dataset, we can only rely on an estimate of the full gradient.
+\citet{harikandeh2015stopwasting} analyzed this scenario under concavity \textcolor{red}{assumptions} of $g$, showing that SVRG is robust to inexact computation of the full gradient.
+In particular, it is still possible to recover the original convergence rate is the error decreases at an appropriate rate.
+{\color{red}Later on we will show how the estimation accuracy impacts on the convergence results with smooth objective.}
 \todopir{This should be argumented more via stop wasting my gradients}

### 2018-02-05 svrpg.tex
fixed some refs

@@
 \end{description}
-To deal with non-convexity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax (see App.~\ref{app:lsmoothness}).\todopirout{search for ref}
+To deal with non-convexity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as the truncated Gaussian (see App.~\ref{app:gauss}).
 Because of the infinite dataset, we can only rely on an estimate of the full gradient.
 \citet{harikandeh2015stopwasting} analyzed this scenario under concavity \textcolor{red}{assumptions} of $g$, showing that SVRG is robust to inexact computation of the full gradient.
-In particular, it is still possible to recover the original convergence rate is the error decreases at an appropriate rate.
+In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate.
 {\color{red}Later on we will show how the estimation accuracy impacts on the convergence results with smooth objective.}
 \todopir{This should be argumented more via stop wasting my gradients}
-Finally, the unbiasedness of Eq.~\ref{E:svrg.gradients.correction} is broken by the non-stationarity of the optimization problem.
+Finally, the unbiasedness of Eq.~\ref{E:svrg.gradient.correction} is broken by the non-stationarity of the optimization problem.
 To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.

### 2018-02-05 svrpg.tex
updated algorithm section

--- 2018-02-04 svrpg.tex
@@
 \newcommand{\HESSLOG}{F}
+% short forms 
+\newcommand{\wt}[1]{\widetilde{#1}}
+\newcommand{\wh}[1]{\widehat{#1}}
+\newcommand{\wo}[1]{\overline{#1}}
+\newcommand{\wb}[1]{\overline{#1}}
 %%%%%%
@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{SVRPG}
+\icmltitlerunning{Stochastic Variance-Reduced Policy Gradient}
 
@@
 \twocolumn[
-\icmltitle{Stochastic Variance Reduced Policy Gradient}
+\icmltitle{Stochastic Variance-Reduced Policy Gradient}
 
@@
 \begin{align} \label{E:policygradient.estimate}
-        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} \nabla f(\tau_i), \quad \tau_i \in \mathcal{D}_N^k.
+        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} g(\tau_i|\vtheta), \quad \tau_i \in \mathcal{D}_N^k.
 \end{align}
 The most widespread gradient estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
-We refer the reader to App.~\ref{A:gradient_estimators} for a formal definition.
+REINFORCE is such that $g(\tau|\vtheta) = \nabla \log\pi(\tau|\vtheta) \mathcal{R}(\tau)$,  we refer the reader to App.~\ref{A:gradient_estimators} for details.
 
@@
 There has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
-Altough these techniques has been mainly derived for finite-sum problems, we will show in Sec.~\ref{sec:alg} how they can be used in RL.
+Although these techniques has been mainly derived for finite-sum problems, we will show in Sec.~\ref{sec:alg} how they can be used in RL.
 The next section has the aim to describe variance reduction techniques for finite-sum problems. In particular, we will present the SVRG algorithm that is at the core of this work.
@@
 \begin{align*}
-        \max_{\vtheta} \left\{ f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g_i(\vtheta)\right\}.
-\end{align*}
-This kind of optimization is very common in machine learning, where each $g_i$ may correspond to a data sample $x_i$ from a dataset $\mathcal{D}_N$ of size $N$ (\ie $g_i(\vtheta) = g(x_i|\vtheta)$). 
+        \max_{\vtheta} \left\{ f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}z_i(\vtheta)\right\}.
+\end{align*}
+This kind of optimization is very common in machine learning, where each $z_i$ may correspond to a data sample $x_i$ from a dataset $\mathcal{D}_N$ of size $N$ (\ie $z_i(\vtheta) = z(x_i|\vtheta)$). 
 %% In this case, we adopt the following, more meaningful notation:
@@
 %% \end{align*}
-A common requirement is that $g$ must be smooth and {\color{red} concave} in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} Under this hypothesis, full gradient ascent~\citep{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.\footnote{In the following we will omit the subscript for the $\ell_2$-norm.}
+A common requirement is that $z$ must be smooth and {\color{red} concave} in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
+Under this hypothesis, full gradient (FG) ascent~\citep{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.\footnote{In the following we will omit the subscript for the $\ell_2$-norm.}
 This can be too expensive for large values of $N$.
@@
 Between snapshots, the parameter is updated with a gradient estimate corrected using stochastic gradient, for any $t \in \{0,\ldots,m-1\}$
-\begin{align}\label{E:svrg.gradient.correction}
-        \blacktriangledown f(\vtheta_{t+1}) = 
-	\nabla f(\tilde{\vtheta}) + 
-	\nabla g(x_i | \vtheta_t) -
-	\nabla g(x_i | \tilde{\vtheta}),
-\end{align} 
+\begin{equation}\label{E:svrg.gradient.correction}
+        \blacktriangledown f(\vtheta_{t+1}) = \nabla f(\tilde{\vtheta}) + \nabla z(x_i | \vtheta_t) - \nabla z(x_i | \tilde{\vtheta}),
+\end{equation} 
 where $x_i$ is sampled uniformly at random from $\mathcal{D}_N$ (\ie $x_i \sim \mathcal{U}(\mathcal{D}_N)$).
@@
 Pseudocode of SVRG for both the convex and the non-convex case is provided in Alg.~\ref{alg:svrg}.
-\todopir{SVRG for non-convex optimization requires to mix the parameter $\widetilde{\vtheta}^{s+1}$?}
+\todopir{SVRG for non-convex optimization requires to mix the parameter $\wt{\vtheta}^{s+1}$?}
 
@@
 		\FOR{$s=0$ {\bfseries to} $S-1$}
-		\STATE $\vtheta_0^{s+1} := \widetilde{\vtheta}^s := \vtheta_{m}^s$
+		\STATE $\vtheta_0^{s+1} := \wt{\vtheta}^s := \vtheta_{m}^s$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
@@
         \STATE $x_i \sim \mathcal{U}\left(\mathcal{D}_N\right)$
-		\STATE $v^{s+1} = 
-			\nabla f(\widetilde{\vtheta}^s) + 
+		\STATE $v^{s+1}_t = 
+			\nabla f(\wt{\vtheta}^s) + 
 			\nabla g(x_i|\vtheta_t^{s+1}) -
-			\nabla g(x_i|\widetilde{\vtheta}^{s})
+			\nabla g(x_i|\wt{\vtheta}^{s})
 		$
-        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}$
+        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
 		\ENDFOR
@@
 
-\section{Algorithm}\label{sec:alg}
+\section{SVRG in Reinforcement Learning}\label{sec:alg}
 In online RL problems the usual approach is to tune the batch size of SG in order to find the optimal trade-off between variance and speed.
@@
 \begin{description}
-        \item[Non-convexity:] the objective function $J(\vtheta)$ is typically non-convex;
+        \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
         \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The actual dataset is infinite and coincide with $\pi_{\vtheta}(\tau)$.
@@
 \end{description}
-To deal with non-convexity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax (see App.~\ref{app:lsmoothness}.\todopirout{search for ref}
+To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}, see App.~\ref{app:lsmoothness} for the derivation.\todopir{this should be shown in appendix}
 Because of the infinite dataset, we can only rely on an estimate of the full gradient.
-\todopir{This should be argumented more via stop wasting my gradients}
-Finally, the unbiasedness of Eq.~\ref{E:svrg.gradients.correction} is broken by the non-stationarity of the optimization problem.
+\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being \textcolor{red}{concave}, showing that SVRG is robust to inexact computation of the full gradient.
+In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate.
+{\color{red}Later on we will show how the estimation accuracy impacts on the convergence results with non-concave objective.}
+Finally, the unbiasedness of Eq.~\ref{E:svrg.gradient.correction} is broken by the non-stationarity of the optimization problem.
 To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
 %Finally, we employ importance weighting~\citep[\eg]{rubinstein1981simulation,precup2000eligibility} to guarantee proper sampling of trajectories.
-We focus on the REINFORCE gradient estimator for simplicity, but the extension to G(PO)MDP is straightforward.
-The proposed SVRG-like gradient estimate, which we call Stochastic Variance Reduced Policy Gradient (SVRPG), is the following:
-\begin{align*}
-	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
-		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
-		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i),
-\end{align*}
-where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from a snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\vert\tilde{\vtheta})}{\pi(\tau\vert\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$.
+
+We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG).
+% For simplicity, we focus on the REINFORCE estimator but the extension to G(PO)MDP is straightforward.
+Between snapshots, the SVRPG estimate is corrected using importance weighting, for any $t \in \{0,\ldots,m-1\}$
+\begin{align*}
+        \blacktriangledown J(\vtheta_{t+1}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
+\end{align*}
+% \begin{align*}
+% 	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
+% 		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
+% 		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i),
+% \end{align*}
+where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{\pi(\tau|\wt{\vtheta})}{\pi(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$.
 Our update is still fundamentally on-policy, since the weighting concerns only the correction term.
-However, this partial "off-policyness" is enough to introduce variance.
-To mitigate it, we use mini-batches of trajectories of size B to compute the second and the third term, as follows:
-\begin{align*}
-&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) +\\
-&\quad\frac{1}{B}\left[\sum_{i=0}^{B-1}
-\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
-- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i)\right].
-\end{align*}
-The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case \cite{konevcny2016mini} \cite{harikandeh2015stopwasting}.
-It is easy to show that the SVRPG gradient estimator has the following, desirable properties (detailed proofs in Appendix \ref{app:proofs}):
-\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
-The SVRPG estimator is unbiased:
+However, this partial ``off-policyness'' is enough to introduce variance.
+This is a well-known issue of importance sampling~\citep[\eg][]{thomas2015high}.
+To mitigate it, we use mini-batches of trajectories of size $B \ll N$ to average the correction, \ie
+\begin{align}\label{E:svrpg.estimate.batch}
+        \blacktriangledown J(\vtheta_{t+1}) &= \wh{\nabla}_N J(\wt{\vtheta})\\ \notag
+                                            & \quad{} + \frac{1}{B} \sum_{i=0}^{B-1} \left[
+        g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
+        \right].
+\end{align}
+%\begin{align*}
+%&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) +\\
+%&\quad\frac{1}{B}\left[\sum_{i=0}^{B-1}
+%\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
+%- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i)\right].
+%\end{align*}
+The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}.
+It is easy to show that the SVRPG estimator has the following, desirable properties (see App.~\ref{app:proofs}):
+\begin{lemma}
+Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$.
+Then, the SVRG estimate in~\eqref{E:svrpg.estimate.batch} is \emph{unbiased}
 \[
@@
 \]
-\end{restatable}
-\begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
-At convergence, the variance of the SVRPG estimator is equal to the variance of the REINFORCE estimator, regardless of the mini-batch size $B$:
+and regardless of the mini-batch size $B$:
 \[
 	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
-	\mathbb{V}ar\left[\gradApp{\vtheta^*}{N}\right].
+    \mathbb{V}ar\left[\wh{\nabla}_N^{\textsc{RF}}J(\vtheta^*)\right].
 \]
-\end{restatable}
-In particular, the latter suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can actually achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance, at least asymptotically.
-The resulting SVRPG policy optimization method is detailed in Algorithm \ref{alg:svrpg}.
+\end{lemma}
+
+%\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
+%The SVRPG estimator is unbiased:
+%\[
+%\mathop{\mathbb{E}}
+%\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
+%\]
+%\end{restatable}
+%\begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
+%At convergence, the variance of the SVRPG estimator is equal to the variance of the REINFORCE estimator, regardless of the mini-batch size $B$:
+%\[
+%	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
+%	\mathbb{V}ar\left[\gradApp{\vtheta^*}{N}\right].
+%\]
+%\end{restatable}
+In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can actually achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
+The resulting SVRPG policy optimization method is detailed in Alg.~\ref{alg:svrpg}.
 To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient, the mini-batch size $B$, and the step size $\alpha$.
-Adaptive selection of these meta-parameters is also possible, and is explored in Section \ref{sec:stopping}.
-However, the next section will focus on providing general converge guarantees for Algorithm \ref{alg:svrpg}.
+Adaptive selection of these meta-parameters is also possible, and is explored in Sec.~\ref{sec:stopping}.
+However, the next section will provide general converge guarantees for SVRPG as in Alg.~\ref{alg:svrpg}.
 
 \begin{algorithm}[tb]
-	\caption{SVRPG}
+        \caption{SVRPG based on REINFORCE}
 	\label{alg:svrpg}
 	\begin{algorithmic}
-		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m-1}^0$, batch size $N$, mini-batch size B
+		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m}^0$, batch size $N$, mini-batch size B
 		\FOR{$s=0$ {\bfseries to} $S-1$}
-		\STATE Sample N trajectories $\tau_j$ from $\pi(\cdot\vert\tilde{\vtheta}^{s})$
+        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $\pi(\cdot\vert\tilde{\vtheta}^{s})$
 		\STATE $\gradApp{\tilde{\vtheta}^{s}}{N} = \frac{1}{N}\sum_{j=0}^{N-1}\score{}{\tau_j\vert\tilde{\vtheta}^{s}}\Reward(\tau_j)$
-		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m-1}^s$
+		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m}^s$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
-		\STATE Sample $B$ trajectories $\tau_i$ from 				$\pi(\cdot\vert\vtheta_t^{s+1})$
-		\STATE 
-		\begin{align*}
-		\blacktriangledown J(\vtheta_t^{s+1}) = 
-		&\gradApp{\tilde{\vtheta}^s}{N} \\
-		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
-		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
-		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
-		\end{align*}
-		\STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha\blacktriangledown J(\vtheta_t^{s+1})$
+        \STATE Sample $B$ trajectories $\{\tau_i\}$ from $\pi(\cdot\vert\vtheta_t^{s+1})$
+        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
+        \begin{aligned}[t]
+                & \Reward(\tau_i) \Big[ \score{}{\tau_i|\vtheta_t^{s+1}} \\
+                &- \omega(\tau|\vtheta^{s+1}_t, \wt{\vtheta}^s)\score{}{\tau_i|\wt{\vtheta}^{s}} \Big]
+        \end{aligned}$
+        \STATE $v^{s+1}_t = \gradApp{\wt{\vtheta}^{s}}{N} + c^{s+1}_t$
+%%		\begin{align*}
+%%		\blacktriangledown J(\vtheta_t^{s+1}) = 
+%%		&\gradApp{\tilde{\vtheta}^s}{N} \\
+%%		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
+%%		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
+%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
+%%		\end{align*}
+        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
 		\ENDFOR
@@
 
-\section{Convergence Guarantees}\label{sec:conv}
+\section{Convergence Guarantees of SVRPG}\label{sec:conv}
 In this section we adapt existing convergence guarantees for nonconvex SVRG \cite{reddi2016stochastic} \cite{allen2016variance} to the Policy Gradient framework.
@@
 
-\unbias*
+%\unbias*
+\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
+The SVRPG estimator is unbiased:
+\[
+\mathop{\mathbb{E}}
+\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
+\]
+\end{restatable}
 \begin{proof}
@@
 
-\zerovar*
+% \zerovar*
+\begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
+At convergence, the variance of the SVRPG estimator is equal to the variance of the REINFORCE estimator, regardless of the mini-batch size $B$:
+\[
+	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
+	\mathbb{V}ar\left[\gradApp{\vtheta^*}{N}\right].
+\]
+\end{restatable}
 \begin{proof}

### 2018-02-05 svrpg.tex

@@
 \end{description}
-To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}, see App.~\ref{app:lsmoothness} for the derivation.\todopir{this should be shown in appendix}
+To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}, see App.~\ref{app:gauss} for the derivation.\todopir{this should be shown in appendix}
 Because of the infinite dataset, we can only rely on an estimate of the full gradient.

### 2018-02-05 svrpg.tex
Updated

@@
 \section{Convergence Guarantees of SVRPG}\label{sec:conv}
-In this section we adapt existing convergence guarantees for nonconvex SVRG \cite{reddi2016stochastic} \cite{allen2016variance} to the Policy Gradient framework.
-Each of the three challenges presented at the beginning of Section \ref{sec:alg} can potentially prevent convergence, so we need additional assumptions:
-\begin{enumerate}
-\item \textit{Non-convexity}:
-	as already mentioned, a common assumption for nonconvex optimization is the L-smoothness of the objective. We actually require the following, which is stronger:
+In this section, we state the convergence guarantees for SVRPG with REINFORCE structure (see Alg.~\ref{alg:svrpg}).
+We mainly leverage on the recent analysis of \textcolor{red}{nonconcave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
+Each of the three challenges presented at the beginning of Sec.~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
+In App.~\ref{app:gauss} we show how the widespread Gaussian policy satisfies these assumptions.
+
+\textit{1) Non-concavity.} A common assumption in this case it to assume the objective function to be $L$-smooth.
+However, in RL we can consider the following assumption which is sufficient for the $L$-smoothness of the objective (see Lem.~\ref{lemma:lsmooth}).
 	\begin{restatable}[On policy derivatives]{assumption}{boundedscore}\label{ass:bounded_score}
 		For each trajectory $\tau$, value of $\vtheta$ and dimensions $i,j$ there are positive constants $G,F<\infty$ such that:
-		\begin{align*}
-		&\left|\nabla_{\theta_i}\log\pi_{\vtheta}(\tau)\right| \leq \GRADLOG \\
-		&\left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(\tau)\right| \leq \HESSLOG
-		\end{align*}
+\[
+		\left|\nabla_{\theta_i}\log\pi_{\vtheta}(\tau)\right| \leq \GRADLOG, \qquad
+        \left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(\tau)\right| \leq \HESSLOG.
+\]
 	\end{restatable}
-	This property is a sufficient condition for the L-smoothness of the objective (see Lemma \ref{lemma:lsmooth} in the appendix).
-\item \textit{Approximation}:
-since we can only compute an estimate of the full gradient, we require its variance to be bounded:
+
+\textit{2) FG Approximation.}
+Since we cannot compute an exact full gradient, we require the variance of the estimator to be bounded.
+This assumption is in spirit similar to the one in~\citep{harikandeh2015stopwasting}.
 	\begin{restatable}[On the variance of the gradient estimator]{assumption}{varreinforce}\label{ass:REINFORCE}
@@
 		\]
+        \todopir{Why 1 as subscript?}
 	\end{restatable}
-The problem of full gradient approximation for SVRG has been addressed in the supervised learning framework \cite{harikandeh2015stopwasting} with results similar to ours.
-\item \textit{Non-stationarity}:
-the variance introduced by importance sampling must be bounded:
+
+\textit{3) Non-stationarity.} 
+Similarly to what done in SL~\citep{cortes2010learning}, we require the variance of the importance weight to be bounded.
 	\begin{restatable}[On the variance of importance weights]{assumption}{varweights}\label{ass:M2}
@@
 	\end{restatable}
-Differently from the previous two, this assumption can be enforced by a proper handling of the epoch size $m$.
-\end{enumerate}
-
-In Appendix \ref{app:gauss} we show how the commonly used Gaussian policy satisfies these assumptions.
-We can now state the following key result:
+Differently from Ass.~\ref{ass:bounded_score} and ~\ref{ass:REINFORCE}, Ass.~\ref{ass:M2} can be enforced by a proper handling of the epoch size $m$.
+
+We can now state the convergence guarantees for SVRPG.
+\todopir{Is it specific for REINFORCE?}
 \begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
-Under Assumptions \ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
+Under Ass.~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Alg.~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
 \begin{align*}
 	&\EVV[]
-	{\norm[]{\nabla J(\vtheta_A)}^2} 
+	{\norm[2]{\nabla J(\vtheta_A)}^2} 
 		\leq
@@
 \end{restatable}
-Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
-In Section \ref{sec:stopping}, instead, we propose a joint selection of step size $\alpha$ and epoch size $m$ which is more suited for applications.
-The $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on non-convex SVRG, \eg \cite{reddi2016stochastic}; the $\mathcal{O}(\nicefrac{1}{N})$ term is due to full gradient approximation and is analogous to the one in \cite{harikandeh2015stopwasting}; the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
+% Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
+Refer to App.~\ref{app:proofs} for a detailed proof involving definition of the constants and the meta-parameter constraints.
+%The theoretical settings may be too conservative.
+In Sec.~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$ which is more suited for applications. 
+By analysing the upper-bound in Thm.~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on \textcolor{red}{non-concave} SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
 To achieve true convergence, both the batch size $N$ and the mini-batch size $B$ should increase over time. 
-In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by full gradient approximation and importance sampling respectively.
-Since the optimal configuration may be task-dependent, we leave this choice to the empirical analysis of Section \ref{sec:exp}.
-
-\section{Hyperparameters}\label{sec:stopping}
-In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
-The first one is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
-The second one is essetial to control the variance introduced by the importance weights, mainly because the farther you go from the snapshot policy the higher the variance of the importance weights will be.
+In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by FG approximation and importance sampling, respectively.
+Since the optimal configuration may be task-dependent, we leave this choice to the empirical analysis of Sec.~\ref{sec:exp}.
+
+\section{Remarks on SVRPG}
+The convergence guarantees presented in the previous section come with requirements on the meta-parameters (\ie $m$ and $\alpha$) that may be too conservative for practical applications.
+Here we provide a practical and automatic way to choose the step size $\alpha$ and the number of sub-iterations $m$ performed between snapshots.
+Additionally, we provide variants of SVRPG exploiting variance-reduction techniques for importance weights.
+Despite lacking of theoretical guarantees, we will show that these methods are able to outperform the base SVRPG.
+
+\subsection{Meta-Parameters Selection}\label{sec:stopping}
+% In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
+At this point, the open question is how to properly select the step size $\alpha$ and the sub-iteration number $m$ in practical applications.
+The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
+The second one is essential to control the variance introduced by the importance weights, mainly because the farther you go from the snapshot policy the higher the variance of the importance weights will be.
 
@@
 
-\section{Practical Variants}\label{sec:prac}
+\subsection{Practical Variants}\label{sec:prac}
 \todomat{Missing opening to section}

### 2018-02-05 svrpg.tex
Updated

@@
 Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$.
-We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.\todopirout{See if it used}
+We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
 The search for a locally optimal policy is performed by means of gradient ascent, where the policy gradient
@@
 \subsection{Stochastic Variance-Reduced Gradient}
-Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $g_i(\cdot|\vtheta)$:
+Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $z_i(\cdot|\vtheta)$:
 \begin{align*}
@@
 A common requirement is that $z$ must be smooth and {\color{red} concave} in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
-Under this hypothesis, full gradient (FG) ascent~\citep{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.\footnote{In the following we will omit the subscript for the $\ell_2$-norm.}
+Under this hypothesis, full gradient (FG) ascent~\citep{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $\mathcal{O}(\nicefrac{N}{\epsilon})$, \ie it requires $k = \mathcal{O}(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.\footnote{In the following we will omit the subscript for the $\ell_2$-norm.}
 This can be too expensive for large values of $N$.
 Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling.
-The resulting IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$~\cite{nemirovskii1983problem,reddi2016stochastic}.
+The resulting IFO complexity is $\mathcal{O}(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$~\cite{nemirovskii1983problem,reddi2016stochastic}.
 Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga} and Finito~\cite{defazio2014finito},
@@
 The idea of SVRG is to alternate full and stochastic gradient updates. 
-Each $m = O(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i g(x_i|\widetilde{\vtheta})$.
+Each $m = \mathcal{O}(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i g(x_i|\widetilde{\vtheta})$.
 Between snapshots, the parameter is updated with a gradient estimate corrected using stochastic gradient, for any $t \in \{0,\ldots,m-1\}$
@@
 The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient.
-The resulting IFO complexity is $O(N+\nicefrac{\sqrt{N}}{\epsilon})$ \cite{reddi2016stochastic}.
+The resulting IFO complexity is $\mathcal{O}(N+\nicefrac{\sqrt{N}}{\epsilon})$ \cite{reddi2016stochastic}.
 
@@
 Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the \textcolor{red}{concave} case~\citep{nesterov2013introductory,ghadimi2013stochastic}.
-Also in this case, SVRG yields an advantageous trade-off, with an IFO complexity of $O(N + \nicefrac{N^{\nicefrac{2}{3}}}{N})$~\citep{reddi2016stochastic}.
+Also in this case, SVRG yields an advantageous trade-off, with an IFO complexity of $\mathcal{O}(N + \nicefrac{N^{\nicefrac{2}{3}}}{N})$~\citep{reddi2016stochastic}.
 The only additional requirement is to select $\vtheta^*$ uniformly at random among all the $\vtheta_k$ instead of simply setting it to the final value.
@@
 In this section, we state the convergence guarantees for SVRPG with REINFORCE structure (see Alg.~\ref{alg:svrpg}).
-We mainly leverage on the recent analysis of \textcolor{red}{nonconcave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
+We mainly leverage on the recent analysis of \textcolor{red}{non-concave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
 Each of the three challenges presented at the beginning of Sec.~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
@@
 \subsection{Meta-Parameters Selection}\label{sec:stopping}
+\todopir{I'm here!}
 % In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.

### 2018-02-05 svrpg.tex
Updated

@@
 
-Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$.
+Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^d\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$.
 We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
@@
 The idea of SVRG is to alternate full and stochastic gradient updates. 
-Each $m = \mathcal{O}(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i g(x_i|\widetilde{\vtheta})$.
+Each $m = \mathcal{O}(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i z(x_i|\widetilde{\vtheta})$.
 Between snapshots, the parameter is updated with a gradient estimate corrected using stochastic gradient, for any $t \in \{0,\ldots,m-1\}$
@@
 \]
+\todopir{Definition of Var??}
 \end{lemma}
@@
 	\begin{restatable}[On the variance of the gradient estimator]{assumption}{varreinforce}\label{ass:REINFORCE}
-		There is a constant $V<\infty$ such that, for each policy $\pol$:
+		There is a constant $V<\infty$ such that, for any policy $\pol$:
 		\[
@@
 	\begin{restatable}[On the variance of importance weights]{assumption}{varweights}\label{ass:M2}
-		There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
+		There is a constant $W<\infty$ such that, for each pair of policies encountered in Alg.~\ref{alg:svrpg} and for each possible trajectory,
 		\[
-		\mathbb{V}ar\left(\omega(\tau)\right) \leq \VARIS.
+                \mathbb{V}ar\left(\omega(\tau| \vtheta_1, \vtheta_2)\right) \leq \VARIS, \quad \forall \vtheta_1,\vtheta_2 \in \realspace^d , \tau \sim \pi_{\vtheta_1}.
 		\]

### 2018-02-05 svrpg.tex
Updated

@@
 % In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
-At this point, the open question is how to properly select the step size $\alpha$ and the sub-iteration number $m$ in practical applications.
-The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
-The second one is essential to control the variance introduced by the importance weights, mainly because the farther you go from the snapshot policy the higher the variance of the importance weights will be.
-
-A common way to reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter.
-Adam helps to stabilize the gradient estimate by computing adaptive learning rates for each parameter based on the history of the gradients' variance.
-
-Due to this feature we can use two different adam in order to properly leverage the structure of the proposed SVRPG update.
+% At this point, the open question is how to properly select the step size $\alpha$ and the sub-iteration number $m$ in practical applications.
+The step size $\alpha$ is crucial to balance the variance due to the sub-iterations and the variance associated to the estimator at the snapshot.
+% The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
+The number $m$ of sub-iterations controls the variance introduced by the importance weights.
+Low $m$ values are associated to low variance but induce lots of snapshot points (\ie lots of FG computations).
+High $m$ values allow to move policy $\pi_{\vtheta_t}$ away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing high variance of the importance weights.
+% the farther the distance between the snapshot policy and the current one, the higher the variance of the importance weights.
+
+\textbf{Adaptive step size.}
+A common way to deal with noisy gradients is to use adaptive strategies to compute the step size.
+ADAptive Moment estimation (ADAM)~\citep{kingma2014adam} stabilizes the parameter update by computing learning rates for each parameter based on an incremental estimate of the gradient variance.
+Due to this feature we would like to incorporate ADAM in the structure of the SVRPG update.
+Recall that SVRPG is characterized by two updates: I) FG update in the snapshot; II) corrected gradient update in the sub-iterations.
+Given this structure, we suggest to use two ADAM estimators:
+\begin{align*}
+        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
+        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_t v_t^{s+1}
+\end{align*}
+where $\alpha^{\textsc{FG}}$ is associated to the snapshot and $\alpha^{\textsc{SI}}$ with the sub-iterations.
+By doing so we decouple the contribution of the variance introduced by the approximate FG from the one introduced by the sub-iterations.
+Note that these two terms have different orders of magnitude since are estimated with a different number of trajectories ($B \ll N$) and w/o importance weights.
+The use of two ADAM estimators allows to capture and exploit this property.
+
+%Due to this feature we can use two different adam in order to properly leverage the structure of the proposed SVRPG update.
 The first adam will be associated to the snapshot updates and the second one will be associated to the sub-iteration updates.

### 2018-02-05 svrpg.tex
Updated
Updated algorithms and notation

@@
 The most widespread gradient estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
-REINFORCE is such that $g(\tau|\vtheta) = \nabla \log\pi(\tau|\vtheta) \mathcal{R}(\tau)$,  we refer the reader to App.~\ref{A:gradient_estimators} for details.
+For example, REINFORCE is such that $g(\tau|\vtheta) = \nabla \log\pi(\tau|\vtheta) \mathcal{R}(\tau)$. 
+Although REINFORCE definition is simpler, G(PO)MDP is often preferred due to its lower variance.
+We refer the reader to App.~\ref{A:gradient_estimators} for details.
 
@@
 \begin{equation}\label{E:svrg.gradient.correction}
-        \blacktriangledown f(\vtheta_{t+1}) = \nabla f(\tilde{\vtheta}) + \nabla z(x_i | \vtheta_t) - \nabla z(x_i | \tilde{\vtheta}),
+        \blacktriangledown f(\vtheta_{t}) = \nabla f(\tilde{\vtheta}) + \nabla z(x_i | \vtheta_t) - \nabla z(x_i | \tilde{\vtheta}),
 \end{equation} 
 where $x_i$ is sampled uniformly at random from $\mathcal{D}_N$ (\ie $x_i \sim \mathcal{U}(\mathcal{D}_N)$).
+Note that $t=0$ corresponds to a FG step (\ie $\blacktriangledown f(\vtheta_0) = \nabla f(\wt{\vtheta})$) since $\vtheta_0 := \wt{\vtheta}$.
 The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient.
@@
 Pseudocode of SVRG for both the convex and the non-convex case is provided in Alg.~\ref{alg:svrg}.
-\todopir{SVRG for non-convex optimization requires to mix the parameter $\wt{\vtheta}^{s+1}$?}
 
@@
 	\begin{algorithmic}
-            \STATE {\bfseries Input:} a dataset $\mathcal{D}_N$, number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m}^0$
+            \STATE {\bfseries Input:} a dataset $\mathcal{D}_N$, number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{0}^0:=\wt{\vtheta}^0$
 		\FOR{$s=0$ {\bfseries to} $S-1$}
-		\STATE $\vtheta_0^{s+1} := \wt{\vtheta}^s := \vtheta_{m}^s$
+        \STATE $\wt{\mu} = f(\wt{\vtheta}^s)$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
@@
         \STATE $x_i \sim \mathcal{U}\left(\mathcal{D}_N\right)$
-		\STATE $v^{s+1}_t = 
-			\nabla f(\wt{\vtheta}^s) + 
-			\nabla g(x_i|\vtheta_t^{s+1}) -
+		\STATE $v^{s}_t = 
+%			\nabla f(\wt{\vtheta}^s) + 
+			\wt{\mu} + 
+			\nabla g(x_i|\vtheta_t^{s}) -
 			\nabla g(x_i|\wt{\vtheta}^{s})
 		$
-        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
+        \STATE $\vtheta_{t+1}^{s} = \vtheta_t^{s} + \alpha v^{s}_t$
 		\ENDFOR
+        \STATE $\vtheta_0^{s+1} := \wt{\vtheta}^{s+1} = \vtheta_{m}^s$
 		\ENDFOR
-		\STATE {\bfseries Convex case: return} $\vtheta_{m}^S$
-		\STATE {\bfseries Non-Convex case: return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
+        \STATE \underline{Concave case:} \textbf{return} $\vtheta_{0}^S$
+        \STATE \underline{Non-Concave case:} \textbf{return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $\{[0,S[\times[0,m[\} \cup \{(S,0)\}$
 	\end{algorithmic}
@@
 
-We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG).
+We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG) for a generic policy gradient estimator $g$ (see Eq.~\ref{E:policygradient.estimate}).
 % For simplicity, we focus on the REINFORCE estimator but the extension to G(PO)MDP is straightforward.
@@
 \begin{align*}
-        \blacktriangledown J(\vtheta_{t+1}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
+        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
 \end{align*}
@@
 where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{\pi(\tau|\wt{\vtheta})}{\pi(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$.
+Similarly to SVRG, we have that $\vtheta_0 := \wt{\vtheta}$ and the update is a FG step.
 Our update is still fundamentally on-policy, since the weighting concerns only the correction term.
@@
 \begin{align}\label{E:svrpg.estimate.batch}
-        \blacktriangledown J(\vtheta_{t+1}) &= \wh{\nabla}_N J(\wt{\vtheta})\\ \notag
+        \blacktriangledown J(\vtheta_{t}) &:= v_t= \wh{\nabla}_N J(\wt{\vtheta})\\ \notag
                                             & \quad{} + \frac{1}{B} \sum_{i=0}^{B-1} \left[
@@
 Adaptive selection of these meta-parameters is also possible, and is explored in Sec.~\ref{sec:stopping}.
-However, the next section will provide general converge guarantees for SVRPG as in Alg.~\ref{alg:svrpg}.
+However, the next section will provide general converge guarantees for SVRPG with REINFORCE estimator.
 
 \begin{algorithm}[tb]
-        \caption{SVRPG based on REINFORCE}
+        \caption{SVRPG}
 	\label{alg:svrpg}
 	\begin{algorithmic}
-		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m}^0$, batch size $N$, mini-batch size B
+        \STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{0}^0 := \wt{\vtheta}^0$
 		\FOR{$s=0$ {\bfseries to} $S-1$}
         \STATE Sample $N$ trajectories $\{\tau_j\}$ from $\pi(\cdot\vert\tilde{\vtheta}^{s})$
-		\STATE $\gradApp{\tilde{\vtheta}^{s}}{N} = \frac{1}{N}\sum_{j=0}^{N-1}\score{}{\tau_j\vert\tilde{\vtheta}^{s}}\Reward(\tau_j)$
-		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m}^s$
+        \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\ref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
         \STATE Sample $B$ trajectories $\{\tau_i\}$ from $\pi(\cdot\vert\vtheta_t^{s+1})$
-        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
-        \begin{aligned}[t]
-                & \Reward(\tau_i) \Big[ \score{}{\tau_i|\vtheta_t^{s+1}} \\
-                &- \omega(\tau|\vtheta^{s+1}_t, \wt{\vtheta}^s)\score{}{\tau_i|\wt{\vtheta}^{s}} \Big]
-        \end{aligned}$
-        \STATE $v^{s+1}_t = \gradApp{\wt{\vtheta}^{s}}{N} + c^{s+1}_t$
+        \STATE $c^s_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^s) - \omega(\tau_i|\vtheta^s_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta})^s\right)$
+%        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
+%        \begin{aligned}[t]
+%                & \Reward(\tau_i) \Big[ \score{}{\tau_i|\vtheta_t^{s+1}} \\
+%                &- \omega(\tau|\vtheta^{s+1}_t, \wt{\vtheta}^s)\score{}{\tau_i|\wt{\vtheta}^{s}} \Big]
+%        \end{aligned}$
+        \STATE $v^{s}_t = \wt{\mu} + c^s_t$ % \gradApp{\wt{\vtheta}^{s}}{N} + c^{s}_t$
 %%		\begin{align*}
@@
 %%		\end{align*}
-        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
+        \STATE $\vtheta_{t+1}^{s} = \vtheta_t^{s} + \alpha v^{s}_t$
 		\ENDFOR
+        \STATE $\vtheta_0^{s+1} := \tilde{\vtheta}^{s+1} = \vtheta_{m}^s$
 		\ENDFOR
-		\STATE {\bfseries return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
+        \STATE {\bfseries return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $\{[0,S[\times[0,m[\} \cup \{(S,0)\}$
 	\end{algorithmic}
@@
 \section{Convergence Guarantees of SVRPG}\label{sec:conv}
-In this section, we state the convergence guarantees for SVRPG with REINFORCE structure (see Alg.~\ref{alg:svrpg}).
+In this section, we state the convergence guarantees for SVRPG with REINFORCE estimator (refer to Alg.~\ref{alg:svrpg} with $g(\tau|\vtheta) =  \nabla \log\pi(\tau|\vtheta) \mathcal{R}(\tau)$).\footnote{Similar results can be derived for G(PO)MDP, see App.~\ref{app:}.}
 We mainly leverage on the recent analysis of \textcolor{red}{non-concave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
@@
 We can now state the convergence guarantees for SVRPG.
-\todopir{Is it specific for REINFORCE?}
-\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
+\begin{restatable}[Convergence of the REINFORCE-SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
+Consider the REINFORCE estimator in Eq.~\ref{E:policygradient.estimate}.
 Under Ass.~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Alg.~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
@@
 
+\textbf{G(PO)MDP.} A similar result to the one stated in Thm~\ref{theo:convergence} can be derived for the G(PO)MDP estimator. A sketch of the proof is reported in~\ref{app:gpomdpproof}.
+In practice, due to the lower variance G(PO)MDP is preferred to REINFORCE.
+For this reason we will use G(PO)MDP for the experiments (see Sec.~\ref{sec:exp}).
+\todopir{sketch of proof for GPOMDP in appendix}
+
 \section{Remarks on SVRPG}
@@
 Additionally, we provide variants of SVRPG exploiting variance-reduction techniques for importance weights.
-Despite lacking of theoretical guarantees, we will show that these methods are able to outperform the base SVRPG.
+Despite lacking of theoretical guarantees, we will show that these methods are able to outperform the baseline SVRPG (Alg.~\ref{alg:svrpg}).
 
 \subsection{Meta-Parameters Selection}\label{sec:stopping}
-\todopir{I'm here!}
 % In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.

### 2018-02-05 svrpg.tex
updated definition of variance and notation

--- 2018-02-05 svrpg.tex
@@
 \[
-	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
-    \mathbb{V}ar\left[\wh{\nabla}_N^{\textsc{RF}}J(\vtheta^*)\right].
+	\Var\left[\gradBlack{\vtheta^*}\right] = 
+    \Var\left[\wh{\nabla}_N^{\textsc{RF}}J(\vtheta^*)\right].
 \]
-\todopir{Definition of Var??}
 \end{lemma}
+For any vector $\mathbf{x}$, we use $\Var[\mathbf{x}]$ to denote the trace of the covariance matrix, \ie $\Tr\EVV[]{(\mathbf{x}-\EVV[]{\mathbf{x}})(\mathbf{x}-\EVV[]{\mathbf{x}})^T}$.
 
@@
 		\[
-		\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq \VARRF.
+			\Var\left[g(\cdot\vert\vtheta)\right] \leq \VARRF.
 		\]
-        \todopir{Why 1 as subscript?}
 	\end{restatable}
@@
 		\[
-                \mathbb{V}ar\left(\omega(\tau| \vtheta_1, \vtheta_2)\right) \leq \VARIS, \quad \forall \vtheta_1,\vtheta_2 \in \realspace^d , \tau \sim \pi_{\vtheta_1}.
+                \mathbb{V}ar\left[\omega(\tau| \vtheta_1, \vtheta_2)\right] \leq \VARIS, \quad \forall \vtheta_1,\vtheta_2 \in \realspace^d , \tau \sim \pi_{\vtheta_1}.
 		\]
@@
 We give some additional definitions which will be useful in the proofs.
+
 \begin{definition}
@@
 \begin{definition}
+For random vectors X, Y:
+\begin{align*}
+	\Covts{X}{Y} &\coloneqq \Tr\left(\Ets{(X-\Ets{X})(Y-\Ets{Y})^T}\right), \\
+	\Varts{X} &\coloneqq \Covts{X}{Y},
+\end{align*}
+where $\Tr(\cdot)$ denotes the trace of a matrix.
+\end{definition}
+
+\begin{definition}
 The full gradient estimation error is:
@@
 \[
-	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
-	\mathbb{V}ar\left[\gradApp{\vtheta^*}{N}\right].
+	\Var\left[\gradBlack{\vtheta^*}\right] = 
+	\Var\left[\gradApp{\vtheta^*}{N}\right].
 \]
@@
 +L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
++\frac{1}{N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
 \nonumber 
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	+\frac{1}{N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
 	\nonumber\\
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	+\frac{1}{N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
 	&\qquad+\Ets{\norm[]{
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	+\frac{1}{N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
 	&\qquad+
@@
 	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	+\frac{1}{N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
 	\nonumber\\
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	+\frac{1}{N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
 	+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}\nonumber\\
@@
 \leq
-\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
+\frac{1}{2N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
 \end{align*}
@@
 	\dotprod{\Ets{e_s}}{\Ets{f(\vtheta_t^{s+1})}}
-	+\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}  \nonumber\\
+	+\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}  \nonumber\\
 	&= 
@@
 	&\qquad+
-	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})} \label{eq:8}
+	\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})} \label{eq:8}
 \end{align}
@@
 	=
-	\left|\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right|  
+	\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right|  
 	\nonumber\\
 	&\leq
-	\Tr\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right| \nonumber \\
+	\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}} \nonumber\\
 	%
 	&\leq	
-	\Tr\left(\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}}\right) \nonumber\\
+	\frac{\alpha_t}{2}\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Varts[t-1]{f(\vtheta_t^{s+1})}\label{eq:9}\\
+	%
+	&=
+	\frac{1}{2N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Varts[t-1]{f(\vtheta_t^{s+1})} \label{eq:10}\\
 	%
 	&\leq
-	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\label{eq:9}\\
-	%
-	&=
-	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \label{eq:10}\\
-	%
-	&\leq
-	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},
+	\frac{1}{2N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},
 	\nonumber
@@
 	&\qquad-
-	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}, \label{eq:13}
+	\frac{\alpha_t}{2N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}, \label{eq:13}
 \end{align}
@@
 &\qquad+
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
+\frac{\alpha_t}{N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
 %
@@
 &\qquad+
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+\frac{\alpha_t}{N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
 %
@@
 &\qquad
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
+\frac{\alpha_t}{N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
 \end{align}
@@
 	&\qquad-
-	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
+	\frac{\alpha_t}{2N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
 	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:17}\\
@@
 	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
-	-\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	-\frac{\alpha_t}{2N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
 	&\qquad -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
@@
 	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
-	-\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	-\frac{\alpha_t}{2N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
 	&\qquad -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2} -c_{t+1}\Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
@@
 	&\qquad
-	-c_{t+1}\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:18}\\
+	-c_{t+1}\frac{\alpha_t}{N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:18}\\
 	%
@@
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
 	%
@@
 	-\alpha_t^2\left(\frac{L}{2}+c_{t+1}\right)\left(\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	+\frac{1}{N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
 	\right.\nonumber\\
@@
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \label{eq:19}\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \label{eq:19}\\
 	%
@@
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
 	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2\VARIS}{2(1-\gamma)^2b} \nonumber\\
@@
 	+\Psi_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
-	-\frac{d_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	-\frac{d_t}{N}\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
 	-\frac{f_t}{b}\VARIS,\nonumber\\

### 2018-02-05 svrpg.tex
changed notation of 1-sample gradient estimator
Added adaptive sub-iterations

--- 2018-02-05 svrpg.tex
@@
 A common requirement is that $z$ must be smooth and {\color{red} concave} in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
-Under this hypothesis, full gradient (FG) ascent~\citep{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $\mathcal{O}(\nicefrac{N}{\epsilon})$, \ie it requires $k = \mathcal{O}(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.\footnote{In the following we will omit the subscript for the $\ell_2$-norm.}
+Under this hypothesis, full gradient (FG) ascent~\citep{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $\mathcal{O}(\nicefrac{N}{\epsilon})$, \ie it requires $k = \mathcal{O}(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.
+% \footnote{In the following we will omit the subscript for the $\ell_2$-norm.}
 This can be too expensive for large values of $N$.
@@
 More recently, some extensions of variance reduction algorithms to the non-convex objectives have been proposed~\citep[\eg][]{reddi2016stochastic,allen2016variance,reddi2016stochastic,reddi2016fast}.
-In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
+In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[2]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[2]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
 Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the \textcolor{red}{concave} case~\citep{nesterov2013introductory,ghadimi2013stochastic}.
@@
 In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can actually achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
-The resulting SVRPG policy optimization method is detailed in Alg.~\ref{alg:svrpg}.
+The resulting SVRPG policy optimization method is detailed in Alg.~\ref{alg:svrpg}.\footnote{Note that when $t=0$ we can avoid to sample trajectories and instead perform directly the policy update using the FG.
+For sake of clarity, we have not explicitly differentiate this step from the standard update rule in Alg.~\ref{alg:svrpg}.}
+
 To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient, the mini-batch size $B$, and the step size $\alpha$.
@@
         \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\ref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
+%        \STATE $ \vtheta^s_1 = \wt{\vtheta}^s + \alpha \wt{\mu}$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
@@
 \section{Remarks on SVRPG}
-The convergence guarantees presented in the previous section come with requirements on the meta-parameters (\ie $m$ and $\alpha$) that may be too conservative for practical applications.
+The convergence guarantees presented in the previous section come with requirements on the meta-parameters (\ie $\alpha$ and $m$) that may be too conservative for practical applications.
 Here we provide a practical and automatic way to choose the step size $\alpha$ and the number of sub-iterations $m$ performed between snapshots.
@@
 \begin{align*}
-        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
-        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_t v_t^{s+1}
-\end{align*}
-where $\alpha^{\textsc{FG}}$ is associated to the snapshot and $\alpha^{\textsc{SI}}$ with the sub-iterations.
+        \vtheta^{s}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
+        \vtheta^{s}_{t+1} &= \vtheta^{s}_t + \alpha^{\textsc{SI}}_{s,t} v_t^{s}
+\end{align*}
+where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s,t}$ with the sub-iterations.
+{\color{red} Note that the ADAM estimator used in the sub-iterations is reset at each snapshot with the FG estimate (see App.~\ref{app:adam} for details).}
+\todopir{Add ADAM definition and explain in more details}
 By doing so we decouple the contribution of the variance introduced by the approximate FG from the one introduced by the sub-iterations.
@@
 
-%Due to this feature we can use two different adam in order to properly leverage the structure of the proposed SVRPG update.
-The first adam will be associated to the snapshot updates and the second one will be associated to the sub-iteration updates.
-By doing so we are decoupling the contribution to the variance introduced by the snapshots from that one introduced by the sub-iterations.
-These two variance contributions lie, respectively, on two different order of magnitude because the number of trajectories used to perform the estimates for the snapshot is greater than that used for a sub-iteration.Therefore, using two adam optimizers allows us to properly adapt the learning rate in each situation.
-For the above mentioned reasons let us denote by $\alpha_{snap}$ the learning rate associated to the last snapshot and by  $\alpha_{sub-iter}$ the learning rate associated to the current sub-iteration.
-
-We have already said that as the number of sub-iterations $m$ increases the variance of the importance weights increases too.
-Therefore, if $m$ is not properly set the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
-
-We have seen, sperimentally, that a fixed epoch size is not the optimal choice, but an adaptive one would better exploit the total amount of data at hand.
-Hence, we permorm a new snapshot anytime the weighted learning rate proposed by the adam optimizer in a sub-iteration is smaller than the weighted one proposed at the immediately previous snapshot.
-This stopping condition is subject to a constraint on the maximum number of sub-iterations performable, which is problem dependent.
-More precisely we stop performing sub-iterations when:
-
-\[\frac{\alpha_{snap}}{N}>\frac{\alpha_{sub-iter}}{B}\]
-
-These two weights ${\bf{\it N}}$ and ${\bf{\it B}}$ represent, respectively, the number of trajectories used for a snapshot and the number of trajectories used for an update of the parameters in a sub-iteration.
-Therefore, in this stopping condition, we use them to take into account the trajectories efficiency.
-The less trajectories we use to perform an update, the better it is from a data usage point of view.
-Hence we want to facilitate sub-iterations with respect to snapshots.
-
-Another insight about this adaptive stopping condition lies within the value of the two different learning rate proposed by adam.
-Whenever the learning rate of a sub-iteration becomes too small with respect to that one proposed at the previous snapshot we are experiencing high variance.
-Hence it could be better to perform a new snapshot so to mitigate this issue.
-
-\subsection{Practical Variants}\label{sec:prac}
+\textbf{Adaptive Sub-Iterations.}
+It is easy to imagine that a predefined schedule (\eg $m$ fixed in advance or changed with a policy independent process) may poorly perform due to the high variability of the updates.
+In particular, fixed a value $m$ the variance of the updates in the sub-iterations depends on the snapshot policy and sampled trajectories.
+This variability is in part captured by the ADAM estimate.
+Hence, we propose to take a snapshot (\ie interrupt the sub-iterations) when the step size $\alpha^{\textsc{SI}}$ proposed by ADAM for the sub-iterations is smaller than the one for the FG (\ie $\alpha^{\textsc{FG}}$).
+If the latter condition is verified, it amounts to say that the noise in the corrected gradient has overcome the information of the FG.
+Formally, the stopping condition is as follow
+\[
+        \text{If }        \frac{\alpha^{\textsc{FG}}}{N} > \frac{\alpha^{\textsc{SI}}}{B} \textbf{ then } \text{take snapshot}
+\]
+where we have introduced $N$ and $B$ to take into account the trajectory efficiency (\ie weighted advantage).
+The less the number of trajectories used to update the policy, the better.
+This weighted approach comes from the fact that, as mentioned above, there is no advantage in having $m$ too small.
+Thus, we aim to facilitate sub-iterations rather than snapshots.
+
+\subsection{Normalized Importance Sampling}\label{sec:prac}
+\todopir{TODO: This section should be generic for normalized importance sampling trying to explain when and where are the advantages of doing this. Explicit formulation can be postponed in appendix.}
 \todomat{Missing opening to section}

### 2018-02-05 svrpg.tex
updated notation in Lemma B.4

--- 2018-02-05 svrpg.tex
@@
 +L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
++\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
 \nonumber 
@@
 	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
-			+\frac{1}{B}\sum_{i=0}^{B-1} \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) 
+			+\frac{1}{B}\sum_{i=0}^{B-1} g(\tau_i\vert\vtheta_t^{s+1}) 
 			-\frac{1}{B}\sum_{i=0}^{B-1}
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
+			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
 	%
@@
 			+\frac{1}{B}\sum_{i=0}^{B-1}\left( 
-			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)
+			g(\tau_i\vert\vtheta_t^{s+1}) -
+			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)
 			\right.\right.\nonumber\\&\qquad\left.\left.
@@
 		\frac{1}{B}\sum_{i=0}^{B-1}\left(
-		\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)
+		g(\tau_i\vert\vtheta_t^{s+1}) -
+			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)
 		\right.\right.\nonumber\\&\qquad\left.\left.
@@
 			\frac{1}{B}\sum_{i=0}^{B-1}\left(
-			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}\right\|^2\right] 
+			g(\tau_i\vert\vtheta_t^{s+1}) -
+				\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}\right\|^2\right] 
 	\nonumber\\
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
 	\nonumber\\
@@
 		\frac{1}{B}\sum_{i=0}^{B-1}\left(
-		\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-		\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)
+		g(\tau_i\vert\vtheta_t^{s+1}) -
+		\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)
 		\right.\right.\nonumber\\&\qquad\left.\left.
@@
 			\frac{1}{B}\sum_{i=0}^{B-1}\left(
-			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}\right\|^2\right] 
+			g(\tau_i\vert\vtheta_t^{s+1}) -
+			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}\right\|^2\right] 
 		\label{eq:1}\\%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	&\qquad+\Ets{\norm[]{
 			\frac{1}{B}\sum_{i=0}^{B-1}\left(
-			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}^2} \label{eq:2}\\
+			g(\tau_i\vert\vtheta_t^{s+1}) -
+			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}^2} \label{eq:2}\\
 	%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	&\qquad+
@@
 			\Ets{\norm[]{
-			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
+			g(\tau_i\vert\vtheta_t^{s+1}) -
+			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
 	%
 	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
 	\nonumber\\
@@
 			\frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
-			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
-			+\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
+			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
+			-g(\tau_i\vert\tilde{\vtheta}^s)
+			+g(\tau_i\vert\tilde{\vtheta}^s) 
+			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
 	%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
 	\nonumber\\
@@
 			\frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
-			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
+			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
+			-g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
 	&\qquad
 			+\frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
+			\Ets{\norm[]{g(\tau_i\vert\tilde{\vtheta}^s) 
+			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
 	%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
 	+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}\nonumber\\
@@
 		\frac{1}{B^2}\sum_{i=0}^{B-1}
-		\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \label{eq:3}
+		\Ets{\norm[]{g(\tau_i\vert\tilde{\vtheta}^s) 
+			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \label{eq:3}
 % ...
@@
 	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
 	\nonumber\\
@@
 where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (\ref{eq:3}) is from L-smoothness,
-(\ref{eq:4}) is from Assumption \ref{ass:bounded_score}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
+(\ref{eq:4}) is from Assumption \ref{ass:bounded_score} 
+{\todomatout{We could generalize to GPOMDP by defining some more constants}\color{red}and the definition of REINFORCE}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
 \end{proof}
@@
 \leq
-\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
+\frac{1}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
 \end{align*}
@@
 	=
-	\left|\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right|  
+	\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right|  
 	\nonumber\\
 	&\leq
-	\Tr\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right| \nonumber \\
+	\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}} \nonumber\\
 	%
 	&\leq	
-	\Tr\left(\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}}\right) \nonumber\\
+	\frac{\alpha_t}{2}\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Varts[t-1]{f(\vtheta_t^{s+1})}\label{eq:9}\\
+	%
+	&=
+	\frac{1}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Varts[t-1]{f(\vtheta_t^{s+1})} \label{eq:10}\\
 	%
 	&\leq
-	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\label{eq:9}\\
-	%
-	&=
-	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \label{eq:10}\\
-	%
-	&\leq
-	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},
+	\frac{1}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},
 	\nonumber
@@
 	&\qquad-
-	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}, \label{eq:13}
+	\frac{\alpha_t}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}, \label{eq:13}
 \end{align}
@@
 &\qquad+
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
+\frac{\alpha_t}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
 %
@@
 &\qquad+
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+\frac{\alpha_t}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
 %
@@
 &\qquad
-\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
+\frac{\alpha_t}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
 \end{align}
@@
 	&\qquad-
-	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
+	\frac{\alpha_t}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
 	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:17}\\
@@
 	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
-	-\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	-\frac{\alpha_t}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	&\qquad -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
@@
 	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
-	-\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	-\frac{\alpha_t}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	&\qquad -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2} -c_{t+1}\Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
@@
 	&\qquad
-	-c_{t+1}\frac{\alpha_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:18}\\
+	-c_{t+1}\frac{\alpha_t}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:18}\\
 	%
@@
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	%
@@
 	-\alpha_t^2\left(\frac{L}{2}+c_{t+1}\right)\left(\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
 	\right.\nonumber\\
@@
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \label{eq:19}\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \label{eq:19}\\
 	%
@@
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2\VARIS}{2(1-\gamma)^2b} \nonumber\\
@@
 	+\Psi_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
-	-\frac{d_t}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
+	-\frac{d_t}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
 	-\frac{f_t}{b}\VARIS,\nonumber\\

### 2018-02-05 svrpg.tex
Added abstract and checked intro

@@
 \begin{abstract}
-Why we need variance-reduced gradient in RL?
+        We present and analyse stochastic variance-reduced gradient (SVRG) in the context of Markov Decision Processes (MDPs).
+        We first show how to account for I) non-concativity of the objective function; II) approximations in the full gradient computation; and III) non-stationarity of the sampling process.
+        These characteristics make the application of SVRG to MDPs challenging.
+        We propose SVRPG, a stochastic variance-reduced policy gradient algorithm that leverages on importance weight to preserve the unbiasedness of the gradient estimate.
+        We further analyse its convergence guarantees under common assumptions on the MDP.
+        Finally, we investigate practical variants of SVRPG and we empirically evaluate them on continuous MDPs.
 \end{abstract}
@@
 Even in this case, stochastic optimization is often used for reasons of efficiency.
-The idea of stochastic gradient ascent \cite{nesterov2013introductory} is to iteratively focus on a random subset of the available data to obtain an approximate improvement direction.
+The idea of stochastic gradient (SG) ascent \cite{nesterov2013introductory} is to iteratively focus on a random subset of the available data to obtain an approximate improvement direction.
 At the level of the single iteration, this can be much less expensive than taking into account all the data.
-However, the sub-sampling of data samples is a source of variance that can potentially compromise convergence, so that per-iteration efficiency and convergence rate must be traded off with proper handling of meta-parameters.
+However, the sub-sampling of data is a source of variance that can potentially compromise convergence, so that per-iteration efficiency and convergence rate must be traded off with proper handling of meta-parameters.
 Variance-reduced gradient algorithms such as SAG \cite{roux2012stochastic}, SVRG \cite{johnson2013accelerating} and SAGA \cite{defazio2014saga} offer better ways of addressing this trade-off, with significant results both in theory and practice.
-Although designed explicitly for ERM, this algorithms address a problem that affects more general machine learning problems.
+Although designed explicitly for ERM, these algorithms address a problem that affects more general machine learning problems.
 In RL, stochastic optimization is rarely a matter of choice, since data must be actively sampled by interacting with an initially unknown environment.
@@
 The fundamental principle of this methods is to optimize a parametric policy by means of stochastic gradient ascent.
-Compared to other applications of SGD, the cost of collecting samples can be very high since it requires to interact with the environment.
+Compared to other applications of SG, the cost of collecting samples can be very high since it requires to interact with the environment.
 This makes SVRG-like methods potentially much more efficient than, \eg batch learning.
@@
 In Section \ref{sec:conv} we provide convergence guarantees for our algorithm which also offers insights on the challenges of variance reduction in RL.
-In Section \ref{sec:stopping} we suggests how to set the hyperparameters of SVRPG, while in Section \ref{sec:prac} we discuss some practical variants of the algorithm.
-Finally, in Section \ref{sec:exp} we empirically evaluate the performance of our method on popular continuous RL benchmark tasks.
+In Section \ref{sec:stopping} we suggests how to set the meta-parameters of SVRPG, while in Section \ref{sec:prac} we discuss some practical variants of the algorithm.
+Finally, in Section \ref{sec:exp} we empirically evaluate the performance of our method on popular continuous RL tasks.
 
 \section{Preliminaries}\label{sec:pre}
-We start providing the essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
+We start providing the essential background on policy gradient methods and stochastic variance-reduced gradient methods for finite-sum optimization.
 %In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
@@
 The most widespread gradient estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
-For example, REINFORCE is such that $g(\tau|\vtheta) = \nabla \log\pi(\tau|\vtheta) \mathcal{R}(\tau)$. 
 Although REINFORCE definition is simpler, G(PO)MDP is often preferred due to its lower variance.
@@
 \end{description}
-To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}, see App.~\ref{app:gauss} for the derivation.\todopir{this should be shown in appendix}
+To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}, see App.~\ref{app:gauss} for the derivation.
 Because of the infinite dataset, we can only rely on an estimate of the full gradient.
@@
 \end{align}
+We want just to underlying that $\EVV[\tau_i \sim \pi_{\vtheta_t}]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$~\citep[\eg][]{jurvcivcek2012reinforcement}.\footnote{The reader can refer to App.~\ref{} for off-policy variants of REINFORCE and G(PO)MDP.}
+This property will be used to prove Lem.~\ref{L:svrpg.properties}.
 %\begin{align*}
@@
 It is easy to show that the SVRPG estimator has the following, desirable properties (see App.~\ref{app:proofs}):
-\begin{lemma}
+\begin{lemma}\label{L:svrpg.properties}
 Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$.
@@
 \]
-and regardless of the mini-batch size $B$:
+and regardless of the mini-batch size $B$:\footnote{
+For any vector $\mathbf{x}$, we use $\Var[\mathbf{x}]$ to denote the trace of the covariance matrix, \ie $\Tr\EVV[]{(\mathbf{x}-\EVV[]{\mathbf{x}})(\mathbf{x}-\EVV[]{\mathbf{x}})^T}$.}
 \[
@@
 \end{lemma}
-For any vector $\mathbf{x}$, we use $\Var[\mathbf{x}]$ to denote the trace of the covariance matrix, \ie $\Tr\EVV[]{(\mathbf{x}-\EVV[]{\mathbf{x}})(\mathbf{x}-\EVV[]{\mathbf{x}})^T}$.
 
@@
 For this reason we will use G(PO)MDP for the experiments (see Sec.~\ref{sec:exp}).
-\todopir{sketch of proof for GPOMDP in appendix}
+\todopir{sketch of proof for GPOMDP in appendix, probably can be unified with the one of REINFORCE}
 
@@
 \todopir{TODO: This section should be generic for normalized importance sampling trying to explain when and where are the advantages of doing this. Explicit formulation can be postponed in appendix.}
-\todomat{Missing opening to section}
 We introduce a term with importance weights to guarantee proper sampling of trajectories.
@@
 \section{Related Work}
+\todopir{I think that this section is worth here}
 
@@
 
+\subsection{On-policy}
 The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
@@
 \end{align*}
+
+\subsection{Off-Policy}
+\todopir{Make ref to Luca's thesis and describe in brief the derivations}
 
@@
 	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
-			+\frac{1}{B}\sum_{i=0}^{B-1} g(\tau_i\vert\vtheta_t^{s+1}) 
+			+\frac{1}{B}\sum_{i=0}^{B-1} \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) 
 			-\frac{1}{B}\sum_{i=0}^{B-1}
-			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
 	%
@@
 			+\frac{1}{B}\sum_{i=0}^{B-1}\left( 
-			g(\tau_i\vert\vtheta_t^{s+1}) -
-			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)
+			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)
 			\right.\right.\nonumber\\&\qquad\left.\left.
@@
 		\frac{1}{B}\sum_{i=0}^{B-1}\left(
-		g(\tau_i\vert\vtheta_t^{s+1}) -
-			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)
+		\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)
 		\right.\right.\nonumber\\&\qquad\left.\left.
@@
 			\frac{1}{B}\sum_{i=0}^{B-1}\left(
-			g(\tau_i\vert\vtheta_t^{s+1}) -
-				\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}\right\|^2\right] 
+			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}\right\|^2\right] 
 	\nonumber\\
@@
 		\frac{1}{B}\sum_{i=0}^{B-1}\left(
-		g(\tau_i\vert\vtheta_t^{s+1}) -
-		\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)
+		\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+		\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)
 		\right.\right.\nonumber\\&\qquad\left.\left.
@@
 			\frac{1}{B}\sum_{i=0}^{B-1}\left(
-			g(\tau_i\vert\vtheta_t^{s+1}) -
-			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}\right\|^2\right] 
+			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}\right\|^2\right] 
 		\label{eq:1}\\%
@@
 			\frac{1}{B}\sum_{i=0}^{B-1}\left(
-			g(\tau_i\vert\vtheta_t^{s+1}) -
-			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}^2} \label{eq:2}\\
+			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}^2} \label{eq:2}\\
 	%
@@
 			\Ets{\norm[]{
-			g(\tau_i\vert\vtheta_t^{s+1}) -
-			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
+			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
+			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
 	%
@@
 			\frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
-			-g(\tau_i\vert\tilde{\vtheta}^s)
-			+g(\tau_i\vert\tilde{\vtheta}^s) 
-			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
+			\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
+			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
+			+\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
+			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
 	%
@@
 			\frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
-			-g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
+			\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
+			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
 	&\qquad
 			+\frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[]{g(\tau_i\vert\tilde{\vtheta}^s) 
-			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
+			\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
+			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
 	%
@@
 		\frac{1}{B^2}\sum_{i=0}^{B-1}
-		\Ets{\norm[]{g(\tau_i\vert\tilde{\vtheta}^s) 
-			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \label{eq:3}
+		\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
+			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \label{eq:3}
 % ...
@@
 where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (\ref{eq:3}) is from L-smoothness,
-(\ref{eq:4}) is from Assumption \ref{ass:bounded_score} 
-{\todomatout{We could generalize to GPOMDP by defining some more constants}\color{red}and the definition of REINFORCE}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
+(\ref{eq:4}) is from Assumption \ref{ass:bounded_score}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
 \end{proof}

### 2018-02-05 svrpg.tex

@@
 			\frac{1}{B}\sum_{i=0}^{B-1}\left(
-			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}\right\|^2\right] 
+			g(\tau_i\vert\vtheta_t^{s+1}) -
+			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}\right\|^2\right] 
 		\label{eq:1}\\%
@@
 			\frac{1}{B}\sum_{i=0}^{B-1}\left(
-			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}^2} \label{eq:2}\\
+			g(\tau_i\vert\vtheta_t^{s+1}) -
+			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}^2} \label{eq:2}\\
 	%
@@
 			\Ets{\norm[]{
-			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
-			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
+			g(\tau_i\vert\vtheta_t^{s+1}) -
+			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
 	%
@@
 			\frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
-			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
-			+\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
+			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
+			-g(\tau_i\vert\tilde{\vtheta}^s)
+			+g(\tau_i\vert\tilde{\vtheta}^s) 
+			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
 	%
@@
 			\frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
-			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
+			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
+			-g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
 	&\qquad
 			+\frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
-			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
+			\Ets{\norm[]{g(\tau_i\vert\tilde{\vtheta}^s) 
+			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
 	%

### 2018-02-06 svrpg.tex
sezione 5.2

@@
 \subsection{Normalized Importance Sampling}\label{sec:prac}
-\todopir{TODO: This section should be generic for normalized importance sampling trying to explain when and where are the advantages of doing this. Explicit formulation can be postponed in appendix.}
-We introduce a term with importance weights to guarantee proper sampling of trajectories.
-This term's variance is subjected to the variance of importance weights. 
-There are some methods to reduce this variance.
-
-As reported in Section \ref{subsec:PolicyGradient} we use G(PO)MDP gradient estimator~\cite{baxter2001infinite}, which gives us the same guarantees we have for the Reinforce Estimator.
-One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}, allowing a reduction of the gradient estimator variance.
-The off-policy gradient estimator becomes:
+As we already mentioned in section \ref{sec:stopping} importance weights means a lot of variance to deal with. Therefore, in oreder to cope with this issue the literature propose two main tecniques:
+\begin{enumerate}
+\item Self-normalization (?ciatazione bib?).
+\item Importance weights trucation (?citazione bib?).
+\end{enumerate}
+Both of them offer a variance reduction over the importance weights in exchange of a bias introduction.
+Hence, if the bias introduced by those techniques is smaller than the variance reduction they cause, it is convenient to use them in practice.
+So far, sperimentally, we have not seen any advantage in using importance weights truncation, hence we discard this technique for the previously mentioned purpose.
+Whereas for what concern self-normalization we expirienced some improvements using it (see section \ref{sec:exp} for the sperimental point of view).
+Self-normalization is easy to use, but depends on the chosen reinforcement learning algorithm.
+Therefore, here we give the definition of this technique in the context of G(PO)MDP where we have used the weighted per-decision importance sampling estimator \cite{precup2000eligibility}:
 \begin{align*}

### 2018-02-06 svrpg.tex
Updated

--- 2018-02-05 svrpg.tex
@@
 Although designed explicitly for ERM, these algorithms address a problem that affects more general machine learning problems.
+
 In RL, stochastic optimization is rarely a matter of choice, since data must be actively sampled by interacting with an initially unknown environment.
@@
 Unfortunately, RL has a series of difficulties that are not present in ERM.
-First, in SL the objective can often be designed to be strongly concave.
+First, in SL the objective can often be designed to be strongly concave (we aim to maximize).
 This is not possible in RL, so we have to deal with non-concavities.
@@
 This also rules out SAG and SAGA because of their storage requirements, which leaves SVRG as the most promising choice.
-Finally, the distribution used to sample data is not under direct control of the algorithm designer, but is a function of policy parameters that changes over time as the policy is optimized, which is a form of nonstationarity.
-
+Finally, the distribution used to sample data is not under direct control of the algorithm designer, but is a function of policy parameters that changes over time as the policy is optimized, which is a form of non-stationarity.
 % Related work?
+SVRG has been used in RL as a efficient technique for optimizing the per-iteration problem in Trust-Region Policy Optimization~\citep{xu2017svrgtrpo} or for policy evaluation~\citep{du2017svrgpe}.
+In both the cases, the optimization problems faced resemble the SL scenario and are not characterized by all the previously mentioned issues.
 
@@
 \end{align}
-We want just to underlying that $\EVV[\tau_i \sim \pi_{\vtheta_t}]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$~\citep[\eg][]{jurvcivcek2012reinforcement}.\footnote{The reader can refer to App.~\ref{} for off-policy variants of REINFORCE and G(PO)MDP.}
+We want just to stress that $\EVV[\tau_i \sim \pi_{\vtheta_t}]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to App.~\ref{} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}
 This property will be used to prove Lem.~\ref{L:svrpg.properties}.
@@
 Adaptive selection of these meta-parameters is also possible, and is explored in Sec.~\ref{sec:stopping}.
-However, the next section will provide general converge guarantees for SVRPG with REINFORCE estimator.
+However, the next section will provide general converge guarantees for SVRPG \textcolor{red}{with REINFORCE estimator}.
 
@@
 \section{Convergence Guarantees of SVRPG}\label{sec:conv}
-In this section, we state the convergence guarantees for SVRPG with REINFORCE estimator (refer to Alg.~\ref{alg:svrpg} with $g(\tau|\vtheta) =  \nabla \log\pi(\tau|\vtheta) \mathcal{R}(\tau)$).\footnote{Similar results can be derived for G(PO)MDP, see App.~\ref{app:}.}
+In this section, we state the convergence guarantees for SVRPG with REINFORCE estimator.\footnote{Similar results can be derived for G(PO)MDP, see App.~\ref{app:}.}
 We mainly leverage on the recent analysis of \textcolor{red}{non-concave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
@@
 
-\section{Discussion}
+\section{Conclusion}
+In this paper we introduce SVRPG, a variant of SVRG specifically designed for RL problems.
+RL has a series of difficulties non-common in SL.
+Among them non-concavity and approximate estimate of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
+Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy dependent.
+We have shown that by exploiting importance weighting techniques it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
+SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
+The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
+In the second case it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
+Additional, as far as we know, we are the first to report convergence guarantees for SVRG in the RL scenario.
+Empirically, we have shown that practical variants of the theoretical SVRPG are able to outperform classical actor-only approaches on several benchmarks.
 

### 2018-02-06 svrpg.tex
updated proof

@@
 \section{Convergence Guarantees of SVRPG}\label{sec:conv}
-In this section, we state the convergence guarantees for SVRPG with REINFORCE estimator.\footnote{Similar results can be derived for G(PO)MDP, see App.~\ref{app:}.}
+In this section, we state the convergence guarantees for SVRPG {\color{red}with REINFORCE estimator}.\footnote{\color{red}Similar results can be derived for G(PO)MDP, see App.~\ref{app:}.}
 We mainly leverage on the recent analysis of \textcolor{red}{non-concave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
@@
 	\begin{restatable}[On policy derivatives]{assumption}{boundedscore}\label{ass:bounded_score}
-		For each trajectory $\tau$, value of $\vtheta$ and dimensions $i,j$ there are positive constants $G,F<\infty$ such that:
+		For each state-action pair $(s,a)$, value of $\vtheta$ and dimensions $i,j$ there are positive constants $G,F<\infty$ such that:
 \[
-		\left|\nabla_{\theta_i}\log\pi_{\vtheta}(\tau)\right| \leq \GRADLOG, \qquad
-        \left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(\tau)\right| \leq \HESSLOG.
+		\left|\nabla_{\theta_i}\log\pi_{\vtheta}(a\vert s)\right| \leq \GRADLOG, \qquad
+        \left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(a \vert s)\right| \leq \HESSLOG.
 \]
@@
 We can now state the convergence guarantees for SVRPG.
-\begin{restatable}[Convergence of the REINFORCE-SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
-Consider the REINFORCE estimator in Eq.~\ref{E:policygradient.estimate}.
+\begin{restatable}[Convergence of the {\color{red}REINFORCE-}SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
+{\color{red}Consider the REINFORCE estimator in Eq.~\ref{E:policygradient.estimate}.}
 Under Ass.~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Alg.~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
@@
 where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s,t}$ with the sub-iterations.
-{\color{red} Note that the ADAM estimator used in the sub-iterations is reset at each snapshot with the FG estimate (see App.~\ref{app:adam} for details).}
+{\color{red} Note that the ADAM estimator used in the sub-iterations is reset at each snapshot \todomatout{actually not reset} with the FG estimate (see App.~\ref{app:adam} for details).}
 \todopir{Add ADAM definition and explain in more details}
@@
 \begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
-At convergence, the variance of the SVRPG estimator is equal to the variance of the REINFORCE estimator, regardless of the mini-batch size $B$:
+At convergence, the variance of the SVRPG estimator is equal to the variance of the gradient estimator, regardless of the mini-batch size $B$:
 \[
@@
 &= \int_{\Tspace}\pol(\tau)\score{\theta}{\tau}\score{\theta}{\tau}^T\Reward(\tau)\de \tau + \int_{\Tspace}\pol(\tau)\Dij{}{\theta}\log\pol(\tau)\Reward(\tau)\de \tau \nonumber\\
-&\leq \sup_{\tau} \left\{\left|\Reward(\tau)\right|\right\} \left(\GRADLOG^2+\HESSLOG\right) \label{eq:0}\\
-&= \frac{R}{1-\gamma}\left(\GRADLOG^2+\HESSLOG\right),\nonumber
+&\leq \sup_{\tau} \left\{\left|\Reward(\tau)\right|\right\} \left(H^2\GRADLOG^2+H\HESSLOG\right) \label{eq:0}\\
+&= \frac{1-\gamma^H}{1-\gamma}RH\left(H\GRADLOG^2+\HESSLOG\right),\nonumber
 \end{align}
@@
 
+\begin{restatable}[]{lemma}{bounded-g}\label{lemma:gbound}
+Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, for every $\tau\in\Tspace$ and $\vtheta\in\Theta$, there is a positive constant $\Gamma<\infty$ such that:
+\[
+	\norm[]{g(\tau\vert\vtheta)}^2 \leq \Gamma.
+\]
+\end{restatable}
+\begin{proof}
+For REINFORCE we have, from Assumption \ref{ass:bounded_score}:
+\begin{align*}
+	\norm[]{g(\tau\vert\vtheta)}^2 &=
+	\norm[]{\score{\vtheta}{\tau}R(\tau)}^2 \\
+	&=\norm[]{\left(\sum_{t=0}^{H-1}\score{\vtheta}{a_t\vert s_t}\right)\left(\sum_{t=0}^{H-1}\gamma^t r_t\right)}^2\leq H^2G^2\frac{(1-\gamma^H)^2}{(1-\gamma)^2}R^2\dim(\vTheta) \coloneqq \Gamma
+\end{align*}
+For G(PO)MDP, we do not have a compact expression for $g$, but since it is derived from REINFORCE by neglecting some terms of the kind $\score{\vtheta}{a_t\vert s_t}\gamma^t r_t$ \cite{baxter2001infinite}\cite{peters2008reinforcement}, the above bound still holds.
+\end{proof}
 %Lemma 2
@@
 \nonumber 
-+\frac{G^2R^2W}{(1-\gamma)^2B}
++\frac{\Gamma\VARIS}{B}
 \]
@@
 	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\GRADLOG^2\frac{R^2}{(1-\gamma)^2}\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \label{eq:4}\\
+	+\Gamma\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \label{eq:4}\\
 	%
@@
 	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\GRADLOG^2\frac{R^2}{(1-\gamma)^2b^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
+	+\frac{\Gamma}{B^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
 	%
@@
 	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\GRADLOG^2\frac{R^2\VARRF}{(1-\gamma)^2B}, \label{eq:5}
+	+\frac{\Gamma\VARIS}{B}, \label{eq:5}
 \end{align}
 where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (\ref{eq:3}) is from L-smoothness,
-(\ref{eq:4}) is from Assumption \ref{ass:bounded_score} 
-{\todomatout{We could generalize to GPOMDP by defining some more constants}\color{red}and the definition of REINFORCE}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
+(\ref{eq:4}) is from Lemma \ref{lemma:gbound}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
 \end{proof}
@@
 	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_tV}{N\Psi_t}
-	+\frac{f_tW}{b\Psi_t},
+	+\frac{f_tW}{B\Psi_t},
 \]
@@
 	&c_{m-1} = 0, \\
-	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\alpha_t^2\right) + \alpha_t\frac{L^2}{b}+\frac{\alpha_t^2L^3}{2}, \\
+	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\alpha_t^2\right) + \alpha_t\frac{L^2}{B}+\frac{\alpha_t^2L^3}{2}, \\
 	&\Psi_t = \alpha_t\left(1-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
 	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right), \\
-	&f_t = \alpha_t^2\frac{G^2R^2(L+2c_{t+1})}{2(1-\gamma)^2}.
+	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2}.
 \end{align*}
@@
 	&\left.\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+G^2\frac{R^2}{(1-\gamma)^2b}W\right)
+	+\frac{\Gamma W}{B}\right)
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
@@
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
-	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2\VARIS}{2(1-\gamma)^2b} \nonumber\\
+	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})\Gamma\VARIS}{2B} \nonumber\\
 	%
@@
 \end{align}
-where (\ref{eq:17}) is from (\ref{eq:13}), (\ref{eq:18}) is from (\ref{eq:16}), (\ref{eq:19}) is from Lemma \ref{lemma:aux0}, and (\ref{eq:20}) is from Assumption \ref{ass:REINFORCE}.
+where (\ref{eq:17}) is from (\ref{eq:13}), (\ref{eq:18}) is from (\ref{eq:16}), (\ref{eq:19}) is from Lemma \ref{lemma:aux2}, and (\ref{eq:20}) is from Assumption \ref{ass:REINFORCE}.
 To complete the proof, besides rearranging terms, we have to ensure that $\Psi_t>0$ for each $t$. This gives the constraints on $\alpha_t$ and $\beta_t$.

### 2018-02-06 svrpg.tex
fixed some constants in proof
edited section 5.2

@@
 where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s,t}$ with the sub-iterations.
-{\color{red} Note that the ADAM estimator used in the sub-iterations is reset at each snapshot \todomatout{actually not reset} with the FG estimate (see App.~\ref{app:adam} for details).}
+
 \todopir{Add ADAM definition and explain in more details}
@@
 \subsection{Normalized Importance Sampling}\label{sec:prac}
-As we already mentioned in section \ref{sec:stopping} importance weights means a lot of variance to deal with. Therefore, in oreder to cope with this issue the literature propose two main tecniques:
-\begin{enumerate}
-\item Self-normalization (?ciatazione bib?).
-\item Importance weights trucation (?citazione bib?).
-\end{enumerate}
-Both of them offer a variance reduction over the importance weights in exchange of a bias introduction.
-Hence, if the bias introduced by those techniques is smaller than the variance reduction they cause, it is convenient to use them in practice.
-So far, sperimentally, we have not seen any advantage in using importance weights truncation, hence we discard this technique for the previously mentioned purpose.
-Whereas for what concern self-normalization we expirienced some improvements using it (see section \ref{sec:exp} for the sperimental point of view).
+As we already mentioned in section \ref{sec:stopping} importance weights means a lot of variance to deal with.
+Therefore, in oreder to cope with this issue the literature propose the Self-normalization techniques \cite{precup2000eligibility}.
+This technique offers a variance reduction over the importance weights in exchange of a bias introduction.
+Hence, if the bias introduced by this is smaller than the variance reduction it causes, it is convenient to use it in practice.
+We expirienced some improvements using it (see section \ref{sec:exp} for the sperimental point of view).
 Self-normalization is easy to use, but depends on the chosen reinforcement learning algorithm.
@@
 	&c_{m-1} = 0, \\
-	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\alpha_t^2\right) + \alpha_t\frac{L^2}{B}+\frac{\alpha_t^2L^3}{2}, \\
+	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\frac{\alpha_t^2L^2}{B}\right) + \alpha_tL^2+\frac{\alpha_t^2L^3}{2B}, \\
 	&\Psi_t = \alpha_t\left(1-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
@@
 	\right.\nonumber\\
-	&\left.\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	&\left.\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 	+\frac{\Gamma W}{B}\right)
@@
 	%
-	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta+\alpha_t^2\right)+\alpha_t\frac{L^2}{b}+\frac{\alpha_t^2L^3}{2}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta+\frac{\alpha_t^2L^2}{B}\right)+\alpha_tL^2+\frac{\alpha_t^2L^3}{2B}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad

### 2018-02-06 svrpg.tex
Edited Experimental Details

@@
 The episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
-	\item \emph{Mujoco Swimmer}: 12-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
+	\item \emph{Mujoco Swimmer}: 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
 2-dimensional action space: the two momentums applied on actuated joints.
@@
 	\hline
-	State space dimension  & 4 & 8 & 20 \\
+	State space dimension  & 4 & 13 & 20 \\
 	Control space dimension & 1 & 2 & 6 \\
@@
 	Discount $(\gamma)$& 0.99 & 0.995 & 0.99 \\
-	Total trajectories& 10000 & 10000 & 10000 \\
+	Total trajectories& 10000 & 20000 & 50000 \\
 	\hline  

### 2018-02-06 svrpg.tex
improved some notation + added some proof comments and todos

@@
 \newcommand{\Ets}[2][t]{\mathbb{E}_{#1\vert s}\left[#2\right]}
+\newcommand{\Es}[1]{\mathbb{E}_{s}\left[#1\right]}
 \newcommand{\Covts}[3][t]{{\mathbb{C}\text{ov}}_{#1\vert s}\left(#2,#3\right)}
+\newcommand{\Covs}[2]{{\mathbb{C}\text{ov}}_{s}\left(#1,#2\right)}
 \newcommand{\Varts}[2][t]{{\mathbb{V}\text{ar}}_{#1\vert s}\left[#2\right]}
+\newcommand{\Vars}[1]{{\mathbb{V}\text{ar}}_{s}\left[#1\right]}
 \newcommand{\gradBlack}[1]{\blacktriangledown J(#1)}
@@
 The resulting IFO complexity is $\mathcal{O}(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$~\cite{nemirovskii1983problem,reddi2016stochastic}.
+\todomat{We are not interested in IFO complexity, report convergence rate instead}
 Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga} and Finito~\cite{defazio2014finito},
@@
 \end{algorithm}
+\todomat{Mention the fact that we return last parameter in practice}
 
@@
 \begin{definition}
+For a random variable X:
+\[
+	\Es{X} = \EVV[\tau_j\sim\pi(\cdot\vert\tilde{\vtheta}^s)
+		\forall j\in\mathbf{N}]{X\vert\tilde{\vtheta}^s},
+\]
+where $\tilde{\vtheta}^s$ is defined in Algorithm \ref{alg:svrpg} and $\mathbf{N} = [0,\dots,N)$.
+\end{definition}
+
+\begin{definition}
 For a random variable $X$:
@@
 \begin{align*}
-	\Covts{X}{Y} &\coloneqq \Tr\left(\Ets{(X-\Ets{X})(Y-\Ets{Y})^T}\right), \\
-	\Varts{X} &\coloneqq \Covts{X}{Y},
-\end{align*}
-where $\Tr(\cdot)$ denotes the trace of a matrix.
+	\Covs{X}{Y} &\coloneqq \Tr\left(\Ets{(X-\Ets{X})(Y-\Ets{Y})^T}\right), \\
+	\Vars{X} &\coloneqq \Covs{X}{Y},
+\end{align*}
+where $\Tr(\cdot)$ denotes the trace of a matrix. $\Covts[t]{X}{Y}$ and $\Varts[t]{X}$ are defined in the same way from $\Ets[t]{X}$.
 \end{definition}
@@
 +L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
++\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
 \nonumber 
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
-	+\Ets[]{\norm[]{\gradApp{\tilde{\vtheta}^s}{N} - \Ets[]{\gradApp{\tilde{\vtheta}^s}{N}}}^2} \nonumber\\
+	+\Es{\norm[]{\gradApp{\tilde{\vtheta}^s}{N} - \Es{\gradApp{\tilde{\vtheta}^s}{N}}}^2} \nonumber\\
 	&\qquad+ 
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
+	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
 	\nonumber\\
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
+	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	&\qquad+\Ets{\norm[]{
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
+	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	&\qquad+
@@
 	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
+	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
 	\nonumber\\
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
+	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
 	+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}\nonumber\\
@@
 \begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
-Under Assumption \ref{ass:bounded_score}, for any function $f(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$:
-\begin{align*}
-\left|\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
--\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
+Under Assumption \ref{ass:bounded_score}, for any function $\varphi(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$:
+\begin{align*}
+\left|\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
+-\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
 \right|
 \leq
-\frac{1}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
+\frac{1}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2}
 \end{align*}
@@
 \begin{align}
-	\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
+	\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
 	&=
-	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
-	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \label{eq:6}\\
+	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} +
+	\Ets[t-1]{\dotprod{e_s}{\varphi(\vtheta_t^{s+1})}} \label{eq:6}\\
 	&=
-	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
-	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \label{eq:7}\\
+	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} +
+	\Ets[t-1]{\dotprod{e_s}{\varphi(\vtheta_t^{s+1})}} \label{eq:7}\\
 	&=
-	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
+	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
 	&\qquad+
-	\dotprod{\Ets{e_s}}{\Ets{f(\vtheta_t^{s+1})}}
-	+\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}  \nonumber\\
+	\dotprod{\Ets{e_s}}{\Ets{\varphi(\vtheta_t^{s+1})}}
+	+\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}  \label{eq:new1}\\
 	&= 
-	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
+	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
 	&\qquad+
-	\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})} \label{eq:8}
+	\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})} \label{eq:8}
 \end{align}
@@
 \begin{align}
-	\left|\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
-	\right.&-\left.\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}\right| 
+	\left|\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
+	\right.&-\left.\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}\right| 
 	=
-	\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right|  
+	\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}\right|  
 	\nonumber\\
 	&\leq
-	\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}} \nonumber\\
+	\sqrt{\Vars{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{\varphi(\vtheta_t^{s+1})}} \nonumber\\
 	%
 	&\leq	
-	\frac{\alpha_t}{2}\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Varts[t-1]{f(\vtheta_t^{s+1})}\label{eq:9}\\
+	\frac{1}{2}\Vars{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})}\label{eq:9}\\
 	%
 	&=
-	\frac{1}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Varts[t-1]{f(\vtheta_t^{s+1})} \label{eq:10}\\
+	\frac{1}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})} \label{eq:10}\\
 	%
 	&\leq
-	\frac{1}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},
+	\frac{1}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2},
 	\nonumber
 \end{align}
-where (\ref{eq:6}) is from Definition \ref{def:ideal}; (\ref{eq:7}) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (\ref{eq:8}) is from $\Ets{e_s}=0$; (\ref{eq:9}) is from Young's inequality; and (\ref{eq:10}) is from the definition of $\gradApp{\vtheta}{N}$.
+where (\ref{eq:6}) is from Definition \ref{def:ideal}; (\ref{eq:7}) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $\varphi(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$;
+(\ref{eq:new1}) is from the fact that $\gradJ{\tilde{\vtheta}^s}$ is constant \wrt $\Vars{\cdot}$;
+(\ref{eq:8}) is from $\Ets{e_s}=0$; (\ref{eq:9}) is from Young's inequality; and (\ref{eq:10}) is from the definition of $\gradApp{\vtheta}{N}$.
 \end{proof}
@@
 	&\qquad-
-	\frac{\alpha_t}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}, \label{eq:13}
-\end{align}
+	\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}, \label{eq:13}
+\end{align}
+\todomat{check last inequality}
 where (\ref{eq:11}) is from the L-smoothness of $J(\vtheta)$, (\ref{eq:12}) is from the SVRPG update, and (\ref{eq:13}) is from Lemma \ref{lemma:aux0}.
@@
 &\qquad+
-\frac{\alpha_t}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
+\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
 %
@@
 &\qquad+
-\frac{\alpha_t}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
 %
@@
 &\qquad
-\frac{\alpha_t}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
+\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
 \end{align}
@@
 	&\qquad-
-	\frac{\alpha_t}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
+	\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
 	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:17}\\
@@
 	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
-	-\frac{\alpha_t}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
+	-\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	&\qquad -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
@@
 	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
-	-\frac{\alpha_t}{2N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
+	-\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	&\qquad -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2} -c_{t+1}\Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
@@
 	&\qquad
-	-c_{t+1}\frac{\alpha_t}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:18}\\
+	-c_{t+1}\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:18}\\
 	%
@@
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	%
@@
 	-\alpha_t^2\left(\frac{L}{2}+c_{t+1}\right)\left(\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
+	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
 	\right.\nonumber\\
@@
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \label{eq:19}\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \label{eq:19}\\
 	%
@@
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})\Gamma\VARIS}{2B} \nonumber\\
@@
 	+\Psi_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
-	-\frac{d_t}{N}\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)}
+	-\frac{d_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
 	-\frac{f_t}{b}\VARIS,\nonumber\\

### 2018-02-06 svrpg.tex
Added related work

@@
 \todomat{why should we use it? What advantages do we expect? At what cost? Pointer to experiments}
+
 \section{Related Work}
-\todopir{I think that this section is worth here}
+Despite the huge interest received in SL, variance-reduced gradient approaches has not attracted the RL community.
+As far as we know, there are just two applications of SVRG in RL.
+The first approach~\citep{du2017svrgpe} aims to exploit SVRG for policy evaluation.
+The policy evaluation problem is simpler than the one faced in this paper (control problem).
+In particular, since the goal is to evaluate just the performance of a predefined policy, the optimization problem is stationary.
+The setting considered in the paper is the one of policy evaluation by minimizing the empirical mean squared projected Bellman error (MSPBE) with a linear approximation of the value function.
+\citet{du2017svrgpe} shown that this problem can equivalently reformulated as a convex-concave saddle-point problem that is characterized by a finite-sum structure.
+This problem can be solved using a variant of SVRG~\citep{Palaniappan2016svrgsaddle} for which convergence guarantees have been provided.
+In the second approach~\citep{xu2017svrgtrpo}, SVRG is just an efficient approach to solve the optimization problem faced by Trust Region Policy Optimization (TRPO) at each iteration.
+We can just talk of a direct application of SVRG to a problem having finite-sum structure since no specific structure of the RL problem is exploited.
+It is worth to mention that, for practical reasons, the authors proposed to use a Newton conjugate gradient method with SVRG.
+
+Concerning importance weighting techniques, RL has made wide use of them for off-policy problems~\citep[\eg]{precup2000eligibility,thomas2015high}.
+However, as mentioned before, SVRPG cannot be compared to such methods since it is practically on-policy.
+Here, importance weighting is just a statistical tool used to preserve the unbiasedness of the corrected gradient.
 
@@
 In this paper we introduce SVRPG, a variant of SVRG specifically designed for RL problems.
-RL has a series of difficulties non-common in SL.
+The control problem considered in the paper has a series of difficulties non-common in SL.
 Among them non-concavity and approximate estimate of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
@@
 We have shown that by exploiting importance weighting techniques it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
+We have additionally shown that under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
 SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
@@
 In the second case it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
-Additional, as far as we know, we are the first to report convergence guarantees for SVRG in the RL scenario.
-Empirically, we have shown that practical variants of the theoretical SVRPG are able to outperform classical actor-only approaches on several benchmarks.
+Finally, we have empirically shown that practical variants of the theoretical SVRPG are able to outperform classical actor-only approaches on several benchmarks.
 

### 2018-02-06 svrpg.tex
Changes to the abstract and the first two sections

--- 2018-02-06 svrpg.tex
@@
 \begin{abstract}
-        We present and analyse stochastic variance-reduced gradient (SVRG) in the context of Markov Decision Processes (MDPs).
-        We first show how to account for I) non-concativity of the objective function; II) approximations in the full gradient computation; and III) non-stationarity of the sampling process.
-        These characteristics make the application of SVRG to MDPs challenging.
-        We propose SVRPG, a stochastic variance-reduced policy gradient algorithm that leverages on importance weight to preserve the unbiasedness of the gradient estimate.
-        We further analyse its convergence guarantees under common assumptions on the MDP.
-        Finally, we investigate practical variants of SVRPG and we empirically evaluate them on continuous MDPs.
+        In this paper, we propose a stochastic variance-reduced version of policy gradient for solving Markov Decision Processes (MDPs).
+        Stochastic variance-reduced gradient (SVRG) methods have proven to be very successful in supervised learning.
+        However, their adaptation to policy gradient is not straightforward and needs to account for I) a non-concave objective function; II) approximations in the full gradient computation; and III) a non-stationary sampling process.
+        The result is SVRPG, a stochastic variance-reduced policy gradient algorithm that leverages on importance weight to preserve the unbiasedness of the gradient estimate.
+        We further analyse its convergence guarantees under standard assumptions on the MDP.
+        Finally, we suggest practical variants of SVRPG, and we empirically evaluate them on continuous MDPs.
 \end{abstract}
@@
 %Similarities between RL and SL
-On a very general level, machine learning addresses the problem of an agent that must select the right actions to solve a task.
+On a very general level, artificial intelligence addresses the problem of an agent that must select the right actions to solve a task.
 The approach of Reinforcement Learning~\citep{sutton1998reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal.
@@
 However, the two approaches share many challenges and tools.
-The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or a policy of behaviour directly.
-Moreover, when the tasks are characterized by large or continuous state-action spaces, RL needs to borrow the powerful function approximators that are the main subject of study of SL, such as neural networks.
+The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or directly a policy defining agent's behaviour.
+Furthermore, when the tasks are characterized by large or continuous state-action spaces, RL needs the powerful function approximators (\eg neural networks) that are the main subject of study of SL.
 % SL needs SVRG so RL
 In a typical SL setting, a performance function $J(\vtheta)$ has to be optimized \wrt to model parameters $\vtheta$.
-The set of data that are available for training is often a subset of all the cases of interest, which may even be infinite, leading to optimization of finite sums to approximate expected performance over an unknown data distribution.
-When generalization to the complete dataset is not taken into consideration we talk about Empirical Risk Minization (ERM).
+The set of data that are available for training is often a subset of all the cases of interest, which may even be infinite, leading to optimization of finite sums that approximate the expected performance over an unknown data distribution.
+When generalization to the complete dataset is not taken into consideration, we talk about Empirical Risk Minimization (ERM).
 Even in this case, stochastic optimization is often used for reasons of efficiency.
@@
 However, the sub-sampling of data is a source of variance that can potentially compromise convergence, so that per-iteration efficiency and convergence rate must be traded off with proper handling of meta-parameters.
-Variance-reduced gradient algorithms such as SAG \cite{roux2012stochastic}, SVRG \cite{johnson2013accelerating} and SAGA \cite{defazio2014saga} offer better ways of addressing this trade-off, with significant results both in theory and practice.
+Variance-reduced gradient algorithms such as SAG \cite{roux2012stochastic}, SVRG \cite{johnson2013accelerating} and SAGA \cite{defazio2014saga} offer better ways of solving this trade-off, with significant results both in theory and practice.
 Although designed explicitly for ERM, these algorithms address a problem that affects more general machine learning problems.
@@
 In RL, stochastic optimization is rarely a matter of choice, since data must be actively sampled by interacting with an initially unknown environment.
-In this scenario, containing the variance of the estimates is a problem that cannot be avoided, which makes variance-reduced algorithms very interesting.
+In this scenario, limiting the variance of the estimates is a necessity that cannot be avoided, which makes variance-reduced algorithms very interesting.
 % Potential advantages of SVRG
 Among RL approaches, policy gradient \cite{sutton2000policy} is the one that bears the closest similarity to SL solutions.
-The fundamental principle of this methods is to optimize a parametric policy by means of stochastic gradient ascent.
+The fundamental principle of these methods is to optimize a parametric policy through stochastic gradient ascent.
 Compared to other applications of SG, the cost of collecting samples can be very high since it requires to interact with the environment.
@@
 First, in SL the objective can often be designed to be strongly concave (we aim to maximize).
-This is not possible in RL, so we have to deal with non-concavities.
-Then, as mentioned, the dataset is not initially available and may even be infinite, which  makes approximations unavoidable.
-This also rules out SAG and SAGA because of their storage requirements, which leaves SVRG as the most promising choice.
-Finally, the distribution used to sample data is not under direct control of the algorithm designer, but is a function of policy parameters that changes over time as the policy is optimized, which is a form of non-stationarity.
+This is not the case for RL, so we have to deal with non-concave objective functions.
+Then, as mentioned before, the dataset is not initially available and may even be infinite, which makes approximations unavoidable.
+This rules out SAG and SAGA because of their storage requirements, which leaves SVRG as the most promising choice.
+Finally, the distribution used to sample data is not under direct control of the algorithm designer, but it is a function of policy parameters that changes over time as the policy is optimized, which is a form of non-stationarity.
 % Related work?
-SVRG has been used in RL as a efficient technique for optimizing the per-iteration problem in Trust-Region Policy Optimization~\citep{xu2017svrgtrpo} or for policy evaluation~\citep{du2017svrgpe}.
-In both the cases, the optimization problems faced resemble the SL scenario and are not characterized by all the previously mentioned issues.
+SVRG has been used in RL as an efficient technique for optimizing the per-iteration problem in Trust-Region Policy Optimization~\citep{xu2017svrgtrpo} or for policy evaluation~\citep{du2017svrgpe}.
+In both the cases, the optimization problems faced resemble the SL scenario and are not affected by all the previously mentioned issues.
 
 %Paper structure
-After providing a background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} we address the aforementioned difficulties to design SVRPG, a variant of SVRG for the policy gradient framework.
+After providing background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} we address, as mentioned above, the difficulties in designing SVRPG, a variant of SVRG for the policy gradient framework.
 In Section \ref{sec:conv} we provide convergence guarantees for our algorithm which also offers insights on the challenges of variance reduction in RL.
-In Section \ref{sec:stopping} we suggests how to set the meta-parameters of SVRPG, while in Section \ref{sec:prac} we discuss some practical variants of the algorithm.
+In Section \ref{sec:stopping} we suggest how to set the meta-parameters of SVRPG, while in Section \ref{sec:prac} we discuss some practical variants of the algorithm.
 Finally, in Section \ref{sec:exp} we empirically evaluate the performance of our method on popular continuous RL tasks.
@@
 \section{Preliminaries}\label{sec:pre}
-We start providing the essential background on policy gradient methods and stochastic variance-reduced gradient methods for finite-sum optimization.
+In this section, we provide the essential background on policy gradient methods and stochastic variance-reduced gradient methods for finite-sum optimization.
 %In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.
@@
 % In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
-We consider episodic MDPs with effective horizon $H$.\footnote{The episode duration is a random variable but the optimal policy is able to reach the target state (\ie absorbing state) in at most $H$ steps.
+We consider episodic MDPs with effective horizon $H$.\footnote{The episode duration is a random variable, but the optimal policy can reach the target state (\ie absorbing state) in at most $H$ steps.
 This has not to be confused with a finite horizon problem where the optimal policy is non-stationary.} In this setting, we can limit our attention to trajectories of length $H$.
@@
 We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
-The search for a locally optimal policy is performed by means of gradient ascent, where the policy gradient
+The search for a locally optimal policy is performed through gradient ascent, where the policy gradient
 %of $J(\vtheta)$ \wrt the policy parameters 
@@
 \begin{align} \label{E:policygradient}
-        \gradJ{\vtheta} = \EVV[\tau \sim \pol]{\score{\vtheta}{\tau}\Reward(\tau)}
+        \gradJ{\vtheta} = \EVV[\tau \sim \pol]{\score{\vtheta}{\tau}\Reward(\tau)}.
         %\int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
@@
         \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} g(\tau_i|\vtheta), \quad \tau_i \in \mathcal{D}_N^k.
-\end{align}
+\end{align}~\todomarcout{chi e' g?}
 The most widespread gradient estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
-Although REINFORCE definition is simpler, G(PO)MDP is often preferred due to its lower variance.
-We refer the reader to App.~\ref{A:gradient_estimators} for details.
-
-The main limitation of plain policy gradient has been the high variance of these estimators.
-The naive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples, \ie by interacting with the environment.
+Although REINFORCE definition is simpler than the G(PO)MDP one, the latter is often preferred due to its lower variance.
+We refer the reader to Appendix~\ref{A:gradient_estimators} for details.
+
+The main limitation of plain policy gradient is the high variance of these estimators.
+The na\"ive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples, \ie by interacting with the environment.
 For this reason, literature has focused on the introduction of baselines aiming to reduce the variance~\citep[\eg][]{Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.\footnote{A baseline is a function $f(s,a)$ such that $\EVV[a \sim \pol]{f(s,a)} = 0$, for any $s\in\Sspace$.}
 There has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
-Although these techniques has been mainly derived for finite-sum problems, we will show in Sec.~\ref{sec:alg} how they can be used in RL.
+Although these techniques have been mainly derived for finite-sum problems, we will show in Sec.~\ref{sec:alg} how they can be used in RL.
 The next section has the aim to describe variance reduction techniques for finite-sum problems. In particular, we will present the SVRG algorithm that is at the core of this work.
@@
 %% \end{align*}
-A common requirement is that $z$ must be smooth and {\color{red} concave} in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
+A common requirement is that $z$ must be smooth and concave in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
 Under this hypothesis, full gradient (FG) ascent~\citep{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $\mathcal{O}(\nicefrac{N}{\epsilon})$, \ie it requires $k = \mathcal{O}(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.
@@
 The resulting IFO complexity is $\mathcal{O}(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$~\cite{nemirovskii1983problem,reddi2016stochastic}.
-Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga} and Finito~\cite{defazio2014finito},
+Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga}, and Finito~\cite{defazio2014finito}.
 The common idea is to reuse past gradient computations to reduce the variance of the current estimate.
-In particular, Stochastic Variance-Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which can become a problem when deep and/or wide neural networks are employed.
+In particular, Stochastic Variance-Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which is a significant advantage when deep and/or wide neural networks are employed.
 
 The idea of SVRG is to alternate full and stochastic gradient updates. 
-Each $m = \mathcal{O}(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i z(x_i|\widetilde{\vtheta})$.
+Each $m = \mathcal{O}(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i \nabla z(x_i|\widetilde{\vtheta})$.
 Between snapshots, the parameter is updated with a gradient estimate corrected using stochastic gradient, for any $t \in \{0,\ldots,m-1\}$
@@
 
-More recently, some extensions of variance reduction algorithms to the non-convex objectives have been proposed~\citep[\eg][]{reddi2016stochastic,allen2016variance,reddi2016stochastic,reddi2016fast}.
+More recently, some extensions of variance reduction algorithms to the non-concave objectives have been proposed~\citep[\eg][]{reddi2016stochastic,allen2016variance,reddi2016stochastic,reddi2016fast}.
 In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[2]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[2]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
-Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the \textcolor{red}{concave} case~\citep{nesterov2013introductory,ghadimi2013stochastic}.
-Also in this case, SVRG yields an advantageous trade-off, with an IFO complexity of $\mathcal{O}(N + \nicefrac{N^{\nicefrac{2}{3}}}{N})$~\citep{reddi2016stochastic}.
+Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the concave case~\citep{nesterov2013introductory,ghadimi2013stochastic}.
+Also in this case, SVRG yields an advantageous trade-off, with an IFO complexity of $\mathcal{O}(N + \nicefrac{N^{\nicefrac{2}{3}}}{\epsilon})$~\citep{reddi2016stochastic}.
 The only additional requirement is to select $\vtheta^*$ uniformly at random among all the $\vtheta_k$ instead of simply setting it to the final value.
-Pseudocode of SVRG for both the convex and the non-convex case is provided in Alg.~\ref{alg:svrg}.
+Pseudocode of SVRG for both the concave and the non-concave case is provided in Algorithm~\ref{alg:svrg}.
 
@@
 		\FOR{$s=0$ {\bfseries to} $S-1$}
-        \STATE $\wt{\mu} = f(\wt{\vtheta}^s)$
+        \STATE $\wt{\mu} = \nabla f(\wt{\vtheta}^s)$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
@@
 \section{Convergence Guarantees of SVRPG}\label{sec:conv}
-In this section, we state the convergence guarantees for SVRPG {\color{red}with REINFORCE estimator}.\footnote{\color{red}Similar results can be derived for G(PO)MDP, see App.~\ref{app:}.}
+In this section, we state the convergence guarantees for SVRPG with REINFORCE estimator.\footnote{Similar results can be derived for G(PO)MDP, see App.~\ref{app:}.}
 We mainly leverage on the recent analysis of \textcolor{red}{non-concave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
@@
 	\begin{restatable}[On policy derivatives]{assumption}{boundedscore}\label{ass:bounded_score}
-		For each state-action pair $(s,a)$, value of $\vtheta$ and dimensions $i,j$ there are positive constants $G,F<\infty$ such that:
+		For each trajectory $\tau$, value of $\vtheta$ and dimensions $i,j$ there are positive constants $G,F<\infty$ such that:
 \[
-		\left|\nabla_{\theta_i}\log\pi_{\vtheta}(a\vert s)\right| \leq \GRADLOG, \qquad
-        \left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(a \vert s)\right| \leq \HESSLOG.
+		\left|\nabla_{\theta_i}\log\pi_{\vtheta}(\tau)\right| \leq \GRADLOG, \qquad
+        \left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(\tau)\right| \leq \HESSLOG.
 \]
@@
 We can now state the convergence guarantees for SVRPG.
-\begin{restatable}[Convergence of the {\color{red}REINFORCE-}SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
-{\color{red}Consider the REINFORCE estimator in Eq.~\ref{E:policygradient.estimate}.}
+\begin{restatable}[Convergence of the REINFORCE-SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
+Consider the REINFORCE estimator in Eq.~\ref{E:policygradient.estimate}.
 Under Ass.~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Alg.~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
@@
 where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s,t}$ with the sub-iterations.
-
+{\color{red} Note that the ADAM estimator used in the sub-iterations is reset at each snapshot with the FG estimate (see App.~\ref{app:adam} for details).}
 \todopir{Add ADAM definition and explain in more details}
@@
 \subsection{Normalized Importance Sampling}\label{sec:prac}
-As we already mentioned in section \ref{sec:stopping} importance weights means a lot of variance to deal with.
-Therefore, in oreder to cope with this issue the literature propose the Self-normalization techniques \cite{precup2000eligibility}.
-This technique offers a variance reduction over the importance weights in exchange of a bias introduction.
-Hence, if the bias introduced by this is smaller than the variance reduction it causes, it is convenient to use it in practice.
-We expirienced some improvements using it (see section \ref{sec:exp} for the sperimental point of view).
+As we already mentioned in section \ref{sec:stopping} importance weights means a lot of variance to deal with. Therefore, in oreder to cope with this issue the literature propose two main tecniques:
+\begin{enumerate}
+\item Self-normalization (?ciatazione bib?).
+\item Importance weights trucation (?citazione bib?).
+\end{enumerate}
+Both of them offer a variance reduction over the importance weights in exchange of a bias introduction.
+Hence, if the bias introduced by those techniques is smaller than the variance reduction they cause, it is convenient to use them in practice.
+So far, sperimentally, we have not seen any advantage in using importance weights truncation, hence we discard this technique for the previously mentioned purpose.
+Whereas for what concern self-normalization we expirienced some improvements using it (see section \ref{sec:exp} for the sperimental point of view).
 Self-normalization is easy to use, but depends on the chosen reinforcement learning algorithm.
@@
 \begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
-At convergence, the variance of the SVRPG estimator is equal to the variance of the gradient estimator, regardless of the mini-batch size $B$:
+At convergence, the variance of the SVRPG estimator is equal to the variance of the REINFORCE estimator, regardless of the mini-batch size $B$:
 \[
@@
 &= \int_{\Tspace}\pol(\tau)\score{\theta}{\tau}\score{\theta}{\tau}^T\Reward(\tau)\de \tau + \int_{\Tspace}\pol(\tau)\Dij{}{\theta}\log\pol(\tau)\Reward(\tau)\de \tau \nonumber\\
-&\leq \sup_{\tau} \left\{\left|\Reward(\tau)\right|\right\} \left(H^2\GRADLOG^2+H\HESSLOG\right) \label{eq:0}\\
-&= \frac{1-\gamma^H}{1-\gamma}RH\left(H\GRADLOG^2+\HESSLOG\right),\nonumber
+&\leq \sup_{\tau} \left\{\left|\Reward(\tau)\right|\right\} \left(\GRADLOG^2+\HESSLOG\right) \label{eq:0}\\
+&= \frac{R}{1-\gamma}\left(\GRADLOG^2+\HESSLOG\right),\nonumber
 \end{align}
@@
 
-\begin{restatable}[]{lemma}{bounded-g}\label{lemma:gbound}
-Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, for every $\tau\in\Tspace$ and $\vtheta\in\Theta$, there is a positive constant $\Gamma<\infty$ such that:
-\[
-	\norm[]{g(\tau\vert\vtheta)}^2 \leq \Gamma.
-\]
-\end{restatable}
-\begin{proof}
-For REINFORCE we have, from Assumption \ref{ass:bounded_score}:
-\begin{align*}
-	\norm[]{g(\tau\vert\vtheta)}^2 &=
-	\norm[]{\score{\vtheta}{\tau}R(\tau)}^2 \\
-	&=\norm[]{\left(\sum_{t=0}^{H-1}\score{\vtheta}{a_t\vert s_t}\right)\left(\sum_{t=0}^{H-1}\gamma^t r_t\right)}^2\leq H^2G^2\frac{(1-\gamma^H)^2}{(1-\gamma)^2}R^2\dim(\vTheta) \coloneqq \Gamma
-\end{align*}
-For G(PO)MDP, we do not have a compact expression for $g$, but since it is derived from REINFORCE by neglecting some terms of the kind $\score{\vtheta}{a_t\vert s_t}\gamma^t r_t$ \cite{baxter2001infinite}\cite{peters2008reinforcement}, the above bound still holds.
-\end{proof}
 %Lemma 2
@@
 \nonumber 
-+\frac{\Gamma\VARIS}{B}
++\frac{G^2R^2W}{(1-\gamma)^2B}
 \]
@@
 	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\Gamma\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \label{eq:4}\\
+	+\GRADLOG^2\frac{R^2}{(1-\gamma)^2}\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \label{eq:4}\\
 	%
@@
 	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\frac{\Gamma}{B^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
+	+\GRADLOG^2\frac{R^2}{(1-\gamma)^2b^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
 	%
@@
 	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\frac{\Gamma\VARIS}{B}, \label{eq:5}
+	+\GRADLOG^2\frac{R^2\VARRF}{(1-\gamma)^2B}, \label{eq:5}
 \end{align}
 where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (\ref{eq:3}) is from L-smoothness,
-(\ref{eq:4}) is from Lemma \ref{lemma:gbound}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
+(\ref{eq:4}) is from Assumption \ref{ass:bounded_score} 
+{\todomatout{We could generalize to GPOMDP by defining some more constants}\color{red}and the definition of REINFORCE}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
 \end{proof}
@@
 	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_tV}{N\Psi_t}
-	+\frac{f_tW}{B\Psi_t},
+	+\frac{f_tW}{b\Psi_t},
 \]
@@
 	&c_{m-1} = 0, \\
-	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\frac{\alpha_t^2L^2}{B}\right) + \alpha_tL^2+\frac{\alpha_t^2L^3}{2B}, \\
+	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\alpha_t^2\right) + \alpha_t\frac{L^2}{b}+\frac{\alpha_t^2L^3}{2}, \\
 	&\Psi_t = \alpha_t\left(1-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
 	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right), \\
-	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2}.
+	&f_t = \alpha_t^2\frac{G^2R^2(L+2c_{t+1})}{2(1-\gamma)^2}.
 \end{align*}
@@
 	\right.\nonumber\\
-	&\left.\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-	+\frac{\Gamma W}{B}\right)
+	&\left.\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	+G^2\frac{R^2}{(1-\gamma)^2b}W\right)
 	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
@@
 	%
-	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta+\frac{\alpha_t^2L^2}{B}\right)+\alpha_tL^2+\frac{\alpha_t^2L^3}{2B}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta+\alpha_t^2\right)+\alpha_t\frac{L^2}{b}+\frac{\alpha_t^2L^3}{2}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad
@@
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Varts[]{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
-	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})\Gamma\VARIS}{2B} \nonumber\\
+	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2\VARIS}{2(1-\gamma)^2b} \nonumber\\
 	%
@@
 \end{align}
-where (\ref{eq:17}) is from (\ref{eq:13}), (\ref{eq:18}) is from (\ref{eq:16}), (\ref{eq:19}) is from Lemma \ref{lemma:aux2}, and (\ref{eq:20}) is from Assumption \ref{ass:REINFORCE}.
+where (\ref{eq:17}) is from (\ref{eq:13}), (\ref{eq:18}) is from (\ref{eq:16}), (\ref{eq:19}) is from Lemma \ref{lemma:aux0}, and (\ref{eq:20}) is from Assumption \ref{ass:REINFORCE}.
 To complete the proof, besides rearranging terms, we have to ensure that $\Psi_t>0$ for each $t$. This gives the constraints on $\alpha_t$ and $\beta_t$.
@@
 The episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
-	\item \emph{Mujoco Swimmer}: 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
+	\item \emph{Mujoco Swimmer}: 12-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
 2-dimensional action space: the two momentums applied on actuated joints.
@@
 	\hline
-	State space dimension  & 4 & 13 & 20 \\
+	State space dimension  & 4 & 8 & 20 \\
 	Control space dimension & 1 & 2 & 6 \\
@@
 	Discount $(\gamma)$& 0.99 & 0.995 & 0.99 \\
-	Total trajectories& 10000 & 20000 & 50000 \\
+	Total trajectories& 10000 & 10000 & 10000 \\
 	\hline  

### 2018-02-06 svrpg.tex

--- 2018-02-06 svrpg.tex
@@
 Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling.
-The resulting IFO complexity is $\mathcal{O}(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$~\cite{nemirovskii1983problem,reddi2016stochastic}.
-\todomat{We are not interested in IFO complexity, report convergence rate instead}
-Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga} and Finito~\cite{defazio2014finito},
+The resulting IFO complexity is $\mathcal{O}(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$~\cite{nemirovskii1983problem,reddi2016stochastic}.\todomat{We are not interested in IFO complexity, report convergence rate instead}
+Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga}, and Finito~\cite{defazio2014finito}.
 The common idea is to reuse past gradient computations to reduce the variance of the current estimate.
-In particular, Stochastic Variance-Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which can become a problem when deep and/or wide neural networks are employed.
+In particular, Stochastic Variance-Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which is a significant advantage when deep and/or wide neural networks are employed.
 
 The idea of SVRG is to alternate full and stochastic gradient updates. 
-Each $m = \mathcal{O}(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i z(x_i|\widetilde{\vtheta})$.
+Each $m = \mathcal{O}(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i \nabla z(x_i|\widetilde{\vtheta})$.
 Between snapshots, the parameter is updated with a gradient estimate corrected using stochastic gradient, for any $t \in \{0,\ldots,m-1\}$

### 2018-02-07 svrpg.tex
Updated definition of g

@@
 \begin{align} \label{E:policygradient.estimate}
-        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} g(\tau_i|\vtheta), \quad \tau_i \in \mathcal{D}_N^k.
-\end{align}~\todomarcout{chi e' g?}
+        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} g(\tau_i|\vtheta), \quad \tau_i \in \mathcal{D}_N^k,
+\end{align}
+where $g(\tau_i|\vtheta)$ is an estimate of $\score{\vtheta}{\tau_i}\Reward(\tau_i)$.
 The most widespread gradient estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
 Although REINFORCE definition is simpler than the G(PO)MDP one, the latter is often preferred due to its lower variance.
-We refer the reader to Appendix~\ref{A:gradient_estimators} for details.
+We refer the reader to Appendix~\ref{A:gradient_estimators} for details and a formal definition of $g$.
 
@@
 \section{Policy Gradient Estimators} \label{A:gradient_estimators}
-
+We present a brief overview of two most widespread gradient estimators (REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}) both in on-policy and off-policy settings.
 \subsection{On-policy}

### 2018-02-07 svrpg.tex
Updated definition of g

@@
 At each iteration $k >0$, a batch $\mathcal{D}_N^k = \{\tau_i\}_{i=0}^N$ of $N>0$ trajectories is collected using policy $\pi_{\vtheta_k}$.
-The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $\mathcal{D}_N^k$, \ie
+The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $\mathcal{D}_N^k$.
+The most common policy gradient estimators can be expressed as follows
 \begin{align} \label{E:policygradient.estimate}
@@
 \end{align}
-where $g(\tau_i|\vtheta)$ is an estimate of $\score{\vtheta}{\tau_i}\Reward(\tau_i)$.
-The most widespread gradient estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
+where $g(\tau_i|\vtheta)$ is an estimate of $\score{\vtheta}{\tau}\Reward(\tau)$, \eg REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
 Although REINFORCE definition is simpler than the G(PO)MDP one, the latter is often preferred due to its lower variance.

### 2018-02-07 svrpg.tex
Added definition of grad estimators (on/off)

@@
 The na\"ive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples, \ie by interacting with the environment.
-For this reason, literature has focused on the introduction of baselines aiming to reduce the variance~\citep[\eg][]{Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.\footnote{A baseline is a function $f(s,a)$ such that $\EVV[a \sim \pol]{f(s,a)} = 0$, for any $s\in\Sspace$.}
+For this reason, literature has focused on the introduction of baselines aiming to reduce the variance~\citep[\eg][]{Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.\footnote{A baseline is a function $f(s,a)$ such that $\EVV[a \sim \pol]{f(s,a)} = 0$, for any $s\in\Sspace$.
+\todopir{check the definition of baseline}}
 There has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
@@
 We present a brief overview of two most widespread gradient estimators (REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}) both in on-policy and off-policy settings.
-\subsection{On-policy}
+Let $\tau = \{\langle s_t,a_t \rangle\}_{t=0}^{H}= \{z_t\}_{t=0}^{H} = z_{0:H}$ is a $(H+1)$-steps trajectory.
+Given that $\tau$ depends on the MDP $M=\{\mathcal{S},\mathcal{A}, p, r, \gamma, \rho\}$ and the actual policy $\pi$, the trajectory is
+said drawn from density distribution $\Delta(\tau|\pi,M)$ defined as:
+\[
+        \Delta(\tau|\pi,M) = \rho(s_0) \pi(z_0) \prod_{k=1}^{H} p(s_k|z_{k-1})\pi(z_k).
+\]
+We can now recall the definition of policy performance 
+\[
+        J(\pol) = \EVV[\tau \sim \Delta(\cdot|\pol)]{\Reward(\tau)|M},
+\]
+where $\Reward(\tau) = \sum_{t=0}^{H}\gamma^t r(z_t)$.
+The policy gradient $\nabla J(\vtheta)$ is
+\begin{equation}\label{E:onpolicygradient}
+        \gradJ{\vtheta} = \EVV[\tau \sim \Delta(\cdot|\pol)]{\nabla \log\Delta(\tau|\vtheta)\Reward(\tau)} = \EVV[\tau \sim \Delta(\cdot|\pol)]{\sum_{j=0}^{H} \gamma^j \mathcal{R}(z_j) \sum_{i=0}^{H} \nabla \log\pi(z_i)}.
+\end{equation}
+
+\textbf{On-policy setting.}
+Consider a policy $\pol$ and let $\mathcal{D}_N = \{\tau_i\}_{i=1}^N$ be a dataset collected using policy $\pol$.
 The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
 \begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
-\end{align*}
-where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline provided $\EVV[a \sim \pol]{b(s,a)} = 0$, for any $s\in\Sspace$.
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}
+\underbracket{
+\left(\sum_{h=0}^{H}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\sum_{h=0}^{H}\gamma^h r(s_h^n,a_h^n) - b\right)
+}_{g(\tau_n|\vtheta)}
+,
+\end{align*}
+where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline provided $\EVV[a \sim \pol]{b(s,a)} = 0$, for any $s\in\Sspace$.\todopir{check definition of baseline}
 The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance \cite{zhao2011analysis} while preserving the unbiasedness:
 \begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\gamma^h r_h^n - b\right).
-\end{align*}
-
-\subsection{Off-Policy}
+\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}
+\underbracket{
+\sum_{h=0}^{H}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\gamma^h r(s_h^n,a_h^n) - b\right)
+}_{g(\tau_n|\vtheta)}.
+\end{align*}
+G(PO)MDP can be seen as a more efficient implementation
+of the REINFORCE algorithm. In fact the latter does not perform
+an optimal credit assignment, since it ignores that the reward at time $t$ does
+not depend on the action performed after time $t$. G(PO)MDP overcomes this
+issue taking into account the causality of rewards in the REINFORCE definition
+of policy gradient.
+
+\textbf{Off-Policy setting.}
+In off-policy setting two policies, called behavioural $\pi^B$ and target $\pi^T$, are involved.
+The first is used to select actions for the interaction with the system, while the second is used to evaluate the agent performance and it is improved in each update.
+Suppose now that we aim to estimate the performance of the target policy $\pi^T$ but we have samples collected using policy $\pi^B$.
+We can use importance weight correction to correct the shift in the distribution and obtain an unbiased estimate of $J(\pi^T)$:
+\[
+        J(\pi^T) = \EVV[\tau \sim \Delta(\cdot|\pi^T)]{\mathcal{R}(\tau)} = \EVV[\tau \sim \Delta(\cdot|\pi^B)]{\omega(\tau) \mathcal{R}(\tau)}
+\]
+where $\omega(\tau) = \frac{\Delta(\tau,\pi^T)}{\Delta(\tau|\pi^B)} = \omega(z_{0:H}) = \prod_{w=0}^{H} \omega(z_w)$ and $\omega(z_w) = \frac{\pi^T(a_w|s_w)}{\pi^B(a_w|s_w)}$
+
+The definition of the off-policy version of~\eqref{E:onpolicygradient} is~\citep[\eg][]{jurvcivcek2012reinforcement}
+\begin{equation}\label{E:offpolicygradient}
+        \nabla J(\pi^T) = \EVV[\tau \sim \Delta(\cdot|\pi^B)]{\omega(\tau) \nabla \log \Delta(\tau|\pi^T) \mathcal{R}(\tau)}
+\end{equation}
+For $\omega(\tau)$ being well defined the behavioural policy should have non-zero probability of selecting any action in every state \ie $\pi^B(a|s) > 0$ for any $(s,a)\in \mathcal{S} \times \mathcal{A}$.
+
+The off-policy version of REINFORCE is easily obtained by taking the empirical average of~\eqref{E:offpolicygradient}:
+\[
+        \nabla J(\pi^T) = \frac{1}{N} \sum_{n=1}^{N} \omega(\tau^n)
+        \left(\sum_{h=0}^{H}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\sum_{h=0}^{H}\gamma^h r(s_h^n,a_h^n)\right).
+\]
+
+The G(PO)MDP off-policy estimator is defined as follows
+\[
+        \nabla J(\pi^T) = \frac{1}{N} \sum_{n=1}^{N}
+        \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k)\right) \gamma^h r(z_h^n) \omega(z_{0:h})
+\]
+\todopir{Add proof of Def 2 from Luca's thesis}
+
 \todopir{Make ref to Luca's thesis and describe in brief the derivations}

### 2018-02-07 svrpg.tex
Updated

@@
 \end{align*}
-where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s,t}$ with the sub-iterations.
-
+where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s,t}$ with the sub-iterations (see App.~\ref{app:practicalsvrpg} for details).
 \todopir{Add ADAM definition and explain in more details}
@@
 
+\section{Practical SVRPG versions}\label{app:practicalsvrpg}
+\begin{algorithm}[h]
+        \begin{algorithmic}
+        \STATE \textbf{Input:} A gradient estimate $g_t$ and parameters $\beta_1$, $\beta_2$, $\epsilon$ and $\alpha$.
+        \STATE $m_t = \beta_1 m_{t-1} + (1 - \beta_1) g_t$
+        \STATE $v_t = \beta_2 v_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
+        \STATE $\hat{m}_t = \dfrac{m_t}{1 - \beta^t_1}$
+        \STATE $\hat{v}_t = \dfrac{v_t}{1 - \beta^t_2}$
+        \STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{v}_t} + \epsilon} \hat{m}_t$
+        \STATE \textbf{Return:} The increment $\Delta(g_t)$  of the parameters.
+\end{algorithmic}
+ \caption{
+ \label{A:adam}
+ Adam}
+\end{algorithm}
+
 \section{Experimental Details}\label{app:exp}

### 2018-02-07 svrpg.tex
Updated
made changes to section 1,3,4; fixed proof

--- 2018-02-07 svrpg.tex
@@
 A common requirement is that $z$ must be smooth and concave in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
-Under this hypothesis, full gradient (FG) ascent~\citep{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $\mathcal{O}(\nicefrac{N}{\epsilon})$, \ie it requires $k = \mathcal{O}(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.
+Under this hypothesis, full gradient (FG) ascent ~\citep{cauchy1847methode} with a constant step size achieves a convergence rate of $\mathcal{O}(\nicefrac{1}{\epsilon})$, \ie it requires $T = \mathcal{O}(\nicefrac{1}{\epsilon})$ iterations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.
 % \footnote{In the following we will omit the subscript for the $\ell_2$-norm.}
-This can be too expensive for large values of $N$.
+However, each iteration requires $N$ gradient computations, which can be too expensive for large values of $N$.
 Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling.
-The resulting IFO complexity is $\mathcal{O}(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$~\cite{nemirovskii1983problem,reddi2016stochastic}.\todomat{We are not interested in IFO complexity, report convergence rate instead}
+As a consequence, the lower per-iteration cost is paid with a worse convergence rate of $\mathcal{O}(\nicefrac{1}{\epsilon^2})$ \cite{nemirovskii1983problem}.
 Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga}, and Finito~\cite{defazio2014finito}.
@@
 Note that $t=0$ corresponds to a FG step (\ie $\blacktriangledown f(\vtheta_0) = \nabla f(\wt{\vtheta})$) since $\vtheta_0 := \wt{\vtheta}$.
-The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient.
-The resulting IFO complexity is $\mathcal{O}(N+\nicefrac{\sqrt{N}}{\epsilon})$ \cite{reddi2016stochastic}.
+The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, achieving a convergence rate of $O(\nicefrac{1}{\epsilon})$ without resorting to plain full gradient.
 
@@
 In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[2]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[2]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
-Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the concave case~\citep{nesterov2013introductory,ghadimi2013stochastic}.
-Also in this case, SVRG yields an advantageous trade-off, with an IFO complexity of $\mathcal{O}(N + \nicefrac{N^{\nicefrac{2}{3}}}{\epsilon})$~\citep{reddi2016stochastic}.
+Under this hypothesis, the convergence rate of full gradient  and of stochastic gradient ascent are the same as in the concave case~\citep{nesterov2013introductory,ghadimi2013stochastic}.
+Also in this case, SVRG achieves the $O(\nicefrac{1}{\epsilon})$ rate \cite{reddi2016stochastic}.
 The only additional requirement is to select $\vtheta^*$ uniformly at random among all the $\vtheta_k$ instead of simply setting it to the final value.
 Pseudocode of SVRG for both the concave and the non-concave case is provided in Algorithm~\ref{alg:svrg}.
+\todomat{check citations for convergence rates}
 
@@
 \end{algorithm}
-\todomat{Mention the fact that we return last parameter in practice}
 
@@
 The resulting SVRPG policy optimization method is detailed in Alg.~\ref{alg:svrpg}.\footnote{Note that when $t=0$ we can avoid to sample trajectories and instead perform directly the policy update using the FG.
-For sake of clarity, we have not explicitly differentiate this step from the standard update rule in Alg.~\ref{alg:svrpg}.}
+For sake of clarity, we have not explicitly differentiate this step from the standard update rule in Alg.~\ref{alg:svrpg}.} Note that, in our implementation, we simply return the last value of the parameter $\tilde{\vtheta}^S$ instead of a random one, as is often done in practice.
+\todomat{add reference}
 
@@
 		\frac{\zeta}{N}
-		+\frac{\xi}{B}
-\end{align*}
+		+\frac{\xi}{B},
+\end{align*}
+where $\psi,\zeta,\xi$ depend only on $\GRADLOG,\HESSLOG,\VARRF,\VARIS,\alpha$ and $m$.
 \end{restatable}
@@
 In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by FG approximation and importance sampling, respectively.
-Since the optimal configuration may be task-dependent, we leave this choice to the empirical analysis of Sec.~\ref{sec:exp}.
-
+Once the last two terms can be neglected, the number of trajectories needed to achieve $\norm[2]{\gradJ{\vtheta}}^2\leq\epsilon$ is $O(\frac{B+\nicefrac{N}{m}}{\epsilon})$.
+Advantage over batch gradient ascent can be achieved with properly selected meta-parameters.
+In the following we propose practical ways to do so.
+
+{\color{red}
 \textbf{G(PO)MDP.} A similar result to the one stated in Thm~\ref{theo:convergence} can be derived for the G(PO)MDP estimator. A sketch of the proof is reported in~\ref{app:gpomdpproof}.
@@
 For this reason we will use G(PO)MDP for the experiments (see Sec.~\ref{sec:exp}).
-\todopir{sketch of proof for GPOMDP in appendix, probably can be unified with the one of REINFORCE}
+\todopir{sketch of proof for GPOMDP in appendix, probably can be unified with the one of REINFORCE}}
 
@@
 \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
++\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 +\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
@@
 	&c_{m-1} = 0, \\
-	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\frac{\alpha_t^2L^2}{B}\right) + \alpha_tL^2+\frac{\alpha_t^2L^3}{2B}, \\
-	&\Psi_t = \alpha_t\left(1-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
-	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right), \\
+	&c_t = c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}, \\
+	&\Psi_t = \alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
+	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right), \\
 	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2}.
@@
 \begin{align*}
-&0\leq\alpha_t < \frac{2(1-\nicefrac{c_{t+1}}{\beta_t})}{L+2c_{t+1}} \\
-&\beta_t > c_{t+1}.
+&0\leq\alpha_t < \frac{2(1-\nicefrac{2c_{t+1}}{\beta_t})}{L+2c_{t+1}} \\
+&\beta_t > 2c_{t+1}.
 \end{align*}
@@
 	&\qquad-
-	\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}, \label{eq:13}
+	\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_t}{2}\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}, \label{eq:13}
 \end{align}
@@
 &\qquad+
-\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
+\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
 %
@@
 &\qquad+
-\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
 %
@@
 &\qquad
-\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
+\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
 \end{align}
@@
 	%
-	&\geq	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\geq	\Ets{J(\vtheta_t^{s+1})+\frac{\alpha_t}{2}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-
-	\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
+	\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
 	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:17}\\
 	%
-	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
+	&\geq \Ets{J(\vtheta_t^{s+1})+\frac{\alpha_t}{2}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
 	-\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
-	&\qquad -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
-	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
-	%
-	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
-	-\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
-	&\qquad -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2} -c_{t+1}\Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	&\qquad -c_{t+1}\Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 	\nonumber\\
@@
 	&\qquad
-	-c_{t+1}\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:18}\\
-	%
-	&= \Ets{J(\vtheta_t^{s+1})} - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta\right)+\alpha_tL^2\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	-c_{t+1}\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -c_{t+1}\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:18}\\
+	%
+	&= \Ets{J(\vtheta_t^{s+1})} - c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad-\alpha_t^2\left(\frac{L}{2}+c_{t+1}\right)\Ets{\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}
-	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	+\frac{\alpha_t}{2}\left(1-\frac{2c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	%
-	&\geq  \Ets{J(\vtheta_t^{s+1})} - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta\right)+\alpha_tL^2\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	&\geq  \Ets{J(\vtheta_t^{s+1})} - c_{t+1}\left(1+\alpha_t\beta+\alpha_t\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad
@@
 	+\frac{\Gamma W}{B}\right)
-	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	+\frac{\alpha_t}{2}\left(1-2\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \label{eq:19}\\
 	%
-	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta+\frac{\alpha_t^2L^2}{B}\right)+\alpha_tL^2+\frac{\alpha_t^2L^3}{2B}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(1+\alpha_t\beta+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad
-	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
+	+\alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})\Gamma\VARIS}{2B} \nonumber\\

### 2018-02-07 svrpg.tex

@@
 			\wt{\mu} + 
-			\nabla g(x_i|\vtheta_t^{s}) -
-			\nabla g(x_i|\wt{\vtheta}^{s})
+			\nabla z(x_i|\vtheta_t^{s}) -
+			\nabla z(x_i|\wt{\vtheta}^{s})
 		$
@@
 
-Concerning importance weighting techniques, RL has made wide use of them for off-policy problems~\citep[\eg]{precup2000eligibility,thomas2015high}.
+Concerning importance weighting techniques, RL has made wide use of them for off-policy problems~\citep[\eg][]{precup2000eligibility,thomas2015high}.
 However, as mentioned before, SVRPG cannot be compared to such methods since it is practically on-policy.

### 2018-02-07 svrpg.tex
improved treatment of convergence rate in Sec. 2.2

@@
 A common requirement is that $z$ must be smooth and concave in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
-Under this hypothesis, full gradient (FG) ascent ~\citep{cauchy1847methode} with a constant step size achieves a convergence rate of $\mathcal{O}(\nicefrac{1}{\epsilon})$, \ie it requires $T = \mathcal{O}(\nicefrac{1}{\epsilon})$ iterations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.
-% \footnote{In the following we will omit the subscript for the $\ell_2$-norm.}
+Under this hypothesis, full gradient (FG) ascent ~\citep{cauchy1847methode} with a constant step size achieves a linear convergence rate \cite{nesterov2013introductory}.
 However, each iteration requires $N$ gradient computations, which can be too expensive for large values of $N$.
 Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling.
-As a consequence, the lower per-iteration cost is paid with a worse convergence rate of $\mathcal{O}(\nicefrac{1}{\epsilon^2})$ \cite{nemirovskii1983problem}.
+As a consequence, the lower per-iteration cost is paid with a worse, sub-linear convergence rate \cite{nemirovskii1983problem}.
 Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga}, and Finito~\cite{defazio2014finito}.
@@
 Note that $t=0$ corresponds to a FG step (\ie $\blacktriangledown f(\vtheta_0) = \nabla f(\wt{\vtheta})$) since $\vtheta_0 := \wt{\vtheta}$.
-The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, achieving a convergence rate of $O(\nicefrac{1}{\epsilon})$ without resorting to plain full gradient.
+The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, achieving a linear convergence rate without resorting to plain full gradient.
 
@@
 In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[2]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[2]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
-Under this hypothesis, the convergence rate of full gradient  and of stochastic gradient ascent are the same as in the concave case~\citep{nesterov2013introductory,ghadimi2013stochastic}.
-Also in this case, SVRG achieves the $O(\nicefrac{1}{\epsilon})$ rate \cite{reddi2016stochastic}.
+Under this hypothesis, the convergence rate of SG is $O(\nicefrac{1}{\sqrt{T}})$ \cite{ghadimi2013stochastic}, \ie $T=O(\nicefrac{1}{\epsilon^2})$ iterations are required to get $\norm[2]{\nabla f(\vtheta)}^2\leq\epsilon$.
+Again, SVRG achieves the same rate as FG \cite{reddi2016stochastic}, which is $O(\frac{1}{T})$ in this case \cite{nesterov2013introductory}.
 The only additional requirement is to select $\vtheta^*$ uniformly at random among all the $\vtheta_k$ instead of simply setting it to the final value.
 Pseudocode of SVRG for both the concave and the non-concave case is provided in Algorithm~\ref{alg:svrg}.
-\todomat{check citations for convergence rates}
 

### 2018-02-07 svrpg.tex
Modified section 5.2, inserted 5.3

@@
 A common requirement is that $z$ must be smooth and concave in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
-Under this hypothesis, full gradient (FG) ascent ~\citep{cauchy1847methode} with a constant step size achieves a linear convergence rate \cite{nesterov2013introductory}.
-However, each iteration requires $N$ gradient computations, which can be too expensive for large values of $N$.
+Under this hypothesis, full gradient (FG) ascent~\citep{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $\mathcal{O}(\nicefrac{N}{\epsilon})$, \ie it requires $k = \mathcal{O}(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[2]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.
+% \footnote{In the following we will omit the subscript for the $\ell_2$-norm.}
+This can be too expensive for large values of $N$.
 Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling.
-As a consequence, the lower per-iteration cost is paid with a worse, sub-linear convergence rate \cite{nemirovskii1983problem}.
+The resulting IFO complexity is $\mathcal{O}(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$~\cite{nemirovskii1983problem,reddi2016stochastic}.\todomat{We are not interested in IFO complexity, report convergence rate instead}
 Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga}, and Finito~\cite{defazio2014finito}.
@@
 Note that $t=0$ corresponds to a FG step (\ie $\blacktriangledown f(\vtheta_0) = \nabla f(\wt{\vtheta})$) since $\vtheta_0 := \wt{\vtheta}$.
-The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, achieving a linear convergence rate without resorting to plain full gradient.
+The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient.
+The resulting IFO complexity is $\mathcal{O}(N+\nicefrac{\sqrt{N}}{\epsilon})$ \cite{reddi2016stochastic}.
 
@@
 In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[2]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[2]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
-Under this hypothesis, the convergence rate of SG is $O(\nicefrac{1}{\sqrt{T}})$ \cite{ghadimi2013stochastic}, \ie $T=O(\nicefrac{1}{\epsilon^2})$ iterations are required to get $\norm[2]{\nabla f(\vtheta)}^2\leq\epsilon$.
-Again, SVRG achieves the same rate as FG \cite{reddi2016stochastic}, which is $O(\frac{1}{T})$ in this case \cite{nesterov2013introductory}.
+Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the concave case~\citep{nesterov2013introductory,ghadimi2013stochastic}.
+Also in this case, SVRG yields an advantageous trade-off, with an IFO complexity of $\mathcal{O}(N + \nicefrac{N^{\nicefrac{2}{3}}}{\epsilon})$~\citep{reddi2016stochastic}.
 The only additional requirement is to select $\vtheta^*$ uniformly at random among all the $\vtheta_k$ instead of simply setting it to the final value.
@@
 			\wt{\mu} + 
-			\nabla z(x_i|\vtheta_t^{s}) -
-			\nabla z(x_i|\wt{\vtheta}^{s})
+			\nabla g(x_i|\vtheta_t^{s}) -
+			\nabla g(x_i|\wt{\vtheta}^{s})
 		$
@@
 The resulting SVRPG policy optimization method is detailed in Alg.~\ref{alg:svrpg}.\footnote{Note that when $t=0$ we can avoid to sample trajectories and instead perform directly the policy update using the FG.
-For sake of clarity, we have not explicitly differentiate this step from the standard update rule in Alg.~\ref{alg:svrpg}.} Note that, in our implementation, we simply return the last value of the parameter $\tilde{\vtheta}^S$ instead of a random one, as is often done in practice.
-\todomat{add reference}
+For sake of clarity, we have not explicitly differentiate this step from the standard update rule in Alg.~\ref{alg:svrpg}.}
 
@@
 		\frac{\zeta}{N}
-		+\frac{\xi}{B},
-\end{align*}
-where $\psi,\zeta,\xi$ depend only on $\GRADLOG,\HESSLOG,\VARRF,\VARIS,\alpha$ and $m$.
+		+\frac{\xi}{B}
+\end{align*}
 \end{restatable}
@@
 In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by FG approximation and importance sampling, respectively.
-Once the last two terms can be neglected, the number of trajectories needed to achieve $\norm[2]{\gradJ{\vtheta}}^2\leq\epsilon$ is $O(\frac{B+\nicefrac{N}{m}}{\epsilon})$.
-Advantage over batch gradient ascent can be achieved with properly selected meta-parameters.
-In the following we propose practical ways to do so.
-
-{\color{red}
+Since the optimal configuration may be task-dependent, we leave this choice to the empirical analysis of Sec.~\ref{sec:exp}.
+
 \textbf{G(PO)MDP.} A similar result to the one stated in Thm~\ref{theo:convergence} can be derived for the G(PO)MDP estimator. A sketch of the proof is reported in~\ref{app:gpomdpproof}.
@@
 For this reason we will use G(PO)MDP for the experiments (see Sec.~\ref{sec:exp}).
-\todopir{sketch of proof for GPOMDP in appendix, probably can be unified with the one of REINFORCE}}
+\todopir{sketch of proof for GPOMDP in appendix, probably can be unified with the one of REINFORCE}
 
@@
 where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s,t}$ with the sub-iterations (see App.~\ref{app:practicalsvrpg} for details).
-\todopir{Add ADAM definition and explain in more details}
 By doing so we decouple the contribution of the variance introduced by the approximate FG from the one introduced by the sub-iterations.
@@
 As we already mentioned in section \ref{sec:stopping} importance weights means a lot of variance to deal with.
-Therefore, in oreder to cope with this issue the literature propose the Self-normalization techniques \cite{precup2000eligibility}.
+Therefore, in order to cope with this issue the literature propose the Self-normalization techniques.
 This technique offers a variance reduction over the importance weights in exchange of a bias introduction.
 Hence, if the bias introduced by this is smaller than the variance reduction it causes, it is convenient to use it in practice.
-We expirienced some improvements using it (see section \ref{sec:exp} for the sperimental point of view).
-Self-normalization is easy to use, but depends on the chosen reinforcement learning algorithm.
-Therefore, here we give the definition of this technique in the context of G(PO)MDP where we have used the weighted per-decision importance sampling estimator \cite{precup2000eligibility}:
-\begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\vert s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
-\end{align*} 
-where:
-\begin{align*}
-\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
- }{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
-\end{align*}
-\begin{align*}
-\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
-}{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
-\end{align*}
-\todomat{why should we use it? What advantages do we expect? At what cost? Pointer to experiments}
+As regards our algorithm, the term
+$
+ \frac{1}{B} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
+\right]
+$  \eqref{E:svrpg.estimate.batch} become:
+\begin{align*}
+\frac{1}{ \sum_{i=0}^{B-1}\omega(\tau_i|\vtheta_t, \wt{\vtheta})} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
+\right].
+\end{align*}
+The rest of terms remain unchanged.
+We experienced some improvements using it (see section \ref{sec:exp} for the experimental point of view), in particular in our experiments we use this technique in the context of G(PO)MDP defined as "weighted per-decision importance sampling estimator" \cite{precup2000eligibility}.
+
+\subsection{Baseline}
+As reported in Sec. \ref{subsec:PolicyGradient} a known method for variance reduction in Reinforcement Learning is the use of baseline.
+This method is independent to SVRPO.
+We decide to test SVRPO with an Actor Critic baseline: the baseline is a states dependent linear function with a time-varying feature vector \cite{duan2016benchmarking}.
+Our choice is to use a single baseline for each term of expression \ref{E:svrpg.estimate.batch}. This is update only in the snapshot: the baseline is estimated using the N trajectories.
 
@@
 
-Concerning importance weighting techniques, RL has made wide use of them for off-policy problems~\citep[\eg][]{precup2000eligibility,thomas2015high}.
+Concerning importance weighting techniques, RL has made wide use of them for off-policy problems~\citep[\eg]{precup2000eligibility,thomas2015high}.
 However, as mentioned before, SVRPG cannot be compared to such methods since it is practically on-policy.
@@
 \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
++L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 +\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
@@
 	&c_{m-1} = 0, \\
-	&c_t = c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}, \\
-	&\Psi_t = \alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
-	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right), \\
+	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\frac{\alpha_t^2L^2}{B}\right) + \alpha_tL^2+\frac{\alpha_t^2L^3}{2B}, \\
+	&\Psi_t = \alpha_t\left(1-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
+	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right), \\
 	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2}.
@@
 \begin{align*}
-&0\leq\alpha_t < \frac{2(1-\nicefrac{2c_{t+1}}{\beta_t})}{L+2c_{t+1}} \\
-&\beta_t > 2c_{t+1}.
+&0\leq\alpha_t < \frac{2(1-\nicefrac{c_{t+1}}{\beta_t})}{L+2c_{t+1}} \\
+&\beta_t > c_{t+1}.
 \end{align*}
@@
 	&\qquad-
-	\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_t}{2}\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}, \label{eq:13}
+	\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}, \label{eq:13}
 \end{align}
@@
 &\qquad+
-\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
+\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
 %
@@
 &\qquad+
-\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
 %
@@
 &\qquad
-\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
+\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
 \end{align}
@@
 	%
-	&\geq	\Ets{J(\vtheta_t^{s+1})+\frac{\alpha_t}{2}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\geq	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-
-	\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
+	\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
 	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:17}\\
 	%
-	&\geq \Ets{J(\vtheta_t^{s+1})+\frac{\alpha_t}{2}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
+	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
 	-\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
-	&\qquad -c_{t+1}\Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	&\qquad -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}
+	-c_{t+1}\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+	%
+	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
+	-\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
+	&\qquad -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2} -c_{t+1}\Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 	\nonumber\\
@@
 	&\qquad
-	-c_{t+1}\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -c_{t+1}\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:18}\\
-	%
-	&= \Ets{J(\vtheta_t^{s+1})} - c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	-c_{t+1}\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -c_{t+1}\frac{1}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:18}\\
+	%
+	&= \Ets{J(\vtheta_t^{s+1})} - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta\right)+\alpha_tL^2\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad-\alpha_t^2\left(\frac{L}{2}+c_{t+1}\right)\Ets{\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}
-	+\frac{\alpha_t}{2}\left(1-\frac{2c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	%
-	&\geq  \Ets{J(\vtheta_t^{s+1})} - c_{t+1}\left(1+\alpha_t\beta+\alpha_t\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	&\geq  \Ets{J(\vtheta_t^{s+1})} - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta\right)+\alpha_tL^2\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad
@@
 	+\frac{\Gamma W}{B}\right)
-	+\frac{\alpha_t}{2}\left(1-2\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \label{eq:19}\\
 	%
-	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(1+\alpha_t\beta+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta+\frac{\alpha_t^2L^2}{B}\right)+\alpha_tL^2+\frac{\alpha_t^2L^3}{2B}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad
-	+\alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
+	+\alpha_t\left(1-\frac{c_{t+1}}{\beta}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})\Gamma\VARIS}{2B} \nonumber\\

### 2018-02-07 svrpg.tex
Modified section 5.2

@@
 As we already mentioned in section \ref{sec:stopping} importance weights means a lot of variance to deal with.
-Therefore, in order to cope with this issue the literature propose the Self-normalization techniques.
+Therefore, in order to cope with this issue the literature propose the Self-normalization technique.
 This technique offers a variance reduction over the importance weights in exchange of a bias introduction.
 Hence, if the bias introduced by this is smaller than the variance reduction it causes, it is convenient to use it in practice.
-As regards our algorithm, the term
-$
- \frac{1}{B} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
-\right]
-$  \eqref{E:svrpg.estimate.batch} become:
-\begin{align*}
-\frac{1}{ \sum_{i=0}^{B-1}\omega(\tau_i|\vtheta_t, \wt{\vtheta})} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
+Introducing self-normalization in the context of our algorithm, we switch from equation \ref{E:svrpg.estimate.batch} to:
+\begin{align*}
+\blacktriangledown J(\vtheta_{t}) := v_t= \wh{\nabla}_N J(\wt{\vtheta})
+ + \frac{1}{B} \sum_{i=0}^{B-1} \left[
+g(\tau_i|\vtheta_t)\right]\\ 
+- \frac{1}{ \sum_{i=0}^{B-1}\omega(\tau_i|\vtheta_t, \wt{\vtheta})} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
 \right].
 \end{align*}
-The rest of terms remain unchanged.
-We experienced some improvements using it (see section \ref{sec:exp} for the experimental point of view), in particular in our experiments we use this technique in the context of G(PO)MDP defined as "weighted per-decision importance sampling estimator" \cite{precup2000eligibility}.
+We experienced some improvements using it (see section \ref{sec:exp} for the experimental point of view).
 
@@
 In each experiment, the length of episodes is 500 steps.
+
+In particular in our experiments we use this technique in the context of G(PO)MDP defined as "weighted per-decision importance sampling estimator" \cite{precup2000eligibility}.
 

### 2018-02-07 svrpg.tex
Updated

--- 2018-02-07 svrpg.tex
@@
 where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s,t}$ with the sub-iterations (see App.~\ref{app:practicalsvrpg} for details).
+\todopir{Add ADAM definition and explain in more details}
 By doing so we decouple the contribution of the variance introduced by the approximate FG from the one introduced by the sub-iterations.
@@
 As we already mentioned in section \ref{sec:stopping} importance weights means a lot of variance to deal with.
-Therefore, in order to cope with this issue the literature propose the Self-normalization techniques.
+Therefore, in oreder to cope with this issue the literature propose the Self-normalization techniques \cite{precup2000eligibility}.
 This technique offers a variance reduction over the importance weights in exchange of a bias introduction.
 Hence, if the bias introduced by this is smaller than the variance reduction it causes, it is convenient to use it in practice.
-As regards our algorithm, the term
-$
- \frac{1}{B} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
-\right]
-$  \eqref{E:svrpg.estimate.batch} become:
-\begin{align*}
-\frac{1}{ \sum_{i=0}^{B-1}\omega(\tau_i|\vtheta_t, \wt{\vtheta})} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
-\right].
-\end{align*}
-The rest of terms remain unchanged.
-We experienced some improvements using it (see section \ref{sec:exp} for the experimental point of view), in particular in our experiments we use this technique in the context of G(PO)MDP defined as "weighted per-decision importance sampling estimator" \cite{precup2000eligibility}.
-
-\subsection{Baseline}
-As reported in Sec. \ref{subsec:PolicyGradient} a known method for variance reduction in Reinforcement Learning is the use of baseline.
-This method is independent to SVRPO.
-We decide to test SVRPO with an Actor Critic baseline: the baseline is a states dependent linear function with a time-varying feature vector \cite{duan2016benchmarking}.
-Our choice is to use a single baseline for each term of expression \ref{E:svrpg.estimate.batch}. This is update only in the snapshot: the baseline is estimated using the N trajectories.
+We expirienced some improvements using it (see section \ref{sec:exp} for the sperimental point of view).
+Self-normalization is easy to use, but depends on the chosen reinforcement learning algorithm.
+Therefore, here we give the definition of this technique in the context of G(PO)MDP where we have used the weighted per-decision importance sampling estimator \cite{precup2000eligibility}:
+\begin{align*}
+\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\vert s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
+\end{align*} 
+where:
+\begin{align*}
+\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
+ }{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
+\end{align*}
+\begin{align*}
+\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
+}{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
+\end{align*}
+\todomat{why should we use it? What advantages do we expect? At what cost? Pointer to experiments}
 
@@
 \left(\sum_{h=0}^{H}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\sum_{h=0}^{H}\gamma^h r(s_h^n,a_h^n) - b\right)
-}_{g(\tau_n|\vtheta)}
+}_{g(\tau_n|\vtheta):=\nabla \log\Delta(\tau_n|\vtheta)\Reward(\tau_n)}
 ,
@@
 
-\textbf{Off-Policy setting.}
+\textbf{Off-policy setting.}
 In off-policy setting two policies, called behavioural $\pi^B$ and target $\pi^T$, are involved.
@@
 \begin{equation}\label{E:offpolicygradient}
-        \nabla J(\pi^T) = \EVV[\tau \sim \Delta(\cdot|\pi^B)]{\omega(\tau) \nabla \log \Delta(\tau|\pi^T) \mathcal{R}(\tau)}
+        \nabla J(\pi^T) = \EVV[\tau \sim \Delta(\cdot|\pi^B)]{\omega(\tau) \nabla \log \Delta(\tau|\pi^T) \mathcal{R}(\tau)} = \EVV[\tau \sim \Delta(\cdot|\pi^B)]{\omega(\tau) g(\tau|\vtheta)}. 
 \end{equation}
 For $\omega(\tau)$ being well defined the behavioural policy should have non-zero probability of selecting any action in every state \ie $\pi^B(a|s) > 0$ for any $(s,a)\in \mathcal{S} \times \mathcal{A}$.
+Equation~\ref{E:offpolicygradient} is important for proving Theorem~\ref{theo:convergence} since it provides a common representation of REINFORCE and G(PO)MDP.
 
@@
 \section{Proofs}\label{app:proofs}
-In this Appendix we prove all the claims made in the paper, with the primary objective of proving Theorem \ref{theo:convergence}.
+In this section, we prove all the claims made in the paper, with the primary objective of proving Theorem \ref{theo:convergence}.
 Our proof is adapted from the one of Theorem 2 from \cite{reddi2016stochastic} and has a very similar structure, but with all the additional challenges and assumptions described in Section \ref{sec:conv}.
@@
 \begin{align*}
-	&\mathbb{E}_{t\vert s}\left[X\right] \coloneqq 
+        \mathbb{E}_{t\vert s}\left[X\right] &\coloneqq 
 		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\vert\tilde{\vtheta}^s)\forall j \in {\it {\bf N}} \\ \tau_{i,h}\sim\pi(\cdot\vert\vtheta^{s+1}_i) \forall h \in {\it {\bf B}}, \text{ for $i=0,\dots,t-1$}}}{\left[X \vert \tilde{\vtheta^s}\right]} \\

### 2018-02-07 svrpg.tex
Auto stash before merge of "master" and "origin/master"
corrections on section 5.3

--- 2018-02-07 svrpg.tex
@@
 \subsection{Baseline}
-As reported in Sec. \ref{subsec:PolicyGradient} a known method for variance reduction in Reinforcement Learning is the use of baseline.
-This method is independent to SVRPO.
-We decide to test SVRPO with an Actor Critic baseline: the baseline is a states dependent linear function with a time-varying feature vector \cite{duan2016benchmarking}.
-Our choice is to use a single baseline for each term of expression \ref{E:svrpg.estimate.batch}. This is update only in the snapshot: the baseline is estimated using the N trajectories.
+As reported in Sec. \ref{subsec:PolicyGradient} a well known variance reduction tool in Reinforcement Learning is the baseline.
+This tool does not depend on SVRPO.
+More precicely we decided to use an Actor Critic baseline, which is a state dependent linear function with a time-varying feature vector \cite{duan2016benchmarking}.
+Furthermore, the baseline is unique and computed every snapshot using the N trajectories at hand.
 

### 2018-02-07 svrpg.tex
Modified section 5.2

@@
 As reported in Sec. \ref{subsec:PolicyGradient} a well known variance reduction tool in Reinforcement Learning is the baseline.
-This tool does not depend on SVRPO.
+This tool does not depend on SVRPG.
 More precicely we decided to use an Actor Critic baseline, which is a state dependent linear function with a time-varying feature vector \cite{duan2016benchmarking}.
@@
 \section{Experiments}\label{sec:exp}
+In practical experiments we use G(PO)MDP as 
 We have designed our experiments to investigate whether:

### 2018-02-07 svrpg.tex
added Lemma on Lipschitz continuity of g

--- 2018-02-07 svrpg.tex
@@
 As we already mentioned in section \ref{sec:stopping} importance weights means a lot of variance to deal with.
-Therefore, in order to cope with this issue the literature propose the Self-normalization technique.
+Therefore, in oreder to cope with this issue the literature propose the Self-normalization techniques \cite{precup2000eligibility}.
 This technique offers a variance reduction over the importance weights in exchange of a bias introduction.
 Hence, if the bias introduced by this is smaller than the variance reduction it causes, it is convenient to use it in practice.
-Introducing self-normalization in the context of our algorithm, we switch from equation \ref{E:svrpg.estimate.batch} to:
-\begin{align*}
-\blacktriangledown J(\vtheta_{t}) := v_t= \wh{\nabla}_N J(\wt{\vtheta})
- + \frac{1}{B} \sum_{i=0}^{B-1} \left[
-g(\tau_i|\vtheta_t)\right]\\ 
-- \frac{1}{ \sum_{i=0}^{B-1}\omega(\tau_i|\vtheta_t, \wt{\vtheta})} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
-\right].
-\end{align*}
-We experienced some improvements using it (see section \ref{sec:exp} for the experimental point of view).
-
-\subsection{Baseline}
-As reported in Sec. \ref{subsec:PolicyGradient} a well known variance reduction tool in Reinforcement Learning is the baseline.
-This tool does not depend on SVRPO.
-More precicely we decided to use an Actor Critic baseline, which is a state dependent linear function with a time-varying feature vector \cite{duan2016benchmarking}.
-Furthermore, the baseline is unique and computed every snapshot using the N trajectories at hand.
+We expirienced some improvements using it (see section \ref{sec:exp} for the sperimental point of view).
+Self-normalization is easy to use, but depends on the chosen reinforcement learning algorithm.
+Therefore, here we give the definition of this technique in the context of G(PO)MDP where we have used the weighted per-decision importance sampling estimator \cite{precup2000eligibility}:
+\begin{align*}
+\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\vert s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
+\end{align*} 
+where:
+\begin{align*}
+\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
+ }{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
+\end{align*}
+\begin{align*}
+\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
+}{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
+\end{align*}
+\todomat{why should we use it? What advantages do we expect? At what cost? Pointer to experiments}
 
@@
 
-In particular in our experiments we use this technique in the context of G(PO)MDP defined as "weighted per-decision importance sampling estimator" \cite{precup2000eligibility}.
-
 \section{Conclusion}
@@
 \begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
-	Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth for some positive Lipschitz constant $L$.
+	Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth for some positive Lipschitz constant $L_1$.
 \end{restatable}
@@
 
+\begin{restatable}[]{lemma}{l-smooth-g}\label{lemma:gsmooth}
+Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, $g(\tau\vert\vtheta)$ is Lipschitz continuous with Lipschitz constant $L_2$, \ie for any trajectory $\tau\in\Tspace$:
+\[
+	\norm[]{g(\tau\vert\vtheta)-g(\tau\vert\vtheta')}^2 \leq L_2\norm[]{\vtheta-\vtheta'}^2.
+\]
+\end{restatable}
+\begin{proof}
+For both REINFORCE and G(PO)MDP, $g(\tau\vert\vtheta)$ is a linear combination of terms of the kind $\score{\vtheta}{a_t\vert s_t}\gamma^t r_t$ \cite{peters2008reinforcement}.
+These terms have bounded gradient from the second inequality of Assumption \ref{ass:bounded_score} and the fact that $|r_t|\leq R$.
+Bounded gradient implies Lipschitz continuity.
+Finally, the linear combination of Lipschitz continuous functions is Lipschitz continuous.
+\end{proof}
+
 \begin{restatable}[]{lemma}{bounded-g}\label{lemma:gbound}
@@
 \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
++\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 +\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
@@
 	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
-	+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}\nonumber\\
+	+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}\nonumber\\
 	&\qquad+
 		\frac{1}{B^2}\sum_{i=0}^{B-1}
-		\Ets{\norm[]{g(\tau_i\vert\tilde{\vtheta}^s) 
-			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \label{eq:3}
+		\Ets{\norm[]{(1 
+			-\omega(\tau_i))g(\tau_i\vert\tilde{\vtheta}^s)}^2} \label{eq:3}
 % ...
@@
 	\nonumber\\
-	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 	+\Gamma\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \label{eq:4}\\
@@
 	\nonumber\\
-	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 	+\frac{\Gamma}{B^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
@@
 	\nonumber\\
-	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 	+\frac{\Gamma\VARIS}{B}, \label{eq:5}
 \end{align}
-where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (\ref{eq:3}) is from L-smoothness,
+where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (\ref{eq:3}) is from Lemma \ref{lemma:gsmooth},
 (\ref{eq:4}) is from Lemma \ref{lemma:gbound}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
@@
 	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right), \\
-	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2}.
-\end{align*}
+	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2},
+\end{align*}
+where $L=\max\left\{L1,L2\right\}$, \ie the greater of the Lipschitz constants from Lemmas \ref{lemma:lsmooth} and \ref{lemma:gsmooth}.
+
 In particular, the following constraints on $\alpha_t$ and $\beta_t$ are sufficient:
@@
 \end{align}
-\todomat{check last inequality}
 where (\ref{eq:11}) is from the L-smoothness of $J(\vtheta)$, (\ref{eq:12}) is from the SVRPG update, and (\ref{eq:13}) is from Lemma \ref{lemma:aux0}.

### 2018-02-07 svrpg.tex

@@
 As we already mentioned in section \ref{sec:stopping} importance weights means a lot of variance to deal with.
-Therefore, in oreder to cope with this issue the literature propose the Self-normalization techniques \cite{precup2000eligibility}.
+Therefore, in order to cope with this issue the literature propose the Self-normalization technique.
 This technique offers a variance reduction over the importance weights in exchange of a bias introduction.
 Hence, if the bias introduced by this is smaller than the variance reduction it causes, it is convenient to use it in practice.
-We expirienced some improvements using it (see section \ref{sec:exp} for the sperimental point of view).
-Self-normalization is easy to use, but depends on the chosen reinforcement learning algorithm.
-Therefore, here we give the definition of this technique in the context of G(PO)MDP where we have used the weighted per-decision importance sampling estimator \cite{precup2000eligibility}:
-\begin{align*}
-\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\vert s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
-\end{align*} 
-where:
-\begin{align*}
-\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
- }{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
-\end{align*}
-\begin{align*}
-\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
-}{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
-\end{align*}
-\todomat{why should we use it? What advantages do we expect? At what cost? Pointer to experiments}
+Introducing self-normalization in the context of our algorithm, we switch from equation \ref{E:svrpg.estimate.batch} to:
+\begin{align*}
+\blacktriangledown J(\vtheta_{t}) := v_t= \wh{\nabla}_N J(\wt{\vtheta})
+ + \frac{1}{B} \sum_{i=0}^{B-1} \left[
+g(\tau_i|\vtheta_t)\right]\\ 
+- \frac{1}{ \sum_{i=0}^{B-1}\omega(\tau_i|\vtheta_t, \wt{\vtheta})} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
+\right].
+\end{align*}
+We experienced some improvements using it (see section \ref{sec:exp} for the experimental point of view).
+
+\subsection{Baseline}
+As reported in Sec. \ref{subsec:PolicyGradient} a well known variance reduction tool in Reinforcement Learning is the baseline.
+This tool does not depend on SVRPG.
+More precicely we decided to use an Actor Critic baseline, which is a state dependent linear function with a time-varying feature vector \cite{duan2016benchmarking}.
+Furthermore, the baseline is unique and computed every snapshot using the N trajectories at hand.
 

### 2018-02-07 svrpg.tex
changed point of view on random return: actually perfect for on-line learning

@@
 %The theoretical settings may be too conservative.
-In Sec.~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$ which is more suited for applications. 
 By analysing the upper-bound in Thm.~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on \textcolor{red}{non-concave} SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
@@
 Once the last two terms can be neglected, the number of trajectories needed to achieve $\norm[2]{\gradJ{\vtheta}}^2\leq\epsilon$ is $O(\frac{B+\nicefrac{N}{m}}{\epsilon})$.
-Advantage over batch gradient ascent can be achieved with properly selected meta-parameters.
-In the following we propose practical ways to do so.
+In this sense, advantage over batch gradient ascent can be achieved with properly selected meta-parameters.
+In Sec.~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$.
+Finally, note from the return statement of Algorithm \ref{alg:svrpg} that $J(\vtheta_A)$ can be seen as the average performance of all the policies tried by the algorithm.
+This is particularly meaningful in the context of on-line learning that we are addressing in this paper.
 

### 2018-02-07 svrpg.tex
modified algs to have s+1

--- 2018-02-07 svrpg.tex
@@
 \begin{equation}\label{E:svrg.gradient.correction}
-        \blacktriangledown f(\vtheta_{t}) = \nabla f(\tilde{\vtheta}) + \nabla z(x_i | \vtheta_t) - \nabla z(x_i | \tilde{\vtheta}),
+        \blacktriangledown f(\vtheta_{t}) := v_t = \nabla f(\tilde{\vtheta}) + \nabla z(x_i | \vtheta_t) - \nabla z(x_i | \tilde{\vtheta}),
 \end{equation} 
@@
 	\begin{algorithmic}
-            \STATE {\bfseries Input:} a dataset $\mathcal{D}_N$, number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{0}^0:=\wt{\vtheta}^0$
+            \STATE {\bfseries Input:} a dataset $\mathcal{D}_N$, number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m}^0:=\wt{\vtheta}^0$
 		\FOR{$s=0$ {\bfseries to} $S-1$}
+        \STATE $\vtheta_0^{s+1} := \wt{\vtheta}^{s} = \vtheta_{m}^s$
         \STATE $\wt{\mu} = \nabla f(\wt{\vtheta}^s)$
@@
         \STATE $x_i \sim \mathcal{U}\left(\mathcal{D}_N\right)$
-		\STATE $v^{s}_t = 
+		\STATE $v^{s+1}_t = 
 %			\nabla f(\wt{\vtheta}^s) + 
 			\wt{\mu} + 
-			\nabla z(x_i|\vtheta_t^{s}) -
+			\nabla z(x_i|\vtheta_t^{s+1}) -
 			\nabla z(x_i|\wt{\vtheta}^{s})
 		$
-        \STATE $\vtheta_{t+1}^{s} = \vtheta_t^{s} + \alpha v^{s}_t$
+        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
 		\ENDFOR
-        \STATE $\vtheta_0^{s+1} := \wt{\vtheta}^{s+1} = \vtheta_{m}^s$
 		\ENDFOR
-        \STATE \underline{Concave case:} \textbf{return} $\vtheta_{0}^S$
-        \STATE \underline{Non-Concave case:} \textbf{return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $\{[0,S[\times[0,m[\} \cup \{(S,0)\}$
+        \STATE \underline{Concave case:} \textbf{return} $\vtheta_{m}^S$
+        \STATE \underline{Non-Concave case:} \textbf{return} $\vtheta_t^{s+1}$ with $(s,t)$ picked uniformly at random from $\{[0,S-1]\times[0,m-1]\}$
 	\end{algorithmic}
@@
         \blacktriangledown J(\vtheta_{t}) &:= v_t= \wh{\nabla}_N J(\wt{\vtheta})\\ \notag
-                                            & \quad{} + \frac{1}{B} \sum_{i=0}^{B-1} \left[
+                                            & \quad{} + 
+        \underbracket{
+        \frac{1}{B} \sum_{i=0}^{B-1} \left[
         g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
-        \right].
+\right]}_{c_t}.
 \end{align}
@@
 	\begin{algorithmic}
-        \STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{0}^0 := \wt{\vtheta}^0$
+        \STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{0}^m := \wt{\vtheta}^0$
 		\FOR{$s=0$ {\bfseries to} $S-1$}
+        \STATE $\vtheta_0^{s+1} := \tilde{\vtheta}^{s} = \vtheta_{m}^s$
         \STATE Sample $N$ trajectories $\{\tau_j\}$ from $\pi(\cdot\vert\tilde{\vtheta}^{s})$
         \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\ref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
-%        \STATE $ \vtheta^s_1 = \wt{\vtheta}^s + \alpha \wt{\mu}$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
         \STATE Sample $B$ trajectories $\{\tau_i\}$ from $\pi(\cdot\vert\vtheta_t^{s+1})$
-        \STATE $c^s_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^s) - \omega(\tau_i|\vtheta^s_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta})^s\right)$
-%        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
-%        \begin{aligned}[t]
-%                & \Reward(\tau_i) \Big[ \score{}{\tau_i|\vtheta_t^{s+1}} \\
-%                &- \omega(\tau|\vtheta^{s+1}_t, \wt{\vtheta}^s)\score{}{\tau_i|\wt{\vtheta}^{s}} \Big]
+%        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)$
+        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
+        \begin{aligned}[t]
+                \Big( & g(\tau_i|\vtheta_t^{s+1})\\ 
+                      & \quad{} - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \Big)
+        \end{aligned}$
+%        \STATE $\begin{aligned}[t]
+%                &v^{s+1}_t = \wt{\mu}\\ 
+%                & +\frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)
 %                \end{aligned}$
-        \STATE $v^{s}_t = \wt{\mu} + c^s_t$ % \gradApp{\wt{\vtheta}^{s}}{N} + c^{s}_t$
+        \STATE $v^{s+1}_t = \wt{\mu} + c^{s+1}_t$ % \gradApp{\wt{\vtheta}^{s}}{N} + c^{s}_t$
 %%		\begin{align*}
@@
 %%		\end{align*}
-        \STATE $\vtheta_{t+1}^{s} = \vtheta_t^{s} + \alpha v^{s}_t$
+        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
 		\ENDFOR
-        \STATE $\vtheta_0^{s+1} := \tilde{\vtheta}^{s+1} = \vtheta_{m}^s$
 		\ENDFOR
-        \STATE {\bfseries return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $\{[0,S[\times[0,m[\} \cup \{(S,0)\}$
+        \STATE {\bfseries return} $\vtheta_t^{s+1}$ with $(s,t)$ picked uniformly at random from $\{[0,S-1]\times[0,m-1]\}$
 	\end{algorithmic}
@@
 \subsection{Baseline}
-As reported in Sec. \ref{subsec:PolicyGradient} a well known variance reduction tool in Reinforcement Learning is the baseline.
-This tool does not depend on SVRPG.
-More precicely we decided to use an Actor Critic baseline, which is a state dependent linear function with a time-varying feature vector \cite{duan2016benchmarking}.
-Furthermore, the baseline is unique and computed every snapshot using the N trajectories at hand.
+As reported in Sec. \ref{subsec:PolicyGradient} a known method for variance reduction in Reinforcement Learning is the use of baseline.
+This method is independent to SVRPO.
+We decide to test SVRPO with an Actor Critic baseline: the baseline is a states dependent linear function with a time-varying feature vector \cite{duan2016benchmarking}.
+Our choice is to use a single baseline for each term of expression \ref{E:svrpg.estimate.batch}. This is update only in the snapshot: the baseline is estimated using the N trajectories.
 
@@
 \section{Experiments}\label{sec:exp}
-In practical experiments we use G(PO)MDP as 
 We have designed our experiments to investigate whether:
@@
 where the sequence $\tilde{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}, ${\it {\bf N}} = [0,\dots,N)$, and ${\it {\bf B}} = [0,\dots,B)$.
+
 \end{definition}
@@
 \begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
-	Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth for some positive Lipschitz constant $L_1$.
+	Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth for some positive Lipschitz constant $L$.
 \end{restatable}
@@
 
-\begin{restatable}[]{lemma}{l-smooth-g}\label{lemma:gsmooth}
-Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, $g(\tau\vert\vtheta)$ is Lipschitz continuous with Lipschitz constant $L_2$, \ie for any trajectory $\tau\in\Tspace$:
-\[
-	\norm[]{g(\tau\vert\vtheta)-g(\tau\vert\vtheta')}^2 \leq L_2\norm[]{\vtheta-\vtheta'}^2.
-\]
-\end{restatable}
-\begin{proof}
-For both REINFORCE and G(PO)MDP, $g(\tau\vert\vtheta)$ is a linear combination of terms of the kind $\score{\vtheta}{a_t\vert s_t}\gamma^t r_t$ \cite{peters2008reinforcement}.
-These terms have bounded gradient from the second inequality of Assumption \ref{ass:bounded_score} and the fact that $|r_t|\leq R$.
-Bounded gradient implies Lipschitz continuity.
-Finally, the linear combination of Lipschitz continuous functions is Lipschitz continuous.
-\end{proof}
-
 \begin{restatable}[]{lemma}{bounded-g}\label{lemma:gbound}
@@
 \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
++\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 +\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
@@
 	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
-	+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}\nonumber\\
+	+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}\nonumber\\
 	&\qquad+
 		\frac{1}{B^2}\sum_{i=0}^{B-1}
-		\Ets{\norm[]{(1 
-			-\omega(\tau_i))g(\tau_i\vert\tilde{\vtheta}^s)}^2} \label{eq:3}
+		\Ets{\norm[]{g(\tau_i\vert\tilde{\vtheta}^s) 
+			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \label{eq:3}
 % ...
@@
 	\nonumber\\
-	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 	+\Gamma\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \label{eq:4}\\
@@
 	\nonumber\\
-	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 	+\frac{\Gamma}{B^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
@@
 	\nonumber\\
-	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	&\qquad+\frac{L^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 	+\frac{\Gamma\VARIS}{B}, \label{eq:5}
 \end{align}
-where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (\ref{eq:3}) is from Lemma \ref{lemma:gsmooth},
+where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (\ref{eq:3}) is from L-smoothness,
 (\ref{eq:4}) is from Lemma \ref{lemma:gbound}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
@@
 	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right), \\
-	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2},
-\end{align*}
-where $L=\max\left\{L1,L2\right\}$, \ie the greater of the Lipschitz constants from Lemmas \ref{lemma:lsmooth} and \ref{lemma:gsmooth}.
-
+	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2}.
+\end{align*}
 In particular, the following constraints on $\alpha_t$ and $\beta_t$ are sufficient:
@@
 \end{align}
+\todomat{check last inequality}
 where (\ref{eq:11}) is from the L-smoothness of $J(\vtheta)$, (\ref{eq:12}) is from the SVRPG update, and (\ref{eq:13}) is from Lemma \ref{lemma:aux0}.

### 2018-02-07 svrpg.tex
Modified Experiments

--- 2018-02-07 svrpg.tex
@@
 %The theoretical settings may be too conservative.
+In Sec.~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$ which is more suited for applications. 
 By analysing the upper-bound in Thm.~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on \textcolor{red}{non-concave} SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
@@
 Once the last two terms can be neglected, the number of trajectories needed to achieve $\norm[2]{\gradJ{\vtheta}}^2\leq\epsilon$ is $O(\frac{B+\nicefrac{N}{m}}{\epsilon})$.
-In this sense, advantage over batch gradient ascent can be achieved with properly selected meta-parameters.
-In Sec.~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$.
-Finally, note from the return statement of Algorithm \ref{alg:svrpg} that $J(\vtheta_A)$ can be seen as the average performance of all the policies tried by the algorithm.
-This is particularly meaningful in the context of on-line learning that we are addressing in this paper.
+Advantage over batch gradient ascent can be achieved with properly selected meta-parameters.
+In the following we propose practical ways to do so.
 
@@
 \section{Experiments}\label{sec:exp}
-In practical experiments we use G(PO)MDP as 
-We have designed our experiments to investigate whether:
+In this section we address the performance evaluation of our policy gradient estimator defined in equation \ref{E:svrpg.estimate.batch} against that one of equation \ref{E:policygradient.estimate}.
+This comparison is performed using G(PO)MDP as estimator of $\score{\vtheta}{\tau}\Reward(\tau)$.
+
+We want to investigate whether:
 \begin{enumerate}

### 2018-02-07 svrpg.tex
Updated untile lemma b.5

--- 2018-02-07 svrpg.tex
@@
 \newcommand{\trajdistr}{\pi_{\vtheta}(\tau)}
-\newcommand{\score}[2]{\nabla\log\pi_{#1}(#2)}
+\newcommand{\score}[2]{\nabla\log p_{#1}(#2)}
 \newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
@@
 It is easy to show that the SVRPG estimator has the following, desirable properties (see App.~\ref{app:proofs}):
-\begin{lemma}\label{L:svrpg.properties}
-Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$.
-Then, the SVRG estimate in~\eqref{E:svrpg.estimate.batch} is \emph{unbiased}
+\begin{lemma}\label{L:svrpg.properties} \todopirout{extended to any unbiased estimator}
+        {\color{blue} Let $\wh{\nabla}_N J(\vtheta)$ be an unbiased estimator of~\eqref{E:policygradient} }
+%Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} 
+and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$. Then, the SVRG estimate in~\eqref{E:svrpg.estimate.batch} is \emph{unbiased}
 \[
@@
 	\Var\left[\gradBlack{\vtheta^*}\right] = 
-    \Var\left[\wh{\nabla}_N^{\textsc{RF}}J(\vtheta^*)\right].
+    \Var\left[\wh{\nabla}_N J(\vtheta^*)\right].
 \]
@@
 \]
-where $\omega(\tau) = \frac{\Delta(\tau,\pi^T)}{\Delta(\tau|\pi^B)} = \omega(z_{0:H}) = \prod_{w=0}^{H} \omega(z_w)$ and $\omega(z_w) = \frac{\pi^T(a_w|s_w)}{\pi^B(a_w|s_w)}$
+where $\omega(\tau|\pi^B,\pi^T) = \frac{\Delta(\tau,\pi^T)}{\Delta(\tau|\pi^B)} = \omega(z_{0:H}|\pi^B,\pi^T) = \prod_{w=0}^{H} \omega(z_w|\pi^B,\pi^T)$ and $\omega(z_w|\pi^B,\pi^T) = \frac{\pi^T(a_w|s_w)}{\pi^B(a_w|s_w)}$
 
@@
 \begin{equation}\label{E:offpolicygradient}
-        \nabla J(\pi^T) = \EVV[\tau \sim \Delta(\cdot|\pi^B)]{\omega(\tau) \nabla \log \Delta(\tau|\pi^T) \mathcal{R}(\tau)} = \EVV[\tau \sim \Delta(\cdot|\pi^B)]{\omega(\tau) g(\tau|\vtheta)}. 
+        \nabla J(\pi^T) = \EVV[\tau \sim \Delta(\cdot|\pi^B)]{\omega(\tau|\pi^B,\pi^T) \nabla \log \Delta(\tau|\pi^T) \mathcal{R}(\tau)} = \EVV[\tau \sim \Delta(\cdot|\pi^B)]{\omega(\tau|\pi^B,\pi^T) g(\tau|\pi^T)}.
 \end{equation}
@@
 
+Note that in the following we will make wide use of the following properties.
+\begin{assumption}\label{asm:unbiasedness}
+        We consider an estimate $\wh{\nabla}_N J(\vtheta)$ as in Eq.~\ref{E:policygradient.estimate} such that
+\begin{enumerate}
+        \item \textit{On-policy Unbiased Estimator.} 
+                \[
+                        \EVV[\tau_i \sim \pi(\cdot|\vtheta)]{\wh{\nabla}_N J(\vtheta)} = \EVV[\tau_i \sim \pi(\cdot|\vtheta)]{\frac{1}{N} \sum_{i=0}^N g(\tau_i|\vtheta)} = \nabla J(\vtheta)
+                \]
+        \item \textit{Off-policy Unbiased Estimator.}
+                \[
+        \EVV[\tau_i \sim \Delta(\cdot|\pi^B)]{\frac{1}{N} \sum_{i=0}^N \omega(\tau_i|\pi^B,\vtheta) g(\tau_i|\vtheta) } 
+        = \nabla J(\vtheta) 
+                \]
+\end{enumerate}
+\end{assumption}
+Note that these assumptions are verified by REINFORCE and G(PO)MDP.
+
 \subsection*{Definitions}
@@
 \end{definition}
+
+We introduce the notation $\tau_{i,h}$ to denote the $h$-th trajectory collected using policy $\vtheta^{s+1}_i$ where $s$ will be clear from the context.
 
@@
 			\vert\vtheta_0^{s+1}}
-		\vert\tilde{\vtheta}^s},
+		\vert\tilde{\vtheta}^s}\\
+        &= \EVV[t-1|s]{\EVV[\tau_{t,h}\sim \pi(\cdot|\vtheta^{s+1}_t)]{X|\vtheta^{s+1}_t}}
 \end{align*}
 where the sequence $\tilde{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}, ${\it {\bf N}} = [0,\dots,N)$, and ${\it {\bf B}} = [0,\dots,B)$. 
+\todopir{what about $\vtheta^{s+1}_0$ and $\wt{\vtheta}^s$? They are equal. How this affect previous definition?}
 
@@
 	\gradJ{\tilde{\vtheta}^s}
-	+ \nabla\log\pi(\tau_i \vert \vtheta_t^{s+1})\Reward(\tau_i) 
-	- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta}^s)\Reward(\tau_i) \\
+%	+ \nabla\log\pi(\tau_i \vert \vtheta_t^{s+1})\Reward(\tau_i) 
+    + g(\tau_i|\vtheta^{s+1}_t)
+%	- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta}^s)\Reward(\tau_i) 
+    - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i|\wt{\vtheta}^s)
+    \\
 	&= \gradBlack{\vtheta_t^{s+1}} - \gradApp{\tilde{\vtheta}^s}{N} + \gradJ{\tilde{\vtheta}^s} \\
@@
 \begin{restatable}{lemma}{unbias}\label{lemma:unbias}
-The SVRPG estimator is unbiased:
+Consider Assumption~\ref{asm:unbiasedness}. Then, the SVRPG estimator is unbiased:
 \[
@@
 \begin{align*}
-\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\tilde{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} - \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i)\score{\tilde{\vtheta}}{\tau_i}R(\tau_i)} \\
-&= \gradJ{\tilde{\vtheta}} + \gradJ{\vtheta} - \gradJ{\tilde{\vtheta}} = \gradJ{\vtheta}.
+        \EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\tilde{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} - \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
+&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
 \end{align*}
@@
 \begin{align*}
-\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[\score{\vtheta^*}{\tau}R(\tau) - \cancel{\omega(\tau)}\score{\vtheta^*}{\tau}R(\tau)\right] \\
+\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*) - \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right] \\
 &= \Var\left[\gradApp{\vtheta^*}{N}\right].
@@
 \Dij{J(\vtheta)}{\theta} 
-&= \int_{\Tspace}\Dij{}{\theta}\pol(\tau)\Reward(\tau)\de \tau
+&= \int_{\Tspace}\Dij{}{\theta}p(\tau|\vtheta)\Reward(\tau)\de \tau
 \nonumber\\ 
-&= \int_{\Tspace}\pol(\tau)\score{\theta}{\tau}\score{\theta}{\tau}^T\Reward(\tau)\de \tau + \int_{\Tspace}\pol(\tau)\Dij{}{\theta}\log\pol(\tau)\Reward(\tau)\de \tau \nonumber\\
-&\leq \sup_{\tau} \left\{\left|\Reward(\tau)\right|\right\} \left(H^2\GRADLOG^2+H\HESSLOG\right) \label{eq:0}\\
+&= \int_{\Tspace}p(\tau|\vtheta)\score{\vtheta}{\tau}\score{\vtheta}{\tau}^T\Reward(\tau)\de \tau + \int_{\Tspace}\pol(\tau)\Dij{}{\theta}\log p(\tau|\vtheta)\Reward(\tau)\de \tau \nonumber\\
+&\leq \sup_{\tau \in \mathcal{T}} \left\{\left|\Reward(\tau)\right|\right\} \left(H^2\GRADLOG^2+H\HESSLOG\right) \label{eq:0}\\
 &= \frac{1-\gamma^H}{1-\gamma}RH\left(H\GRADLOG^2+\HESSLOG\right),\nonumber
@@
 \[
-	\norm[]{g(\tau\vert\vtheta)-g(\tau\vert\vtheta')}^2 \leq L_2\norm[]{\vtheta-\vtheta'}^2.
-\]
+	\norm[2]{g(\tau\vert\vtheta)-g(\tau\vert\vtheta')}^2 \leq L_2\norm[2]{\vtheta-\vtheta'}^2.
+\]
+\todopir{I would denote $L_J$ and $L_g$ rather than 1 and 2}
 \end{restatable}
 \begin{proof}
-For both REINFORCE and G(PO)MDP, $g(\tau\vert\vtheta)$ is a linear combination of terms of the kind $\score{\vtheta}{a_t\vert s_t}\gamma^t r_t$ \cite{peters2008reinforcement}.
+        For both REINFORCE and G(PO)MDP, $g(\tau\vert\vtheta)$ is a linear combination of terms of the kind $\nabla \log \pi_{\vtheta}(a_t\vert s_t)\gamma^t r_t$ \cite{peters2008reinforcement}.
 These terms have bounded gradient from the second inequality of Assumption \ref{ass:bounded_score} and the fact that $|r_t|\leq R$.
@@
 \[
-	\norm[]{g(\tau\vert\vtheta)}^2 \leq \Gamma.
+	\norm[2]{g(\tau\vert\vtheta)}^2 \leq \Gamma.
 \]
@@
 \begin{align*}
-	\norm[]{g(\tau\vert\vtheta)}^2 &=
-	\norm[]{\score{\vtheta}{\tau}R(\tau)}^2 \\
-	&=\norm[]{\left(\sum_{t=0}^{H-1}\score{\vtheta}{a_t\vert s_t}\right)\left(\sum_{t=0}^{H-1}\gamma^t r_t\right)}^2\leq H^2G^2\frac{(1-\gamma^H)^2}{(1-\gamma)^2}R^2\dim(\vTheta) \coloneqq \Gamma
-\end{align*}
-For G(PO)MDP, we do not have a compact expression for $g$, but since it is derived from REINFORCE by neglecting some terms of the kind $\score{\vtheta}{a_t\vert s_t}\gamma^t r_t$ \cite{baxter2001infinite}\cite{peters2008reinforcement}, the above bound still holds.
+	\norm[2]{g(\tau\vert\vtheta)}^2 &=
+	\norm[2]{\score{\vtheta}{\tau}R(\tau)}^2 \\
+	&=\norm[2]{\left(\sum_{t=0}^{H-1}\score{\vtheta}{a_t\vert s_t}\right)\left(\sum_{t=0}^{H-1}\gamma^t r_t\right)}^2\leq H^2G^2\frac{(1-\gamma^H)^2}{(1-\gamma)^2}R^2\dim(\vTheta) \coloneqq \Gamma
+\end{align*}
+For G(PO)MDP, we do not have a compact expression for $g$, but since it is derived from REINFORCE by neglecting some terms of the kind $\nabla \log \pi_{\vtheta}{a_t | s_t}\gamma^t r_t$ \cite{baxter2001infinite,peters2008reinforcement}, the above bound still holds.
+\todopir{we need to prove it formally if we have time}
 \end{proof}
@@
 \[
-\Ets{\gradBlack{\vtheta_t^{s+1}}} \leq
-\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+        \Ets{\norm[2]{\gradBlack{\vtheta_t^{s+1}}}^2} \leq
+\Ets[t-1]{\norm[2]{\gradJ{\vtheta_t^{s+1}}}^2} 
++\frac{L_2^2}{B}\Ets[t-1]{\norm[2]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 +\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
@@
 \begin{proof}
+For ease of notation denote $\omega(\tau_i) := \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^{s})$. Then,
+\begingroup
+\allowdisplaybreaks
 	\begin{align}
-	\Ets{\gradBlack{\vtheta_t^{s+1}}} 
-	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
+%            \Ets{\norm[2]{\gradBlack{\vtheta_t^{s+1}}}^2}
+            \mathbb{E}_{t|s}\big[&\norm[2]{\gradBlack{\vtheta_t^{s+1}}}^2\big]
+	= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
 			+\frac{1}{B}\sum_{i=0}^{B-1} g(\tau_i\vert\vtheta_t^{s+1}) 
@@
 			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)
-			\right.\right.\nonumber\\&\qquad\left.\left.
-			-\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
-			+\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}\right\|^2\right] \nonumber\\
+			% \right.\right.\nonumber
+            % \\&\qquad\left.\left.
+            \pm \gradJ{\vtheta_t^{s+1}} \pm \gradJ{\tilde{\vtheta}^s} 
+			% -\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
+			% +\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}
+    \right\|^2\right] \nonumber\\
 	%
@@
 			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)
-		\right.\right.\nonumber\\&\qquad\left.\left.
+	%	\right.\right.\nonumber\\&\qquad\left.\left.
 		- \Ets{
@@
 		\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)
-		\right.\right.\nonumber\\&\qquad\left.\left.
+		% \right.\right.\nonumber\\&\qquad\left.\left.
 		- \Ets{
@@
 		\label{eq:1}\\%
-	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
-	&\qquad+\Ets{\norm[]{
+	&\leq \Ets[t-1]{\norm[2]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} 
+%    \nonumber\\ &\qquad
+    +\Ets{\norm[2]{
 			\frac{1}{B}\sum_{i=0}^{B-1}\left(
@@
 	%
-	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
-	&\qquad+
-			\frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[]{
+	&\leq \Ets[t-1]{\norm[2]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
+%     \nonumber\\ 	&\qquad
+			+ \frac{1}{B^2}\sum_{i=0}^{B-1}
+			\Ets{\norm[2]{
 			g(\tau_i\vert\vtheta_t^{s+1}) -
-			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
-	%
-	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
-	\nonumber\\
-	&\qquad+
-			\frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
-			-g(\tau_i\vert\tilde{\vtheta}^s)
-			+g(\tau_i\vert\tilde{\vtheta}^s) 
-			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
+    \omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s) \pm g(\tau_i|\wt{\vtheta}^s) }^2} \nonumber\\
+	%
+%%	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+%%	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
+%%	\nonumber\\
+%%	&\qquad+
+%%			\frac{1}{B^2}\sum_{i=0}^{B-1}
+%%			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
+%%			-g(\tau_i\vert\tilde{\vtheta}^s)
+%%			+g(\tau_i\vert\tilde{\vtheta}^s) 
+%%			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
 	%
@@
 			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
-			-g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
-	&\qquad
+			-g(\tau_i\vert\tilde{\vtheta}^s)}^2} 
+%        \nonumber\\ 	&\qquad
 			+\frac{1}{B^2}\sum_{i=0}^{B-1}
@@
 	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
-	+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}\nonumber\\
-	&\qquad+
+	\nonumber\\
+	&\qquad
+    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+    +
 		\frac{1}{B^2}\sum_{i=0}^{B-1}
 		\Ets{\norm[]{(1 
-			-\omega(\tau_i))g(\tau_i\vert\tilde{\vtheta}^s)}^2} \label{eq:3}
+			-\omega(\tau_i))g(\tau_i\vert\tilde{\vtheta}^s)}^2} \label{eq:3}\\
 % ...
-\end{align}
-\begin{align}
+%\end{align}
+%\begin{align}
 % ...
@@
 	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
-	\nonumber\\
-	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+%	\nonumber\\&\qquad
+    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 	+\frac{\Gamma}{B^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
@@
 	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
-	\nonumber\\
-	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+% 	\nonumber\\ &\qquad
+    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
 	+\frac{\Gamma\VARIS}{B}, \label{eq:5}
 \end{align}
-where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (\ref{eq:3}) is from Lemma \ref{lemma:gsmooth},
+\endgroup
+where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$\todopirout{ref}, (\ref{eq:3}) is from Lemma \ref{lemma:gsmooth},
 (\ref{eq:4}) is from Lemma \ref{lemma:gbound}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
+\todopir{Explain step $E[\|\|^2] \leq Var$}
 \end{proof}

### 2018-02-07 svrpg.tex
Minor changes to Section 3

--- 2018-02-07 svrpg.tex
@@
 The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $\mathcal{D}_N^k$.
-The most common policy gradient estimators can be expressed as follows
+The most common policy gradient estimators (\eg REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}) can be expressed as follows
 \begin{align} \label{E:policygradient.estimate}
@@
 \end{align}
-where $g(\tau_i|\vtheta)$ is an estimate of $\score{\vtheta}{\tau}\Reward(\tau)$, \eg REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
+where $g(\tau_i|\vtheta)$ is an estimate of $\score{\vtheta}{\tau_i}\Reward(\tau_i)$.
 Although REINFORCE definition is simpler than the G(PO)MDP one, the latter is often preferred due to its lower variance.
@@
 \section{SVRG in Reinforcement Learning}\label{sec:alg}
-In online RL problems the usual approach is to tune the batch size of SG in order to find the optimal trade-off between variance and speed.
-Recall that compared to SL, the samples are not fixed in advance but need to be collected at each policy change.
+In online RL problems, the usual approach is to tune the batch size of SG to find the optimal trade-off between variance and speed.
+Recall that, compared to SL, the samples are not fixed in advance but we need to collect them at each policy change.
 Since this operation may be costly, we would like to minimize the number of interactions with the environment.
-Due to this issues, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
-However, a direct application of SVRG to RL is not possible due to the following properties:
+For these reasons, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
+However, a direct application of SVRG to RL is not possible due to the following issues:
 % In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence.
@@
         \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
-        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The actual dataset is infinite and coincide with $\pi_{\vtheta}(\tau)$.
+        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem.
+The objective function is an expected value over the trajectory density $\pi_{\vtheta}(\tau)$ of the total discounted reward.
 %    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
-        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the parameter $\vtheta$ influences the sampling process.
+        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the policy parameter $\vtheta$ influences the sampling process.
 \end{description}
@@
 As reported in Sec. \ref{subsec:PolicyGradient} a well known variance reduction tool in Reinforcement Learning is the baseline.
-This tool does not depend on SVRPG.
+This tool does not depend on SVRPO.
 More precicely we decided to use an Actor Critic baseline, which is a state dependent linear function with a time-varying feature vector \cite{duan2016benchmarking}.
@@
 \section{Experiments}\label{sec:exp}
-In this section we address the performance evaluation of our policy gradient estimator defined in equation \ref{E:svrpg.estimate.batch} against that one of equation \ref{E:policygradient.estimate}.
-This comparison is performed using G(PO)MDP as estimator of $\score{\vtheta}{\tau}\Reward(\tau)$.
-
-We want to investigate whether:
+We have designed our experiments to investigate whether:
 \begin{enumerate}

### 2018-02-07 svrpg.tex
minor changes to Sec 5.2 and 5.3

--- 2018-02-07 svrpg.tex
@@
 \subsection{Normalized Importance Sampling}\label{sec:prac}
-As we already mentioned in section \ref{sec:stopping} importance weights means a lot of variance to deal with.
-Therefore, in order to cope with this issue the literature propose the Self-normalization technique.
-This technique offers a variance reduction over the importance weights in exchange of a bias introduction.
-Hence, if the bias introduced by this is smaller than the variance reduction it causes, it is convenient to use it in practice.
+As mentioned in section \ref{sec:stopping}, importance weights are an additional source of variance. A common way to cope with this issue is self-normalization.
+This technique can reduce the variance of the importance weights at the cost of introducing some bias. Whether the trade-off is advantageous depends on the specific task.  
 Introducing self-normalization in the context of our algorithm, we switch from equation \ref{E:svrpg.estimate.batch} to:
 \begin{align*}
-\blacktriangledown J(\vtheta_{t}) := v_t= \wh{\nabla}_N J(\wt{\vtheta})
+\blacktriangledown J(\vtheta_{t}) = \wh{\nabla}_N J(\wt{\vtheta})
  + \frac{1}{B} \sum_{i=0}^{B-1} \left[
@@
 \end{align*}
-We experienced some improvements using it (see section \ref{sec:exp} for the experimental point of view).
+In Section \ref{sec:exp} we show some cases in which self-normalization provides a performance improvement.
+\todomat{add reference, maybe just Precup}
 
 \subsection{Baseline}
-As reported in Sec. \ref{subsec:PolicyGradient} a well known variance reduction tool in Reinforcement Learning is the baseline.
-This tool does not depend on SVRPG.
-More precicely we decided to use an Actor Critic baseline, which is a state dependent linear function with a time-varying feature vector \cite{duan2016benchmarking}.
-Furthermore, the baseline is unique and computed every snapshot using the N trajectories at hand.
+As reported in Sec. \ref{subsec:PolicyGradient}, a well known variance reduction tool in Reinforcement Learning is the employment of baselines in policy gradient estimation.
+This tool is orthogonal to the variance reduction technique proposed in this paper.
+In Section \ref{sec:exp} we try to combine SVRPG with a non-stationary, state-dependent, learned linear baseline like the one used in \cite{duan2016benchmarking}, with the variant that baseline parameters are updated only at the snapshot, with trajectories sampled from $\pi_{\tilde{\vtheta}^s}$.
 

### 2018-02-07 svrpg.tex
Updated tau distribution

--- 2018-02-07 svrpg.tex
@@
 A trajectory $\tau$ is a sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following a stationary policy, where $s_0 \sim \rho$.
-With abuse of notation, we denote with $\pi(\tau)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories, and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
+We denote with $p(\tau|\pi)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories (see Appendix~\ref{A:gradient_estimators} for the definition), and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
 %
@@
 %
-Policies can be ranked based on their expected total reward: $J(\pi) = \EVV[\tau \sim \pi]{\Reward(\tau)|M}$.
+Policies can be ranked based on their expected total reward: $J(\pi) = \EVV[\tau \sim p(\cdot|\pi)]{\Reward(\tau)|M}$.
 Solving an MDP $M$ means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.
@@
 Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^d\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$.
-We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
+We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $p_{\vtheta}(\tau)$ with $p(\tau|\vtheta)$ for the sake of readability.
 The search for a locally optimal policy is performed through gradient ascent, where the policy gradient
@@
 \begin{align} \label{E:policygradient}
-        \gradJ{\vtheta} = \EVV[\tau \sim \pol]{\score{\vtheta}{\tau}\Reward(\tau)}.
+        \gradJ{\vtheta} = \EVV[\tau \sim p(\cdot|\vtheta)]{\score{\vtheta}{\tau}\Reward(\tau)}.
         %\int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
@@
 The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $\mathcal{D}_N^k$.
-The most common policy gradient estimators (\eg REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}) can be expressed as follows
+The most common policy gradient estimators can be expressed as follows
 \begin{align} \label{E:policygradient.estimate}
@@
 \end{align}
-where $g(\tau_i|\vtheta)$ is an estimate of $\score{\vtheta}{\tau_i}\Reward(\tau_i)$.
+where $g(\tau_i|\vtheta)$ is an estimate of $\score{\vtheta}{\tau_i}\Reward(\tau_i)$, \eg REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
 Although REINFORCE definition is simpler than the G(PO)MDP one, the latter is often preferred due to its lower variance.
@@
 \section{SVRG in Reinforcement Learning}\label{sec:alg}
-In online RL problems, the usual approach is to tune the batch size of SG to find the optimal trade-off between variance and speed.
-Recall that, compared to SL, the samples are not fixed in advance but we need to collect them at each policy change.
+In online RL problems the usual approach is to tune the batch size of SG in order to find the optimal trade-off between variance and speed.
+Recall that compared to SL, the samples are not fixed in advance but need to be collected at each policy change.
 Since this operation may be costly, we would like to minimize the number of interactions with the environment.
-For these reasons, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
-However, a direct application of SVRG to RL is not possible due to the following issues:
+Due to this issues, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
+However, a direct application of SVRG to RL is not possible due to the following properties:
 % In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence.
@@
         \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
-        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem.
-The objective function is an expected value over the trajectory density $\pi_{\vtheta}(\tau)$ of the total discounted reward.
+        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The actual dataset is infinite and coincide with $\pi_{\vtheta}(\tau)$.
 %    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
-        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the policy parameter $\vtheta$ influences the sampling process.
+        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the parameter $\vtheta$ influences the sampling process.
 \end{description}
@@
 % \end{align*}
-where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{\pi(\tau|\wt{\vtheta})}{\pi(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$.
+where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{p(\tau|\wt{\vtheta})}{p(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$.
 Similarly to SVRG, we have that $\vtheta_0 := \wt{\vtheta}$ and the update is a FG step.
@@
 \end{align}
-We want just to stress that $\EVV[\tau_i \sim \pi_{\vtheta_t}]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to App.~\ref{} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}
+We want just to stress that $\EVV[\tau_i \sim p(\cdot|\vtheta_t)]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to App.~\ref{} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}
 This property will be used to prove Lem.~\ref{L:svrpg.properties}.
@@
         \STATE $\vtheta_0^{s+1} := \tilde{\vtheta}^{s} = \vtheta_{m}^s$
-        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $\pi(\cdot\vert\tilde{\vtheta}^{s})$
+        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\tilde{\vtheta}^{s})$
         \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\ref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
-        \STATE Sample $B$ trajectories $\{\tau_i\}$ from $\pi(\cdot\vert\vtheta_t^{s+1})$
+        \STATE Sample $B$ trajectories $\{\tau_i\}$ from $p(\cdot\vert\vtheta_t^{s+1})$
 %        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)$
@@
 		\[
-                \mathbb{V}ar\left[\omega(\tau| \vtheta_1, \vtheta_2)\right] \leq \VARIS, \quad \forall \vtheta_1,\vtheta_2 \in \realspace^d , \tau \sim \pi_{\vtheta_1}.
+                \mathbb{V}ar\left[\omega(\tau| \vtheta_1, \vtheta_2)\right] \leq \VARIS, \quad \forall \vtheta_1,\vtheta_2 \in \realspace^d , \tau \sim p(\cdot|\vtheta_1).
 		\]

### 2018-02-07 svrpg.tex
images

@@
 \usepackage{graphicx}
+\graphicspath{ {images/} }
 \usepackage{subfigure}
@@
 This comparison is performed using G(PO)MDP as estimator of $\score{\vtheta}{\tau}\Reward(\tau)$.
-
-We want to investigate whether:
-\begin{enumerate}
-\item The algorithm reaches high value of the reward function faster than GPOMDP.
-\item Averagely SVRPG keep the reward function more stable than GPOMDP.
-\end{enumerate}
+For the sake of simplicity and with a little abuse of notation we refer to \ref{E:svrpg.estimate.batch} as SVRPG and to \ref{E:policygradient.estimate} as G(PO)MDP.
+In all the experiments we used gaussian policies and bootsrap confidence intervals $\alpha = 10\% $ over ten different runs for evaluation.
+Each run with a different policy initialization which is kept fixed when changing the algorithm.
+The maximum number of trajectories at hand is fixed for each task.
 
@@
 
-In all the experiments we compete with GPOMDP standard algorithm. GPOMDP uses a number of trajectories per batch that is equal to the number of trajectories used in SVRPG mini-batch size.
-The neural network and the learning rate are the same for GPOMDP and SVRPG.
-In SVRPG we used two different Adam parameters: the first one for Snapshot and the second one for sub-iterations.
-We used the stopping condition defined in Section \ref{sec:stopping}.
-We fixed the maximum number of trajectories that both the algorithms can use as shown in appendices.
-
-The neural networks is composed of one layer with 8 neurons for Cart-Pole environment and by two layers with 32x32 neurons for Swimmer and for Half Cheetah environments.
-In each experiment, the length of episodes is 500 steps.
-
-In particular in our experiments we use this technique in the context of G(PO)MDP defined as "weighted per-decision importance sampling estimator" \cite{precup2000eligibility}.
+\begin{figure}[t]
+    \includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP}
+    \caption{Self-Normalized SVRPG vs GPOMDP in swimmer}
+    \label{fig:swimmerone}
+\end{figure}
+\begin{figure}[t]
+    \includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
+    \caption{Self-Normalized SVRPG vs Non Self-Normalized SVRPG in swimmer}
+    \label{fig:swimmertwo}
+\end{figure}
+\begin{figure}[t]
+    \includegraphics[width=0.5\textwidth]{half_cheetah_self_norm_baseline_SVRPG_vs_GPOMDP100_baseline}
+    \caption{Self-Normalized SVRPG with baseline vs GPOMDP with baseline in half cheetah}
+    \label{fig:halfcheetah}
+\end{figure}
+\begin{figure}[t]
+    \includegraphics[width=0.5\textwidth]{half_cheetah_self_norm_baseline_SVRPG_vs_GPOMDP100_baseline}
+    \caption{SVRPG vs GPOMDP in cart-pole}
+    \label{fig:cartpole}
+\end{figure}
 

### 2018-02-07 svrpg.tex
corrections on images

--- 2018-02-07 svrpg.tex
@@
 Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^d\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$.
-We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $p_{\vtheta}(\tau)$ with $p(\tau|\vtheta)$ for the sake of readability.
+For sake of brevity, we will denote the performance of a parametric policy with $J(\vtheta)$ and the probability of a trajectory $\tau$ with $p(\tau|\vtheta)$ (in some occasions, $p(\tau|\vtheta)$ will be replaced by $p_{\vtheta}(\tau)$ for the sake of readability).
 The search for a locally optimal policy is performed through gradient ascent, where the policy gradient
@@
 At each iteration $k >0$, a batch $\mathcal{D}_N^k = \{\tau_i\}_{i=0}^N$ of $N>0$ trajectories is collected using policy $\pi_{\vtheta_k}$.
-The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $\mathcal{D}_N^k$.
-The most common policy gradient estimators can be expressed as follows
+The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of Eq.~\eqref{E:policygradient} using $\mathcal{D}_N^k$.
+The most common policy gradient estimators (\eg REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}) can be expressed as follows
 \begin{align} \label{E:policygradient.estimate}
@@
 \end{align}
-where $g(\tau_i|\vtheta)$ is an estimate of $\score{\vtheta}{\tau_i}\Reward(\tau_i)$, \eg REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
-Although REINFORCE definition is simpler than the G(PO)MDP one, the latter is often preferred due to its lower variance.
+where $g(\tau_i|\vtheta)$ is an estimate of $\score{\vtheta}{\tau_i}\Reward(\tau_i)$.
+Although the REINFORCE definition is simpler than the G(PO)MDP one, the latter is usually preferred due to its lower variance.
 We refer the reader to Appendix~\ref{A:gradient_estimators} for details and a formal definition of $g$.
@@
 There has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
-Although these techniques have been mainly derived for finite-sum problems, we will show in Sec.~\ref{sec:alg} how they can be used in RL.
+Although these techniques have been mainly derived for finite-sum problems, we will show in Section~\ref{sec:alg} how they can be used in RL.
 The next section has the aim to describe variance reduction techniques for finite-sum problems. In particular, we will present the SVRG algorithm that is at the core of this work.
@@
         \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
-        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The actual dataset is infinite and coincide with $\pi_{\vtheta}(\tau)$.
+        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem.
+The objective function is an expected value over the trajectory density $p_{\vtheta}(\tau)$ of the total discounted reward, for which we would need an infinite dataset.
 %    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
-        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the parameter $\vtheta$ influences the sampling process.
+        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the policy parameter $\vtheta$ influences the sampling process.
 \end{description}
-To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}, see App.~\ref{app:gauss} for the derivation.
+To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}, see Appendix~\ref{app:gauss} for the derivation.
 Because of the infinite dataset, we can only rely on an estimate of the full gradient.
-\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being \textcolor{red}{concave}, showing that SVRG is robust to inexact computation of the full gradient.
+\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being \textcolor{red}{concave}, showing that SVRG is robust to an inexact computation of the full gradient.
 In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate.
 {\color{red}Later on we will show how the estimation accuracy impacts on the convergence results with non-concave objective.}
-Finally, the unbiasedness of Eq.~\ref{E:svrg.gradient.correction} is broken by the non-stationarity of the optimization problem.
+Finally, the non-stationarity of the optimization problem introduces a bias into the SVRG estimator in Eq.~\eqref{E:svrg.gradient.correction}.
 To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
@@
 
-We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG) for a generic policy gradient estimator $g$ (see Eq.~\ref{E:policygradient.estimate}).
+We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG) for a generic policy gradient estimator $g$ (see Eq.~\eqref{E:policygradient.estimate}).
 % For simplicity, we focus on the REINFORCE estimator but the extension to G(PO)MDP is straightforward.
@@
 \begin{align*}
-        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
+        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta}),
 \end{align*}
@@
 % \end{align*}
-where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{p(\tau|\wt{\vtheta})}{p(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$.
-Similarly to SVRG, we have that $\vtheta_0 := \wt{\vtheta}$ and the update is a FG step.
-Our update is still fundamentally on-policy, since the weighting concerns only the correction term.
+where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in Eq.~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{p(\tau|\wt{\vtheta})}{p(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$.
+Similarly to SVRG, we have that $\vtheta_0 := \wt{\vtheta}$, and the update is a FG step.
+Our update is still fundamentally on-policy since the weighting concerns only the correction term.
 However, this partial ``off-policyness'' is enough to introduce variance.
@@
 \end{align}
-We want just to stress that $\EVV[\tau_i \sim p(\cdot|\vtheta_t)]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to App.~\ref{} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}
-This property will be used to prove Lem.~\ref{L:svrpg.properties}.
+We want just to stress that $\EVV[\tau_i \sim p(\cdot|\vtheta_t)]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to Appendix~\ref{} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}
+This property will be used to prove Lemma~\ref{L:svrpg.properties}.
 %\begin{align*}
@@
 The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}.
-It is easy to show that the SVRPG estimator has the following, desirable properties (see App.~\ref{app:proofs}):
+It is easy to show that the SVRPG estimator has the following, desirable properties (see Appendix~\ref{app:proofs}):
 \begin{lemma}\label{L:svrpg.properties} \todopirout{extended to any unbiased estimator}
@@
 %\end{restatable}
-In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can actually achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
-The resulting SVRPG policy optimization method is detailed in Alg.~\ref{alg:svrpg}.\footnote{Note that when $t=0$ we can avoid to sample trajectories and instead perform directly the policy update using the FG.
-For sake of clarity, we have not explicitly differentiate this step from the standard update rule in Alg.~\ref{alg:svrpg}.}
-
-To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient, the mini-batch size $B$, and the step size $\alpha$.
-Adaptive selection of these meta-parameters is also possible, and is explored in Sec.~\ref{sec:stopping}.
+In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
+Algorithm~\ref{alg:svrpg} presents the details of the SVRPG method for policy optimization.\footnote{Note that when $t=0$ we can avoid to sample trajectories and instead directly perform the policy update using the FG.
+For the sake of clarity, we have not explicitly differentiated this step from the standard update rule in Algorithm~\ref{alg:svrpg}.}
+
+To use SVRPG in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient, the mini-batch size $B$, and the step size $\alpha$.
+Adaptive selection of these meta-parameters is also possible, and it is explored in Section~\ref{sec:stopping}.
 However, the next section will provide general converge guarantees for SVRPG \textcolor{red}{with REINFORCE estimator}.
@@
         \STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\tilde{\vtheta}^{s})$
-        \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\ref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
+        \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\eqref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
 		\FOR{$t=0$ {\bfseries to} $m-1$}
@@
 \section{Convergence Guarantees of SVRPG}\label{sec:conv}
-In this section, we state the convergence guarantees for SVRPG {\color{red}with REINFORCE estimator}.\footnote{\color{red}Similar results can be derived for G(PO)MDP, see App.~\ref{app:}.}
+In this section, we state the convergence guarantees for SVRPG {\color{red}with REINFORCE estimator}.\footnote{\color{red}Similar results can be derived for G(PO)MDP, see Appendix~\ref{app:}.}
 We mainly leverage on the recent analysis of \textcolor{red}{non-concave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
-Each of the three challenges presented at the beginning of Sec.~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
-In App.~\ref{app:gauss} we show how the widespread Gaussian policy satisfies these assumptions.
-
-\textit{1) Non-concavity.} A common assumption in this case it to assume the objective function to be $L$-smooth.
-However, in RL we can consider the following assumption which is sufficient for the $L$-smoothness of the objective (see Lem.~\ref{lemma:lsmooth}).
+Each of the three challenges presented at the beginning of Section~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
+In Appendix~\ref{app:gauss} we show how the traditional Gaussian policy satisfies these assumptions.
+
+\textit{1) Non-concavity.} A common assumption, in this case, is to assume the objective function to be $L$-smooth.
+However, in RL we can consider the following assumption which is sufficient for the $L$-smoothness of the objective (see Lemma~\ref{lemma:lsmooth}).
 	\begin{restatable}[On policy derivatives]{assumption}{boundedscore}\label{ass:bounded_score}
@@
 Since we cannot compute an exact full gradient, we require the variance of the estimator to be bounded.
-This assumption is in spirit similar to the one in~\citep{harikandeh2015stopwasting}.
+This assumption is similar in spirit to the one in~\citep{harikandeh2015stopwasting}.
 	\begin{restatable}[On the variance of the gradient estimator]{assumption}{varreinforce}\label{ass:REINFORCE}
@@
 	\begin{restatable}[On the variance of importance weights]{assumption}{varweights}\label{ass:M2}
-		There is a constant $W<\infty$ such that, for each pair of policies encountered in Alg.~\ref{alg:svrpg} and for each possible trajectory,
+		There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm~\ref{alg:svrpg} and for each possible trajectory,
 		\[
@@
 	\end{restatable}
-Differently from Ass.~\ref{ass:bounded_score} and ~\ref{ass:REINFORCE}, Ass.~\ref{ass:M2} can be enforced by a proper handling of the epoch size $m$.
+Differently from Assumptions~\ref{ass:bounded_score} and ~\ref{ass:REINFORCE}, Assumption~\ref{ass:M2} can be enforced by a proper handling of the epoch size $m$.
 
@@
 \begin{restatable}[Convergence of the {\color{red}REINFORCE-}SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
-{\color{red}Consider the REINFORCE estimator in Eq.~\ref{E:policygradient.estimate}.}
-Under Ass.~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Alg.~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
+{\color{red}Consider the REINFORCE estimator in Eq.~\eqref{E:policygradient.estimate}.}
+Under Assumptions~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
 \begin{align*}
@@
 % Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
-Refer to App.~\ref{app:proofs} for a detailed proof involving definition of the constants and the meta-parameter constraints.
+Refer to Appendix~\ref{app:proofs} for a detailed proof involving the definition of the constants and the meta-parameter constraints.
 %The theoretical settings may be too conservative.
-By analysing the upper-bound in Thm.~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on \textcolor{red}{non-concave} SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
+By analysing the upper-bound in Theorem~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on \textcolor{red}{non-concave} SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
 To achieve true convergence, both the batch size $N$ and the mini-batch size $B$ should increase over time. 
@@
 In this sense, advantage over batch gradient ascent can be achieved with properly selected meta-parameters.
-In Sec.~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$.
-Finally, note from the return statement of Algorithm \ref{alg:svrpg} that $J(\vtheta_A)$ can be seen as the average performance of all the policies tried by the algorithm.
-This is particularly meaningful in the context of on-line learning that we are addressing in this paper.
+In Section~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$.
+Finally, from the return statement of Algorithm \ref{alg:svrpg}, it is worth noting that $J(\vtheta_A)$ can be seen as the average performance of all the policies tried by the algorithm.
+This is particularly meaningful in the context of online learning that we are considering in this paper.
 
@@
 In practice, due to the lower variance G(PO)MDP is preferred to REINFORCE.
-For this reason we will use G(PO)MDP for the experiments (see Sec.~\ref{sec:exp}).
+For this reason, we will use G(PO)MDP for the experiments (see Sec.~\ref{sec:exp}).
 \todopir{sketch of proof for GPOMDP in appendix, probably can be unified with the one of REINFORCE}}
@@
 Additionally, we provide variants of SVRPG exploiting variance-reduction techniques for importance weights.
-Despite lacking of theoretical guarantees, we will show that these methods are able to outperform the baseline SVRPG (Alg.~\ref{alg:svrpg}).
-
-\subsection{Meta-Parameters Selection}\label{sec:stopping}
+Despite lacking theoretical guarantees, we will show that these methods can outperform the baseline SVRPG (Algorithm~\ref{alg:svrpg}).
+
+\subsection{Meta-Parameter Selection}\label{sec:stopping}
 % In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
 % At this point, the open question is how to properly select the step size $\alpha$ and the sub-iteration number $m$ in practical applications.
-The step size $\alpha$ is crucial to balance the variance due to the sub-iterations and the variance associated to the estimator at the snapshot.
+The step size $\alpha$ is crucial to balance the variance due to the sub-iterations and the variance associated with the estimator at the snapshot.
 % The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
-The number $m$ of sub-iterations controls the variance introduced by the importance weights.
-Low $m$ values are associated to low variance but induce lots of snapshot points (\ie lots of FG computations).
-High $m$ values allow to move policy $\pi_{\vtheta_t}$ away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing high variance of the importance weights.
+The number $m$ of sub-iterations influences the variance introduced by the importance weights.
+Low $m$ values are associated with low variance but induce lots of snapshot points (\ie lots of FG computations).
+High $m$ values may move policy $\pi_{\vtheta_t}$ far away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing high variance of the importance weights.
 % the farther the distance between the snapshot policy and the current one, the higher the variance of the importance weights.
@@
 \textbf{Adaptive step size.}
-A common way to deal with noisy gradients is to use adaptive strategies to compute the step size.
+A standard way to deal with noisy gradients is to use adaptive strategies to compute the step size.
 ADAptive Moment estimation (ADAM)~\citep{kingma2014adam} stabilizes the parameter update by computing learning rates for each parameter based on an incremental estimate of the gradient variance.
-Due to this feature we would like to incorporate ADAM in the structure of the SVRPG update.
-Recall that SVRPG is characterized by two updates: I) FG update in the snapshot; II) corrected gradient update in the sub-iterations.
+Due to this feature, we would like to incorporate ADAM in the structure of the SVRPG update.
+Recall that SVRPG performs two different updates of the parameters $\vtheta$: I) FG update in the snapshot; II) corrected gradient update in the sub-iterations.
 Given this structure, we suggest to use two ADAM estimators:
@@
         \vtheta^{s}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
-        \vtheta^{s}_{t+1} &= \vtheta^{s}_t + \alpha^{\textsc{SI}}_{s,t} v_t^{s}
-\end{align*}
-where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s,t}$ with the sub-iterations (see App.~\ref{app:practicalsvrpg} for details).
-By doing so we decouple the contribution of the variance introduced by the approximate FG from the one introduced by the sub-iterations.
+        \vtheta^{s}_{t+1} &= \vtheta^{s}_t + \alpha^{\textsc{SI}}_{s,t} v_t^{s},
+\end{align*}
+where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s,t}$ with the sub-iterations (see Appendix~\ref{app:practicalsvrpg} for details).
+By doing so, we decouple the contribution of the variance due to the approximate FG from the one introduced by the sub-iterations.
 Note that these two terms have different orders of magnitude since are estimated with a different number of trajectories ($B \ll N$) and w/o importance weights.
@@
 \textbf{Adaptive Sub-Iterations.}
-It is easy to imagine that a predefined schedule (\eg $m$ fixed in advance or changed with a policy independent process) may poorly perform due to the high variability of the updates.
-In particular, fixed a value $m$ the variance of the updates in the sub-iterations depends on the snapshot policy and sampled trajectories.
-This variability is in part captured by the ADAM estimate.
-Hence, we propose to take a snapshot (\ie interrupt the sub-iterations) when the step size $\alpha^{\textsc{SI}}$ proposed by ADAM for the sub-iterations is smaller than the one for the FG (\ie $\alpha^{\textsc{FG}}$).
+It is easy to imagine that a predefined schedule (\eg $m$ fixed in advance or changed with a policy-independent process) may poorly perform due to the high variability of the updates.
+In particular, given a fixed number of sub-iterations $m$, the variance of the updates in the sub-iterations depends on the snapshot policy and the sampled trajectories.
+Since the ADAM estimate partly captures such variability,  we propose to take a new snapshot (\ie interrupt the sub-iterations) whenever the step size $\alpha^{\textsc{SI}}$ proposed by ADAM for the sub-iterations is smaller than the one for the FG (\ie $\alpha^{\textsc{FG}}$).
 If the latter condition is verified, it amounts to say that the noise in the corrected gradient has overcome the information of the FG.
@@
 \[
-        \text{If }        \frac{\alpha^{\textsc{FG}}}{N} > \frac{\alpha^{\textsc{SI}}}{B} \textbf{ then } \text{take snapshot}
+        \text{If }        \frac{\alpha^{\textsc{FG}}}{N} > \frac{\alpha^{\textsc{SI}}}{B} \textbf{ then } \text{take snapshot,}
 \]
@@
 \subsection{Normalized Importance Sampling}\label{sec:prac}
-As mentioned in section \ref{sec:stopping}, importance weights are an additional source of variance. A common way to cope with this issue is self-normalization.
+As mentioned in Section~\ref{sec:stopping}, importance weights are an additional source of variance. A standard way to cope with this issue is self-normalization.
 This technique can reduce the variance of the importance weights at the cost of introducing some bias. Whether the trade-off is advantageous depends on the specific task.  
-Introducing self-normalization in the context of our algorithm, we switch from equation \ref{E:svrpg.estimate.batch} to:
+Introducing self-normalization in the context of our algorithm, we switch from Eq.~\eqref{E:svrpg.estimate.batch} to:
 \begin{align*}
@@
 \subsection{Baseline}
-As reported in Sec. \ref{subsec:PolicyGradient}, a well known variance reduction tool in Reinforcement Learning is the employment of baselines in policy gradient estimation.
+As reported in Section~\ref{subsec:PolicyGradient}, a popular variance reduction tool in Reinforcement Learning is the employment of baselines in policy gradient estimation.
 This tool is orthogonal to the variance reduction technique proposed in this paper.
-In Section \ref{sec:exp} we try to combine SVRPG with a non-stationary, state-dependent, learned linear baseline like the one used in \cite{duan2016benchmarking}, with the variant that baseline parameters are updated only at the snapshot, with trajectories sampled from $\pi_{\tilde{\vtheta}^s}$.
+In Section~\ref{sec:exp} we try to combine SVRPG with a non-stationary, state-dependent, learned linear baseline like the one used in~\cite{duan2016benchmarking}, with the variant that baseline parameters are updated only at the snapshot, with trajectories sampled from $\pi_{\tilde{\vtheta}^s}$.
 
 \section{Related Work}
-Despite the huge interest received in SL, variance-reduced gradient approaches has not attracted the RL community.
+Despite the considerable interest received in SL, variance-reduced gradient approaches have not attracted the RL community.
 As far as we know, there are just two applications of SVRG in RL.
 The first approach~\citep{du2017svrgpe} aims to exploit SVRG for policy evaluation.
-The policy evaluation problem is simpler than the one faced in this paper (control problem).
+The policy evaluation problem is more straightforward than the one faced in this paper (control problem).
 In particular, since the goal is to evaluate just the performance of a predefined policy, the optimization problem is stationary.
 The setting considered in the paper is the one of policy evaluation by minimizing the empirical mean squared projected Bellman error (MSPBE) with a linear approximation of the value function.
-\citet{du2017svrgpe} shown that this problem can equivalently reformulated as a convex-concave saddle-point problem that is characterized by a finite-sum structure.
+\citet{du2017svrgpe} shown that this problem can be equivalently reformulated as a convex-concave saddle-point problem that is characterized by a finite-sum structure.
 This problem can be solved using a variant of SVRG~\citep{Palaniappan2016svrgsaddle} for which convergence guarantees have been provided.
-In the second approach~\citep{xu2017svrgtrpo}, SVRG is just an efficient approach to solve the optimization problem faced by Trust Region Policy Optimization (TRPO) at each iteration.
+The second approach~\citep{xu2017svrgtrpo} uses SVRG as a practical method to solve the optimization problem faced by Trust Region Policy Optimization (TRPO) at each iteration.
 We can just talk of a direct application of SVRG to a problem having finite-sum structure since no specific structure of the RL problem is exploited.
@@
 Concerning importance weighting techniques, RL has made wide use of them for off-policy problems~\citep[\eg][]{precup2000eligibility,thomas2015high}.
-However, as mentioned before, SVRPG cannot be compared to such methods since it is practically on-policy.
+However, as mentioned before, SVRPG cannot be compared to such methods since it is in all respects an on-policy algorithm.
 Here, importance weighting is just a statistical tool used to preserve the unbiasedness of the corrected gradient.
@@
 \section{Conclusion}
-In this paper we introduce SVRPG, a variant of SVRG specifically designed for RL problems.
+In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.
 The control problem considered in the paper has a series of difficulties non-common in SL.
-Among them non-concavity and approximate estimate of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
+Among them, non-concavity and approximate estimates of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
 Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy dependent.
-We have shown that by exploiting importance weighting techniques it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
-We have additionally shown that under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
+We have shown that by exploiting importance weighting techniques, it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
+We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
 SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
 The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
-In the second case it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
-Finally, we have empirically shown that practical variants of the theoretical SVRPG are able to outperform classical actor-only approaches on several benchmarks.
+In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
+Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on several benchmarks.
 

### 2018-02-07 svrpg.tex

@@
 
-In all the experiments we compete with GPOMDP standard algorithm. GPOMDP uses a number of trajectories per batch that is equal to the number of trajectories used in SVRPG mini-batch size.
-The neural network and the learning rate are the same for GPOMDP and SVRPG.
-In SVRPG we used two different Adam parameters: the first one for Snapshot and the second one for sub-iterations.
-We used the stopping condition defined in Section \ref{sec:stopping}.
-We fixed the maximum number of trajectories that both the algorithms can use as shown in appendices.
-
-The neural networks is composed of one layer with 8 neurons for Cart-Pole environment and by two layers with 32x32 neurons for Swimmer and for Half Cheetah environments.
-In each experiment, the length of episodes is 500 steps.
-
-In particular in our experiments we use this technique in the context of G(PO)MDP defined as "weighted per-decision importance sampling estimator" \cite{precup2000eligibility}.
+\begin{figure}[t]
+    \includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP}
+    \caption{Self-Normalized SVRPG vs GPOMDP in swimmer}
+    \label{fig:swimmerone}
+\end{figure}
+\begin{figure}[t]
+    \includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
+    \caption{Self-Normalized SVRPG vs Non Self-Normalized SVRPG in swimmer}
+    \label{fig:swimmertwo}
+\end{figure}
+\begin{figure}[t]
+    \includegraphics[width=0.5\textwidth]{half_cheetah_self_norm_baseline_SVRPG_vs_GPOMDP100_baseline}
+    \caption{Self-Normalized SVRPG with baseline vs GPOMDP with baseline in half cheetah}
+    \label{fig:halfcheetah}
+\end{figure}
+\begin{figure}[t]
+    \includegraphics[width=0.5\textwidth]{cart_pole_SVRPG_vs_GPOMDP}
+    \caption{SVRPG vs GPOMDP in cart-pole}
+    \label{fig:cartpole}
+\end{figure}
 

### 2018-02-07 svrpg.tex
Working on appendix

--- 2018-02-07 svrpg.tex
@@
 \usepackage{graphicx}
-\graphicspath{ {images/} }
 \usepackage{subfigure}
@@
 \begin{equation}\label{E:svrg.gradient.correction}
-        \blacktriangledown f(\vtheta_{t}) := v_t = \nabla f(\tilde{\vtheta}) + \nabla z(x_i | \vtheta_t) - \nabla z(x_i | \tilde{\vtheta}),
+        \blacktriangledown f(\vtheta_{t}) := v_t = \nabla f(\wt{\vtheta}) + \nabla z(x_i | \vtheta_t) - \nabla z(x_i | \wt{\vtheta}),
 \end{equation} 
@@
 % \begin{align*}
-% 	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
+% 	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \wt{\vtheta})\Reward(\tau_j) \\
 % 		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
-% 		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i),
+% 		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \wt{\vtheta})\Reward(\tau_i),
 % \end{align*}
@@
 %\begin{align*}
-%&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) +\\
+%&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \wt{\vtheta})\Reward(\tau_j) +\\
 %&\quad\frac{1}{B}\left[\sum_{i=0}^{B-1}
 %\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
-%- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i)\right].
+%- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \wt{\vtheta})\Reward(\tau_i)\right].
 %\end{align*}
@@
 		\FOR{$s=0$ {\bfseries to} $S-1$}
-        \STATE $\vtheta_0^{s+1} := \tilde{\vtheta}^{s} = \vtheta_{m}^s$
-        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\tilde{\vtheta}^{s})$
+        \STATE $\vtheta_0^{s+1} := \wt{\vtheta}^{s} = \vtheta_{m}^s$
+        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\wt{\vtheta}^{s})$
         \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\ref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
@@
 %%		\blacktriangledown J(\vtheta_t^{s+1}) = 
-%%		&\gradApp{\tilde{\vtheta}^s}{N} \\
+%%		&\gradApp{\wt{\vtheta}^s}{N} \\
 %%		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
 %%		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
-%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
+%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \wt{\vtheta}^{s}}\Reward(\tau_i)\right]
 %%		\end{align*}
@@
 This tool is orthogonal to the variance reduction technique proposed in this paper.
-In Section \ref{sec:exp} we try to combine SVRPG with a non-stationary, state-dependent, learned linear baseline like the one used in \cite{duan2016benchmarking}, with the variant that baseline parameters are updated only at the snapshot, with trajectories sampled from $\pi_{\tilde{\vtheta}^s}$.
+In Section \ref{sec:exp} we try to combine SVRPG with a non-stationary, state-dependent, learned linear baseline like the one used in \cite{duan2016benchmarking}, with the variant that baseline parameters are updated only at the snapshot, with trajectories sampled from $\pi_{\wt{\vtheta}^s}$.
 
@@
 This comparison is performed using G(PO)MDP as estimator of $\score{\vtheta}{\tau}\Reward(\tau)$.
-For the sake of simplicity and with a little abuse of notation we refer to \ref{E:svrpg.estimate.batch} as SVRPG and to \ref{E:policygradient.estimate} as G(PO)MDP.
-In all the experiments we used gaussian policies and bootsrap confidence intervals $\alpha = 10\% $ over ten different runs for evaluation.
-Each run with a different policy initialization which is kept fixed when changing the algorithm.
-The maximum number of trajectories at hand is fixed for each task.
+
+We want to investigate whether:
+\begin{enumerate}
+\item The algorithm reaches high value of the reward function faster than GPOMDP.
+\item Averagely SVRPG keep the reward function more stable than GPOMDP.
+\end{enumerate}
 
@@
 
-\begin{figure}[t]
-    \includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP}
-    \caption{Self-Normalized SVRPG vs GPOMDP in swimmer}
-    \label{fig:swimmerone}
-\end{figure}
-\begin{figure}[t]
-    \includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
-    \caption{Self-Normalized SVRPG vs Non Self-Normalized SVRPG in swimmer}
-    \label{fig:swimmertwo}
-\end{figure}
-\begin{figure}[t]
-    \includegraphics[width=0.5\textwidth]{half_cheetah_self_norm_baseline_SVRPG_vs_GPOMDP100_baseline}
-    \caption{Self-Normalized SVRPG with baseline vs GPOMDP with baseline in half cheetah}
-    \label{fig:halfcheetah}
-\end{figure}
-\begin{figure}[t]
-    \includegraphics[width=0.5\textwidth]{half_cheetah_self_norm_baseline_SVRPG_vs_GPOMDP100_baseline}
-    \caption{SVRPG vs GPOMDP in cart-pole}
-    \label{fig:cartpole}
-\end{figure}
+In all the experiments we compete with GPOMDP standard algorithm. GPOMDP uses a number of trajectories per batch that is equal to the number of trajectories used in SVRPG mini-batch size.
+The neural network and the learning rate are the same for GPOMDP and SVRPG.
+In SVRPG we used two different Adam parameters: the first one for Snapshot and the second one for sub-iterations.
+We used the stopping condition defined in Section \ref{sec:stopping}.
+We fixed the maximum number of trajectories that both the algorithms can use as shown in appendices.
+
+The neural networks is composed of one layer with 8 neurons for Cart-Pole environment and by two layers with 32x32 neurons for Swimmer and for Half Cheetah environments.
+In each experiment, the length of episodes is 500 steps.
+
+In particular in our experiments we use this technique in the context of G(PO)MDP defined as "weighted per-decision importance sampling estimator" \cite{precup2000eligibility}.
 
@@
 \[
-	\Es{X} = \EVV[\tau_j\sim\pi(\cdot\vert\tilde{\vtheta}^s)
-		\forall j\in\mathbf{N}]{X\vert\tilde{\vtheta}^s},
-\]
-where $\tilde{\vtheta}^s$ is defined in Algorithm \ref{alg:svrpg} and $\mathbf{N} = [0,\dots,N)$.
+	\Es{X} = \EVV[\tau_j\sim\pi(\cdot\vert\wt{\vtheta}^s)
+		\forall j\in\mathbf{N}]{X\vert\wt{\vtheta}^s},
+\]
+where $\wt{\vtheta}^s$ is defined in Algorithm \ref{alg:svrpg} and $\mathbf{N} = [0,\dots,N)$.
 \end{definition}
@@
         \mathbb{E}_{t\vert s}\left[X\right] &\coloneqq 
-		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\vert\tilde{\vtheta}^s)\forall j \in {\it {\bf N}} \\ \tau_{i,h}\sim\pi(\cdot\vert\vtheta^{s+1}_i) \forall h \in {\it {\bf B}}, \text{ for $i=0,\dots,t-1$}}}{\left[X \vert \tilde{\vtheta^s}\right]} \\
-	&\coloneqq \EVV[\tau_j\sim\pi(\cdot\vert\tilde{\vtheta}^s)\forall j \in {\it {\bf N}}]{
+		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\vert\wt{\vtheta}^s)\forall j \in {\it {\bf N}} \\ \tau_{i,h}\sim\pi(\cdot\vert\vtheta^{s+1}_i) \forall h \in {\it {\bf B}}, \text{ for $i=0,\dots,t-1$}}}{\left[X \vert \wt{\vtheta^s}\right]} \\
+	&\coloneqq \EVV[\tau_j\sim\pi(\cdot\vert\wt{\vtheta}^s)\forall j \in {\it {\bf N}}]{
 			\EVV[\tau_{0,h}\sim\pi(\cdot\vert\vtheta_0^{s+1}) \forall h \in {\it {\bf B}}]
@@
 			\vert\vtheta_0^{s+1}}
-		\vert\tilde{\vtheta}^s}\\
+		\vert\wt{\vtheta}^s}\\
         &= \EVV[t-1|s]{\EVV[\tau_{t,h}\sim \pi(\cdot|\vtheta^{s+1}_t)]{X|\vtheta^{s+1}_t}}
 \end{align*}
-where the sequence $\tilde{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}, ${\it {\bf N}} = [0,\dots,N)$, and ${\it {\bf B}} = [0,\dots,B)$. 
+where the sequence $\wt{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}, ${\it {\bf N}} = [0,\dots,N)$, and ${\it {\bf B}} = [0,\dots,B)$. 
 \todopir{what about $\vtheta^{s+1}_0$ and $\wt{\vtheta}^s$? They are equal. How this affect previous definition?}
@@
 \end{definition}
-Intuitively, the $\Ets{\cdot}$ operator computes the expected value with respect to the sampling of trajectories from the snapshot $\tilde{\vtheta}^s$ up to the $t$-th iteration included.
+Intuitively, the $\Ets{\cdot}$ operator computes the expected value with respect to the sampling of trajectories from the snapshot $\wt{\vtheta}^s$ up to the $t$-th iteration included.
 Note that the order in which expected values are taken is important since each $\vtheta_{t}^{s+1}$ is function of previously sampled trajectories and is used to sample new ones.
@@
 \[
-	e_s \coloneqq \gradApp{\tilde{\vtheta}^s}{N} - \gradJ{\tilde{\vtheta}^s} 
+	e_s \coloneqq \gradApp{\wt{\vtheta}^s}{N} - \gradJ{\wt{\vtheta}^s} 
 \]
@@
 	\gradIdeal{\vtheta_t^{s+1}} &\coloneqq 
-	\gradJ{\tilde{\vtheta}^s}
+	\gradJ{\wt{\vtheta}^s}
 %	+ \nabla\log\pi(\tau_i \vert \vtheta_t^{s+1})\Reward(\tau_i) 
     + g(\tau_i|\vtheta^{s+1}_t)
-%	- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta}^s)\Reward(\tau_i) 
+%	- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \wt{\vtheta}^s)\Reward(\tau_i) 
     - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i|\wt{\vtheta}^s)
     \\
-	&= \gradBlack{\vtheta_t^{s+1}} - \gradApp{\tilde{\vtheta}^s}{N} + \gradJ{\tilde{\vtheta}^s} \\
+	&= \gradBlack{\vtheta_t^{s+1}} - \gradApp{\wt{\vtheta}^s}{N} + \gradJ{\wt{\vtheta}^s} \\
 	&= \gradBlack{\vtheta_t^{s+1}} - e_s
@@
 \begin{align*}
-        \EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\tilde{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} - \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
+        \EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} - \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
 &= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
@@
 \begin{proof}
-As $\vtheta\to\vtheta^*$, also $\tilde{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
+As $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
 \begin{align*}
@@
 \Ets[t-1]{\norm[2]{\gradJ{\vtheta_t^{s+1}}}^2} 
-+\frac{L_2^2}{B}\Ets[t-1]{\norm[2]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
++\frac{L_2^2}{B}\Ets[t-1]{\norm[2]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
++\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}
 \nonumber 
@@
             \mathbb{E}_{t|s}\big[&\norm[2]{\gradBlack{\vtheta_t^{s+1}}}^2\big]
-	= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
+	= \Ets{\norm[]{\gradApp{\wt{\vtheta}^s}{N}
 			+\frac{1}{B}\sum_{i=0}^{B-1} g(\tau_i\vert\vtheta_t^{s+1}) 
 			-\frac{1}{B}\sum_{i=0}^{B-1}
-    \omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
-	%
-	&= \mathbb{E}_{t\vert s}\left[\left\|\gradApp{\tilde{\vtheta}^s}{N}
+    \omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)}^2} \nonumber\\
+	%
+	&= \mathbb{E}_{t\vert s}\left[\left\|\gradApp{\wt{\vtheta}^s}{N}
 			+\frac{1}{B}\sum_{i=0}^{B-1}\left( 
 			g(\tau_i\vert\vtheta_t^{s+1}) -
-			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)
+			\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)\right)
 			% \right.\right.\nonumber
             % \\&\qquad\left.\left.
-            \pm \gradJ{\vtheta_t^{s+1}} \pm \gradJ{\tilde{\vtheta}^s} 
-			% -\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
-			% +\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}
+            \pm \gradJ{\vtheta_t^{s+1}} \pm \gradJ{\wt{\vtheta}^s} 
+			% -\gradJ{\vtheta_t^{s+1}} + \gradJ{\wt{\vtheta}^s}
+			% +\gradJ{\vtheta_t^{s+1}} - \gradJ{\wt{\vtheta}^s}
     \right\|^2\right] \nonumber\\
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
-	+\Es{\norm[]{\gradApp{\tilde{\vtheta}^s}{N} - \Es{\gradApp{\tilde{\vtheta}^s}{N}}}^2} \nonumber\\
+	+\Es{\norm[]{\gradApp{\wt{\vtheta}^s}{N} - \Es{\gradApp{\wt{\vtheta}^s}{N}}}^2} \nonumber\\
 	&\qquad+ 
@@
 		g(\tau_i\vert\vtheta_t^{s+1}) -
-			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)
+			\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)\right)
 	%	\right.\right.\nonumber\\&\qquad\left.\left.
@@
 			g(\tau_i\vert\vtheta_t^{s+1}) -
-				\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}\right\|^2\right] 
+				\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)\right)}\right\|^2\right] 
 	\nonumber\\
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
+	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}
 	\nonumber\\
@@
 		g(\tau_i\vert\vtheta_t^{s+1}) -
-		\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)
+		\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)\right)
 		% \right.\right.\nonumber\\&\qquad\left.\left.
@@
 			g(\tau_i\vert\vtheta_t^{s+1}) -
-			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}\right\|^2\right] 
+			\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)\right)}\right\|^2\right] 
 		\label{eq:1}\\%
 	&\leq \Ets[t-1]{\norm[2]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} 
+	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} 
 %    \nonumber\\ &\qquad
@@
 			g(\tau_i\vert\vtheta_t^{s+1}) -
-			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}^2} \label{eq:2}\\
+			\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)\right)}^2} \label{eq:2}\\
 	%
 	&\leq \Ets[t-1]{\norm[2]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
+	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}
 %     \nonumber\\ 	&\qquad
@@
 			g(\tau_i\vert\vtheta_t^{s+1}) -
-    \omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s) \pm g(\tau_i|\wt{\vtheta}^s) }^2} \nonumber\\
+    \omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s) \pm g(\tau_i|\wt{\vtheta}^s) }^2} \nonumber\\
 	%
 %%	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-%%	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
+%%	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}
 %%	\nonumber\\
@@
 %%			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
-%%			-g(\tau_i\vert\tilde{\vtheta}^s)
-%%			+g(\tau_i\vert\tilde{\vtheta}^s) 
-%%			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
+%%			-g(\tau_i\vert\wt{\vtheta}^s)
+%%			+g(\tau_i\vert\wt{\vtheta}^s) 
+%%			-\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)}^2} \nonumber\\
 	%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
+	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}
 	\nonumber\\
@@
 			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
-			-g(\tau_i\vert\tilde{\vtheta}^s)}^2} 
+			-g(\tau_i\vert\wt{\vtheta}^s)}^2} 
 %        \nonumber\\ 	&\qquad
 			+\frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[]{g(\tau_i\vert\tilde{\vtheta}^s) 
-			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\
+			\Ets{\norm[]{g(\tau_i\vert\wt{\vtheta}^s) 
+			-\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)}^2} \nonumber\\
 	%
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
+	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}
 	\nonumber\\
 	&\qquad
-    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
     +
@@
 		\Ets{\norm[]{(1 
-			-\omega(\tau_i))g(\tau_i\vert\tilde{\vtheta}^s)}^2} \label{eq:3}\\
+			-\omega(\tau_i))g(\tau_i\vert\wt{\vtheta}^s)}^2} \label{eq:3}\\
 % ...
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
+	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}
 	\nonumber\\
-	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
 	+\Gamma\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \label{eq:4}\\
@@
 	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
+	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}
 %	\nonumber\\&\qquad
-    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
 	+\frac{\Gamma}{B^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
@@
 	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
-	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
+	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}
 % 	\nonumber\\ &\qquad
-    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
 	+\frac{\Gamma\VARIS}{B}, \label{eq:5}
@@
 \leq
-\frac{1}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2}
-\end{align*}
+\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2}
+\end{align*}
+\todopir{I have to check this one!!!}
 \end{restatable}
@@
 	\dotprod{\Ets{e_s}}{\Ets{\varphi(\vtheta_t^{s+1})}}
-	+\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}  \label{eq:new1}\\
+	+\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}  \label{eq:new1}\\
 	&= 
@@
 	&\qquad+
-	\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})} \label{eq:8}
+	\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})} \label{eq:8}
 \end{align}
@@
 	=
-	\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}\right|  
+	\left|\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}\right|  
 	\nonumber\\
 	&\leq
-	\sqrt{\Vars{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{\varphi(\vtheta_t^{s+1})}} \nonumber\\
+	\sqrt{\Vars{\gradApp{\wt{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{\varphi(\vtheta_t^{s+1})}} \nonumber\\
 	%
 	&\leq	
-	\frac{1}{2}\Vars{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})}\label{eq:9}\\
+	\frac{1}{2}\Vars{\gradApp{\wt{\vtheta}^s}{N}} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})}\label{eq:9}\\
 	%
 	&=
-	\frac{1}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})} \label{eq:10}\\
+	\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})} \label{eq:10}\\
 	%
 	&\leq
-	\frac{1}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2},
+	\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2},
 	\nonumber
@@
 where (\ref{eq:6}) is from Definition \ref{def:ideal}; (\ref{eq:7}) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $\varphi(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$;
-(\ref{eq:new1}) is from the fact that $\gradJ{\tilde{\vtheta}^s}$ is constant \wrt $\Vars{\cdot}$;
+(\ref{eq:new1}) is from the fact that $\gradJ{\wt{\vtheta}^s}$ is constant \wrt $\Vars{\cdot}$;
 (\ref{eq:8}) is from $\Ets{e_s}=0$; (\ref{eq:9}) is from Young's inequality; and (\ref{eq:10}) is from the definition of $\gradApp{\vtheta}{N}$.
@@
 \begin{align*}
-	&R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \\
+	&R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}, \\
 	&c_{m-1} = 0, \\
@@
 	&\qquad-
-	\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} -\frac{\alpha_t}{2}\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}, \label{eq:13}
+	\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} -\frac{\alpha_t}{2}\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}, \label{eq:13}
 \end{align}
-where (\ref{eq:11}) is from the L-smoothness of $J(\vtheta)$, (\ref{eq:12}) is from the SVRPG update, and (\ref{eq:13}) is from Lemma \ref{lemma:aux0}.
+where~\eqref{eq:11} is from the L-smoothness of $J(\vtheta)$ and \eqref{eq:12} is from the SVRPG update.
+{\color{blue} Inequality~\ref{eq:13} follows from Lemma~\ref{lemma:aux0} by noticing that $\nabla J(\vtheta^{s+1}_t)$ is a deterministic function given $\vtheta^{s+1}_t$.
+As a consequence we can directly apply Lemma~\ref{lemma:aux0} to lower-bound $\EVV[t|s]{\langle \nabla J(\vtheta^{s+1}_t, \blacktriangledown J(\vtheta^{s+1}_t \rangle}$ by $\norm[2]{\nabla J(\vtheta^{s+1}_t}^2 - f(\vtheta^{s+1}_t, \wt{\vtheta}^s)$.}
 
 Next we have:
+\begingroup
+\allowdisplaybreaks
 \begin{align}
-\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} 
-&= \Ets{\norm[]{\vtheta_{t+1}^{s+1}- \vtheta_t^{s+1} + \vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
-&=\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_{t}^{s+1}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\dotprod{\vtheta_{t+1}^{s+1}-\vtheta_{t}^{s+1}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber \\
-&= \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \label{eq:14}\\
-&\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2+2\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber\\ 
+\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\wt{\vtheta}^s}^2} 
+&= \Ets{\norm[]{\vtheta_{t+1}^{s+1}- \vtheta_t^{s+1} + \vtheta_t^{s+1}-\wt{\vtheta}^s}^2} \nonumber\\
+&=\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_{t}^{s+1}}^2+\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2+2\dotprod{\vtheta_{t+1}^{s+1}-\vtheta_{t}^{s+1}}{\vtheta_t^{s+1}-\wt{\vtheta}^s}} \label{eq:14a} \\
+&= \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2+2\alpha_t\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\wt{\vtheta}^s}} \label{eq:14}\\
+&\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2+2\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\wt{\vtheta}^s}} \nonumber\\ 
 &\qquad+
-\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \label{eq:15}\\
-%
-&\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-+2\alpha_t\Ets{\left|\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\tilde{\vtheta}^s}\right|} \nonumber\\ 
+\frac{\alpha_t}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2} \label{eq:15}\\
+%
+&\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
++2\alpha_t\Ets{\left|\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\wt{\vtheta}^s}\right|} \nonumber\\ 
 &\qquad+
-\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
-%
-&\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-+2\alpha_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}} \nonumber\\ 
+\frac{\alpha_t}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2} \nonumber\\
+%
+&\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
++2\alpha_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}} \nonumber\\ 
 &\qquad+
-\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
-%
-&\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
-+2\alpha_t\Ets{\frac{1}{2\beta}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta}{2}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\ 
+\frac{\alpha_t}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2} \nonumber\\
+%
+&\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
++2\alpha_t\Ets{\frac{1}{2\beta}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta}{2}\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2} \label{eq:16a}\\ 
 &\qquad
-\frac{\alpha_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \label{eq:16}
++\frac{\alpha_t}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\alpha_t\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}, \label{eq:16}
 \end{align}
-where (\ref{eq:14}) is from the SVRPG update, (\ref{eq:15}) is from Lemma \ref{lemma:aux0}, and (\ref{eq:16}) is from Young's inequality in the 'Peter-Paul' variant.
+\endgroup
+where~\eqref{eq:14a} is obtained using the triangular inequality,~\eqref{eq:14} is from the SVRPG update,~\eqref{eq:15} is from Lemma~\ref{lemma:aux0},~\eqref{eq:16a} is from Cauchy-Schwarz inequality, and~\eqref{eq:16} is from Young's inequality in the `Peter-Paul' variant.
 Finally:
+\begingroup
+\allowdisplaybreaks
 \begin{align}
@@
 \end{align}
+\endgroup
 where (\ref{eq:17}) is from (\ref{eq:13}), (\ref{eq:18}) is from (\ref{eq:16}), (\ref{eq:19}) is from Lemma \ref{lemma:aux2}, and (\ref{eq:20}) is from Assumption \ref{ass:REINFORCE}.

### 2018-02-07 svrpg.tex

--- 2018-02-07 svrpg.tex
@@
 		\FOR{$s=0$ {\bfseries to} $S-1$}
-        \STATE $\vtheta_0^{s+1} := \tilde{\vtheta}^{s} = \vtheta_{m}^s$
-        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\tilde{\vtheta}^{s})$
+        \STATE $\vtheta_0^{s+1} := \wt{\vtheta}^{s} = \vtheta_{m}^s$
+        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\wt{\vtheta}^{s})$
         \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\eqref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
@@
 This tool is orthogonal to the variance reduction technique proposed in this paper.
-In Section~\ref{sec:exp} we try to combine SVRPG with a non-stationary, state-dependent, learned linear baseline like the one used in~\cite{duan2016benchmarking}, with the variant that baseline parameters are updated only at the snapshot, with trajectories sampled from $\pi_{\tilde{\vtheta}^s}$.
+In Section~\ref{sec:exp} we try to combine SVRPG with a non-stationary, state-dependent, learned linear baseline like the one used in~\cite{duan2016benchmarking}, with the variant that baseline parameters are updated only at the snapshot, with trajectories sampled from $\pi_{\wt{\vtheta}^s}$.
 

### 2018-02-08 svrpg.tex
Updated
experiments section corrections

@@
 The maximum number of trajectories at hand is fixed for each task.
-
-In order to pursue the above mentioned objectives, we evaluate the behavior of the reward function averaged on 10 runs for each experiment.
-
-We decide to use the following models of Rllab: \emph{Cart-Pole Balancing}, \emph{Mujoco Swimmer}, \emph{Mujoco Half Cheetah}.
+In \emph{figure} \ref{fig:swimmerone} we can see how SVRPG outperforms G(PO)MDP over the swimmer task.
+Clearly the mean is nearly always higher for our algorithm, moreover near the end we see that te confidence intervals are also disjoint.
+In \emph{figure} \ref{fig:swimmertwo} we addressed the comparison between SVRPG and its self-normalized version still for the swimmer task.
+We can see an inprovement on the average return, but the confidence intervals are never disjoint.
+In \emph{figure} \ref{fig:halfcheetah} the difference between our algorithm and G(PO)MDP over the half-cheetah task is not statistically significant (both of them use the same baseline even though is not optimal for SVRPG see \ref{} for further hints over this topic), however we can see a certain improvement on the mean of the objective function.
+Eventually in \emph{figure} \ref{fig:cartpole} we can clearly see that SVRPG outperforms GPOMDP in the cart-pole task.
 
@@
 
-\newpage
-\mbox{}
-\newpage
+\clearpage
 \onecolumn
@@
 As a consequence we can directly apply Lemma~\ref{lemma:aux0} to lower-bound $\EVV[t|s]{\langle \nabla J(\vtheta^{s+1}_t, \blacktriangledown J(\vtheta^{s+1}_t \rangle}$ by $\norm[2]{\nabla J(\vtheta^{s+1}_t}^2 - f(\vtheta^{s+1}_t, \wt{\vtheta}^s)$.}
+\todopir{is \eqref{eq:11} true just for smooth functions? Does it require convexity/concavity? Ref?}
 

### 2018-02-08 svrpg.tex
Modified experiments in appendices and cited precup in section 5.2

@@
 \subsection{Normalized Importance Sampling}\label{sec:prac}
-As mentioned in Section~\ref{sec:stopping}, importance weights are an additional source of variance. A standard way to cope with this issue is self-normalization.
+As mentioned in Section~\ref{sec:stopping}, importance weights are an additional source of variance. A standard way to cope with this issue is self-normalization \citep{precup2000eligibility}.
 This technique can reduce the variance of the importance weights at the cost of introducing some bias. Whether the trade-off is advantageous depends on the specific task.  
@@
 In Section \ref{sec:exp} we show some cases in which self-normalization provides a performance improvement.
-\todomat{add reference, maybe just Precup}
 
@@
 	Snapshot trajectories & 100 & 100 & 100 \\
-	Sub-iterations trajectories/GPOMDP batch & 10 & 10 & 10 \\
+	Sub-iterations trajectories & 10 & 10 & 10 \\
 	Max sub-iterations & 50 & 20 & 20 \\

### 2018-02-08 svrpg.tex
Updated

--- 2018-02-07 svrpg.tex
@@
 Let $\tau = \{\langle s_t,a_t \rangle\}_{t=0}^{H}= \{z_t\}_{t=0}^{H} = z_{0:H}$ is a $(H+1)$-steps trajectory.
-Given that $\tau$ depends on the MDP $M=\{\mathcal{S},\mathcal{A}, p, r, \gamma, \rho\}$ and the actual policy $\pi$, the trajectory is
-said drawn from density distribution $\Delta(\tau|\pi,M)$ defined as:
-\[
-        \Delta(\tau|\pi,M) = \rho(s_0) \pi(z_0) \prod_{k=1}^{H} p(s_k|z_{k-1})\pi(z_k).
+Given that $\tau$ depends on the MDP $M=\{\mathcal{S},\mathcal{A}, \mathcal{P}, \Reward, \gamma, \rho\}$ and the actual policy $\pi$, the trajectory is
+said drawn from density distribution $p(\tau|\pi,M)$ defined as:
+\[
+        p(\tau|\pi,M) = \rho(s_0) \pi(z_0) \prod_{k=1}^{H} \mathcal{P}(s_k|z_{k-1})\pi(z_k).
 \]
@@
 \[
-        J(\pol) = \EVV[\tau \sim \Delta(\cdot|\pol)]{\Reward(\tau)|M},
-\]
-where $\Reward(\tau) = \sum_{t=0}^{H}\gamma^t r(z_t)$.
+        J(\pol) = \EVV[\tau \sim p(\cdot|\pol)]{\Reward(\tau)|M},
+\]
+where $\Reward(\tau) = \sum_{t=0}^{H}\gamma^t \Reward(z_t)$.
 The policy gradient $\nabla J(\vtheta)$ is
 \begin{equation}\label{E:onpolicygradient}
-        \gradJ{\vtheta} = \EVV[\tau \sim \Delta(\cdot|\pol)]{\nabla \log\Delta(\tau|\vtheta)\Reward(\tau)} = \EVV[\tau \sim \Delta(\cdot|\pol)]{\sum_{j=0}^{H} \gamma^j \mathcal{R}(z_j) \sum_{i=0}^{H} \nabla \log\pi(z_i)}.
+        \gradJ{\vtheta} = \EVV[\tau \sim p(\cdot|\pol)]{\nabla \log\Delta(\tau|\vtheta)\Reward(\tau)} = \EVV[\tau \sim p(\cdot|\pol)]{\sum_{j=0}^{H} \gamma^j \mathcal{R}(z_j) \sum_{i=0}^{H} \nabla \log\pi(z_i)}.
 \end{equation}
@@
 \underbracket{
-\left(\sum_{h=0}^{H}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\sum_{h=0}^{H}\gamma^h r(s_h^n,a_h^n) - b\right)
-}_{g(\tau_n|\vtheta):=\nabla \log\Delta(\tau_n|\vtheta)\Reward(\tau_n)}
+        \left(\sum_{h=0}^{H}\nabla \log \pol(a_h^n\vert s_h^n) \right)\left(\sum_{h=0}^{H}\gamma^h \Reward(s_h^n,a_h^n) - b\right)
+}_{g(\tau_n|\vtheta):=\nabla \log p(\tau_n|\vtheta)\Reward(\tau_n)}
 ,
@@
 \underbracket{
-\sum_{h=0}^{H}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\gamma^h r(s_h^n,a_h^n) - b\right)
+        \sum_{h=0}^{H}\left(\sum_{k=0}^{h} \nabla \log \pol (a_h^n\vert s_h^n) \right)\left(\gamma^h \Reward(s_h^n,a_h^n) - b\right)
 }_{g(\tau_n|\vtheta)}.
@@
 \[
-        J(\pi^T) = \EVV[\tau \sim \Delta(\cdot|\pi^T)]{\mathcal{R}(\tau)} = \EVV[\tau \sim \Delta(\cdot|\pi^B)]{\omega(\tau) \mathcal{R}(\tau)}
-\]
-where $\omega(\tau|\pi^B,\pi^T) = \frac{\Delta(\tau,\pi^T)}{\Delta(\tau|\pi^B)} = \omega(z_{0:H}|\pi^B,\pi^T) = \prod_{w=0}^{H} \omega(z_w|\pi^B,\pi^T)$ and $\omega(z_w|\pi^B,\pi^T) = \frac{\pi^T(a_w|s_w)}{\pi^B(a_w|s_w)}$
+        J(\pi^T) = \EVV[\tau \sim p(\cdot|\pi^T)]{\mathcal{R}(\tau)} = \EVV[\tau \sim p(\cdot|\pi^B)]{\omega(\tau) \mathcal{R}(\tau)}
+\]
+where $\omega(\tau|\pi^B,\pi^T) = \frac{p(\tau,\pi^T)}{p(\tau|\pi^B)} = \omega(z_{0:H}|\pi^B,\pi^T) = \prod_{w=0}^{H} \omega(z_w|\pi^B,\pi^T)$ and $\omega(z_w|\pi^B,\pi^T) = \frac{\pi^T(a_w|s_w)}{\pi^B(a_w|s_w)}$
 
@@
 \begin{equation}\label{E:offpolicygradient}
-        \nabla J(\pi^T) = \EVV[\tau \sim \Delta(\cdot|\pi^B)]{\omega(\tau|\pi^B,\pi^T) \nabla \log \Delta(\tau|\pi^T) \mathcal{R}(\tau)} = \EVV[\tau \sim \Delta(\cdot|\pi^B)]{\omega(\tau|\pi^B,\pi^T) g(\tau|\pi^T)}.
+        \nabla J(\pi^T) = \EVV[\tau \sim p(\cdot|\pi^B)]{\omega(\tau|\pi^B,\pi^T) \nabla \log p(\tau|\pi^T) \mathcal{R}(\tau)} = \EVV[\tau \sim p(\cdot|\pi^B)]{\omega(\tau|\pi^B,\pi^T) g(\tau|\pi^T)}.
 \end{equation}
-For $\omega(\tau)$ being well defined the behavioural policy should have non-zero probability of selecting any action in every state \ie $\pi^B(a|s) > 0$ for any $(s,a)\in \mathcal{S} \times \mathcal{A}$.
+For $\omega(\tau|\pi^B,\pi^T)$ being well defined the behavioural policy should have non-zero probability of selecting any action in every state \ie $\pi^B(a|s) > 0$ for any $(s,a)\in \mathcal{S} \times \mathcal{A}$.
 Equation~\ref{E:offpolicygradient} is important for proving Theorem~\ref{theo:convergence} since it provides a common representation of REINFORCE and G(PO)MDP.
@@
 \[
-        \nabla J(\pi^T) = \frac{1}{N} \sum_{n=1}^{N} \omega(\tau^n)
-        \left(\sum_{h=0}^{H}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\sum_{h=0}^{H}\gamma^h r(s_h^n,a_h^n)\right).
+        \nabla J(\pi^T) = \frac{1}{N} \sum_{n=1}^{N} \omega(\tau^n|\pi^B, \pi^T)
+        \left(\sum_{h=0}^{H} \nabla \log \pi^{T}(a_h^n\vert s_h^n) \right)\left(\sum_{h=0}^{H}\gamma^h \Reward(s_h^n,a_h^n)\right).
 \]
@@
         \nabla J(\pi^T) = \frac{1}{N} \sum_{n=1}^{N}
-        \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k)\right) \gamma^h r(z_h^n) \omega(z_{0:h})
+        \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) \gamma^h \Reward(z_h^n) \omega(z_{0:h}|\pi^B,\pi^T)
 \]
@@
 {\color{blue} Inequality~\ref{eq:13} follows from Lemma~\ref{lemma:aux0} by noticing that $\nabla J(\vtheta^{s+1}_t)$ is a deterministic function given $\vtheta^{s+1}_t$.
-As a consequence we can directly apply Lemma~\ref{lemma:aux0} to lower-bound $\EVV[t|s]{\langle \nabla J(\vtheta^{s+1}_t, \blacktriangledown J(\vtheta^{s+1}_t \rangle}$ by $\norm[2]{\nabla J(\vtheta^{s+1}_t}^2 - f(\vtheta^{s+1}_t, \wt{\vtheta}^s)$.}
+As a consequence we can directly apply Lemma~\ref{lemma:aux0} to lower-bound $\EVV[t|s]{\langle \nabla J(\vtheta^{s+1}_t), \blacktriangledown J(\vtheta^{s+1}_t) \rangle}$ by $\norm[2]{\nabla J(\vtheta^{s+1}_t}^2 - f(\vtheta^{s+1}_t, \wt{\vtheta}^s)$.}
+\todopir{is \eqref{eq:11} true just for smooth functions? Does it require convexity/concavity? Ref?}
 

### 2018-02-08 svrpg.tex
Updated

@@
 \begin{align}
-\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\wt{\vtheta}^s}^2} 
-&= \Ets{\norm[]{\vtheta_{t+1}^{s+1}- \vtheta_t^{s+1} + \vtheta_t^{s+1}-\wt{\vtheta}^s}^2} \nonumber\\
+        \mathbb{E}_{t|s}\bigg[ & \norm[]{\vtheta_{t+1}^{s+1}-\wt{\vtheta}^s}^2 \bigg]
+= \Ets{\norm[]{\vtheta_{t+1}^{s+1}- \vtheta_t^{s+1} + \vtheta_t^{s+1}-\wt{\vtheta}^s}^2} \nonumber\\
 &=\Ets{\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_{t}^{s+1}}^2+\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2+2\dotprod{\vtheta_{t+1}^{s+1}-\vtheta_{t}^{s+1}}{\vtheta_t^{s+1}-\wt{\vtheta}^s}} \label{eq:14a} \\
@@
 &\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
-+2\alpha_t\Ets{\frac{1}{2\beta}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta}{2}\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2} \label{eq:16a}\\ 
++2\alpha_t\Ets{\frac{1}{2\beta_t}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta_t}{2}\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2} \label{eq:16a}\\ 
 &\qquad
@@
 where~\eqref{eq:14a} is obtained using the triangular inequality,~\eqref{eq:14} is from the SVRPG update,~\eqref{eq:15} is from Lemma~\ref{lemma:aux0},~\eqref{eq:16a} is from Cauchy-Schwarz inequality, and~\eqref{eq:16} is from Young's inequality in the `Peter-Paul' variant.
-Finally:
+Let's us consider the following function:
+\[
+	R_{t+1}^{s+1} \coloneqq \Ets{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2}. 
+\]
+The objective is now to provide a lower bound to it.
 \begingroup
@@
 \begin{align}
-	R_{t+1}^{s+1} &= \Ets{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
-	%
-	&\geq	\Ets{J(\vtheta_t^{s+1})+\frac{\alpha_t}{2}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	R_{t+1}^{s+1} 
+    % &= \Ets{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
+	%
+	%&\geq	\Ets{J(\vtheta_t^{s+1})+\frac{\alpha_t}{2}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
+	&\geq	\Ets{J(\vtheta_t^{s+1}) - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}
+    + \EVV[t-1|s]{\frac{\alpha_t}{2}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
+    \nonumber\\
 	&\qquad-
@@
 	\nonumber\\
-	&\qquad-2c_{t+1}\alpha_t\Ets{\frac{1}{2\beta}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta}{2}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\ 
+	&\qquad-2c_{t+1}\alpha_t\Ets{\frac{1}{2\beta_t}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta_t}{2}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\ 
 	&\qquad
@@
 	&\qquad-\alpha_t^2\left(\frac{L}{2}+c_{t+1}\right)\Ets{\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}
-	+\frac{\alpha_t}{2}\left(1-\frac{2c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	+\frac{\alpha_t}{2}\left(1-\frac{2c_{t+1}}{\beta_t}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	%
-	&\geq  \Ets{J(\vtheta_t^{s+1})} - c_{t+1}\left(1+\alpha_t\beta+\alpha_t\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	&\geq  \Ets{J(\vtheta_t^{s+1})} - c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad
@@
 	+\frac{\Gamma W}{B}\right)
-	+\frac{\alpha_t}{2}\left(1-2\frac{c_{t+1}}{\beta}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	+\frac{\alpha_t}{2}\left(1-2\frac{c_{t+1}}{\beta_t}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \label{eq:19}\\
 	%
-	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(1+\alpha_t\beta+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad
-	+\alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	+\alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
@@
 \endgroup
-where (\ref{eq:17}) is from (\ref{eq:13}), (\ref{eq:18}) is from (\ref{eq:16}), (\ref{eq:19}) is from Lemma \ref{lemma:aux2}, and (\ref{eq:20}) is from Assumption \ref{ass:REINFORCE}.
+where~\eqref{eq:17} is from~\eqref{eq:13} noticing that $\EVV[t|s]{\norm[]{\nabla J(\vtheta^{s+1}_t)}^2} = \EVV[t|s]{\norm[]{\nabla J(\vtheta^{s+1}_t)}^2}$,~\eqref{eq:18} is from~\eqref{eq:16},~\eqref{eq:19} is from Lemma~\ref{lemma:aux2}, and (\ref{eq:20}) is from Assumption \ref{ass:REINFORCE}.
 To complete the proof, besides rearranging terms, we have to ensure that $\Psi_t>0$ for each $t$. This gives the constraints on $\alpha_t$ and $\beta_t$.

### 2018-02-08 svrpg.tex
fixed expected values in Lemma 9

@@
 \[
-	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
+	\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
 	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_tV}{N\Psi_t}
@@
 \begin{align*}
-	&R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}, \\
+	&R_t^{s+1}\coloneqq \Ets[t-1]{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}, \\
 	&c_{m-1} = 0, \\
@@
 \end{align}
-where~\eqref{eq:11} is from the L-smoothness of $J(\vtheta)$ and \eqref{eq:12} is from the SVRPG update.
-{\color{blue} Inequality~\ref{eq:13} follows from Lemma~\ref{lemma:aux0} by noticing that $\nabla J(\vtheta^{s+1}_t)$ is a deterministic function given $\vtheta^{s+1}_t$.
-As a consequence we can directly apply Lemma~\ref{lemma:aux0} to lower-bound $\EVV[t|s]{\langle \nabla J(\vtheta^{s+1}_t), \blacktriangledown J(\vtheta^{s+1}_t) \rangle}$ by $\norm[2]{\nabla J(\vtheta^{s+1}_t}^2 - f(\vtheta^{s+1}_t, \wt{\vtheta}^s)$.}
-\todopir{is \eqref{eq:11} true just for smooth functions? Does it require convexity/concavity? Ref?}
+where~\eqref{eq:11} is from the L-smoothness of $J(\vtheta)$ \cite{nesterov2013introductory} and \eqref{eq:12} is from the SVRPG update.
+Inequality~\ref{eq:13} follows from Lemma~\ref{lemma:aux0} by noticing that $\nabla J(\vtheta^{s+1}_t)$ is a deterministic function given $\vtheta^{s+1}_t$.
+As a consequence, we can directly apply Lemma~\ref{lemma:aux0} with $\vphi(\vtheta_t^{s+1})\coloneqq\gradJ{\vtheta_t^{s+1}}$.
 
@@
 &\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
-+2\alpha_t\Ets{\left|\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\wt{\vtheta}^s}\right|} \nonumber\\ 
++2\alpha_t\Ets[t-1]{\left|\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_t^{s+1}-\wt{\vtheta}^s}\right|} \nonumber\\ 
 &\qquad+
@@
 &\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
-+2\alpha_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}} \nonumber\\ 
++2\alpha_t\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}} \nonumber\\ 
 &\qquad+
@@
 &\leq \Ets{\alpha_t^2\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2+\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
-+2\alpha_t\Ets{\frac{1}{2\beta_t}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta_t}{2}\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2} \label{eq:16a}\\ 
++2\alpha_t\Ets[t-1]{\frac{1}{2\beta_t}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta_t}{2}\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2} \label{eq:16a}\\ 
 &\qquad
@@
 \endgroup
-where~\eqref{eq:14a} is obtained using the triangular inequality,~\eqref{eq:14} is from the SVRPG update,~\eqref{eq:15} is from Lemma~\ref{lemma:aux0},~\eqref{eq:16a} is from Cauchy-Schwarz inequality, and~\eqref{eq:16} is from Young's inequality in the `Peter-Paul' variant.
-Let's us consider the following function:
+where~\eqref{eq:14a} is obtained using the triangular inequality,~\eqref{eq:14} is from the SVRPG update,~\eqref{eq:15} is from Lemma~\ref{lemma:aux0} with $\vphi(\vtheta_t^{s+1})\coloneqq\vtheta_t^{s+1}-\tilde{\vtheta}^s$,~\eqref{eq:16a} is from Cauchy-Schwarz inequality, and~\eqref{eq:16} is from Young's inequality in the `Peter-Paul' variant.
+Let us consider the following function:
 \[
@@
 	R_{t+1}^{s+1} 
-    % &= \Ets{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\
-	%
-	%&\geq	\Ets{J(\vtheta_t^{s+1})+\frac{\alpha_t}{2}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\geq	\Ets{J(\vtheta_t^{s+1}) - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}
@@
 	%
-	&\geq \Ets{J(\vtheta_t^{s+1})+\frac{\alpha_t}{2}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
+	&\geq \Ets{J(\vtheta_t^{s+1}) - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} 
+	+ \Ets[t-1]{\frac{\alpha_t}{2}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 }
 	-\frac{\alpha_t}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
@@
 	\nonumber\\
-	&\qquad-2c_{t+1}\alpha_t\Ets{\frac{1}{2\beta_t}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta_t}{2}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\ 
+	&\qquad-2c_{t+1}\alpha_t\Ets[t-1]{\frac{1}{2\beta_t}\norm[]{\gradJ{\vtheta_t^{s+1}}}^2+\frac{\beta_t}{2}\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2} \nonumber\\ 
 	&\qquad
@@
 	%
-	&= \Ets{J(\vtheta_t^{s+1})} - c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	&= \Ets[t-1]{J(\vtheta_t^{s+1})} - c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t\right)\Ets[t-1]{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad-\alpha_t^2\left(\frac{L}{2}+c_{t+1}\right)\Ets{\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}
-	+\frac{\alpha_t}{2}\left(1-\frac{2c_{t+1}}{\beta_t}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	+\frac{\alpha_t}{2}\left(1-\frac{2c_{t+1}}{\beta_t}\right)\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
 	%
-	&\geq  \Ets{J(\vtheta_t^{s+1})} - c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t\right)\Ets{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	&\geq  \Ets[t-1]{J(\vtheta_t^{s+1})} - c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t\right)\Ets[t-1]{\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad
@@
 	+\frac{\Gamma W}{B}\right)
-	+\frac{\alpha_t}{2}\left(1-2\frac{c_{t+1}}{\beta_t}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	+\frac{\alpha_t}{2}\left(1-2\frac{c_{t+1}}{\beta_t}\right)\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \label{eq:19}\\
 	%
-	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
+	& = \Ets[t-1]{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\
 	&\qquad
-	+\alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
+	+\alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
 	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
@@
 	&= R_t^{s+1}
-	+\Psi_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
+	+\Psi_t\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
 	-\frac{d_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
@@
 	&\geq R_t^{s+1}
-	+\Psi_t\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
+	+\Psi_t\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
 	-\frac{d_t}{N}\VARRF
@@
 \endgroup
-where~\eqref{eq:17} is from~\eqref{eq:13} noticing that $\EVV[t|s]{\norm[]{\nabla J(\vtheta^{s+1}_t)}^2} = \EVV[t|s]{\norm[]{\nabla J(\vtheta^{s+1}_t)}^2}$,~\eqref{eq:18} is from~\eqref{eq:16},~\eqref{eq:19} is from Lemma~\ref{lemma:aux2}, and (\ref{eq:20}) is from Assumption \ref{ass:REINFORCE}.
+where~\eqref{eq:17} is from~\eqref{eq:13} noticing that $\EVV[t|s]{\norm[]{\nabla J(\vtheta^{s+1}_t)}^2} = \EVV[t-1|s]{\norm[]{\nabla J(\vtheta^{s+1}_t)}^2}$,~\eqref{eq:18} is from~\eqref{eq:16},~\eqref{eq:19} is from Lemma~\ref{lemma:aux2}, and (\ref{eq:20}) is from Assumption \ref{ass:REINFORCE}.
 To complete the proof, besides rearranging terms, we have to ensure that $\Psi_t>0$ for each $t$. This gives the constraints on $\alpha_t$ and $\beta_t$.

### 2018-02-08 svrpg.tex
Updated
generalized results to GPOMDP; fixed alpha constraint

@@
 Because of the infinite dataset, we can only rely on an estimate of the full gradient.
-\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being \textcolor{red}{concave}, showing that SVRG is robust to an inexact computation of the full gradient.
+\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being concave, showing that SVRG is robust to an inexact computation of the full gradient.
 In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate.
@@
 It is easy to show that the SVRPG estimator has the following, desirable properties (see Appendix~\ref{app:proofs}):
-\begin{lemma}\label{L:svrpg.properties} \todopirout{extended to any unbiased estimator}
-        {\color{blue} Let $\wh{\nabla}_N J(\vtheta)$ be an unbiased estimator of~\eqref{E:policygradient} }
+\begin{lemma}\label{L:svrpg.properties}
+        Let $\wh{\nabla}_N J(\vtheta)$ be an unbiased estimator of~\eqref{E:policygradient}
 %Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} 
@@
 Adaptive selection of these meta-parameters is also possible, and it is explored in Section~\ref{sec:stopping}.
-However, the next section will provide general converge guarantees for SVRPG \textcolor{red}{with REINFORCE estimator}.
+However, the next section will provide general converge guarantees for SVRPG.
 
@@
 		\ENDFOR
-        \STATE {\bfseries return} $\vtheta_t^{s+1}$ with $(s,t)$ picked uniformly at random from $\{[0,S-1]\times[0,m-1]\}$
+        \STATE {\bfseries return} $\vtheta_A\coloneqq\vtheta_t^{s+1}$ with $(s,t)$ picked uniformly at random from $\{[0,S-1]\times[0,m-1]\}$
 	\end{algorithmic}
@@
 \section{Convergence Guarantees of SVRPG}\label{sec:conv}
-In this section, we state the convergence guarantees for SVRPG {\color{red}with REINFORCE estimator}.\footnote{\color{red}Similar results can be derived for G(PO)MDP, see Appendix~\ref{app:}.}
-We mainly leverage on the recent analysis of \textcolor{red}{non-concave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
+In this section, we state the convergence guarantees for SVRPG with REINFORCE or G(PO)MDP gradient estimator.
+We mainly leverage on the recent analysis of non-concave SVRG~\cite{reddi2016stochastic,allen2016variance}.
 Each of the three challenges presented at the beginning of Section~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
@@
 We can now state the convergence guarantees for SVRPG.
-\begin{restatable}[Convergence of the {\color{red}REINFORCE-}SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
-{\color{red}Consider the REINFORCE estimator in Eq.~\eqref{E:policygradient.estimate}.}
+\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
+Assume the REINFORCE or the G(PO)MDP gradient estimator is used in Equation $\ref{E:policygradient.estimate}$.
 Under Assumptions~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
@@
 %The theoretical settings may be too conservative.
-By analysing the upper-bound in Theorem~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on \textcolor{red}{non-concave} SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
+By analysing the upper-bound in Theorem~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on non-concave SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
 To achieve true convergence, both the batch size $N$ and the mini-batch size $B$ should increase over time. 
@@
 This is particularly meaningful in the context of online learning that we are considering in this paper.
-
-{\color{red}
-\textbf{G(PO)MDP.} A similar result to the one stated in Thm~\ref{theo:convergence} can be derived for the G(PO)MDP estimator. A sketch of the proof is reported in~\ref{app:gpomdpproof}.
-In practice, due to the lower variance G(PO)MDP is preferred to REINFORCE.
-For this reason, we will use G(PO)MDP for the experiments (see Sec.~\ref{sec:exp}).
-\todopir{sketch of proof for GPOMDP in appendix, probably can be unified with the one of REINFORCE}}
 
@@
 \begin{align*}
-&0\leq\alpha_t < \frac{2(1-\nicefrac{2c_{t+1}}{\beta_t})}{L+2c_{t+1}} \\
+&0\leq\alpha_t < \frac{1-\nicefrac{2c_{t+1}}{\beta_t}}{L+2c_{t+1}} \\
 &\beta_t > 2c_{t+1}.

### 2018-02-08 svrpg.tex
fixed Lipschitz constant notation

@@
         \mathbb{E}_{t\vert s}\left[X\right] &\coloneqq 
-		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\vert\wt{\vtheta}^s)\forall j \in {\it {\bf N}} \\ \tau_{i,h}\sim\pi(\cdot\vert\vtheta^{s+1}_i) \forall h \in {\it {\bf B}}, \text{ for $i=0,\dots,t-1$}}}{\left[X \vert \wt{\vtheta^s}\right]} \\
+		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\vert\wt{\vtheta}^s)\forall j \in {\it {\bf N}} \\ \tau_{i,h}\sim\pi(\cdot\vert\vtheta^{s+1}_i) \forall h \in {\it {\bf B}}, \text{ for $i=0,\dots,t$}}}{\left[X \vert \wt{\vtheta^s}\right]} \\
 	&\coloneqq \EVV[\tau_j\sim\pi(\cdot\vert\wt{\vtheta}^s)\forall j \in {\it {\bf N}}]{
@@
 where the sequence $\wt{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}, ${\it {\bf N}} = [0,\dots,N)$, and ${\it {\bf B}} = [0,\dots,B)$.
-\todopir{what about $\vtheta^{s+1}_0$ and $\wt{\vtheta}^s$? They are equal. How this affect previous definition?}
+To avoid inconsistencies, we also define $\Ets[(-1)]{X} \coloneqq \Es{X}$.
 
@@
 \begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
-	Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth for some positive Lipschitz constant $L_1$.
+	Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth for some positive Lipschitz constant $L_J$.
 \end{restatable}
@@
 \begin{restatable}[]{lemma}{l-smooth-g}\label{lemma:gsmooth}
-Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, $g(\tau\vert\vtheta)$ is Lipschitz continuous with Lipschitz constant $L_2$, \ie for any trajectory $\tau\in\Tspace$:
-\[
-	\norm[2]{g(\tau\vert\vtheta)-g(\tau\vert\vtheta')}^2 \leq L_2\norm[2]{\vtheta-\vtheta'}^2.
-\]
-\todopir{I would denote $L_J$ and $L_g$ rather than 1 and 2}
+Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, $g(\tau\vert\vtheta)$ is Lipschitz continuous with Lipschitz constant $L_g$, \ie for any trajectory $\tau\in\Tspace$:
+\[
+	\norm[2]{g(\tau\vert\vtheta)-g(\tau\vert\vtheta')}^2 \leq L_g\norm[2]{\vtheta-\vtheta'}^2.
+\]
 \end{restatable}
@@
 \Ets[t-1]{\norm[2]{\gradJ{\vtheta_t^{s+1}}}^2} 
-+\frac{L_2^2}{B}\Ets[t-1]{\norm[2]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
++\frac{L_g^2}{B}\Ets[t-1]{\norm[2]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
 +\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}
@@
 	&\qquad
-    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
+    +\frac{L_g^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
     +
@@
 	\nonumber\\
-	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
+	&\qquad+\frac{L_g^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
 	+\Gamma\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \label{eq:4}\\
@@
 %	\nonumber\\&\qquad
-    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
+    +\frac{L_g^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
 	+\frac{\Gamma}{B^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\
@@
 % 	\nonumber\\ &\qquad
-    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
+    +\frac{L_g^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
 	+\frac{\Gamma\VARIS}{B}, \label{eq:5}
@@
 \end{align*}
-where $L=\max\left\{L1,L2\right\}$, \ie the greater of the Lipschitz constants from Lemmas \ref{lemma:lsmooth} and \ref{lemma:gsmooth}.
+where $L=\max\left\{L_J,L_g\right\}$, \ie the greater of the Lipschitz constants from Lemmas \ref{lemma:lsmooth} and \ref{lemma:gsmooth}.
 

### 2018-02-08 svrpg.tex
Updated

--- 2018-02-08 svrpg.tex
@@
 Let us consider the following function:
-\[
+\begin{equation}\label{E:lyapunov.function}
 	R_{t+1}^{s+1} \coloneqq \Ets{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2}. 
-\]
+\end{equation}
 The objective is now to provide a lower bound to it.
@@
 	+\alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right)\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \nonumber\\
-	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
-	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})\Gamma\VARIS}{2B} \nonumber\\
+	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} 
+%    \nonumber\\ &\qquad
+    -\alpha_t^2\frac{(L+2c_{t+1})\Gamma\VARIS}{2B} \nonumber\\
 	%
@@
 	-\frac{d_t}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}
-	-\frac{f_t}{b}\VARIS,\nonumber\\
+	-\frac{f_t}{B}\VARIS,\nonumber\\
 	&\geq R_t^{s+1}
@@
 	-\frac{d_t}{N}\VARRF
-	-\frac{f_t}{b}\VARIS, \label{eq:20}
+	-\frac{f_t}{B}\VARIS, \label{eq:20}
 \end{align}
@@
 \begin{align*}
-& \psi \coloneqq \min_t\{\Psi_t\}, \\
-& \zeta \coloneqq \frac{\max_t\{d_t\}V}{\psi} \\
-& \xi \coloneqq \frac{\max_t\{f_t\}W}{\psi}.
-\end{align*}
-The values of all other constants and the constraints on the step size are provided in Lemma \ref{lemma:aux1}.
-
-Starting from Lemma \ref{lemma:aux1}, summing over iterations of an epoch $s$ we get:
-\begin{align}
+& \psi \coloneqq \min_t\{\Psi_t\}, 
+& \zeta \coloneqq \frac{\max_t\{d_t\}V}{\psi}, 
+&& \xi \coloneqq \frac{\max_t\{f_t\}W}{\psi},
+\end{align*}
+where $\Psi$, $d_t$ and $f_t$ are defined in Lemma~\ref{lemma:aux1}.
+% The values of all other constants and the constraints on the step size are provided in Lemma \ref{lemma:aux1}.
+Starting from Lemma \ref{lemma:aux1}, summing over iterations of an epoch $s$ and using telescopic sum we obtain
+\begin{align*}
 \sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
  \frac{\sum_{t=0}^{m-1}\left(R_{t+1}^{s+1} - R_t^{s+1}\right)}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
- %
- &= \frac{\sum_{t=0}^{m-1}\left(\Ets[t+1]{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} - \Ets[t]{J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2}\right)}{\psi} \nonumber\\
- &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
- %
- &= \frac{\Ets[m]{\sum_{t=0}^{m-1}\left(J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2\right)}}{\psi} \nonumber\\
- &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
- %
- &= \frac{\Ets[m]{J(\vtheta_{m}^{s+1}) - c_{m}\norm[]{\vtheta_{m}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{0}^{s+1}) - c_{0}\norm[]{\vtheta_{0}^{s+1}-\tilde{\vtheta}^s}^2}}{\psi} \nonumber\\
- &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \label{eq:21}\\
- %
- &= \frac{\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B}, \label{eq:22}
- \end{align}
-where (\ref{eq:21}) is obtained by telescoping the sum and (\ref{eq:22}) from the fact that $c_m=0$ and $\vtheta_0^{s+1}=\tilde{\vtheta}^s$.
+ & = \frac{R^{s+1}_m - R^{s+1}_0}{\psi}  + \frac{m\zeta}{N} + \frac{m\xi}{B}
+\end{align*}
+By using the definition of $R^s_t$ in~\eqref{E:lyapunov.function}, the fact that $c_m = 0$ and $\vtheta^{s+1}_0 = \wt{\vtheta}^s = \vtheta^s_m$, we can state that
+\todopir{Why $c_m =0$? In the previous theorem  was $c_{m-1} = 0$}
+\begin{align*}
+        R^{s+1}_m - R^{s+1}_0 
+        &= \EVV[m|s]{J(\vtheta^{s+1}_{m})- c_m \norm[]{\vtheta^{s+1}_{m} - \wt{\vtheta}^s}^2 } - \EVV[0|s]{J(\vtheta^{s+1}_{0}) - c_0 \norm[]{\vtheta^{s+1}_0 - \wt{\vtheta}^s}^2}\\
+        &= \EVV[m|s]{J(\vtheta^{s+1}_{m})} - \EVV[0|s]{J(\wt{\vtheta}^s)}
+        = \EVV[m|s]{J(\wt{\vtheta}^{s+1}) - J(\wt{\vtheta}^s)}
+\end{align*}
 Next, summing over epochs:
@@
 \end{align}
-where the expectation in (\ref{eq:23}) is \wrt all the trajectories sampled in a run of Algorithm \ref{alg:svrpg} and (\ref{eq:24}) is from the definition of $\vtheta^*$.
-
+where the expectation in~\eqref{eq:23} is \wrt all the trajectories sampled in a run of Algorithm~\ref{alg:svrpg} and~\eqref{eq:24} is from the definition of $\vtheta^*$ (\ie the policy performance maximizer).
 Finally, we consider the expectation \wrt all sources of randomness, including the uniform sampling of the output parameter:
-\begin{align}
+\begin{align*}
 \EVV[]{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
-&=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} \nonumber\\
-&\leq
-\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B} \nonumber.
-\end{align}
+&=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
+\leq
+\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B}.
+\end{align*}
+
+%% \begin{align}
+%% \sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}^2}}&\leq
+%%  \frac{\sum_{t=0}^{m-1}\left(R_{t+1}^{s+1} - R_t^{s+1}\right)}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
+%%  %
+%%  &= \frac{\sum_{t=0}^{m-1}\left(\Ets[t+1]{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} - \Ets[t]{J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2}\right)}{\psi} \nonumber\\
+%%  &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
+%%  %
+%%  &= \frac{\Ets[m]{\sum_{t=0}^{m-1}\left(J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2\right)}}{\psi} \nonumber\\
+%%  &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
+%%  %
+%%  &= \frac{\Ets[m]{J(\vtheta_{m}^{s+1}) - c_{m}\norm[]{\vtheta_{m}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{0}^{s+1}) - c_{0}\norm[]{\vtheta_{0}^{s+1}-\tilde{\vtheta}^s}^2}}{\psi} \nonumber\\
+%%  &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \label{eq:21}\\
+%%  %
+%%  &= \frac{\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B}, \label{eq:22}
+%%  \end{align}
+%% where (\ref{eq:21}) is obtained by telescoping the sum and (\ref{eq:22}) from the fact that $c_m=0$ and $\vtheta_0^{s+1}=\tilde{\vtheta}^s$.
+%% Next, summing over epochs:
+%% \begin{align}
+%% \sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
+%% \frac{\sum_{s=0}^{S-1}\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \nonumber\\
+%% %
+%% &\leq
+%% \frac{\EVV[]{J(\tilde{\vtheta}^{S}) - J(\tilde{\vtheta}^{0})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}
+%%  \label{eq:23}\\
+%% %
+%% &\leq
+%% \frac{J(\vtheta^*) - J(\vtheta^0)}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}, \label{eq:24}
+%% \end{align}
+%% where the expectation in (\ref{eq:23}) is \wrt all the trajectories sampled in a run of Algorithm \ref{alg:svrpg} and (\ref{eq:24}) is from the definition of $\vtheta^*$.
+%%  
+%% Finally, we consider the expectation \wrt all sources of randomness, including the uniform sampling of the output parameter:
+%% \begin{align}
+%% \EVV[]{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
+%% &=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} \nonumber\\
+%% &\leq
+%% \frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B} \nonumber.
+%% \end{align}
 \end{proof}

### 2018-02-08 svrpg.tex
Updated

@@
 %\end{restatable}
+Previous results hold both for REINFORCE and G(PO)MDP.
 In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
@@
 For the sake of clarity, we have not explicitly differentiated this step from the standard update rule in Algorithm~\ref{alg:svrpg}.}
+
+\todopir{mention that even if for theoretical guarantees we should return a random solution, this is not a problem for us because we are evaluating the performance on-line. And is always improving}
 
@@
 \begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
-Assume the REINFORCE or the G(PO)MDP gradient estimator is used in Equation $\ref{E:policygradient.estimate}$.
+Assume the REINFORCE or the G(PO)MDP gradient estimator is used in SVRPG (see Equation~\eqref{E:svrpg.estimate.batch}).
 Under Assumptions~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
@@
 \end{algorithm}
+\todopir{Explain how we use adam and when/if it is reset on snapshots}
 

### 2018-02-08 svrpg.tex
moved proof of Lemma 3.1 to main paper

--- 2018-02-08 svrpg.tex
@@
 The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}.
-It is easy to show that the SVRPG estimator has the following, desirable properties (see Appendix~\ref{app:proofs}):
+It is easy to show that the SVRPG estimator has the following, desirable properties:
 \begin{lemma}\label{L:svrpg.properties}
@@
 and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$. Then, the SVRG estimate in~\eqref{E:svrpg.estimate.batch} is \emph{unbiased}
-\[
+\begin{equation}\label{eq:unbiased}
 \mathop{\mathbb{E}}
 \left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
-\]
+\end{equation}
 and regardless of the mini-batch size $B$:\footnote{
 For any vector $\mathbf{x}$, we use $\Var[\mathbf{x}]$ to denote the trace of the covariance matrix, \ie $\Tr\EVV[]{(\mathbf{x}-\EVV[]{\mathbf{x}})(\mathbf{x}-\EVV[]{\mathbf{x}})^T}$.}
-\[
+\begin{equation}\label{eq:zerovar}
 	\Var\left[\gradBlack{\vtheta^*}\right] = 
     \Var\left[\wh{\nabla}_N J(\vtheta^*)\right].
-\]
+\end{equation}
 \end{lemma}
-
-%\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
-%The SVRPG estimator is unbiased:
-%\[
-%\mathop{\mathbb{E}}
-%\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
-%\]
-%\end{restatable}
-%\begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
-%At convergence, the variance of the SVRPG estimator is equal to the variance of the REINFORCE estimator, regardless of the mini-batch size $B$:
-%\[
-%	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
-%	\mathbb{V}ar\left[\gradApp{\vtheta^*}{N}\right].
-%\]
-%\end{restatable}
+\begin{proof}
+Claim (\ref{eq:unbiased}) is from the unbiasedness of the estimators:
+\begin{align*}
+\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} \\
+&\qquad- \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
+&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
+\end{align*}
+As for claim (\ref{eq:zerovar}), note that as $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
+\begin{align*}
+\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*)\right. \\
+&\left.- \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right]
+= \Var\left[\gradApp{\vtheta^*}{N}\right].
+\end{align*}
+Note that it is important that the trajectories used in the second and the third term are the same for the variance to vanish.
+\end{proof}
 In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
@@
 
-\subsection*{Basic Lemmas}
-We prove two basic properties of the SVRPG update.
-
-%\unbias*
-\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
-Consider Assumption~\ref{asm:unbiasedness}. Then, the SVRPG estimator is unbiased:
-\[
-\mathop{\mathbb{E}}
-\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
-\]
-\end{restatable}
-\begin{proof}
-\begin{align*}
-        \EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} - \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
-&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
-\end{align*}
-Note that the importance weight is necessary to guarantee unbiasedness, since the $\tau_i$ are sampled from $\pi_{\vtheta}$.
-\end{proof}
-
-% \zerovar*
-\begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
-At convergence, the variance of the SVRPG estimator is equal to the variance of the gradient estimator, regardless of the mini-batch size $B$:
-\[
-	\Var\left[\gradBlack{\vtheta^*}\right] = 
-	\Var\left[\gradApp{\vtheta^*}{N}\right].
-\]
-\end{restatable}
-\begin{proof}
-As $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
-\begin{align*}
-\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*) - \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right] \\
-&= \Var\left[\gradApp{\vtheta^*}{N}\right].
-\end{align*}
-Note that it is important that the trajectories used in the second and the third term are the same for the variance to vanish.
-\end{proof}
+%\subsection*{Basic Lemmas}
+%We prove two basic properties of the SVRPG update.
+%
+%%\unbias*
+%\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
+%Consider Assumption~\ref{asm:unbiasedness}. Then, the SVRPG estimator is unbiased:
+%\[
+%\mathop{\mathbb{E}}
+%\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
+%\]
+%\end{restatable}
+%\begin{proof}
+%\begin{align*}
+%        \EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} - \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
+%&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
+%\end{align*}
+%Note that the importance weight is necessary to guarantee unbiasedness, since the $\tau_i$ are sampled from $\pi_{\vtheta}$.
+%\end{proof}
+%
+%% \zerovar*
+%\begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
+%At convergence, the variance of the SVRPG estimator is equal to the variance of the gradient estimator, regardless of the mini-batch size $B$:
+%\[
+%	\Var\left[\gradBlack{\vtheta^*}\right] = 
+%	\Var\left[\gradApp{\vtheta^*}{N}\right].
+%\]
+%\end{restatable}
+%\begin{proof}
+%As $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
+%\begin{align*}
+%\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*) - \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right] \\
+%&= \Var\left[\gradApp{\vtheta^*}{N}\right].
+%\end{align*}
+%Note that it is important that the trajectories used in the second and the third term are the same for the variance to vanish.
+%\end{proof}
 

### 2018-02-08 svrpg.tex

@@
 \end{proof}
+Previous results hold for both REINFORCE and G(PO)MDP.
 In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
@@
 	&R_t^{s+1}\coloneqq \Ets[t-1]{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}, \\
-	&c_{m-1} = 0, \\
+	&c_{m} = 0, \\
 	&c_t = c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}, \\
@@
 \end{align*}
-By using the definition of $R^s_t$ in~\eqref{E:lyapunov.function}, the fact that $c_m = 0$ and $\vtheta^{s+1}_0 = \wt{\vtheta}^s = \vtheta^s_m$, we can state that
-\todopir{Why $c_m =0$? In the previous theorem  was $c_{m-1} = 0$}
+By using the definition of $R^s_t$ in~\eqref{E:lyapunov.function}, the fact that $c_m = 0$ and $\vtheta^{s+1}_0 = \wt{\vtheta}^s = \vtheta^s_m$, we can state that:
 \begin{align*}

### 2018-02-08 svrpg.tex
updated all sections except Experiments

@@
 However, the two approaches share many challenges and tools.
-The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or directly a policy defining agent's behaviour.
+The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or directly a policy defining the agent's behaviour.
 Furthermore, when the tasks are characterized by large or continuous state-action spaces, RL needs the powerful function approximators (\eg neural networks) that are the main subject of study of SL.
@@
 %Paper structure
-After providing background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} we address, as mentioned above, the difficulties in designing SVRPG, a variant of SVRG for the policy gradient framework.
+After providing background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} propose SVRPG, a variant of SVRG for the policy gradient framework, addressing all the difficulties mentioned above.
 In Section \ref{sec:conv} we provide convergence guarantees for our algorithm which also offers insights on the challenges of variance reduction in RL.
@@
 
-The idea of SVRG is to alternate full and stochastic gradient updates. 
+The idea of SVRG (Algorithm~\ref{alg:svrg}) is to alternate full and stochastic gradient updates. 
 Each $m = \mathcal{O}(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i \nabla z(x_i|\widetilde{\vtheta})$.
-Between snapshots, the parameter is updated with a gradient estimate corrected using stochastic gradient, for any $t \in \{0,\ldots,m-1\}$
+Between snapshots, the parameter is updated with $\blacktriangledown f(\vtheta)$, a gradient estimate corrected using stochastic gradient. For any $t \in \{0,\ldots,m-1\}$:
 \begin{equation}\label{E:svrg.gradient.correction}
-        \blacktriangledown f(\vtheta_{t}) := v_t = \nabla f(\wt{\vtheta}) + \nabla z(x_i | \vtheta_t) - \nabla z(x_i | \wt{\vtheta}),
+        \blacktriangledown f(\vtheta_{t}) := v_t = \nabla f(\wt{\vtheta}) + \nabla z(x | \vtheta_t) - \nabla z(x | \wt{\vtheta}),
 \end{equation} 
-where $x_i$ is sampled uniformly at random from $\mathcal{D}_N$ (\ie $x_i \sim \mathcal{U}(\mathcal{D}_N)$).
+where $x$ is sampled uniformly at random from $\mathcal{D}_N$ (\ie $x \sim \mathcal{U}(\mathcal{D}_N)$).
 Note that $t=0$ corresponds to a FG step (\ie $\blacktriangledown f(\vtheta_0) = \nabla f(\wt{\vtheta})$) since $\vtheta_0 := \wt{\vtheta}$.
@@
 The only additional requirement is to select $\vtheta^*$ uniformly at random among all the $\vtheta_k$ instead of simply setting it to the final value.
-Pseudocode of SVRG for both the concave and the non-concave case is provided in Algorithm~\ref{alg:svrg}.
 
@@
 %		\STATE Pick $i$ uniformly at random from $[1,N]$
-        \STATE $x_i \sim \mathcal{U}\left(\mathcal{D}_N\right)$
+        \STATE $x \sim \mathcal{U}\left(\mathcal{D}_N\right)$
 		\STATE $v^{s+1}_t = 
@@
 			\wt{\mu} + 
-			\nabla z(x_i|\vtheta_t^{s+1}) -
-			\nabla z(x_i|\wt{\vtheta}^{s})
+			\nabla z(x|\vtheta_t^{s+1}) -
+			\nabla z(x|\wt{\vtheta}^{s})
 		$
@@
 \end{description}
-To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}, see Appendix~\ref{app:gauss} for the derivation.
+To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as Gaussian\footnote{see Appendix~\ref{app:gauss} for more details on the Gaussian policy case.} and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}.
 Because of the infinite dataset, we can only rely on an estimate of the full gradient.
@@
 In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate.
-{\color{red}Later on we will show how the estimation accuracy impacts on the convergence results with non-concave objective.}
+{\color{red}In Section \ref{sec:conv} we will show how the estimation accuracy impacts on the convergence results with non-concave objective.}
 Finally, the non-stationarity of the optimization problem introduces a bias into the SVRG estimator in Eq.~\eqref{E:svrg.gradient.correction}.
@@
 
-We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG) for a generic policy gradient estimator $g$ (see Eq.~\eqref{E:policygradient.estimate}).
-% For simplicity, we focus on the REINFORCE estimator but the extension to G(PO)MDP is straightforward.
-Between snapshots, the SVRPG estimate is corrected using importance weighting, for any $t \in \{0,\ldots,m-1\}$
-\begin{align*}
-        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta}),
-\end{align*}
-% \begin{align*}
-% 	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \wt{\vtheta})\Reward(\tau_j) \\
-% 		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
-% 		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \wt{\vtheta})\Reward(\tau_i),
-% \end{align*}
-where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in Eq.~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{p(\tau|\wt{\vtheta})}{p(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$.
+\begin{algorithm}[tb]
+	\caption{SVRPG}
+	\label{alg:svrpg}
+	\begin{algorithmic}
+		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{m}^0 := \wt{\vtheta}^0$
+		\FOR{$s=0$ {\bfseries to} $S-1$}
+		\STATE $\vtheta_0^{s+1} := \wt{\vtheta}^{s} = \vtheta_{m}^s$
+		\STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\wt{\vtheta}^{s})$
+		\STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\eqref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
+		\FOR{$t=0$ {\bfseries to} $m-1$}
+		\STATE Sample $B$ trajectories $\{\tau_i\}$ from $p(\cdot\vert\vtheta_t^{s+1})$
+		%        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)$
+		\STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
+		\begin{aligned}[t]
+		\Big( & g(\tau_i|\vtheta_t^{s+1})\\ 
+		& \quad{} - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \Big)
+		\end{aligned}$
+		%        \STATE $\begin{aligned}[t]
+		%                &v^{s+1}_t = \wt{\mu}\\ 
+		%                & +\frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)
+		%                \end{aligned}$
+		\STATE $v^{s+1}_t = \wt{\mu} + c^{s+1}_t$ % \gradApp{\wt{\vtheta}^{s}}{N} + c^{s}_t$
+		%%		\begin{align*}
+		%%		\blacktriangledown J(\vtheta_t^{s+1}) = 
+		%%		&\gradApp{\wt{\vtheta}^s}{N} \\
+		%%		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
+		%%		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
+		%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \wt{\vtheta}^{s}}\Reward(\tau_i)\right]
+		%%		\end{align*}
+		\STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
+		\ENDFOR
+		\ENDFOR
+		\STATE {\bfseries return} $\vtheta_A\coloneqq\vtheta_t^{s+1}$ with $(s,t)$ picked uniformly at random from $\{[0,S-1]\times[0,m-1]\}$
+	\end{algorithmic}
+\end{algorithm}
+
+We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG) for a generic policy gradient estimator $g$. Pseudo-code is provided in Algorithm \ref{alg:svrpg}.
+The overall structure is the same as Algorithm \ref{alg:svrg}, 
+but the snapshot gradient is not exact and the gradient estimate used between snapshots is corrected using importance weighting:
+\begin{align*}
+        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau|\vtheta_t) - \omega(\tau|\vtheta_t, \wt{\vtheta}) g(\tau|\wt{\vtheta})
+\end{align*}
+for any $t \in \{0,\ldots,m-1\}$,
+where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in Eq.~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{p(\tau|\wt{\vtheta})}{p(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$.
 Similarly to SVRG, we have that $\vtheta_0 := \wt{\vtheta}$, and the update is a FG step.
 Our update is still fundamentally on-policy since the weighting concerns only the correction term.
-However, this partial ``off-policyness'' is enough to introduce variance.
+However, this partial ``off-policyness'' represents an additional source of variance.
 This is a well-known issue of importance sampling~\citep[\eg][]{thomas2015high}.
@@
 In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
-Algorithm~\ref{alg:svrpg} presents the details of the SVRPG method for policy optimization.\footnote{Note that when $t=0$ we can avoid to sample trajectories and instead directly perform the policy update using the FG.
-For the sake of clarity, we have not explicitly differentiated this step from the standard update rule in Algorithm~\ref{alg:svrpg}.}
-
 \todopir{mention that even if for theoretical guarantees we should return a random solution, this is not a problem for us because we are evaluating the performance on-line. And is always improving}
-
-To use SVRPG in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient, the mini-batch size $B$, and the step size $\alpha$.
-Adaptive selection of these meta-parameters is also possible, and it is explored in Section~\ref{sec:stopping}.
-However, the next section will provide general converge guarantees for SVRPG.
-
-\begin{algorithm}[tb]
-        \caption{SVRPG}
-	\label{alg:svrpg}
-	\begin{algorithmic}
-        \STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{0}^m := \wt{\vtheta}^0$
-		\FOR{$s=0$ {\bfseries to} $S-1$}
-        \STATE $\vtheta_0^{s+1} := \wt{\vtheta}^{s} = \vtheta_{m}^s$
-        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\wt{\vtheta}^{s})$
-        \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\eqref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
-		\FOR{$t=0$ {\bfseries to} $m-1$}
-        \STATE Sample $B$ trajectories $\{\tau_i\}$ from $p(\cdot\vert\vtheta_t^{s+1})$
-%        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)$
-        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
-        \begin{aligned}[t]
-                \Big( & g(\tau_i|\vtheta_t^{s+1})\\ 
-                      & \quad{} - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \Big)
-        \end{aligned}$
-%        \STATE $\begin{aligned}[t]
-%                &v^{s+1}_t = \wt{\mu}\\ 
-%                & +\frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)
-%                \end{aligned}$
-        \STATE $v^{s+1}_t = \wt{\mu} + c^{s+1}_t$ % \gradApp{\wt{\vtheta}^{s}}{N} + c^{s}_t$
-%%		\begin{align*}
-%%		\blacktriangledown J(\vtheta_t^{s+1}) = 
-%%		&\gradApp{\wt{\vtheta}^s}{N} \\
-%%		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
-%%		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
-%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \wt{\vtheta}^{s}}\Reward(\tau_i)\right]
-%%		\end{align*}
-        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
-		\ENDFOR
-		\ENDFOR
-        \STATE {\bfseries return} $\vtheta_A\coloneqq\vtheta_t^{s+1}$ with $(s,t)$ picked uniformly at random from $\{[0,S-1]\times[0,m-1]\}$
-	\end{algorithmic}
-\end{algorithm}
 
@@
 	\end{restatable}
-Differently from Assumptions~\ref{ass:bounded_score} and ~\ref{ass:REINFORCE}, Assumption~\ref{ass:M2} can be enforced by a proper handling of the epoch size $m$.
+Differently from Assumptions~\ref{ass:bounded_score} and ~\ref{ass:REINFORCE}, Assumption~\ref{ass:M2} must be enforced by a proper handling of the epoch size $m$.
 
@@
 % At this point, the open question is how to properly select the step size $\alpha$ and the sub-iteration number $m$ in practical applications.
-The step size $\alpha$ is crucial to balance the variance due to the sub-iterations and the variance associated with the estimator at the snapshot.
+The step size $\alpha$ is crucial to balance variance reduction and efficiency.
 % The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
-The number $m$ of sub-iterations influences the variance introduced by the importance weights.
-Low $m$ values are associated with low variance but induce lots of snapshot points (\ie lots of FG computations).
-High $m$ values may move policy $\pi_{\vtheta_t}$ far away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing high variance of the importance weights.
+Epoch length $m$ influences the variance introduced by the importance weights: low values of $m$ are associated with low variance but increase the frequency of snapshot points (which means many FG computations); high values of $m$ may move policy $\pi_{\vtheta_t}$ far away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing high variance in the importance weights.
+We propose to jointly adapt the two meta-parameters.
 % the farther the distance between the snapshot policy and the current one, the higher the variance of the importance weights.
@@
 Recall that SVRPG performs two different updates of the parameters $\vtheta$: I) FG update in the snapshot; II) corrected gradient update in the sub-iterations.
-Given this structure, we suggest to use two ADAM estimators:
-\begin{align*}
-        \vtheta^{s}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
-        \vtheta^{s}_{t+1} &= \vtheta^{s}_t + \alpha^{\textsc{SI}}_{s,t} v_t^{s},
+Given this structure, we suggest to use two separate ADAM estimators:
+\begin{align*}
+        \vtheta^{s+1}_0 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
+        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s,t} v_t^{s},
 \end{align*}
@@
 
-\textbf{Adaptive Sub-Iterations.}
+\textbf{Adaptive epoch length.}
 It is easy to imagine that a predefined schedule (\eg $m$ fixed in advance or changed with a policy-independent process) may poorly perform due to the high variability of the updates.
@@
 \[
-        \text{If }        \frac{\alpha^{\textsc{FG}}}{N} > \frac{\alpha^{\textsc{SI}}}{B} \textbf{ then } \text{take snapshot,}
+        \textbf{If }        \frac{\alpha^{\textsc{FG}}}{N} > \frac{\alpha^{\textsc{SI}}}{B} \textbf{ then } \text{take snapshot,}
 \]
@@
 In Section \ref{sec:exp} we show some cases in which self-normalization provides a performance improvement.
-
-\subsection{Baseline}
-As reported in Section~\ref{subsec:PolicyGradient}, a popular variance reduction tool in Reinforcement Learning is the employment of baselines in policy gradient estimation.
-This tool is orthogonal to the variance reduction technique proposed in this paper.
-In Section~\ref{sec:exp} we try to combine SVRPG with a non-stationary, state-dependent, learned linear baseline like the one used in~\cite{duan2016benchmarking}, with the variant that baseline parameters are updated only at the snapshot, with trajectories sampled from $\pi_{\wt{\vtheta}^s}$.
 
@@
 The second approach~\citep{xu2017svrgtrpo} uses SVRG as a practical method to solve the optimization problem faced by Trust Region Policy Optimization (TRPO) at each iteration.
-We can just talk of a direct application of SVRG to a problem having finite-sum structure since no specific structure of the RL problem is exploited.
+This is just a direct application of SVRG to a problem having finite-sum structure since no specific structure of the RL problem is exploited.
 It is worth to mention that, for practical reasons, the authors proposed to use a Newton conjugate gradient method with SVRG.
@@
 In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.
-The control problem considered in the paper has a series of difficulties non-common in SL.
+The control problem considered in the paper has a series of difficulties that are not common in SL.
 Among them, non-concavity and approximate estimates of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
-Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy dependent.
+Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy-dependent.
 We have shown that by exploiting importance weighting techniques, it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
 We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
-SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
+{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
 The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
 In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
+}
+\todomat{Repetition of Section 6?}
 Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on several benchmarks.

### 2018-02-08 svrpg.tex
updated section 5

@@
 
-\subsection{Stochastic Variance-Reduced Gradient}
+\subsection{Stochastic Variance-Reduced Gradient}\label{sec:svrg}
 Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $z_i(\cdot|\vtheta)$:
@@
 
+\subsection{Zeroth update}
+As noted in Section \ref{sec:alg}, the update performed at the beginning of each epoch is equivalent to a full-gradient update.
+In our setting, where collecting samples is particularly expensive, the $B$ trajectories collected using the snapshot trajectory $\pi_{\tilde{\vtheta}^s}$ feel like a waste of data.
+In practice, we just perform an approximate full gradient update using the $N$ trajectories sampled to compute $\gradApp{\tilde{\vtheta}^s}{N}$, \ie
+\begin{align*}
+	&\vtheta_{1}^{s+1} = \tilde{\vtheta}^s + \alpha\gradApp{\tilde{\vtheta}^s}{N} \\
+	&\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v_t^{s+1} \text{ for $t=1,\dots,m-1$}.
+\end{align*}
+In the rest of this section we assume this practical variant.
+
 \subsection{Meta-Parameter Selection}\label{sec:stopping}
@@
 \begin{align*}
-        \vtheta^{s+1}_0 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
-        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s,t} v_t^{s},
-\end{align*}
-where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s,t}$ with the sub-iterations (see Appendix~\ref{app:practicalsvrpg} for details).
+        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
+        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} v_t^{s+1} \text{ for $t=1,\dots,m-1$},
+\end{align*}
+where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s+1,t}$ with the sub-iterations (see Appendix~\ref{app:practicalsvrpg} for details).
 By doing so, we decouple the contribution of the variance due to the approximate FG from the one introduced by the sub-iterations.

### 2018-02-08 svrpg.tex
Updated

@@
 
+\todopir{Maybe we can talk about variance reduction baselines}
+
 \section{Experiments}\label{sec:exp}
@@
 \]
-\todopir{Add proof of Def 2 from Luca's thesis}
-
-\todopir{Make ref to Luca's thesis and describe in brief the derivations}
+% \todopir{Add proof of Def 2 from Luca's thesis}
+
+% \todopir{Make ref to Luca's thesis and describe in brief the derivations}
 
@@
 \end{align*}
-\todopir{I have to check this one!!!}
 \end{restatable}
@@
 \end{align}
+where~\eqref{eq:6} is from Definition~\ref{def:ideal};~\eqref{eq:7} is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $\varphi(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$;~\eqref{eq:new1} is from the fact that $\gradJ{\wt{\vtheta}^s}$ is constant \wrt $\Vars{\cdot}$;~\eqref{eq:8} is from $\Ets{e_s}=0$.
+\todopir{In Eq~\ref{eq:7} the first expected value is on $t-1$? In any case even $t$ is fine.}
 Hence:
@@
 	&\leq
-	\sqrt{\Vars{\gradApp{\wt{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{\varphi(\vtheta_t^{s+1})}} \nonumber\\
+    \sqrt{\Vars{\gradApp{\wt{\vtheta}^s}{N}}} \cdot \sqrt{\Varts[t-1]{\varphi(\vtheta_t^{s+1})}} \label{eq:9a}\\
 	%
@@
 \end{align}
-where (\ref{eq:6}) is from Definition \ref{def:ideal}; (\ref{eq:7}) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $\varphi(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$;
-(\ref{eq:new1}) is from the fact that $\gradJ{\wt{\vtheta}^s}$ is constant \wrt $\Vars{\cdot}$;
-(\ref{eq:8}) is from $\Ets{e_s}=0$; (\ref{eq:9}) is from Young's inequality; and (\ref{eq:10}) is from the definition of $\gradApp{\vtheta}{N}$.
+where~\eqref{eq:9a} comes from Cauchy-Schwarz inequality
+\[
+        |\Cov(X,Y)| = |\EVV[]{(X-\mu_X)\transpose{(Y-\mu_Y)}}| \leq \EVV[]{(X-\mu_X)^2}^{1/2}\EVV[]{(Y-\mu_Y)^2}^{1/2} = \sqrt{\Var(X) \Var(Y)},
+\]
+\eqref{eq:9} is from Young's inequality, and~\eqref{eq:10} is from the definition of $\gradApp{\vtheta}{N}$.
+\todopir{comment the last step.}
 \end{proof}
@@
 
-\paragraph{Multi-dimensional actions}
-When actions are multi-dimensional, factored Gaussian policies are typically employed, so the results extends trivially from the one-dimensional case.
+\paragraph{Multi-dimensional actions.}
+When actions are multi-dimensional, factored Gaussian policies are typically employed, so the results extend trivially from the one-dimensional case.
 Actual multi-variate Gaussian distributions would require more calculations, but we do not expect substantially different results.
 
-\paragraph{Non-linear mean}
+\paragraph{Non-linear mean.}
 In complex continuous tasks, $\mu(s)$ often represents a deep neural network, or multi-layer perceptron, where $\vtheta$ are the weights of the network.
@@
 
-\paragraph{Adaptive variance}
+\paragraph{Adaptive variance.}
 It is a common practice to learn also the variance of the policy in order to adapt the degree of exploration.

### 2018-02-08 svrpg.tex
expanded section 7

--- 2018-02-08 svrpg.tex
@@
 \end{align*}
-In the rest of this section we assume this practical variant.
+In the following we will always assume this practical variant is employed.
 
@@
 \section{Experiments}\label{sec:exp}
-In this section we address the performance evaluation of our policy gradient estimator defined in equation \ref{E:svrpg.estimate.batch} against that one of equation \ref{E:policygradient.estimate}.
-This comparison is performed using G(PO)MDP as estimator of $\score{\vtheta}{\tau}\Reward(\tau)$.
-For the sake of simplicity and with a little abuse of notation we refer to \ref{E:svrpg.estimate.batch} as SVRPG and to \ref{E:policygradient.estimate} as G(PO)MDP.
-In all the experiments we used gaussian policies and bootsrap confidence intervals $\alpha = 10\% $ over ten different runs for evaluation.
-Each run with a different policy initialization which is kept fixed when changing the algorithm.
-The maximum number of trajectories at hand is fixed for each task.
-In \emph{figure} \ref{fig:swimmerone} we can see how SVRPG outperforms G(PO)MDP over the swimmer task.
-Clearly the mean is nearly always higher for our algorithm, moreover near the end we see that te confidence intervals are also disjoint.
-In \emph{figure} \ref{fig:swimmertwo} we addressed the comparison between SVRPG and its self-normalized version still for the swimmer task.
-We can see an inprovement on the average return, but the confidence intervals are never disjoint.
-In \emph{figure} \ref{fig:halfcheetah} the difference between our algorithm and G(PO)MDP over the half-cheetah task is not statistically significant (both of them use the same baseline even though is not optimal for SVRPG see \ref{} for further hints over this topic), however we can see a certain improvement on the mean of the objective function.
-Eventually in \emph{figure} \ref{fig:cartpole} we can clearly see that SVRPG outperforms GPOMDP in the cart-pole task.
+In this section we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks.
+This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases.
+For our algorithm, we use a batch size $N=100$, a mini-batch size $B=10$, and the jointly adaptive step size $\alpha$ and epoch length $m$ proposed in Section \ref{sec:stopping}.
+Since the aim of this comparison is to show the improvement that SVRG-flavored variance reduction brings to SG in the policy gradient framework, we set the batch size of the baseline policy gradient algorithm to $B$.
+In this sense, we measure the improvement yielded by computing snapshot gradients and using them to adjust parameter updates.
+Since we evaluate on-line performance over the number of sampled trajectories, the cost of computing such snapshot gradients is automatically taken into consideration.
+To make the comparison fair, we also use Adam in the baseline PG algorithm, which we will denote simply as G(PO)MDP in the following.
+In all the experiments we used deep Gaussian policies with adaptive standard deviation (details on network architecture in Appendix \ref{app:exp}).
+Each experiment is run 10 times with a random policy initialization, different each time, shared by all the algorithms under comparison.
+The length of the experiment, \ie the total number of trajectories, is fixed for each task.
+Performance is evaluated by using test-trajectories on a subset of the policies considered during the learning process.
+Of course test trajectories are not counted, since they are only meant to improve the accuracy of reported results and do not affect learning.
+We provide average performance with 90\% bootstrap confidence intervals.
+Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}\footnote{https://github.com/rll/rllab}, on which our agents are also based.
+
+\paragraph{Cart-pole.} Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on the continuous cart-pole balancing task.
+Here the performance gain is evident.
+Notice also how the variance of our algorithm is visibly smaller.
+
+\begin{figure}[t]
+	\includegraphics[width=0.5\textwidth]{cart_pole_SVRPG_vs_GPOMDP}
+	\caption{SVRPG vs G(PO)MDP in the Cart-pole task}
+	\label{fig:cartpole}
+\end{figure}
+
+\paragraph{Swimmer} The mujoco swimmer task is far more difficult than cart-pole.
+The longer horizon and the more complex dynamics can have a dangerous impact on the variance of importance weights.
+In this case, the self-normalization technique proposed in Section \ref{sec:selfnorm} brings an improvement, as shown in Figure \ref{fig:swimmertwo}.
+Although the confidence intervals are never disjoint, the average of the self-normalized version is always higher, which we consider enough to employ the variant in the subsequent comparison.
+Figure \ref{fig:swimmerone} shows self-normalized SVRPG against G(PO)MDP.
+On average, our algorithm dominates for almost the entire task.
+Toward the end of the task, the improvement becomes statistically significant.
+Also in this case, the different width of confidence intervals suggest that SVRPG is more robust, especially in the early iterations.
+
+\begin{figure}[t]
+	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
+	\caption{Self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task}
+	\label{fig:swimmertwo}
+\end{figure}
 
@@
 	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP}
-    \caption{Self-Normalized SVRPG vs GPOMDP in swimmer}
+	\caption{Self-Normalized SVRPG vs G(PO)MDP in the Swimmer task.}
 	\label{fig:swimmerone}
 \end{figure}
-\begin{figure}[t]
-    \includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
-    \caption{Self-Normalized SVRPG vs Non Self-Normalized SVRPG in swimmer}
-    \label{fig:swimmertwo}
-\end{figure}
-\begin{figure}[t]
-    \includegraphics[width=0.5\textwidth]{half_cheetah_self_norm_baseline_SVRPG_vs_GPOMDP100_baseline}
-    \caption{Self-Normalized SVRPG with baseline vs GPOMDP with baseline in half cheetah}
-    \label{fig:halfcheetah}
-\end{figure}
-\begin{figure}[t]
-    \includegraphics[width=0.5\textwidth]{cart_pole_SVRPG_vs_GPOMDP}
-    \caption{SVRPG vs GPOMDP in cart-pole}
-    \label{fig:cartpole}
-\end{figure}
+
+%\begin{figure}[t]
+%    \includegraphics[width=0.5\textwidth]{half_cheetah_self_norm_baseline_SVRPG_vs_GPOMDP100_baseline}
+%    \caption{Self-Normalized SVRPG with baseline vs GPOMDP with baseline in half cheetah}
+%    \label{fig:halfcheetah}
+%\end{figure}
 
@@
 \todomat{Repetition of Section 6?}
-Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on several benchmarks.
+Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
 

### 2018-02-08 svrpg.tex
minor fix in Lemma B6

@@
 \left|\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
--\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
+-\Ets[t-1]{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
 \right|
@@
 	&=
-	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} +
+	\Ets[t-1]{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} +
 	\Ets[t-1]{\dotprod{e_s}{\varphi(\vtheta_t^{s+1})}} \label{eq:7}\\
 	&=
-	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
+	\Ets[t-1]{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
 	&\qquad+
-	\dotprod{\Ets{e_s}}{\Ets{\varphi(\vtheta_t^{s+1})}}
+	\dotprod{\Ets[t-1]{e_s}}{\Ets[t-1]{\varphi(\vtheta_t^{s+1})}}
 	+\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}  \label{eq:new1}\\
 	&= 
-	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
+	\Ets[t-1]{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
 	&\qquad+
@@
 \end{align}
-where~\eqref{eq:6} is from Definition~\ref{def:ideal};~\eqref{eq:7} is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $\varphi(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$;~\eqref{eq:new1} is from the fact that $\gradJ{\wt{\vtheta}^s}$ is constant \wrt $\Vars{\cdot}$;~\eqref{eq:8} is from $\Ets{e_s}=0$.
-\todopir{In Eq~\ref{eq:7} the first expected value is on $t-1$? In any case even $t$ is fine.}
+where~\eqref{eq:6} is from Definition~\ref{def:ideal};~\eqref{eq:7} is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $\varphi(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$;~\eqref{eq:new1} is from the fact that $\gradJ{\wt{\vtheta}^s}$ is constant \wrt $\Vars{\cdot}$;~\eqref{eq:8} is from $\Ets[\cdot]{e_s}=0$.
 Hence:

### 2018-02-08 svrpg.tex
reverted last change
Minor fixes

--- 2018-02-08 svrpg.tex
@@
 \begin{abstract}
-        In this paper, we propose a stochastic variance-reduced version of policy gradient for solving Markov Decision Processes (MDPs).
+        In this paper, we propose a novel reinforcement-learning algorithm consisting in a stochastic variance-reduced version of policy gradient for solving Markov Decision Processes (MDPs).
         Stochastic variance-reduced gradient (SVRG) methods have proven to be very successful in supervised learning.
@@
 In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate.
-{\color{red}In Section \ref{sec:conv} we will show how the estimation accuracy impacts on the convergence results with non-concave objective.}
+In Section \ref{sec:conv} we will show how the estimation accuracy impacts on the convergence results with a non-concave objective.
 Finally, the non-stationarity of the optimization problem introduces a bias into the SVRG estimator in Eq.~\eqref{E:svrg.gradient.correction}.
@@
 \end{align}
-We want just to stress that $\EVV[\tau_i \sim p(\cdot|\vtheta_t)]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to Appendix~\ref{} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}
+It is worth noting that the full gradient and the correction term have the same expected value: $\EVV[\tau_i \sim p(\cdot|\vtheta_t)]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to Appendix~\ref{A:gradient_estimators} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}
 This property will be used to prove Lemma~\ref{L:svrpg.properties}.
@@
 %\end{align*}
-The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}.
+The use of mini-batches is also common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}.
 It is easy to show that the SVRPG estimator has the following, desirable properties:
@@
 \end{align*}
-Note that it is important that the trajectories used in the second and the third term are the same for the variance to vanish.
+Note that it is essential that the trajectories used in the second and the third term are the same for the variance to vanish.
 \end{proof}
@@
 Each of the three challenges presented at the beginning of Section~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
-In Appendix~\ref{app:gauss} we show how the traditional Gaussian policy satisfies these assumptions.
+In Appendix~\ref{app:gauss} we show how Gaussian policies satisfy these assumptions.
 
@@
 Once the last two terms can be neglected, the number of trajectories needed to achieve $\norm[2]{\gradJ{\vtheta}}^2\leq\epsilon$ is $O(\frac{B+\nicefrac{N}{m}}{\epsilon})$.
-In this sense, advantage over batch gradient ascent can be achieved with properly selected meta-parameters.
+In this sense, an advantage over batch gradient ascent can be achieved with properly selected meta-parameters.
 In Section~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$.
@@
 \end{align*}
-In the following we will always assume this practical variant is employed.
+In the following, we will always assume this practical variant is employed.
 
@@
 % The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
-Epoch length $m$ influences the variance introduced by the importance weights: low values of $m$ are associated with low variance but increase the frequency of snapshot points (which means many FG computations); high values of $m$ may move policy $\pi_{\vtheta_t}$ far away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing high variance in the importance weights.
-We propose to jointly adapt the two meta-parameters.
+Epoch length $m$ influences the variance introduced by the importance weights.
+Low values of $m$ are associated with small variance but increase the frequency of snapshot points (which means many FG computations).
+High values of $m$ may move policy $\pi_{\vtheta_t}$ far away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing large variance in the importance weights.
+We propose to adapt the two meta-parameters jointly.
 % the farther the distance between the snapshot policy and the current one, the higher the variance of the importance weights.
@@
 Recall that SVRPG performs two different updates of the parameters $\vtheta$: I) FG update in the snapshot; II) corrected gradient update in the sub-iterations.
-Given this structure, we suggest to use two separate ADAM estimators:
+Given this structure, we suggest using two separate ADAM estimators:
 \begin{align*}
@@
 \end{align*}
-where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s+1,t}$ with the sub-iterations (see Appendix~\ref{app:practicalsvrpg} for details).
+where $\alpha^{\textsc{FG}}_{s}$ is associated with the snapshot and $\alpha^{\textsc{SI}}_{s+1,t}$ with the sub-iterations (see Appendix~\ref{app:practicalsvrpg} for details).
 By doing so, we decouple the contribution of the variance due to the approximate FG from the one introduced by the sub-iterations.
@@
 If the latter condition is verified, it amounts to say that the noise in the corrected gradient has overcome the information of the FG.
-Formally, the stopping condition is as follow
+Formally, the stopping condition is as follows
 \[
@@
 \end{align*}
-In Section \ref{sec:exp} we show some cases in which self-normalization provides a performance improvement.
+In Section \ref{sec:exp} we show that self-normalization can provide a performance improvement.
 
@@
 \section{Experiments}\label{sec:exp}
-In this section we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks.
+In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks.
 This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases.
@@
 Performance is evaluated by using test-trajectories on a subset of the policies considered during the learning process.
-Of course test trajectories are not counted, since they are only meant to improve the accuracy of reported results and do not affect learning.
+Of course, test trajectories are not counted, since they are only meant to improve the accuracy of reported results and do not affect learning.
 We provide average performance with 90\% bootstrap confidence intervals.
@@
 where~\eqref{eq:6} is from Definition~\ref{def:ideal};~\eqref{eq:7} is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $\varphi(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$;~\eqref{eq:new1} is from the fact that $\gradJ{\wt{\vtheta}^s}$ is constant \wrt $\Vars{\cdot}$;~\eqref{eq:8} is from $\Ets{e_s}=0$.
-\todopir{In Eq~\ref{eq:7} the first expected value is on $t-1$? In any case even $t$ is fine.}
 Hence:

### 2018-02-08 svrpg.tex
version for first submission (minus todos)

@@
 
-\subsection{Zeroth update}
+\subsection{Zeroth Update}
 As noted in Section \ref{sec:alg}, the update performed at the beginning of each epoch is equivalent to a full-gradient update.
@@
 \paragraph{Cart-pole.} Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on the continuous cart-pole balancing task.
-Here the performance gain is evident.
-Notice also how the variance of our algorithm is visibly smaller.
+Despite using more trajectories for each parameter, our algorithm shows faster convergence, which can be ascribed to the
+better quality of updates due to variance reduction.
 
@@
 The longer horizon and the more complex dynamics can have a dangerous impact on the variance of importance weights.
-In this case, the self-normalization technique proposed in Section \ref{sec:selfnorm} brings an improvement, as shown in Figure \ref{fig:swimmertwo}.
-Although the confidence intervals are never disjoint, the average of the self-normalized version is always higher, which we consider enough to employ the variant in the subsequent comparison.
+In this case, the self-normalization technique proposed in Section \ref{sec:prac} brings an improvement, as shown in Figure \ref{fig:swimmertwo}.
+Although the confidence intervals are never disjoint, the average of the self-normalized version is always higher, which we consider enough to employ the variant in the next comparison.
 Figure \ref{fig:swimmerone} shows self-normalized SVRPG against G(PO)MDP.
 On average, our algorithm dominates for almost the entire task.
+Also here, we note an increase of speed in early iterations.
 Toward the end of the task, the improvement becomes statistically significant.
-Also in this case, the different width of confidence intervals suggest that SVRPG is more robust, especially in the early iterations.
 

### 2018-02-09 svrpg.tex
Corrected learning rate in appendices

--- 2018-02-08 svrpg.tex
@@
 	NN Hidden layers & 8 & 32x32 & 32x32 \\
-	Learning rate & $10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
+	Learning rate SVRPO & $5\cdotp10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
+	Learning rate GPOMDP & $10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
 	Snapshot trajectories & 100 & 100 & 100 \\

### 2018-02-09 svrpg.tex
Updated

@@
 \subsection{Normalized Importance Sampling}\label{sec:prac}
-As mentioned in Section~\ref{sec:stopping}, importance weights are an additional source of variance. A standard way to cope with this issue is self-normalization \citep{precup2000eligibility}.
-This technique can reduce the variance of the importance weights at the cost of introducing some bias. Whether the trade-off is advantageous depends on the specific task.  
+As mentioned in Section~\ref{sec:stopping}, importance weights are an additional source of variance.
+A standard way to cope with this issue is self-normalization~\citep[\eg][]{precup2000eligibility,owenmcbook}.
+This technique can reduce the variance of the importance weights at the cost of introducing some bias~\citep[][Chapter 9]{owenmcbook}.
+Whether the trade-off is advantageous depends on the specific task.  
 Introducing self-normalization in the context of our algorithm, we switch from Eq.~\eqref{E:svrpg.estimate.batch} to:
 \begin{align*}
-\blacktriangledown J(\vtheta_{t}) = \wh{\nabla}_N J(\wt{\vtheta})
- + \frac{1}{B} \sum_{i=0}^{B-1} \left[
-g(\tau_i|\vtheta_t)\right]\\ 
-- \frac{1}{ \sum_{i=0}^{B-1}\omega(\tau_i|\vtheta_t, \wt{\vtheta})} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
+        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + \frac{1}{B} \sum_{i=0}^{B-1} \left[g(\tau_i|\vtheta_t)\right]\\ 
+                                          &\qquad{} - \frac{1}{\Omega} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
 \right].
 \end{align*}
+where $\Omega = \sum_{i=0}^{B-1}\omega(\tau_i|\vtheta_t, \wt{\vtheta})$.
 In Section \ref{sec:exp} we show that self-normalization can provide a performance improvement.

### 2018-02-09 svrpg.tex
Updated
finished Appenddix B (Proofs)

@@
 \end{align*}
-where $\Tr(\cdot)$ denotes the trace of a matrix. $\Covts[t]{X}{Y}$ and $\Varts[t]{X}$ are defined in the same way from $\Ets[t]{X}$.
+where $\Tr(\cdot)$ denotes the trace of a matrix. From the linearity of expected value we have the following:
+\begin{equation}\label{neweq:1}
+\Vars{X} = \Es{\norm[]{X-\Es{X}}^2} \leq \Es{\norm[]{X}^2}
+\end{equation}
+$\Covts[t]{X}{Y}$ and $\Varts[t]{X}$ are defined in the same way from $\Ets[t]{X}$.
 \end{definition}
@@
 For G(PO)MDP, we do not have a compact expression for $g$, but since it is derived from REINFORCE by neglecting some terms of the kind $\nabla \log \pi_{\vtheta}{a_t | s_t}\gamma^t r_t$ \cite{baxter2001infinite,peters2008reinforcement}, the above bound still holds.
-\todopir{we need to prove it formally if we have time}
 \end{proof}
@@
 	%
-	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
+	+\Vars{\gradApp{\wt{\vtheta}^s}{N}} \nonumber\\
+	&\qquad+ 
+	\mathbb{E}_{t\vert s}\left[\left\|
+	\frac{1}{B}\sum_{i=0}^{B-1}\left(
+	g(\tau_i\vert\vtheta_t^{s+1}) -
+	\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)\right)
+	%	\right.\right.\nonumber\\&\qquad\left.\left.
+	- \Ets{
+		\frac{1}{B}\sum_{i=0}^{B-1}\left(
+		g(\tau_i\vert\vtheta_t^{s+1}) -
+		\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)\right)}\right\|^2\right] 
+	\label{neweq:2}\\
+	%
+	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
 	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}
@@
 \endgroup
-where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$\todopirout{ref}, (\ref{eq:3}) is from Lemma \ref{lemma:gsmooth},
+where (\ref{neweq:2}) is from (\ref{neweq:1}), (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from (\ref{neweq:1}), (\ref{eq:3}) is from Lemma \ref{lemma:gsmooth}, 
 (\ref{eq:4}) is from Lemma \ref{lemma:gbound}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
-\todopir{Explain step $E[\|\|^2] \leq Var$}
 \end{proof}
@@
 	\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2},
-	\nonumber
+	\label{neweq:3}
 \end{align}
@@
 \]
-\eqref{eq:9} is from Young's inequality, and~\eqref{eq:10} is from the definition of $\gradApp{\vtheta}{N}$.
-\todopir{comment the last step.}
+\eqref{eq:9} is from Young's inequality,~\eqref{eq:10} is from the definition of $\gradApp{\vtheta}{N}$, and \eqref{neweq:3} is from \eqref{neweq:1}.
 \end{proof}

### 2018-02-09 svrpg.tex
put correct cartpole plot in paper
Updated

--- 2018-02-09 svrpg.tex
@@
 The na\"ive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples, \ie by interacting with the environment.
-For this reason, literature has focused on the introduction of baselines aiming to reduce the variance~\citep[\eg][]{Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.\footnote{A baseline is a function $f(s,a)$ such that $\EVV[a \sim \pol]{f(s,a)} = 0$, for any $s\in\Sspace$.
-\todopir{check the definition of baseline}}
+For this reason, literature has focused on the introduction of baselines (\ie functions $b : \mathcal{S} \times \mathcal{A} \to \realspace$) aiming to reduce the variance~\citep[\eg][]{Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.
 There has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
@@
 
+There are been a surge of work investigating variance reduction techniques for policy gradient methods.
+% The most widespread technique of reducing the variance preserving the unbiasedness of the gradient is through baselines~\citep[\eg][]{williams1992simple,weaver2001optimal,peters2008reinforcement,Thomas2017actionbaseline,wu2018variance}.
+The specific structure of the policy gradient allows to incorporated a baseline (\ie a function $b :\mathcal{S}\times\mathcal{A} \to \realspace$) without affect the unbiasedness of the gradient~\citep[\eg][]{williams1992simple,weaver2001optimal,peters2008reinforcement,Thomas2017actionbaseline,wu2018variance}.
+Although the baseline can be arbitrarily selected, literature often refer to the optimal baseline as the one minimizing the variance of the estimate.
+Despite these approaches and SVRG has the same goal, they are substantially different.
+In particular, the proposed SVRPG does not make explicit use of the structure of the policy gradient framework.
+This suggests that would be possible to integrate an ad-hoc SVRPG baseline in order to further reduce the variance of the estimate.
+But since this paper is about the applicability of SVRG technique to RL we consider this topic as future work.
+Additionally, the experiments show the advantage of using simple SVRPG against G(PO)MPD with baseline (see experiment of \textcolor{blue}{XXX} in Section~\ref{sec:exp}).
+
 Concerning importance weighting techniques, RL has made wide use of them for off-policy problems~\citep[\eg][]{precup2000eligibility,thomas2015high}.
@@
 Here, importance weighting is just a statistical tool used to preserve the unbiasedness of the corrected gradient.
-
-\todopir{Maybe we can talk about variance reduction baselines}
 
@@
 \underbracket{
-        \left(\sum_{h=0}^{H}\nabla \log \pol(a_h^n\vert s_h^n) \right)\left(\sum_{h=0}^{H}\gamma^h \Reward(s_h^n,a_h^n) - b\right)
+        \left(\sum_{h=0}^{H}\nabla \log \pol(z_h^n) \right)\left(\sum_{h=0}^{H}\gamma^h \Reward(z_h^n) - b\right)
 }_{g(\tau_n|\vtheta):=\nabla \log p(\tau_n|\vtheta)\Reward(\tau_n)}
@@
 \underbracket{
-        \sum_{h=0}^{H}\left(\sum_{k=0}^{h} \nabla \log \pol (a_h^n\vert s_h^n) \right)\left(\gamma^h \Reward(s_h^n,a_h^n) - b\right)
+        \sum_{h=0}^{H}\left(\sum_{k=0}^{h} \nabla \log \pol (z_h^n) \right)\left(\gamma^h \Reward(z_h^n) - b\right)
 }_{g(\tau_n|\vtheta)}.
 \end{align*}
-G(PO)MDP can be seen as a more efficient implementation
-of the REINFORCE algorithm. In fact the latter does not perform
-an optimal credit assignment, since it ignores that the reward at time $t$ does
-not depend on the action performed after time $t$. G(PO)MDP overcomes this
-issue taking into account the causality of rewards in the REINFORCE definition
-of policy gradient.
+G(PO)MDP can be seen as a more efficient implementation of the REINFORCE algorithm. 
+In fact the latter does not perform an optimal credit assignment, since it ignores that the reward at time $t$ does not depend on the action performed after time $t$.
+G(PO)MDP overcomes this issue taking into account the causality of rewards in the REINFORCE definition of policy gradient.
 
@@
         \nabla J(\pi^T) = \frac{1}{N} \sum_{n=1}^{N} \omega(\tau^n|\pi^B, \pi^T)
-        \left(\sum_{h=0}^{H} \nabla \log \pi^{T}(a_h^n\vert s_h^n) \right)\left(\sum_{h=0}^{H}\gamma^h \Reward(s_h^n,a_h^n)\right).
-\]
-
+        \underbracket{
+                \left(\sum_{h=0}^{H} \nabla \log \pi^{T}(z_h^n) \right)\left(\sum_{h=0}^{H}\gamma^h \Reward(z_h^n)\right)
+        }_{g(\tau_n|\pi^T)}.
+\]
+%
 The G(PO)MDP off-policy estimator is defined as follows
@@
         \nabla J(\pi^T) = \frac{1}{N} \sum_{n=1}^{N}
+        \underbracket{
                 \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) \gamma^h \Reward(z_h^n) \omega(z_{0:h}|\pi^B,\pi^T)
+        }_{\omega(\tau_n|\pi^B,\pi^T)g(\tau_n|\pi^T)}.
 \]

### 2018-02-09 svrpg.tex

@@
 \begin{figure}[t]
-	\includegraphics[width=0.5\textwidth]{cart_pole_SVRPG_vs_GPOMDP}
+	\includegraphics[width=0.5\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP}
 	\caption{SVRPG vs G(PO)MDP in the Cart-pole task}

### 2018-02-09 svrpg.tex
added table with average return

--- 2018-02-09 svrpg.tex
@@
 \usepackage[makeroom]{cancel}
+\usepackage{placeins}
 
@@
 \section{Experiments}\label{sec:exp}
-In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks.
+In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer.
 This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases.
@@
 Performance is evaluated by using test-trajectories on a subset of the policies considered during the learning process.
-Of course, test trajectories are not counted, since they are only meant to improve the accuracy of reported results and do not affect learning.
 We provide average performance with 90\% bootstrap confidence intervals.
 Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}\footnote{https://github.com/rll/rllab}, on which our agents are also based.
-
-\paragraph{Cart-pole.} Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on the continuous cart-pole balancing task.
-Despite using more trajectories for each parameter, our algorithm shows faster convergence, which can be ascribed to the
-better quality of updates due to variance reduction.
+More details on meta-parameters and exhaustive task descriptions are provided in Appendix \ref{app:exp}.
 
 \begin{figure}[t]
-	\includegraphics[width=0.5\textwidth]{cart_pole_SVRPG_vs_GPOMDP}
-	\caption{SVRPG vs G(PO)MDP in the Cart-pole task}
+	\includegraphics[width=0.5\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP}
+	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% confidence intervals}
 	\label{fig:cartpole}
@@
 
-\paragraph{Swimmer} The mujoco swimmer task is far more difficult than cart-pole.
-The longer horizon and the more complex dynamics can have a dangerous impact on the variance of importance weights.
+\begin{figure}[t]
+	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
+	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task, with 90\% confidence intervals.}
+	\label{fig:swimmertwo}
+\end{figure}
+
+\begin{figure}[t]
+	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP}
+	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs G(PO)MDP in the Swimmer task, with 90\% confidence intervals}
+	\label{fig:swimmerone}
+\end{figure}
+
+\begin{table*}[t]
+	\caption{
+		\label{tab:1}
+		Comparison between G(PO)MDP and SVRPG on different tasks. Average return over the whole learning process is reported, together with normalized standard deviation and 90\% confidence intervals.}
+	\centering
+	\begin{tabular}{@{}lccccccc@{}} 
+		\toprule
+		\phantom{abc} & \multicolumn{3}{c}{G(PO)MDP} & \phantom{abc} & \multicolumn{3}{c}{SVRPG} \\
+		\cmidrule{2-4} \cmidrule{6-8}
+		\phantom{abc} & Avg Return & Norm. Std & C.I. & \phantom{abc}
+		& Avg Return & Norm. Std & C.I.
+		\\\cmidrule{2-8}
+		Cart-pole & 618.49 & 21.09 & [570.76, 666.20] & \phantom{abc} 
+		& \textbf{736.31} & 16.51 & [698.96, 773.65] \\ 
+		\phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} \\
+		Swimmer & 39.86 & 3.84 & [31.18, \:\:\: 48.53] & \phantom{abc} & 43.33 & 3.27 & [37.41, \:\:\: 49.25]\\
+		\bottomrule
+	\end{tabular}
+\end{table*}
+
+Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on a continuous variant of the classical Cart-pole task, which is a 2D balancing task.
+Despite using more trajectories for each parameter, our algorithm shows faster convergence, which can be ascribed to the better quality of updates due to variance reduction.
+
+The Swimmer task is a 3D continuous-control locomotion task.
+This is more difficult than cart-pole.
+In particular, the longer horizon and the more complex dynamics can have a dangerous impact on the variance of importance weights.
 In this case, the self-normalization technique proposed in Section \ref{sec:prac} brings an improvement, as shown in Figure \ref{fig:swimmertwo}.
@@
 
-\begin{figure}[t]
-	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
-	\caption{Self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task}
-	\label{fig:swimmertwo}
-\end{figure}
-
-\begin{figure}[t]
-	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP}
-	\caption{Self-Normalized SVRPG vs G(PO)MDP in the Swimmer task.}
-	\label{fig:swimmerone}
-\end{figure}
+Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for on-line learning~\citep[\eg][]{duan2016benchmarking}.
+This corresponds to the area below the curves in the figures.
+Note also that the convergence result Theorem \ref{theo:convergence} is precisely about average return.
+We also report normalized standard deviation and 90\% bootstrap confidence intervals.
+In all the tasks, our algorithm has higher average return and lower normalized standard deviation.
+In the case of Cart-pole, disjoint confidence intervals show that the advantage is statistically significant.
 
@@
 
+\FloatBarrier
 \bibliography{svrpg}

### 2018-02-09 svrpg.tex
added comment on random return at the end of sec. 3

@@
 In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
-\todopir{mention that even if for theoretical guarantees we should return a random solution, this is not a problem for us because we are evaluating the performance on-line. And is always improving}
+{\color{blue}Note that the randomized return value of Algorithm \ref{alg:svrpg} does not affect on-line learning at all, but will be used as a theoretical tool in the next section.}
 
@@
 }
-\todomat{Repetition of Section 6?}
+\todomatout{Repetition of Section 6?}
 Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.

### 2018-02-09 svrpg.tex
Corrected some values

@@
 		\\\cmidrule{2-8}
-		Cart-pole & 618.49 & 21.09 & [570.76, 666.20] & \phantom{abc} 
-		& \textbf{736.31} & 16.51 & [698.96, 773.65] \\ 
+		Cart-pole & 618.49 & 21.09 & [579.82, 657.16] & \phantom{abc} 
+		& \textbf{736.31} & 16.51 & [706.04, 766.57] \\ 
 		\phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} \\
-		Swimmer & 39.86 & 3.84 & [31.18, \:\:\: 48.53] & \phantom{abc} & 43.33 & 3.27 & [37.41, \:\:\: 49.25]\\
+		Swimmer & 39.86 & 3.84 & [31.18, \:\:\: 48.53] & \phantom{abc} & 49.60 & 3.27 & [43.60, \:\:\: 55.60]\\
 		\bottomrule

### 2018-02-09 svrpg.tex
Corrected again

@@
 		\phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} \\
-		Swimmer & 39.86 & 3.84 & [31.18, \:\:\: 48.53] & \phantom{abc} & 49.60 & 3.27 & [43.60, \:\:\: 55.60]\\
+		Swimmer & 39.86 & 3.84 & [32.82, \:\:\: 46.89] & \phantom{abc} & 49.60 & 3.27 & [43.60, \:\:\: 55.60]\\
 		\bottomrule

### 2018-02-09 svrpg.tex
Updated

@@
 \usepackage[makeroom]{cancel}
-\usepackage{placeins}
 
@@
 In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
-{\color{blue}Note that the randomized return value of Algorithm \ref{alg:svrpg} does not affect on-line learning at all, but will be used as a theoretical tool in the next section.}
+\todopir{mention that even if for theoretical guarantees we should return a random solution, this is not a problem for us because we are evaluating the performance on-line. And is always improving}
 
@@
 \section{Experiments}\label{sec:exp}
-In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer.
+In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks.
 This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases.
@@
 Performance is evaluated by using test-trajectories on a subset of the policies considered during the learning process.
+Of course, test trajectories are not counted, since they are only meant to improve the accuracy of reported results and do not affect learning.
 We provide average performance with 90\% bootstrap confidence intervals.
 Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}\footnote{https://github.com/rll/rllab}, on which our agents are also based.
-More details on meta-parameters and exhaustive task descriptions are provided in Appendix \ref{app:exp}.
+
+\paragraph{Cart-pole.} Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on the continuous cart-pole balancing task.
+Despite using more trajectories for each parameter, our algorithm shows faster convergence, which can be ascribed to the
+better quality of updates due to variance reduction.
 
@@
 	\includegraphics[width=0.5\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP}
-	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% confidence intervals}
+	\caption{SVRPG vs G(PO)MDP in the Cart-pole task}
 	\label{fig:cartpole}
@@
 
-\begin{figure}[t]
-	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
-	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task, with 90\% confidence intervals.}
-	\label{fig:swimmertwo}
-\end{figure}
-
-\begin{figure}[t]
-	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP}
-	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs G(PO)MDP in the Swimmer task, with 90\% confidence intervals}
-	\label{fig:swimmerone}
-\end{figure}
-
-\begin{table*}[t]
-	\caption{
-		\label{tab:1}
-		Comparison between G(PO)MDP and SVRPG on different tasks. Average return over the whole learning process is reported, together with normalized standard deviation and 90\% confidence intervals.}
-	\centering
-	\begin{tabular}{@{}lccccccc@{}} 
-		\toprule
-		\phantom{abc} & \multicolumn{3}{c}{G(PO)MDP} & \phantom{abc} & \multicolumn{3}{c}{SVRPG} \\
-		\cmidrule{2-4} \cmidrule{6-8}
-		\phantom{abc} & Avg Return & Norm. Std & C.I. & \phantom{abc}
-		& Avg Return & Norm. Std & C.I.
-		\\\cmidrule{2-8}
-		Cart-pole & 618.49 & 21.09 & [579.82, 657.16] & \phantom{abc} 
-		& \textbf{736.31} & 16.51 & [706.04, 766.57] \\ 
-		\phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} \\
-		Swimmer & 39.86 & 3.84 & [32.82, \:\:\: 46.89] & \phantom{abc} & 49.60 & 3.27 & [43.60, \:\:\: 55.60]\\
-		\bottomrule
-	\end{tabular}
-\end{table*}
-
-Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on a continuous variant of the classical Cart-pole task, which is a 2D balancing task.
-Despite using more trajectories for each parameter, our algorithm shows faster convergence, which can be ascribed to the better quality of updates due to variance reduction.
-
-The Swimmer task is a 3D continuous-control locomotion task.
-This is more difficult than cart-pole.
-In particular, the longer horizon and the more complex dynamics can have a dangerous impact on the variance of importance weights.
+\paragraph{Swimmer} The mujoco swimmer task is far more difficult than cart-pole.
+The longer horizon and the more complex dynamics can have a dangerous impact on the variance of importance weights.
 In this case, the self-normalization technique proposed in Section \ref{sec:prac} brings an improvement, as shown in Figure \ref{fig:swimmertwo}.
@@
 
-Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for on-line learning~\citep[\eg][]{duan2016benchmarking}.
-This corresponds to the area below the curves in the figures.
-Note also that the convergence result Theorem \ref{theo:convergence} is precisely about average return.
-We also report normalized standard deviation and 90\% bootstrap confidence intervals.
-In all the tasks, our algorithm has higher average return and lower normalized standard deviation.
-In the case of Cart-pole, disjoint confidence intervals show that the advantage is statistically significant.
+\begin{figure}[t]
+	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
+	\caption{Self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task}
+	\label{fig:swimmertwo}
+\end{figure}
+
+\begin{figure}[t]
+	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP}
+	\caption{Self-Normalized SVRPG vs G(PO)MDP in the Swimmer task.}
+	\label{fig:swimmerone}
+\end{figure}
 
@@
 }
-\todomatout{Repetition of Section 6?}
+\todomat{Repetition of Section 6?}
 Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
 
-\FloatBarrier
 \bibliography{svrpg}
@@
 \end{align*}
-where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline provided $\EVV[a \sim \pol]{b(s,a)} = 0$, for any $s\in\Sspace$.\todopir{check definition of baseline}
+where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and $b : \mathcal{S}\times\mathcal{A} \to \realspace$~\citep[\eg][]{Thomas2017actionbaseline}.
 The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance \cite{zhao2011analysis} while preserving the unbiasedness:

### 2018-02-09 svrpg.tex
added Lemma in appendix B to explain better a passage

--- 2018-02-09 svrpg.tex
@@
 		\\\cmidrule{2-8}
-		Cart-pole & 618.49 & 21.09 & [579.82, 657.16] & \phantom{abc} 
-		& \textbf{736.31} & 16.51 & [706.04, 766.57] \\ 
+		Cart-pole & 618.49 & 21.09 & [570.76, 666.20] & \phantom{abc} 
+		& \textbf{736.31} & 16.51 & [698.96, 773.65] \\ 
 		\phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} \\
-		Swimmer & 39.86 & 3.84 & [32.82, \:\:\: 46.89] & \phantom{abc} & 49.60 & 3.27 & [43.60, \:\:\: 55.60]\\
+		Swimmer & 39.86 & 3.84 & [31.18, \:\:\: 48.53] & \phantom{abc} & 43.33 & 3.27 & [37.41, \:\:\: 49.25]\\
 		\bottomrule
@@
 
-\begin{definition}
+\begin{definition}\label{def:var}
 For random vectors X, Y:
@@
 \begin{equation}\label{neweq:1}
-\Vars{X} = \Es{\norm[]{X-\Es{X}}^2} \leq \Es{\norm[]{X}^2}
+\Vars{X} = \Es{\norm[]{X-\Es{X}}^2}
 \end{equation}
@@
 \end{proof}
+
+%New Lemma on Variance
+\begin{restatable}[]{lemma}{varineq}\label{lemma:varineq}
+For any random vector X, the variance (as defined in Definition \ref{def:var}), can be bounded as follows:
+\[
+	\Vars{X} \leq \Es{\norm[]{X}^2}.
+\]
+\end{restatable}
+\begin{proof}
+By using basic properties of expected value and scalar variance:
+\begin{align*}
+	\Vars{X} = \Es{\norm[]{X-\Es{X}}^2} &= \Es{\sum_{i=1}^{\dim(X)}\left(X_i-\Es{X_i}\right)^2} = \sum_{i=1}^{\dim(X)}\Es{\left(X_i-\Es{X_i}\right)^2} \\
+	&\leq \sum_{i=1}^{\dim(X)}\Es{X_i^2} = \Es{\sum_{i=1}^{\dim(X)}X_i^2} = \Es{\norm[]{X}^2}.
+\end{align*}
+\end{proof}
+
 %Lemma 2
@@
 \endgroup
-where (\ref{neweq:2}) is from (\ref{neweq:1}), (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from (\ref{neweq:1}), (\ref{eq:3}) is from Lemma \ref{lemma:gsmooth}, 
+where (\ref{neweq:2}) is from (\ref{neweq:1}), (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from Lemma \ref{lemma:varineq}, (\ref{eq:3}) is from Lemma \ref{lemma:gsmooth},
 (\ref{eq:4}) is from Lemma \ref{lemma:gbound}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
@@
 \]
-\eqref{eq:9} is from Young's inequality,~\eqref{eq:10} is from the definition of $\gradApp{\vtheta}{N}$, and \eqref{neweq:3} is from \eqref{neweq:1}.
+\eqref{eq:9} is from Young's inequality,~\eqref{eq:10} is from the definition of $\gradApp{\vtheta}{N}$, and \eqref{neweq:3} is from Lemma \ref{lemma:varineq}.
 \end{proof}

### 2018-02-09 svrpg.tex

@@
 		\\\cmidrule{2-8}
-		Cart-pole & 618.49 & 21.09 & [570.76, 666.20] & \phantom{abc} 
-		& \textbf{736.31} & 16.51 & [698.96, 773.65] \\ 
+		Cart-pole & 618.49 & 21.09 & [579.82, 657.16] & \phantom{abc} 
+		& \textbf{736.31} & 16.51 & [706.04, 766.57] \\ 
 		\phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} \\
-		Swimmer & 39.86 & 3.84 & [31.18, \:\:\: 48.53] & \phantom{abc} & 43.33 & 3.27 & [37.41, \:\:\: 49.25]\\
+		Swimmer & 39.86 & 3.84 & [32.82, \:\:\: 46.89] & \phantom{abc} & 49.60 & 3.27 & [43.60, \:\:\: 55.60]\\
 		\bottomrule

### 2018-02-09 svrpg.tex
Updated

--- 2018-02-09 svrpg.tex
@@
 The na\"ive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples, \ie by interacting with the environment.
-For this reason, literature has focused on the introduction of baselines (\ie functions $b : \mathcal{S} \times \mathcal{A} \to \realspace$) aiming to reduce the variance~\citep[\eg][]{Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.
-There has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
+For this reason, literature has focused on the introduction of baselines (\ie functions $b : \mathcal{S} \times \mathcal{A} \to \realspace$) aiming to reduce the variance~\citep[\eg][]{williams1992simple,Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.
+{\color{red}
+This baselines are usually designed in order to minimize the variance of the gradient estimate but even them need to be estimated from data, partially reducing they effectiveness.
+On the other hand, there has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
 Although these techniques have been mainly derived for finite-sum problems, we will show in Section~\ref{sec:alg} how they can be used in RL.
+In particular, we will shown that the proposed SVRPG algorithm can take the best of both worlds (\ie SL and RL) since it can be plugged over a policy gradient estimate using baselines.
 The next section has the aim to describe variance reduction techniques for finite-sum problems. In particular, we will present the SVRG algorithm that is at the core of this work.
+}
 
@@
 \end{description}
-To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as Gaussian\footnote{see Appendix~\ref{app:gauss} for more details on the Gaussian policy case.} and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}.
+To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as Gaussian\footnote{See Appendix~\ref{app:gauss} for more details on the Gaussian policy case.} and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}.
 Because of the infinite dataset, we can only rely on an estimate of the full gradient.
@@
 The overall structure is the same as Algorithm \ref{alg:svrg}, 
-but the snapshot gradient is not exact and the gradient estimate used between snapshots is corrected using importance weighting:
+but the snapshot gradient is not exact and the gradient estimate used between snapshots is corrected using importance weighting:\footnote{\color{red} Note that $g$ can be any unbiased estimator w/o baseline.
+The unbiasedness is required for theoretical results (\eg Appendix~\ref{A:gradient_estimators}).}
 \begin{align*}
@@
 Although the baseline can be arbitrarily selected, literature often refer to the optimal baseline as the one minimizing the variance of the estimate.
+Nevertheless, even the baseline needs to be estimated from data. This fact may partially reduce its effectiveness by introducing variance.
 Despite these approaches and SVRG has the same goal, they are substantially different.
-In particular, the proposed SVRPG does not make explicit use of the structure of the policy gradient framework.
+In particular, the proposed SVRPG does not make explicit use of the structure of the policy gradient framework and it is independent from the underline gradient estimate (\ie w/o baseline).
 This suggests that would be possible to integrate an ad-hoc SVRPG baseline in order to further reduce the variance of the estimate.
 But since this paper is about the applicability of SVRG technique to RL we consider this topic as future work.
-Additionally, the experiments show the advantage of using simple SVRPG against G(PO)MPD with baseline (see experiment of \textcolor{blue}{XXX} in Section~\ref{sec:exp}).
+Additionally, the experiments show the advantage of using simple SVRPG against G(PO)MPD with baseline (see the \textcolor{blue}{half-cheetah} domain in Section~\ref{sec:exp}).
 
@@
 \end{align*}
-For G(PO)MDP, we do not have a compact expression for $g$, but since it is derived from REINFORCE by neglecting some terms of the kind $\nabla \log \pi_{\vtheta}{a_t | s_t}\gamma^t r_t$ \cite{baxter2001infinite,peters2008reinforcement}, the above bound still holds.
+For G(PO)MDP, we do not have a compact expression for $g$, but since it is derived from REINFORCE by neglecting some terms of the kind $\nabla \log \pi_{\vtheta}(a_t | s_t)\gamma^t r_t$ \cite{baxter2001infinite,peters2008reinforcement}, the above bound still holds.
 \end{proof}

### 2018-02-09 svrpg.tex
added baseline paragraph to sec. 7

--- 2018-02-09 svrpg.tex
@@
 
+\paragraph{Preliminary results on actor-critic.}As mentioned, another variance-reduction technique in policy gradient is to employ baselines, or \textit{critics}.
+This tool is orthogonal to the methods described in this paper, and the theoretical results of Section \ref{sec:conv} are general in this sense.
+In the experiments described so far, we compared against the so-called \textit{actor-only} G(PO)MDP, \ie without baseline.
+In order to move towards a more general understanding of the problem of variance in policy gradient, we also test SVRPG in an \textit{actor-critic} scenario.
+To do so, we use the more challenging mujoco Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer.
+Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms.
+Results are promising, showing that a combination of baseline usage and SVRG-like variance reduction can yield an improvement that the two techniques alone are not able to achieve.
+
 %\begin{figure}[t]	

### 2018-02-09 svrpg.tex
mentioned baseline in proofs
Updated section 5

--- 2018-02-09 svrpg.tex
@@
 Here we provide a practical and automatic way to choose the step size $\alpha$ and the number of sub-iterations $m$ performed between snapshots.
-Additionally, we provide variants of SVRPG exploiting variance-reduction techniques for importance weights.
-Despite lacking theoretical guarantees, we will show that these methods can outperform the baseline SVRPG (Algorithm~\ref{alg:svrpg}).
-
-\subsection{Zeroth Update}
+Additionally, we provide a variant of SVRPG exploiting a variance-reduction technique for importance weights.
+Despite lacking theoretical guarantees, we will show in Section~\ref{sec:exp} that this method can outperform the baseline SVRPG (Algorithm~\ref{alg:svrpg}).
+
+\subsection{Full Gradient Update}
 As noted in Section \ref{sec:alg}, the update performed at the beginning of each epoch is equivalent to a full-gradient update.
-In our setting, where collecting samples is particularly expensive, the $B$ trajectories collected using the snapshot trajectory $\pi_{\tilde{\vtheta}^s}$ feel like a waste of data.
-In practice, we just perform an approximate full gradient update using the $N$ trajectories sampled to compute $\gradApp{\tilde{\vtheta}^s}{N}$, \ie
-\begin{align*}
-	&\vtheta_{1}^{s+1} = \tilde{\vtheta}^s + \alpha\gradApp{\tilde{\vtheta}^s}{N} \\
-	&\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v_t^{s+1} \text{ for $t=1,\dots,m-1$}.
+In our setting, where collecting samples is particularly expensive, the $B$ trajectories collected using the snapshot trajectory $\pi_{\wt{\vtheta}^s}$ feels like a waste of data (the term $\sum_i g(\tau_i) - \omega(\tau_i) g(\tau_i) =0$ since $\vtheta_0 = \wt{\vtheta}$).
+In practice, we just perform an approximate full gradient update using the $N$ trajectories sampled to compute $\gradApp{\wt{\vtheta}^s}{N}$, \ie
+\begin{align*}
+	&\vtheta_{1}^{s+1} = \wt{\vtheta}^s + \alpha\gradApp{\wt{\vtheta}^s}{N} \\
+	&\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha 
+%        v_t^{s+1}
+        \blacktriangledown J(\vtheta^{s+1}_t)
+        \text{ for $t=1,\dots,m-1$}.
 \end{align*}
@@
 % At this point, the open question is how to properly select the step size $\alpha$ and the sub-iteration number $m$ in practical applications.
-The step size $\alpha$ is crucial to balance variance reduction and efficiency.
+The step size $\alpha$ is crucial to balance variance reduction and efficiency, while
 % The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
-Epoch length $m$ influences the variance introduced by the importance weights.
+the epoch length $m$ influences the variance introduced by the importance weights.
 Low values of $m$ are associated with small variance but increase the frequency of snapshot points (which means many FG computations).
@@
         \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
-        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} v_t^{s+1} \text{ for $t=1,\dots,m-1$},
+        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} 
+%        v_t^{s+1}
+        \blacktriangledown J(\vtheta^{s+1}_t)
+        \text{ for $t=1,\dots,m-1$},
 \end{align*}

### 2018-02-09 svrpg.tex

@@
 Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
+\todomat{add future work}
 
@@
 These terms have bounded gradient from the second inequality of Assumption \ref{ass:bounded_score} and the fact that $|r_t|\leq R$.
-Bounded gradient implies Lipschitz continuity.
-Finally, the linear combination of Lipschitz continuous functions is Lipschitz continuous.
+If a baseline is used in REINFORCE or G(PO)MDP, we only need the additional assumption that said baseline is bounded.
+        Bounded gradient implies Lipschitz continuity. Finally, the linear combination of Lipschitz continuous functions is Lipschitz continuous.
 \end{proof}
@@
 For G(PO)MDP, we do not have a compact expression for $g$, but since it is derived from REINFORCE by neglecting some terms of the kind $\nabla \log \pi_{\vtheta}(a_t | s_t)\gamma^t r_t$ \cite{baxter2001infinite,peters2008reinforcement}, the above bound still holds.
+If a baseline is used in REINFORCE or G(PO)MDP, we only need the additional assumption that said baseline is bounded.
 \end{proof}

### 2018-02-09 svrpg.tex
Corrected appendices experiments setting

--- 2018-02-09 svrpg.tex
@@
 \section{Experimental Details}\label{app:exp}
-Setting specifications for all the used environments:
+Used environments:
 \begin{enumerate}
@@
 
+Refer to \cite{duan2016benchmarking}  for more details about GPOMDP in Half Cheetah task.
 All the experiments' parameters are reported in the following table:
@@
 	Control space dimension & 1 & 2 & 6 \\
-	NN Hidden layers & 8 & 32x32 & 32x32 \\
-	Learning rate SVRPO & $5\cdotp10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
-	Learning rate GPOMDP & $10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
-	Snapshot trajectories & 100 & 100 & 100 \\
-	Sub-iterations trajectories & 10 & 10 & 10 \\
-	Max sub-iterations & 50 & 20 & 20 \\
-	Trajectories lengths& 500 & 500 & 500 \\
+	NN Hidden layers & 8 & 32x32 & 100x50x25 \\
+	Adam learning rate SVRPO & $5\cdotp10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
+	Adam learning rate GPOMDP & $10^{-2}$ & $10^{-3}$ & $10^{-2}$ \\
+	Snapshot batch size & 100 & 100 & 100 \\
+	Sub-iterations batch size & 10 & 10 & 10 \\
+	GPOMDP batch size & 10 & 10 & 100 \\
+	Max number of sub-iterations & 50 & 20 & 20 \\
+	Episodes horizon& 100 & 500 & 500 \\
+	Baseline& No & No & Yes \\
 	Discount $(\gamma)$& 0.99 & 0.995 & 0.99 \\
-	Total trajectories& 10000 & 20000 & 50000 \\
+	Total number of trajectories& 10000 & 20000 & 50000 \\
 	\hline  

### 2018-02-09 svrpg.tex
Updated

--- 2018-02-09 svrpg.tex
@@
 \section{Experiments}\label{sec:exp}
-In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer.
-This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases.
+In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer.\todopir{ref, even Benchmarking Deep Reinforcement Learning for Continuous Control is fine}
+We consider G(PO)MDP since has a smaller variance than REINFORCE.
+% This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases. 
 For our algorithm, we use a batch size $N=100$, a mini-batch size $B=10$, and the jointly adaptive step size $\alpha$ and epoch length $m$ proposed in Section \ref{sec:stopping}.
@@
 In all the experiments we used deep Gaussian policies with adaptive standard deviation (details on network architecture in Appendix \ref{app:exp}).
-Each experiment is run 10 times with a random policy initialization, different each time, shared by all the algorithms under comparison.
+Each experiment is run $10$ times with a random policy initialization and seed but this initialization is shared among the algorithms under comparison.
 The length of the experiment, \ie the total number of trajectories, is fixed for each task.
@@
 We provide average performance with 90\% bootstrap confidence intervals.
-Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}\footnote{https://github.com/rll/rllab}, on which our agents are also based.
+Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}, on which our agents are also based.\footnote{rllab can be found at \url{https://github.com/rll/rllab}, while the link to our implementation is not reported for blindness but it is already publically available.}
 More details on meta-parameters and exhaustive task descriptions are provided in Appendix \ref{app:exp}.
@@
 Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on a continuous variant of the classical Cart-pole task, which is a 2D balancing task.
-Despite using more trajectories for each parameter, our algorithm shows faster convergence, which can be ascribed to the better quality of updates due to variance reduction.
+Despite using more trajectories for each parameter,\todopirout{I do not understand this sentence} our algorithm shows faster convergence, which can be ascribed to the better quality of updates due to variance reduction.
 
@@
 
-Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for on-line learning~\citep[\eg][]{duan2016benchmarking}.
+Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for on-line learning since it is related to an average regret~\citep[\eg][]{duan2016benchmarking}.
 This corresponds to the area below the curves in the figures.
-Note also that the convergence result Theorem \ref{theo:convergence} is precisely about average return.
+Note that the convergence result Theorem \ref{theo:convergence} is precisely about average return.
 We also report normalized standard deviation and 90\% bootstrap confidence intervals.
 In all the tasks, our algorithm has higher average return and lower normalized standard deviation.
-In the case of Cart-pole, disjoint confidence intervals show that the advantage is statistically significant.
-
-\paragraph{Preliminary results on actor-critic.}As mentioned, another variance-reduction technique in policy gradient is to employ baselines, or \textit{critics}.
+In Cart-pole, disjoint confidence intervals show that the advantage is statistically significant.
+
+\textbf{Preliminary results on actor-critic.} %As mentioned, 
+Another variance-reduction technique in policy gradient is to employ baselines, or \textit{critics}.
 This tool is orthogonal to the methods described in this paper, and the theoretical results of Section \ref{sec:conv} are general in this sense.
@@
 In order to move towards a more general understanding of the problem of variance in policy gradient, we also test SVRPG in an \textit{actor-critic} scenario.
-To do so, we use the more challenging mujoco Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer.
+To do so, we use the more challenging MuJoCo~\citep{todorov2012mujoco} Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer.
 Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms.

### 2018-02-09 svrpg.tex
added half-cheetah data, note on REINFORCE/PGT in rllab and minor fixes

--- 2018-02-09 svrpg.tex
@@
 	\begin{restatable}[On policy derivatives]{assumption}{boundedscore}\label{ass:bounded_score}
-		For each state-action pair $(s,a)$, value of $\vtheta$ and dimensions $i,j$ there are positive constants $G,F<\infty$ such that:
+		For each state-action pair $(s,a)$, any value of $\vtheta$ and all parameter components $i,j$ there are positive constants $G,F<\infty$ such that:
 \[
@@
 By analysing the upper-bound in Theorem~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on non-concave SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
-To achieve true convergence, both the batch size $N$ and the mini-batch size $B$ should increase over time. 
+To achieve asymptotic convergence, the batch size $N$ and the mini-batch size $B$ should increase over time. 
 In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by FG approximation and importance sampling, respectively.
@@
 	\includegraphics[width=0.5\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP}
-	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% confidence intervals}
+	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% t-student confidence intervals.
+Results denoted with $\dagger$ are taken from \cite{duan2016benchmarking}.}
 	\label{fig:cartpole}
@@
 		\label{tab:1}
-		Comparison between G(PO)MDP and SVRPG on different tasks. Average return over the whole learning process is reported, together with normalized standard deviation and 90\% confidence intervals.}
+		Comparison between G(PO)MDP and SVRPG on different tasks.
+Average return over the whole learning process is reported, together with normalized standard deviation and 90\% confidence intervals.
+Results denoted with $\dagger$ are taken from \cite{duan2016benchmarking}.}
 	\centering
@@
 		\\\cmidrule{2-8}
-		Cart-pole & 618.49 & 21.09 & [579.82, 657.16] & \phantom{abc} 
-		& \textbf{736.31} & 16.51 & [706.04, 766.57] \\ 
-		\phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} & \phantom{abc} \\
-		Swimmer & 39.86 & 3.84 & [32.82, \:\:\: 46.89] & \phantom{abc} & 49.60 & 3.27 & [43.60, \:\:\: 55.60]\\
+		Cart-pole & $618.49$ & $21.09$ & $[579.82, 657.16]$ & \phantom{abc} 
+		& $736.31$ & $16.51$ & $[706.04, 766.57]$ \\
+		Swimmer & $39.86$ & $3.84$ & $[32.82, 46.89]$ & \phantom{abc} & $49.60$ & $3.27$ & $[43.60, 55.60]$\\
+		Half-cheetah & $1183.1^\dagger$ & $69.20^\dagger$ & $[1035.58, 1330.62]^\dagger$ & \phantom{abc} & $1480.95$ & $73.21$ & $[1324.87, 1637.03]$\\
 		\bottomrule
@@
 To do so, we use the more challenging mujoco Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer.
-Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms.
+Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP\footnote{\cite{duan2016benchmarking} report results on REINFORCE.
+However, inspection on \textit{rllab} code and documentation reveals that it is actually PGT \cite{sutton2000policy}, which is equivalent to G(PO)MDP \citep[shown by][]{peters2008reinforcement}.
+Using the name REINFORCE in a general way is inaccurate, but widespread.} on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms.
 Results are promising, showing that a combination of baseline usage and SVRG-like variance reduction can yield an improvement that the two techniques alone are not able to achieve.
@@
 \section{Policy Gradient Estimators} \label{A:gradient_estimators}
-We present a brief overview of two most widespread gradient estimators (REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}) both in on-policy and off-policy settings.
+We present a brief overview of the two most widespread gradient estimators (REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}) both in on-policy and off-policy settings.
 Let $\tau = \{\langle s_t,a_t \rangle\}_{t=0}^{H}= \{z_t\}_{t=0}^{H} = z_{0:H}$ is a $(H+1)$-steps trajectory.
@@
 \section{Experimental Details}\label{app:exp}
-Used environments:
+Setting specifications for all the used environments:
 \begin{enumerate}
@@
 
-Refer to \cite{duan2016benchmarking}  for more details about GPOMDP in Half Cheetah task.
 All the experiments' parameters are reported in the following table:
@@
 	Control space dimension & 1 & 2 & 6 \\
-	NN Hidden layers & 8 & 32x32 & 100x50x25 \\
-	Adam learning rate SVRPO & $5\cdotp10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
-	Adam learning rate GPOMDP & $10^{-2}$ & $10^{-3}$ & $10^{-2}$ \\
-	Snapshot batch size & 100 & 100 & 100 \\
-	Sub-iterations batch size & 10 & 10 & 10 \\
-	GPOMDP batch size & 10 & 10 & 100 \\
-	Max number of sub-iterations & 50 & 20 & 20 \\
-	Episodes horizon& 100 & 500 & 500 \\
-	Baseline& No & No & Yes \\
+	NN Hidden layers & 8 & 32x32 & 32x32 \\
+	Learning rate SVRPO & $5\cdotp10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
+	Learning rate GPOMDP & $10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
+	Snapshot trajectories & 100 & 100 & 100 \\
+	Sub-iterations trajectories & 10 & 10 & 10 \\
+	Max sub-iterations & 50 & 20 & 20 \\
+	Trajectories lengths& 500 & 500 & 500 \\
 	Discount $(\gamma)$& 0.99 & 0.995 & 0.99 \\
-	Total number of trajectories& 10000 & 20000 & 50000 \\
+	Total trajectories& 10000 & 20000 & 50000 \\
 	\hline  

### 2018-02-09 svrpg.tex
Added ref

--- 2018-02-09 svrpg.tex
@@
 \section{Experiments}\label{sec:exp}
-In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer.\todopir{ref, even Benchmarking Deep Reinforcement Learning for Continuous Control is fine}
+In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer~\citep[\eg][]{duan2016benchmarking}.
 We consider G(PO)MDP since has a smaller variance than REINFORCE.

### 2018-02-09 svrpg.tex
learning rates and baseline

--- 2018-02-09 svrpg.tex
@@
 \end{algorithm}
-\todopir{Explain how we use adam and when/if it is reset on snapshots}
-
+\subsection{Learning Rates}
+Let us give a deeper insight on the two different learning rate used by our algorithm.
+We have already said that we use two different ADAM one related to the snapshots and another one related to the sub-iterations.
+In this way tha ADAM associated to the snapshots will take into account only the snapshots history through the parameters $m_{t-1}$ and $v_{t-1}$.
+The same reasoning holds for the sub-iterations.
+This, of course, allows our algorithm to take into account the different variabilities in a more accurate manner.
+\subsection{Baseline}
+For all batch gradient-based algorithms, we use the same time-varying feature encoding for the linear baseline:
+\begin{align*}
+\psi_{s,t}=concat(s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1)
+\end{align*}
+where $s$ is the state vector and $\odot$ is the element-wise product.
 \section{Experimental Details}\label{app:exp}

### 2018-02-09 svrpg.tex
added footnote on PGT/REINFORCE

--- 2018-02-09 svrpg.tex
@@
 \section{Experiments}\label{sec:exp}
-In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer~\citep[\eg][]{duan2016benchmarking}.
+In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer.\todopir{ref, even Benchmarking Deep Reinforcement Learning for Continuous Control is fine}
 We consider G(PO)MDP since has a smaller variance than REINFORCE.
@@
 To do so, we use the more challenging MuJoCo~\citep{todorov2012mujoco} Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer.
-Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms.
+Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP\footnote{\cite{duan2016benchmarking} report results on REINFORCE.
+However, inspection on \textit{rllab} code and documentation reveals that it is actually PGT \cite{sutton2000policy}, which is equivalent to G(PO)MDP \citep[shown by][]{peters2008reinforcement}.
+Using the name REINFORCE in a general way is inaccurate, but widespread.} on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms.
 Results are promising, showing that a combination of baseline usage and SVRG-like variance reduction can yield an improvement that the two techniques alone are not able to achieve.
-
-%\begin{figure}[t]	
-%    \includegraphics[width=0.5\textwidth]{half_cheetah_self_norm_baseline_SVRPG_vs_GPOMDP100_baseline}
-%    \caption{Self-Normalized SVRPG with baseline vs GPOMDP with baseline in half cheetah}
-%    \label{fig:halfcheetah}
-%\end{figure}
 

### 2018-02-09 svrpg.tex
typo

@@
 Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on a continuous variant of the classical Cart-pole task, which is a 2D balancing task.
-Despite using more trajectories for each parameter,\todopirout{I do not understand this sentence} our algorithm shows faster convergence, which can be ascribed to the better quality of updates due to variance reduction.
+{\color{blue}Despite using more trajectories for each parameter update}, our algorithm shows faster convergence, which can be ascribed to the better quality of updates due to variance reduction.
 

### 2018-02-09 svrpg.tex
Updated

--- 2018-02-09 svrpg.tex
@@
 The na\"ive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples, \ie by interacting with the environment.
-For this reason, literature has focused on the introduction of baselines (\ie functions $b : \mathcal{S} \times \mathcal{A} \to \realspace$) aiming to reduce the variance~\citep[\eg][]{williams1992simple,Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.
+For this reason, literature has focused on the introduction of baselines (\ie functions $b : \mathcal{S} \times \mathcal{A} \to \realspace$) aiming to reduce the variance~\citep[\eg][]{williams1992simple,Peters2008reinf,Thomas2017actionbaseline,wu2018variance}, see Appendix~\ref{A:gradient_estimators} for a formal definition of $b$.
 {\color{red}
-This baselines are usually designed in order to minimize the variance of the gradient estimate but even them need to be estimated from data, partially reducing they effectiveness.
+These baselines are usually designed in order to minimize the variance of the gradient estimate but even them need to be estimated from data, partially reducing they effectiveness.
 On the other hand, there has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
@@
 To do so, we use the more challenging MuJoCo~\citep{todorov2012mujoco} Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer.
-Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP\footnote{\cite{duan2016benchmarking} report results on REINFORCE.
-However, inspection on \textit{rllab} code and documentation reveals that it is actually PGT \cite{sutton2000policy}, which is equivalent to G(PO)MDP \citep[shown by][]{peters2008reinforcement}.
-Using the name REINFORCE in a general way is inaccurate, but widespread.} on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms.
+Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms.
 Results are promising, showing that a combination of baseline usage and SVRG-like variance reduction can yield an improvement that the two techniques alone are not able to achieve.
+
+%\begin{figure}[t]	
+%    \includegraphics[width=0.5\textwidth]{half_cheetah_self_norm_baseline_SVRPG_vs_GPOMDP100_baseline}
+%    \caption{Self-Normalized SVRPG with baseline vs GPOMDP with baseline in half cheetah}
+%    \label{fig:halfcheetah}
+%\end{figure}
 
@@
 We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
-{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
-The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
-In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
-}
-\todomatout{Repetition of Section 6?}
+%{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
+%The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
+%In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
+%}
+%\todomatout{Repetition of Section 6?}
 Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
-\todomat{add future work}
+Preliminary results support the effectiveness of SVRPG also with a commonly used baseline for the policy gradient.
+Despite that, we believe that it will be possible to derive a baseline specifically designed for SVRPG in order to jointly exploit the RL structure and the SVRG idea.
+\todomat{Other future work}
 

### 2018-02-09 svrpg.tex
added a future work
Updated

@@
 \begin{align*}
-        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
-        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} 
+        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\left(\wh{\nabla}_N J(\wt{\vtheta}^s) \right)\\
+        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t}\left( 
 %        v_t^{s+1}
-        \blacktriangledown J(\vtheta^{s+1}_t)
+        \blacktriangledown J(\vtheta^{s+1}_t)\right)
         \text{ for $t=1,\dots,m-1$},
@@
 \begin{equation}\label{E:onpolicygradient}
-        \gradJ{\vtheta} = \EVV[\tau \sim p(\cdot|\pol)]{\nabla \log\Delta(\tau|\vtheta)\Reward(\tau)} = \EVV[\tau \sim p(\cdot|\pol)]{\sum_{j=0}^{H} \gamma^j \mathcal{R}(z_j) \sum_{i=0}^{H} \nabla \log\pi(z_i)}.
+        \gradJ{\vtheta} = \EVV[\tau \sim p(\cdot|\pol)]{\nabla \log p(\tau|\vtheta)\Reward(\tau)} = \EVV[\tau \sim p(\cdot|\pol)]{\sum_{j=0}^{H} \gamma^j \mathcal{R}(z_j) \sum_{i=0}^{H} \nabla \log\pi(z_i)}.
 \end{equation}
@@
 \underbracket{
-        \left(\sum_{h=0}^{H}\nabla \log \pol(z_h^n) \right)\left(\sum_{h=0}^{H}\gamma^h \Reward(z_h^n) - b\right)
+        \left(\sum_{h=0}^{H}\nabla \log \pol(z_h^n) \right)\left(\sum_{h=0}^{H}\gamma^h 
+%        \Reward(z_h^n)
+        r_h^n
+- b(z_h^n)\right)
 }_{g(\tau_n|\vtheta):=\nabla \log p(\tau_n|\vtheta)\Reward(\tau_n)}
@@
 \end{align*}
-where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and $b : \mathcal{S}\times\mathcal{A} \to \realspace$~\citep[\eg][]{Thomas2017actionbaseline}.
+where subscripts denote the time step, superscripts denote the trajectory, $r_h^n$ is the reward actually collected at time $h$ from trajectory $\tau^n$ and $b : \mathcal{S}\times\mathcal{A} \to \realspace$~\citep[\eg][]{Thomas2017actionbaseline}.
 The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance \cite{zhao2011analysis} while preserving the unbiasedness:
@@
 \underbracket{
-        \sum_{h=0}^{H}\left(\sum_{k=0}^{h} \nabla \log \pol (z_h^n) \right)\left(\gamma^h \Reward(z_h^n) - b\right)
+        \sum_{h=0}^{H}\left(\sum_{k=0}^{h} \nabla \log \pol (z_h^n) \right)\left(\gamma^h 
+%        \Reward(z_h^n)
+        r_h^n
+        - b(z^n_h)\right)
 }_{g(\tau_n|\vtheta)}.

### 2018-02-09 svrpg.tex
baseline

--- 2018-02-09 svrpg.tex
@@
 The na\"ive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples, \ie by interacting with the environment.
-For this reason, literature has focused on the introduction of baselines (\ie functions $b : \mathcal{S} \times \mathcal{A} \to \realspace$) aiming to reduce the variance~\citep[\eg][]{williams1992simple,Peters2008reinf,Thomas2017actionbaseline,wu2018variance}, see Appendix~\ref{A:gradient_estimators} for a formal definition of $b$.
+For this reason, literature has focused on the introduction of baselines (\ie functions $b : \mathcal{S} \times \mathcal{A} \to \realspace$) aiming to reduce the variance~\citep[\eg][]{williams1992simple,Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.
 {\color{red}
-These baselines are usually designed in order to minimize the variance of the gradient estimate but even them need to be estimated from data, partially reducing they effectiveness.
+This baselines are usually designed in order to minimize the variance of the gradient estimate but even them need to be estimated from data, partially reducing they effectiveness.
 On the other hand, there has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
@@
 We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
-%{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
-%The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
-%In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
-%}
-%\todomatout{Repetition of Section 6?}
+{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
+The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
+In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
+}
+\todomatout{Repetition of Section 6?}
 Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
-Preliminary results support the effectiveness of SVRPG also with a commonly used baseline for the policy gradient.
-Despite that, we believe that it will be possible to derive a baseline specifically designed for SVRPG in order to jointly exploit the RL structure and the SVRG idea.
-\todomat{Other future work}
+\todomat{add future work}
 
@@
 \end{algorithm}
-\todopir{Explain how we use adam and when/if it is reset on snapshots}
-
+\subsection{Learning Rates}
+Let us give a deeper insight on the two different learning rate used by our algorithm.
+We have already said that we use two different ADAM one related to the snapshots and another one related to the sub-iterations.
+In this way tha ADAM associated to the snapshots will take into account only the snapshots history through the parameters $m_{t-1}$ and $v_{t-1}$.
+The same reasoning holds for the sub-iterations.
+This, of course, allows our algorithm to take into account the different variabilities in a more accurate manner.
+\subsection{Baseline}
+For all batch gradient-based algorithms, we use the same time-varying feature encoding for the linear baseline:
+\begin{align*}
+\psi_{s,t}=concat(s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1)
+\end{align*}
+where $s$ is the state vector and $\odot$ is the element-wise product. In the case of our algorithm the model representing the baseline is fitted every snapshot so we have:\\
+\begin{align*}
+        \nabla J(\pi^T) =\frac{1}{N} \sum_{n=1}^{N}
+                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(y_k^n)\right) (\gamma^h \Reward(y_h^n)-b_{sn})
+          + \frac{1}{B} \sum_{n=1}^{B}
+                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\\ - \frac{1}{B} \sum_{n=1}^{B}
+                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\omega(z_{0:h}|\pi^B,\pi^T)
+\end{align*}
+Where $b_{sn}$ is the prediction given by the baseline at the snapshot, whereas $b_{si}$ is the prediction given by the baseline at a given subiteration. 
 \section{Experimental Details}\label{app:exp}

### 2018-02-09 svrpg.tex

--- 2018-02-09 svrpg.tex
@@
 Despite that, we believe that it will be possible to derive a baseline specifically designed for SVRPG in order to jointly exploit the RL structure and the SVRG idea.
-\todomat{Other future work}
+Future work should also focus on making batch sizes $N$ and $B$ adaptive, as suggested in \cite{papini2017adaptive}.
 
@@
 \end{align*}
-where $s$ is the state vector and $\odot$ is the element-wise product.
+where $s$ is the state vector and $\odot$ is the element-wise product. In the case of our algorithm the model representing the baseline is fitted every snapshot so we have:\\
+\begin{align*}
+        \nabla J(\pi^T) =\frac{1}{N} \sum_{n=1}^{N}
+                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(y_k^n)\right) (\gamma^h \Reward(y_h^n)-b_{sn})
+          + \frac{1}{B} \sum_{n=1}^{B}
+                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\\ - \frac{1}{B} \sum_{n=1}^{B}
+                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\omega(z_{0:h}|\pi^B,\pi^T)
+\end{align*}
+Where $b_{sn}$ is the prediction given by the baseline at the snapshot, whereas $b_{si}$ is the prediction given by the baseline at a given subiteration. 
 \section{Experimental Details}\label{app:exp}

### 2018-02-09 svrpg.tex
Updated
moved proof of SVRPG properties to appendix

--- 2018-02-09 svrpg.tex
@@
 It is easy to show that the SVRPG estimator has the following, desirable properties:
-\begin{lemma}\label{L:svrpg.properties}
+\begin{restatable}[]{lemma}{properties}\label{L:svrpg.properties}
         Let $\wh{\nabla}_N J(\vtheta)$ be an unbiased estimator of~\eqref{E:policygradient}
@@
 \end{equation}
-\end{lemma}
-\begin{proof}
-Claim (\ref{eq:unbiased}) is from the unbiasedness of the estimators:
-\begin{align*}
-\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} \\
-&\qquad- \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
-&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
-\end{align*}
-As for claim (\ref{eq:zerovar}), note that as $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
-\begin{align*}
-\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*)\right. \\
-&\left.- \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right]
-= \Var\left[\gradApp{\vtheta^*}{N}\right].
-\end{align*}
-Note that it is essential that the trajectories used in the second and the third term are the same for the variance to vanish.
-\end{proof}
+\end{restatable}
+
 Previous results hold for both REINFORCE and G(PO)MDP.
@@
 \begin{align*}
-        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\left(\wh{\nabla}_N J(\wt{\vtheta}^s) \right)\\
-        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t}\left( 
+        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
+        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} 
 %        v_t^{s+1}
-        \blacktriangledown J(\vtheta^{s+1}_t)\right)
+        \blacktriangledown J(\vtheta^{s+1}_t)
         \text{ for $t=1,\dots,m-1$},
@@
 \begin{equation}\label{E:onpolicygradient}
-        \gradJ{\vtheta} = \EVV[\tau \sim p(\cdot|\pol)]{\nabla \log p(\tau|\vtheta)\Reward(\tau)} = \EVV[\tau \sim p(\cdot|\pol)]{\sum_{j=0}^{H} \gamma^j \mathcal{R}(z_j) \sum_{i=0}^{H} \nabla \log\pi(z_i)}.
+        \gradJ{\vtheta} = \EVV[\tau \sim p(\cdot|\pol)]{\nabla \log\Delta(\tau|\vtheta)\Reward(\tau)} = \EVV[\tau \sim p(\cdot|\pol)]{\sum_{j=0}^{H} \gamma^j \mathcal{R}(z_j) \sum_{i=0}^{H} \nabla \log\pi(z_i)}.
 \end{equation}
@@
 \underbracket{
-        \left(\sum_{h=0}^{H}\nabla \log \pol(z_h^n) \right)\left(\sum_{h=0}^{H}\gamma^h 
-%        \Reward(z_h^n)
-        r_h^n
-- b(z_h^n)\right)
+        \left(\sum_{h=0}^{H}\nabla \log \pol(z_h^n) \right)\left(\sum_{h=0}^{H}\gamma^h \Reward(z_h^n) - b\right)
 }_{g(\tau_n|\vtheta):=\nabla \log p(\tau_n|\vtheta)\Reward(\tau_n)}
@@
 \end{align*}
-where subscripts denote the time step, superscripts denote the trajectory, $r_h^n$ is the reward actually collected at time $h$ from trajectory $\tau^n$ and $b : \mathcal{S}\times\mathcal{A} \to \realspace$~\citep[\eg][]{Thomas2017actionbaseline}.
+where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and $b : \mathcal{S}\times\mathcal{A} \to \realspace$~\citep[\eg][]{Thomas2017actionbaseline}.
 The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance \cite{zhao2011analysis} while preserving the unbiasedness:
@@
 \underbracket{
-        \sum_{h=0}^{H}\left(\sum_{k=0}^{h} \nabla \log \pol (z_h^n) \right)\left(\gamma^h 
-%        \Reward(z_h^n)
-        r_h^n
-        - b(z^n_h)\right)
+        \sum_{h=0}^{H}\left(\sum_{k=0}^{h} \nabla \log \pol (z_h^n) \right)\left(\gamma^h \Reward(z_h^n) - b\right)
 }_{g(\tau_n|\vtheta)}.
@@
 %\end{proof}
+
+\subsection*{Basic Properties of SVRPG}
+\properties
+\begin{proof}
+	Claim (\ref{eq:unbiased}) is from the unbiasedness of the estimators:
+	\begin{align*}
+	\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} \\
+	&\qquad- \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
+	&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
+	\end{align*}
+	As for claim (\ref{eq:zerovar}), note that as $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
+	\begin{align*}
+	\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*)\right. \\
+	&\left.- \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right]
+	= \Var\left[\gradApp{\vtheta^*}{N}\right].
+	\end{align*}
+	Note that it is essential that the trajectories used in the second and the third term are the same for the variance to vanish.
+\end{proof}
 
@@
 \section{Practical SVRPG versions}\label{app:practicalsvrpg}
+In this appendix we provide more details on the practical variants of our algorithm.
+\subsection{Learning Rates}
 \begin{algorithm}[h]
@@
 \end{algorithm}
-\subsection{Learning Rates}
-Let us give a deeper insight on the two different learning rate used by our algorithm.
-We have already said that we use two different ADAM one related to the snapshots and another one related to the sub-iterations.
-In this way tha ADAM associated to the snapshots will take into account only the snapshots history through the parameters $m_{t-1}$ and $v_{t-1}$.
+Let us give a deeper insight on the two different learning rate schedules used by our algorithm.
+We report pseudo-code of the original ADAM in Algorithm \ref{A:adam}.
+As mentioned, we use two distinct instances of ADAM: one related to the snapshots, and one to the sub-iterations.
+In this way the ADAM associated to the snapshots will take into account only the history of gradient moments at the snapshots.
 The same reasoning holds for the sub-iterations.

### 2018-02-09 svrpg.tex

@@
 \begin{align*}
-        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
-        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} 
+        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\left(\wh{\nabla}_N J(\wt{\vtheta}^s) \right)\\
+        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} \Big( 
 %        v_t^{s+1}
-        \blacktriangledown J(\vtheta^{s+1}_t)
+        \blacktriangledown J(\vtheta^{s+1}_t)\Big)
         \text{ for $t=1,\dots,m-1$},

### 2018-02-09 svrpg.tex
fixed lemma renaming
Updated

@@
                 \[
-                        \EVV[\tau_i \sim \pi(\cdot|\vtheta)]{\wh{\nabla}_N J(\vtheta)} = \EVV[\tau_i \sim \pi(\cdot|\vtheta)]{\frac{1}{N} \sum_{i=0}^N g(\tau_i|\vtheta)} = \nabla J(\vtheta)
+                        \EVV[\tau_i \sim p(\cdot|\vtheta)]{\wh{\nabla}_N J(\vtheta)} = \EVV[\tau_i \sim p(\cdot|\vtheta)]{\frac{1}{N} \sum_{i=0}^N g(\tau_i|\vtheta)} = \nabla J(\vtheta)
                 \]
@@
                 \[
-        \EVV[\tau_i \sim \Delta(\cdot|\pi^B)]{\frac{1}{N} \sum_{i=0}^N \omega(\tau_i|\pi^B,\vtheta) g(\tau_i|\vtheta) } 
+        \EVV[\tau_i \sim p(\cdot|\pi^B)]{\frac{1}{N} \sum_{i=0}^N \omega(\tau_i|\pi^B,\vtheta) g(\tau_i|\vtheta) } 
         = \nabla J(\vtheta) 
@@
 We give some additional definitions which will be useful in the proofs.
+In the rest of the proof we will always consider the $\ell_2$-norm when not explicitly stated.
 
@@
 \subsection*{Basic Properties of SVRPG}
-\properties
+\properties*
 \begin{proof}
@@
 %New Lemma on Variance
-\begin{restatable}[]{lemma}{varineq}\label{lemma:varineq}
+The following lemma is trivial but we report for sake of clarity.
+
+\begin{restatable}[Variance is upper-bounded by second moment.]{lemma}{varineq}\label{lemma:varineq}
 For any random vector X, the variance (as defined in Definition \ref{def:var}), can be bounded as follows:
@@
 \begin{proof}
-By using basic properties of expected value and scalar variance:
+        By using Equation~\eqref{neweq:1} and the definition of variance of a scalar random variable:
 \begin{align*}
 	\Vars{X} = \Es{\norm[]{X-\Es{X}}^2} &= \Es{\sum_{i=1}^{\dim(X)}\left(X_i-\Es{X_i}\right)^2} = \sum_{i=1}^{\dim(X)}\Es{\left(X_i-\Es{X_i}\right)^2} \\
-	&\leq \sum_{i=1}^{\dim(X)}\Es{X_i^2} = \Es{\sum_{i=1}^{\dim(X)}X_i^2} = \Es{\norm[]{X}^2}.
+                                     &=\sum_{i=1}^{\dim(X)} \Es{X_i^2} - \left(\Es{X_i} \right)^2 \leq \sum_{i=1}^{\dim(X)}\Es{X_i^2}
+%    = \Es{\sum_{i=1}^{\dim(X)}X_i^2} 
+                                        = \Es{\norm[]{X}^2}.
 \end{align*}
@@
 		\label{eq:1}\\%
-	&\leq \Ets[t-1]{\norm[2]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
 	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} 
 %    \nonumber\\ &\qquad
-    +\Ets{\norm[2]{
+    +\Ets{\norm[]{
 			\frac{1}{B}\sum_{i=0}^{B-1}\left(
@@
 	%
-	&\leq \Ets[t-1]{\norm[2]{\gradJ{\vtheta_t^{s+1}}}^2} 
+	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
 	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}
@@
 			+ \frac{1}{B^2}\sum_{i=0}^{B-1}
-			\Ets{\norm[2]{
+			\Ets{\norm[]{
 			g(\tau_i\vert\vtheta_t^{s+1}) -
@@
 \endgroup
-where (\ref{neweq:2}) is from (\ref{neweq:1}), (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from Lemma \ref{lemma:varineq}, (\ref{eq:3}) is from Lemma \ref{lemma:gsmooth},
+where (\ref{neweq:2}) is from (\ref{neweq:1}), (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$,~\ref{eq:2} is from Lemma~\ref{lemma:varineq}, (\ref{eq:3}) is from Lemma \ref{lemma:gsmooth},
 (\ref{eq:4}) is from Lemma \ref{lemma:gbound}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.
@@
 2-dimensional action space: the two momentums applied on actuated joints.
-The reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2] a^2$.
+The reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2]{a}^2$.
 The episodes terminates when the number of time steps T is greater than 500.
@@
 6-dimensional action space: the 6 momentums applied on actuated joints.
-The reward function is defined as $r(s, a) := v_x - 0.05\norm[2] a^2$.
+The reward function is defined as $r(s, a) := v_x - 0.05\norm[2]{a}^2$.
 The episodes terminates when the number of time steps T is greater than 500.

### 2018-02-09 svrpg.tex
updating appendix D

--- 2018-02-09 svrpg.tex
@@
 		\STATE \textbf{Input:} A gradient estimate $g_t$ and parameters $\beta_1$, $\beta_2$, $\epsilon$ and $\alpha$.
-		\STATE $m_t = \beta_1 m_{t-1} + (1 - \beta_1) g_t$
-		\STATE $v_t = \beta_2 v_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
-		\STATE $\hat{m}_t = \dfrac{m_t}{1 - \beta^t_1}$
-		\STATE $\hat{v}_t = \dfrac{v_t}{1 - \beta^t_2}$
-		\STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{v}_t} + \epsilon} \hat{m}_t$
+		\STATE $\kappa_t = \beta_1 \kappa_{t-1} + (1 - \beta_1) g_t$
+		\STATE $\nu_t = \beta_2 \nu_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
+		\STATE $\hat{\kappa}_t = \dfrac{\kappa_t}{1 - \beta^t_1}$
+		\STATE $\hat{\nu}_t = \dfrac{\nu_t}{1 - \beta^t_2}$
+		\STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{\nu}_t} + \epsilon} \hat{\kappa}_t$
 		\STATE \textbf{Return:} The increment $\Delta(g_t)$  of the parameters.
@@
 We report pseudo-code of the original ADAM in Algorithm \ref{A:adam}.
-As mentioned, we use two distinct instances of ADAM: one related to the snapshots, and one to the sub-iterations.
-In this way the ADAM associated to the snapshots will take into account only the history of gradient moments at the snapshots.
-The same reasoning holds for the sub-iterations.
-This, of course, allows our algorithm to take into account the different variabilities in a more accurate manner.
+As mentioned, we use two distinct instances of ADAM to manage different sources of variance: one related to the snapshots, and one to the sub-iterations.
+In this way the ADAM associated to the snapshots takes into account only the history of gradient moments at the snapshots.
+Using the notation in Algorithm \ref{A:adam}, we can explicitly define our gradient updates:
+\begin{align*}
+\vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\left(\wh{\nabla}_N J(\wt{\vtheta}^s) \right)\\
+\vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} \Big( 
+%        v_t^{s+1}
+\blacktriangledown J(\vtheta^{s+1}_t)\Big)
+\text{ for $t=1,\dots,m-1$},
+\end{align*}
+
 \subsection{Baseline}

### 2018-02-09 svrpg.tex
Minor fixes to related works

--- 2018-02-09 svrpg.tex
@@
 It is easy to show that the SVRPG estimator has the following, desirable properties:
-\begin{restatable}[]{lemma}{properties}\label{L:svrpg.properties}
+\begin{lemma}\label{L:svrpg.properties}
         Let $\wh{\nabla}_N J(\vtheta)$ be an unbiased estimator of~\eqref{E:policygradient}
@@
 \end{equation}
-\end{restatable}
-
+\end{lemma}
+\begin{proof}
+Claim (\ref{eq:unbiased}) is from the unbiasedness of the estimators:
+\begin{align*}
+\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} \\
+&\qquad- \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
+&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
+\end{align*}
+As for claim (\ref{eq:zerovar}), note that as $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
+\begin{align*}
+\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*)\right. \\
+&\left.- \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right]
+= \Var\left[\gradApp{\vtheta^*}{N}\right].
+\end{align*}
+Note that it is essential that the trajectories used in the second and the third term are the same for the variance to vanish.
+\end{proof}
 Previous results hold for both REINFORCE and G(PO)MDP.
@@
         \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\left(\wh{\nabla}_N J(\wt{\vtheta}^s) \right)\\
-        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} \Big( 
+        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t}\left( 
 %        v_t^{s+1}
-        \blacktriangledown J(\vtheta^{s+1}_t)\Big)
+        \blacktriangledown J(\vtheta^{s+1}_t)\right)
         \text{ for $t=1,\dots,m-1$},
@@
 
-There are been a surge of work investigating variance reduction techniques for policy gradient methods.
+In the recent past, there has been a surge of studies investigating variance reduction techniques for policy gradient methods.
 % The most widespread technique of reducing the variance preserving the unbiasedness of the gradient is through baselines~\citep[\eg][]{williams1992simple,weaver2001optimal,peters2008reinforcement,Thomas2017actionbaseline,wu2018variance}.
-The specific structure of the policy gradient allows to incorporated a baseline (\ie a function $b :\mathcal{S}\times\mathcal{A} \to \realspace$) without affect the unbiasedness of the gradient~\citep[\eg][]{williams1992simple,weaver2001optimal,peters2008reinforcement,Thomas2017actionbaseline,wu2018variance}.
-Although the baseline can be arbitrarily selected, literature often refer to the optimal baseline as the one minimizing the variance of the estimate.
+The specific structure of the policy gradient allows incorporating a baseline (\ie a function $b :\mathcal{S}\times\mathcal{A} \to \realspace$) without affecting the unbiasedness of the gradient~\citep[\eg][]{williams1992simple,weaver2001optimal,peters2008reinforcement,Thomas2017actionbaseline,wu2018variance}.
+Although the baseline can be arbitrarily selected, literature often refers to the optimal baseline as the one minimizing the variance of the estimate.
 Nevertheless, even the baseline needs to be estimated from data. This fact may partially reduce its effectiveness by introducing variance.
-Despite these approaches and SVRG has the same goal, they are substantially different.
-In particular, the proposed SVRPG does not make explicit use of the structure of the policy gradient framework and it is independent from the underline gradient estimate (\ie w/o baseline).
-This suggests that would be possible to integrate an ad-hoc SVRPG baseline in order to further reduce the variance of the estimate.
-But since this paper is about the applicability of SVRG technique to RL we consider this topic as future work.
-Additionally, the experiments show the advantage of using simple SVRPG against G(PO)MPD with baseline (see the \textcolor{blue}{half-cheetah} domain in Section~\ref{sec:exp}).
-
-Concerning importance weighting techniques, RL has made wide use of them for off-policy problems~\citep[\eg][]{precup2000eligibility,thomas2015high}.
+Even if these approaches share the same goal as SVRG, they are substantially different.
+In particular, the proposed SVRPG does not make explicit use of the structure of the policy gradient framework, and it is independent of the underlying gradient estimate (\ie w/o baseline).
+This suggests that would be possible to integrate an ad-hoc SVRPG baseline to further reduce the variance of the estimate.
+Since this paper is about the applicability of SVRG technique to RL, we consider this topic as future work.
+Additionally, the experiments show that SVRPG has an advantage over G(PO)MPD even when the baseline is used (see the half-cheetah domain in Section~\ref{sec:exp}).
+
+Concerning importance weighting techniques, RL has made extensive use of them for off-policy problems~\citep[\eg][]{precup2000eligibility,thomas2015high}.
 However, as mentioned before, SVRPG cannot be compared to such methods since it is in all respects an on-policy algorithm.
@@
 %\end{proof}
-
-\subsection*{Basic Properties of SVRPG}
-\properties*
-\begin{proof}
-	Claim (\ref{eq:unbiased}) is from the unbiasedness of the estimators:
-	\begin{align*}
-	\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} \\
-	&\qquad- \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
-	&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
-	\end{align*}
-	As for claim (\ref{eq:zerovar}), note that as $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
-	\begin{align*}
-	\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*)\right. \\
-	&\left.- \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right]
-	= \Var\left[\gradApp{\vtheta^*}{N}\right].
-	\end{align*}
-	Note that it is essential that the trajectories used in the second and the third term are the same for the variance to vanish.
-\end{proof}
 
@@
 \section{Practical SVRPG versions}\label{app:practicalsvrpg}
-In this appendix we provide more details on the practical variants of our algorithm.
-\subsection{Learning Rates}
 \begin{algorithm}[h]
@@
         \STATE \textbf{Input:} A gradient estimate $g_t$ and parameters $\beta_1$, $\beta_2$, $\epsilon$ and $\alpha$.
-		\STATE $\kappa_t = \beta_1 \kappa_{t-1} + (1 - \beta_1) g_t$
-		\STATE $\nu_t = \beta_2 \nu_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
-		\STATE $\hat{\kappa}_t = \dfrac{\kappa_t}{1 - \beta^t_1}$
-		\STATE $\hat{\nu}_t = \dfrac{\nu_t}{1 - \beta^t_2}$
-		\STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{\nu}_t} + \epsilon} \hat{\kappa}_t$
+        \STATE $m_t = \beta_1 m_{t-1} + (1 - \beta_1) g_t$
+        \STATE $v_t = \beta_2 v_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
+        \STATE $\hat{m}_t = \dfrac{m_t}{1 - \beta^t_1}$
+        \STATE $\hat{v}_t = \dfrac{v_t}{1 - \beta^t_2}$
+        \STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{v}_t} + \epsilon} \hat{m}_t$
         \STATE \textbf{Return:} The increment $\Delta(g_t)$  of the parameters.
@@
 \end{algorithm}
-Let us give a deeper insight on the two different learning rate schedules used by our algorithm.
-We report pseudo-code of the original ADAM in Algorithm \ref{A:adam}.
-As mentioned, we use two distinct instances of ADAM to manage different sources of variance: one related to the snapshots, and one to the sub-iterations.
-In this way the ADAM associated to the snapshots takes into account only the history of gradient moments at the snapshots.
-Using the notation in Algorithm \ref{A:adam}, we can explicitly define our gradient updates:
-\begin{align*}
-\vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\left(\wh{\nabla}_N J(\wt{\vtheta}^s) \right)\\
-\vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} \Big( 
-%        v_t^{s+1}
-\blacktriangledown J(\vtheta^{s+1}_t)\Big)
-\text{ for $t=1,\dots,m-1$},
-\end{align*}
-
+\subsection{Learning Rates}
+Let us give a deeper insight on the two different learning rate used by our algorithm.
+We have already said that we use two different ADAM one related to the snapshots and another one related to the sub-iterations.
+In this way tha ADAM associated to the snapshots will take into account only the snapshots history through the parameters $m_{t-1}$ and $v_{t-1}$.
+The same reasoning holds for the sub-iterations.
+This, of course, allows our algorithm to take into account the different variabilities in a more accurate manner.
 \subsection{Baseline}
@@
 \end{align*}
-where $s$ is the state vector and $\odot$ is the element-wise product. In the case of our algorithm the model representing the baseline is fitted every snapshot so we have:\\
-\begin{align*}
-        \nabla J(\pi^T) =\frac{1}{N} \sum_{n=1}^{N}
-                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(y_k^n)\right) (\gamma^h \Reward(y_h^n)-b_{sn})
-          + \frac{1}{B} \sum_{n=1}^{B}
-                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\\ - \frac{1}{B} \sum_{n=1}^{B}
-                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\omega(z_{0:h}|\pi^B,\pi^T)
-\end{align*}
-Where $b_{sn}$ is the prediction given by the baseline at the snapshot, whereas $b_{si}$ is the prediction given by the baseline at a given subiteration. 
+where $s$ is the state vector and $\odot$ is the element-wise product.
 \section{Experimental Details}\label{app:exp}

### 2018-02-09 svrpg.tex

--- 2018-02-09 svrpg.tex
@@
 		\STATE \textbf{Input:} A gradient estimate $g_t$ and parameters $\beta_1$, $\beta_2$, $\epsilon$ and $\alpha$.
-		\STATE $\kappa_t = \beta_1 \kappa_{t-1} + (1 - \beta_1) g_t$
-		\STATE $\nu_t = \beta_2 \nu_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
-		\STATE $\hat{\kappa}_t = \dfrac{\kappa_t}{1 - \beta^t_1}$
-		\STATE $\hat{\nu}_t = \dfrac{\nu_t}{1 - \beta^t_2}$
-		\STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{\nu}_t} + \epsilon} \hat{\kappa}_t$
+		\STATE $m_t = \beta_1 m_{t-1} + (1 - \beta_1) g_t$
+		\STATE $v_t = \beta_2 v_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
+		\STATE $\hat{m}_t = \dfrac{m_t}{1 - \beta^t_1}$
+		\STATE $\hat{v}_t = \dfrac{v_t}{1 - \beta^t_2}$
+		\STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{v}_t} + \epsilon} \hat{m}_t$
 		\STATE \textbf{Return:} The increment $\Delta(g_t)$  of the parameters.
@@
 We report pseudo-code of the original ADAM in Algorithm \ref{A:adam}.
-As mentioned, we use two distinct instances of ADAM to manage different sources of variance: one related to the snapshots, and one to the sub-iterations.
-In this way the ADAM associated to the snapshots takes into account only the history of gradient moments at the snapshots.
-Using the notation in Algorithm \ref{A:adam}, we can explicitly define our gradient updates:
-\begin{align*}
-\vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\left(\wh{\nabla}_N J(\wt{\vtheta}^s) \right)\\
-\vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} \Big( 
-%        v_t^{s+1}
-\blacktriangledown J(\vtheta^{s+1}_t)\Big)
-\text{ for $t=1,\dots,m-1$},
-\end{align*}
-
+As mentioned, we use two distinct instances of ADAM: one related to the snapshots, and one to the sub-iterations.
+In this way the ADAM associated to the snapshots will take into account only the history of gradient moments at the snapshots.
+The same reasoning holds for the sub-iterations.
+This, of course, allows our algorithm to take into account the different variabilities in a more accurate manner.
 \subsection{Baseline}

### 2018-02-09 svrpg.tex
Fixes to experiments and conclusions

--- 2018-02-09 svrpg.tex
@@
 In all the experiments we used deep Gaussian policies with adaptive standard deviation (details on network architecture in Appendix \ref{app:exp}).
-Each experiment is run $10$ times with a random policy initialization and seed but this initialization is shared among the algorithms under comparison.
+Each experiment is run $10$ times with a random policy initialization and seed, but this initialization is shared among the algorithms under comparison.
 The length of the experiment, \ie the total number of trajectories, is fixed for each task.
@@
 We provide average performance with 90\% bootstrap confidence intervals.
-Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}, on which our agents are also based.\footnote{rllab can be found at \url{https://github.com/rll/rllab}, while the link to our implementation is not reported for blindness but it is already publically available.}
+Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}, on which our agents are also based.\footnote{rllab can be found at \url{https://github.com/rll/rllab}, while the link to our implementation is not reported for blindness, but it is already publically available.}
 More details on meta-parameters and exhaustive task descriptions are provided in Appendix \ref{app:exp}.
@@
 Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on a continuous variant of the classical Cart-pole task, which is a 2D balancing task.
-{\color{blue}Despite using more trajectories for each parameter update}, our algorithm shows faster convergence, which can be ascribed to the better quality of updates due to variance reduction.
+Despite using more trajectories on average for each parameter update, our algorithm shows faster convergence, which can be ascribed to the better quality of updates due to variance reduction.
 
 The Swimmer task is a 3D continuous-control locomotion task.
-This is more difficult than cart-pole.
+This task is more difficult than cart-pole.
 In particular, the longer horizon and the more complex dynamics can have a dangerous impact on the variance of importance weights.
-In this case, the self-normalization technique proposed in Section \ref{sec:prac} brings an improvement, as shown in Figure \ref{fig:swimmertwo}.
-Although the confidence intervals are never disjoint, the average of the self-normalized version is always higher, which we consider enough to employ the variant in the next comparison.
+In this case, the self-normalization technique proposed in Section \ref{sec:prac} brings an improvement (even if not statistically significant), as shown in Figure \ref{fig:swimmertwo}.
+% Although the confidence intervals are never disjoint, the average of the self-normalized version is always higher, which we consider enough to employ the variant in the next comparison. 
 Figure \ref{fig:swimmerone} shows self-normalized SVRPG against G(PO)MDP.
-On average, our algorithm dominates for almost the entire task.
-Also here, we note an increase of speed in early iterations.
-Toward the end of the task, the improvement becomes statistically significant.
-
-Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for on-line learning since it is related to an average regret~\citep[\eg][]{duan2016benchmarking}.
-This corresponds to the area below the curves in the figures.
-Note that the convergence result Theorem \ref{theo:convergence} is precisely about average return.
-We also report normalized standard deviation and 90\% bootstrap confidence intervals.
-In all the tasks, our algorithm has higher average return and lower normalized standard deviation.
+Our algorithm outperforms G(PO)MDP for almost the entire learning process.
+Also here, we note an increase of speed in early iterations, and, toward the end of the learning process, the improvement becomes statistically significant.
+
+Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for online learning since it is related to an average regret~\citep[\eg][]{duan2016benchmarking}.
+This metric corresponds to the area below the curves in the figures.
+Note that the convergence result in Theorem \ref{theo:convergence} is precisely about the average return.
+We also report normalized standard deviations and 90\% $t$-confidence intervals.
+In both tasks, our algorithm has higher average return and lower normalized standard deviation.
 In Cart-pole, disjoint confidence intervals show that the advantage is statistically significant.
@@
 \textbf{Preliminary results on actor-critic.} %As mentioned, 
-Another variance-reduction technique in policy gradient is to employ baselines, or \textit{critics}.
+Another variance-reduction technique in policy gradient is to employ baselines or \textit{critics}.
 This tool is orthogonal to the methods described in this paper, and the theoretical results of Section \ref{sec:conv} are general in this sense.
-In the experiments described so far, we compared against the so-called \textit{actor-only} G(PO)MDP, \ie without baseline.
-In order to move towards a more general understanding of the problem of variance in policy gradient, we also test SVRPG in an \textit{actor-critic} scenario.
-To do so, we use the more challenging MuJoCo~\citep{todorov2012mujoco} Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer.
-Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP\footnote{\cite{duan2016benchmarking} report results on REINFORCE.
+In the experiments described so far, we compared against the so-called \textit{actor-only} G(PO)MDP, \ie without the baseline.
+To move towards a more general understanding of the problem of variance in policy gradient, we also test SVRPG in an \textit{actor-critic} scenario.
+To do so, we consider the more challenging MuJoCo~\citep{todorov2012mujoco} Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer.
+Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP\footnote{\citet{duan2016benchmarking} report results on REINFORCE.
 However, inspection on \textit{rllab} code and documentation reveals that it is actually PGT \cite{sutton2000policy}, which is equivalent to G(PO)MDP \citep[shown by][]{peters2008reinforcement}.
 Using the name REINFORCE in a general way is inaccurate, but widespread.} on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms.
-Results are promising, showing that a combination of baseline usage and SVRG-like variance reduction can yield an improvement that the two techniques alone are not able to achieve.
+Results are promising, showing that a combination of the baseline usage and SVRG-like variance reduction can yield an improvement that the two techniques alone are not able to achieve.
 
@@
 Preliminary results support the effectiveness of SVRPG also with a commonly used baseline for the policy gradient.
-Despite that, we believe that it will be possible to derive a baseline specifically designed for SVRPG in order to jointly exploit the RL structure and the SVRG idea.
+Despite that, we believe that it will be possible to derive a baseline designed explicitly for SVRPG to exploit the RL structure and the SVRG idea jointly.
 Future work should also focus on making batch sizes $N$ and $B$ adaptive, as suggested in \cite{papini2017adaptive}.

### 2018-02-09 svrpg.tex
Updated

--- 2018-02-09 svrpg.tex
@@
 \graphicspath{ {images/} }
-\usepackage{subfigure}
+%\usepackage{subfigure}
 \usepackage{booktabs} % for professional tables
@@
 
-\begin{figure}[t]
-	\includegraphics[width=0.5\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP}
+\begin{figure*}[t]
+    \begin{minipage}[b]{0.3\textwidth}
+	\includegraphics[width=.95\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP_tex.pdf}
 	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% t-student confidence intervals.
@@
 	\label{fig:cartpole}
-\end{figure}
-
-\begin{figure}[t]
-	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
+    \end{minipage}
+    \begin{minipage}[b]{0.3\textwidth}
+	\includegraphics[width=0.95\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG_tex.pdf}
 	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task, with 90\% confidence intervals.}
 	\label{fig:swimmertwo}
-\end{figure}
-
-\begin{figure}[t]
-	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP}
+    \end{minipage}
+    \hfill
+        \begin{minipage}[b]{.3\textwidth}
+	\includegraphics[width=0.95\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP_tex.pdf}
 	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs G(PO)MDP in the Swimmer task, with 90\% confidence intervals}
 	\label{fig:swimmerone}
-\end{figure}
+    \end{minipage}
+\end{figure*}
 

### 2018-02-09 svrpg.tex
Fit in 8 pages

@@
 
+\makeatletter
+\g@addto@macro\normalsize{%
+ \setlength\abovedisplayskip{6pt}
+ \setlength\belowdisplayskip{6pt}
+}
+\makeatother
+
 % The \icmltitle you define below is probably too long as a header.
@@
 
+\vspace{-0.05in}
 \section{Introduction}
+\vspace{-0.05in}
 %Similarities between RL and SL
@@
 
+\vspace{-0.05in}
 \section{Preliminaries}\label{sec:pre}
+\vspace{-0.05in}
 In this section, we provide the essential background on policy gradient methods and stochastic variance-reduced gradient methods for finite-sum optimization.
@@
 
+\vspace{-0.05in}
 \subsection{Policy Gradient}\label{subsec:PolicyGradient}
+\vspace{-0.05in}
 A Reinforcement Learning task~\citep{sutton1998reinforcement} can be modelled with a discrete-time continuous Markov Decision Process (MDP) $M = \{\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\}$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$;
@@
 
+\vspace{-0.05in}
 \subsection{Stochastic Variance-Reduced Gradient}\label{sec:svrg}
+\vspace{-0.05in}
 Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $z_i(\cdot|\vtheta)$:
@@
 
+\vspace{-0.05in}
 \section{SVRG in Reinforcement Learning}\label{sec:alg}
+\vspace{-0.05in}
 In online RL problems, the usual approach is to tune the batch size of SG to find the optimal trade-off between variance and speed.
@@
 \end{lemma}
-\begin{proof}
-Claim (\ref{eq:unbiased}) is from the unbiasedness of the estimators:
-\begin{align*}
-\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} \\
-&\qquad- \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
-&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
-\end{align*}
-As for claim (\ref{eq:zerovar}), note that as $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
-\begin{align*}
-\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*)\right. \\
-&\left.- \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right]
-= \Var\left[\gradApp{\vtheta^*}{N}\right].
-\end{align*}
-Note that it is essential that the trajectories used in the second and the third term are the same for the variance to vanish.
-\end{proof}
+%\begin{proof}
+%Claim (\ref{eq:unbiased}) is from the unbiasedness of the estimators:
+%\begin{align*}
+%\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} \\
+%&\qquad- \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
+%&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
+%\end{align*}
+%As for claim (\ref{eq:zerovar}), note that as $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
+%\begin{align*}
+%\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*)\right. \\
+%&\left.- \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right]
+%= \Var\left[\gradApp{\vtheta^*}{N}\right].
+%\end{align*}
+%Note that it is essential that the trajectories used in the second and the third term are the same for the variance to vanish.
+%\end{proof}
 Previous results hold for both REINFORCE and G(PO)MDP.
@@
 
+\vspace{-0.05in}
 \section{Convergence Guarantees of SVRPG}\label{sec:conv}
+\vspace{-0.05in}
 In this section, we state the convergence guarantees for SVRPG with REINFORCE or G(PO)MDP gradient estimator.
@@
 
+\vspace{-0.05in}
 \section{Remarks on SVRPG}
+\vspace{-0.05in}
 The convergence guarantees presented in the previous section come with requirements on the meta-parameters (\ie $\alpha$ and $m$) that may be too conservative for practical applications.
@@
 
+\vspace{-0.05in}
 \subsection{Full Gradient Update}
+\vspace{-0.05in}
 As noted in Section \ref{sec:alg}, the update performed at the beginning of each epoch is equivalent to a full-gradient update.
@@
 
+\vspace{-0.05in}
 \subsection{Meta-Parameter Selection}\label{sec:stopping}
+\vspace{-0.05in}
 % In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
@@
 
+\vspace{-0.05in}
 \subsection{Normalized Importance Sampling}\label{sec:prac}
+\vspace{-0.05in}
 As mentioned in Section~\ref{sec:stopping}, importance weights are an additional source of variance.
@@
 
+\vspace{-0.05in}
 \section{Related Work}
+\vspace{-0.05in}
 Despite the considerable interest received in SL, variance-reduced gradient approaches have not attracted the RL community.
@@
 
+\vspace{-0.05in}
 \section{Experiments}\label{sec:exp}
+\vspace{-0.05in}
 In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer~\citep[\eg][]{duan2016benchmarking}.
@@
 \begin{figure*}[t]
-    \begin{minipage}[b]{0.3\textwidth}
-	\includegraphics[width=.95\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP_tex.pdf}
-	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% t-student confidence intervals.
-Results denoted with $\dagger$ are taken from \cite{duan2016benchmarking}.}
+    \begin{minipage}[t]{0.32\textwidth}
+	\includegraphics[width=.98\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP_tex.pdf}
+	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% t-student confidence intervals.}
 	\label{fig:cartpole}
     \end{minipage}
-    \begin{minipage}[b]{0.3\textwidth}
-	\includegraphics[width=0.95\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG_tex.pdf}
+    \hfill
+    \begin{minipage}[t]{0.32\textwidth}
+	\includegraphics[width=0.98\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG_tex.pdf}
 	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task, with 90\% confidence intervals.}
@@
     \hfill
-        \begin{minipage}[b]{.3\textwidth}
-	\includegraphics[width=0.95\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP_tex.pdf}
+    \begin{minipage}[t]{.32\textwidth}
+	\includegraphics[width=0.98\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP_tex.pdf}
 	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs G(PO)MDP in the Swimmer task, with 90\% confidence intervals}
@@
 
+\vspace{-0.05in}
 \section{Conclusion}
+\vspace{-0.05in}
 In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.

### 2018-02-09 svrpg.tex
updated app. D,E

--- 2018-02-09 svrpg.tex
@@
 \graphicspath{ {images/} }
-%\usepackage{subfigure}
+\usepackage{subfigure}
 \usepackage{booktabs} % for professional tables
@@
 
-In the recent past, there has been a surge of studies investigating variance reduction techniques for policy gradient methods.
+There are been a surge of work investigating variance reduction techniques for policy gradient methods.
 % The most widespread technique of reducing the variance preserving the unbiasedness of the gradient is through baselines~\citep[\eg][]{williams1992simple,weaver2001optimal,peters2008reinforcement,Thomas2017actionbaseline,wu2018variance}.
-The specific structure of the policy gradient allows incorporating a baseline (\ie a function $b :\mathcal{S}\times\mathcal{A} \to \realspace$) without affecting the unbiasedness of the gradient~\citep[\eg][]{williams1992simple,weaver2001optimal,peters2008reinforcement,Thomas2017actionbaseline,wu2018variance}.
-Although the baseline can be arbitrarily selected, literature often refers to the optimal baseline as the one minimizing the variance of the estimate.
+The specific structure of the policy gradient allows to incorporated a baseline (\ie a function $b :\mathcal{S}\times\mathcal{A} \to \realspace$) without affect the unbiasedness of the gradient~\citep[\eg][]{williams1992simple,weaver2001optimal,peters2008reinforcement,Thomas2017actionbaseline,wu2018variance}.
+Although the baseline can be arbitrarily selected, literature often refer to the optimal baseline as the one minimizing the variance of the estimate.
 Nevertheless, even the baseline needs to be estimated from data. This fact may partially reduce its effectiveness by introducing variance.
-Even if these approaches share the same goal as SVRG, they are substantially different.
-In particular, the proposed SVRPG does not make explicit use of the structure of the policy gradient framework, and it is independent of the underlying gradient estimate (\ie w/o baseline).
-This suggests that would be possible to integrate an ad-hoc SVRPG baseline to further reduce the variance of the estimate.
-Since this paper is about the applicability of SVRG technique to RL, we consider this topic as future work.
-Additionally, the experiments show that SVRPG has an advantage over G(PO)MPD even when the baseline is used (see the half-cheetah domain in Section~\ref{sec:exp}).
-
-Concerning importance weighting techniques, RL has made extensive use of them for off-policy problems~\citep[\eg][]{precup2000eligibility,thomas2015high}.
+Despite these approaches and SVRG has the same goal, they are substantially different.
+In particular, the proposed SVRPG does not make explicit use of the structure of the policy gradient framework and it is independent from the underline gradient estimate (\ie w/o baseline).
+This suggests that would be possible to integrate an ad-hoc SVRPG baseline in order to further reduce the variance of the estimate.
+But since this paper is about the applicability of SVRG technique to RL we consider this topic as future work.
+Additionally, the experiments show the advantage of using simple SVRPG against G(PO)MPD with baseline (see the \textcolor{blue}{half-cheetah} domain in Section~\ref{sec:exp}).
+
+Concerning importance weighting techniques, RL has made wide use of them for off-policy problems~\citep[\eg][]{precup2000eligibility,thomas2015high}.
 However, as mentioned before, SVRPG cannot be compared to such methods since it is in all respects an on-policy algorithm.
@@
 
-\begin{figure*}[t]
-    \begin{minipage}[b]{0.3\textwidth}
-	\includegraphics[width=.95\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP_tex.pdf}
+\begin{figure}[t]
+	\includegraphics[width=0.5\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP}
 	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% t-student confidence intervals.
@@
 	\label{fig:cartpole}
-    \end{minipage}
-    \begin{minipage}[b]{0.3\textwidth}
-	\includegraphics[width=0.95\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG_tex.pdf}
+\end{figure}
+
+\begin{figure}[t]
+	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
 	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task, with 90\% confidence intervals.}
 	\label{fig:swimmertwo}
-    \end{minipage}
-    \hfill
-        \begin{minipage}[b]{.3\textwidth}
-	\includegraphics[width=0.95\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP_tex.pdf}
+\end{figure}
+
+\begin{figure}[t]
+	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP}
 	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs G(PO)MDP in the Swimmer task, with 90\% confidence intervals}
 	\label{fig:swimmerone}
-    \end{minipage}
-\end{figure*}
+\end{figure}
 
@@
 Let us give a deeper insight on the two different learning rate schedules used by our algorithm.
-We report pseudo-code of the original ADAM in Algorithm \ref{A:adam}.
-As mentioned, we use two distinct instances of ADAM: one related to the snapshots, and one to the sub-iterations.
-In this way the ADAM associated to the snapshots will take into account only the history of gradient moments at the snapshots.
-The same reasoning holds for the sub-iterations.
-This, of course, allows our algorithm to take into account the different variabilities in a more accurate manner.
+We report pseudo-code of the original ADAM \cite{kingma2014adam} in Algorithm \ref{A:adam}.
+As mentioned, we use two distinct instances of ADAM to manage different sources of variance: one related to the snapshots, and one to the sub-iterations.
+In this way the ADAM associated to the snapshots takes into account only the history of gradient moments at the snapshots.
+By using Algorithm \ref{A:adam} as a subroutine $\textbf{ADAM}(g,\alpha,\beta)$, we can explicitly define our gradient updates:
+\begin{align*}
+\vtheta^{s+1}_1 &= \wt{\vtheta}^s + \textbf{ADAM}\left(\wh{\nabla}_N J(\wt{\vtheta}^s),\beta_1,\beta_2,2\alpha\right),\\
+\vtheta^{s+1}_{t+1} &= \textbf{ADAM}\Big( 
+%        v_t^{s+1}
+\blacktriangledown J(\vtheta^{s+1}_t),\beta_1,\beta_2,\alpha\Big)
+\text{ for $t=1,\dots,m-1$},
+\end{align*}
+where separate histories are kept for estimated first moments $\kappa_{FG},\kappa_{IS}$ and estimated second moments $\nu_{FG},\nu_{IS}$.
+The meta-parameters $\alpha,\beta_1,\beta_2$ are constant and set to default values or with minor manual tuning (see table \ref{table:metaparams}).
+Note that we double $\alpha$ for the snapshot ADAM, since we can rely on a larger number of trajectories ($N$ instead of $B$) to control the variance.
+
 \subsection{Baseline}
-For all batch gradient-based algorithms, we use the same time-varying feature encoding for the linear baseline:
-\begin{align*}
-\psi_{s,t}=concat(s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1)
-\end{align*}
-where $s$ is the state vector and $\odot$ is the element-wise product. In the case of our algorithm the model representing the baseline is fitted every snapshot so we have:\\
-\begin{align*}
-        \nabla J(\pi^T) =\frac{1}{N} \sum_{n=1}^{N}
-                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(y_k^n)\right) (\gamma^h \Reward(y_h^n)-b_{sn})
-          + \frac{1}{B} \sum_{n=1}^{B}
-                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\\ - \frac{1}{B} \sum_{n=1}^{B}
-                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\omega(z_{0:h}|\pi^B,\pi^T)
-\end{align*}
-Where $b_{sn}$ is the prediction given by the baseline at the snapshot, whereas $b_{si}$ is the prediction given by the baseline at a given subiteration. 
+The baseline used in the Half Cheetah experiment is the one used in \cite{duan2016benchmarking}. It is a linear state-value function estimator, or critic. 
+The (time-varying) feature encoding for the linear baseline is:
+\begin{align*}
+\vphi(s,t)=[s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1],
+\end{align*}
+where $s\in\mathbb{R}^d$ is the state vector and $\odot$ is the element-wise product. The baseline is then:
+\[
+b(s_t,a_t) = \mathbf{\lambda}^T\vphi(s_t,t).
+\]
+The baseline is fitted from scratch at each policy gradient iteration, with least squares, to match state-value function $V^{\pi}(s)$.
+When used with SVRPG, the critic parameter $\mathbf{\lambda}$ is updated only at the snapshot.
+
+%\begin{align*}
+%        \nabla J(\pi^T) =\frac{1}{N} \sum_{n=1}^{N}
+%                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(y_k^n)\right) (\gamma^h \Reward(y_h^n)-b_{sn})
+%          + \frac{1}{B} \sum_{n=1}^{B}
+%                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\\ - \frac{1}{B} \sum_{n=1}^{B}
+%                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\omega(z_{0:h}|\pi^B,\pi^T)
+%\end{align*}
+%Where $b_{sn}$ is the prediction given by the baseline at the snapshot, whereas $b_{si}$ is the prediction given by the baseline at a given subiteration.
+
 \section{Experimental Details}\label{app:exp}
-Used environments:
+We describe the RL tasks of Section \ref{sec:exp} in more detail:
 \begin{enumerate}
-	\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, cart velocity $\dot{x}$ and pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body.
+	\item \emph{Cart-Pole Balancing} : an inverted pendulum mounted on a cart must be kept standing by moving the cart backward or forward ;4-dimensional state space: cart position x, pole angle $\theta$, cart velocity $\dot{x}$ and pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body.
 Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$.
 The episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
-    \item \emph{Mujoco Swimmer}: 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
+    \item \emph{Mujoco Swimmer}: a snake-like robot immersed in a fluid must move forward; 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
 2-dimensional action space: the two momentums applied on actuated joints.
@@
 The episodes terminates when the number of time steps T is greater than 500.
-    \item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles.
+    \item \emph{Mujoco Half Cheetah}: a planar biped robot must move forward; 20-dimensional state space: 9 links and 6 actuated joints angles.
 6-dimensional action space: the 6 momentums applied on actuated joints.
@@
 
-Refer to \cite{duan2016benchmarking}  for more details about GPOMDP in Half Cheetah task.
-All the experiments' parameters are reported in the following table:
-
+All the parameters used in the experiments, including neural network architectures, are reported in the following table:
+
+\begin{table}[H]\caption{Parameters used in the experiments of Section \ref{sec:exp}.}\label{table:metaparams}
 	\centering
@@
 		\hline
-	State space dimension  & 4 & 13 & 20 \\
-	Control space dimension & 1 & 2 & 6 \\
-	NN Hidden layers & 8 & 32x32 & 100x50x25 \\
-	Adam learning rate SVRPO & $5\cdotp10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
-	Adam learning rate GPOMDP & $10^{-2}$ & $10^{-3}$ & $10^{-2}$ \\
-	Snapshot batch size & 100 & 100 & 100 \\
-	Sub-iterations batch size & 10 & 10 & 10 \\
-	GPOMDP batch size & 10 & 10 & 100 \\
+		NN hidden weights & 8 & 32x32 & 100x50x25 \\
+		NN activation & tanh & tanh & tanh \\
+		Adam $\alpha$ (SVRPG) & $5\cdotp10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
+		Adam $\alpha$ (GPOMDP) & $10^{-2}$ & $10^{-3}$ & $10^{-2}$ \\
+		Adam $\beta_1$ & 0.9 & 0.9 & 0.9 \\
+		Adam $\beta_2$ & 0.99 & 0.99 & 0.99 \\ 
+		Snapshot batch size $N$ (SVRPG) & 100 & 100 & 100 \\
+		Mini-batch size $B$ (SVRPG) & 10 & 10 & 10 \\
+		Batch size (GPOMDP) & 10 & 10 & 100 \\
 		Max number of sub-iterations & 50 & 20 & 20 \\
-	Episodes horizon& 100 & 500 & 500 \\
+		Task horizon& 100 & 500 & 500 \\
 		Baseline& No & No & Yes \\
-	Discount $(\gamma)$& 0.99 & 0.995 & 0.99 \\
+		Discount factor $\gamma$& 0.99 & 0.995 & 0.99 \\
 		Total number of trajectories& 10000 & 20000 & 50000 \\
@@
 	\end{tabular}
+\end{table}
+Where not specified, meta-parameters are shared among G(PO)MDP and SVRPG.
+Refer to \cite{duan2016benchmarking} for more details about G(PO)MDP (REINFORCE in the paper) on the Half Cheetah task.
 

### 2018-02-09 svrpg.tex
Updated

--- 2018-02-09 svrpg.tex
@@
 \begin{align*}
-\psi_{s,t}=concat(s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1)
+        \psi(s,t)=concat(s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1)
 \end{align*}
 where $s$ is the state vector and $\odot$ is the element-wise product.
+As a consequence, the baseline for both SVRPG and G(PO)MDP is defined as follows:
+\[
+        b(s,a) = \transpose{\boldsymbol{\lambda}} \psi(s,t)
+\]
 \section{Experimental Details}\label{app:exp}
@@
 Refer to \cite{duan2016benchmarking}  for more details about GPOMDP in Half Cheetah task.
-All the experiments' parameters are reported in the following table:
-
+All the experiments' parameters are reported in Table~\ref{T:exp_settings}.
+
+\begin{table}[b]
 \centering
+\caption{Settings of the experiments.}
+\label{T:exp_settings}
 \begin{tabular}{| l | c  c  c |}
@@
 \end{tabular}
+\end{table}
+In Figure~\ref{fig:cartpole_big},~\ref{fig:swimmertwo_big} and~\ref{fig:swimmerone_big} we have reported the same results presented in the main paper but on bigger figures.
+
+\begin{figure}[t]
+        \centering
+    \begin{minipage}[t]{0.48\textwidth}
+	\includegraphics[width=.98\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP_tex.pdf}
+	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% t-student confidence intervals.}
+	\label{fig:cartpole_big}
+    \end{minipage}
+    \hfill
+    \begin{minipage}[t]{0.48\textwidth}
+	\includegraphics[width=0.98\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG_tex.pdf}
+	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task, with 90\% confidence intervals.}
+	\label{fig:swimmertwo_big}
+    \end{minipage}
+\end{figure}
+\begin{figure}[t]
+        \centering
+	\includegraphics[width=0.47\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP_tex.pdf}
+	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs G(PO)MDP in the Swimmer task, with 90\% confidence intervals}
+	\label{fig:swimmerone_big}
+\end{figure}
 

### 2018-02-09 svrpg.tex
Updated

--- 2018-02-09 svrpg.tex
@@
 		\STATE \textbf{Input:} A gradient estimate $g_t$ and parameters $\beta_1$, $\beta_2$, $\epsilon$ and $\alpha$.
-        \STATE $m_t = \beta_1 m_{t-1} + (1 - \beta_1) g_t$
-        \STATE $v_t = \beta_2 v_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
-        \STATE $\hat{m}_t = \dfrac{m_t}{1 - \beta^t_1}$
-        \STATE $\hat{v}_t = \dfrac{v_t}{1 - \beta^t_2}$
-        \STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{v}_t} + \epsilon} \hat{m}_t$
+		\STATE $\kappa_t = \beta_1 \kappa_{t-1} + (1 - \beta_1) g_t$
+		\STATE $\nu_t = \beta_2 \nu_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
+		\STATE $\hat{\kappa}_t = \dfrac{\kappa_t}{1 - \beta^t_1}$
+		\STATE $\hat{\nu}_t = \dfrac{\nu_t}{1 - \beta^t_2}$
+		\STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{\nu}_t} + \epsilon} \hat{\kappa}_t$
 		\STATE \textbf{Return:} The increment $\Delta(g_t)$  of the parameters.
@@
 \end{algorithm}
-\subsection{Learning Rates}
-Let us give a deeper insight on the two different learning rate used by our algorithm.
-We have already said that we use two different ADAM one related to the snapshots and another one related to the sub-iterations.
-In this way tha ADAM associated to the snapshots will take into account only the snapshots history through the parameters $m_{t-1}$ and $v_{t-1}$.
-The same reasoning holds for the sub-iterations.
-This, of course, allows our algorithm to take into account the different variabilities in a more accurate manner.
+Let us give a deeper insight on the two different learning rate schedules used by our algorithm.
+We report pseudo-code of the original ADAM \cite{kingma2014adam} in Algorithm \ref{A:adam}.
+As mentioned, we use two distinct instances of ADAM to manage different sources of variance: one related to the snapshots, and one to the sub-iterations.
+In this way the ADAM associated to the snapshots takes into account only the history of gradient moments at the snapshots.
+By using Algorithm \ref{A:adam} as a subroutine $\textbf{ADAM}(g,\alpha,\beta)$, we can explicitly define our gradient updates:
+
+\begin{align*}
+\vtheta^{s+1}_1 &= \wt{\vtheta}^s + \textbf{ADAM}\left(\wh{\nabla}_N J(\wt{\vtheta}^s),\beta_1,\beta_2,2\alpha\right),\\
+\vtheta^{s+1}_{t+1} &= \textbf{ADAM}\Big( 
+%        v_t^{s+1}
+\blacktriangledown J(\vtheta^{s+1}_t),\beta_1,\beta_2,\alpha\Big)
+\text{ for $t=1,\dots,m-1$},
+\end{align*}
+
+where separate histories are kept for estimated first moments $\kappa_{FG},\kappa_{IS}$ and estimated second moments $\nu_{FG},\nu_{IS}$.
+The meta-parameters $\alpha,\beta_1,\beta_2$ are constant and set to default values or with minor manual tuning (see table \ref{table:metaparams}).
+Note that we double $\alpha$ for the snapshot ADAM, since we can rely on a larger number of trajectories ($N$ instead of $B$) to control the variance.
+
 \subsection{Baseline}
-For all batch gradient-based algorithms, we use the same time-varying feature encoding for the linear baseline:
-\begin{align*}
-\psi_{s,t}=concat(s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1)
-\end{align*}
-where $s$ is the state vector and $\odot$ is the element-wise product.
+The baseline used in the Half Cheetah experiment is the one used in \cite{duan2016benchmarking}. It is a linear state-value function estimator, or critic. 
+The (time-varying) feature encoding for the linear baseline is:
+\begin{align*}
+\vphi(s,t)=[s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1],
+\end{align*}
+where $s\in\mathbb{R}^d$ is the state vector and $\odot$ is the element-wise product. The baseline is then:
+\[
+b(s_t,a_t) = \mathbf{\lambda}^T\vphi(s_t,t).
+\]
+The baseline is fitted from scratch at each policy gradient iteration, with least squares, to match state-value function $V^{\pi}(s)$.
+When used with SVRPG, the critic parameter $\mathbf{\lambda}$ is updated only at the snapshot.
+
+%\begin{align*}
+%        \nabla J(\pi^T) =\frac{1}{N} \sum_{n=1}^{N}
+%                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(y_k^n)\right) (\gamma^h \Reward(y_h^n)-b_{sn})
+%          + \frac{1}{B} \sum_{n=1}^{B}
+%                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\\ - \frac{1}{B} \sum_{n=1}^{B}
+%                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\omega(z_{0:h}|\pi^B,\pi^T)
+%\end{align*}
+%Where $b_{sn}$ is the prediction given by the baseline at the snapshot, whereas $b_{si}$ is the prediction given by the baseline at a given subiteration.
+
 \section{Experimental Details}\label{app:exp}
-Used environments:
+We describe the RL tasks of Section \ref{sec:exp} in more detail:
 \begin{enumerate}
-	\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, cart velocity $\dot{x}$ and pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body.
+	\item \emph{Cart-Pole Balancing} : an inverted pendulum mounted on a cart must be kept standing by moving the cart backward or forward ;4-dimensional state space: cart position x, pole angle $\theta$, cart velocity $\dot{x}$ and pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body.
 Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$.
 The episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
-	\item \emph{Mujoco Swimmer}: 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
+    \item \emph{Mujoco Swimmer}: a snake-like robot immersed in a fluid must move forward; 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
 2-dimensional action space: the two momentums applied on actuated joints.
-The reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2] a^2$.
+The reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2]{a}^2$.
 The episodes terminates when the number of time steps T is greater than 500.
-	\item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles.
+    \item \emph{Mujoco Half Cheetah}: a planar biped robot must move forward; 20-dimensional state space: 9 links and 6 actuated joints angles.
 6-dimensional action space: the 6 momentums applied on actuated joints.
-The reward function is defined as $r(s, a) := v_x - 0.05\norm[2] a^2$.
+The reward function is defined as $r(s, a) := v_x - 0.05\norm[2]{a}^2$.
 The episodes terminates when the number of time steps T is greater than 500.

### 2018-02-09 svrpg.tex
Minor fixes.

--- 2018-02-09 svrpg.tex
@@
 For this reason, literature has focused on the introduction of baselines (\ie functions $b : \mathcal{S} \times \mathcal{A} \to \realspace$) aiming to reduce the variance~\citep[\eg][]{williams1992simple,Peters2008reinf,Thomas2017actionbaseline,wu2018variance}, see Appendix~\ref{A:gradient_estimators} for a formal definition of $b$.
-{\color{red}
-These baselines are usually designed in order to minimize the variance of the gradient estimate but even them need to be estimated from data, partially reducing they effectiveness.
+These baselines are usually designed to minimize the variance of the gradient estimate, but even them need to be estimated from data, partially reducing their effectiveness.
 On the other hand, there has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
 Although these techniques have been mainly derived for finite-sum problems, we will show in Section~\ref{sec:alg} how they can be used in RL.
-In particular, we will shown that the proposed SVRPG algorithm can take the best of both worlds (\ie SL and RL) since it can be plugged over a policy gradient estimate using baselines.
+In particular, we will show that the proposed SVRPG algorithm can take the best of both worlds (\ie SL and RL) since it can be plugged onto a policy gradient estimate using baselines.
 The next section has the aim to describe variance reduction techniques for finite-sum problems. In particular, we will present the SVRG algorithm that is at the core of this work.
-}
 
@@
 The overall structure is the same as Algorithm \ref{alg:svrg}, 
-but the snapshot gradient is not exact and the gradient estimate used between snapshots is corrected using importance weighting:\footnote{\color{red} Note that $g$ can be any unbiased estimator w/o baseline.
+but the snapshot gradient is not exact and the gradient estimate used between snapshots is corrected using importance weighting:\footnote{Note that $g$ can be any unbiased estimator, with or without baseline.
 The unbiasedness is required for theoretical results (\eg Appendix~\ref{A:gradient_estimators}).}
@@
 In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
-{\color{blue}Note that the randomized return value of Algorithm \ref{alg:svrpg} does not affect on-line learning at all, but will be used as a theoretical tool in the next section.}
+Note that the randomized return value of Algorithm \ref{alg:svrpg} does not affect online learning at all, but will be used as a theoretical tool in the next section.
 
@@
 \begin{align*}
-        \psi(s,t)=concat(s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1)
+\psi_{s,t}=concat(s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1)
 \end{align*}
 where $s$ is the state vector and $\odot$ is the element-wise product.
-As a consequence, the baseline for both SVRPG and G(PO)MDP is defined as follows:
-\[
-        b(s,a) = \transpose{\boldsymbol{\lambda}} \psi(s,t)
-\]
 \section{Experimental Details}\label{app:exp}
@@
 Refer to \cite{duan2016benchmarking}  for more details about GPOMDP in Half Cheetah task.
-All the experiments' parameters are reported in Table~\ref{T:exp_settings}.
-
-\begin{table}[b]
+All the experiments' parameters are reported in the following table:
+
 \centering
-\caption{Settings of the experiments.}
-\label{T:exp_settings}
 \begin{tabular}{| l | c  c  c |}
@@
 \end{tabular}
-\end{table}
-In Figure~\ref{fig:cartpole_big},~\ref{fig:swimmertwo_big} and~\ref{fig:swimmerone_big} we have reported the same results presented in the main paper but on bigger figures.
-
-\begin{figure}[t]
-        \centering
-    \begin{minipage}[t]{0.48\textwidth}
-	\includegraphics[width=.98\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP_tex.pdf}
-	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% t-student confidence intervals.}
-	\label{fig:cartpole_big}
-    \end{minipage}
-    \hfill
-    \begin{minipage}[t]{0.48\textwidth}
-	\includegraphics[width=0.98\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG_tex.pdf}
-	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task, with 90\% confidence intervals.}
-	\label{fig:swimmertwo_big}
-    \end{minipage}
-\end{figure}
-\begin{figure}[t]
-        \centering
-	\includegraphics[width=0.47\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP_tex.pdf}
-	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs G(PO)MDP in the Swimmer task, with 90\% confidence intervals}
-	\label{fig:swimmerone_big}
-\end{figure}
 

### 2018-02-09 svrpg.tex
stressed more the convergence rate

@@
         The result is SVRPG, a stochastic variance-reduced policy gradient algorithm that leverages on importance weight to preserve the unbiasedness of the gradient estimate.
-        We further analyse its convergence guarantees under standard assumptions on the MDP.
+        {\color{blue}Under standard assumptions on the MDP, we provide convergence guarantees for SVRPG with a convergence rate that is linear under increasing batch sizes.}
         Finally, we suggest practical variants of SVRPG, and we empirically evaluate them on continuous MDPs.
@@
 After providing background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} propose SVRPG, a variant of SVRG for the policy gradient framework, addressing all the difficulties mentioned above.
-In Section \ref{sec:conv} we provide convergence guarantees for our algorithm which also offers insights on the challenges of variance reduction in RL.
+{\color{blue} In Section \ref{sec:conv} we provide convergence guarantees for our algorithm, and we show a convergence rate that has an $O(\nicefrac{1}{T})$ dependence on the number $T$ of iterations}.
 In Section \ref{sec:stopping} we suggest how to set the meta-parameters of SVRPG, while in Section \ref{sec:prac} we discuss some practical variants of the algorithm.
@@
 	\includegraphics[width=.98\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP_tex.pdf}
-	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% t-student confidence intervals.}
+	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% confidence intervals.}
 	\label{fig:cartpole}

### 2018-02-09 svrpg.tex
Minor fixes.

--- 2018-02-09 svrpg.tex
@@
 
-Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^d\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$.
+Policy gradient methods restrict the search for the best performing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^d\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$.
 For sake of brevity, we will denote the performance of a parametric policy with $J(\vtheta)$ and the probability of a trajectory $\tau$ with $p(\tau|\vtheta)$ (in some occasions, $p(\tau|\vtheta)$ will be replaced by $p_{\vtheta}(\tau)$ for the sake of readability).
@@
 	\begin{restatable}[On policy derivatives]{assumption}{boundedscore}\label{ass:bounded_score}
-		For each state-action pair $(s,a)$, any value of $\vtheta$ and all parameter components $i,j$ there are positive constants $G,F<\infty$ such that:
+		For each state-action pair $(s,a)$, any value of $\vtheta$, and all parameter components $i,j$ there exist constants $0 \leq G,F<\infty$ such that:
 \[
@@
 	\begin{restatable}[On the variance of importance weights]{assumption}{varweights}\label{ass:M2}
-		There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm~\ref{alg:svrpg} and for each possible trajectory,
+		There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm~\ref{alg:svrpg} and for each trajectory,
 		\[
@@
 \end{align*}
-In the following, we will always assume this practical variant is employed.
+In the following, we will always use this practical variant.
 
@@
 High values of $m$ may move policy $\pi_{\vtheta_t}$ far away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing large variance in the importance weights.
-We propose to adapt the two meta-parameters jointly.
+We will jointly set the two meta-parameters.
 % the farther the distance between the snapshot policy and the current one, the higher the variance of the importance weights.
@@
 By doing so, we decouple the contribution of the variance due to the approximate FG from the one introduced by the sub-iterations.
-Note that these two terms have different orders of magnitude since are estimated with a different number of trajectories ($B \ll N$) and w/o importance weights.
+Note that these two terms have different orders of magnitude since are estimated with a different number of trajectories ($B \ll N$) and the estimator in the snapshot does not require importance weights.
 The use of two ADAM estimators allows to capture and exploit this property.

### 2018-02-09 svrpg.tex
Minor fixes.

--- 2018-02-09 svrpg.tex
@@
 \g@addto@macro\normalsize{%
- \setlength\abovedisplayskip{6pt}
- \setlength\belowdisplayskip{6pt}
+%  \setlength\abovedisplayskip{6pt}
+%  \setlength\belowdisplayskip{6pt}
 }
@@
         The result is SVRPG, a stochastic variance-reduced policy gradient algorithm that leverages on importance weight to preserve the unbiasedness of the gradient estimate.
-        We further analyse its convergence guarantees under standard assumptions on the MDP.
+        Under standard assumptions on the MDP, we provide convergence guarantees for SVRPG with a convergence rate that is linear under increasing batch sizes.
         Finally, we suggest practical variants of SVRPG, and we empirically evaluate them on continuous MDPs.
@@
 After providing background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} propose SVRPG, a variant of SVRG for the policy gradient framework, addressing all the difficulties mentioned above.
-In Section \ref{sec:conv} we provide convergence guarantees for our algorithm which also offers insights on the challenges of variance reduction in RL.
+In Section \ref{sec:conv} we provide convergence guarantees for our algorithm, and we show a convergence rate that has an $O(\nicefrac{1}{T})$ dependence on the number $T$ of iterations.
 In Section \ref{sec:stopping} we suggest how to set the meta-parameters of SVRPG, while in Section \ref{sec:prac} we discuss some practical variants of the algorithm.
@@
 The less the number of trajectories used to update the policy, the better.
-This weighted approach comes from the fact that, as mentioned above, there is no advantage in having $m$ too small.
-Thus, we aim to facilitate sub-iterations rather than snapshots.
+Including the batch sizes in the stopping condition allows us to optimize the trade-off between the quality of the updates and the cost of performing them.
+% This weighted approach comes from the fact that, as mentioned above, there is no advantage in having $m$ too small.
+% Thus, we aim to facilitate sub-iterations rather than snapshots.
 
@@
 Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for online learning since it is related to an average regret~\citep[\eg][]{duan2016benchmarking}.
-This metric corresponds to the area below the curves in the figures.
+This metric is related to the area below the curves.
 Note that the convergence result in Theorem \ref{theo:convergence} is precisely about the average return.
@@
 \textbf{Preliminary results on actor-critic.} %As mentioned, 
-Another variance-reduction technique in policy gradient is to employ baselines or \textit{critics}.
+Another variance-reduction technique in policy gradient consists of using baselines or \textit{critics}.
 This tool is orthogonal to the methods described in this paper, and the theoretical results of Section \ref{sec:conv} are general in this sense.
 In the experiments described so far, we compared against the so-called \textit{actor-only} G(PO)MDP, \ie without the baseline.
-To move towards a more general understanding of the problem of variance in policy gradient, we also test SVRPG in an \textit{actor-critic} scenario.
+To move towards a more general understanding of the variance issue in policy gradient, we also test SVRPG in an \textit{actor-critic} scenario.
 To do so, we consider the more challenging MuJoCo~\citep{todorov2012mujoco} Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer.

### 2018-02-09 svrpg.tex
Fix of typos

@@
 Although these techniques have been mainly derived for finite-sum problems, we will show in Section~\ref{sec:alg} how they can be used in RL.
-In particular, we will show that the proposed SVRPG algorithm can take the best of both worlds (\ie SL and RL) since it can be plugged onto a policy gradient estimate using baselines.
+In particular, we will show that the proposed SVRPG algorithm can take the best of both worlds (\ie SL and RL) since it can be plugged into a policy gradient estimate using baselines.
 The next section has the aim to describe variance reduction techniques for finite-sum problems. In particular, we will present the SVRG algorithm that is at the core of this work.
@@
 Note that $t=0$ corresponds to a FG step (\ie $\blacktriangledown f(\vtheta_0) = \nabla f(\wt{\vtheta})$) since $\vtheta_0 := \wt{\vtheta}$.
-The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, achieving a linear convergence rate without resorting to plain full gradient.
+The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, achieving a linear convergence rate without resorting to a plain full gradient.
 
@@
 % In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence.
-I would say that SG is not a choice by referring to papers that uses batches}
+I would say that SG is not a choice by referring to papers that use batches}
 % We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
@@
 G(PO)MDP can be seen as a more efficient implementation of the REINFORCE algorithm. 
-In fact the latter does not perform an optimal credit assignment, since it ignores that the reward at time $t$ does not depend on the action performed after time $t$.
+In fact, the latter does not perform an optimal credit assignment since it ignores that the reward at time $t$ does not depend on the action performed after time $t$.
 G(PO)MDP overcomes this issue taking into account the causality of rewards in the REINFORCE definition of policy gradient.
@@
 
-Next we have:
+Next, we have:
 \begingroup
@@
 The meta-parameters $\alpha,\beta_1,\beta_2$ are constant and set to default values or with minor manual tuning (see table \ref{table:metaparams}).
-Note that we double $\alpha$ for the snapshot ADAM, since we can rely on a larger number of trajectories ($N$ instead of $B$) to control the variance.
+Note that we double $\alpha$ for the snapshot ADAM since we can rely on a larger number of trajectories ($N$ instead of $B$) to control the variance.
 
@@
 Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$.
-The episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
+The episodes terminate when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
     \item \emph{Mujoco Swimmer}: a snake-like robot immersed in a fluid must move forward; 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles.
@@
 The reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2]{a}^2$.
-The episodes terminates when the number of time steps T is greater than 500.
+The episodes terminate when the number of time steps T is greater than 500.
     \item \emph{Mujoco Half Cheetah}: a planar biped robot must move forward; 20-dimensional state space: 9 links and 6 actuated joints angles.
@@
 The reward function is defined as $r(s, a) := v_x - 0.05\norm[2]{a}^2$.
-The episodes terminates when the number of time steps T is greater than 500.
+The episodes terminate when the number of time steps T is greater than 500.
 \end{enumerate}

### 2018-02-09 svrpg.tex
mentioned natural grad in future work

--- 2018-02-09 svrpg.tex
@@
 Despite that, we believe that it will be possible to derive a baseline designed explicitly for SVRPG to exploit the RL structure and the SVRG idea jointly.
-Future work should also focus on making batch sizes $N$ and $B$ adaptive, as suggested in \cite{papini2017adaptive}.
+{\color{blue}Another possible improvement would be to employ the natural gradient \cite{kakade2002natural} to better control the effects of parameter updates to the variance of importance weights.} Future work should also focus on making batch sizes $N$ and $B$ adaptive, as suggested in \cite{papini2017adaptive}.
 

### 2018-02-09 svrpg.tex
Updated

--- 2018-02-09 svrpg.tex
@@
 These baselines are usually designed to minimize the variance of the gradient estimate, but even them need to be estimated from data, partially reducing their effectiveness.
-On the other hand, there has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
+On the other hand, there has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning (SL).
 Although these techniques have been mainly derived for finite-sum problems, we will show in Section~\ref{sec:alg} how they can be used in RL.
@@
 A common requirement is that $z$ must be smooth and concave in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
-Under this hypothesis, full gradient (FG) ascent ~\citep{cauchy1847methode} with a constant step size achieves a linear convergence rate \cite{nesterov2013introductory}.
+Under this hypothesis, full gradient (FG) ascent~\citep{cauchy1847methode} with a constant step size achieves a linear convergence rate in the number $T$ of iterations (\ie parameter updates)~\citep{nesterov2013introductory}.
 However, each iteration requires $N$ gradient computations, which can be too expensive for large values of $N$.
@@
 
-More recently, some extensions of variance reduction algorithms to the non-concave objectives have been proposed~\citep[\eg][]{reddi2016stochastic,allen2016variance,reddi2016stochastic,reddi2016fast}.
+More recently, some extensions of variance reduction algorithms to the non-concave objectives have been proposed~\citep[\eg][]{allen2016variance,reddi2016stochastic,reddi2016fast}.
 In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[2]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[2]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
@@
 Even if these approaches share the same goal as SVRG, they are substantially different.
-In particular, the proposed SVRPG does not make explicit use of the structure of the policy gradient framework, and it is independent of the underlying gradient estimate (\ie w/o baseline).
+In particular, the proposed SVRPG does not make explicit use of the structure of the policy gradient framework, and it is independent of the underlying gradient estimate (\ie with or without baseline).
 This suggests that would be possible to integrate an ad-hoc SVRPG baseline to further reduce the variance of the estimate.

### 2018-02-09 svrpg.tex
Fit in 8 pages

--- 2018-02-09 svrpg.tex
@@
 	\includegraphics[width=.98\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP_tex.pdf}
+	\vspace{-0.1in}
 	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% confidence intervals.}
@@
 	\includegraphics[width=0.98\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG_tex.pdf}
+	\vspace{-0.1in}
 	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task, with 90\% confidence intervals.}
@@
 	\includegraphics[width=0.98\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP_tex.pdf}
+	\vspace{-0.1in}
 	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs G(PO)MDP in the Swimmer task, with 90\% confidence intervals}
@@
     \end{minipage}
+    \vspace{-0.15in}
 \end{figure*}
@@
 	\end{tabular}
+    \vspace{-0.05in}
 \end{table*}
@@
 Despite that, we believe that it will be possible to derive a baseline designed explicitly for SVRPG to exploit the RL structure and the SVRG idea jointly.
-Future work should also focus on making batch sizes $N$ and $B$ adaptive, as suggested in \cite{papini2017adaptive}.
+Another possible improvement would be to employ the natural gradient~\cite{kakade2002natural} to better control the effects of parameter updates on the variance of importance weights.
+Future work should also focus on making batch sizes $N$ and $B$ adaptive, as suggested in~\cite{papini2017adaptive}.
 

### 2018-02-09 svrpg.tex
mentioned videos in appendix, restored subsection structure of app. D

--- 2018-02-09 svrpg.tex
@@
 These baselines are usually designed to minimize the variance of the gradient estimate, but even them need to be estimated from data, partially reducing their effectiveness.
-On the other hand, there has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning (SL).
+On the other hand, there has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
 Although these techniques have been mainly derived for finite-sum problems, we will show in Section~\ref{sec:alg} how they can be used in RL.
@@
 A common requirement is that $z$ must be smooth and concave in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
-Under this hypothesis, full gradient (FG) ascent~\citep{cauchy1847methode} with a constant step size achieves a linear convergence rate in the number $T$ of iterations (\ie parameter updates)~\citep{nesterov2013introductory}.
+Under this hypothesis, full gradient (FG) ascent ~\citep{cauchy1847methode} with a constant step size achieves a linear convergence rate \cite{nesterov2013introductory}.
 However, each iteration requires $N$ gradient computations, which can be too expensive for large values of $N$.
@@
 
-More recently, some extensions of variance reduction algorithms to the non-concave objectives have been proposed~\citep[\eg][]{allen2016variance,reddi2016stochastic,reddi2016fast}.
+More recently, some extensions of variance reduction algorithms to the non-concave objectives have been proposed~\citep[\eg][]{reddi2016stochastic,allen2016variance,reddi2016stochastic,reddi2016fast}.
 In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[2]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[2]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
@@
 Even if these approaches share the same goal as SVRG, they are substantially different.
-In particular, the proposed SVRPG does not make explicit use of the structure of the policy gradient framework, and it is independent of the underlying gradient estimate (\ie with or without baseline).
+In particular, the proposed SVRPG does not make explicit use of the structure of the policy gradient framework, and it is independent of the underlying gradient estimate (\ie w/o baseline).
 This suggests that would be possible to integrate an ad-hoc SVRPG baseline to further reduce the variance of the estimate.
@@
 
-\section{Practical SVRPG versions}\label{app:practicalsvrpg}
+\section{Practical SVRPG Versions}\label{app:practicalsvrpg}
+We provide more details on the practical variants of SVRPG.
+
+\subsection{Adpative Step Size}
 \begin{algorithm}[h]
@@
 
+\paragraph{Videos:} The supplementary materials include videos showing the behavior of the final SVRPG policy on the three considered tasks.
+
 \end{document}

### 2018-02-09 svrpg.tex
typo

--- 2018-02-09 svrpg.tex
@@
         However, their adaptation to policy gradient is not straightforward and needs to account for I) a non-concave objective function; II) approximations in the full gradient computation; and III) a non-stationary sampling process.
-        The result is SVRPG, a stochastic variance-reduced policy gradient algorithm that leverages on importance weight to preserve the unbiasedness of the gradient estimate.
+        The result is SVRPG, a stochastic variance-reduced policy gradient algorithm that leverages on importance weights to preserve the unbiasedness of the gradient estimate.
         Under standard assumptions on the MDP, we provide convergence guarantees for SVRPG with a convergence rate that is linear under increasing batch sizes.
@@
 
+\paragraph{Videos:} The supplementary materials include videos showing the behavior of the final SVRPG policies on the three considered tasks.
+
 \end{document}

### 2018-02-09 svrpg.tex
typo
fix in appendices

@@
         However, their adaptation to policy gradient is not straightforward and needs to account for I) a non-concave objective function; II) approximations in the full gradient computation; and III) a non-stationary sampling process.
-        The result is SVRPG, a stochastic variance-reduced policy gradient algorithm that leverages on importance weights to preserve the unbiasedness of the gradient estimate.
+        The result is SVRPG, a stochastic variance-reduced policy gradient algorithm that leverages on importance weight to preserve the unbiasedness of the gradient estimate.
         Under standard assumptions on the MDP, we provide convergence guarantees for SVRPG with a convergence rate that is linear under increasing batch sizes.
@@
 \begin{align*}
-\vtheta^{s+1}_1 &= \wt{\vtheta}^s + \textbf{ADAM}\left(\wh{\nabla}_N J(\wt{\vtheta}^s),\beta_1,\beta_2,2\alpha\right),\\
+\vtheta^{s+1}_1 &= \wt{\vtheta}^s + \textbf{ADAM}\left(\wh{\nabla}_N J(\wt{\vtheta}^s),\beta_1,\beta_2,\alpha\right),\\
 \vtheta^{s+1}_{t+1} &= \textbf{ADAM}\Big( 
 %        v_t^{s+1}
-\blacktriangledown J(\vtheta^{s+1}_t),\beta_1,\beta_2,\alpha\Big)
+\blacktriangledown J(\vtheta^{s+1}_t),\beta_1,\beta_2,\frac{\alpha}{2}\Big)
 \text{ for $t=1,\dots,m-1$},
@@
 The meta-parameters $\alpha,\beta_1,\beta_2$ are constant and set to default values or with minor manual tuning (see table \ref{table:metaparams}).
-Note that we double $\alpha$ for the snapshot ADAM since we can rely on a larger number of trajectories ($N$ instead of $B$) to control the variance.
+Note that we double the sub-iterations' learning rate for the snapshot ADAM since we can rely on a larger number of trajectories ($N$ instead of $B$) to control the variance.
 

### 2018-02-09 svrpg.tex

@@
 
-\paragraph{Videos:} The supplementary materials include videos showing the behavior of the final SVRPG policies on the three considered tasks.
+\paragraph{Videos:} The supplementary materials include videos showing the behavior of the final SVRPG agents on the three considered tasks.
 

### 2018-04-12 rebuttals.txt
rebuttals

We would like to thank all the reviewers for their feedback and in-depth comments.

Reviewer 1:

The heuristics proposed in the paper were not developed with the main purpose of beating the baseline, but to make the algorithm work (learn) in a stable way. The intrinsic difficulties of adapting SVRG to the RL framework have as a consequence a certain sensitivity to meta-parameters. Our contribution should be intended as two-fold: a theoretical part supporting the usage of SVRG in RL, and a more empirical part highlighting the practical difficulties and proposing a way to tackle them. We think the community would benefit from a picture of both the opportunities and the risks of this technique. In the final version, we will make this more clear.

The proof to Lemma 3.1. was meant to be in the appendix and is missing by mistake. We will definitively include it. The variance property is true at convergence because the O(1/B) terms cancel out as both the parameter and the snapshot parameter approach the optimum. This is a nice property of SVRG mentioned also in Section 2 of the original SVRG paper.

The performance of GPOMDP on Swimmer is lower than the one reported in Duan et al. because we do not employ the linear critic used in that paper. Since critic and SVRG are both ways to reduce the variance of the gradient estimate, and a combination of the two is non-trivial (how do you critique snapshot policies?) we turned the actor off to measure the net advantage of SVRG. Later, in the Half-Cheetah experiment, we include the actor in both GPOMDP and SVRPG.

We reported Duan et al. values only for the Half-Cheetah experiment because it is the only case in which we used the actor. We recognize that this is confusing. In fact, we now have reproduced the GPOMDP/Half-Cheetah result and we will simply report our results, mentioning that the ones on Half-Cheetah are coherent with what is reported in Duan et al., while the ones on Cart-Pole and Swimmer cannot be compared since we did not use the critic.

The version of Cart-Pole we use is the same of Duan et al, and this should be mentioned, although the task is already described in Appendix E. The lower scores are due to the fact that we do not include a critic in GPOMDP.

As a first attempt to bring the SVRG technique to the policy optimization framework, we do not claim SVRPG has the maturity of algorithms like TRPO. Recent policy optimization algorithms employ a series of tools, e.g. natural gradients and exploration bonuses, that are outside the scope of this paper but represent promising future-work material.

We thank the reviewer for pointing out typos.

Reviewer 2:

1), 2) The impact of Theorem 4, besides the theoretical guarantees, is to show the interplay of the meta-parameters more than prescript specific values. In nonconvex optimization, it is often the case that the meta-parameters prescribed by the theory are prohibitively conservative to be used in applications. For instance, in Reddi et al. the authors understand from their theoretical analysis that a constant step-size should be used, but empirically search for the specific value that yield the best practical performance. Even in the original SVRG paper, common learning rate scheduling policies are employed in the experiments. In RL, an example is the learning rate suggested in [Kakade, Sham, and John Langford. "Approximately optimal approximate reinforcement learning." ICML. Vol. 2. 2002.], which guarantees monotonic improvement but is prohibitively small for practical usage. However, the ideas embedded in that theoretical analysis ultimately lead to TRPO, which achieves great performance by relaxing several theoretical constraints. In our work, we both provide theoretical guarantees and propose a practical way to adapt meta-parameters. The statement of Theorem 4 (and some subtelties of its proof) heavily guided the choice of this heuristics, and we will make this more clear in the final version.
3) \theta^* is a global optimizer and we should state it explicitly.
4) \theta_m^0, \tilde\theta^0 and \theta_0^1 are all the same. Superscripts denote epochs and subscripts inner iterations, as should be clear from the pseudocode. We will check for notation inconsistencies in the text.
5) Plain PG is an instance of Stochastic Gradient Ascent, which indeed has a 1/\sqrt{T} rate.
6) We consider the horizon H to be part of the problem, but we should mention that problems with long horizons can be more difficult since we employ importance weighting.

On experiments:
1) The comparison is fair, figures have the number of trajectories on the horizontal axis, and the extra trajectories collected by  SVRPG at the snapshot are indeed counted. Our algorithm manages to be faster anyway because the quality of the updates is higher.
2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work since adapting batch sizes in the policy gradient framework is a  difficult problem that have been only slightly explored in the RL literature so far (cfr. Papini et al., "Adaptive batch size for safe policy gradients." NIPS 2017).
3) This is because we used the rllab variant of Cart-Pole (Duan et al., 2016), where the agent collects a reward of ~10 at each time-step, as described in Appendix E.
4) For SVRPG, \alpha is selected with the Adam variant described in Section 5.2. For GPOMDP, we use Adam. All meta-parameters are reported in Appendix E and have be chosen to make the comparison fair.

Reviewer 3:

We thank the reviewer for the positive feedback and the suggestions on related works, which will be useful for future developments of our method. [Bietti et al. 2016] seems particularly relevant and we should mention it in the final version.

### 2018-04-13 rebuttals.txt
Updated

@@
 The heuristics proposed in the paper were not developed with the main purpose of beating the baseline, but to make the algorithm work (learn) in a stable way.
-The intrinsic difficulties of adapting SVRG to the RL framework have as a consequence a certain sensitivity to meta-parameters.
+The intrinsic difficulties of adapting SVRG to the RL framework have, as a consequence, a certain sensitivity to the meta-parameters.
 Our contribution should be intended as two-fold: a theoretical part supporting the usage of SVRG in RL, and a more empirical part highlighting the practical difficulties and proposing a way to tackle them.
@@
 
-The proof to Lemma 3.1.
-was meant to be in the appendix and is missing by mistake.
+The proof to Lemma 3.1 was meant to be in the appendix and is missing by mistake.
 We will definitively include it.
@@
 This is a nice property of SVRG mentioned also in Section 2 of the original SVRG paper.
+(METTIAMO LINK ANONIMO?)
 
@@
 For instance, in Reddi et al.
-the authors understand from their theoretical analysis that a constant step-size should be used, but empirically search for the specific value that yield the best practical performance.
-Even in the original SVRG paper, common learning rate scheduling policies are employed in the experiments.
+the authors understand from their theoretical analysis that a constant step-size should be used, but they empirically search for the specific value that yields the best practical performance.
+Even in the original SVRG paper, common learning rate scheduling policies (ANNEALING SCHEMA??) are employed in the experiments.
 In RL, an example is the learning rate suggested in [Kakade, Sham, and John Langford.
@@
 In our work, we both provide theoretical guarantees and propose a practical way to adapt meta-parameters.
-The statement of Theorem 4 (and some subtelties of its proof) heavily guided the choice of this heuristics, and we will make this more clear in the final version.
+The statement of Theorem 4 (and some subtleties of its proof) heavily guided the choice of this heuristics, and we will make this more clear in the final version.
 3) \theta^* is a global optimizer and we should state it explicitly.
@@
 Our algorithm manages to be faster anyway because the quality of the updates is higher.
-2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work since adapting batch sizes in the policy gradient framework is a  difficult problem that have been only slightly explored in the RL literature so far (cfr.
+2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work since adapting batch sizes in the policy gradient framework is a  difficult problem that has been only slightly explored in the RL literature so far (cfr.
 Papini et al., "Adaptive batch size for safe policy gradients." NIPS 2017).
 3) This is because we used the rllab variant of Cart-Pole (Duan et al., 2016), where the agent collects a reward of ~10 at each time-step, as described in Appendix E.
-4) For SVRPG, \alpha is selected with the Adam variant described in Section 5.2. For GPOMDP, we use Adam. All meta-parameters are reported in Appendix E and have be chosen to make the comparison fair.
+4) For SVRPG, \alpha is selected with the Adam variant described in Section 5.2.
+For GPOMDP, we use Adam.
+All meta-parameters are reported in Appendix E and have been chosen to make the comparison fair.
 
@@
 We thank the reviewer for the positive feedback and the suggestions on related works, which will be useful for future developments of our method.
-[Bietti et al.
-2016] seems particularly relevant and we should mention it in the final version.
+We will mention the suggested papers/directions in the final version, especially [Bietti et al.
+2016] seems particularly relevant.

### 2018-04-13 rebuttals.txt
updated rebuttals

--- 2018-04-12 rebuttals.txt
@@
 because we do not employ the linear critic used in that paper.
-Since critic and SVRG are both ways to reduce the variance of the gradient estimate, and a combination of the two is non-trivial (how do you critique snapshot policies?) we turned the actor off to measure the net advantage of SVRG.
-Later, in the Half-Cheetah experiment, we include the actor in both GPOMDP and SVRPG.
+Since critic and SVRG are both ways to reduce the variance of the gradient estimate, and a combination of the two is non-trivial (how do you critique snapshot policies?) we turned the critic off to measure the net advantage of SVRG.
+Later, in the Half-Cheetah experiment, we include the critic in both GPOMDP and SVRPG.
 
 We reported Duan et al.
-values only for the Half-Cheetah experiment because it is the only case in which we used the actor.
+values only for the Half-Cheetah experiment because it is the only case in which we used the critic (in order to evaluate the combination of SVRG and critic).
 We recognize that this is confusing.
@@
 The version of Cart-Pole we use is the same of Duan et al, and this should be mentioned, although the task is already described in Appendix E.
-The lower scores are due to the fact that we do not include a critic in GPOMDP.
+The lower scores are due to the fact that we do not include a critic in GPOMDP, to make the comparison with SVRPG (that does not use the critic on Cart-Pole) fair.
 
@@
 1) The comparison is fair, figures have the number of trajectories on the horizontal axis, and the extra trajectories collected by  SVRPG at the snapshot are indeed counted.
-Our algorithm manages to be faster anyway because the quality of the updates is higher.
+Our algorithm updates the policy less often, but manages to be faster anyway because the quality of the updates is higher.
 2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work since adapting batch sizes in the policy gradient framework is a  difficult problem that have been only slightly explored in the RL literature so far (cfr.

### 2018-04-13 rebuttals.txt

@@
 3) This is because we used the rllab variant of Cart-Pole (Duan et al., 2016), where the agent collects a reward of ~10 at each time-step, as described in Appendix E.
-4) For SVRPG, \alpha is selected with the Adam variant described in Section 5.2. For GPOMDP, we use Adam. All meta-parameters are reported in Appendix E and have be chosen to make the comparison fair.
+4) For SVRPG, \alpha is selected with the Adam variant described in Section 5.2.
+For GPOMDP, we use Adam.
+All meta-parameters are reported in Appendix E and have been chosen to make the comparison fair.
 

### 2018-04-14 rebuttals.txt
updated rebuttals

@@
 The intrinsic difficulties of adapting SVRG to the RL framework have, as a consequence, a certain sensitivity to the meta-parameters.
-Our contribution should be intended as two-fold: a theoretical part supporting the usage of SVRG in RL, and a more empirical part highlighting the practical difficulties and proposing a way to tackle them.
-We think the community would benefit from a picture of both the opportunities and the risks of this technique.
-In the final version, we will make this more clear.
+Our contribution should be intended as two-fold: a theoretical part supporting the usage of SVRG in RL, and a more empirical part highlighting the additional challenges of variance reduction in RL, where we also show how these difficulties can be overcome with proper heuristics.
+We think the community would benefit from a picture of both the opportunities and the risks of applying SVRG to policy gradient learning.
 
@@
 The variance property is true at convergence because the O(1/B) terms cancel out as both the parameter and the snapshot parameter approach the optimum.
-This is a nice property of SVRG mentioned also in Section 2 of the original SVRG paper.
-(METTIAMO LINK ANONIMO?)
+More formally: "As \theta -> \theta^*, also \tilde\theta -> \theta^*; hence, by continuity of J: Var(SVRG) -> Var(\hat\nablaJ) + 1/B*Var[g(\theta^*) - 1*g(\theta^*)] = Var(\hat\nabla)".
+The importance weight is one because the two parameters are the same.
+See also a similar proof in Section 2 of the original SVRG paper.
 
@@
 because we do not employ the linear critic used in that paper.
-Since critic and SVRG are both ways to reduce the variance of the gradient estimate, and a combination of the two is non-trivial (how do you critique snapshot policies?) we turned the critic off to measure the net advantage of SVRG.
-Later, in the Half-Cheetah experiment, we include the critic in both GPOMDP and SVRPG.
+Since critic and SVRG are both ways to reduce the variance of the gradient estimate, and a combination of the two is non-trivial (how do you critique snapshot policies?) we turned the critic off in both GPOMDP and SVRPG to measure the net advantage of SVRG.
+Later, in the Half-Cheetah experiment, we include the critic in both algorithms.
 
 We reported Duan et al.
-values only for the Half-Cheetah experiment because it is the only case in which we used the critic (in order to evaluate the combination of SVRG and critic).
-We recognize that this is confusing.
-In fact, we now have reproduced the GPOMDP/Half-Cheetah result and we will simply report our results, mentioning that the ones on Half-Cheetah are coherent with what is reported in Duan et al., while the ones on Cart-Pole and Swimmer cannot be compared since we did not use the critic.
+values only for the Half-Cheetah experiment because it is the only case in which we used the critic.
+This represent a further step: evaluating the combination of SVRG and critic as variance reduction techniques.
+We recognize that the current presentation is confusing.
+In fact, we now have reproduced the GPOMDP/Half-Cheetah result and we will simply report our results in the final version.
+We will mention that experiments on Half-Cheetah are coherent with Duan et al., while the ones on Cart-Pole and Swimmer cannot be compared since we did not use the critic.
 
-The version of Cart-Pole we use is the same of Duan et al, and this should be mentioned, although the task is already described in Appendix E.
-The lower scores are due to the fact that we do not include a critic in GPOMDP, to make the comparison with SVRPG (that does not use the critic on Cart-Pole) fair.
+The version of Cart-Pole we use is the same of Duan et al, and this should be mentioned, although the task is thoroughly described in Appendix E.
+As for Swimmer, the lower scores are due to the fact that we do not include a critic in either GPOMDP or SVRPG.
 
@@
 Reviewer 2:
+We see that meta-parameters are a delicate aspect of our algorithm.
+The pragmatic role of Theorem 4 is to show the interplay of these parameters more than prescript specific values.
+In our work, we both provide general theoretical guarantees and propose a practical way to select meta-parameters.
+The statement of Theorem 4 (and some subtleties of its proof) heavily guided the choice of this heuristics, and we will make this more clear in the final version.
+More specifically:
 
-1), 2) The impact of Theorem 4, besides the theoretical guarantees, is to show the interplay of the meta-parameters more than prescript specific values.
-In nonconvex optimization, it is often the case that the meta-parameters prescribed by the theory are prohibitively conservative to be used in applications.
+1) Variance reduction algorithms for supervised learning often provide explicit guidelines on how to set the epoch size, e.g.
+O(N).
+In RL, this meta-parameter is more delicate since it is linked to the variance introduced by importance sampling.
+Theory would prescribe a large epoch size to avoid instabilities, but this would hinder the efficiency of the algorithm.
+The value optimizing this tradeoff is difficult to find and changes over time.
+It is also problem dependent since it also depends on the geometry of policy space.is reasons, we designed an adaptive algorithm, which is heuristic but based on insights from the theory.
+In the final version we explain this process better, and show examples of manual selection of m for sensitivity analysis.
+Future work will try to bridge the gap between theory and practice, e.g.
+by employing natural gradient or more sophisticated importance weighting techniques to reduce the sensitivity of SVRPG to this meta-parameter.
+2) In nonconvex optimization, it is often the case that the step sizes prescribed by theory are prohibitively conservative to be used in applications.
 For instance, in Reddi et al.
 the authors understand from their theoretical analysis that a constant step-size should be used, but they empirically search for the specific value that yields the best practical performance.
-Even in the original SVRG paper, common learning rate scheduling policies (ANNEALING SCHEMA??) are employed in the experiments.
-In RL, an example is the learning rate suggested in [Kakade, Sham, and John Langford.
-"Approximately optimal approximate reinforcement learning." ICML.
-Vol.
-2.
-2002.], which guarantees monotonic improvement but is prohibitively small for practical usage.
-However, the ideas embedded in that theoretical analysis ultimately lead to TRPO, which achieves great performance by relaxing several theoretical constraints.
-In our work, we both provide theoretical guarantees and propose a practical way to adapt meta-parameters.
-The statement of Theorem 4 (and some subtleties of its proof) heavily guided the choice of this heuristics, and we will make this more clear in the final version.
-3) \theta^* is a global optimizer and we should state it explicitly.
+Even in the original SVRG paper, common annealing schemas are employed in the experiments.
+3) In Theorem 4.1, \theta^* denotes a global optimizer, although only convergence to a local optimum is guaranteed. We will make this more clear.
 4) \theta_m^0, \tilde\theta^0 and \theta_0^1 are all the same.
@@
 We will check for notation inconsistencies in the text.
-5) Plain PG is an instance of Stochastic Gradient Ascent, which indeed has a 1/\sqrt{T} rate.
+5) Also under these assumptions, plain PG is an instance of Stochastic Gradient Ascent, which has a 1/\sqrt{T} rate.
 6) We consider the horizon H to be part of the problem, but we should mention that problems with long horizons can be more difficult since we employ importance weighting.
+In the final version we will show an example where we vary the horizon of a task for sensitivity analysis.
 

### 2018-04-14 svrpg.tex
updated rebuttals

@@
 It is easy to show that the SVRPG estimator has the following, desirable properties:
-\begin{lemma}\label{L:svrpg.properties}
+
+\begin{restatable}[]{lemma}{svrpgprop}\label{L:svrpg.properties}
         Let $\wh{\nabla}_N J(\vtheta)$ be an unbiased estimator of~\eqref{E:policygradient}
@@
 \end{equation}
-\end{lemma}
+\end{restatable}
 %\begin{proof}
@@
 
-%\subsection*{Basic Lemmas}
-%We prove two basic properties of the SVRPG update.
-%
-%%\unbias*
+\subsection*{Basic Lemmas}
+We prove two basic properties of the SVRPG update.
+
+\svrpgprop*
 %\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
@@
 %\end{restatable}
-%\begin{proof}
-%\begin{align*}
-%        \EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} - \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
-%&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
-%\end{align*}
-%Note that the importance weight is necessary to guarantee unbiasedness, since the $\tau_i$ are sampled from $\pi_{\vtheta}$.
+\begin{proof}
+\begin{align*}
+        \EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} - \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
+&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
+\end{align*}
+Note that the importance weight is necessary to guarantee unbiasedness, since the $\tau_i$ are sampled from $\pi_{\vtheta}$.
 %\end{proof}
-%
-%% \zerovar*
 %\begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
@@
 %\begin{proof}
-%As $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
-%\begin{align*}
-%\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*) - \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right] \\
-%&= \Var\left[\gradApp{\vtheta^*}{N}\right].
-%\end{align*}
-%Note that it is important that the trajectories used in the second and the third term are the same for the variance to vanish.
-%\end{proof}
+As $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
+	\begin{align*}
+\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*) - \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right] \\
+&= \Var\left[\gradApp{\vtheta^*}{N}\right].
+\end{align*}
+Note that it is important that the trajectories used in the second and the third term are the same for the variance to vanish.
+\end{proof}
 

### 2018-04-16 rebuttals.txt
Updated

@@
 The intrinsic difficulties of adapting SVRG to the RL framework have, as a consequence, a certain sensitivity to the meta-parameters.
-Our contribution should be intended as two-fold: a theoretical part supporting the usage of SVRG in RL, and a more empirical part highlighting the additional challenges of variance reduction in RL, where we also show how these difficulties can be overcome with proper heuristics.
+Our contribution should be intended as two-fold: a theoretical part supporting the use of SVRG in RL, and a more empirical part highlighting the additional challenges of variance reduction in RL (where we also show how these difficulties can be overcome with proper heuristics).
 We think the community would benefit from a picture of both the opportunities and the risks of applying SVRG to policy gradient learning.
@@
 values only for the Half-Cheetah experiment because it is the only case in which we used the critic.
-This represent a further step: evaluating the combination of SVRG and critic as variance reduction techniques.
+This represents a further step: evaluating the combination of SVRG and critic as variance reduction techniques.
 We recognize that the current presentation is confusing.
@@
 In our work, we both provide general theoretical guarantees and propose a practical way to select meta-parameters.
-The statement of Theorem 4 (and some subtleties of its proof) heavily guided the choice of this heuristics, and we will make this more clear in the final version.
+The statement of Theorem 4 (and some subtleties of its proof) is heavily guided the choice of this heuristics, and we will make this more clear in the final version.
 More specifically:
@@
 In RL, this meta-parameter is more delicate since it is linked to the variance introduced by importance sampling.
-Theory would prescribe a large epoch size to avoid instabilities, but this would hinder the efficiency of the algorithm.
+The theory would prescribe a large epoch size to avoid instabilities, but this would hinder the efficiency of the algorithm.
 The value optimizing this tradeoff is difficult to find and changes over time.
-It is also problem dependent since it also depends on the geometry of policy space.is reasons, we designed an adaptive algorithm, which is heuristic but based on insights from the theory.
-In the final version we explain this process better, and show examples of manual selection of m for sensitivity analysis.
+It is also problem dependent since it also depends on the geometry of policy space.
+For these reasons, we designed an adaptive algorithm, which is heuristic but based on insights from the theory.
+In the final version, we will explain this process better and we will show examples of manual selection of m for sensitivity analysis.
 Future work will try to bridge the gap between theory and practice, e.g.
 by employing natural gradient or more sophisticated importance weighting techniques to reduce the sensitivity of SVRPG to this meta-parameter.
-2) In nonconvex optimization, it is often the case that the step sizes prescribed by theory are prohibitively conservative to be used in applications.
+2) In nonconvex optimization, it is often the case that the step sizes prescribed by the theory are prohibitively conservative to be used in applications.
 For instance, in Reddi et al.
@@
 6) We consider the horizon H to be part of the problem, but we should mention that problems with long horizons can be more difficult since we employ importance weighting.
-In the final version we will show an example where we vary the horizon of a task for sensitivity analysis.
+In the final version, we will show an example where we vary the horizon of a task for sensitivity analysis.
 
@@
 1) The comparison is fair, figures have the number of trajectories on the horizontal axis, and the extra trajectories collected by  SVRPG at the snapshot are indeed counted.
-Our algorithm updates the policy less often, but manages to be faster anyway because the quality of the updates is higher.
-2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work since adapting batch sizes in the policy gradient framework is a  difficult problem that have been only slightly explored in the RL literature so far (cfr.
+Our algorithm updates the policy less often but manages to be faster anyway because the quality of the updates is higher.
+2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work since adapting batch sizes in the policy gradient framework is a  difficult problem that has been only slightly explored in the RL literature so far (cfr.
 Papini et al., "Adaptive batch size for safe policy gradients." NIPS 2017).

### 2018-04-16 rebuttals_mat.txt
Updated with new version

--- 2018-04-16 rebuttals.txt
@@
 
-Reviewer 1:
+AR1
+The heuristics proposed in the paper were not developed with the main purpose of beating the baseline, but to make the algorithm work (learn) in a stable way.
+The intrinsic difficulties of adapting SVRG to RL have, as a consequence, a certain sensitivity to the meta-parameters.
+Our contribution should be intended as two-fold: a theoretical part supporting the use of SVRG in RL, and a more empirical part highlighting the additional challenges of variance reduction in RL (where we also show how these difficulties can be overcome with proper heuristics).
+We think the community would benefit from a picture of both the opportunities and the risks of applying SVRG to PG.
 
-The heuristics proposed in the paper were not developed with the main purpose of beating the baseline, but to make the algorithm work (learn) in a stable way.
-The intrinsic difficulties of adapting SVRG to the RL framework have, as a consequence, a certain sensitivity to the meta-parameters.
-Our contribution should be intended as two-fold: a theoretical part supporting the use of SVRG in RL, and a more empirical part highlighting the additional challenges of variance reduction in RL (where we also show how these difficulties can be overcome with proper heuristics).
-We think the community would benefit from a picture of both the opportunities and the risks of applying SVRG to policy gradient learning.
-
-The proof to Lemma 3.1 was meant to be in the appendix and is missing by mistake.
-We will definitively include it.
+The proof to Lem 3.1 was meant to be in the appendix and is missing by mistake.
 The variance property is true at convergence because the O(1/B) terms cancel out as both the parameter and the snapshot parameter approach the optimum.
@@
 The importance weight is one because the two parameters are the same.
-See also a similar proof in Section 2 of the original SVRG paper.
+See also a similar proof in Sec 2 of the original SVRG paper.
 
@@
 Later, in the Half-Cheetah experiment, we include the critic in both algorithms.
-
 We reported Duan et al.
@@
 We will mention that experiments on Half-Cheetah are coherent with Duan et al., while the ones on Cart-Pole and Swimmer cannot be compared since we did not use the critic.
-
-The version of Cart-Pole we use is the same of Duan et al, and this should be mentioned, although the task is thoroughly described in Appendix E.
+The version of Cart-Pole we use is the same of Duan et al, and this should be mentioned, although the task is thoroughly described in App E.
 As for Swimmer, the lower scores are due to the fact that we do not include a critic in either GPOMDP or SVRPG.
@@
 
-Reviewer 2:
+AR2
 We see that meta-parameters are a delicate aspect of our algorithm.
-The pragmatic role of Theorem 4 is to show the interplay of these parameters more than prescript specific values.
+The pragmatic role of Thm 4 is to show the interplay of these parameters more than prescript specific values.
 In our work, we both provide general theoretical guarantees and propose a practical way to select meta-parameters.
-The statement of Theorem 4 (and some subtleties of its proof) is heavily guided the choice of this heuristics, and we will make this more clear in the final version.
-More specifically:
+The statement of Thm 4 (and some subtleties of its proof) heavily guided the choice of this heuristics, and we will make this more clear in the final version.
 
@@
 The theory would prescribe a large epoch size to avoid instabilities, but this would hinder the efficiency of the algorithm.
-The value optimizing this tradeoff is difficult to find and changes over time.
-It is also problem dependent since it also depends on the geometry of policy space.
+The optimal value is usually time and problem dependent (depends on the geometry of the policy space).
 For these reasons, we designed an adaptive algorithm, which is heuristic but based on insights from the theory.
-In the final version, we will explain this process better and we will show examples of manual selection of m for sensitivity analysis.
-Future work will try to bridge the gap between theory and practice, e.g.
-by employing natural gradient or more sophisticated importance weighting techniques to reduce the sensitivity of SVRPG to this meta-parameter.
-2) In nonconvex optimization, it is often the case that the step sizes prescribed by the theory are prohibitively conservative to be used in applications.
-For instance, in Reddi et al.
-the authors understand from their theoretical analysis that a constant step-size should be used, but they empirically search for the specific value that yields the best practical performance.
-Even in the original SVRG paper, common annealing schemas are employed in the experiments.
-3) In Theorem 4.1, \theta^* denotes a global optimizer, although only convergence to a local optimum is guaranteed. We will make this more clear.
-4) \theta_m^0, \tilde\theta^0 and \theta_0^1 are all the same.
-Superscripts denote epochs and subscripts inner iterations, as should be clear from the pseudocode.
-We will check for notation inconsistencies in the text.
-5) Also under these assumptions, plain PG is an instance of Stochastic Gradient Ascent, which has a 1/\sqrt{T} rate.
-6) We consider the horizon H to be part of the problem, but we should mention that problems with long horizons can be more difficult since we employ importance weighting.
-In the final version, we will show an example where we vary the horizon of a task for sensitivity analysis.
+In the final version, we will comment on this and we will do a sensitivity analysis for m.
+Future work will try to bridge the gap between theory and practice.
+2) In nonconvex opt., it is often the case that the step sizes prescribed by the theory are prohibitively conservative to be used in applications.
+For example, Reddi et al.
+performed empirical search for the step size in the experiments rather then use the theoretical one.
+Even in the original SVRG paper, common annealing schemas are employed.
+3) In Thm 4.1, \theta^* denotes a global optimizer, although only convergence to a local optimum is guaranteed. We will make this more clear.
+4) \theta_m^0, \tilde\theta^0 and \theta_0^1 are all the same. Superscripts denote epochs and subscripts inner iterations (see pseudocode). 
+5) Also under these assumptions, plain PG is an instance of Stochastic Gradient, which has a 1/\sqrt{T} rate.
+6) We consider the horizon H to be part of the problem.
+Long horizons can be more difficult since we employ importance weighting we will mention this in the final version and we will a sensitivity analysis (if space will allow us).
 
@@
 Our algorithm updates the policy less often but manages to be faster anyway because the quality of the updates is higher.
-2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work since adapting batch sizes in the policy gradient framework is a  difficult problem that has been only slightly explored in the RL literature so far (cfr.
-Papini et al., "Adaptive batch size for safe policy gradients." NIPS 2017).
-3) This is because we used the rllab variant of Cart-Pole (Duan et al., 2016), where the agent collects a reward of ~10 at each time-step, as described in Appendix E.
-4) For SVRPG, \alpha is selected with the Adam variant described in Section 5.2.
-For GPOMDP, we use Adam.
-All meta-parameters are reported in Appendix E and have been chosen to make the comparison fair.
+2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work.
+Adapting batch sizes in PG is a difficult problem that has been only slightly explored in the RL literature so far (cfr.
+Papini et al.
+2017).
+3) This is because we used the rllab variant of Cart-Pole (Duan et al., 2016), where the agent collects a reward of ~10 at each time-step, as described in App E.
+4) For SVRPG, \alpha is selected with the Adam variant described in Sec 5.2. For GPOMDP, we use Adam. All meta-parameters are reported in App E and have been chosen to make the comparison fair.
 
-Reviewer 3:
-
+AR3
 We thank the reviewer for the positive feedback and the suggestions on related works, which will be useful for future developments of our method.

### 2018-04-16 rebuttals_mat.txt
Updated

@@
 The heuristics proposed in the paper were not developed with the main purpose of beating the baseline, but to make the algorithm work (learn) in a stable way.
-The intrinsic difficulties of adapting SVRG to RL have, as a consequence, a certain sensitivity to the meta-parameters.
 Our contribution should be intended as two-fold: a theoretical part supporting the use of SVRG in RL, and a more empirical part highlighting the additional challenges of variance reduction in RL (where we also show how these difficulties can be overcome with proper heuristics).
@@
 More formally: "As \theta -> \theta^*, also \tilde\theta -> \theta^*; hence, by continuity of J: Var(SVRG) -> Var(\hat\nablaJ) + 1/B*Var[g(\theta^*) - 1*g(\theta^*)] = Var(\hat\nabla)".
-The importance weight is one because the two parameters are the same.
-See also a similar proof in Sec 2 of the original SVRG paper.
+The importance weight is one because the two parameters are the same (see also Sec 2 of the original SVRG paper).
 
@@
 because we do not employ the linear critic used in that paper.
-Since critic and SVRG are both ways to reduce the variance of the gradient estimate, and a combination of the two is non-trivial (how do you critique snapshot policies?) we turned the critic off in both GPOMDP and SVRPG to measure the net advantage of SVRG.
-Later, in the Half-Cheetah experiment, we include the critic in both algorithms.
+Since critic and SVRG are both ways to reduce the variance of the gradient estimate, and a combination of the two is non-trivial we turned the critic off in both GPOMDP and SVRPG to measure the net advantage of SVRG.
 We reported Duan et al.
@@
 We will mention that experiments on Half-Cheetah are coherent with Duan et al., while the ones on Cart-Pole and Swimmer cannot be compared since we did not use the critic.
-The version of Cart-Pole we use is the same of Duan et al, and this should be mentioned, although the task is thoroughly described in App E.
+Cart-Pole we used is the same of Duan et al, and this should be mentioned, although the task is thoroughly described in App E.
 As for Swimmer, the lower scores are due to the fact that we do not include a critic in either GPOMDP or SVRPG.
@@
 The pragmatic role of Thm 4 is to show the interplay of these parameters more than prescript specific values.
-In our work, we both provide general theoretical guarantees and propose a practical way to select meta-parameters.
+We both provided general theoretical guarantees and proposed a practical way to select meta-parameters.
 The statement of Thm 4 (and some subtleties of its proof) heavily guided the choice of this heuristics, and we will make this more clear in the final version.
 
-1) Variance reduction algorithms for supervised learning often provide explicit guidelines on how to set the epoch size, e.g.
+1) Variance reduction algorithms for SL often provide explicit guidelines on how to set the epoch size, e.g.
 O(N).
@@
 The theory would prescribe a large epoch size to avoid instabilities, but this would hinder the efficiency of the algorithm.
-The optimal value is usually time and problem dependent (depends on the geometry of the policy space).
+The optimal value is usually time and problem dependent (eg the geometry of the policy space).
 For these reasons, we designed an adaptive algorithm, which is heuristic but based on insights from the theory.
@@
 Our algorithm updates the policy less often but manages to be faster anyway because the quality of the updates is higher.
-2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work.
-Adapting batch sizes in PG is a difficult problem that has been only slightly explored in the RL literature so far (cfr.
-Papini et al.
-2017).
+2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work. This topic has been faced, for example, in (Papini et al. 2017) for PG.
 3) This is because we used the rllab variant of Cart-Pole (Duan et al., 2016), where the agent collects a reward of ~10 at each time-step, as described in App E.

### 2018-04-16 rebuttals.txt
rebuttals are now 6000 chars

@@
 
-Reviewer 1:
+AR1
+The heuristics proposed in the paper were not developed with the main purpose of beating the baseline, but to make the algorithm work in a stable way.
+Our contribution should be intended as two-fold: a theoretical part supporting the use of SVRG in RL, and an empirical part highlighting the additional challenges of variance reduction in RL (where we also show how these difficulties can be overcome with proper heuristics).
+We think the community would benefit from a picture of both the opportunities and the risks of applying SVRG to PG.
 
-The heuristics proposed in the paper were not developed with the main purpose of beating the baseline, but to make the algorithm work (learn) in a stable way.
-The intrinsic difficulties of adapting SVRG to the RL framework have, as a consequence, a certain sensitivity to the meta-parameters.
-Our contribution should be intended as two-fold: a theoretical part supporting the use of SVRG in RL, and a more empirical part highlighting the additional challenges of variance reduction in RL (where we also show how these difficulties can be overcome with proper heuristics).
-We think the community would benefit from a picture of both the opportunities and the risks of applying SVRG to policy gradient learning.
-
-The proof to Lemma 3.1 was meant to be in the appendix and is missing by mistake.
-We will definitively include it.
+The proof to Lem 3.1 was meant to be in the appendix and is missing by mistake.
 The variance property is true at convergence because the O(1/B) terms cancel out as both the parameter and the snapshot parameter approach the optimum.
 More formally: "As \theta -> \theta^*, also \tilde\theta -> \theta^*; hence, by continuity of J: Var(SVRG) -> Var(\hat\nablaJ) + 1/B*Var[g(\theta^*) - 1*g(\theta^*)] = Var(\hat\nabla)".
-The importance weight is one because the two parameters are the same.
-See also a similar proof in Section 2 of the original SVRG paper.
+The importance weight is one because the two parameters are the same (see also Sec 2 of the original SVRG paper).
 
@@
 because we do not employ the linear critic used in that paper.
-Since critic and SVRG are both ways to reduce the variance of the gradient estimate, and a combination of the two is non-trivial (how do you critique snapshot policies?) we turned the critic off in both GPOMDP and SVRPG to measure the net advantage of SVRG.
-Later, in the Half-Cheetah experiment, we include the critic in both algorithms.
-
+Since critic and SVRG are both ways to reduce the variance of the gradient estimate, and a combination of the two is non-trivial we turned the critic off in both GPOMDP and SVRPG to measure the net advantage of SVRG.
 We reported Duan et al.
@@
 We will mention that experiments on Half-Cheetah are coherent with Duan et al., while the ones on Cart-Pole and Swimmer cannot be compared since we did not use the critic.
-
-The version of Cart-Pole we use is the same of Duan et al, and this should be mentioned, although the task is thoroughly described in Appendix E.
+Cart-Pole we used is the same of Duan et al, and this should be mentioned, although the task is thoroughly described in App E.
 As for Swimmer, the lower scores are due to the fact that we do not include a critic in either GPOMDP or SVRPG.
@@
 Recent policy optimization algorithms employ a series of tools, e.g.
-natural gradients and exploration bonuses, that are outside the scope of this paper but represent promising future-work material.
+natural gradients and exploration bonus, that are outside the scope of this paper but represent promising future-work material.
 
@@
 
-Reviewer 2:
+AR2
 We see that meta-parameters are a delicate aspect of our algorithm.
-The pragmatic role of Theorem 4 is to show the interplay of these parameters more than prescript specific values.
-In our work, we both provide general theoretical guarantees and propose a practical way to select meta-parameters.
-The statement of Theorem 4 (and some subtleties of its proof) is heavily guided the choice of this heuristics, and we will make this more clear in the final version.
-More specifically:
+The pragmatic role of Thm 4 is to show the interplay of these parameters more than prescript specific values.
+We both provide general theoretical guarantees and propose a practical way to select meta-parameters.
+The statement of Thm 4 (and some subtleties of its proof) heavily guided the choice of this heuristics, and we will make this more clear in the final version.
 
-1) Variance reduction algorithms for supervised learning often provide explicit guidelines on how to set the epoch size, e.g.
+1) Variance reduction algorithms for SL often provide explicit guidelines on how to set the epoch size, e.g.
 O(N).
@@
 The theory would prescribe a large epoch size to avoid instabilities, but this would hinder the efficiency of the algorithm.
-The value optimizing this tradeoff is difficult to find and changes over time.
-It is also problem dependent since it also depends on the geometry of policy space.
+The optimal value is usually time and problem dependent (eg the geometry of the policy space).
 For these reasons, we designed an adaptive algorithm, which is heuristic but based on insights from the theory.
-In the final version, we will explain this process better and we will show examples of manual selection of m for sensitivity analysis.
-Future work will try to bridge the gap between theory and practice, e.g.
-by employing natural gradient or more sophisticated importance weighting techniques to reduce the sensitivity of SVRPG to this meta-parameter.
-2) In nonconvex optimization, it is often the case that the step sizes prescribed by the theory are prohibitively conservative to be used in applications.
-For instance, in Reddi et al.
-the authors understand from their theoretical analysis that a constant step-size should be used, but they empirically search for the specific value that yields the best practical performance.
-Even in the original SVRG paper, common annealing schemas are employed in the experiments.
-3) In Theorem 4.1, \theta^* denotes a global optimizer, although only convergence to a local optimum is guaranteed. We will make this more clear.
-4) \theta_m^0, \tilde\theta^0 and \theta_0^1 are all the same.
-Superscripts denote epochs and subscripts inner iterations, as should be clear from the pseudocode.
-We will check for notation inconsistencies in the text.
-5) Also under these assumptions, plain PG is an instance of Stochastic Gradient Ascent, which has a 1/\sqrt{T} rate.
-6) We consider the horizon H to be part of the problem, but we should mention that problems with long horizons can be more difficult since we employ importance weighting.
-In the final version, we will show an example where we vary the horizon of a task for sensitivity analysis.
+In the final version, we will comment on this and we will do a sensitivity analysis for m.
+Future work will try to bridge the gap between theory and practice.
+2) In nonconvex opt., it is often the case that the step sizes prescribed by the theory are prohibitively conservative to be used in applications.
+For example, Reddi et al.
+performed empirical search for the step size in the experiments rather then use the theoretical one.
+Even in the original SVRG paper, common annealing schemas are employed.
+3) In Thm 4.1, \theta^* denotes a global optimizer, although only convergence to a local optimum is guaranteed. We will make this more clear.
+4) \theta_m^0, \tilde\theta^0 and \theta_0^1 are all the same. Superscripts denote epochs and subscripts inner iterations (see pseudocode). 
+5) Also under these assumptions, plain PG is an instance of Stochastic Gradient, which has a 1/\sqrt{T} rate.
+6) We consider the horizon H to be part of the problem.
+Long horizons can be more difficult since we employ importance weighting we will mention this in the final version and we will a sensitivity analysis (if space will allow us).
 
@@
 Our algorithm updates the policy less often but manages to be faster anyway because the quality of the updates is higher.
-2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work since adapting batch sizes in the policy gradient framework is a  difficult problem that has been only slightly explored in the RL literature so far (cfr.
-Papini et al., "Adaptive batch size for safe policy gradients." NIPS 2017).
-3) This is because we used the rllab variant of Cart-Pole (Duan et al., 2016), where the agent collects a reward of ~10 at each time-step, as described in Appendix E.
-4) For SVRPG, \alpha is selected with the Adam variant described in Section 5.2.
-For GPOMDP, we use Adam.
-All meta-parameters are reported in Appendix E and have been chosen to make the comparison fair.
+2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work. This topic has been faced, for example, in (Papini et al. 2017) for PG.
+3) This is because we used the rllab variant of Cart-Pole (Duan et al., 2016), where the agent collects a reward of ~10 at each time-step, as described in App E.
+4) For SVRPG, \alpha is selected with the Adam variant described in Sec 5.2. For GPOMDP, we use Adam. All meta-parameters are reported in App E and have been chosen to make the comparison fair.
 
-Reviewer 3:
-
+AR3
 We thank the reviewer for the positive feedback and the suggestions on related works, which will be useful for future developments of our method.

### 2018-04-16 rebuttals.txt
Minor fixes

@@
 6) We consider the horizon H to be part of the problem.
-Long horizons can be more difficult since we employ importance weighting we will mention this in the final version and we will a sensitivity analysis (if space will allow us).
+Long horizons can be more difficult since we employ importance weighting.
+We will mention this in the final version and we will perform a sensitivity analysis.
 

### 2018-04-18 reviews/rebuttals.txt
Moved rebuttal in subfolder

--- 2018-04-16 rebuttals.txt
@@
-We would like to thank all the reviewers for their feedback and in-depth comments.
+We would like to thank the reviewers for their feedback and in-depth comments.
 
@@
 This represents a further step: evaluating the combination of SVRG and critic as variance reduction techniques.
-We recognize that the current presentation is confusing.
-In fact, we now have reproduced the GPOMDP/Half-Cheetah result and we will simply report our results in the final version.
+We see that the current presentation is confusing.
+In fact, we now have reproduced the GPOMDP/Cheetah result and we will simply report our results in the final version.
 We will mention that experiments on Half-Cheetah are coherent with Duan et al., while the ones on Cart-Pole and Swimmer cannot be compared since we did not use the critic.
@@
 Recent policy optimization algorithms employ a series of tools, e.g.
-natural gradients and exploration bonus, that are outside the scope of this paper but represent promising future-work material.
+natural gradient and exploration bonus, that are outside the scope of this paper but represent promising future-work material.
 

### 2018-05-21 svrpg.tex
minor changes/additions suggested by reviewers

@@
 %Paper structure
-After providing background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} propose SVRPG, a variant of SVRG for the policy gradient framework, addressing all the difficulties mentioned above.
+After providing background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} we propose SVRPG, a variant of SVRG for the policy gradient framework, addressing all the difficulties mentioned above.
 In Section \ref{sec:conv} we provide convergence guarantees for our algorithm, and we show a convergence rate that has an $O(\nicefrac{1}{T})$ dependence on the number $T$ of iterations.
@@
 As a consequence, the lower per-iteration cost is paid with a worse, sub-linear convergence rate \cite{nemirovskii1983problem}.
-Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga}, and Finito~\cite{defazio2014finito}.
+Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga}, Finito~\cite{defazio2014finito}, and MISO~\cite{mairal2015incremental}.
 The common idea is to reuse past gradient computations to reduce the variance of the current estimate.
@@
 In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate.
+\citet{bietti2017stochastic} performed a similar analysis on MISO.
 In Section \ref{sec:conv} we will show how the estimation accuracy impacts on the convergence results with a non-concave objective.
@@
 	\begin{algorithmic}
-		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{m}^0 := \wt{\vtheta}^0$
+		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{m}^0 \coloneqq \wt{\vtheta}^0 \coloneqq \vtheta_0$
 		\FOR{$s=0$ {\bfseries to} $S-1$}
@@
 \textit{3) Non-stationarity.} 
-Similarly to what done in SL~\citep{cortes2010learning}, we require the variance of the importance weight to be bounded.
+Similarly to what is done in SL~\citep{cortes2010learning}, we require the variance of the importance weight to be bounded.
 	\begin{restatable}[On the variance of importance weights]{assumption}{varweights}\label{ass:M2}
@@
 \end{align*}
-where $\psi,\zeta,\xi$ depend only on $\GRADLOG,\HESSLOG,\VARRF,\VARIS,\alpha$ and $m$.
+where $\vtheta^*$ is a global optimum and $\psi,\zeta,\xi$ depend only on $\GRADLOG,\HESSLOG,\VARRF,\VARIS,\alpha$ and $m$.
 \end{restatable}

### 2018-05-23 svrpg.tex
added cheetah figure, removed table

@@
 % Use the following line for the initial blind version submitted for review:
-\usepackage{icml2018}
+%\usepackage{icml2018}
 
 % If accepted, instead use the following line for the camera-ready submission:
-%\usepackage[accepted]{icml2018}
+\usepackage[accepted]{icml2018}
 
@@
 \usepackage{placeins}
+\usepackage{subcaption}
 
@@
 In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer~\citep[\eg][]{duan2016benchmarking}.
-We consider G(PO)MDP since has a smaller variance than REINFORCE.
+We consider G(PO)MDP since it has a smaller variance than REINFORCE.
 % This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases. 
@@
 To make the comparison fair, we also use Adam in the baseline PG algorithm, which we will denote simply as G(PO)MDP in the following.
-In all the experiments we used deep Gaussian policies with adaptive standard deviation (details on network architecture in Appendix \ref{app:exp}).
+In all the experiments, we use deep Gaussian policies with adaptive standard deviation (details on network architecture in Appendix \ref{app:exp}).
 Each experiment is run $10$ times with a random policy initialization and seed, but this initialization is shared among the algorithms under comparison.
@@
 We provide average performance with 90\% bootstrap confidence intervals.
-Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}, on which our agents are also based.\footnote{rllab can be found at \url{https://github.com/rll/rllab}, while the link to our implementation is not reported for blindness, but it is already publically available.}
+Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}, on which our agents are also based.\footnote{Code available at \url{github.com/Dam930/rllab}.}
 More details on meta-parameters and exhaustive task descriptions are provided in Appendix \ref{app:exp}.
@@
 \begin{figure*}[t]
-    \begin{minipage}[t]{0.32\textwidth}
+    \begin{subfigure}[b]{0.49\textwidth}
 	\includegraphics[width=.98\textwidth]{images/cart_pole_SVRPG_vs_GPOMDP_tex.pdf}
 	\vspace{-0.1in}
-	\caption{On-line performance over sampled trajectories of SVRPG vs G(PO)MDP in the Cart-pole task, with 90\% confidence intervals.}
+	\caption{SVRPG vs G(PO)MDP on Cart-pole.}
 	\label{fig:cartpole}
-    \end{minipage}
-    \hfill
-    \begin{minipage}[t]{0.32\textwidth}
+    \end{subfigure}\hfill
+    \begin{subfigure}[b]{0.49\textwidth}
 	\includegraphics[width=0.98\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG_tex.pdf}
 	\vspace{-0.1in}
-	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task, with 90\% confidence intervals.}
+	\caption{Self-Normalized SVRPG vs SVRPG on Swimmer.}
 	\label{fig:swimmertwo}
-    \end{minipage}
-    \hfill
-    \begin{minipage}[t]{.32\textwidth}
+    \end{subfigure}
+
+    \begin{subfigure}[b]{.49\textwidth}
 	\includegraphics[width=0.98\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP_tex.pdf}
 	\vspace{-0.1in}
-	\caption{On-line performance over sampled trajectories of self-Normalized SVRPG vs G(PO)MDP in the Swimmer task, with 90\% confidence intervals}
+	\caption{Self-Normalized SVRPG vs G(PO)MDP on Swimmer.}
 	\label{fig:swimmerone}
-    \end{minipage}
-    \vspace{-0.15in}
+    \end{subfigure}\hfill
+    \begin{subfigure}[b]{.49\textwidth}
+    	\includegraphics[width=0.98\textwidth]{half_cheetah_Self_Normalized_SVRPG_vs_GPOMDP.pdf}
+    	\vspace{-0.1in}
+    	\caption{Self-Normalized SVRPG vs G(PO)MDP on Half-Cheetah.}
+    	\label{fig:cheetah}
+    \end{subfigure}
+    \caption{Comparison of on-line performance over sampled trajectories, with 90\% confidence intervals.}
 \end{figure*}
 
-\begin{table*}[t]
-	\caption{
-		\label{tab:1}
-		Comparison between G(PO)MDP and SVRPG on different tasks.
+%
+%\begin{table*}[t]
+%	\caption{
+%		\label{tab:1}
+%		Comparison between G(PO)MDP and SVRPG on different tasks.
 Average return over the whole learning process is reported, together with normalized standard deviation and 90\% confidence intervals.
 Results denoted with $\dagger$ are taken from \cite{duan2016benchmarking}.}
-	\centering
-	\begin{tabular}{@{}lccccccc@{}} 
-		\toprule
-		\phantom{abc} & \multicolumn{3}{c}{G(PO)MDP} & \phantom{abc} & \multicolumn{3}{c}{SVRPG} \\
-		\cmidrule{2-4} \cmidrule{6-8}
-		\phantom{abc} & Avg Return & Norm. Std & C.I. & \phantom{abc}
-		& Avg Return & Norm. Std & C.I.
-		\\\cmidrule{2-8}
-		Cart-pole & $618.49$ & $21.09$ & $[579.82, 657.16]$ & \phantom{abc} 
-		& $736.31$ & $16.51$ & $[706.04, 766.57]$ \\
-		Swimmer & $39.86$ & $3.84$ & $[32.82, 46.89]$ & \phantom{abc} & $49.60$ & $3.27$ & $[43.60, 55.60]$\\
-		Half-cheetah & $1183.1^\dagger$ & $69.20^\dagger$ & $[1035.58, 1330.62]^\dagger$ & \phantom{abc} & $1480.95$ & $73.21$ & $[1324.87, 1637.03]$\\
-		\bottomrule
-	\end{tabular}
-    \vspace{-0.05in}
-\end{table*}
+%	\centering
+%	\begin{tabular}{@{}lccccccc@{}} 
+%		\toprule
+%		\phantom{abc} & \multicolumn{3}{c}{G(PO)MDP} & \phantom{abc} & \multicolumn{3}{c}{SVRPG} \\
+%		\cmidrule{2-4} \cmidrule{6-8}
+%		\phantom{abc} & Avg Return & Norm. Std & C.I. & \phantom{abc}
+%		& Avg Return & Norm. Std & C.I.
+%		\\\cmidrule{2-8}
+%		Cart-pole & $618.49$ & $21.09$ & $[579.82, 657.16]$ & \phantom{abc} 
+%		& $736.31$ & $16.51$ & $[706.04, 766.57]$ \\
+%		Swimmer & $39.86$ & $3.84$ & $[32.82, 46.89]$ & \phantom{abc} & $49.60$ & $3.27$ & $[43.60, 55.60]$\\
+%		Half-cheetah & $1183.1^\dagger$ & $69.20^\dagger$ & $[1035.58, 1330.62]^\dagger$ & \phantom{abc} & $1480.95$ & $73.21$ & $[1324.87, 1637.03]$\\
+%		\bottomrule
+%	\end{tabular}
+%    \vspace{-0.05in}
+%\end{table*}
 
@@
 
-Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for online learning since it is related to an average regret~\citep[\eg][]{duan2016benchmarking}.
+%Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for online learning since it is related to an average regret~\citep[\eg][]{duan2016benchmarking}.
 This metric is related to the area below the curves.
@@
 We also report normalized standard deviations and 90\% $t$-confidence intervals.
-In both tasks, our algorithm has higher average return and lower normalized standard deviation.
-In Cart-pole, disjoint confidence intervals show that the advantage is statistically significant.
+%In both tasks, our algorithm has higher average return and lower normalized standard deviation.
+%In Cart-pole, disjoint confidence intervals show that the advantage is statistically significant.
 
@@
 To do so, we consider the more challenging MuJoCo~\citep{todorov2012mujoco} Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer.
-Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP\footnote{\citet{duan2016benchmarking} report results on REINFORCE.
+Figure \ref{fig:cheetah} compares self-normalized SVRPG and G(PO)MDP on Half-cheetah, using the critic suggested in \cite{duan2016benchmarking} for both algorithms.
+Results are promising, showing that a combination of the baseline usage and SVRG-like variance reduction can yield an improvement that the two techniques alone are not able to achieve.
+Moreover, SVRPG presents a noticeably lower variance.
+The performance of actor-critic G(PO)MDP\footnote{\citet{duan2016benchmarking} report results on REINFORCE.
 However, inspection on \textit{rllab} code and documentation reveals that it is actually PGT \cite{sutton2000policy}, which is equivalent to G(PO)MDP \citep[shown by][]{peters2008reinforcement}.
-Using the name REINFORCE in a general way is inaccurate, but widespread.} on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms.
-Results are promising, showing that a combination of the baseline usage and SVRG-like variance reduction can yield an improvement that the two techniques alone are not able to achieve.
+Using the name REINFORCE in a general way is inaccurate, but widespread.} on Half-Cheetah is coherent with the one reported in \cite{duan2016benchmarking}.
+Other results are not comparable since we did not use the critic.
 

### 2018-05-30 svrpg.tex
Updated

@@
 On a very general level, artificial intelligence addresses the problem of an agent that must select the right actions to solve a task.
-The approach of Reinforcement Learning~\citep{sutton1998reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal.
+The approach of Reinforcement Learning (RL)~\citep{sutton1998reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal.
 This makes RL fundamentally different from Supervised Learning (SL), where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels).
@@
 This rules out SAG and SAGA because of their storage requirements, which leaves SVRG as the most promising choice.
-Finally, the distribution used to sample data is not under direct control of the algorithm designer, but it is a function of policy parameters that changes over time as the policy is optimized, which is a form of non-stationarity.
+Finally, the distribution used to sample data is not under direct control of the algorithm designer, but it is a function of policy parameters that change over time as the policy is optimized, which is a form of non-stationarity.
 % Related work?
@@
 %Paper structure
-After providing background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} we propose SVRPG, a variant of SVRG for the policy gradient framework, addressing all the difficulties mentioned above.
+After providing background on policy gradient and SVRG in Section \ref{sec:pre}, we propose SVRPG, a variant of SVRG for the policy gradient framework, addressing all the difficulties mentioned above (see Section \ref{sec:alg}).
 In Section \ref{sec:conv} we provide convergence guarantees for our algorithm, and we show a convergence rate that has an $O(\nicefrac{1}{T})$ dependence on the number $T$ of iterations.
@@
 % In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
-We consider episodic MDPs with effective horizon $H$.\footnote{The episode duration is a random variable, but the optimal policy can reach the target state (\ie absorbing state) in at most $H$ steps.
+We consider episodic MDPs with effective horizon $H$.\footnote{The episode duration is a random variable, but the optimal policy can reach the target state (\ie absorbing state) in less than $H$ steps.
 This has not to be confused with a finite horizon problem where the optimal policy is non-stationary.} In this setting, we can limit our attention to trajectories of length $H$.
@@
 \vspace{-0.05in}
-Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $z_i(\cdot|\vtheta)$:
+Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $z_i(\vtheta)$:
 \begin{align*}
@@
 The idea of SVRG (Algorithm~\ref{alg:svrg}) is to alternate full and stochastic gradient updates. 
-Each $m = \mathcal{O}(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i \nabla z(x_i|\widetilde{\vtheta})$.
+Each $m = O(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i \nabla z(x_i|\widetilde{\vtheta})$.
 Between snapshots, the parameter is updated with $\blacktriangledown f(\vtheta)$, a gradient estimate corrected using stochastic gradient. For any $t \in \{0,\ldots,m-1\}$:
@@
 More recently, some extensions of variance reduction algorithms to the non-concave objectives have been proposed~\citep[\eg][]{allen2016variance,reddi2016stochastic,reddi2016fast}.
-In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[2]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[2]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
+In this scenario, $f$ is typically required to be $L$-smooth, \ie $\norm[2]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[2]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$.
 Under this hypothesis, the convergence rate of SG is $O(\nicefrac{1}{\sqrt{T}})$ \cite{ghadimi2013stochastic}, \ie $T=O(\nicefrac{1}{\epsilon^2})$ iterations are required to get $\norm[2]{\nabla f(\vtheta)}^2\leq\epsilon$.
 Again, SVRG achieves the same rate as FG \cite{reddi2016stochastic}, which is $O(\frac{1}{T})$ in this case \cite{nesterov2013introductory}.
-The only additional requirement is to select $\vtheta^*$ uniformly at random among all the $\vtheta_k$ instead of simply setting it to the final value.
+The only additional requirement is to select $\vtheta^*$ uniformly at random among all the $\vtheta_k$ instead of simply setting it to the final value ($k$ being the iterations).
 

### 2018-05-31 svrpg.tex
Updated

@@
 and regardless of the mini-batch size $B$:\footnote{
-For any vector $\mathbf{x}$, we use $\Var[\mathbf{x}]$ to denote the trace of the covariance matrix, \ie $\Tr\EVV[]{(\mathbf{x}-\EVV[]{\mathbf{x}})(\mathbf{x}-\EVV[]{\mathbf{x}})^T}$.}
+For any vector $\mathbf{x}$, we use $\Var[\mathbf{x}]$ to denote the trace of the covariance matrix, \ie $\Tr(\EVV[]{(\mathbf{x}-\EVV[]{\mathbf{x}})(\mathbf{x}-\EVV[]{\mathbf{x}})^T})$.}
 \begin{equation}\label{eq:zerovar}
@@
 %The theoretical settings may be too conservative.
-By analysing the upper-bound in Theorem~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on non-concave SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
+By analysing the upper-bound in Theorem~\ref{theo:convergence} we observe that: I) the $O(\nicefrac{1}{T})$ term is coherent with results on non-concave SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $O(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $O(\nicefrac{1}{B})$ term is due to importance weighting.
 To achieve asymptotic convergence, the batch size $N$ and the mini-batch size $B$ should increase over time. 

### 2018-06-11 poster/poster.tex
started working on poster

\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
\documentclass[portrait,a0paper,fontscale=0.312]{baposter}

% For graphs
\usepackage{graphicx}

\usepackage{array}
\usepackage{booktabs}
\usepackage{eso-pic}
\usepackage{layout}
\usepackage{fancybox}

\usepackage{calc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{relsize}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{bm}
\usepackage{url}
\usepackage{xfrac}
\usepackage{natbib}
\usepackage{mathtools}
\usepackage{cancel}
\usepackage{paralist}
\usepackage{nicefrac}
\usepackage[export]{adjustbox} % loads also graphicx
\usepackage{caption}
\usepackage{xfrac}

% \usepackage[centering,includeheadfoot,margin=2cm]{geometry}

\usepackage{multicol}

% \usepackage[linesnumbered,ruled,vlined,noend]{algorithm2e}
% % \usepackage{algorithmicx}
% % \usepackage{algpseudocode}
% \newlength\figureheight
% \newlength\figurewidth
% \setlength{\algomargin}{2em}
% \SetKwComment{Comment}{$\blacktriangleright$\ }{}
% 
% \providecommand{\SetAlgoLined}{\SetLine}
% \providecommand{\DontPrintSemicolon}{\dontprintsemicolon}

\newcommand{\figurewidth}{7cm}
\newcommand{\figureheight}{3cm}

%\usepackage{times}
%\usepackage{helvet}
%\usepackage{bookman}
\usepackage{palatino}

%\newcommand{\captionfont}{\footnotesize}

\usetikzlibrary{calc}

% \newcommand{\SET}[1]  {\ensuremath{\mathcal{#1}}}
% \newcommand{\MAT}[1]  {\ensuremath{\boldsymbol{#1}}}
% \newcommand{\VEC}[1]  {\ensuremath{\boldsymbol{#1}}}
% \newcommand{\Video}{\SET{V}}
% \newcommand{\video}{\VEC{f}}
% \newcommand{\track}{x}
% \newcommand{\Track}{\SET T}
% \newcommand{\LMs}{\SET L}
% \newcommand{\lm}{l}
% \newcommand{\PosE}{\SET P}
% \newcommand{\posE}{\VEC p}
% \newcommand{\negE}{\VEC n}
% \newcommand{\NegE}{\SET N}
% \newcommand{\Occluded}{\SET O}
% \newcommand{\occluded}{o}

\renewcommand{\sfdefault}{lmss}
\sffamily

\newcommand{\listhead}[1] {\textsc{\underline{#1}}}

\definecolor{rouge1}{RGB}{226,0,38}  % red P
\definecolor{orange1}{RGB}{243,154,38}  % orange P
\definecolor{jaune}{RGB}{254,205,27}  % jaune P
\definecolor{blanc}{RGB}{255,255,255} % blanc P

\definecolor{rouge2}{RGB}{230,68,57}  % red S
\definecolor{orange2}{RGB}{236,117,40}  % orange S
\definecolor{taupe}{RGB}{134,113,127} % taupe S
\definecolor{gris}{RGB}{91,94,111} % gris S
\definecolor{bleu1}{RGB}{38,109,131} % bleu S
\definecolor{bleu2}{RGB}{28,50,114} % bleu S
\definecolor{vert1}{RGB}{133,146,66} % vert S
\definecolor{vert3}{RGB}{20,200,66} % vert S
\definecolor{vert2}{RGB}{157,193,7} % vert S
\definecolor{darkyellow}{RGB}{233,165,0}  % orange S
\definecolor{lightgray}{rgb}{0.9,0.9,0.9}
\definecolor{darkgray}{rgb}{0.6,0.6,0.6}

\definecolor{blue900}{HTML}{0D47A1}
\definecolor{blue800}{HTML}{1565C0}

\definecolor{green1}{RGB}{217,255,250}
\definecolor{orange1}{RGB}{250,245,217}

\newcommand{\rcol}[1]{\textcolor{red}{\textit{#1}}}
\newcommand{\gcol}[1]{\textcolor{vert3}{\textit{#1}}}
\newcommand{\bcol}[1]{\textcolor{blue}{\textit{#1}}}
\newcommand{\ycol}[1]{\textcolor{darkyellow}{\textit{#1}}}

\newcommand{\rcolb}[1]{\textcolor{red}{\textit{\textbf{#1}}}}
\newcommand{\gcolb}[1]{\textcolor{vert3}{\textit{\textbf{#1}}}}
\newcommand{\bcolb}[1]{\textcolor{blue}{\textit{\textbf{#1}}}}
\newcommand{\ycolb}[1]{\textcolor{darkyellow}{\textit{\textbf{#1}}}}

\newcommand{\otoprule}{\midrule[\heavyrulewidth]}
\newcommand{\dbacks}[1]{\textbf{\textcolor{red!80!black}{{#1}}}}

\usepackage{tikz,pgfplots}
\pgfplotsset{compat=newest}
\tikzstyle{every picture}+=[remember picture]
\tikzstyle{na} = [baseline=-.5ex]
\everymath{\displaystyle}
\usetikzlibrary{arrows,shapes}
\usetikzlibrary{positioning}

\usepackage{wasysym}

\usepackage{tcolorbox}

\usepackage{algorithm}
\usepackage{algorithmic}

\usepackage{xspace}
\DeclareRobustCommand{\eg}{e.g.,\@\xspace}
\DeclareRobustCommand{\ie}{i.e.,\@\xspace}
\DeclareRobustCommand{\wrt}{w.r.t.\@\xspace}
\DeclareRobustCommand{\wp}{w.p.\@\xspace}

%=============================================
% General commands
%---------------------------------------------
\newcommand{\transpose}[1]{{#1}^\texttt{T}}
\DeclareMathOperator*{\EV}{\mathbb{E}}
\DeclareMathOperator*{\argmax}{\arg\,\max}
\DeclareMathOperator*{\argmin}{\arg\,\min}
\DeclareMathOperator*{\arginf}{\arg\,\inf}
\newcommand{\EVV}[2][x]{\EV_{#1}\left[{#2}\right]}
\newcommand{\norm}[2][\infty]{\left\|#2\right\|_{#1}}
\newcommand{\indfun}[1]{\mathds{1}\left(#1\right)}

\newcommand{\realspace}{\mathbb R}		% realspace
\newcommand{\natspace}{\mathbb N}		% naturalspace
\newcommand{\pdfunc}[1]{p\left(#1\right)}	% density function
\newcommand{\prob}[1]{Pr\left(#1\right)}	% probability
\newcommand{\funoper}[1]{\left[#1\right]}	% function operator
\newcommand{\apx}[1]{\widetilde{#1}}		% approximation symbol
\newcommand{\e}{\boldsymbol{e}} 		% unit vector

\newcommand{\texsub}[1]{\textsc{\tiny #1}}

%=============================================
% RL Commands
%---------------------------------------------
%%% MDPs
\newcommand{\mdp}{\mathcal{M}}				% MDP
\newcommand{\statespace}{\mathcal S}			% state space
\newcommand{\actionspace}{\mathcal A}			% action space
\newcommand{\pmodel}{\mathcal P}			% transition function
\newcommand{\pmodelfun}[1]{\pmodel\left(#1\right)}	% transition function
\newcommand{\rmodel}{\mathcal R}			% reward function
\newcommand{\rmodelfun}[1]{\rmodel\left(#1\right)}	% reward function
\newcommand{\initD}{\mu}					% initial distribution

%% Policy Gradient
\newcommand{\vtheta}{\boldsymbol{\theta}}
\newcommand{\gradJ}[1]{\nabla_{#1}J_\mu(\vtheta)}
\newcommand{\vphi}{\boldsymbol{\phi}}
\newcommand{\DeltaJ}{J(\vtheta') - J(\vtheta)}
\newcommand{\gradApp}[1]{\widehat{\nabla}_{#1}J(\vtheta)}
\newcommand{\gradSim}[1]{\tilde{\nabla}_{#1}J(\vtheta)}
\newcommand{\gradRF}[1]{\hat{\nabla}_{#1}J^{RF}(\vtheta)}
\newcommand{\gradPGT}[1]{\hat{\nabla}_{#1}J^{PGT}(\vtheta)}
\newcommand{\gradGPOMDP}[1]{\hat{\nabla}_{#1}J^{G(PO)MDP}(\vtheta)}
\newcommand{\gradDown}[1]{\underline{\hat{\nabla}_{#1}J}(\vtheta)}
\newcommand{\gradUp}[1]{\overline{\hat{\nabla}_{#1}J}(\vtheta)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Some math symbols used in the text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Multicol Settings
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setlength{\columnsep}{1.5em}
\setlength{\columnseprule}{0mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Save space in lists. Use this after the opening of the list
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\compresslist}{%
\setlength{\itemsep}{1pt}%
\setlength{\parskip}{0pt}%
\setlength{\parsep}{0pt}%
}

\captionsetup{justification=raggedright,singlelinecheck=false}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Begin of Document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Here starts the poster
%%%---------------------------------------------------------------------------
%%% Format it to your taste with the options
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Define some colors

%\definecolor{lightblue}{cmyk}{0.83,0.24,0,0.12}
\definecolor{lightblue}{rgb}{0.145,0.6666,1}

% % Draw a video
% \newlength{\FSZ}
% \newcommand{\drawvideo}[3]{% [0 0.25 0.5 0.75 1 1.25 1.5]
%    \noindent\pgfmathsetlength{\FSZ}{\linewidth/#2}
%    \begin{tikzpicture}[outer sep=0pt,inner sep=0pt,x=\FSZ,y=\FSZ]
%    \draw[color=lightblue!50!black] (0,0) node[outer sep=0pt,inner sep=0pt,text width=\linewidth,minimum height=0] (video) {\noindent#3};
%    \path [fill=lightblue!50!black,line width=0pt] 
%      (video.north west) rectangle ([yshift=\FSZ] video.north east) 
%     \foreach \x in {1,2,...,#2} {
%       {[rounded corners=0.6] ($(video.north west)+(-0.7,0.8)+(\x,0)$) rectangle +(0.4,-0.6)}
%     }
% ;
%    \path [fill=lightblue!50!black,line width=0pt] 
%      ([yshift=-1\FSZ] video.south west) rectangle (video.south east) 
%     \foreach \x in {1,2,...,#2} {
%       {[rounded corners=0.6] ($(video.south west)+(-0.7,-0.2)+(\x,0)$) rectangle +(0.4,-0.6)}
%     }
% ;
%    \foreach \x in {1,...,#1} {
%      \draw[color=lightblue!50!black] ([xshift=\x\linewidth/#1] video.north west) -- ([xshift=\x\linewidth/#1] video.south west);
%    }
%    \foreach \x in {0,#1} {
%      \draw[color=lightblue!50!black] ([xshift=\x\linewidth/#1,yshift=1\FSZ] video.north west) -- ([xshift=\x\linewidth/#1,yshift=-1\FSZ] video.south west);
%    }
%    \end{tikzpicture}
% }
% 
% \hyphenation{resolution occlusions}
% %%
\begin{poster}%
  % Poster Options
  {
  % Show grid to help with alignment
  columns=6,
  grid=false,
  % Column spacing
  colspacing=1em,
  % Color style
  bgColorOne=white,
  bgColorTwo=white,
  borderColor=blue900,
  headerColorOne=blue800,
  headerColorTwo=blue800,
  headerFontColor=white,
  boxColorOne=white,
  boxColorTwo=lightblue,
  % Format of textbox
  textborder=roundedleft,
  % Format of text header
  eyecatcher=true,
  headerborder=closed,
  headerheight=0.095\textheight,
%  textfont=\sc, An example of changing the text font
  headershape=roundedright,
  headershade=shadelr,
  headerfont=\large\bf\textsc, %Sans Serif
  textfont={\setlength{\parindent}{1.5em}},
  boxshade=plain,
%  background=shade-tb,
  background=plain,
  linewidth=2pt
  }
  % Eye Catcher
  % {\includegraphics[height=9em]{./pics/airlab_logo_reflect.png}} 
  {\includegraphics[height=9.0em]{./pics/polilogo/logoPoliBlue_poster.png}}
%   {\hspace{3.5cm}}
  % Title
  {\bf\textsc{Stochastic Variance-Reduced Policy Gradient}\vspace{0.1em}}
  % Authors
  {\textsc{M. Papini, D. Binaghi, G. Canonaco, M. Pirotta and M. Restelli}\\ 
  {\normalsize \texttt{matteo.papini@polimi.it}}
  %{\normalsize \texttt{\{matteo.papini, marcello.restelli\}@polimi.it}},
  %{\normalsize \texttt{\{damiano.binaghi, giuseppe.canonaco\}@mail.polimi.it}}  %{\normalsize \texttt{\{matteo.pirotta\}@inria.fr}}
  }
  % University logo
  {% The makebox allows the title to flow into the logo, this is a hack because of the L shaped logo.
    %\includegraphics[height=9.0em]{./pics/PoliMI.pdf}%\hspace{.5cm}
    \includegraphics[height=4em]{./pics/inria_sc}
  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Now define the boxes that make up the poster
%%%---------------------------------------------------------------------------
%%% Each box has a name and can be placed absolutely or relatively.
%%% The only inconvenience is that you can only specify a relative position 
%%% towards an already declared box. So if you have a box attached to the 
%%% bottom, one to the top and a third one which should be in between, you 
%%% have to specify the top and bottom boxes before you specify the middle 
%%% box.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %
    % A coloured circle useful as a bullet with an adjustably strong filling
    \newcommand{\colouredcircle}{%
      \tikz{\useasboundingbox (-0.2em,-0.32em) rectangle(0.2em,0.32em); \draw[draw=black,fill=lightblue,line width=0.03em] (0,0) circle(0.18em);}}

\newcommand{\HL}[1]{\textcolor{blue}{\textbf{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Problem}{name=problem,column=0,row=0,span=3}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Contributions}{name=contributions,column=3,row=0, span=3,bottomaligned=problem}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{SVRPG}{name=svrpg,column=0,span=6,row=1,below=problem}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

}

\headerbox{Convergence rate}{name=convergence,column=0,span=3,row=2,below=svrpg}{

}

\headerbox{Heuristics}{name=heuristics,column=3,span=3,row=2,below=svrpg}{

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Empirical Results}{name=emp,column=0,row=3,span=6,below=heuristics,above=bottom}{

}

\end{poster}

\end{document}

### 2018-06-25 poster/poster.tex
bozza poster

@@
 \PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-\documentclass[portrait,a0paper,fontscale=0.312]{baposter}
+\documentclass[landscape,a0paper,fontscale=0.312]{baposter}
 
@@
 
+%Specific
+\newcommand{\Ets}[2][t]{\mathbb{E}_{#1\vert s}\left[#2\right]}
+\newcommand{\Es}[1]{\mathbb{E}_{s}\left[#1\right]}
+\newcommand{\Covts}[3][t]{{\mathbb{C}\text{ov}}_{#1\vert s}\left(#2,#3\right)}
+\newcommand{\Covs}[2]{{\mathbb{C}\text{ov}}_{s}\left(#1,#2\right)}
+\newcommand{\Varts}[2][t]{{\mathbb{V}\text{ar}}_{#1\vert s}\left[#2\right]}
+\newcommand{\Vars}[1]{{\mathbb{V}\text{ar}}_{s}\left[#1\right]}
+\newcommand{\gradBlack}[1]{\blacktriangledown J(#1)}
+\newcommand{\gradIdeal}[1]{\dnabla J(#1)}
+\newcommand{\VARRF}{V}
+\newcommand{\GRADLOG}{G}
+\newcommand{\VARIS}{W}
+\newcommand{\HESSLOG}{F}
+% short forms 
+\newcommand{\wt}[1]{\widetilde{#1}}
+\newcommand{\wh}[1]{\widehat{#1}}
+\newcommand{\wo}[1]{\overline{#1}}
+\newcommand{\wb}[1]{\overline{#1}}
+
+%Colors
+\definecolor{crimson}{rgb}{0.86, 0.08, 0.24}
+\definecolor{myblue}{rgb}{0.0, 0.44, 1.0}
+\definecolor{amber2}{rgb}{1.0, 0.49, 0.0}
+\newcommand{\chall}[1]{\textcolor{amber2}{\textbf{#1}}}
+\newcommand{\bad}[1]{\textcolor{crimson}{\textbf{#1}}}
+\newcommand{\enhance}[1]{\textcolor{myblue}{\textbf{#1}}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   % Show grid to help with alignment
-  columns=6,
+  columns=4,
   grid=false,
@@
   % {\includegraphics[height=9em]{./pics/airlab_logo_reflect.png}} 
-  {\includegraphics[height=9.0em]{./pics/polilogo/logoPoliBlue_poster.png}}
+  {\includegraphics[height=7.0em]{./pics/polilogo/logoPoliBlue_poster.png}}
 %   {\hspace{3.5cm}}
@@
   {\textsc{M. Papini, D. Binaghi, G. Canonaco, M. Pirotta and M. Restelli}\\ 
-  {\normalsize \texttt{matteo.papini@polimi.it}}
-  %{\normalsize \texttt{\{matteo.papini, marcello.restelli\}@polimi.it}},
-  %{\normalsize \texttt{\{damiano.binaghi, giuseppe.canonaco\}@mail.polimi.it}}  %{\normalsize \texttt{\{matteo.pirotta\}@inria.fr}}
+  {\normalsize \texttt{\{matteo.papini, marcello.restelli\}@polimi.it}},
+  {\normalsize \texttt{\{damiano.binaghi, giuseppe.canonaco\}@mail.polimi.it}},  {\normalsize \texttt{\{matteo.pirotta\}@inria.fr}}
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{Problem}{name=problem,column=0,row=0,span=3}{
+  \headerbox{Motivation}{name=motiv,column=0,row=0,span=1}{
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\noindent We want to solve \textbf{continuous} Markov Decision Processes (MDPs), \eg robot locomotion
 
+	\vspace{2mm}
+
+	\noindent\textbf{Policy Gradient} (PG): optimize a parametric policy $\pi_{\vtheta}$ via \textbf{gradient ascent} on performance $J(\vtheta)$.
+
+	\vspace{2mm}
+
+	\noindent Two main strategies for gradient computation:
+
+	\begin{itemize}
+	\item \textbf{Full Gradient (FG)}: \bad{sample infefficient} \\$\longrightarrow O(\nicefrac{N}{\epsilon})$
+	\item \textbf{Stochastic Gradient (SG)}: \bad{slow convergence} \\$\longrightarrow O(\nicefrac{1}{\epsilon^2})$
+	\end{itemize}
+
+	\noindent Sample efficiency is crucial in Reinforcement Learning (RL), where collecting samples is extremely expensive $\implies$ \textbf{FG} often unfeasible.
+
+	\vspace{2mm}
+
+	\noindent Slower convergence of \textbf{SG} due to \bad{gradient variance}.
+	\\A solution from the Supervised Learning (SL) literature (finite-sum optimization):
+	\begin{itemize}
+		\item \textbf{Stochastic Variance-Reduced Gradient (SVRG)} $\longrightarrow O(N+\nicefrac{N^{\nicefrac{2}{3}}}{\epsilon})$
+	\end{itemize}
 }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{Contributions}{name=contributions,column=3,row=0, span=3,bottomaligned=problem}{
+\headerbox{Problem}{name=problem,column=0,row=1, span=1,below=motiv}{
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\noindent Can we apply \textbf{SVRG} to policy optimization?\\
+	Compared to SL, RL has three additional challenges:
+	\begin{enumerate}
+		\item \chall{Non-concavity}: policy performance $J(\vtheta)$ is typically a non-concave objective
+		\item \chall{Infinite dataset}: expectation over all possible trajectories cannot be expressed as a finite sum  
+		\item \chall{Non-stationarity}: the data-generating distribution changes as we learn
+	\end{enumerate}
+}
 
+\headerbox{Related Work}{name=related,column=0,row=1, span=1,below=problem,above=bottom}{
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\noindent From the \textbf{SL} literature, \textit{separate} study of: 
+	\begin{itemize}
+		\item \chall{Non-concavity}~\citep{du2017svrgpe,reddi2016stochastic}
+		\item \chall{Infinite dataset}~\citep{harikandeh2015stopwasting,bietti2017stochastic}
+	\end{itemize}
+
+	\noindent From the \textbf{RL} literature:
+	\begin{itemize}
+		\item \textbf{SVRPGE}~\citep{du2017svrgpe}: SVRG for policy \textit{evaluation}
+		\item \textbf{SVRPO}~\cite{xu2017svrgtrpo}: direct application of SVRG to TRPO
+	\end{itemize}
+
+\chall{Nonstationarity} never addressed explicitly!
 }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{SVRPG}{name=svrpg,column=0,span=6,row=1,below=problem}{
+\headerbox{Contributions}{name=contrib,column=1,row=0, span=2}{
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-}
-
-\headerbox{Convergence rate}{name=convergence,column=0,span=3,row=2,below=svrpg}{
-
-}
-
-\headerbox{Heuristics}{name=heuristics,column=3,span=3,row=2,below=svrpg}{
-
+	\begin{itemize}
+		\item We design \enhance{SVRPG}, a variant of SVRG for the PG framework that explicitly addresses the \chall{challenges} of RL.
+		\item We study the \textbf{convergence} properties of SVRPG
+		\item  We devise \textbf{heuristics} for meta-parameter selection in order to develop a practical algorithm.
+		\item We evaluate SVRPG on simulated \textbf{continuous control tasks}. 
+	\end{itemize}
 }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\headerbox{Empirical Results}{name=emp,column=0,row=3,span=6,below=heuristics,above=bottom}{
+  \headerbox{SVRPG}{name=svrpg,column=1,span=2,row=1,below=contrib}{
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{itemize}
+	\item Take any \textbf{unbiased} policy gradient estimator $g(\tau|\vtheta)$ (\eg REINFORCE)
+	\item \textbf{SVRG} idea: combine frequent, inaccurate \textbf{SG} estimations with rare, accurate \textbf{FG} computations.
+	\item Manage \chall{non-concavity} with \textbf{smooth} policies (\eg Gaussian)
+	\item Approximate \chall{infinite dataset} with a large \textbf{batch size} $N$
+	\item Correct \chall{nonstationarity} with \textbf{importance weighting}
+	\item At each \textbf{iteration}, update policy parameter as $\vtheta\gets\vtheta+\alpha\gradBlack{\vtheta}$ with the following \enhance{SVRPG estimator}:	
+\end{itemize}
+%
+\vspace{1cm}
+%
+\begin{center}
+	\tikz[overlay]{
+		\node[draw=myblue, inner sep=20pt, ultra thick] {
+			\Large{$
+			\blacktriangledown J(\vtheta) = \wh{\nabla}_N J(\wt{\vtheta}) + \wh{\nabla}_B J({\vtheta}) - \omega(\vtheta, \wt{\vtheta}) \wh{\nabla}_B J({\wt{\vtheta}})
+			$}
+		};
+	}
+\end{center} 
+%
+\vspace{1.5cm}
+\fbox{
+\begin{minipage}[t]{.3\textwidth}
+\begin{center}
+	\textbf{FG term}
 
+	\begin{equation*}
+	\frac{1}{N}\sum_{i=1}^{N} g(\tau_i|\wt{\vtheta})
+	\end{equation*}\\
+
+	\begin{itemize}
+		\item Trajectories sampled with \textbf{snapshot} policy $\pi_{\wt{\vtheta}}$
+		\item Accurate FG approximation with $N$ samples \\$\longrightarrow$ \bad{expensive!}
+		\item Updated once every $m$ iterations (one \textbf{epoch})
+		to save sample efficiency
+	\end{itemize}
+\end{center}
+\end{minipage}
+}
+%
+\hfill\fbox{
+\begin{minipage}[t]{.3\textwidth}
+	\begin{center}
+		\textbf{SG term}
+
+		\begin{equation*}
+		\frac{1}{B}\sum_{i=1}^{B} g(\tau_i|\vtheta)
+		\end{equation*}
+
+		\begin{itemize}
+			\item Trajectories sampled with current policy $\pi_{\vtheta}$ (\textbf{online})
+			\item Gradient estimation with $B<<N$ samples\\
+			$\longrightarrow$ \bad{high variance!}
+			\item Stabilized by the low-variance FG term
+		\end{itemize}
+	\end{center}
+\end{minipage}
+}
+%
+\hfill\fbox{
+\begin{minipage}[t]{.25\textwidth}
+	\begin{center}
+			\textbf{Correction term}
+
+			\begin{equation*}
+			\frac{1}{B}\sum_{i=1}^{B}\underbrace{\omega(\tau_i|\vtheta,\wt{\vtheta})}_{\frac{p(\tau_i|\wt{\vtheta})}{p(\tau_i|\vtheta)}} g(\tau_i|\vtheta)
+			\end{equation*}
+
+			\begin{itemize}
+				\item Same trajectories as the SG term
+				\item But evaluated in the snapshot parameter!
+				\item \textbf{Importance weighting}
+		\end{itemize}
+	\end{center}
+\end{minipage}
+}
+%
+}
+
+\headerbox{Convergence}{name=convergence,column=1,span=2,row=2, below=svrpg, above=bottom}{
+\begin{align*}
+&\EVV[]
+{\norm[2]{\nabla J(\vtheta_A)}^2} 
+\leq
+\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
+\frac{\zeta}{N}
++\frac{\xi}{B},
+\end{align*}
+
+Serve piu' spazio, ma l'organizzazione del box sopra e' tutta da vedere. Mancano anche le due basic properties
+}
+
+\headerbox{Heuristics}{name=heuristics,column=3,span=1,row=0,height=0.5,height=0.32}{
+	\begin{itemize}
+		\item Adaptive step size
+		\item Adaptive epoch size
+		\item Normalized Importance Weighting
+		\item Critic (baseline)
+	\end{itemize}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\headerbox{Empirical Results}{name=emp,column=3,row=1,span=1,below=heuristics,height=0.45}{
+	Direi che ci stanno 2 grafici, ma devo pensare a quali mettere
+
+	Eventualmente saltano Related Work e le References
+}
+
+\headerbox{References}{name=ref,column=3,row=1,span=1,below=emp,above=bottom}{
+        \tiny
+\bibliographystyle{plainnat}
+\begingroup
+\renewcommand{\section}[2]{}%
+\bibliography{../svrpg.bib}
+\endgroup
 }

### 2018-06-26 poster/poster.tex
Updated

@@
 \PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-\documentclass[landscape,a0paper,fontscale=0.312]{baposter}
+\documentclass[landscape,a0paper,fontscale=0.32]{baposter}
 
@@
 \newcommand{\dbacks}[1]{\textbf{\textcolor{red!80!black}{{#1}}}}
+
+\colorlet{redp}{red!20} % vert S
+\colorlet{greenp}{vert3!20} % vert S\usepackage[framemethod=tikz]{mdframed}
+\colorlet{bluep}{blue!20} % vert S
+\colorlet{blueg}{blue!70!green!90!white}
+
+\newcommand{\hl}[3][\fboxsep1pt]{{#1\colorbox{#2}{#3}}}%
+%\newcommand{\hl}[2]{{\adjustbox{bgcolor=#1}{#2}}}%
+
+\newcommand{\hlr}[1]{\hl{redp}{#1}}
+\newcommand{\hlg}[1]{\hl{greenp}{#1}}
+\newcommand{\hlb}[1]{\hl{bluep}{#1}}
+\newcommand{\hlo}[1]{\hl{orange!50}{#1}}
+\newcommand{\hly}[1]{\hl{yellow!70!black!50!white}{#1}}
+\newcommand{\hlblg}[1]{\hl{blue!75!green!20!white}{#1}}
+\newcommand{\tikzhlblg}[1]{\tikz[baseline]{\node[inner sep=2pt, anchor=base,fill=blue!75!green!20!white]{{#1}};}}
+\newcommand{\tikzhly}[1]{  \tikz[baseline]{\node[inner sep=2pt, anchor=base,fill=yellow!70!black!50!white]{{#1}};}}
+\newcommand{\tikzhlr}[1]{  \tikz[baseline]{\node[inner sep=2pt, anchor=base,fill=redp]{{#1}};}}
+\newcommand{\tikzhlg}[1]{  \tikz[baseline]{\node[inner sep=2pt, anchor=base,fill=greenp]{{#1}};}}
+\newcommand{\tikzhlb}[1]{  \tikz[baseline]{\node[inner sep=2pt, anchor=base,fill=bluep]{{#1}};}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	\noindent We want to solve \textbf{continuous} Markov Decision Processes (MDPs), \eg robot locomotion
+          \noindent $\blacktriangleright$~We want to solve \hlb{continuous} Markov Decision Processes (\hlb{MDPs}), \eg robot locomotion
 
@@
 
-	\noindent\textbf{Policy Gradient} (PG): optimize a parametric policy $\pi_{\vtheta}$ via \textbf{gradient ascent} on performance $J(\vtheta)$.
+    \noindent $\blacktriangleright$~\hlg{\textbf{Policy Gradient} (PG)}: optimize a parametric policy $\pi_{\vtheta}$ via \textbf{gradient ascent} on performance $J(\vtheta)$.
 
@@
 
-	\noindent Two main strategies for gradient computation:
+	\noindent $\blacktriangleright$~Two main strategies for gradient computation:
 
-	\begin{itemize}
+	\begin{itemize}\compresslist
 	\item \textbf{Full Gradient (FG)}: \bad{sample infefficient} \\$\longrightarrow O(\nicefrac{N}{\epsilon})$
@@
 
-	\noindent Sample efficiency is crucial in Reinforcement Learning (RL), where collecting samples is extremely expensive $\implies$ \textbf{FG} often unfeasible.
+    \begin{minipage}{.85\textwidth}
+    \noindent \hly{Sample efficiency} is crucial in Reinforcement Learning (RL), where collecting samples is extremely expensive\\ 
+    \hspace*{.3cm} $\implies$ \textbf{FG} often \underline{unfeasible}
 
 	\vspace{2mm}
+	\noindent Slower convergence of \textbf{SG} due to high \hly{gradient variance}.
+    \end{minipage}
 
-	\noindent Slower convergence of \textbf{SG} due to \bad{gradient variance}.
-	\\A solution from the Supervised Learning (SL) literature (finite-sum optimization):
-	\begin{itemize}
+    \vspace{.3cm}
+    \noindent $\blacktriangleright$~ \hlr{Solution:} from finite-sum optimization in Supervise Learning:
+	\begin{itemize}\compresslist
 		\item \textbf{Stochastic Variance-Reduced Gradient (SVRG)} $\longrightarrow O(N+\nicefrac{N^{\nicefrac{2}{3}}}{\epsilon})$
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\headerbox{Problem}{name=problem,column=0,row=1, span=1,below=motiv}{
+\headerbox{SVRG in RL: challenges}{name=problem,column=0,row=1, span=1,below=motiv}{
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	\noindent Can we apply \textbf{SVRG} to policy optimization?\\
-	Compared to SL, RL has three additional challenges:
+%	\noindent Can we apply \textbf{SVRG} to policy optimization?\\
+        \noindent \hlb{\textbf{Goal:}} replace SG in RL with SVRG\\[.1cm]
+        \hspace*{2cm}\underline{RL introduces three challenges}
+%	Compared to SL, RL has three additional challenges:
 	\begin{enumerate}
@@
 		\item \chall{Infinite dataset}: expectation over all possible trajectories cannot be expressed as a finite sum  
-		\item \chall{Non-stationarity}: the data-generating distribution changes as we learn
+        \item \chall{Non-stationarity}: the data-generating distribution changes as we learn (policy changes)
 	\end{enumerate}
@@
 
-\headerbox{Related Work}{name=related,column=0,row=1, span=1,below=problem,above=bottom}{
+\headerbox{What is out there?}{name=related,column=0,row=1, span=1,below=problem,above=bottom}{
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	\noindent From the \textbf{SL} literature, \textit{separate} study of: 
-	\begin{itemize}
+        \noindent From the \hlb{\textbf{SL}} literature, \textit{separate} study of: 
+	\begin{itemize}\compresslist
 		\item \chall{Non-concavity}~\citep{du2017svrgpe,reddi2016stochastic}
@@
 
-	\noindent From the \textbf{RL} literature:
-	\begin{itemize}
+    \noindent From the \hlb{\textbf{RL}} literature:
+	\begin{itemize}\compresslist
 		\item \textbf{SVRPGE}~\citep{du2017svrgpe}: SVRG for policy \textit{evaluation}
-		\item \textbf{SVRPO}~\cite{xu2017svrgtrpo}: direct application of SVRG to TRPO
+		\item \textbf{SVRPO}~\citep{xu2017svrgtrpo}: direct application of SVRG to TRPO
 	\end{itemize}
 
-\chall{Nonstationarity} never addressed explicitly!
+\chall{Non-stationarity} never addressed explicitly!
 }
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+        {\color{red} Questa possiamo magari toglierla}
 	\begin{itemize}

### 2018-06-26 poster/poster.tex
fixed central box; removed contributions

--- 2018-06-25 poster/poster.tex
@@
 
+\usepackage{multirow}
+
+\usepackage{tabu}
+
 \usepackage{xspace}
@@
 \newcommand{\vtheta}{\boldsymbol{\theta}}
-\newcommand{\gradJ}[1]{\nabla_{#1}J_\mu(\vtheta)}
+\newcommand{\gradJ}[1]{\nabla J({#1})}
 \newcommand{\vphi}{\boldsymbol{\phi}}
@@
 \newcommand{\Covs}[2]{{\mathbb{C}\text{ov}}_{s}\left(#1,#2\right)}
+\newcommand{\Var}[1]{{\mathbb{V}\text{ar}}\left[#1\right]}
 \newcommand{\Varts}[2][t]{{\mathbb{V}\text{ar}}_{#1\vert s}\left[#2\right]}
@@
 \newcommand{\enhance}[1]{\textcolor{myblue}{\textbf{#1}}}
+
+%Fake items
+\newcommand{\tabitem}[1]{~~\llap{\textbullet}~~ #1 \hfill\mbox{}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\headerbox{Contributions}{name=contrib,column=1,row=0, span=2}{
-	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	\begin{itemize}
-		\item We design \enhance{SVRPG}, a variant of SVRG for the PG framework that explicitly addresses the \chall{challenges} of RL.
-		\item We study the \textbf{convergence} properties of SVRPG
-		\item  We devise \textbf{heuristics} for meta-parameter selection in order to develop a practical algorithm.
-		\item We evaluate SVRPG on simulated \textbf{continuous control tasks}. 
-	\end{itemize}
-}
+%\headerbox{Contributions}{name=contrib,column=1,row=0, span=2}{
+%	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%	\begin{itemize}
+%		\item We design \enhance{SVRPG}, a variant of SVRG for the PG framework that explicitly addresses the \chall{challenges} of RL.
+%		\item We study the \textbf{convergence} properties of SVRPG
+%		\item  We devise \textbf{heuristics} for meta-parameter selection in order to develop a practical algorithm.
+%		\item We evaluate SVRPG on simulated \textbf{continuous control tasks}. 
+%	\end{itemize}
+%}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{SVRPG}{name=svrpg,column=1,span=2,row=1,below=contrib}{
+  \headerbox{SVRPG}{name=svrpg,column=1,span=2,
+  	row=0,%1
+  	%below=contrib
+  }{
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\noindent
+We design an \textbf{SVRG}-like algorithm for the \textbf{PG} framework.
+
+\vspace{0.2cm}
+
+\begin{minipage}[t]{.45\textwidth}
+The easy part:
 \begin{itemize}
 	\item Take any \textbf{unbiased} policy gradient estimator $g(\tau|\vtheta)$ (\eg REINFORCE)
-	\item \textbf{SVRG} idea: combine frequent, inaccurate \textbf{SG} estimations with rare, accurate \textbf{FG} computations.
+	\item \textbf{SVRG} idea: combine frequent \textbf{SG} with rare \textbf{FG} into a corrected gradient estimator $\gradBlack{\vtheta}$
+	\item Two time scales: $1$ \textbf{epoch} $=m$ \textbf{iterations}
+\end{itemize}
+\end{minipage}
+\hfill%
+\begin{minipage}[t]{.45\textwidth}
+Solving the RL challenges:
+\begin{enumerate}
 	\item Manage \chall{non-concavity} with \textbf{smooth} policies (\eg Gaussian)
@@
 	\item Correct \chall{nonstationarity} with \textbf{importance weighting}	
-	\item At each \textbf{iteration}, update policy parameter as $\vtheta\gets\vtheta+\alpha\gradBlack{\vtheta}$ with the following \enhance{SVRPG estimator}:	
-\end{itemize}
-%
-\vspace{1cm}
-%
+\end{enumerate}
+\end{minipage}
+
 \begin{center}
-	\tikz[overlay]{
-		\node[draw=myblue, inner sep=20pt, ultra thick] {
-			\Large{$
-			\blacktriangledown J(\vtheta) = \wh{\nabla}_N J(\wt{\vtheta}) + \wh{\nabla}_B J({\vtheta}) - \omega(\vtheta, \wt{\vtheta}) \wh{\nabla}_B J({\wt{\vtheta}})
-			$}
+	Introducing the \enhance{SVRPG} estimator:
+\end{center}
+
+\noindent\dotfill
+\begin{tabu} to \textwidth{X[c]X[2c]X[c]}
+	\multirow{2}{*}{
+		\shortstack{
+			\textbf{FG term}\\
+			$\frac{1}{N}\sum_{i=1}^{N} g(\tau_i|\wt{\vtheta})$\\
+			$\tau_i\sim\wt{\vtheta}$\\
+			$N$ trajectories\\
+			Accurate but \bad{expensive}\\
+			$\downarrow$\\
+			Updated once per \textbf{epoch}
+		}
+	}	 
+	&\tikz[baseline]{ 
+		\node[draw=myblue, very thick, inner sep=5pt]{\Large{$\wh{\nabla}_N J(\wt{\vtheta}) + \wh{\nabla}_B J({\vtheta}) - \omega(\vtheta, \wt{\vtheta}) \wh{\nabla}_B J({\wt{\vtheta}})$}
 		};
 	}  
-\end{center} 
-%
-\vspace{1.5cm}
-\fbox{
-\begin{minipage}[t]{.3\textwidth}
-\begin{center}
-	\textbf{FG term}
+	\vspace{0.5cm} 
+	& \multirow{2}{*}{
+		\shortstack{
+			\textbf{Correction term}\\
+			$\frac{1}{B}\sum_{i=1}^{B}\underbrace{\omega(\tau_i|\vtheta,\wt{\vtheta})}_{\textbf{imp. weight}}g(\tau_i|\wt{\vtheta})$\\
+			Trajectories from SG term \\
+			Ensures SVRG properties\\
+			\bad{Distribution mismatch}\\
+			$\downarrow$\\
+			\textbf{importance weighting}\\
+			$\downarrow$\\
+			\bad{new source of variance}
+		}
+	} \\
+	& \shortstack{
+		\textbf{SG term}\\
+		$\frac{1}{B}\sum_{i=1}^{B} g(\tau_i|\vtheta)$\\
+		$\tau_i\sim\vtheta$\\
+		$B<<N$ trajectories\\
+		Computed at each iteration\\
+		\\Cheap but \bad{high-variance}\\
+		$\downarrow$\\
+		Stabilized by other terms
+	}
+\end{tabu}
 
+\vspace{0.5cm}
+
+\begin{minipage}[t]{.55\textwidth}
+	Fundamental \textbf{SVRG} properties:
+	\begin{enumerate}
+		\item Unbiasedness: $\EVV[]{\gradBlack{\vtheta}} = \gradJ{\vtheta}$
+		\item Vanishing variance: $\Var{\gradBlack{\vtheta}}\to0$ as $\vtheta\to\vtheta^*$
+	\end{enumerate}
+\vfill
+\end{minipage}
+\hfill%
+\begin{minipage}[t]{.4\textwidth}
+\textbf{Importance weighting}
+
+Use samples from $\vtheta$ as taken with $\wt{\vtheta}$:
+\vspace{-0.2cm}
 \begin{equation*}
-	\frac{1}{N}\sum_{i=1}^{N} g(\tau_i|\wt{\vtheta})
-	\end{equation*}\\
-
-	\begin{itemize}
-		\item Trajectories sampled with \textbf{snapshot} policy $\pi_{\wt{\vtheta}}$
-		\item Accurate FG approximation with $N$ samples \\$\longrightarrow$ \bad{expensive!}
-		\item Updated once every $m$ iterations (one \textbf{epoch})
-		to save sample efficiency
-	\end{itemize}
-\end{center}
+	w(\tau|\vtheta,\wt{\vtheta}) = \frac{p(\tau|\wt\vtheta)}{p(\tau|\vtheta)}
+	\quad
+	\genfrac{}{}{0pt}{}{\leftarrow \textit{target distribution}}
+		{\leftarrow \textit{source distribution}}
+\end{equation*}
 \end{minipage}
-}
-%
-\hfill\fbox{
-\begin{minipage}[t]{.3\textwidth}
-	\begin{center}
-		\textbf{SG term}
-
-		\begin{equation*}
-		\frac{1}{B}\sum_{i=1}^{B} g(\tau_i|\vtheta)
-		\end{equation*}
-
-		\begin{itemize}
-			\item Trajectories sampled with current policy $\pi_{\vtheta}$ (\textbf{online})
-			\item Gradient estimation with $B<<N$ samples\\
-			$\longrightarrow$ \bad{high variance!}
-			\item Stabilized by the low-variance FG term
-		\end{itemize}
-	\end{center}
-\end{minipage}
-}
-%
-\hfill\fbox{
-\begin{minipage}[t]{.25\textwidth}
-	\begin{center}
-			\textbf{Correction term}
-
-			\begin{equation*}
-			\frac{1}{B}\sum_{i=1}^{B}\underbrace{\omega(\tau_i|\vtheta,\wt{\vtheta})}_{\frac{p(\tau_i|\wt{\vtheta})}{p(\tau_i|\vtheta)}} g(\tau_i|\vtheta)
-			\end{equation*}
-
-			\begin{itemize}
-				\item Same trajectories as the SG term
-				\item But evaluated in the snapshot parameter!
-				\item \textbf{Importance weighting}
-		\end{itemize}
-	\end{center}
-\end{minipage}
-}
-%
 }

### 2018-06-27 poster/poster.tex
Updated SVRPG box

@@
 
+% box for algorithms
+\newlength{\minipagewidth}
+\newlength{\minipagewidthx}
+\setlength{\minipagewidth}{0.99\textwidth}
+\setlength{\minipagewidthx}{0.99\textwidth}
+\setlength{\fboxsep}{1.5mm}
+\addtolength{\minipagewidth}{-\fboxrule}
+\addtolength{\minipagewidth}{-\fboxrule}
+\addtolength{\minipagewidth}{-\fboxsep}
+\addtolength{\minipagewidth}{-\fboxsep}
+\addtolength{\minipagewidthx}{+\fboxsep}
+\newcommand{\bookbox}[1]{
+\par\medskip\noindent
+\framebox[\columnwidth]{
+\begin{minipage}{\minipagewidth} {#1} \end{minipage} } \par\medskip }
+
+\newcommand{\bookboxx}[1]{\small
+\par\medskip\noindent
+\framebox[0.99\textwidth]{
+\begin{minipage}{0.97\dimexpr\textwidth-\parindent\relax} {#1} \end{minipage} } \par\medskip }
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   {\normalsize \texttt{\{matteo.papini, marcello.restelli\}@polimi.it}},
-  {\normalsize \texttt{\{damiano.binaghi, giuseppe.canonaco\}@mail.polimi.it}},  {\normalsize \texttt{\{matteo.pirotta\}@inria.fr}}
+  {\normalsize \texttt{\{damiano.binaghi, giuseppe.canonaco\}@mail.polimi.it}},  {\normalsize \texttt{matteo.pirotta@inria.fr}}
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{SVRPG}{name=svrpg,column=1,span=2,
+  \headerbox{SVRPG: stochastic variance-reduced policy gradient}{name=svrpg,column=1,span=2,
   	row=0,%1
@@
 
-\begin{minipage}[t]{.45\textwidth}
-The easy part:
-\begin{itemize}
-	\item Take any \textbf{unbiased} policy gradient estimator $g(\tau|\vtheta)$ (\eg REINFORCE)
-	\item \textbf{SVRG} idea: combine frequent \textbf{SG} with rare \textbf{FG} into a corrected gradient estimator $\gradBlack{\vtheta}$
-	\item Two time scales: $1$ \textbf{epoch} $=m$ \textbf{iterations}
-\end{itemize}
+\noindent\begin{minipage}[t]{.32\textwidth}
+        \hlg{The easy part:}\\[.2cm]
+%\begin{itemize}\compresslist
+        $\bullet$~Take any \textbf{unbiased} policy gradient estimator $g(\tau|\vtheta)$ (\eg REINFORCE)\\[.15cm]
+    $\bullet$~\textbf{SVRG} idea: combine \hly{frequent \textbf{SG}} with \hly{rare \textbf{FG}} into a corrected gradient estimator $\gradBlack{\vtheta}$\\[.15cm]
+    $\bullet$~Two-time scales:\\ \hspace*{.8cm} $1$ \textbf{epoch} $=m$ \textbf{iterations}
+%\end{itemize}
 \end{minipage}
 \hfill%
-\begin{minipage}[t]{.45\textwidth}
-Solving the RL challenges:
-\begin{enumerate}
-	\item Manage \chall{non-concavity} with \textbf{smooth} policies (\eg Gaussian)
-	\item Approximate \chall{infinite dataset} with a large \textbf{batch size} $N$
-	\item Correct \chall{nonstationarity} with \textbf{importance weighting}	
-\end{enumerate}
+\begin{minipage}[t]{.33\textwidth}
+        \hlo{Solving the RL challenges:}\\[.2cm]
+%\begin{enumerate}\compresslist
+    $\bullet$~Manage \chall{non-concavity} with \textbf{smooth} policies (\eg Gaussian)\\[.15cm]
+    $\bullet$~Approximate \chall{infinite dataset} with a large \textbf{batch size} $N$\\[.15cm]
+	$\bullet$~Correct \chall{nonstationarity} with \textbf{importance weighting}	
+%\end{enumerate}
+\end{minipage}
+\hfill%
+\begin{minipage}[t]{.31\textwidth}
+        \vspace*{-.4cm}
+        \hspace*{1.8cm} \emph{Algorithm}
+        \vspace*{-.2cm}
+        \bookboxx{
+                \small
+                \textbf{For} $s = 1,\ldots$\\
+                \hspace*{.2cm} Sample $N$ trajectories using $\theta^s$\\
+                \hspace*{.2cm} Compute FG $\wh{\nabla}_N J(\theta^S)$\\
+                \hspace*{.2cm} \textbf{For} $t = 1, \ldots, m-1$\\
+                \hspace*{.4cm} Sample $B$ trajectories using $\theta^{s+1}_t$\\
+                \hspace*{.4cm} Compute SG $\wh{\nabla}_B J(\theta^{s+1}_t)$ and \\
+                \hspace*{.6cm} $c_t^{s+1} = \omega(\theta^{s+1}_t, \theta^s)\wh{\nabla}_B J(\theta^{s+1}_t)$\\
+                \hspace*{.4cm} Compute $\theta^{s+1}_{t+1}$ using $\gradBlack{\theta^{s+1}_t}$
+        }
 \end{minipage}
@@
 \begin{center}
-	Introducing the \enhance{SVRPG} estimator:
+        Introducing the \enhance{SVRPG} estimator $\gradBlack{\vtheta}$:
 \end{center}
@@
 		\shortstack{
-			\textbf{FG term}\\
+                \tikz[baseline]{\node[anchor=base,fill=greenp, inner sep=2pt](FGtext){$\textbf{FG term}$};}\\
 			$\frac{1}{N}\sum_{i=1}^{N} g(\tau_i|\wt{\vtheta})$\\
@@
 	}	 
-	&\tikz[baseline]{ 
-		\node[draw=myblue, very thick, inner sep=5pt]{\Large{$\wh{\nabla}_N J(\wt{\vtheta}) + \wh{\nabla}_B J({\vtheta}) - \omega(\vtheta, \wt{\vtheta}) \wh{\nabla}_B J({\wt{\vtheta}})$}
+	&\tikz[overlay]{ 
+    \node[inner sep=5pt, xshift=-.2cm]{\large{$\gradBlack{\vtheta} := \tikz[baseline]{\node[anchor=base, fill=greenp](FG){$\wh{\nabla}_N J(\wt{\vtheta})$};} + 
+            \tikz[baseline]{\node[anchor=base, fill=redp](SG){$\wh{\nabla}_B J({\vtheta})$};} - 
+    \tikz[baseline]{\node[anchor=base, fill=yellow](C){$\omega(\vtheta, \wt{\vtheta}) \wh{\nabla}_B J({\wt{\vtheta}})$};}$}
 		};
 	}  
-	\vspace{0.5cm} 
+	\vspace{0.9cm} 
 	& \multirow{2}{*}{
 		\shortstack{
-			\textbf{Correction term}\\
+            \tikz[baseline]{\node[anchor=base, fill=yellow, inner sep=2pt](Ctext){$\textbf{Correction term}$};}\\
 			$\frac{1}{B}\sum_{i=1}^{B}\underbrace{\omega(\tau_i|\vtheta,\wt{\vtheta})}_{\textbf{imp. weight}}g(\tau_i|\wt{\vtheta})$\\
@@
 	& \shortstack{
-		\textbf{SG term}\\
+    \tikz[baseline]{\node[anchor=base, fill=redp, inner sep=2pt](SGtext){$\textbf{SG term}$};}\\
 		$\frac{1}{B}\sum_{i=1}^{B} g(\tau_i|\vtheta)$\\
@@
 \end{tabu}
+\tikz[overlay]{
+        \path[->, greenp, thick] (FG.north west) edge[bend right] (FGtext.east);
+        \path[->, yellow, thick] (C.east) edge[bend left] (Ctext.west);
+        \draw[->, redp, thick] (SG.south) -- (SGtext.north);
+}
 
@@
 	\begin{enumerate}
-		\item Unbiasedness: $\EVV[]{\gradBlack{\vtheta}} = \gradJ{\vtheta}$
-		\item Vanishing variance: $\Var{\gradBlack{\vtheta}}\to0$ as $\vtheta\to\vtheta^*$
+            \item \textcolor{myblue}{\textbf{Unbiasedness:}} $\EVV[]{\gradBlack{\vtheta}} = \gradJ{\vtheta}$
+            \item \textcolor{myblue}{\textbf{Vanishing variance:}} $\Var{\gradBlack{\vtheta}}\to0$ as $\vtheta\to\vtheta^*$
 	\end{enumerate}
@@
 \begin{minipage}[t]{.4\textwidth}
-\textbf{Importance weighting}
+        \hly{\textbf{Importance weighting}}
 

### 2018-06-27 poster/poster.tex
edited central column

@@
 \colorlet{greenp}{vert3!20} % vert S\usepackage[framemethod=tikz]{mdframed}
-\colorlet{bluep}{blue!20} % vert S
+\colorlet{bluep}{blue!12} % vert S
 \colorlet{blueg}{blue!70!green!90!white}
-
+\definecolor{yellowp}{rgb}{0.99, 0.99, 0.59}
+\colorlet{orangep}{orange!40}
 \newcommand{\hl}[3][\fboxsep1pt]{{#1\colorbox{#2}{#3}}}%
@@
 \newcommand{\hlb}[1]{\hl{bluep}{#1}}
-\newcommand{\hlo}[1]{\hl{orange!50}{#1}}
-\newcommand{\hly}[1]{\hl{yellow!70!black!50!white}{#1}}
+\newcommand{\hlo}[1]{\hl{orangep}{#1}}
+\newcommand{\hly}[1]{\hl{yellow!70!black!40!white}{#1}}
 \newcommand{\hlblg}[1]{\hl{blue!75!green!20!white}{#1}}
@@
 
-\usepackage{tcolorbox}
+%\usepackage{tcolorbox}
 
@@
 \usepackage{tabu}
+\usepackage{schemata}
 
@@
 \begin{minipage}{0.97\dimexpr\textwidth-\parindent\relax} {#1} \end{minipage} } \par\medskip }
+
+%Reduce linespace in biblio
+\setlength{\bibsep}{2pt}
+
+\usepackage[many]{tcolorbox}
+\usetikzlibrary{decorations.pathreplacing}
+%\tcbset{%
+%	enhanced,
+%	coltitle=black, 
+%	detach title,
+%	left=10mm,
+%	overlay={
+%		\node[rotate=90, minimum width=1cm, anchor=south,yshift=-0.8cm] at (frame.east) {\tcbtitle};
+%	}
+%}
+\newtcolorbox{rightbrace1}{%
+	boxsep=0cm,
+	left=.05cm,
+	top=.05cm,
+	bottom=.05cm,
+	grow to left by=-.35cm,
+	colback=black!10,
+	enhanced jigsaw, 
+	breakable, % allow page breaks
+	frame hidden, % hide the default frame
+	width=6.1cm,
+	overlay={%
+		\draw [
+		fill=none, % fill paper
+		decoration={brace,amplitude=0.5em},
+		decorate,
+		ultra thick,
+		darkgray,
+		]
+		% right line
+		(frame.north east)--(frame.south east);
+		\node[rotate=270, minimum width=1cm, anchor=south,yshift=0.35cm] at (frame.east) {iteration};
+	},
+	% paragraph skips obeyed within tcolorbox
+	parbox=false,
+}
+%
+\newtcolorbox{rightbrace2}{%
+	boxsep=0cm,
+	left=.05cm,
+	top=.05cm,
+	bottom=.05cm,
+	grow to left by=-.35cm,
+	colback=black!5!white,
+	enhanced jigsaw, 
+	breakable, % allow page breaks
+	frame hidden, % hide the default frame
+	width=7.6cm,
+	overlay={%
+		\draw [
+		fill=none, % fill paper
+		decoration={brace,amplitude=0.5em},
+		decorate,
+		ultra thick,
+		darkgray,
+		]
+		% right line
+		(frame.north east)--(frame.south east);
+		\node[rotate=270, minimum width=1cm, anchor=south,yshift=0.3cm] at (frame.east) {epoch};
+	},
+	% paragraph skips obeyed within tcolorbox
+	parbox=false,
+}
 
@@
     \begin{minipage}{.85\textwidth}
-    \noindent \hly{Sample efficiency} is crucial in Reinforcement Learning (RL), where collecting samples is extremely expensive\\ 
+    \noindent \hlr{Sample efficiency} is crucial in Reinforcement Learning (RL), where collecting samples is extremely expensive\\ 
     \hspace*{.3cm} $\implies$ \textbf{FG} often \underline{unfeasible}
@@
 	\vspace{2mm}
-	\noindent Slower convergence of \textbf{SG} due to high \hly{gradient variance}.
+	\noindent Slower convergence of \textbf{SG} due to high \hlr{gradient variance}.
     \end{minipage}
@@
     \vspace{.3cm}
-    \noindent $\blacktriangleright$~ \hlr{Solution:} from finite-sum optimization in Supervise Learning:
+    \noindent $\blacktriangleright$~ \hly{Solution} from finite-sum optimization in Supervised Learning (SL):
 	\begin{itemize}\compresslist
@@
 	\begin{itemize}\compresslist
-		\item \chall{Non-concavity}~\citep{du2017svrgpe,reddi2016stochastic}
+		\item \chall{Non-concavity}~\citep{allen2016variance,reddi2016stochastic}
 		\item \chall{Infinite dataset}~\citep{harikandeh2015stopwasting,bietti2017stochastic}
@@
 	\begin{itemize}\compresslist
-		\item \textbf{SVRPGE}~\citep{du2017svrgpe}: SVRG for policy \textit{evaluation}
+		\item \textbf{SVRPGE}~\citep{du2017svrgpe}: policy \textit{evaluation}
 		\item \textbf{SVRPO}~\citep{xu2017svrgtrpo}: direct application of SVRG to TRPO
@@
 
-\noindent\begin{minipage}[t]{.32\textwidth}
-        \hlg{The easy part:}\\[.2cm]
+\noindent\begin{minipage}[t]{.48\textwidth}
+\begin{tcolorbox}[left=0cm,right=0cm,top=0cm,bottom=0cm,colback=white, boxsep=.1cm, colframe=yellow!70!black!40!white, grow to left by=0cm, grow to right by=0cm]
+\vspace*{-.11cm}\hspace*{-.11cm}\hlb{The easy part:}\\[.2cm]
 %\begin{itemize}\compresslist
         $\bullet$~Take any \textbf{unbiased} policy gradient estimator $g(\tau|\vtheta)$ (\eg REINFORCE)\\[.15cm]
-    $\bullet$~\textbf{SVRG} idea: combine \hly{frequent \textbf{SG}} with \hly{rare \textbf{FG}} into a corrected gradient estimator $\gradBlack{\vtheta}$\\[.15cm]
-    $\bullet$~Two-time scales:\\ \hspace*{.8cm} $1$ \textbf{epoch} $=m$ \textbf{iterations}
+    $\bullet$~\textbf{SVRG} idea: combine \hlr{frequent \textbf{SG}} with \\\hlg{rare \textbf{FG}} into a corrected gradient estimator $\gradBlack{\vtheta}$\\[.15cm]
+    $\bullet$~Two-time scales: $1$ \textbf{epoch} $=m$ \textbf{iterations}
 %\end{itemize}
-\end{minipage}
-\hfill%
-\begin{minipage}[t]{.33\textwidth}
-        \hlo{Solving the RL challenges:}\\[.2cm]
+\end{tcolorbox}
+\end{minipage}\hfill%
+\begin{minipage}[t]{.48\textwidth}
+\begin{tcolorbox}[left=0cm,right=0cm,top=0cm,bottom=0cm,colback=white, boxsep=.1cm, colframe=orangep, grow to left by=0cm, grow to right by=0cm]
+\vspace*{-.11cm}\hspace*{-.11cm}\hlo{Solving the RL challenges:}\\[.2cm]
 %\begin{enumerate}\compresslist
@@
 %\end{enumerate}
-\end{minipage}
-\hfill%
-\begin{minipage}[t]{.31\textwidth}
-        \vspace*{-.4cm}
-        \hspace*{1.8cm} \emph{Algorithm}
-        \vspace*{-.2cm}
-        \bookboxx{
-                \small
-                \textbf{For} $s = 1,\ldots$\\
-                \hspace*{.2cm} Sample $N$ trajectories using $\theta^s$\\
-                \hspace*{.2cm} Compute FG $\wh{\nabla}_N J(\theta^S)$\\
-                \hspace*{.2cm} \textbf{For} $t = 1, \ldots, m-1$\\
-                \hspace*{.4cm} Sample $B$ trajectories using $\theta^{s+1}_t$\\
-                \hspace*{.4cm} Compute SG $\wh{\nabla}_B J(\theta^{s+1}_t)$ and \\
-                \hspace*{.6cm} $c_t^{s+1} = \omega(\theta^{s+1}_t, \theta^s)\wh{\nabla}_B J(\theta^{s+1}_t)$\\
-                \hspace*{.4cm} Compute $\theta^{s+1}_{t+1}$ using $\gradBlack{\theta^{s+1}_t}$
-        }
+\end{tcolorbox}
 \end{minipage}
@@
             \tikz[baseline]{\node[anchor=base, fill=redp](SG){$\wh{\nabla}_B J({\vtheta})$};} - 
-    \tikz[baseline]{\node[anchor=base, fill=yellow](C){$\omega(\vtheta, \wt{\vtheta}) \wh{\nabla}_B J({\wt{\vtheta}})$};}$}
+    \tikz[baseline]{\node[anchor=base, fill=yellowp](C){$\omega(\vtheta, \wt{\vtheta}) \wh{\nabla}_B J({\wt{\vtheta}})$};}$}
 		};
@@
 		\shortstack{
-            \tikz[baseline]{\node[anchor=base, fill=yellow, inner sep=2pt](Ctext){$\textbf{Correction term}$};}\\
+            \tikz[baseline]{\node[anchor=base, fill=yellowp, inner sep=2pt](Ctext){$\textbf{Correction term}$};}\\
 			$\frac{1}{B}\sum_{i=1}^{B}\underbrace{\omega(\tau_i|\vtheta,\wt{\vtheta})}_{\textbf{imp. weight}}g(\tau_i|\wt{\vtheta})$\\
@@
         \path[->, greenp, thick] (FG.north west) edge[bend right] (FGtext.east);
-        \path[->, yellow, thick] (C.east) edge[bend left] (Ctext.west);
+        \path[->, yellowp, thick] (C.east) edge[bend left] (Ctext.west);
         \draw[->, redp, thick] (SG.south) -- (SGtext.north);
@@
 
-\vspace{0.5cm}
+\vspace{0.2cm}
+\noindent
+\begin{minipage}[t]{.48\textwidth}
+	\begin{tcolorbox}[left=0cm,right=0cm,top=0cm,bottom=0cm,colback=white, boxsep=.1cm, colframe=bluep, grow to left by=0cm, grow to right by=0cm]
+		\vspace*{-.11cm}\hspace*{-.2cm}
+		\hlb{Fundamental SVRG properties:}
 
-\begin{minipage}[t]{.55\textwidth}
-	Fundamental \textbf{SVRG} properties:
 	\begin{enumerate}
 	\item \textcolor{myblue}{\textbf{Unbiasedness:}} $\EVV[]{\gradBlack{\vtheta}} = \gradJ{\vtheta}$
-            \item \textcolor{myblue}{\textbf{Vanishing variance:}} $\Var{\gradBlack{\vtheta}}\to0$ as $\vtheta\to\vtheta^*$
+	\item \textcolor{myblue}{\textbf{Vanishing variance:}} $\Var{\gradBlack{\vtheta}}\to0\:\:$ as $\:\:\vtheta\to\vtheta^*$
 	\end{enumerate}
-\vfill
+	\end{tcolorbox}
 \end{minipage}
 \hfill%
-\begin{minipage}[t]{.4\textwidth}
-        \hly{\textbf{Importance weighting}}
+\begin{minipage}[t]{.48\textwidth}
+\begin{tcolorbox}[left=0cm,right=0cm,top=0cm,bottom=0cm,colback=white, boxsep=.1cm, colframe=yellow!70!black!40!white, grow to left by=0cm, grow to right by=0cm]
+\vspace*{-.11cm}\hspace*{-.2cm}
+\hly{Importance weighting}
 
@@
 \end{equation*}
+\end{tcolorbox}
 \end{minipage}
@@
 
-\headerbox{Convergence}{name=convergence,column=1,span=2,row=2, below=svrpg, above=bottom}{
+\headerbox{Convergence}{name=convergence,column=2,span=1,row=2, below=svrpg, above=bottom}{
+\noindent
+\chall{Non-concavity }$\longrightarrow$\textit{ local} maximum
+
+\noindent\hlb{Linear} rate under increasing batch sizes:
 \begin{align*}
 &\EVV[]
-{\norm[2]{\nabla J(\vtheta_A)}^2} 
-\leq
-\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
-\frac{\zeta}{N}
-+\frac{\xi}{B},
+{\norm[]{\nabla J(\vtheta)}^2} \leq
+\tikz[baseline]{\node[anchor=base, fill=bluep,inner sep=5](Nterm){$\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T}$};}
++
+\tikz[baseline]{\node[anchor=base, fill=orangep,inner sep=5](Nterm){$\frac{\zeta}{N}$};} + 
+\tikz[baseline]{\node[anchor=base, fill=orangep,inner sep=5](Bterm){$\frac{\xi}{B}$};}
 \end{align*}
 
-Serve piu' spazio, ma l'organizzazione del box sopra e' tutta da vedere. Mancano anche le due basic properties
+\vspace*{1cm}
+
+\noindent
+\begin{tabular}{cc}
+\shortstack{ \tikz[baseline]{\node[anchor=base, inner sep=2](Ntext){\chall{Infinite-dataset} error};}\\ 
+from FG approximation\\
+$\downarrow$ 
+\\take \textit{large} $N$
+}&
+\hfill%
+\shortstack{
+\vspace*{-.1cm}
+\tikz[baseline]{\node[anchor=base, inner sep=2](Btext){\chall{Non-stationarity} error};}\\ 
+from importance weighting\\
+	$\downarrow$ 
+	\\take \textit{large} $B$ (still $<<N$)
+}
+\end{tabular}
+
+\tikz[overlay]{
+	\draw[orangep, thick,->] plot[smooth,tension=.3] coordinates{(Nterm.south) ($(Nterm.south) + (0,-.5cm)$) ($(Ntext.north) + (0,.5cm)$)
+	 (Ntext.north)};
+	\draw[orangep, thick,->] plot[smooth,tension=.3] coordinates{(Bterm.south) ($(Bterm.south) + (0,-.7cm)$) ($(Btext.north) + (0,.5cm)$)
+		(Btext.north)};
 }
 
-\headerbox{Heuristics}{name=heuristics,column=3,span=1,row=0,height=0.5,height=0.32}{
+\begin{itemize}
+\item Constants ($\psi,\zeta,\xi$) depend only on \hlg{step size $\alpha$} and \hlg{epoch size $m$}
+\item Theory dictates \hlr{very conservative} meta-parameters ($\alpha,m,N,B)$  to guarantee convergence
+\end{itemize}
+}
+
+\headerbox{Algorithm}{name=algo,column=1,span=1,row=1,below=svrpg}{
+\noindent
+\small
+%\begin{minipage}[t]{.7\textwidth}
+\textbf{For} $s = 1,\ldots$
+\vspace*{-.2cm}
+\begin{rightbrace2}
+	Sample $N$ trajectories using $\wt{\theta}$\\
+	Compute FG $=\wh{\nabla}_N J(\wt{\theta})$\\
+	\textbf{For} $t = 1, \ldots, m$
+	\vspace*{-.15cm}
+	\begin{rightbrace1}
+	Sample $B$ trajectories using $\theta$\\
+	Compute SG $=\wh{\nabla}_B J(\theta)$\\
+	Compute correction $=\omega(\theta, \wt{\theta})\wh{\nabla}_B J(\wt{\theta})$\\
+	Update $\theta\gets\theta+\alpha\gradBlack{\theta}$
+	\end{rightbrace1}
+	\vspace*{-.15cm}
+	\hspace*{-.5cm}Update $\wt{\theta}\gets\theta$
+\end{rightbrace2}
+\vspace*{-.2cm}
+}
+
+\headerbox{Heuristics}{name=heuristics,column=3,span=1,row=0}{
 	\begin{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\headerbox{Empirical Results}{name=emp,column=3,row=1,span=1,below=heuristics,height=0.45}{
+\headerbox{Empirical Results}{name=emp,column=3,row=1,span=1,below=heuristics}{
 	Direi che ci stanno 2 grafici, ma devo pensare a quali mettere
@@
 
-\headerbox{References}{name=ref,column=3,row=1,span=1,below=emp,above=bottom}{
+\headerbox{References}{name=ref,column=1,row=2,span=1,below=algo,above=bottom}{
         \tiny
@@
 \renewcommand{\section}[2]{}%
-\bibliography{../svrpg.bib}
+\bibliography{poster.bib}
 \endgroup

### 2018-06-27 poster/poster.tex
first complete draft

@@
 \usepackage{tabu}
-\usepackage{schemata}
 
@@
 	parbox=false,
+}
+
+\usepackage{stackengine}
+\usepackage{scalerel}
+\newcommand\dangersign[1][2ex]{%
+	\renewcommand\stacktype{L}%
+	\scaleto{\stackon[1.5pt]{\color{crimson}$\triangle$}{\tiny \bfseries{!}}}{#1}%
 }
@@
 \noindent\begin{minipage}[t]{.48\textwidth}
-\begin{tcolorbox}[left=0cm,right=0cm,top=0cm,bottom=0cm,colback=white, boxsep=.1cm, colframe=yellow!70!black!40!white, grow to left by=0cm, grow to right by=0cm]
+\begin{tcolorbox}[left=0cm,right=0cm,top=0cm,bottom=0cm,colback=white, boxsep=.1cm, colframe=bluep, grow to left by=0cm, grow to right by=0cm]
 \vspace*{-.11cm}\hspace*{-.11cm}\hlb{The easy part:}\\[.2cm]
@@
 \headerbox{Heuristics}{name=heuristics,column=3,span=1,row=0}{
+\noindent
+Heuristics can improve performance in practice
 	\begin{itemize}
-		\item Adaptive step size
-		\item Adaptive epoch size
-		\item Normalized Importance Weighting
-		\item Critic (baseline)
+		\item \hlb{Step size:} \textbf{ADAM} to counteract gradient variance\\
+		\dangersign \hspace*{.05cm} FG and SG variance have completely different magnitude $\rightarrow$ use two separate annealing schedules $\alpha_{FG}$ and $\alpha_{SG}$
+		\item \hlb{Epoch size:} cut epoch when the effective step size becomes too small
+		\begin{equation*}
+			\frac{\alpha_{FG}}{N} > \frac{\alpha_{SG}}{B} \implies \textbf{new epoch}
+		\end{equation*}
+		\item \hlb{Normalized importance weights:} reduce variance at the price of introducing a small bias
+		\item \hlb{Critic:} an orthogonal variance reduction technique from the \textbf{PG} literature
 	\end{itemize}
@@
 \headerbox{Empirical Results}{name=emp,column=3,row=1,span=1,below=heuristics}{
-	Direi che ci stanno 2 grafici, ma devo pensare a quali mettere
 
-	Eventualmente saltano Related Work e le References
+\noindent
+\begin{tikzpicture}
+	\node[inner sep=0pt] (picture1) {\includegraphics[width=.98\textwidth]{images/cartpole.pdf}};
+	\node[above left] at ($(picture1.north west) + (2.7cm,-.6cm)$) {\textbf{Cart-Pole}};
+\end{tikzpicture}%
+
+\noindent
+\begin{tikzpicture}
+\node[inner sep=0pt] (picture2) {\includegraphics[width=0.98\textwidth]{images/swimmer.pdf}};
+\node[above left] at ($(picture2.north west) + (2.8cm,-.5cm)$) {\textbf{Swimmer}};
+\end{tikzpicture}%
+
+\vspace*{-.3cm}
+
+\noindent
+\begin{tikzpicture}
+\node[inner sep=0pt] (picture3) {\includegraphics[width=0.98\textwidth]{images/cheetah.pdf}};
+\node[above left] at ($(picture3.north west) + (3.4cm,-.8cm)$) {\textbf{Half-Cheetah}};
+\end{tikzpicture}%
 }

### 2018-06-27 poster/poster.tex
changed some colors

@@
 \colorlet{blueg}{blue!70!green!90!white}
-\definecolor{yellowp}{rgb}{0.99, 0.99, 0.59}
+\definecolor{yellowp}{rgb}{0.93, 0.93, 0.42}
 \colorlet{orangep}{orange!40}
@@
 \newcommand{\hlo}[1]{\hl{orangep}{#1}}
-\newcommand{\hly}[1]{\hl{yellow!70!black!40!white}{#1}}
+\newcommand{\hly}[1]{\hl{yellowp}{#1}}
 \newcommand{\hlblg}[1]{\hl{blue!75!green!20!white}{#1}}
@@
 %Reduce linespace in biblio
-\setlength{\bibsep}{2pt}
+\setlength{\bibsep}{1.5pt}
 
@@
         $\bullet$~Take any \textbf{unbiased} policy gradient estimator $g(\tau|\vtheta)$ (\eg REINFORCE)\\[.15cm]
-    $\bullet$~\textbf{SVRG} idea: combine \hlr{frequent \textbf{SG}} with \\\hlg{rare \textbf{FG}} into a corrected gradient estimator $\gradBlack{\vtheta}$\\[.15cm]
+    $\bullet$~\textbf{SVRG} idea: combine \hlr{frequent \textbf{SG}} with \\\hlg{rare \textbf{FG}} into a \hly{corrected} gradient estimator $\gradBlack{\vtheta}$\\[.15cm]
     $\bullet$~Two-time scales: $1$ \textbf{epoch} $=m$ \textbf{iterations}
@@
 \begin{center}
-        Introducing the \enhance{SVRPG} estimator $\gradBlack{\vtheta}$:
+        Introducing the \large{\enhance{SVRPG}} estimator $\gradBlack{\vtheta}$:
 \end{center}
@@
 	\begin{enumerate}
-	\item \textcolor{myblue}{\textbf{Unbiasedness:}} $\EVV[]{\gradBlack{\vtheta}} = \gradJ{\vtheta}$
-	\item \textcolor{myblue}{\textbf{Vanishing variance:}} $\Var{\gradBlack{\vtheta}}\to0\:\:$ as $\:\:\vtheta\to\vtheta^*$
+	\item \textbf{Unbiasedness:} $\EVV[]{\gradBlack{\vtheta}} = \gradJ{\vtheta}$
+	\item \textbf{Vanishing variance:} $\Var{\gradBlack{\vtheta}}\to0\:\:$ as $\:\:\vtheta\to\vtheta^*$
 	\end{enumerate}
@@
 \begin{minipage}[t]{.48\textwidth}
-\begin{tcolorbox}[left=0cm,right=0cm,top=0cm,bottom=0cm,colback=white, boxsep=.1cm, colframe=yellow!70!black!40!white, grow to left by=0cm, grow to right by=0cm]
+\begin{tcolorbox}[left=0cm,right=0cm,top=0cm,bottom=0cm,colback=white, boxsep=.1cm, colframe=yellowp, grow to left by=0cm, grow to right by=0cm]
 \vspace*{-.11cm}\hspace*{-.2cm}
@@
 \begin{itemize}
-\item Constants ($\psi,\zeta,\xi$) depend only on \hlg{step size $\alpha$} and \hlg{epoch size $m$}
+\item Constants ($\psi,\zeta,\xi$) depend only on \hlg{step size $\alpha$} and \hly{epoch size $m$}
 \item Theory dictates \hlr{very conservative} meta-parameters ($\alpha,m,N,B)$  to guarantee convergence
@@
 	\begin{itemize}
-		\item \hlb{Step size:} \textbf{ADAM} to counteract gradient variance\\
+		\item \hlg{Step size:} \textbf{ADAM} to counteract gradient variance\\
 		\dangersign \hspace*{.05cm} FG and SG variance have completely different magnitude $\rightarrow$ use two separate annealing schedules $\alpha_{FG}$ and $\alpha_{SG}$
-		\item \hlb{Epoch size:} cut epoch when the effective step size becomes too small
+		\item \hly{Epoch size:} cut epoch when the effective step size becomes too small
 		\begin{equation*}
@@
 		\item \hlb{Normalized importance weights:} reduce variance at the price of introducing a small bias
-		\item \hlb{Critic:} an orthogonal variance reduction technique from the \textbf{PG} literature
+		\item \hlr{Critic:} an orthogonal variance reduction technique from the \textbf{PG} literature
 	\end{itemize}
